### 2020-05-23 demo.Rmd
initial

---
title: "An interdisciplinary examination of stress and injury occurrence in athletes"
runtitle: "Short Title"
documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Dr Harry Fisher
    affiliation: '1'
    etal: Fisher # First author's last name. 
    email: harryfisher21@gmail.com # Indicates corresponding Author
    institution: Cardiff Metropolitan University
    street: Cyncoed Road
    zip: CF23 6XD
    country: United Kingdom
  - name: Dr Marianne Gittoes
    affiliation: '1'
  - name: Professor Lynne Evans
    affiliation: '1'
  - name: Miss Leah Bitchell
    affiliation: '1'  
  - name: Dr Richard Mullen  
    affiliation: '2'
  - name: Dr Marco Scutari
    affiliation: '3'    
affiliation:
  - id: '1'
    department: School of Sport and Health Science
    institution: Cardiff Metropolitan University
    city: Cardiff
    country: United Kingdom
  - id: '2' 
    department: Division of Sport, Health & Exercise Sciences
    institution: Brunel University
    city: London
    country: United Kingdom
  - id: '3' 
    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
    city: Manno
    country: Switzerland

output: rticles::frontiers_article
csl: frontiers.csl
bibliography: library.bib
---

\begin{abstract}

This paper adopts an interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period. At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
d reported any injuries they had sustained since the last data collection. Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study. Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury. Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury. In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury. Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.

\tiny
 \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}

\end{abstract}

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE)
#source("r-files/functions.R")
#source("r-files/methods_code.R")
#source("r-files/results_code.R")
```

# Introduction {-}

Over the last four decades sport related injuries have received increased research attention [@Ivarsson2017].
This attention is unsurprising given the high incidence [@Rosa2014; @Sheu2016],
and undesirable physical and psychological effects of sports injuries [@Leddy1994; @Brewer2012].
To mitigate against both the increasing incidence and	undesirable consequences of injury, research  has identified several psychological [@Slimani2018],
anatomical [@Murphy2003],
biomechanical [@Neely1998; @Hughes2014]
and environmental [@Meeuwisse2007]
factors associated with sports injury occurrence.
Indeed, several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009],
of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].

\begin{figure}[!h]
\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
\label{fig:fig1}
\end{figure}

Williams and Andersen's [@Williams1998] model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) will contribute to their response, either interactively or in isolation.
Central to the model is the stress response, which reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation. The model predicts that athletes who have a history of many stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater risk of injury.

Within Williams and Andersen’s [@Williams1998] model, major life events, a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007];
specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
However, personality traits and coping resources have also been	found to predict injury, with for example, athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to athletes with the opposite profile. [@Smith1990; @Lavallee1996; @Ivarsson2010].
However, the amount of variance explained by the psychosocial factors proposed by the model has been modest, typically between 5 - 30% [@Galambos2005; @Ivarsson2010]; suggesting other factors are also likely to contribute to injury occurrence.

While the psychosocial factors proposed in Williams and Andersen’s [@Williams1998] model have received the most research attention, the mechanisms through which these factors are proposed to exert their effect have remained under-investigated in the literature. 
To elaborate, the  model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
However, to date, the research has largely focused on attentional deficits [@Andersen1999; @Rogers2005; @Wilkerson2012a; @Swanik2007].
For example, Andersen and Williams [@Andersen1999]
measured peripheral and central vision during high and low stress conditions and found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following season.
Indeed, Rodgers and Landers [@Rogers2005] supported Andersen and Williams's [@Andersen1999] earlier findings reporting that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
However, the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control, remains to be explored [cf. @Williams1998].

One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
[@Meeuwisse2007; @Wiese-Bjornstal2009].
For example, a large body of research indicates that training-related stress is also likely to be related to the stress response and injury occurrence [@Lee2017; @Djaoui2017],
and may account for the unexplained variance from the psychological predictors of injury.
Appaneal and Perna [@Appaneal2014]
proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model and to address some of these issues.
To elaborate, the BMSAIH aimed to clarify the mediating pathways between the stress response and injury, consider other health outcomes and behavioural factors that impact sports participation,
and integrate the impact of training on athletes’ health [@Appaneal2014].
The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p. 74] to a range of undesirable health outcomes including illness and injury.
Consequently, the BMSAIH provides a framework for future research to build on Williams and Andersen's [@Williams1998] model, by including other physiological markers of training-related stress, which together may provide greater insight into the injury process.

Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other research has identified additional markers of training-related stress that are associated with an increased risk of injury; for example,
heart rate variability [@Bellenger2016; @Williams2017],
postural stability [@Romero-Franco2014]
and muscle stiffness [@Pruyn2015].
However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
Recently, Bittencourt et al. [@Bittencourt2016]
suggested that to better understand the multifactorial nature of sports injuries, research needs to move away from studying risk factors in isolation and instead adopt a complex systems approach to injury.
Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p. 3],
where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.

A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables. One technique that provides a solution to this challenge is Bayesian network (BN) modelling. BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network) of interest [@Scutari2014]. Once constructed, the learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?). Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].

To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
Given the exploratory nature of the current study the following objectives were defined:

- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
- Examine the relationships between the markers of stress and injury using a prospective, repeated measures design.
- Evaluate the relationships between the markers of stress and injury using Bayesian network modelling that captures the complex nature of injury occurrence.

### 2020-05-23 demo.tex
initial

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersHLTH}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\def\Authors{
  Dr Harry Fisher\,\textsuperscript{1*},
  Dr Marianne Gittoes\,\textsuperscript{1},
  Professor Lynne Evans\,\textsuperscript{1},
  Miss Leah Bitchell\,\textsuperscript{1},
  Dr Richard Mullen\,\textsuperscript{2},
  Dr Marco Scutari\,\textsuperscript{3}}

\def\Address{

  \textsuperscript{1} School of Sport and Health Science, Cardiff Metropolitan University,  Cardiff,   United Kingdom

  \textsuperscript{2} Division of Sport, Health \& Exercise Sciences, Brunel University,  London,   United Kingdom

  \textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA),  Manno,   Switzerland
  }

  \def\corrAuthor{Dr Harry Fisher}\def\corrAddress{Cardiff Metropolitan University\\Cyncoed Road\\, CF23 6XD United Kingdom}\def\corrEmail{\href{mailto:harryfisher21@gmail.com}{\nolinkurl{harryfisher21@gmail.com}}}
  \def\firstAuthorLast{Fisher {et~al.}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Short Title]{An interdisciplinary examination of stress and injury occurrence in
athletes}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

This paper adopts an interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period. At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
d reported any injuries they had sustained since the last data collection. Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study. Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury. Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury. In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury. Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.

\tiny
 \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}

\end{abstract}

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Over the last four decades sport related injuries have received
increased research attention (Ivarsson et al., 2017). This attention is
unsurprising given the high incidence (Rosa et al., 2014; Sheu et al.,
2016), and undesirable physical and psychological effects of sports
injuries (Leddy et al., 1994; Brewer, 2012). To mitigate against both
the increasing incidence and undesirable consequences of injury,
research has identified several psychological (Slimani et al., 2018),
anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes,
2014) and environmental (Meeuwisse et al., 2007) factors associated with
sports injury occurrence. Indeed, several models of injury causation
have been proposed that highlight the multifactorial nature of injury
occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009),
of which one of the most widely cited was developed by Williams and
Anderson (Fig \ref{fig:fig1}; Andersen and Williams, 1988; Williams and
Andersen, 1998).

\begin{figure}[!h]
\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
\label{fig:fig1}
\end{figure}

Williams and Andersen's (Williams and Andersen, 1998) model proposed
that when faced with a potentially stressful athletic situation, an
athlete's personality traits (e.g., hardiness, locus of control and
competitive trait anxiety), history of stressors (e.g., major life
events and previous injuries) and coping resources (e.g., general coping
behaviours) will contribute to their response, either interactively or
in isolation. Central to the model is the stress response, which
reflects the bi-directional relationship between athletes' appraisal of,
and response to, a stressful athletic situation. The model predicts that
athletes who have a history of many stressors, personality traits that
intensify the stress response and few coping resources, will exhibit
greater attentional (e.g., peripheral narrowing) and/or physiological
(e.g., increased muscle tension) responses that put these individuals at
greater risk of injury.

Within Williams and Andersen's (Williams and Andersen, 1998) model,
major life events, a component of an athlete's history of stressors,
most consistently predicts injury occurrence (Williams and Andersen,
2007); specifically, major life events with a negative, as opposed to
positive, valence (Passer and Seese, 1983; Maddison and Prapavessis,
2005). However, personality traits and coping resources have also been
found to predict injury, with for example, athletes more likely to
sustain an injury if they have poor social support and psychological
coping skills, and high trait anxiety and elevated competitive state
anxiety; compared to athletes with the opposite profile. (Smith et al.,
1990; Lavallée and Flint, 1996; Ivarsson and Johnson, 2010). However,
the amount of variance explained by the psychosocial factors proposed by
the model has been modest, typically between 5 - 30\% (Galambos et al.,
2005; Ivarsson and Johnson, 2010); suggesting other factors are also
likely to contribute to injury occurrence.

While the psychosocial factors proposed in Williams and Andersen's
(Williams and Andersen, 1998) model have received the most research
attention, the mechanisms through which these factors are proposed to
exert their effect have remained under-investigated in the literature.
To elaborate, the model suggests that injuries are likely to occur
through either increased physiological arousal resulting in increased
muscle tension and reduced flexibility or attentional deficits caused by
increased distractibility and peripheral narrowing. However, to date,
the research has largely focused on attentional deficits (Andersen and
Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et
al., 2007). For example, Andersen and Williams (Andersen and Williams,
1999) measured peripheral and central vision during high and low stress
conditions and found athletes with high life event stress coupled with
low social support had greater peripheral narrowing under stressful
conditions compared to athletes with the opposing profile; these
athletes went on to sustain an increased number of injuries during the
following season. Indeed, Rodgers and Landers (Rogers and Landers, 2005)
supported Andersen and Williams's (Andersen and Williams, 1999) earlier
findings reporting that peripheral narrowing under stress mediated 8.1\%
of the relationship between negative life events and injury. However,
the remaining variance between negative life events and athletic injury
through the other proposed mechanisms, such as increased muscle tension
and reduced motor control, remains to be explored (cf.~Williams and
Andersen, 1998).

One challenge faced by researchers addressing the sports injury problem
with a psychological lens is the multifactorial nature of injury, and
the possible contribution of other non-psychological factors to the
stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009). For
example, a large body of research indicates that training-related stress
is also likely to be related to the stress response and injury
occurrence (Lee et al., 2017; Djaoui et al., 2017), and may account for
the unexplained variance from the psychological predictors of injury.
Appaneal and Perna (Appaneal and Perna, 2014) proposed the
biopsychosocial model of stress athletic injury and health (BMSAIH) to
serve as an extension to Williams and Andersen's (Williams and Andersen,
1998) model and to address some of these issues. To elaborate, the
BMSAIH aimed to clarify the mediating pathways between the stress
response and injury, consider other health outcomes and behavioural
factors that impact sports participation, and integrate the impact of
training on athletes' health (Appaneal and Perna, 2014). The central
tenet of the BMSAIH is that psychosocial distress (e.g., negative life
events) may act synergistically with training-related stress as a result
of high-intensity and high-volume sports training, and ``widen the
window of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of
undesirable health outcomes including illness and injury. Consequently,
the BMSAIH provides a framework for future research to build on Williams
and Andersen's (Williams and Andersen, 1998) model, by including other
physiological markers of training-related stress, which together may
provide greater insight into the injury process.

Although research supporting the BMSAIH has mainly focused on the
relationship between hormonal responses to training and injury
occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other
research has identified additional markers of training-related stress
that are associated with an increased risk of injury; for example, heart
rate variability (Bellenger et al., 2016; Williams et al., 2017),
postural stability (Romero-Franco et al., 2014) and muscle stiffness
(Pruyn et al., 2015). However, these markers are often studied in
isolation without an assessment of the psychosocial factors that are
known to contribute to injury, thereby limiting our understanding of how
psychosocially and physiologically derived stress may contribute
synergistically to injury occurrence. Recently, Bittencourt et
al.~(Bittencourt et al., 2016) suggested that to better understand the
multifactorial nature of sports injuries, research needs to move away
from studying risk factors in isolation and instead adopt a complex
systems approach to injury. Such an approach posits that injury may
arise from a complex ``web of determinants'' (Bittencourt et al., 2016,
3), where different factors interact in unpredictable and unplanned
ways, but result in a global outcome pattern of either adaptation or
injury.

A challenge when adopting a complex systems approach is using an
appropriate analysis technique that is able to capture the uncertainty
and complexity of the relationships between different factors. One
technique that provides a solution to this challenge is Bayesian network
(BN) modelling. BN's allow the construction of graphical probabilistic
models using the underlying structure that connects different variables
(nodes in the network) of interest (Scutari and Denis, 2014). Once
constructed, the learned structure can be used for inference by
obtaining the posterior probabilities of a particular node for a given
query (e.g., if the value of Node A is x and the value of Node B is y,
what is the probability Node C being value z?). Furthermore, BN's do not
distinguish between dependent and independent variables as they are a
form of unsupervised learning, which is a strength over regression or
structure equation models when the underlying relationship in the
network may not be known (Olmedilla et al., 2018).

To summarise, despite offering a possible framework to build on the
research stemming from Williams and Andersen's (Williams and Andersen,
1998) model, there remains and opportunity to explore other
physiological stress-related markers proposed by the BMSAIH, in addition
to the already well-established psychological characteristics known to
be related to injury (Appaneal and Perna, 2014). Furthermore, research
has typically not captured changes in both psychosocial factors and
stress-related physiological markers that may occur between initial
measurement and injury occurrence. Given the exploratory the current
study had the following objectives:

\begin{itemize}
\tightlist
\item
  Identify suitable markers of stress that can be easily captured in a
  large cohort of athletes in a field based setting.
\item
  Examine the relationships between the markers of stress and injury
  using a prospective, repeated measures design.
\item
  Evaluate the relationships between the markers of stress and injury
  using Bayesian network modelling that captures the complex nature of
  injury occurrence.
\end{itemize}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Andersen1988}{}%
Andersen, M. B., and Williams, J. M. (1988). A model of stress and
athletic injury: Prediction and prevention. \emph{Journal of Sport and
Exercise Psychology} 10, 294--306.
doi:\href{https://doi.org/10.1123/jsep.10.3.294}{10.1123/jsep.10.3.294}.

\leavevmode\hypertarget{ref-Andersen1999}{}%
Andersen, M. B., and Williams, J. M. (1999). Athletic injury,
psychosocial factors and perceptual changes during stress. \emph{Journal
of Sports Sciences} 17, 735--741.
doi:\href{https://doi.org/10.1080/026404199365597}{10.1080/026404199365597}.

\leavevmode\hypertarget{ref-Appaneal2014}{}%
Appaneal, R. N., and Perna, F. M. (2014). ``Biopsychosocial model of
injury,'' in \emph{Encyclopedia of sport and exercise psychology}, eds.
R. C. Eklund and G. Tenenbaum (Thousand Oaks, CA: Sage), 74--77.

\leavevmode\hypertarget{ref-Bellenger2016}{}%
Bellenger, C. R., Fuller, J. T., Thomson, R. L., Davison, K., Robertson,
E. Y., and Buckley, J. D. (2016). Monitoring athletic training status
through autonomic heart rate regulation: A systematic review and
meta-analysis. \emph{Sports Medicine} 46, 1461--1486.
doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}.

\leavevmode\hypertarget{ref-Bittencourt2016}{}%
Bittencourt, N. F. N., Meeuwisse, W. H., Mendonça, L. D.,
Nettel-Aguirre, A., Ocarino, J. M., and Fonseca, S. T. (2016). Complex
systems approach for sports injuries: Moving from risk factor
identification to injury pattern recognition - narrative review and new
concept. \emph{British Journal of Sports Medicine} 50, 1309--1314.
doi:\href{https://doi.org/10.1136/bjsports-2015-095850}{10.1136/bjsports-2015-095850}.

\leavevmode\hypertarget{ref-Brewer2012}{}%
Brewer, B. W. (2012). ``Psychology of sport injury rehabilitation,'' in
\emph{Handbook of sport psychology}, eds. G. Tenenbaum and R. C. Eklund
(Hoboken, NJ, USA: Wiley), 404--424.
doi:\href{https://doi.org/10.1002/9781118270011.ch18}{10.1002/9781118270011.ch18}.

\leavevmode\hypertarget{ref-Djaoui2017}{}%
Djaoui, L., Haddad, M., Chamari, K., and Dellal, A. (2017). Monitoring
training load and fatigue in soccer players with physiological markers.
\emph{Physiology and Behavior} 181, 86--94.
doi:\href{https://doi.org/10.1016/j.physbeh.2017.09.004}{10.1016/j.physbeh.2017.09.004}.

\leavevmode\hypertarget{ref-Galambos2005}{}%
Galambos, S. A., Terry, P. C., Moyle, G. M., and Locke, S. A. (2005).
Psychological predictors of injury among elite athletes. \emph{British
Journal of Sports Medicine} 39, 351--354.
doi:\href{https://doi.org/10.1136/bjsm.2005.018440}{10.1136/bjsm.2005.018440}.

\leavevmode\hypertarget{ref-Hughes2014}{}%
Hughes, G. (2014). A review of recent perspectives on biomechanical risk
factors associated with anterior cruciate ligament injury.
\emph{Research in Sports Medicine} 22, 193--212.
doi:\href{https://doi.org/10.1080/15438627.2014.881821}{10.1080/15438627.2014.881821}.

\leavevmode\hypertarget{ref-Ivarsson2010}{}%
Ivarsson, A., and Johnson, U. (2010). Psychological factors as
predictors of injuries among senior soccer players. A prospective study.
\emph{Journal of Sports Science and Medicine} 9, 347--352. Available at:
\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3761721/}.

\leavevmode\hypertarget{ref-Ivarsson2017}{}%
Ivarsson, A., Johnson, U., Andersen, M. B., Tranaeus, U., Stenling, A.,
and Lindwall, M. (2017). Psychosocial factors and sport injuries:
Meta-analyses for prediction and prevention. \emph{Sports Medicine} 47,
353--365.
doi:\href{https://doi.org/10.1007/s40279-016-0578-x}{10.1007/s40279-016-0578-x}.

\leavevmode\hypertarget{ref-Kumar2001}{}%
Kumar, S. (2001). Theories of musculoskeletal injury causation.
\emph{Ergonomics} 44, 17--47.
doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}.

\leavevmode\hypertarget{ref-Lavallee1996}{}%
Lavallée, L., and Flint, F. (1996). The relationship of stress,
competitive anxiety, mood state, and social support to athletic injury.
\emph{Journal of Athletic Training} 31, 296--299. Available at:
\url{http://www.ncbi.nlm.nih.gov/pubmed/16558413}.

\leavevmode\hypertarget{ref-Leddy1994}{}%
Leddy, M. H., Lambert, M. J., and Ogles, B. M. (1994). Psychological
consequences of athletic injury among high-level competitors.
\emph{Research Quarterly for Exercise and Sport} 65, 347--354.
doi:\href{https://doi.org/10.1080/02701367.1994.10607639}{10.1080/02701367.1994.10607639}.

\leavevmode\hypertarget{ref-Lee2017}{}%
Lee, E. C., Fragala, M. S., Kavouras, S. A., Queen, R. M., Pryor, J. L.,
and Casa, D. J. (2017). Biomarkers in sports and exercise: Tracking
health, performance, and recovery in athletes. \emph{Journal of Strength
and Conditioning Research} 31, 2920--2937.
doi:\href{https://doi.org/10.1519/JSC.0000000000002122}{10.1519/JSC.0000000000002122}.

\leavevmode\hypertarget{ref-Maddison2005}{}%
Maddison, R., and Prapavessis, H. (2005). A psychological approach to
the prediction and prevention of athletic injury. \emph{Journal of Sport
and Exercise Psychology} 27, 289--310.
doi:\href{https://doi.org/10.1123/jsep.27.3.289}{10.1123/jsep.27.3.289}.

\leavevmode\hypertarget{ref-Meeuwisse2007}{}%
Meeuwisse, W. H., Tyreman, H., Hagel, B., and Emery, C. (2007). A
dynamic model of etiology in sport injury: The recursive nature of risk
and causation. \emph{Clinical Journal of Sport Medicine} 17, 215--219.
doi:\href{https://doi.org/10.1097/JSM.0b013e3180592a48}{10.1097/JSM.0b013e3180592a48}.

\leavevmode\hypertarget{ref-Murphy2003}{}%
Murphy, D. F., Connolly, D. A. J., and Beynnon, B. D. (2003). Risk
factors for lower extremity injury: A review of the literature.
\emph{British Journal of Sports Medicine} 37, 13--29.
doi:\href{https://doi.org/10.1136/bjsm.37.1.13}{10.1136/bjsm.37.1.13}.

\leavevmode\hypertarget{ref-Neely1998}{}%
Neely, F. G. (1998). Biomechanical risk factors for exercise-related
lower limb injuries. \emph{Sports Medicine} 26, 395--413.
doi:\href{https://doi.org/10.2165/00007256-199826060-00003}{10.2165/00007256-199826060-00003}.

\leavevmode\hypertarget{ref-Olmedilla2018}{}%
Olmedilla, A., Rubio, V. J., Fuster-Parra, P., Pujals, C., and
García-Mas, A. (2018). A Bayesian approach to sport injuries likelihood:
Does player's self-efficacy and environmental factors plays the main
role? \emph{Frontiers in Psychology} 9, 1--10.
doi:\href{https://doi.org/10.3389/fpsyg.2018.01174}{10.3389/fpsyg.2018.01174}.

\leavevmode\hypertarget{ref-Passer1983a}{}%
Passer, M. W., and Seese, M. D. (1983). Life stress and athletic injury:
Examination of positive versus negative events and three moderator
variables. \emph{Journal of Human Stress} 9, 11--16.
doi:\href{https://doi.org/10.1080/0097840X.1983.9935025}{10.1080/0097840X.1983.9935025}.

\leavevmode\hypertarget{ref-Perna2003}{}%
Perna, F. M., Antoni, M. H., Baum, A., Gordon, P., and Schneiderman, N.
(2003). Cognitive behavioral stress management effects on injury and
illness among competitive athletes: A randomized clinical trial.
\emph{Annals of Behavioral Medicine} 25, 66--73. Available at:
\url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}.

\leavevmode\hypertarget{ref-Perna1995}{}%
Perna, F. M., and McDowell, S. L. (1995). Role of psychological stress
in cortisol recovery from exhaustive exercise among elite athletes.
\emph{International Journal of Behavioral Medicine} 2, 13--26.
doi:\href{https://doi.org/10.1207/s15327558ijbm0201_2}{10.1207/s15327558ijbm0201\_2}.

\leavevmode\hypertarget{ref-Perna1997}{}%
Perna, F., Schneiderman, N., and LaPerriere, A. (1997). Psychological
stress, exercise and immunity. \emph{International Journal of Sports
Medicine} 18, 78--83.
doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}.

\leavevmode\hypertarget{ref-Pruyn2015}{}%
Pruyn, E. C., Watsford, M. L., and Murphy, A. J. (2015). Differences in
lower-body stiffness between levels of netball competition.
\emph{Journal of Strength and Conditioning Research} 29, 1197--1202.
doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}.

\leavevmode\hypertarget{ref-Rogers2005}{}%
Rogers, T., and Landers, D. M. (2005). Mediating effects of peripheral
vision in the life event stress/athletic injury relationship.
\emph{Journal of Sport and Exercise Psychology} 27, 271--288.
doi:\href{https://doi.org/10.1002/9781444303650}{10.1002/9781444303650}.

\leavevmode\hypertarget{ref-Romero-Franco2014}{}%
Romero-Franco, N., Gallego-Izquierdo, T., Martínez-López, E. J.,
Hita-Contreras, F., Osuna-Pére, Catalina, M., and Martínez-Amat, A.
(2014). Postural stability and subsequent sports injuries during indoor
season of athletes. \emph{Journal of Physical Therapy Science} 26,
683--687.
doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}.

\leavevmode\hypertarget{ref-Rosa2014}{}%
Rosa, B. B., Asperti, A. M., Helito, C. P., Demange, M. K., Fernandes,
T. L., and Hernandez, A. J. (2014). Epidemiology of sports injuries on
collegiate athletes at a single center. \emph{Acta Ortopédica
Brasileira} 22, 321--324.
doi:\href{https://doi.org/10.1590/1413-78522014220601007}{10.1590/1413-78522014220601007}.

\leavevmode\hypertarget{ref-Scutari2014}{}%
Scutari, M., and Denis, J.-B. (2014). \emph{Bayesian networks: with
examples in R}. 1st ed. Chapman \& Hall/CRC Available at:
\url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}.

\leavevmode\hypertarget{ref-Sheu2016}{}%
Sheu, Y., Chen, L. H., and Hedegaard, H. (2016). Sports- and
Recreation-related Injury Episodes in the United States, 2011-2014.
\emph{National Health Statistics Reports}, 1--12. Available at:
\url{https://www.ncbi.nlm.nih.gov/pubmed/27906643}.

\leavevmode\hypertarget{ref-Slimani2018}{}%
Slimani, M., Bragazzi, N. L., Znazen, H., Paravlic, A., Azaiez, F., and
Tod, D. (2018). Psychosocial predictors and psychological prevention of
soccer injuries: A systematic review and meta-analysis of the
literature. \emph{Physical Therapy in Sport} 32, 293--300.
doi:\href{https://doi.org/10.1016/j.ptsp.2018.05.006}{10.1016/j.ptsp.2018.05.006}.

\leavevmode\hypertarget{ref-Smith1990}{}%
Smith, R. E., Smoll, F. L., and Ptacek, J. T. (1990). Conjunctive
moderator variables in vulnerability and resiliency research: Life
stress, social support and coping skills, and adolescent sport injuries.
\emph{Journal of Personality and Social Psychology} 58, 360--370.
doi:\href{https://doi.org/10.1037/0022-3514.58.2.360}{10.1037/0022-3514.58.2.360}.

\leavevmode\hypertarget{ref-Swanik2007}{}%
Swanik, C. B., Covassin, T., Stearne, D. J., and Schatz, P. (2007). The
relationship between neurocognitive function and noncontact anterior
cruciate ligament injuries. \emph{American Journal of Sports Medicine}
35, 943--948.
doi:\href{https://doi.org/10.1177/0363546507299532}{10.1177/0363546507299532}.

\leavevmode\hypertarget{ref-Wiese-Bjornstal2009}{}%
Wiese-Bjornstal, D. M. (2009). Sport injury and college athlete health
across the lifespan. \emph{Journal of Intercollegiate Sport} 2, 64--80.
doi:\href{https://doi.org/10.1123/jis.2.1.64}{10.1123/jis.2.1.64}.

\leavevmode\hypertarget{ref-Wilkerson2012a}{}%
Wilkerson, G. B. (2012). Neurocognitive reaction time predicts lower
extremity sprains and strains. \emph{International Journal of Athletic
Therapy and Training} 17, 4--9.
doi:\href{https://doi.org/10.1123/ijatt.17.6.4}{10.1123/ijatt.17.6.4}.

\leavevmode\hypertarget{ref-Williams2007}{}%
Williams, J. M., and Andersen, M. B. (2007). ``Psychosocial antecedents
of sport injury and interventions for risk reduction,'' in
\emph{Handbook of sport psychology}, eds. G. Tenenbaum and R. C. Eklund
(Hoboken, NJ, USA: Wiley), 379--403.
doi:\href{https://doi.org/10.1002/9781118270011.ch17}{10.1002/9781118270011.ch17}.

\leavevmode\hypertarget{ref-Williams1998}{}%
Williams, J. M., and Andersen, M. B. (1998). Psychosocial antecedents of
sport injury: review and critique of the stress and injury model.
\emph{Journal of Applied Sport Psychology} 10, 5--25.
doi:\href{https://doi.org/10.1080/10413209808406375}{10.1080/10413209808406375}.

\leavevmode\hypertarget{ref-Williams2017}{}%
Williams, S., Booton, T., Watson, M., Rowland, D., and Altini, M.
(2017). Heart rate variability is a moderating factor in the
workload-injury relationship of competitive crossfit™ athletes.
\emph{Journal of Sports Science and Medicine} 16, 443--449.

\end{document}

### 2020-05-23 full_paper_v2.Rmd
initial

---
title: "An interdisciplinary examination of stress and injury occurrence in athletes"
runtitle: "Short Title"
documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Dr Harry Fisher
    affiliation: '1'
    etal: Fisher # First author's last name. 
    email: harryfisher21@gmail.com # Indicates corresponding Author
    institution: Cardiff Metropolitan University
    street: Cyncoed Road
    zip: CF23 6XD
    country: United Kingdom
  - name: Dr Marianne Gittoes
    affiliation: '1'
  - name: Professor Lynne Evans
    affiliation: '1'
  - name: Miss Leah Bitchell
    affiliation: '1'  
  - name: Dr Richard Mullen  
    affiliation: '2'
  - name: Dr Marco Scutari
    affiliation: '3'    
affiliation:
  - id: '1'
    department: School of Sport and Health Science
    institution: Cardiff Metropolitan University
    city: Cardiff
    country: United Kingdom
  - id: '2' 
    department: Division of Sport, Health & Exercise Sciences
    institution: Brunel University
    city: London
    country: United Kingdom
  - id: '3' 
    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
    city: Manno
    country: Switzerland

output: rticles::frontiers_article
csl: frontiers.csl
bibliography: library.bib
---

\begin{abstract}

This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period. At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
d reported any injuries they had sustained since the last data collection. Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study. Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury. Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury. In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury. Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.

\tiny
 \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}

\end{abstract}

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE)
source("r-files/functions.R")
source("r-files/methods_code.R")
source("r-files/results_code.R")
```

# Introduction {-}

Over the last four decades sport related injuries have received increased research attention [@Ivarsson2017].
This attention is unsurprising given the high incidence [@Rosa2014; @Sheu2016],
and undesirable physical and psychological effects of sports injuries [@Leddy1994; @Brewer2012].
To mitigate against both the increasing incidence and	undesirable consequences of injury, research  has identified several psychological [@Slimani2018],
anatomical [@Murphy2003],
biomechanical [@Neely1998; @Hughes2014]
and environmental [@Meeuwisse2007]
factors associated with sports injury occurrence.
Indeed, several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009],
of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].

\begin{figure}[!h]
\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
\label{fig:fig1}
\end{figure}

Williams and Andersen's [@Williams1998] model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) will contribute to their response, either interactively or in isolation.
Central to the model is the stress response, which reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation. The model predicts that athletes who have a history of many stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater risk of injury.

Within Williams and Andersen’s [@Williams1998] model, major life events, a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007];
specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
However, personality traits and coping resources have also been	found to predict injury, with for example, athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to athletes with the opposite profile. [@Smith1990; @Lavallee1996; @Ivarsson2010].
However, the amount of variance explained by the psychosocial factors proposed by the model has been modest, typically between 5 - 30% [@Galambos2005; @Ivarsson2010]; suggesting other factors are also likely to contribute to injury occurrence.

While the psychosocial factors proposed in Williams and Andersen’s [@Williams1998] model have received the most research attention, the mechanisms through which these factors are proposed to exert their effect have remained under-investigated in the literature. 
To elaborate, the  model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
However, to date, the research has largely focused on attentional deficits [@Andersen1999; @Rogers2005; @Wilkerson2012a; @Swanik2007].
For example, Andersen and Williams [@Andersen1999]
measured peripheral and central vision during high and low stress conditions and found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following season.
Indeed, Rodgers and Landers [@Rogers2005] supported Andersen and Williams's [@Andersen1999] earlier findings reporting that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
However, the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control, remains to be explored [cf. @Williams1998].

One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
[@Meeuwisse2007; @Wiese-Bjornstal2009].
For example, a large body of research indicates that training-related stress is also likely to be related to the stress response and injury occurrence [@Lee2017; @Djaoui2017],
and may account for the unexplained variance from the psychological predictors of injury.
Appaneal and Perna [@Appaneal2014]
proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model and to address some of these issues.
To elaborate, the BMSAIH aimed to clarify the mediating pathways between the stress response and injury, consider other health outcomes and behavioural factors that impact sports participation,
and integrate the impact of training on athletes’ health [@Appaneal2014].
The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p. 74] to a range of undesirable health outcomes including illness and injury.
Consequently, the BMSAIH provides a framework for future research to build on Williams and Andersen's [@Williams1998] model, by including other physiological markers of training-related stress, which together may provide greater insight into the injury process.

Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other research has identified additional markers of training-related stress that are associated with an increased risk of injury; for example,
heart rate variability [@Bellenger2016; @Williams2017],
postural stability [@Romero-Franco2014]
and muscle stiffness [@Pruyn2015].
However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
Recently, Bittencourt et al. [@Bittencourt2016]
suggested that to better understand the multifactorial nature of sports injuries, research needs to move away from studying risk factors in isolation and instead adopt a complex systems approach to injury.
Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p. 3],
where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.

A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables. One technique that provides a solution to this challenge is Bayesian network (BN) modelling. BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014]. The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?). Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].

To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
Given the exploratory nature of the current study the following objectives were defined:

- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.

# Methods {-}

```{r methodssetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Participants

A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study. Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
Participants self-rated competitive level ranged from recreational to international standard. 
A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
Participants were engaged in training for their respective sports for at least five hours per week.
Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.

```{r table1, results='asis'}

kable(
  npars %>% 
    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
    slice(c(1,3,8, 4, 9:11)) %>% 
  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
                          name == "Height" ~ "Height (cm)",
                          name == "Weight" ~ "Body mass (kg)",
                          TRUE ~ name)),
  caption = "Participant characteristics.",
  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
  align = "c",
  escape = F
) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Demographics M (SD)", 1, 4) %>%
  pack_rows("Current competitive level n (%)", 5, 7) %>%
  row_spec(0, bold = TRUE)
```

## Measures

### Major life events

A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
The LESCA comprises 69 items that reflect possible life events that participants may have experienced. 
Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$). 
Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
A score for total life events was calculated by summing the absolute values for both negative and positive events.
Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. 
Petrie also provided evidence of discriminant, convergent and predictive validity. 
The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
The composite reliability for the LESCA in this study was 0.84.

### Reinforcement Sensitivity Theory Personality Questionnaire

A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total score that was subsequently used for further analysis.
The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`. Further details regarding the revised RST are in S1 Appendix.

### Heart rate variability

A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].

### Musculoskeletal properties

A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure muscle stiffness.
The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016]. 
The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].

### Postural stability

Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
Participants were asked to maintain each stance for a total of 20-seconds.
Participants hands were placed on hips at the level of the iliac crests. 
A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.

\begin{figure}[!h]
\caption{{\bf mBESS positions (A-F).}
Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
\label{fig:fig2}
\end{figure}

The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
b) opening eyes; 
c) stepping, stumbling, or falling; 
d) moving the thigh into more than 30 degrees of flexion or abduction; 
e) lifting the forefoot or heel; and 
f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
To calculate limb asymmetry, the DL and NDL leg score was calculated by summing the DL and NDL errors respectively.
A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.

### Injury

Participants self-reported any injuries they sustained at each data collection during the study period.
An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
Injury status (did / did not sustain an injury) served as the main outcome measure.

## Procedure

At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week. 
Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.

A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018. 
Each participant was asked to attend a total of four data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
Participants provided informed consent before data collection commenced.

\begin{figure}[!h]
\caption{{\bf Study design.}
For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
\label{fig:fig3}
\end{figure}

For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
Participants were randomly assigned to one of the two groups and remained in those groups across all time points.

\begin{figure}[!h]
\caption{{\bf Session protocol.}
Outline of the protocol for each data collection.}
\label{fig:fig4}
\end{figure}

### Questionnaires

The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed  on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis purposes.

### HRV

To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines. 
Participants were seated and asked to remain as still as possible for the duration of the recording. 
No attempt was made to control participants respiratory frequency or tidal volume [@Denver2007]. 
Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.

Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices. 
Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996]. RMSSD was calculated as:

\begin{equation} 
  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
\end{equation}

Where N is the length of the time series, and $RR_i$ the RR interval between beats $i$ and $i-1$, where each beat position corresponds to the beat detection instant.

### Muscle stiffness

To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb. 
The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique  to determine the exact location of each site [@Chuang2012].
The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle. 
The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.

After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site. 
A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues. The device was programmed to deliver five consecutive impulses, separated by a one second interval [@Morgan2018]. 
For each impulse, the device computed stiffness values, with the median of the five values being saved by the device for further analysis. 
In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted. 
Sets of measurements above 3 % were measured again to ensure the reliability of the data. 
The CV was calculated in real time by the device after each set of measurements. 
Measurements saved on the device were uploaded to a computer using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis. 
For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis.

### Postural stability

Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds. 
Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
The video recording was started prior to the first stance position and stopped after all stances had been completed.
Each completed mBESS protocol took approximately 4 minutes.
Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.

## Data Analysis

Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014]. 
A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988]. 
However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012]. 
In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.

\begin{figure}[H]
\caption{{\bf Example network.}
A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
\label{fig:fig5}
\end{figure}

Learning the structure of the network is an important step in BN modelling. The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014]. 
In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.

[^1]: The term "query" in relation to Bayesian Networks stems from Pearl's expert systems theory (1988). A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly. Widely used texts on Bayesian Network analysis [@Koller2009] have widely adopted the terminology in favour of that used in traditional statistics.

BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure. 
Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.

### Data Preparation

To prepare the data for the BN, missing values in the dataset were first imputed. 
Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings. The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random. 
A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].

Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.

For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time. 
Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].

### Network structure

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference. 
Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist). 
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014]. 
To account for the repeated measures design and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
Table \ref{tab:table2} provides an example of the formatted data. 
Participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables. 
Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
All subsequent models used the same blacklist.

```{r table2, results='asis'}
exampledata <- tibble(
  Participant = c("1", "1", "2", "3", "3"),
  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
  " " = "->",
  "X_2" = c("T2", "T3", "T2", "T2", "T3")
)

kable(
  exampledata,
  caption = "Example of the data arrangement used for the network.",
  escape = TRUE
) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE)
```

### Preliminary network structures

Prior to the final network structure presented in the results section, several structures were investigated.
Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
A higher BIC value indicates the structure of a DAG is a better fit to the observed data [@Scutari2014]. 
BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.

Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`). 
Additionally, despite strong evidence in the literature that both negative and total life event stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data. 
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
Total life event score was not included in the final structure.

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure.

Finally, both total score and asymmetry were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.

\begin{table}[H]
\begin{adjustwidth}{-2.25in}{0in}
\caption{\label{tab:table3}Variables included in the final network structure.}
\centering
\begin{tabular}{l|l|l|l}
\hline
{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
\hline
clevel & Current competitive level & Club\_university\_county & National\_international\\
\hline
gender & Gender of the participant & Female & Male\\
\hline
hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
\hline
ind\_team & Participate in an individual or team based sport & Individual & Team\\
\hline
pi & Sustained previous injury & No Injury & Injury\\
\hline
nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
\hline
FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
\hline
BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
\hline
RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
\hline
stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
\hline
rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
\hline
balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
\hline
nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
\hline
nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
\hline
nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
\hline
\end{tabular}
\end{adjustwidth}
\end{table}

Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network. 
To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3. 
Independent variables were then standardized to allow relative changes between variables to be compared. 
The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High". 
Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.

To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
Arcs that were present in at least 30% of the models were included in the averaged model.
The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
Arcs below 0.3 were not included in the model.
The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.

### Network Inference

Conditional probability queries (CPQ) were used to perform inference on both network structures.
To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query. 
The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
Of particular interest in the current study were the variables that were connected to "injured" nodes.
To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.

The second network contained both continuous and discrete data. 
To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.

# Results {-}

# Results

During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

```{r table4, results='asis'}

cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
  kable(
    .,
    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
    escape = F
  ) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
  footnote(general = "Other included bone, skin and brain injuries.") %>%
  kable_styling(latex_options = "HOLD_position")  %>%
  row_spec(0,bold=TRUE)

```

```{r mainnetworkresults}
set.seed(999)
strengthfinder <- function(var1, var2){
  str %>%
    filter(from == var1 & to == var2)
}

from <- c("nlebase", "ind_team", "ind_team", "gender", "gender", "clevel", "pi")
to <- c("RI_1", "hours", "nlebase", "stiffness_1", "stiffness_2", "balance_1", "stiffness_1")
x <- pmap_dfr(.l = list(from = rep(from, 2),
     to = rep(to, 2),
     state = rep(c("High", "Low"), each = length(from))),
         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
  gather(var, state, -prob, -outcome) %>%
  drop_na(state) %>%
  mutate(prob = format(round(prob, digits = 2), nsmall = 2))

from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")

y <- pmap_dfr(.l = list(from = rep(from, 2),
     to = rep(to, 2),
     state = rep(c("High", "Low"), each = length(from))),
         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
  gather(var, state, -prob, -outcome) %>%
  drop_na(state) %>%
  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
```

### First network structure

The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
Several of the explanatory variables showed strong connections with independent variables in the network.
The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.

\begin{figure}[H]
\caption{{\bf Network structure.}
The full Bayesian network structure was plotted using the strength.plot function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables. The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
\label{fig:fig6}
\end{figure}

### Markov blanket for injured_1

```{r mb1}

set.seed(1001)
var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")

query1 <- map_dfr(var, ~ newprobtable(
  tallc,
  .x,
  "injured_1",
  "injured",
  avg30
)) %>%
  gather(variable, state, -prob, -outcome) %>%
  drop_na() %>%
  select(variable, state, prob) %>%
  spread(state, prob) %>%
  select(variable, Low, High) %>%
  dendroTools::round_df(2)

query2 <- newprobtable(
  tallc,
  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
  "injured_1",
  "injured",
  avg30
) %>%
  toptail(n = 3) %>%
  arrange(-prob) %>%
  select(outcome, everything()) %>%
  dendroTools::round_df(2)
```

The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.

```{r}
# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
# q1p <- graphviz.plot(plotsg1,
#                      shape = "rectangle",
#                      render = F,
#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
# 
# strq1 <- str %>% 
#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") & 
#            to %in% c("balance_1", "injured_1")) %>% 
#   filter(strength > 0.30) %>% 
#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
# 
# labels <- as.character(strq1$strength)
# labels[4] <- str_pad(labels[4], side = "left", width = 2)
# 
# edgenames <- edgeNames(q1p)
# names(labels) <- edgenames
# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6, 
#                                     "stiffness_1~injured_1" = 3,
#                                     "clevel~balance_1" = 3),
#                             lty = c("injured_1~balance_1" = "dashed",
#                                     "hours~injured_1" = "dashed"))
# 
# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
# graph.par(list(edges=list(fontsize = 12)))
# 
# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
# renderGraph(q1p)

```

\begin{figure}[H]
\caption{{\bf Markov blanket of injured\_1.}
Arc strengths are included as arc labels.}
\label{fig:fig7}
\end{figure}

```{r table5, results='asis'}
query1 %>%
  select(c(1, 2, 3)) %>%
  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
Negative life events had a negligible effect when moving from the "Low" to "High" state.

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.

```{r table6, results='asis'}

query2 %>%
  select(7, 2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}. 
Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.

```{r table7, results="asis"}
injured1comp %>%
  filter(variable != "outcome") %>% 
  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(  .,
    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
    col.names = c("Variable", "High risk", "Low risk"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

### Markov blanket of injured_2

```{r}
set.seed(1002)
var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2")
query3 <- map_dfr(var2, ~newprobtable(tallc,
                                      .x,
                                      "injured_2",
                                      "injured",
                                      avg30)) %>% 
  gather(variable, state, -prob, - outcome) %>% 
  drop_na() %>% 
  select(variable, state, prob) %>% 
  spread(state, prob) %>% 
  select(variable, Low, High) %>% 
  dendroTools::round_df(2)

query4 <- newprobtable(tallc, 
                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
                       "injured_2",
                       "injured",
                       avg30
) %>% 
  toptail(3) %>% 
  arrange(-prob) %>% 
  select(outcome, everything()) %>% 
  dendroTools::round_df(2)
```

The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
The arc between stiffness_2 and injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
Results of the first query for injured_2 in the "injured" state are presented in table \ref{tab:table8}.
Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.

```{r}
# plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
# q2p <- graphviz.plot(plotsg2,
#                      render = F,
#                      shape = "rectangle",
#                      groups = list(c("gender"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
# 
# strq2 <- str %>% 
#   filter(from %in% c("FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & 
#            to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
#   filter(strength > 0.30) %>% 
#   filter(direction != 0) %>% 
#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
# 
# labels <- as.character(strq2$strength)
# labels[6] <- str_pad(labels[6], side = "left", width = 2)
# edgenames <- c("gender~stiffness_2", "FFFS_1~injured_2",
#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
# names(labels) <- edgenames
# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
#                                     "nlelg_2~injured_2" = 6,
#                                     "clevel~balance_1" = 3,
#                                     "gender~stiffness_2" = 4),
#                             lty = c("FFFS_1~injured_2" = "dashed",
#                                     "injured_2~balance_2" = "dashed",
#                                     "injured_2~rmssd_2" = "dashed"))
# 
# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
# graph.par(list(edges=list(fontsize = 13)))
# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
# renderGraph(q2p)
```

\begin{figure}[H]
\caption{{\bf Markov blanket for injured\_2.} }
\label{fig:fig8}
\end{figure}

```{r table8, results="asis"}

query3 %>%
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".

```{r table9, results="asis" }

query4 %>%
  select(7,2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

## Second network structure - changes within variables

```{r include=FALSE, warning=FALSE}
set.seed(1003)
bmtable <- tidy(bm1, effects = "fixed") %>% 
  select(term:conf.high) %>% 
  dendroTools::round_df(2) %>% 
  mutate_at(vars(conf.low, conf.high), list(~format(round(., 2), nsmall = 2))) %>%
  mutate(conf.low = str_c("[", conf.low, sep = ""),
         conf.high = str_c(conf.high, "]", sep = "")) %>% 
  unite(., "95% CI", c("conf.low", "conf.high"), sep = ", ") %>% 
  mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
         "95% CI" = str_replace_all(`95% CI`, ",", ", ")) %>% 
  slice(1:4) %>% 
  mutate(term = c("Intercept", "RMSSD", "BIS", "RMSSD:BIS"))

```

The network for changes within variables is presented in Fig \ref{fig:fig9}.
An arc between BIS $\rightarrow$ FFFS with strength 1 was present in the network.
Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strength of 0.60.
The arc between RMSSD $\rightarrow$ FFFS was 0.56.
The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a linear regression model.
While the continuous nodes in the second network are already parameterised using linear regressions, it was not possible to investigate the interaction between explanatory variables.
The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.

```{r}
# mb(avg30, "injured_2")
# plotsg2 <- model2network('[nlelg_2][gender][pi][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|pi:gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
# q2p <- graphviz.plot(plotsg2,
#                      render = F,
#                      shape = "rectangle",
#                      groups = list(c("gender", "pi"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
# 
# strq2 <- str %>% 
#   filter(from %in% c("pi", "FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
#   filter(strength > 0.30) %>% 
#   filter(direction != 0) %>% 
#   slice(2:8) %>% 
#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
# 
# labels <- as.character(strq2$strength)
# labels[7] <- str_pad(labels[7], side = "left", width = 2)
# edgenames <- c("pi~stiffness_2", "gender~stiffness_2", "FFFS_1~injured_2",
#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
# names(labels) <- edgenames
# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
#                                     "nlelg_2~injured_2" = 6,
#                                     "clevel~balance_1" = 3,
#                                     "gender~stiffness_2" = 4),
#                             lty = c("FFFS_1~injured_2" = "dashed",
#                                     "injured_2~balance_2" = "dashed",
#                                     "pi~stiffness_2" = "dashed",
#                                     "injured_2~rmssd_2" = "dashed"))
# 
# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
# graph.par(list(edges=list(fontsize = 13)))
# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
# renderGraph(q2p)
```

\begin{figure}[H]
\caption{{\bf Change network.}
Network structure of the changes within variables between time points.}
\label{fig:fig9}
\end{figure}

Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in RMSSD where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).

```{r table10, results="asis"}

bmtable %>%
  kable(
    .,
    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)

```

```{r}
set.seed(1005)
test <- expand.grid(list(c(-1:1),
                         c(-1:1)))
query6 <- map2_dfr(.x = as.numeric(test$Var1),
                   .y = as.numeric(test$Var2), ~tibble(prob= cpquery(fittedcont, 
                                                                     event = (injured == "injured"),
                                                                     evidence = list(nlec = .x,  stiffness = .y), method = "lw"),
                                                       nle = .x,
                                                       stiffness = .y))

query6a <- query6 %>%
  dendroTools::round_df(2)
```

The Markov blanket for the "injured" node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
For stiffness and nlec the values in the nodes represent the standardised change between time point.
Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of `r query6a[9,1]` over the duration of the study.
With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.

```{r}
# plotsg3 <- model2network('[gender][pi][hours][injured|pi:hours][stiffness|gender:injured][nlec|injured]')
# q3p <- graphviz.plot(plotsg3,
#                      render = F,
#                      shape = "rectangle",
#                      groups = list(c("gender", "pi", "hours"), "injured", c("stiffness", "nlec")))
# strq3 <- strchange %>% 
#   filter(from %in% c("pi", "hours", "injured", "gender") & 
#            to %in% c("stiffness", "nlec", "injured")) %>% 
#   filter(strength > 0.30) %>% 
#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
# 
# labels <- as.character(strq3$strength)
# edgenames <- c("pi~injured", "gender~stiffness", "hours~injured",
#                "injured~stiffness", "injured~nlec")
# names(labels) <- edgenames
# nodeRenderInfo(q3p)$fill[c("injured")] = "gold"
# edgeRenderInfo(q3p) <- list(lwd = c("injured~stiffness" = 4, 
#                                     "hours~injured" = 5,
#                                     "gender~stiffness" = 4),
#                             lty = c("pi~injured" = "dashed"))
# 
# q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
# graph.par(list(edges=list(fontsize = 13)))
# nodeRenderInfo(q3p)$fill[c("nlec", "stiffness")] = "tomato"
# nodeRenderInfo(q3p)$fill["injured"] = "gold"
# renderGraph(q3p)
# rm(plotsg3, strq3, labels)

```

\begin{figure}[H]
\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
\label{fig:fig10}
\end{figure}

```{r table11, results="asis" }
set.seed(1006)
test <- expand.grid(list(
  c(-1:1),
  c(-1:1)
))
query6 <- map2_dfr(
  .x = as.numeric(test$Var1),
  .y = as.numeric(test$Var2),
  ~ tibble(
    prob = cpquery(
      fittedcont,
      event = (injured == "injured"),
      evidence = list(nlec = .x, stiffness = .y),
      method = "lw"
    ),
    nle = .x,
    stiffness = .y
  )
)

query6 %>%
  select(nle, stiffness, prob) %>%
  mutate_at(vars(nle, stiffness), ~ case_when(
    . == 1 ~ "+1SD",
    . == 0 ~ "mean",
    . == -1 ~ "-1SD"
  )) %>%
  arrange(-prob) %>%
  mutate(prob = round(prob, digits = 2)) %>%
  select(c(3, 1, 2)) %>%
  kable(
    .,
    col.names = c("Probability", "nlec", "Stiffness"),
    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)

```

Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).

```{r table12, results="asis"}
set.seed(1007)
testt <- expand.grid(list(
  c(-1:1),
  c(-1:1),
  c("Low", "High"),
  c("injury", "no injury")
))

query7 <- pmap_dfr(
  .l = testt,
  ~ tibble(
    prob = cpquery(
      fittedcont,
      event = (injured == "injured"),
      evidence = list(
        nlec = as.numeric(..1),
        stiffness = as.numeric(..2),
        hours = ..3,
        pi = ..4
      ),
      method = "lw"
    ),
    nle = ..1,
    stiffness = ..2,
    hours = ..3,
    pi = ..4
  )
) %>%
  toptail(3) %>%
  arrange(-prob)

query7 %>%
  mutate_at(vars(nle, stiffness), ~ case_when(
    . == 1 ~ "+1SD",
    . == 0 ~ "mean",
    . == -1 ~ "-1SD"
  )) %>%
  select(prob, hours, pi, nle, stiffness) %>%
  rename(Probability = "prob") %>%
  kable(
    .,
    caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
  ) %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

# Discussion

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

The statement about the authors and contributors can be up to several sentences
long, describing the tasks of individual authors referred to by their initials
and should be included at the end of the manuscript before the References
section.

# Acknowledgments {-}

Funding:

# Supplemental Data

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A reference list should be automatically created here. However it won't. Pandoc 
will place the list of references at the end of the document instead. There are 
no convenient solution for now to force Pandoc to do otherwise. The easiest way 
to get around this problem is to edit the LaTeX file created by Pandoc before 
compiling it again using the traditional LaTeX commands.

# Figures {-}

```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

### 2020-05-23 full_paper_v2.tex
initial

--- 2020-05-23 demo.tex
@@
 
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
 \def\Authors{
@@
 This paper adopts an interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
+At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 
-While the psychosocial factors proposed in Williams and Andersen's
-(Williams and Andersen, 1998) model have received the most research
-attention, the mechanisms through which these factors are proposed to
-exert their effect have remained under-investigated in the literature.
-To elaborate, the model suggests that injuries are likely to occur
-through either increased physiological arousal resulting in increased
-muscle tension and reduced flexibility or attentional deficits caused by
-increased distractibility and peripheral narrowing. However, to date,
-the research has largely focused on attentional deficits (Andersen and
-Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et
-al., 2007). For example, Andersen and Williams (Andersen and Williams,
-1999) measured peripheral and central vision during high and low stress
-conditions and found athletes with high life event stress coupled with
-low social support had greater peripheral narrowing under stressful
-conditions compared to athletes with the opposing profile; these
-athletes went on to sustain an increased number of injuries during the
-following season. Indeed, Rodgers and Landers (Rogers and Landers, 2005)
-supported Andersen and Williams's (Andersen and Williams, 1999) earlier
-findings reporting that peripheral narrowing under stress mediated 8.1\%
-of the relationship between negative life events and injury. However,
-the remaining variance between negative life events and athletic injury
-through the other proposed mechanisms, such as increased muscle tension
-and reduced motor control, remains to be explored (cf.~Williams and
-Andersen, 1998).
+While the psychosocial factors proposed in Williams and Andersen (1998)
+have received
 
-One challenge faced by researchers addressing the sports injury problem
-with a psychological lens is the multifactorial nature of injury, and
-the possible contribution of other non-psychological factors to the
-stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009). For
-example, a large body of research indicates that training-related stress
-is also likely to be related to the stress response and injury
+Research that has explored the effects of the psychosocial factors in
+Williams and Andersen's (Williams and Andersen, 1998) model, has failed
+to examine the mechanisms through which these factors are proposed to
+exert their effect. To elaborate, the model suggests that injuries are
+likely to occur through either increased physiological arousal resulting
+in increased muscle tension and reduced flexibility or attentional
+deficits caused by increased distractibility and peripheral narrowing.
+However, to date, the research has largely focused on attentional
+deficits (Andersen and Williams, 1999; Rogers and Landers, 2005;
+Wilkerson, 2012; Swanik et al., 2007). For example, Andersen and
+Williams (Andersen and Williams, 1999) measured peripheral and central
+vision during high and low stress conditions and found athletes with
+high life event stress coupled with low social support had greater
+peripheral narrowing under stressful conditions compared to athletes
+with the opposing profile; these athletes went on to sustain an
+increased number of injuries during the following season. Indeed,
+Rodgers and Landers (Rogers and Landers, 2005) supported Andersen and
+Williams's (Andersen and Williams, 1999) earlier findings reporting that
+peripheral narrowing under stress mediated 8.1\% of the relationship
+between negative life events and injury. However, few attempts have been
+made to explain the remaining variance between negative life events and
+athletic injury through the other proposed mechanisms, such as increased
+muscle tension and reduced motor control (cf.~Williams and Andersen,
+1998).
+
+One possible reason for this oversight is the multifactorial nature of
+injury and the possible contribution of other non-psychological factors
+to the stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009).
+For example, a large body of research indicates that training-related
+stress is also likely to be related to the stress response and injury
 occurrence (Lee et al., 2017; Djaoui et al., 2017), and may account for
@@
 postural stability (Romero-Franco et al., 2014) and muscle stiffness
-(Pruyn et al., 2015). However, these markers are often studied in
+(Pruyn et al., 2015). Unfortunately, these markers are often studied in
 isolation without an assessment of the psychosocial factors that are
@@
 ways, but result in a global outcome pattern of either adaptation or
-injury.
-
-A challenge when adopting a complex systems approach is using an
-appropriate analysis technique that is able to capture the uncertainty
-and complexity of the relationships between different factors. One
-technique that provides a solution to this challenge is Bayesian network
-(BN) modelling. BN's allow the construction of graphical probabilistic
-models using the underlying structure that connects different variables
-(nodes in the network) of interest (Scutari and Denis, 2014). Once
-constructed, the learned structure can be used for inference by
-obtaining the posterior probabilities of a particular node for a given
-query (e.g., if the value of Node A is x and the value of Node B is y,
-what is the probability Node C being value z?). Furthermore, BN's do not
-distinguish between dependent and independent variables as they are a
-form of unsupervised learning, which is a strength over regression or
-structure equation models when the underlying relationship in the
-network may not be known (Olmedilla et al., 2018).
+injury. Complex systems approaches have been used in health care to
+model the large number of risk factors associated with different types
+of diseases (Plsek and Greenhalgh, 2001); however, very few studies have
+attempted to address sport injury occurrence using such an approach
+(Hulme et al., 2018).
 
@@
 research stemming from Williams and Andersen's (Williams and Andersen,
-1998) model, there remains and opportunity to explore other
-physiological stress-related markers proposed by the BMSAIH, in addition
-to the already well-established psychological characteristics known to
-be related to injury (Appaneal and Perna, 2014). Furthermore, research
-has typically not captured changes in both psychosocial factors and
-stress-related physiological markers that may occur between initial
-measurement and injury occurrence. Given the exploratory the current
-study had the following objectives:
+1998) model, researchers have largely overlooked the potential to
+explore other physiological stress-related markers proposed by the
+BMSAIH, in addition to the already well-established psychological
+characteristics known to be related to injury (Appaneal and Perna,
+2014). Furthermore, research has typically not captured changes in both
+psychosocial factors and stress-related physiological markers that may
+occur between initial measurement and injury occurrence. Viewed through
+the lens of a complex systems approach the interaction between
+psychosocial sources of stress, stress-related physiological markers and
+injury occurrence may provide new insight into the injury process.
+Therefore, the purpose of the current study was to examine the
+relationship between psychosocial factors, physiological stress-related
+markers and occurrence of injury in athletes over time.
 
-\begin{itemize}
-\tightlist
-\item
-  Identify suitable markers of stress that can be easily captured in a
-  large cohort of athletes in a field based setting.
-\item
-  Examine the relationships between the markers of stress and injury
-  using a prospective, repeated measures design.
-\item
-  Evaluate the relationships between the markers of stress and injury
-  using Bayesian network modelling that captures the complex nature of
-  injury occurrence.
-\end{itemize}
+\hypertarget{results}{%
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
+
+\hypertarget{subsection-1}{%
+\subsection*{Subsection 1}\label{subsection-1}}
+\addcontentsline{toc}{subsection}{Subsection 1}
+
+You can use \texttt{R} chunks directly to plot graphs.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{x <-}\StringTok{ }\DecValTok{0}\OperatorTok{:}\DecValTok{100}
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{999}\NormalTok{)}
+\NormalTok{y <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(x }\OperatorTok{+}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{sd =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}\StringTok{ }\DecValTok{3}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(x, y)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{subsection-2}{%
+\subsection*{Subsection 2}\label{subsection-2}}
+\addcontentsline{toc}{subsection}{Subsection 2}
+
+Frontiers requires figures to be submitted individually, in the same
+order as they are referred to in the manuscript. Figures will then be
+automatically embedded at the bottom of the submitted manuscript. Kindly
+ensure that each table and figure is mentioned in the text and in
+numerical order. Permission must be obtained for use of copyrighted
+material from other sources (including the web). Please note that it is
+compulsory to follow figure instructions. Figures which are not
+according to the guidelines will cause substantial delay during the
+production process.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{disclosureconflict-of-interest-statement}{%
+\section*{Disclosure/Conflict-of-Interest
+Statement}\label{disclosureconflict-of-interest-statement}}
+\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
+Statement}
+
+The authors declare that the research was conducted in the absence of
+any commercial or financial relationships that could be construed as a
+potential conflict of interest.
+
+\hypertarget{author-contributions}{%
+\section*{Author Contributions}\label{author-contributions}}
+\addcontentsline{toc}{section}{Author Contributions}
+
+The statement about the authors and contributors can be up to several
+sentences long, describing the tasks of individual authors referred to
+by their initials and should be included at the end of the manuscript
+before the References section.
+
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
+
+Funding:
+
+\hypertarget{supplemental-data}{%
+\section{Supplemental Data}\label{supplemental-data}}
+
+Supplementary Material should be uploaded separately on submission, if
+there are Supplementary Figures, please include the caption in the same
+file as the figure. LaTeX Supplementary Material templates can be found
+in the Frontiers LaTeX folder
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+A reference list should be automatically created here. However it won't.
+Pandoc will place the list of references at the end of the document
+instead. There are no convenient solution for now to force Pandoc to do
+otherwise. The easiest way to get around this problem is to edit the
+LaTeX file created by Pandoc before compiling it again using the
+traditional LaTeX commands.
+
+\hypertarget{figures}{%
+\section*{Figures}\label{figures}}
+\addcontentsline{toc}{section}{Figures}
+
+\begin{figure}
+
+{\centering \includegraphics[width=85mm,height=85mm]{full_paper_v2_files/figure-latex/Figure-1-1}
+
+}
+
+\caption{Figure caption}\label{fig:Figure-1}
+\end{figure}
 
@@
 doi:\href{https://doi.org/10.1080/15438627.2014.881821}{10.1080/15438627.2014.881821}.
+
+\leavevmode\hypertarget{ref-Hulme2018}{}%
+Hulme, A., Thompson, J., Nielsen, R. O., Read, G. J. M. M., and Salmon,
+P. M. (2018). Towards a complex systems approach in sports injury
+research: Simulating running-related injury development with agent-based
+modelling. \emph{British Journal of Sports Medicine} 53, 560--569.
+doi:\href{https://doi.org/10.1136/bjsports-2017-098871}{10.1136/bjsports-2017-098871}.
 
@@
 
-\leavevmode\hypertarget{ref-Olmedilla2018}{}%
-Olmedilla, A., Rubio, V. J., Fuster-Parra, P., Pujals, C., and
-García-Mas, A. (2018). A Bayesian approach to sport injuries likelihood:
-Does player's self-efficacy and environmental factors plays the main
-role? \emph{Frontiers in Psychology} 9, 1--10.
-doi:\href{https://doi.org/10.3389/fpsyg.2018.01174}{10.3389/fpsyg.2018.01174}.
-
 \leavevmode\hypertarget{ref-Passer1983a}{}%
@@
 doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}.
+
+\leavevmode\hypertarget{ref-Plsek2001}{}%
+Plsek, P. E., and Greenhalgh, T. (2001). The challenge of complexity in
+health care. \emph{BMJ} 323, 625--628.
+doi:\href{https://doi.org/10.1136/bmj.323.7313.625}{10.1136/bmj.323.7313.625}.
 
@@
 doi:\href{https://doi.org/10.1590/1413-78522014220601007}{10.1590/1413-78522014220601007}.
-
-\leavevmode\hypertarget{ref-Scutari2014}{}%
-Scutari, M., and Denis, J.-B. (2014). \emph{Bayesian networks: with
-examples in R}. 1st ed. Chapman \& Hall/CRC Available at:
-\url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}.
 

### 2020-05-23 full_paper_v2/full_paper_v2.Rmd
initial

--- 2020-05-23 full_paper_v2.Rmd
@@
 csl: frontiers.csl
-bibliography: library.bib
+bibliography: ../library.bib
 ---
@@
 
-This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
+This paper adopts an interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
+At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 \end{abstract}
-
-```{r include=FALSE}
-options(scipen=999, digits = 2)
-knitr::opts_chunk$set(echo = FALSE)
-source("r-files/functions.R")
-source("r-files/methods_code.R")
-source("r-files/results_code.R")
-```
 
@@
 
-While the psychosocial factors proposed in Williams and Andersen’s [@Williams1998] model have received the most research attention, the mechanisms through which these factors are proposed to exert their effect have remained under-investigated in the literature.
+While the psychosocial factors proposed in @Williams1998 have received
+
+Research that has explored the effects of the psychosocial factors in Williams and Andersen’s [@Williams1998] model, has failed to examine the mechanisms through which these factors are proposed to exert their effect.
 To elaborate, the  model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
@@
 Indeed, Rodgers and Landers [@Rogers2005] supported Andersen and Williams's [@Andersen1999] earlier findings reporting that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
-However, the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control, remains to be explored [cf.
+However, few attempts have been made to explain the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control [cf.
 @Williams1998].
 
-One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
+One possible reason for this oversight is the multifactorial nature of injury and the possible contribution of other non-psychological factors to the stress response
 [@Meeuwisse2007; @Wiese-Bjornstal2009].
@@
 and muscle stiffness [@Pruyn2015].
-However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
+Unfortunately, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
 Recently, Bittencourt et al. [@Bittencourt2016]
@@
 where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+Complex systems approaches have been used in health care to model the large number of risk factors  associated with different types of diseases [@Plsek2001];
+however, very few studies have attempted to address sport injury occurrence using such an approach [@Hulme2018].
 
-A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables.
-One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014].
-The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
-Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
-
-To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
+To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, researchers have largely overlooked the potential to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
 Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
-Given the exploratory nature of the current study the following objectives were defined:
-
-- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
-- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
-- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.
-
-# Methods {-}
-
-```{r methodssetup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
-
-## Participants
-
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
-Participants self-rated competitive level ranged from recreational to international standard. 
-A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at least five hours per week.
-Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
-
-```{r table1, results='asis'}
-
-kable(
-  npars %>% 
-    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
-    slice(c(1,3,8, 4, 9:11)) %>% 
-  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
-                          name == "Height" ~ "Height (cm)",
-                          name == "Weight" ~ "Body mass (kg)",
-                          TRUE ~ name)),
-  caption = "Participant characteristics.",
-  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
-  align = "c",
-  escape = F
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Demographics M (SD)", 1, 4) %>%
-  pack_rows("Current competitive level n (%)", 5, 7) %>%
-  row_spec(0, bold = TRUE)
-```
-
-## Measures
-
-### Major life events
-
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
-The LESCA comprises 69 items that reflect possible life events that participants may have experienced. 
-Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
-Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
-A score for total life events was calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. 
-Petrie also provided evidence of discriminant, convergent and predictive validity. 
-The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
-The composite reliability for the LESCA in this study was 0.84.
-
-### Reinforcement Sensitivity Theory Personality Questionnaire
-
-A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
-The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
-Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
-The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
-Further details regarding the revised RST are in S1 Appendix.
-
-### Heart rate variability
-
-A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
-In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].
-
-### Musculoskeletal properties
-
-A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure muscle stiffness.
-The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
-The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
-The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016]. 
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].
-
-### Postural stability
-
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
-In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
-non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
-To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
-Participants were asked to maintain each stance for a total of 20-seconds.
-Participants hands were placed on hips at the level of the iliac crests. 
-A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
-
-The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
-One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
-b) opening eyes; 
-c) stepping, stumbling, or falling; 
-d) moving the thigh into more than 30 degrees of flexion or abduction; 
-e) lifting the forefoot or heel; and 
-f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
-A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
-A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
-To calculate limb asymmetry, the DL and NDL leg score was calculated by summing the DL and NDL errors respectively.
-A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
-To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
-The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.
-
-### Injury
-
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
-Injury status (did / did not sustain an injury) served as the main outcome measure.
-
-## Procedure
-
-At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
-With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
-To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week.
-Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-
-A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
-Each participant was asked to attend a total of four data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
-Participants provided informed consent before data collection commenced.
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
-
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
-To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
-The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
-Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
-
-\begin{figure}[!h]
-\caption{{\bf Session protocol.}
-Outline of the protocol for each data collection.}
-\label{fig:fig4}
-\end{figure}
-
-### Questionnaires
-
-The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed  on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
-The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
-For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis purposes.
-
-### HRV
-
-To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
-Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
-Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines. 
-Participants were seated and asked to remain as still as possible for the duration of the recording. 
-No attempt was made to control participants respiratory frequency or tidal volume [@Denver2007]. 
-Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
-
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices.
-Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996].
-RMSSD was calculated as:
-
-\begin{equation} 
-  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
-\end{equation}
-
-Where N is the length of the time series, and $RR_i$ the RR interval between beats $i$ and $i-1$, where each beat position corresponds to the beat detection instant.
-
-### Muscle stiffness
-
-To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb. 
-The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique  to determine the exact location of each site [@Chuang2012].
-The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle. 
-The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
-A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.
-
-After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site. 
-A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one second interval [@Morgan2018].
-For each impulse, the device computed stiffness values, with the median of the five values being saved by the device for further analysis. 
-In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted. 
-Sets of measurements above 3 % were measured again to ensure the reliability of the data. 
-The CV was calculated in real time by the device after each set of measurements. 
-Measurements saved on the device were uploaded to a computer using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis. 
-For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis.
-
-### Postural stability
-
-Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
-For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds. 
-Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
-The video recording was started prior to the first stance position and stopped after all stances had been completed.
-Each completed mBESS protocol took approximately 4 minutes.
-Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
-The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
-
-## Data Analysis
-
-Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
-A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988].
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
-The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012].
-In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
-
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
-
-Learning the structure of the network is an important step in BN modelling.
-The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
-The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
-The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014].
-In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
-In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
-BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
-
-[^1]: The term "query" in relation to Bayesian Networks stems from Pearl's expert systems theory (1988).
-A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
-Widely used texts on Bayesian Network analysis [@Koller2009] have widely adopted the terminology in favour of that used in traditional statistics.
-
-BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
-For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
-Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
-In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure.
-Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
-The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.
-
-### Data Preparation
-
-To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
-The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
-A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
-In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
-A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
-The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].
-
-Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
-Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
-All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.
-
-For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
-Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
-
-### Network structure
-
-To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
-Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist).
-Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
-To account for the repeated measures design and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
-In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
-However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
-Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
-Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
-Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
-Table \ref{tab:table2} provides an example of the formatted data. 
-Participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
-Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
-In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
-Finally, arc direction was restricted between specific nodes within the explanatory variables. 
-Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
-All subsequent models used the same blacklist.
-
-```{r table2, results='asis'}
-exampledata <- tibble(
-  Participant = c("1", "1", "2", "3", "3"),
-  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-  " " = "->",
-  "X_2" = c("T2", "T3", "T2", "T2", "T3")
-)
-
-kable(
-  exampledata,
-  caption = "Example of the data arrangement used for the network.",
-  escape = TRUE
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-```
-
-### Preliminary network structures
-
-Prior to the final network structure presented in the results section, several structures were investigated.
-Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
-A higher BIC value indicates the structure of a DAG is a better fit to the observed data [@Scutari2014]. 
-BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
-
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
-Additionally, despite strong evidence in the literature that both negative and total life event stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
-Total life event score was not included in the final structure.
-
-The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
-Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
-RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
-Therefore, only RI was included in the final structure.
-
-Finally, both total score and asymmetry were included in the initial network.
-However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
-Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
-
-\begin{table}[H]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{\label{tab:table3}Variables included in the final network structure.}
-\centering
-\begin{tabular}{l|l|l|l}
-\hline
-{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
-\hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Sustained previous injury & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
-\hline
-FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
-\hline
-BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
-\hline
-RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
-\hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
-\hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
-\hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
-For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
-Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
-The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
-To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
-Independent variables were then standardized to allow relative changes between variables to be compared. 
-The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
-Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
-The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
-
-To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
-The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
-Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
-An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
-In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
-Arcs below 0.3 were not included in the model.
-The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
-
-### Network Inference
-
-Conditional probability queries (CPQ) were used to perform inference on both network structures.
-To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
-CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
-The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
-Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
-Of particular interest in the current study were the variables that were connected to "injured" nodes.
-To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
-A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
-
-The second network contained both continuous and discrete data. 
-To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
-Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
+Viewed through the lens of a complex systems approach the interaction between psychosocial sources of stress, stress-related physiological markers and injury occurrence may provide new insight into the injury process.
+Therefore, the purpose of the current study was to examine the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes over time.
 
@@
 
-# Results
+<!-- Comments should be HTML-style comments -->
 
-During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
-Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
+## Subsection 1 {-}
 
-```{r table4, results='asis'}
+You can use `R` chunks directly to plot graphs.
 
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
-  kable(
-    .,
-    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
-    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
-    escape = F
-  ) %>%
-  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
-  kable_styling(latex_options = "HOLD_position")  %>%
-  row_spec(0,bold=TRUE)
-
+```{r graph, echo = TRUE, message=FALSE, fig.show = "hide"}
+x <- 0:100
+set.seed(999)
+y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
+plot(x, y)
 ```
 
-```{r mainnetworkresults}
-set.seed(999)
-strengthfinder <- function(var1, var2){
-  str %>%
-    filter(from == var1 & to == var2)
-}
+## Subsection 2 {-}
 
-from <- c("nlebase", "ind_team", "ind_team", "gender", "gender", "clevel", "pi")
-to <- c("RI_1", "hours", "nlebase", "stiffness_1", "stiffness_2", "balance_1", "stiffness_1")
-x <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-
-from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
-to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")
-
-y <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-```
-
-### First network structure
-
-The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
-Several of the explanatory variables showed strong connections with independent variables in the network.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
-The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
-Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
-Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
-The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
-The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
-In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
-For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
-### Markov blanket for injured_1
-
-```{r mb1}
-
-set.seed(1001)
-var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")
-
-query1 <- map_dfr(var, ~ newprobtable(
-  tallc,
-  .x,
-  "injured_1",
-  "injured",
-  avg30
-)) %>%
-  gather(variable, state, -prob, -outcome) %>%
-  drop_na() %>%
-  select(variable, state, prob) %>%
-  spread(state, prob) %>%
-  select(variable, Low, High) %>%
-  dendroTools::round_df(2)
-
-query2 <- newprobtable(
-  tallc,
-  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
-  "injured_1",
-  "injured",
-  avg30
-) %>%
-  toptail(n = 3) %>%
-  arrange(-prob) %>%
-  select(outcome, everything()) %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
-The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
-
-```{r}
-# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
-# q1p <- graphviz.plot(plotsg1,
-#                      shape = "rectangle",
-#                      render = F,
-#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
-# 
-# strq1 <- str %>% 
-#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") & 
-#            to %in% c("balance_1", "injured_1")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq1$strength)
-# labels[4] <- str_pad(labels[4], side = "left", width = 2)
-# 
-# edgenames <- edgeNames(q1p)
-# names(labels) <- edgenames
-# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
-# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6, 
-#                                     "stiffness_1~injured_1" = 3,
-#                                     "clevel~balance_1" = 3),
-#                             lty = c("injured_1~balance_1" = "dashed",
-#                                     "hours~injured_1" = "dashed"))
-# 
-# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 12)))
-# 
-# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
-# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
-# renderGraph(q1p)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-```{r table5, results='asis'}
-query1 %>%
-  select(c(1, 2, 3)) %>%
-  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
-Negative life events had a negligible effect when moving from the "Low" to "High" state.
-
-The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
-The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
-When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
-Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
-In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.
-
-```{r table6, results='asis'}
-
-query2 %>%
-  select(7, 2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
-The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}. 
-Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.
-
-```{r table7, results="asis"}
-injured1comp %>%
-  filter(variable != "outcome") %>% 
-  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(  .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
-    col.names = c("Variable", "High risk", "Low risk"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-### Markov blanket of injured_2
-
-```{r}
-set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2")
-query3 <- map_dfr(var2, ~newprobtable(tallc,
-                                      .x,
-                                      "injured_2",
-                                      "injured",
-                                      avg30)) %>% 
-  gather(variable, state, -prob, - outcome) %>% 
-  drop_na() %>% 
-  select(variable, state, prob) %>% 
-  spread(state, prob) %>% 
-  select(variable, Low, High) %>% 
-  dendroTools::round_df(2)
-
-query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
-                       "injured_2",
-                       "injured",
-                       avg30
-) %>% 
-  toptail(3) %>% 
-  arrange(-prob) %>% 
-  select(outcome, everything()) %>% 
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
-The arc between stiffness_2 and injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
-Results of the first query for injured_2 in the "injured" state are presented in table \ref{tab:table8}.
-Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
-FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
-"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
-
-```{r}
-# plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & 
-#            to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[6] <- str_pad(labels[6], side = "left", width = 2)
-# edgenames <- c("gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
-
-```{r table9, results="asis" }
-
-query4 %>%
-  select(7,2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-## Second network structure - changes within variables
-
-```{r include=FALSE, warning=FALSE}
-set.seed(1003)
-bmtable <- tidy(bm1, effects = "fixed") %>% 
-  select(term:conf.high) %>% 
-  dendroTools::round_df(2) %>% 
-  mutate_at(vars(conf.low, conf.high), list(~format(round(., 2), nsmall = 2))) %>%
-  mutate(conf.low = str_c("[", conf.low, sep = ""),
-         conf.high = str_c(conf.high, "]", sep = "")) %>% 
-  unite(., "95% CI", c("conf.low", "conf.high"), sep = ", ") %>% 
-  mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
-         "95% CI" = str_replace_all(`95% CI`, ",", ", ")) %>% 
-  slice(1:4) %>% 
-  mutate(term = c("Intercept", "RMSSD", "BIS", "RMSSD:BIS"))
-
-```
-
-The network for changes within variables is presented in Fig \ref{fig:fig9}.
-An arc between BIS $\rightarrow$ FFFS with strength 1 was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strength of 0.60.
-The arc between RMSSD $\rightarrow$ FFFS was 0.56.
-The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a linear regression model.
-While the continuous nodes in the second network are already parameterised using linear regressions, it was not possible to investigate the interaction between explanatory variables.
-The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
-
-```{r}
-# mb(avg30, "injured_2")
-# plotsg2 <- model2network('[nlelg_2][gender][pi][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|pi:gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("pi", "FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   slice(2:8) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[7] <- str_pad(labels[7], side = "left", width = 2)
-# edgenames <- c("pi~stiffness_2", "gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "pi~stiffness_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in RMSSD where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table10, results="asis"}
-
-bmtable %>%
-  kable(
-    .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
-    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r}
-set.seed(1005)
-test <- expand.grid(list(c(-1:1),
-                         c(-1:1)))
-query6 <- map2_dfr(.x = as.numeric(test$Var1),
-                   .y = as.numeric(test$Var2), ~tibble(prob= cpquery(fittedcont, 
-                                                                     event = (injured == "injured"),
-                                                                     evidence = list(nlec = .x,  stiffness = .y), method = "lw"),
-                                                       nle = .x,
-                                                       stiffness = .y))
-
-query6a <- query6 %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for the "injured" node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
-Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of `r query6a[9,1]` over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
-
-```{r}
-# plotsg3 <- model2network('[gender][pi][hours][injured|pi:hours][stiffness|gender:injured][nlec|injured]')
-# q3p <- graphviz.plot(plotsg3,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi", "hours"), "injured", c("stiffness", "nlec")))
-# strq3 <- strchange %>% 
-#   filter(from %in% c("pi", "hours", "injured", "gender") & 
-#            to %in% c("stiffness", "nlec", "injured")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq3$strength)
-# edgenames <- c("pi~injured", "gender~stiffness", "hours~injured",
-#                "injured~stiffness", "injured~nlec")
-# names(labels) <- edgenames
-# nodeRenderInfo(q3p)$fill[c("injured")] = "gold"
-# edgeRenderInfo(q3p) <- list(lwd = c("injured~stiffness" = 4, 
-#                                     "hours~injured" = 5,
-#                                     "gender~stiffness" = 4),
-#                             lty = c("pi~injured" = "dashed"))
-# 
-# q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q3p)$fill[c("nlec", "stiffness")] = "tomato"
-# nodeRenderInfo(q3p)$fill["injured"] = "gold"
-# renderGraph(q3p)
-# rm(plotsg3, strq3, labels)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-```{r table11, results="asis" }
-set.seed(1006)
-test <- expand.grid(list(
-  c(-1:1),
-  c(-1:1)
-))
-query6 <- map2_dfr(
-  .x = as.numeric(test$Var1),
-  .y = as.numeric(test$Var2),
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(nlec = .x, stiffness = .y),
-      method = "lw"
-    ),
-    nle = .x,
-    stiffness = .y
-  )
-)
-
-query6 %>%
-  select(nle, stiffness, prob) %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  arrange(-prob) %>%
-  mutate(prob = round(prob, digits = 2)) %>%
-  select(c(3, 1, 2)) %>%
-  kable(
-    .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
-
-```{r table12, results="asis"}
-set.seed(1007)
-testt <- expand.grid(list(
-  c(-1:1),
-  c(-1:1),
-  c("Low", "High"),
-  c("injury", "no injury")
-))
-
-query7 <- pmap_dfr(
-  .l = testt,
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(
-        nlec = as.numeric(..1),
-        stiffness = as.numeric(..2),
-        hours = ..3,
-        pi = ..4
-      ),
-      method = "lw"
-    ),
-    nle = ..1,
-    stiffness = ..2,
-    hours = ..3,
-    pi = ..4
-  )
-) %>%
-  toptail(3) %>%
-  arrange(-prob)
-
-query7 %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
-  ) %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
+Frontiers requires figures to be submitted individually, in the same order as
+they are referred to in the manuscript. Figures will then be automatically
+embedded at the bottom of the submitted manuscript. Kindly ensure that each
+table and figure is mentioned in the text and in numerical order. Permission
+must be obtained for use of copyrighted material from other sources (including
+the web). Please note that it is compulsory to follow figure instructions.
+Figures which are not according to the guidelines will cause substantial delay
+during the production process.
 

### 2020-05-23 rst_extra.Rmd
initial

---
title: "S1 Appendix - RST-PQ Validation"
bibliography: library.bib
output: pdf_document
csl: plos.csl

header-includes:
  - \usepackage{float}
  - \usepackage{lmodern}
  - \usepackage[T1]{fontenc}
  - \usepackage{setspace}
  - \linespread{2.0}

knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "supplimentary") })

---

The validation of the RST-PQ was examined in a separate programme of PhD research which has yet to be published [@Young2019].
The present author was involved in the validation process as the participants from this study were used to validate the RST-PQ. 
Details of validation process are included here for detail but are not strictly part of this programme of research.

To establish the factorial validity of the 65-item model of the RST-PQ in a population of university sports students, Bayesian Structural Equation Modelling [BSEM; @Muthen2012] was used with responses from sample of 419 university-level athletes [@Young2019]. 
BSEM was used for the validation as it several advantages over the maximal likelihood procedures used in traditional confirmatory factor analysis (CFA). 
In particular, BSEM allows informative variance priors to be specified on cross loadings.
In traditional CFA, cross loadings are held at zero and can lead to a blocked or miss-specified model [@Marsh2009].
By recognising that some small cross-loadings and correlated residuals could be present within the items, BSEM can reduce the possibility of model misspecification [@Gucci2016].
To perform the validation, three models were specified following the recommendations of @Muthen2012.
Model one specified non-informative priors on the factor loadings and exact zeros on the cross-loadings and correlated residuals.
Model two specified non-informative priors on the factor loadings, informative approximate zeros on the cross-loadings and exact zeros on the correlated residuals.

Lastly, model three specified non-informative priors on the factor loadings, and informative approximate zeros on the cross-loadings and correlated residuals.
The variance of the informative priors was set at $\pm$ 0.10, which equates to loadings with 95% limit of $\pm$ .20 and implies weak cross loading and correlated residuals.
Loadings that were outside of the range were identified as being highly correlated or having a large source of residual variance [@Muthen2012].

Estimation of the BSEM models was performed in MPLUS where a Markov chain Monte Carlo (MCMC) algorithm was used with a Gibbs sampler, in which 100,000 iterations were drawn to examine the parameter estimates and model convergence. 
Model convergence was assessed using potential scale reduction factor (PSRF), where values between 1.0 and 1.1 indicated good model convergence (@Gelman1998; @Gelman2013).
Model fit was assessed by Posterior Predictive P (PPP) values and 95 % credibility intervals, where PPP values >= 0.50 and 95 % confidence interval balanced around approximately zero indicated good model fit [@Muthen2012]
To compare the different BSEM models, the deviance information criterion (DIC) was used, where lower values indicate a better model-fit.

The analysis revealed the 65 item RST-PQ to have both good model convergence and model fit. In addition, all the main factor loadings were significant.
However, several problematic items were identified based on weak factor loadings and substantial cross-loadings and correlated residuals.
Consequently, the model was re-specified following an item deletion process [@Young2019] and a 51-item RST-PQ emerged with acceptable model-fit (PPP = .739, 95% posterior predictive confidence intervals = -199.220, 101.356). 
The findings from the BSEM were replicated with a sample of 350 participants from the current research programme (PPP = .787, 95% posterior predictive confidence intervals = -208.405, 90.785).
Consequently, the 51-item questionnaire was adopted as the chosen measure of RST in the current research.

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
<div id="refs"></div>

### 2020-05-23 supplimentary/supplimentary_tables.tex
initial

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplimentary materials},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float}
\usepackage{longtable}
\pagenumbering{gobble}
\usepackage{caption}
\captionsetup[table]{labelformat=empty}
% https://github.com/rstudio/rmarkdown/issues/337
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

% https://github.com/rstudio/rmarkdown/pull/252
\usepackage{titling}
\setlength{\droptitle}{-2em}

\pretitle{\vspace{\droptitle}\centering\huge}
\posttitle{\par}

\preauthor{\centering\large\emph}
\postauthor{\par}

\predate{\centering\large\emph}
\postdate{\par}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Supplimentary materials}
\date{}

\begin{document}
\maketitle

\textbf{S1 Table. Summary of modifications to LESCA items.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-2}}
\centering
\begin{tabular}[t]{>{\raggedleft\arraybackslash}p{0.3cm}|>{\raggedright\arraybackslash}p{4.8cm}|>{\raggedright\arraybackslash}p{4.8cm}|>{\raggedright\arraybackslash}p{4.8cm}}
\hline
Q & Original & Students & Non-students\\
\hline
19 & Beginning a new school experience  (beginning college, transferring college etc) & Beginning a new school experience (beginning university, transferring university etc) & Beginning a new work experience\\
\hline
21 & Academic probation / ineligibility & \textit{No change} & Removed\\
\hline
22 & Being dismissed from dorm or other residence & Being dismissed from halls or other residence & Being asked to vacate house/home\\
\hline
27 & Financial problems concerning school & \textit{No change} & Financial problems\\
\hline
29 & Conflict with room-mate & \textit{No change} & Conflict within household\\
\hline
36 & Suspended from team for non-academic reasons & \textit{No change} & Removed\\
\hline
49 & Being absent from school (classes) because of participation in sport & Being absent from university (classes) because of participation in sport & Being absent from work because of participation in sport\\
\hline
61 & Major change in level of academic performance (doing better or worse) & \textit{No change} & Major change in level of performance at work (doing better or worse)\\
\hline
62 & Making career decisions (applying to graduate school,interviewing for jobs, etc) & Making career decisions (applying for Masters degree, interviewing for jobs, etc) & Making career decisions (applying for Masters degree, interviewing for jobs, etc)\\
\hline
\end{tabular}
\end{table}

\newpage

\textbf{S2 Table. All arc strengths greater than 0.3 included in the first network.}

\begin{longtable}[t]{l|l|r|r}
\caption{\label{tab:unnamed-chunk-3}}\\
\hline
from & to & strength & direction\\
\hline
\endfirsthead
\caption[]{ S3 Table continued}\\
\hline
from & to & strength & direction\\
\hline
\endhead
nlebase & nlelg\_1 & 1.00 & 1.00\\
\hline
nlebase & nlelg\_2 & 1.00 & 1.00\\
\hline
nlelg\_1 & injured\_1 & 1.00 & 1.00\\
\hline
nlelg\_2 & injured\_2 & 1.00 & 1.00\\
\hline
BIS\_1 & FFFS\_1 & 0.98 & 0.68\\
\hline
FFFS\_1 & BIS\_1 & 0.98 & 0.32\\
\hline
ind\_team & hours & 0.95 & 0.50\\
\hline
hours & ind\_team & 0.95 & 0.50\\
\hline
nlebase & RI\_1 & 0.83 & 1.00\\
\hline
ind\_team & nlebase & 0.78 & 1.00\\
\hline
BIS\_2 & FFFS\_2 & 0.73 & 0.71\\
\hline
FFFS\_2 & BIS\_2 & 0.73 & 0.29\\
\hline
gender & stiffness\_1 & 0.72 & 1.00\\
\hline
hours & rmssd\_1 & 0.66 & 1.00\\
\hline
gender & stiffness\_2 & 0.60 & 1.00\\
\hline
nlelg\_1 & BIS\_1 & 0.52 & 0.97\\
\hline
BIS\_1 & nlelg\_1 & 0.52 & 0.03\\
\hline
pi & stiffness\_1 & 0.51 & 1.00\\
\hline
injured\_2 & stiffness\_2 & 0.51 & 0.65\\
\hline
stiffness\_2 & injured\_2 & 0.51 & 0.35\\
\hline
clevel & balance\_1 & 0.49 & 1.00\\
\hline
FFFS\_1 & BIS\_2 & 0.49 & 1.00\\
\hline
nlebase & rmssd\_1 & 0.49 & 1.00\\
\hline
injured\_1 & stiffness\_1 & 0.46 & 0.34\\
\hline
stiffness\_1 & injured\_1 & 0.46 & 0.66\\
\hline
clevel & RI\_2 & 0.46 & 1.00\\
\hline
FFFS\_1 & injured\_2 & 0.44 & 1.00\\
\hline
hours & injured\_1 & 0.40 & 1.00\\
\hline
nlelg\_2 & BIS\_2 & 0.38 & 0.98\\
\hline
BIS\_2 & nlelg\_2 & 0.38 & 0.02\\
\hline
gender & FFFS\_1 & 0.37 & 1.00\\
\hline
injured\_2 & rmssd\_2 & 0.37 & 0.84\\
\hline
rmssd\_2 & injured\_2 & 0.37 & 0.16\\
\hline
pi & nlelg\_2 & 0.37 & 1.00\\
\hline
nlebase & RI\_2 & 0.36 & 1.00\\
\hline
RI\_1 & BIS\_1 & 0.34 & 0.44\\
\hline
BIS\_1 & RI\_1 & 0.34 & 0.56\\
\hline
injured\_2 & balance\_2 & 0.34 & 0.95\\
\hline
balance\_2 & injured\_2 & 0.34 & 0.05\\
\hline
injured\_1 & balance\_1 & 0.34 & 0.68\\
\hline
balance\_1 & injured\_1 & 0.34 & 0.32\\
\hline
gender & FFFS\_2 & 0.33 & 1.00\\
\hline
RI\_1 & balance\_1 & 0.32 & 0.21\\
\hline
balance\_1 & RI\_1 & 0.32 & 0.79\\
\hline
pi & clevel & 0.32 & 0.50\\
\hline
clevel & pi & 0.32 & 0.50\\
\hline
BIS\_1 & FFFS\_2 & 0.32 & 1.00\\
\hline
pi & ind\_team & 0.31 & 0.44\\
\hline
ind\_team & pi & 0.31 & 0.56\\
\hline
pi & stiffness\_2 & 0.31 & 1.00\\
\hline
ind\_team & FFFS\_1 & 0.30 & 1.00\\
\hline
\end{longtable}

\newpage

\textbf{S3 Table. All arc strengths greater than 0.3 included in the second network.}

\begin{longtable}[t]{l|l|r|r}
\caption{\label{tab:unnamed-chunk-4}}\\
\hline
from & to & strength & direction\\
\hline
BIS & FFFS & 1.00 & 0.75\\
\hline
FFFS & BIS & 1.00 & 0.25\\
\hline
hours & injured & 0.72 & 1.00\\
\hline
injured & stiffness & 0.66 & 1.00\\
\hline
rmssd & FFFS & 0.64 & 0.67\\
\hline
FFFS & rmssd & 0.64 & 0.33\\
\hline
clevel & BIS & 0.61 & 1.00\\
\hline
gender & stiffness & 0.57 & 1.00\\
\hline
injured & nlec & 0.45 & 1.00\\
\hline
pi & balance & 0.42 & 1.00\\
\hline
pi & injured & 0.38 & 1.00\\
\hline
RI & BIS & 0.36 & 0.35\\
\hline
BIS & RI & 0.36 & 0.65\\
\hline
stiffness & BIS & 0.34 & 0.92\\
\hline
BIS & stiffness & 0.34 & 0.08\\
\hline
\end{longtable}

\newpage

\textbf{S4 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-5}}
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
prob & clevel & hours & nlelg\_1 & stiffness\_1 & balance\_1\\
\hline
0.53 & club\_university\_county & High & Low & High & High\\
\hline
0.46 & national\_international & High & Low & High & Low\\
\hline
0.44 & national\_international & High & Low & High & High\\
\hline
0.39 & club\_university\_county & Low & High & High & High\\
\hline
0.37 & club\_university\_county & High & Low & High & Low\\
\hline
0.34 & club\_university\_county & High & High & High & High\\
\hline
0.32 & national\_international & Low & High & High & Low\\
\hline
0.3 & national\_international & Low & High & High & High\\
\hline
0.29 & club\_university\_county & Low & High & Low & High\\
\hline
0.29 & club\_university\_county & High & High & Low & High\\
\hline
0.27 & national\_international & High & High & High & Low\\
\hline
0.26 & national\_international & High & High & High & High\\
\hline
0.25 & club\_university\_county & Low & High & High & Low\\
\hline
0.24 & national\_international & Low & High & Low & Low\\
\hline
0.23 & national\_international & High & High & Low & Low\\
\hline
0.22 & national\_international & Low & High & Low & High\\
\hline
0.22 & national\_international & High & High & Low & High\\
\hline
0.21 & club\_university\_county & High & High & High & Low\\
\hline
0.18 & club\_university\_county & Low & High & Low & Low\\
\hline
0.18 & club\_university\_county & High & High & Low & Low\\
\hline
0.17 & club\_university\_county & Low & Low & High & High\\
\hline
0.16 & club\_university\_county & High & Low & Low & High\\
\hline
0.13 & national\_international & Low & Low & High & Low\\
\hline
0.12 & national\_international & High & Low & Low & Low\\
\hline
0.12 & national\_international & Low & Low & High & High\\
\hline
0.12 & national\_international & High & Low & Low & High\\
\hline
0.09 & club\_university\_county & Low & Low & High & Low\\
\hline
0.09 & club\_university\_county & High & Low & Low & Low\\
\hline
0.07 & club\_university\_county & Low & Low & Low & High\\
\hline
0.06 & national\_international & Low & Low & Low & Low\\
\hline
0.05 & national\_international & Low & Low & Low & High\\
\hline
0.04 & club\_university\_county & Low & Low & Low & Low\\
\hline
\end{tabular}
\end{table}

\newpage

\textbf{S5 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_2.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-6}}
\centering
\begin{tabular}[t]{r|l|l|l|l|l}
\hline
prob & FFFS\_1 & nlelg\_2 & stiffness\_2 & rmssd\_2 & balance\_2\\
\hline
0.53 & Low & Low & High & Low & High\\
\hline
0.45 & Low & High & High & Low & High\\
\hline
0.41 & Low & Low & High & High & High\\
\hline
0.39 & Low & Low & High & Low & Low\\
\hline
0.34 & Low & High & High & High & High\\
\hline
0.32 & Low & High & High & Low & Low\\
\hline
0.31 & Low & Low & Low & Low & High\\
\hline
0.29 & Low & Low & High & High & Low\\
\hline
0.28 & Low & High & Low & Low & High\\
\hline
0.25 & High & Low & High & Low & High\\
\hline
0.23 & Low & High & High & High & Low\\
\hline
0.22 & Low & Low & Low & High & High\\
\hline
0.21 & High & High & High & Low & High\\
\hline
0.21 & Low & Low & Low & Low & Low\\
\hline
0.19 & Low & High & Low & High & High\\
\hline
0.18 & Low & High & Low & Low & Low\\
\hline
0.17 & High & Low & High & High & High\\
\hline
0.16 & High & Low & High & Low & Low\\
\hline
0.14 & High & High & High & High & High\\
\hline
0.14 & Low & Low & Low & High & Low\\
\hline
0.13 & High & High & High & Low & Low\\
\hline
0.13 & High & Low & Low & Low & High\\
\hline
0.12 & Low & High & Low & High & Low\\
\hline
0.12 & High & High & Low & Low & High\\
\hline
0.11 & High & Low & High & High & Low\\
\hline
0.08 & High & High & High & High & Low\\
\hline
0.08 & High & Low & Low & High & High\\
\hline
0.08 & High & Low & Low & Low & Low\\
\hline
0.07 & High & High & Low & High & High\\
\hline
0.07 & High & High & Low & Low & Low\\
\hline
0.05 & High & Low & Low & High & Low\\
\hline
0.04 & High & High & Low & High & Low\\
\hline
\end{tabular}
\end{table}

\end{document}

### 2020-05-23 supplimentary_tables.Rmd
initial

---
title: Supplimentary materials
output: 
  pdf_document:
    keep_tex: yes
header-includes:
  - \usepackage{float} 
  - \usepackage{longtable}
  - \pagenumbering{gobble}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "supplimentary") })
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn)

source("r-files/functions.R")
source("r-files/methods_code.R")
source("r-files/results_code.R")
```

\textbf{S1 Table. Summary of modifications to LESCA items.}

```{r}
 tab <- data.frame(stringsAsFactors=FALSE,
 Question = c(19, 21, 22, 27, 29, 36, 49, 61, 62),
 original = c("Beginning a new school experience  (beginning college, transferring college etc)",
                "Academic probation / ineligibility",
                "Being dismissed from dorm or other residence",
                "Financial problems concerning school",
                "Conflict with room-mate",
                "Suspended from team for non-academic reasons",
                "Being absent from school (classes) because of participation in sport",
                "Major change in level of academic performance (doing better or worse)",
                "Making career decisions (applying to graduate school,interviewing for jobs, etc)"),
 "Sportsstudents" = c("Beginning a new school experience (beginning university, transferring university etc)",
                       NA,
                       "Being dismissed from halls or other residence", 
                       NA, 
                       NA, 
                       NA,
                       "Being absent from university (classes) because of participation in sport",
                     NA,
                     "Making career decisions (applying for Masters degree, interviewing for jobs, etc)"),
 "non-student" = c("Beginning a new work experience",
                   "Removed",
                   "Being asked to vacate house/home",
                   "Financial problems",
                   "Conflict within household",
                   "Removed",
                   "Being absent from work because of participation in sport",
                   "Major change in level of performance at work (doing better or worse)",
                   "Making career decisions (applying for Masters degree, interviewing for jobs, etc)"))

tab[is.na(tab)] <- "\\textit{No change}"

kable(
  tab %>%
    mutate(Sportsstudents = ifelse(str_detect(Sportsstudents, "-"), text_spec(Sportsstudents, align = "c"), Sportsstudents)),
  col.names = c("Q", "Original", "Students", "Non-students"),
  escape = F,
  caption = ""
) %>% 
  column_spec(1, width = c("0.3cm")) %>% 
   column_spec(2:4, width = ("4.8cm")) %>% 
  kable_styling(latex_options = "HOLD_position")
```

\newpage

\textbf{S2 Table. All arc strengths greater than 0.3 included in the first network.}

```{r results='asis'}
full <- str %>%
  filter(strength > 0.299 & direction != 0) %>%
  arrange(-strength) %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
   mutate_at(c("from", "to"), ~str_replace_all(., "_", "\\\\_")) %>% 
  dendroTools::round_df(2)

  kable(
    full,
    caption = "",
    longtable = T,
    escape = F,
    format = "latex"
  ) %>% 
  kable_styling(latex_options = c("HOLD_position" , "repeat_header"), repeat_header_text = "S3 Table continued")

```

\textbf{S3 Table. All arc strengths greater than 0.3 included in the second network.}

```{r results='asis'}
strchange %>%
  filter(strength > 0.299 & direction != 0) %>%
  arrange(-strength) %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
   mutate_at(c("from", "to"), ~str_replace_all(., "_", "\\\\_")) %>% 
  dendroTools::round_df(2) %>% 
  kable(
    caption = "",
    longtable = T,
    escape = F,
    format = "latex"
  ) %>% 
  kable_styling(latex_options = ("HOLD_position"))
```

\newpage

\textbf{S4 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}

```{r results='asis'}
set.seed(1001)

tab3 <- newprobtable(
  tallc,
  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
  "injured_1",
  "injured",
  avg30
) %>%
  arrange(-prob) %>%
  select(outcome, everything()) %>%
  dendroTools::round_df(2) %>% 
  select(7, 2:6) %>% 
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>% 
  mutate_all(~str_replace_all(., "_", "\\\\_"))

kable(
  tab3,
  caption = "",
    escape = F
) %>%
  kable_styling(latex_options = "HOLD_position")
```

\newpage

\textbf{S5 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_2.}

```{r results='asis'}

set.seed(1002)

tab4 <- newprobtable(
  tallc,
  c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
  "injured_2",
  "injured",
  avg30
) %>%
  arrange(-prob) %>%
  select(outcome, everything()) %>%
  dendroTools::round_df(2) %>%
  select(7, 2:6) %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_"))

  kable(
    tab4,
    caption = "",
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position")

```

### 2020-05-23 wordmaker.Rmd
initial

--- 2020-05-23 full_paper_v2.Rmd
@@
 ---
-title: "An interdisciplinary examination of stress and injury occurrence in athletes"
-runtitle: "Short Title"
-documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
-author:
-  - name: Dr Harry Fisher
-    affiliation: '1'
-    etal: Fisher # First author's last name. 
-    email: harryfisher21@gmail.com # Indicates corresponding Author
-    institution: Cardiff Metropolitan University
-    street: Cyncoed Road
-    zip: CF23 6XD
-    country: United Kingdom
-  - name: Dr Marianne Gittoes
-    affiliation: '1'
-  - name: Professor Lynne Evans
-    affiliation: '1'
-  - name: Miss Leah Bitchell
-    affiliation: '1'  
-  - name: Dr Richard Mullen  
-    affiliation: '2'
-  - name: Dr Marco Scutari
-    affiliation: '3'    
-affiliation:
-  - id: '1'
-    department: School of Sport and Health Science
-    institution: Cardiff Metropolitan University
-    city: Cardiff
-    country: United Kingdom
-  - id: '2' 
-    department: Division of Sport, Health & Exercise Sciences
-    institution: Brunel University
-    city: London
-    country: United Kingdom
-  - id: '3' 
-    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
-    city: Manno
-    country: Switzerland
-
-output: rticles::frontiers_article
-csl: frontiers.csl
-bibliography: library.bib
+title: "wordmaker"
+output: word_document
 ---
 
-\begin{abstract}
-
-This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
-Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
-Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury.
-Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury.
-In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury.
-Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.
-
-\tiny
- \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}
-
-\end{abstract}
-
-```{r include=FALSE}
-options(scipen=999, digits = 2)
-knitr::opts_chunk$set(echo = FALSE)
-source("r-files/functions.R")
-source("r-files/methods_code.R")
-source("r-files/results_code.R")
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-# Introduction {-}
 
@@
 @Williams1998].
-
-One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
-[@Meeuwisse2007; @Wiese-Bjornstal2009].
-For example, a large body of research indicates that training-related stress is also likely to be related to the stress response and injury occurrence [@Lee2017; @Djaoui2017],
-and may account for the unexplained variance from the psychological predictors of injury.
-Appaneal and Perna [@Appaneal2014]
-proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model and to address some of these issues.
-To elaborate, the BMSAIH aimed to clarify the mediating pathways between the stress response and injury, consider other health outcomes and behavioural factors that impact sports participation,
-and integrate the impact of training on athletes’ health [@Appaneal2014].
-The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p.
-74] to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides a framework for future research to build on Williams and Andersen's [@Williams1998] model, by including other physiological markers of training-related stress, which together may provide greater insight into the injury process.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other research has identified additional markers of training-related stress that are associated with an increased risk of injury; for example,
-heart rate variability [@Bellenger2016; @Williams2017],
-postural stability [@Romero-Franco2014]
-and muscle stiffness [@Pruyn2015].
-However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
-Recently, Bittencourt et al. [@Bittencourt2016]
-suggested that to better understand the multifactorial nature of sports injuries, research needs to move away from studying risk factors in isolation and instead adopt a complex systems approach to injury.
-Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p. 3],
-where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
-
-A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables.
-One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014].
-The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
-Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
-
-To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
-Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
-Given the exploratory nature of the current study the following objectives were defined:
-
-- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
-- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
-- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.
-
-# Methods {-}
-
-```{r methodssetup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
-
-## Participants
-
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
-Participants self-rated competitive level ranged from recreational to international standard. 
-A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at least five hours per week.
-Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
-
-```{r table1, results='asis'}
-
-kable(
-  npars %>% 
-    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
-    slice(c(1,3,8, 4, 9:11)) %>% 
-  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
-                          name == "Height" ~ "Height (cm)",
-                          name == "Weight" ~ "Body mass (kg)",
-                          TRUE ~ name)),
-  caption = "Participant characteristics.",
-  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
-  align = "c",
-  escape = F
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Demographics M (SD)", 1, 4) %>%
-  pack_rows("Current competitive level n (%)", 5, 7) %>%
-  row_spec(0, bold = TRUE)
-```
-
-## Measures
-
-### Major life events
-
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
-The LESCA comprises 69 items that reflect possible life events that participants may have experienced. 
-Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
-Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
-A score for total life events was calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. 
-Petrie also provided evidence of discriminant, convergent and predictive validity. 
-The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
-The composite reliability for the LESCA in this study was 0.84.
-
-### Reinforcement Sensitivity Theory Personality Questionnaire
-
-A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
-The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
-Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
-The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
-Further details regarding the revised RST are in S1 Appendix.
-
-### Heart rate variability
-
-A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
-In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].
-
-### Musculoskeletal properties
-
-A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure muscle stiffness.
-The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
-The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
-The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016]. 
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].
-
-### Postural stability
-
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
-In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
-non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
-To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
-Participants were asked to maintain each stance for a total of 20-seconds.
-Participants hands were placed on hips at the level of the iliac crests. 
-A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
-
-The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
-One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
-b) opening eyes; 
-c) stepping, stumbling, or falling; 
-d) moving the thigh into more than 30 degrees of flexion or abduction; 
-e) lifting the forefoot or heel; and 
-f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
-A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
-A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
-To calculate limb asymmetry, the DL and NDL leg score was calculated by summing the DL and NDL errors respectively.
-A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
-To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
-The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.
-
-### Injury
-
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
-Injury status (did / did not sustain an injury) served as the main outcome measure.
-
-## Procedure
-
-At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
-With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
-To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week.
-Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-
-A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
-Each participant was asked to attend a total of four data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
-Participants provided informed consent before data collection commenced.
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
-
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
-To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
-The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
-Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
-
-\begin{figure}[!h]
-\caption{{\bf Session protocol.}
-Outline of the protocol for each data collection.}
-\label{fig:fig4}
-\end{figure}
-
-### Questionnaires
-
-The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed  on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
-The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
-For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis purposes.
-
-### HRV
-
-To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
-Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
-Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines. 
-Participants were seated and asked to remain as still as possible for the duration of the recording. 
-No attempt was made to control participants respiratory frequency or tidal volume [@Denver2007]. 
-Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
-
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices.
-Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996].
-RMSSD was calculated as:
-
-\begin{equation} 
-  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
-\end{equation}
-
-Where N is the length of the time series, and $RR_i$ the RR interval between beats $i$ and $i-1$, where each beat position corresponds to the beat detection instant.
-
-### Muscle stiffness
-
-To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb. 
-The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique  to determine the exact location of each site [@Chuang2012].
-The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle. 
-The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
-A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.
-
-After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site. 
-A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one second interval [@Morgan2018].
-For each impulse, the device computed stiffness values, with the median of the five values being saved by the device for further analysis. 
-In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted. 
-Sets of measurements above 3 % were measured again to ensure the reliability of the data. 
-The CV was calculated in real time by the device after each set of measurements. 
-Measurements saved on the device were uploaded to a computer using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis. 
-For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis.
-
-### Postural stability
-
-Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
-For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds. 
-Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
-The video recording was started prior to the first stance position and stopped after all stances had been completed.
-Each completed mBESS protocol took approximately 4 minutes.
-Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
-The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
-
-## Data Analysis
-
-Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
-A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988].
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
-The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012].
-In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
-
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
-
-Learning the structure of the network is an important step in BN modelling.
-The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
-The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
-The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014].
-In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
-In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
-BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
-
-[^1]: The term "query" in relation to Bayesian Networks stems from Pearl's expert systems theory (1988).
-A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
-Widely used texts on Bayesian Network analysis [@Koller2009] have widely adopted the terminology in favour of that used in traditional statistics.
-
-BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
-For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
-Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
-In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure.
-Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
-The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.
-
-### Data Preparation
-
-To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
-The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
-A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
-In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
-A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
-The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].
-
-Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
-Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
-All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.
-
-For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
-Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
-
-### Network structure
-
-To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
-Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist).
-Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
-To account for the repeated measures design and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
-In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
-However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
-Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
-Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
-Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
-Table \ref{tab:table2} provides an example of the formatted data. 
-Participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
-Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
-In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
-Finally, arc direction was restricted between specific nodes within the explanatory variables. 
-Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
-All subsequent models used the same blacklist.
-
-```{r table2, results='asis'}
-exampledata <- tibble(
-  Participant = c("1", "1", "2", "3", "3"),
-  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-  " " = "->",
-  "X_2" = c("T2", "T3", "T2", "T2", "T3")
-)
-
-kable(
-  exampledata,
-  caption = "Example of the data arrangement used for the network.",
-  escape = TRUE
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-```
-
-### Preliminary network structures
-
-Prior to the final network structure presented in the results section, several structures were investigated.
-Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
-A higher BIC value indicates the structure of a DAG is a better fit to the observed data [@Scutari2014]. 
-BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
-
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
-Additionally, despite strong evidence in the literature that both negative and total life event stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
-Total life event score was not included in the final structure.
-
-The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
-Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
-RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
-Therefore, only RI was included in the final structure.
-
-Finally, both total score and asymmetry were included in the initial network.
-However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
-Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
-
-\begin{table}[H]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{\label{tab:table3}Variables included in the final network structure.}
-\centering
-\begin{tabular}{l|l|l|l}
-\hline
-{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
-\hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Sustained previous injury & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
-\hline
-FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
-\hline
-BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
-\hline
-RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
-\hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
-\hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
-\hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
-For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
-Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
-The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
-To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
-Independent variables were then standardized to allow relative changes between variables to be compared. 
-The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
-Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
-The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
-
-To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
-The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
-Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
-An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
-In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
-Arcs below 0.3 were not included in the model.
-The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
-
-### Network Inference
-
-Conditional probability queries (CPQ) were used to perform inference on both network structures.
-To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
-CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
-The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
-Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
-Of particular interest in the current study were the variables that were connected to "injured" nodes.
-To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
-A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
-
-The second network contained both continuous and discrete data. 
-To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
-Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-
-# Results {-}
-
-# Results
-
-During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
-Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
-
-```{r table4, results='asis'}
-
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
-  kable(
-    .,
-    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
-    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
-    escape = F
-  ) %>%
-  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
-  kable_styling(latex_options = "HOLD_position")  %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r mainnetworkresults}
-set.seed(999)
-strengthfinder <- function(var1, var2){
-  str %>%
-    filter(from == var1 & to == var2)
-}
-
-from <- c("nlebase", "ind_team", "ind_team", "gender", "gender", "clevel", "pi")
-to <- c("RI_1", "hours", "nlebase", "stiffness_1", "stiffness_2", "balance_1", "stiffness_1")
-x <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-
-from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
-to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")
-
-y <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-```
-
-### First network structure
-
-The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
-Several of the explanatory variables showed strong connections with independent variables in the network.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
-The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
-Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
-Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
-The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
-The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
-In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
-For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
-### Markov blanket for injured_1
-
-```{r mb1}
-
-set.seed(1001)
-var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")
-
-query1 <- map_dfr(var, ~ newprobtable(
-  tallc,
-  .x,
-  "injured_1",
-  "injured",
-  avg30
-)) %>%
-  gather(variable, state, -prob, -outcome) %>%
-  drop_na() %>%
-  select(variable, state, prob) %>%
-  spread(state, prob) %>%
-  select(variable, Low, High) %>%
-  dendroTools::round_df(2)
-
-query2 <- newprobtable(
-  tallc,
-  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
-  "injured_1",
-  "injured",
-  avg30
-) %>%
-  toptail(n = 3) %>%
-  arrange(-prob) %>%
-  select(outcome, everything()) %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
-The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
-
-```{r}
-# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
-# q1p <- graphviz.plot(plotsg1,
-#                      shape = "rectangle",
-#                      render = F,
-#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
-# 
-# strq1 <- str %>% 
-#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") & 
-#            to %in% c("balance_1", "injured_1")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq1$strength)
-# labels[4] <- str_pad(labels[4], side = "left", width = 2)
-# 
-# edgenames <- edgeNames(q1p)
-# names(labels) <- edgenames
-# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
-# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6, 
-#                                     "stiffness_1~injured_1" = 3,
-#                                     "clevel~balance_1" = 3),
-#                             lty = c("injured_1~balance_1" = "dashed",
-#                                     "hours~injured_1" = "dashed"))
-# 
-# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 12)))
-# 
-# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
-# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
-# renderGraph(q1p)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-```{r table5, results='asis'}
-query1 %>%
-  select(c(1, 2, 3)) %>%
-  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
-Negative life events had a negligible effect when moving from the "Low" to "High" state.
-
-The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
-The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
-When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
-Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
-In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.
-
-```{r table6, results='asis'}
-
-query2 %>%
-  select(7, 2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
-The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}. 
-Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.
-
-```{r table7, results="asis"}
-injured1comp %>%
-  filter(variable != "outcome") %>% 
-  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(  .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
-    col.names = c("Variable", "High risk", "Low risk"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-### Markov blanket of injured_2
-
-```{r}
-set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2")
-query3 <- map_dfr(var2, ~newprobtable(tallc,
-                                      .x,
-                                      "injured_2",
-                                      "injured",
-                                      avg30)) %>% 
-  gather(variable, state, -prob, - outcome) %>% 
-  drop_na() %>% 
-  select(variable, state, prob) %>% 
-  spread(state, prob) %>% 
-  select(variable, Low, High) %>% 
-  dendroTools::round_df(2)
-
-query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
-                       "injured_2",
-                       "injured",
-                       avg30
-) %>% 
-  toptail(3) %>% 
-  arrange(-prob) %>% 
-  select(outcome, everything()) %>% 
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
-The arc between stiffness_2 and injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
-Results of the first query for injured_2 in the "injured" state are presented in table \ref{tab:table8}.
-Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
-FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
-"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
-
-```{r}
-# plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & 
-#            to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[6] <- str_pad(labels[6], side = "left", width = 2)
-# edgenames <- c("gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
-
-```{r table9, results="asis" }
-
-query4 %>%
-  select(7,2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-## Second network structure - changes within variables
-
-```{r include=FALSE, warning=FALSE}
-set.seed(1003)
-bmtable <- tidy(bm1, effects = "fixed") %>% 
-  select(term:conf.high) %>% 
-  dendroTools::round_df(2) %>% 
-  mutate_at(vars(conf.low, conf.high), list(~format(round(., 2), nsmall = 2))) %>%
-  mutate(conf.low = str_c("[", conf.low, sep = ""),
-         conf.high = str_c(conf.high, "]", sep = "")) %>% 
-  unite(., "95% CI", c("conf.low", "conf.high"), sep = ", ") %>% 
-  mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
-         "95% CI" = str_replace_all(`95% CI`, ",", ", ")) %>% 
-  slice(1:4) %>% 
-  mutate(term = c("Intercept", "RMSSD", "BIS", "RMSSD:BIS"))
-
-```
-
-The network for changes within variables is presented in Fig \ref{fig:fig9}.
-An arc between BIS $\rightarrow$ FFFS with strength 1 was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strength of 0.60.
-The arc between RMSSD $\rightarrow$ FFFS was 0.56.
-The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a linear regression model.
-While the continuous nodes in the second network are already parameterised using linear regressions, it was not possible to investigate the interaction between explanatory variables.
-The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
-
-```{r}
-# mb(avg30, "injured_2")
-# plotsg2 <- model2network('[nlelg_2][gender][pi][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|pi:gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("pi", "FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   slice(2:8) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[7] <- str_pad(labels[7], side = "left", width = 2)
-# edgenames <- c("pi~stiffness_2", "gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "pi~stiffness_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in RMSSD where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table10, results="asis"}
-
-bmtable %>%
-  kable(
-    .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
-    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r}
-set.seed(1005)
-test <- expand.grid(list(c(-1:1),
-                         c(-1:1)))
-query6 <- map2_dfr(.x = as.numeric(test$Var1),
-                   .y = as.numeric(test$Var2), ~tibble(prob= cpquery(fittedcont, 
-                                                                     event = (injured == "injured"),
-                                                                     evidence = list(nlec = .x,  stiffness = .y), method = "lw"),
-                                                       nle = .x,
-                                                       stiffness = .y))
-
-query6a <- query6 %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for the "injured" node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
-Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of `r query6a[9,1]` over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
-
-```{r}
-# plotsg3 <- model2network('[gender][pi][hours][injured|pi:hours][stiffness|gender:injured][nlec|injured]')
-# q3p <- graphviz.plot(plotsg3,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi", "hours"), "injured", c("stiffness", "nlec")))
-# strq3 <- strchange %>% 
-#   filter(from %in% c("pi", "hours", "injured", "gender") & 
-#            to %in% c("stiffness", "nlec", "injured")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq3$strength)
-# edgenames <- c("pi~injured", "gender~stiffness", "hours~injured",
-#                "injured~stiffness", "injured~nlec")
-# names(labels) <- edgenames
-# nodeRenderInfo(q3p)$fill[c("injured")] = "gold"
-# edgeRenderInfo(q3p) <- list(lwd = c("injured~stiffness" = 4, 
-#                                     "hours~injured" = 5,
-#                                     "gender~stiffness" = 4),
-#                             lty = c("pi~injured" = "dashed"))
-# 
-# q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q3p)$fill[c("nlec", "stiffness")] = "tomato"
-# nodeRenderInfo(q3p)$fill["injured"] = "gold"
-# renderGraph(q3p)
-# rm(plotsg3, strq3, labels)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-```{r table11, results="asis" }
-set.seed(1006)
-test <- expand.grid(list(
-  c(-1:1),
-  c(-1:1)
-))
-query6 <- map2_dfr(
-  .x = as.numeric(test$Var1),
-  .y = as.numeric(test$Var2),
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(nlec = .x, stiffness = .y),
-      method = "lw"
-    ),
-    nle = .x,
-    stiffness = .y
-  )
-)
-
-query6 %>%
-  select(nle, stiffness, prob) %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  arrange(-prob) %>%
-  mutate(prob = round(prob, digits = 2)) %>%
-  select(c(3, 1, 2)) %>%
-  kable(
-    .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
-
-```{r table12, results="asis"}
-set.seed(1007)
-testt <- expand.grid(list(
-  c(-1:1),
-  c(-1:1),
-  c("Low", "High"),
-  c("injury", "no injury")
-))
-
-query7 <- pmap_dfr(
-  .l = testt,
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(
-        nlec = as.numeric(..1),
-        stiffness = as.numeric(..2),
-        hours = ..3,
-        pi = ..4
-      ),
-      method = "lw"
-    ),
-    nle = ..1,
-    stiffness = ..2,
-    hours = ..3,
-    pi = ..4
-  )
-) %>%
-  toptail(3) %>%
-  arrange(-prob)
-
-query7 %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
-  ) %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-# Discussion
-
-# Disclosure/Conflict-of-Interest Statement {-}
-
-<!--  
-Frontiers follows the recommendations by the International Committee of Medical
-Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
-that all financial, commercial or other relationships that might be perceived by
-the academic community as representing a potential conflict of interest must be
-disclosed. If no such relationship exists, authors will be asked to declare that
-the research was conducted in the absence of any commercial or financial
-relationships that could be construed as a potential conflict of interest. When
-disclosing the potential conflict of interest, the authors need to address the
-following points:
-
- - Did you or your institution at any time receive payment or services from a
-   third party for any aspect of the submitted work?
- - Please declare financial relationships with entities that could be perceived
-   to influence, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- - Please declare patents and copyrights, whether pending, issued, licensed
-   and/or receiving royalties relevant to the work.
- - Please state other relationships or activities that readers could perceive to
-   have influenced, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- -->
-
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Author Contributions {-}
-
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-# Acknowledgments {-}
-
-Funding:
-
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# References
-
-A reference list should be automatically created here. However it won't. Pandoc 
-will place the list of references at the end of the document instead. There are 
-no convenient solution for now to force Pandoc to do otherwise. The easiest way 
-to get around this problem is to edit the LaTeX file created by Pandoc before 
-compiling it again using the traditional LaTeX commands.
-
-# Figures {-}
-
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```

### 2020-05-23 r-files/functions.R

@@ -1,0 +1,240 @@
+splitr <- function(x){cut(x, breaks = c(min(x), median(x), max(x)), labels = c("Low", "High"), 
+                          include.lowest = TRUE, ordered_result = TRUE)}
+
+
+## function to prepare data for network
+networkprepare <- function(networkdata){
+  filter(networkdata, time == 1) %>% select(-time) -> t1
+  filter(networkdata, time == 2) %>% select(-time) -> t2
+  filter(networkdata, time == 3) %>% select(-time) -> t3
+  explans <- networkdata %>% 
+    select(c(id, pi:hours)) %>% 
+    names()
+  merge(t1, t2, by = explans, suff = c("_1", "_2")) %>% 
+    select(-id) -> t12
+  merge(t2, t3, by = explans, suff = c("_1", "_2")) %>% 
+    select(-id) -> t23
+  tallc = rbind(t12, t23)
+} # generates tallc files
+
+networkmaker <- function(names){
+  networkdata <- datanet %>% 
+    select(names) %>% 
+    mutate_at(vars(BIS:FFFS, nlebase), splitr)
+  tallc <- networkprepare(networkdata)
+  
+  g1 = tallc %>% select(ends_with("_1")) %>% names()
+  g2 = tallc %>% select(ends_with("_2")) %>% names()
+  
+  explans <- networkdata %>% 
+    select(pi:hours, nlebase) %>% 
+    names()
+  
+  bl <-  expand.grid(from = g2, to = g1, stringsAsFactors = FALSE)
+  bl <- rbind(bl, expand.grid(from = c(g1, g2), 
+                              to = explans, stringsAsFactors = FALSE)) 
+  bl <- rbind(bl, data.frame(from = c("clevel", "nlebase"),
+                             to = c("gender", "ind_team"), stringsAsFactors = FALSE)) 
+  bl <- rbind(bl, data.frame(from = c(g1[2:length(g1)]),
+                             to = c(g2[2:length(g2)]), stringsAsFactors = FALSE))
+  
+  
+  wl = data.frame(from = c(paste(vars_select(names(tallc), starts_with("n"))[2]), 
+                           paste(vars_select(names(tallc), starts_with("n"))[3])),
+                  to = c("injured_1", "injured_2"))
+  
+  cl = makeCluster(4, type = "SOCK")
+  clusterSetRNGStream(cl, 40)
+  
+  str = boot.strength(tallc, algorithm = "tabu",
+                      algorithm.args = list(blacklist = bl, 
[…]

### 2020-05-23 r-files/methods_code.R

@@ -1,0 +1,596 @@
+## Methods
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect)
+library(readxl)
+library(janitor)
+library(irr)
+library(lavaan)
+library(caret)
+library(parallel)
+library(zoo)
+library(sjmisc)
+## Table 1
+
+x <- read_csv("data/new_cleandata_inj.csv")
+
+id_remove <- x %>% 
+  group_by(id) %>% 
+  filter(sum(time) == 1) %>% 
+  pull(id)
+
+x %>% 
+  filter(!id %in% id_remove) %>% 
+  group_by(time, id) %>% 
+  slice(1) %>% 
+  ungroup() %>% 
+  group_by(time) %>% 
+  count()
+
+datanet %>% 
+  group_by(time) %>% 
+  count()
+
+parchar <- read_csv("data/participant_chars.csv") %>% 
+  filter(!id %in% id_remove)
+
+clevel <- parchar %>%
+  group_by(gender) %>%
+  summarise(
+    Recreational = n_perc0(clevel == "Recreational"),
+    University = n_perc0(clevel == "University"),
+    "National/International" = n_perc0(clevel == "National/International")
+  ) %>%
+  gather(name, value, -gender) %>%
+  spread(gender, value) %>%
+  ungroup() %>%
+  arrange(match(name, c("Recreational", "University", "National/International")))
+
+
+npars <- parchar %>%
+  group_by(gender) %>%
+  summarise(
+    Age = mean_sd(age, denote_sd = "paren", digits = 1),
+    Height = mean_sd(height, denote_sd = "paren", digits = 1),
+    Weight = mean_sd(weight, denote_sd = "paren", digits = 1),
+    "Hours per week" = mean_sd(hours, denote_sd = "paren", digits = 1),
+    "Student" = n_perc0(student == "y"),
+    "Non-student" = n_perc0(student == "n"),
+    "Injured" = n_perc0(pic == "1"),
+    "Healthy" = n_perc0(pic == "0")
+  ) %>%
+  gather(name, value, -gender) %>% 
+  arrange(gender)%>%
+  spread(gender, value) %>%
+  rbind(., clevel) %>%
+  rename(
+    "Male (n = 231) " = "male",
+    "Female (n = 120)" = "female"
+  ) %>%
+  ungroup()
+
+rm(clevel)
+
[…]

### 2020-05-23 r-files/results_code.R

@@ -1,0 +1,296 @@
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect,
+               Rgraphviz, sjmisc)
+
+dinj <- read_csv("data/new_cleandata_inj.csv")
+
+injtable <- dinj %>% 
+  select(id, time, variable, gender, injured:typeofinjuryg) %>% 
+  filter(time != 4 & injured == 1) 
+
+
+inj_summary <- injtable %>% 
+  mutate(bodylocation = ifelse(bodypart %in% c("upperarm_shoulder",
+                                               "head_neck",
+                                               "torso",
+                                               "elbow_lowerarm_hand"), "Upper body", "Lower body")) %>% 
+  group_by(gender, bodylocation) %>% 
+  summarise("Joint / ligament" = n_perc0(typeofinjuryg == "joint_ligament"),
+            "Muscle / tendon" = n_perc0(typeofinjuryg == "muscle_tendon"),
+            "Other" = n_perc0(typeofinjuryg %in% c("other", "brain", "bone", "skin"))) %>% 
+  gather(variable, value, -gender, -bodylocation)  %>% 
+  spread(bodylocation, value) 
+
+
+inj_mechanism <- injtable %>% 
+  group_by(gender) %>% 
+  summarise("Acute" = n_perc_x(acu_over == "acute", digits = 0),
+            "Chronic" = n_perc_x(acu_over == "chronic", digits = 0),
+            "Contact" = n_perc_x(contact_non == "contact", digits = 0),
+            "Non-contact" = n_perc_x(contact_non == "noncontact", digits = 0)) %>% 
+  gather(variable, value, -gender)%>% 
+  spread(gender, value) %>% 
+  ungroup()
+
+
+
+tallc <- networkprepare(networkdata)
+## extract names for blacklist + plotting
+
+g1 <- tallc %>%
+  select(ends_with("_1")) %>%
+  names()
+g2 <- tallc %>%
+  select(ends_with("_2")) %>%
+  names()
+
+## get explanatory variable names
+explans <- networkdata %>%
+  select(pi:hours, nlebase) %>%
+  names()
+
+## create blacklist
[…]

### 2020-05-30 demo.Rmd
pre methods update

@@
 title: "An interdisciplinary examination of stress and injury occurrence in athletes"
-runtitle: "Short Title"
+runtitle: "Stress and sports injury"
 documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
@@
 
-This paper adopts an interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
+This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
 At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
@@
 knitr::opts_chunk$set(echo = FALSE)
-#source("r-files/functions.R")
-#source("r-files/methods_code.R")
+source("r-files/functions.R")
+source("r-files/methods_code.R")
 #source("r-files/results_code.R")
@@
 One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network) of interest [@Scutari2014].
-Once constructed, the learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
+BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014].
+The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
 Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
@@
 - Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
-- Examine the relationships between the markers of stress and injury using a prospective, repeated measures design.
-- Evaluate the relationships between the markers of stress and injury using Bayesian network modelling that captures the complex nature of injury occurrence.
+- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
+- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.
+
+# Methods {-}
+
+```{r methodssetup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+```
+
+## Participants {-}
+
+A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
+Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants self-rated competitive level ranged from recreational to international standard. 
+A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
+At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
+Participants were engaged in training for their respective sports for at least five hours per week.
+Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
+
+```{r table1, results='asis'}
+
+kable(
+  npars %>% 
+    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
+    slice(c(1,3,8, 4, 9:11)) %>% 
+  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
+                          name == "Height" ~ "Height (cm)",
+                          name == "Weight" ~ "Body mass (kg)",
+                          TRUE ~ name)),
+  caption = "Participant characteristics.",
+  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
+  align = "c",
+  escape = F
+) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  pack_rows("Demographics M (SD)", 1, 4) %>%
+  pack_rows("Current competitive level n (%)", 5, 7) %>%
+  row_spec(0, bold = TRUE)
+```
+
+## Measures {-}
+
+### Major life events  {-}
+
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
+Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
+The LESCA comprises 69 items that reflect possible life events that participants may have experienced. 
+Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
+Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
+Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
+A score for total life events was calculated by summing the absolute values for both negative and positive events.
+Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. 
+Petrie also provided evidence of discriminant, convergent and predictive validity. 
+The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
+For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+The composite reliability for the LESCA in this study was 0.84.
+
+### Reinforcement Sensitivity Theory Personality Questionnaire  {-}
+
+A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
+The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
+Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
+The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total score that was subsequently used for further analysis.
+The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
+Further details regarding the revised RST are in S1 Appendix.
+
+### Heart rate variability {-}
+
+A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
+In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].
+
+### Musculoskeletal properties {-}
+
+A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure muscle stiffness.
+The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
+The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
+The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016]. 
+The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].
+
+### Postural stability {-}
+
+Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
+In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
+non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
+To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
+Participants were asked to maintain each stance for a total of 20-seconds.
+Participants hands were placed on hips at the level of the iliac crests. 
+A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
+
+\begin{figure}[!h]
+\caption{{\bf mBESS positions (A-F).}
+Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
+\label{fig:fig2}
+\end{figure}
+
+The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
+One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
+b) opening eyes; 
+c) stepping, stumbling, or falling; 
+d) moving the thigh into more than 30 degrees of flexion or abduction; 
+e) lifting the forefoot or heel; and 
+f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
+A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
+A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
+To calculate limb asymmetry, the DL and NDL leg score was calculated by summing the DL and NDL errors respectively.
+A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
+To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
+The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.
+
+### Injury {-}
+
+Participants self-reported any injuries they sustained at each data collection during the study period.
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
+Injury status (did / did not sustain an injury) served as the main outcome measure.
+
+## Procedure {-}
+
+At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
+With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
+To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week.
+Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
+A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
+Each participant was asked to attend a total of 3 data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
+Participants provided informed consent before data collection commenced.
+
+\begin{figure}[!h]
+\caption{{\bf Study design.}
+For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
+\label{fig:fig3}
+\end{figure}
+
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
+To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
+The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
+Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
+
+\begin{figure}[!h]
+\caption{{\bf Session protocol.}
+Outline of the protocol for each data collection.}
+\label{fig:fig4}
+\end{figure}
+
+### Questionnaires {-}
+
+The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed  on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
+The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
+For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
+The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis purposes.
+
+### HRV {-}
+
+To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
+Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
+Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines. 
+Participants were seated and asked to remain as still as possible for the duration of the recording. 
+No attempt was made to control participants respiratory frequency or tidal volume [@Denver2007]. 
+Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
+
+Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices.
+Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996].
+RMSSD was calculated as:
+
+\begin{equation} 
+  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
+\end{equation}
+
+Where N is the length of the time series, and $RR_i$ the RR interval between beats $i$ and $i-1$, where each beat position corresponds to the beat detection instant.
+
+### Muscle stiffness {-}
+
+To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb. 
+The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique  to determine the exact location of each site [@Chuang2012].
+The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle. 
+The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
+A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.
+
+After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site. 
+A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
+The device was programmed to deliver five consecutive impulses, separated by a one second interval [@Morgan2018].
+For each impulse, the device computed stiffness values, with the median of the five values being saved by the device for further analysis. 
+In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted. 
+Sets of measurements above 3 % were measured again to ensure the reliability of the data. 
+The CV was calculated in real time by the device after each set of measurements. 
+Measurements saved on the device were uploaded to a computer using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis. 
+For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis.
+
+### Postural stability {-}
+
+Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
+For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds. 
+Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
+The video recording was started prior to the first stance position and stopped after all stances had been completed.
+Each completed mBESS protocol took approximately 4 minutes.
+Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
+The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
+
+## Data Analysis {-}
+
+Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
+A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
+A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988].
+However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
+The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012].
+In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
+
+\begin{figure}[H]
+\caption{{\bf Example network.}
+A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
+\label{fig:fig5}
+\end{figure}
+
+In order to use discrete networks, continuous variables must first be split into categorical levels.
+When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
+
+Learning the structure of the network is an important step in BN modelling.
+The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
+The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
+The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014].
+In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
+In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
+BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
+
+[^1]: The term "query" in relation to Bayesian Networks stems from Pearl's expert systems theory (1988).
+A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
+Widely used texts on Bayesian Network analysis [@Koller2009] have widely adopted the terminology in favour of that used in traditional statistics.
+
+BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
+For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
+Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
+In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure.
+Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
+The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.
+
+### Data Preparation {-}
+
+Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
+To prepare the data for the BN, missing values in the dataset were first imputed. 
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
+The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
+A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
+In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
+A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
+The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].
+
+A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary table ...) and required no further transformation prior to the median split.
+For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
+Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
+In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
+
+### Network structure {-}
+
+To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
+Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist).
+Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
+To account for the repeated measures design and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
+In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
+However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
+Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
+Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
+Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
+Table \ref{tab:table2} provides an example of the formatted data. 
+Participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
+Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
+In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
+Finally, arc direction was restricted between specific nodes within the explanatory variables. 
+Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
+All subsequent models used the same blacklist.
+
+```{r table2, results='asis'}
+exampledata <- tibble(
+  Participant = c("1", "1", "2", "3", "3"),
+  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
+  " " = "->",
+  "X_2" = c("T2", "T3", "T2", "T2", "T3")
+)
+
+kable(
+  exampledata,
+  caption = "Example of the data arrangement used for the network.",
+  escape = TRUE
+) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0, bold = TRUE)
+```
+
+### Preliminary network structures {-}
+
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explred.
+Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
+A higher BIC value indicates the structure of a DAG is a better fit to the observed data [@Scutari2014]. 
+BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
+
+Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
+Additionally, despite strong evidence in the literature that both negative and total life event stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
+Total life event score was not included in the final structure.
+
+The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
+Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
+RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
+Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
+Therefore, only RI was included in the final structure.
+
+Finally, both total score and asymmetry were included in the initial network.
+However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
+Therefore, balance asymmetry was removed from the final network structure.
+To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
+
+\begin{table}[H]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{\label{tab:table3}Variables included in the final network structure.}
+\centering
+\begin{tabular}{l|l|l|l}
+\hline
+{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
+\hline
+clevel & Current competitive level & Club\_university\_county & National\_international\\
+\hline
+gender & Gender of the participant & Female & Male\\
+\hline
+hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
+\hline
+ind\_team & Participate in an individual or team based sport & Individual & Team\\
+\hline
+pi & Sustained previous injury & No Injury & Injury\\
+\hline
+nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
+\hline
+FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
+\hline
+BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
+\hline
+RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
+\hline
+stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
+\hline
+rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
+\hline
+balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
+\hline
+nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
+For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
+Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
+To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
+The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
+To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
+Independent variables were then standardized to allow relative changes between variables to be compared. 
+The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
+The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
+Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
+The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
+
+To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
+The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
+Arcs that were present in at least 30% of the models were included in the averaged model.
+The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
+An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
+In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
+Arcs below 0.3 were not included in the model.
+The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+
+### Network Inference {-}
+
+Conditional probability queries (CPQ) were used to perform inference on both network structures.
+To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
+The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
+CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
+The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
+Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
+Of particular interest in the current study were the variables that were connected to "injured" nodes.
+To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
+A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
+CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
+
+The second network contained both continuous and discrete data. 
+To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
+The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
+Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.

### 2020-05-30 demo.tex
pre methods update

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
 \def\Authors{
@@
 
-\title[Short Title]{An interdisciplinary examination of stress and injury occurrence in
+\title[Stress and sports injury]{An interdisciplinary examination of stress and injury occurrence in
 athletes}
@@
 appropriate analysis technique that is able to capture the uncertainty
-and complexity of the relationships between different factors. One
+and complexity of the relationships between different variables. One
 technique that provides a solution to this challenge is Bayesian network
@@
 models using the underlying structure that connects different variables
-(nodes in the network) of interest (Scutari and Denis, 2014). Once
-constructed, the learned structure can be used for inference by
-obtaining the posterior probabilities of a particular node for a given
-query (e.g., if the value of Node A is x and the value of Node B is y,
-what is the probability Node C being value z?). Furthermore, BN's do not
-distinguish between dependent and independent variables as they are a
-form of unsupervised learning, which is a strength over regression or
-structure equation models when the underlying relationship in the
-network may not be known (Olmedilla et al., 2018).
+(nodes in the network) (Scutari and Denis, 2014). The learned structure
+can be used for inference by obtaining the posterior probabilities of a
+particular node for a given query (e.g., if the value of Node A is x and
+the value of Node B is y, what is the probability Node C being value
+z?). Furthermore, BN's do not distinguish between dependent and
+independent variables as they are a form of unsupervised learning, which
+is a strength over regression or structure equation models when the
+underlying relationship in the network may not be known (Olmedilla et
+al., 2018).
 
@@
 stress-related physiological markers that may occur between initial
-measurement and injury occurrence. Given the exploratory the current
-study had the following objectives:
+measurement and injury occurrence. Given the exploratory nature of the
+current study the following objectives were defined:
 
@@
 \item
-  Examine the relationships between the markers of stress and injury
-  using a prospective, repeated measures design.
+  Capture the markers of stress and injury occurrence using a
+  prospective, repeated measures design.
 \item
-  Evaluate the relationships between the markers of stress and injury
-  using Bayesian network modelling that captures the complex nature of
-  injury occurrence.
+  Explore and evaluate the relationships between the markers of stress
+  and injury using Bayesian network modelling.
 \end{itemize}
 
+\hypertarget{methods}{%
+\section*{Methods}\label{methods}}
+\addcontentsline{toc}{section}{Methods}
+
+\hypertarget{participants}{%
+\subsection*{Participants}\label{participants}}
+\addcontentsline{toc}{subsection}{Participants}
+
+A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120)
+were initially recruited for the study. Participants had an average age
+of 22.0 \(\pm\) 7.0 years and represented a range of team (football,
+rugby, netball, cricket, lacrosse, basketball and field hockey) and
+individual sports (athletics, tennis, weightlifting, gymnastics, judo,
+swimming and golf from a British University and local sports clubs
+(Table \ref{tab:table1}). Participants self-rated competitive level
+ranged from recreational to international standard. A total of 126
+(49.03\%) participants had sustained an injury in the 12 months prior to
+the start of the study (male: \emph{n} = 114 {[}49\%{]}, female;
+\emph{n} = 48 {[}40\%{]}). At the start of the study, all participants
+were injury free (no modifications to their usual training routine due
+to a sport related medical problem for a minimum of four weeks).
+Participants were engaged in training for their respective sports for at
+least five hours per week. Ethical approval was obtained from the
+University ethics committee prior to the start of the study and all
+participants provided informed consent.
+
+\begin{table}[H]
+
+\caption{\label{tab:table1}Participant characteristics.}
+\centering
+\begin{tabular}[t]{c|c|c}
+\hline
+\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
+\hline
+\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
+\hline
+\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
+\hline
+\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
+\hline
+\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
+\hline
+\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
+\hline
+\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
+\hline
+\hspace{1em}Recreational & 3 (4) & 7 (4)\\
+\hline
+\hspace{1em}University & 45 (56) & 141 (80)\\
+\hline
+\hspace{1em}National/International & 33 (41) & 28 (16)\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{measures}{%
+\subsection*{Measures}\label{measures}}
+\addcontentsline{toc}{subsection}{Measures}
+
+\hypertarget{major-life-events}{%
+\subsubsection*{Major life events}\label{major-life-events}}
+\addcontentsline{toc}{subsubsection}{Major life events}
+
+A modified version of the Life Events Survey for Collegiate Athletes
+(LESCA) was used to measure participants' history of life event stress
+(Petrie, 1992). Modifications were made to the LESCA to ensure the
+suitability of the items for the study population (S1 Table). The LESCA
+comprises 69 items that reflect possible life events that participants
+may have experienced. Example items include, ``Major change in the
+frequency (increased or decreased) of social activities due to
+participation in sport'', ``Major change in the amount (more or less) of
+academic activity (homework, class time, etc)'' and ``Major change in
+level of athletic performance in actual competition (better or worse)''.
+Participants were asked to rate the perceived impact of each life event
+they had experienced within the last 12 months on an 8-point Likert
+scale anchored at -4 (\(extremely\;negative\)) and +4
+(\(extremely\;positive\)). Negative and positive life event scores were
+calculated by summing the negative and positive scores respectively. A
+score for total life events was calculated by summing the absolute
+values for both negative and positive events. Petrie (1992) (Petrie,
+1992) reported test-retest reliabilities at 1-week and 8-weeks with
+values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p
+\textless{} .001) respectively. Petrie also provided evidence of
+discriminant, convergent and predictive validity. The LESCA is the most
+widely used measure of major life events for athletes in the sports
+injury literature. For this study, Composite Reliability (Fornell and
+Larcker, 1981) was preferred to Cronbach's alpha as it does not assume
+parallelity (i.e., all factor loadings are constrained to be equal, and
+all error variances are constrained to be equal) and instead takes into
+consideration the varying factor loadings of the items in the
+questionnaire. The composite reliability for the LESCA in this study was
+0.84.
+
+\hypertarget{reinforcement-sensitivity-theory-personality-questionnaire}{%
+\subsubsection*{Reinforcement Sensitivity Theory Personality
+Questionnaire}\label{reinforcement-sensitivity-theory-personality-questionnaire}}
+\addcontentsline{toc}{subsubsection}{Reinforcement Sensitivity Theory
+Personality Questionnaire}
+
+A revised version of the Reinforcement Sensitivity Theory Personality
+Questionnaire (RST-PQ) was used to measure motivation, emotion,
+personality and their relevance to psychopathology (Corr and Cooper,
+2016). The revised version of the RST-PQ comprises 51 statements that
+measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I
+am the sort of person who easily freezes-up when scared''), Behavioural
+Inhibition System (BIS; e.g., ``When trying to make a decision, I find
+myself constantly chewing it over'') and four Behavioural Approach
+System (BAS) factors; Reward Interest (e.g., ``I regularly try new
+activities just to see if I enjoy them''), Goal Drive Persistence (e.g.,
+``I am very persistent in achieving my goals''), Reward Reactivity
+(e.g., ``I get a special thrill when I am praised for something I've
+done well'') and Impulsivity (e.g., ``I find myself doing things on the
+spur of the moment''). Participants rated each item on a scale from 1
+(\(not\; at\; all\)) to 4 (\(highly\)) to reflect how well each
+statement described their personality in general. The responses to items
+associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed
+to give a total score that was subsequently used for further analysis.
+The composite reliabilities for each subscale were; BIS = 0.92, FFFS =
+0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details
+regarding the revised RST are in S1 Appendix.
+
+\hypertarget{heart-rate-variability}{%
+\subsubsection*{Heart rate variability}\label{heart-rate-variability}}
+\addcontentsline{toc}{subsubsection}{Heart rate variability}
+
+A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap
+(Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00)
+with ECG recordings (Giles et al., 2016), which are considered the gold
+standard for assessing HRV. In addition, HRV indices calculated from IBI
+and ECG data have shown a strong correlation (r = .99) in athletes
+(Caminal et al., 2018) and under spontaneous breathing conditions (Plews
+et al., 2017).
+
+\hypertarget{musculoskeletal-properties}{%
+\subsubsection*{Musculoskeletal
+properties}\label{musculoskeletal-properties}}
+\addcontentsline{toc}{subsubsection}{Musculoskeletal properties}
+
+A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to
+measure muscle stiffness. The MyotonPRO is a non-invasive, handheld
+device that applies a mechanical impulse of 0.40 N for 0.15 ms
+perpendicular to the surface of the skin. The impulse causes natural
+damped oscillations in the tissue, which are recorded by a three-axis
+digital accelerometer sensor in the device. The raw oscillation signal
+is then processed, and the stiffness parameter is calculated
+(Agyapong-Badu et al., 2016). The MyotonPRO has previously been reported
+to be a reliable and valid tool for the measurement of in-vivo tissue
+stiffness properties (Chuang et al., 2013; Pruyn et al., 2016; Nair et
+al., 2014), and has demonstrated good internal consistency (coefficient
+of variation \textless{} 1.4\%) over sets of 10 repetitions (Aird et
+al., 2012).
+
+\hypertarget{postural-stability}{%
+\subsubsection*{Postural stability}\label{postural-stability}}
+\addcontentsline{toc}{subsubsection}{Postural stability}
+
+Postural stability was assessed with a modified version of the balance
+error scoring system (mBESS) based on the protocol recommended by Hunt
+et al. (2009). In total, each trial of the mBESS was performed without
+shoes (McCrory et al., 2013) and included six stances in the following
+order; dominant leg (DL; standing on the dominant foot with the
+non-dominant foot at approximately 30-degrees of hip flexion and
+45-degrees of knee flexion), non-dominant leg (NDL; standing on the
+non-dominant foot with the dominant foot at approximately 30-degrees of
+hip flexion and 45-degrees of knee flexion) and tandem leg stance (TS;
+standing heel-to-toe with the non-dominant foot behind the dominant) on
+firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively
+(Fig \ref{fig:fig2}). To determine leg dominance, participants were
+asked their preferred leg to kick a ball to a target, and the chosen
+limb was labelled as dominant (cf.~Cingel et al., 2017). Participants
+were asked to maintain each stance for a total of 20-seconds.
+Participants hands were placed on hips at the level of the iliac crests.
+A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United
+Kingdom) was used to record each participants performance during the
+mBESS.
+
+\begin{figure}[!h]
+\caption{{\bf mBESS positions (A-F).}
+Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
+\label{fig:fig2}
+\end{figure}
+
+The error identification criteria from the original BESS protocol was
+used by the lead researcher who scored all the BESS trials. One error
+was recorded if any of the following movements were observed during each
+trial: a) lifting hands off iliac crests; b) opening eyes; c) stepping,
+stumbling, or falling; d) moving the thigh into more than 30 degrees of
+flexion or abduction; e) lifting the forefoot or heel; and f) remaining
+out of the testing position for more than 5-seconds (Riemann et al.,
+1999). A maximum score of 10 errors was possible for each stance.
+Multiple errors occurring simultaneously were recorded as one error. A
+participant was given the maximum score of 10 if they remained out of
+the stance position for more than 5-seconds. To calculate limb
+asymmetry, the DL and NDL leg score was calculated by summing the DL and
+NDL errors respectively. A total score was calculated by summing the
+total number of errors recorded on all stances (DL, NLD and TS, on foam
+and firm surfaces). To assess the intra-rater reliability, a single
+measurement, absolute agreement, two-way mixed effects model for the
+intraclass correlation (ICC; Koo and Li, 2016) was used on a sample of
+40 participants from the first time point. The test-retest scoring of
+BESS resulted in a ``good'' to ``excellent'' ICC score (ICC = 0.93, 95\%
+confidence interval = 0.88 - 0.96), indicating the scoring was reliable.
+
+\hypertarget{injury}{%
+\subsubsection*{Injury}\label{injury}}
+\addcontentsline{toc}{subsubsection}{Injury}
+
+Participants self-reported any injuries they sustained at each data
+collection during the study period. An injury was defined as any sports
+related medical problem causing the athlete to miss or modify their
+usual training routine for at least 24 hours (Fuller et al., 2006, 2007;
+Timpka et al., 2014). Minor scrapes and bruises that may require certain
+modifications (e.g., strapping or taping) but did not limit continued
+participation were not considered injuries (cf.~Appaneal et al., 2009).
+Injury status (did / did not sustain an injury) served as the main
+outcome measure.
+
+\hypertarget{procedure}{%
+\subsection*{Procedure}\label{procedure}}
+\addcontentsline{toc}{subsection}{Procedure}
+
+At the start of the UK academic year (September 2016 and 2017), coaches
+of sports teams at a British University and local sports clubs were
+contacted and informed about the study. With the coaches' permission,
+the lead researcher attended training sessions to inform the athletes
+about the overall purpose of the study and the requirements of
+participation. To be eligible for the study athletes had to be injury
+free (no modifications to their usual training routine due to a sport
+related medical problem for a minimum of four weeks) and training a
+minimum of five hours per week. Athletes who met the criteria and
+volunteered to take part in the study were invited to attend scheduled
+testing sessions. A repeated measures prospective cohort design was used
+to assess athletes' major life events, stress-related physiological
+markers and injury status over two consecutive twelve-month periods
+between September 2016 and September 2018. Each participant was asked to
+attend a total of 3 data collections over a twelve-month period, with
+each data collection separated by a four-month interval (Fig
+\ref{fig:fig3}). Participants provided informed consent before data
+collection commenced.
+
+\begin{figure}[!h]
+\caption{{\bf Study design.}
+For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
+\label{fig:fig3}
+\end{figure}
+
+For the first three data collections (T1, T2 and T3), participants
+followed the same protocol in a specific order (Fig \ref{fig:fig4}). To
+ensure all measures could be collected within a viable time-frame,
+participants were separated into two groups. The first group completed
+all computer-based measures followed by all physical measurements,
+whereas the second group completed all physical measurements followed by
+computer-based measures. Participants were randomly assigned to one of
+the two groups and remained in those groups across all time points.
+
+\begin{figure}[!h]
+\caption{{\bf Session protocol.}
+Outline of the protocol for each data collection.}
+\label{fig:fig4}
+\end{figure}
+
+\hypertarget{questionnaires}{%
+\subsubsection*{Questionnaires}\label{questionnaires}}
+\addcontentsline{toc}{subsubsection}{Questionnaires}
+
+The questionnaires, which included demographic information, the LESCA,
+RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed
+on-line (SurveyMonkey Inc., USA, www.surveymonkey.com). The instructions
+for the LESCA were modified at T2 and T3 so that participants reported
+major life events that had occurred since the previous testing session.
+For injury reporting, participants were asked to record any injuries
+that they had sustained since the last data collection. The data were
+downloaded from surverymonkey.com and imported into R (R Core Team,
+2019) for analysis purposes.
+
+\hypertarget{hrv}{%
+\subsubsection*{HRV}\label{hrv}}
+\addcontentsline{toc}{subsubsection}{HRV}
+
+To minimise potential distractions, participants were directed to a
+designated quiet area in the laboratory where IBI data were recorded.
+Participants were instructed to turn off their mobile devices to avoid
+any interference with the Bluetooth sensor. Each chest strap was
+dampened with water and adjusted so it fitted tightly but comfortably,
+as outlined by Polar's guidelines. Participants were seated and asked to
+remain as still as possible for the duration of the recording. No
+attempt was made to control participants respiratory frequency or tidal
+volume (Denver et al., 2007). Inter-beat interval (IBI) data was
+collected for 10-minutes at a sampling frequency of 1000 Hz.
+
+Raw, unfiltered IBI recordings were exported from the Polar Flow web
+service as a space delimited .txt file and imported into R (R Core Team,
+2019) where the \emph{RHRV} package ({\textbf{???}}) was used to
+calculate HRV indices. Raw IBI data was filtered using an adaptive
+threshold filter, and the first 3-minutes and last 2-minutes of each
+recording were discarded, leaving a 5-minutes window that was used to
+calculate the root mean square of successive differences (RMSSD) in RR
+intervals following the recommendations for short term IBI recordings
+(Laborde et al., 2017; Malik et al., 1996). RMSSD was calculated as:
+
+\begin{equation} 
+  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
+\end{equation}
+
+Where N is the length of the time series, and \(RR_i\) the RR interval
+between beats \(i\) and \(i-1\), where each beat position corresponds to
+the beat detection instant.
+
+\hypertarget{muscle-stiffness}{%
+\subsubsection*{Muscle stiffness}\label{muscle-stiffness}}
+\addcontentsline{toc}{subsubsection}{Muscle stiffness}
+
+To assess muscle stiffness, participants lay horizontally on a massage
+bed and four testing sites were identified on each lower limb. The
+muscle belly of the rectus femoris (RF), biceps femoris (BF), medial
+gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified
+using a visual-palpatory technique to determine the exact location of
+each site (Chuang et al., 2012). The visual-palpatory technique required
+the participant to contract the target muscle to aid the lead researcher
+to visually identify the muscle. The participant was then asked to relax
+the muscle and the muscle was palpated to locate the muscle belly. A
+skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was
+used to mark the testing site in the centre of the muscle belly.
+
+After the eight testing sites had been identified, the testing end of
+the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin
+on the testing site. A constant pre-load of 0.18 N was applied for
+initial compression of subcutaneous tissues. The device was programmed
+to deliver five consecutive impulses, separated by a one second interval
+(Morgan et al., 2018). For each impulse, the device computed stiffness
+values, with the median of the five values being saved by the device for
+further analysis. In accordance with Myoton.com, a set of five
+measurements with a coefficient of variation (CV) of less than 3\% was
+accepted. Sets of measurements above 3 \% were measured again to ensure
+the reliability of the data. The CV was calculated in real time by the
+device after each set of measurements. Measurements saved on the device
+were uploaded to a computer using MyotonPRO software and imported in R
+(R Core Team, 2019) for further analysis. For each participant, the sum
+of all eight testing sites was calculated to provide a total lower
+extremity stiffness score and was used for further analysis.
+
+\hypertarget{postural-stability-1}{%
+\subsubsection*{Postural stability}\label{postural-stability-1}}
+\addcontentsline{toc}{subsubsection}{Postural stability}
+
+Instructions for the mBESS were read to each participant and a
+demonstration of the positions was provided by the research assistant.
+For each position, participants were instructed to close their eyes,
+rest their hands on their iliac crests and remain as still as possible
+for 20-seconds. Participants were instructed to return to the testing
+position as quickly as possible if they lost their balance. The video
+recording was started prior to the first stance position and stopped
+after all stances had been completed. Each completed mBESS protocol took
+approximately 4 minutes. Only one trial was performed to avoid
+familiarisation effects across the repeated measurement (cf.~Valovich et
+al., 2003). The video recordings for each participant were imported from
+the recording equipment (Sony DSC-RX10) and the lead researcher scored
+each trial using the error identification criteria.
+
+\hypertarget{data-analysis}{%
+\subsection*{Data Analysis}\label{data-analysis}}
+\addcontentsline{toc}{subsection}{Data Analysis}
+
+Two Bayesian Networks (BN) were used to explore the relationships
+between the psychological measures, physiological markers of stress and
+sports injury. A BN is a graphical representation of a joint probability
+distribution among a set of random variables, and provides a statistical
+model describing the dependencies and conditional independences from
+empirical data in a visually appealing way (Scutari and Denis, 2014). A
+BN consists of arcs and nodes that together are formally known as a
+directed acyclic graph (DAG), where a node is termed a parent of a child
+if there is an arc directed from the former to the latter (Fig
+\ref{fig:fig5}; Pearl, 1988). However, the direction of the arc does not
+necessarily imply causation, and the relationship between variables are
+often described as probabilistic instead of casual (Scutari and Denis,
+2014). The information within a node can be either continuous or
+discrete, and a complete network can contain both continuous and
+discrete nodes; however, discrete networks are the most commonly used
+form of BN (Chen and Pollino, 2012). In discrete networks, conditional
+probabilities for each child node are allocated for each combination of
+the possible states in their parent nodes and can be used to assess the
+strength of a dependency in the network.
+
+\begin{figure}[H]
+\caption{{\bf Example network.}
+A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
+\label{fig:fig5}
+\end{figure}
+
+In order to use discrete networks, continuous variables must first be
+split into categorical levels. When there are a large number of
+variables in the network, limiting the number of levels has the benefit
+of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100
+(10\^{}2) possible parameter combinations, however the same network with
+three levels has 1000 (10\^{}3) possible parameter combinations, the
+latter being significantly more computationally expensive. Using a
+larger number of splits in the data also comes at a cost of reducing the
+statistical power in detecting probabilistic associations, and reduces
+the precision of parameter estimates for the probabilistic associations
+that are detected because it reduces the sample-size-to-parameters ratio
+(Scutari and Denis, 2014).
+
+Learning the structure of the network is an important step in BN
+modelling. The structure of a network can be constructed using expert
+knowledge and/or data-driven algorithm techniques (e.g., search and
+score, such as hill climbing and gradient descent algorithms; Scutari \&
+Denis, 2014). The learned structure can then be used for inference by
+querying the network\footnote{The term ``query'' in relation to Bayesian
+  Networks stems from Pearl's expert systems theory (1988). A query can
+  be submitted to an expert (in this case, the network is the expert) to
+  get an opinion, the expert then updates the querier's beliefs
+  accordingly. Widely used texts on Bayesian Network analysis (Koller
+  and Friedman, 2009) have widely adopted the terminology in favour of
+  that used in traditional statistics.} and obtaining the posterior
+probabilities of a particular node for a given query. The posterior
+distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where
+the learned network \(B\) with structure \(G\) and parameters
+\(\Theta\), are investigated with new evidence \(E\) using the
+information in \(B\) (Scutari and Denis, 2014). In the example network
+presented in Fig \ref{fig:fig5}, new values assigned to each of the
+parent nodes (e.g., both set to ``Low'') could be used to investigate
+what effect the new information has on the state of the child node
+(conditional probability of a particular state of the child node). In a
+more complex network containing many nodes, the outcome of a particular
+node can be assessed conditional on the states of any subset of nodes in
+the network. BNs therefore provide a unique and versatile approach to
+modelling a set of variables to uncover dependency structures within the
+data.
+
+BNs have recently been used in the sport psychology literature
+(Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several
+benefits over traditional statistical analysis. For example, predictions
+can be made about any variable in the network, rather than there being a
+distinction between dependent and independent variables in the data,
+such as in linear regression models that are often used within the sport
+psychology literature (Olmedilla et al., 2018; Bittencourt et al.,
+2016). Furthermore, the structure of a network can be obtained from both
+empirical data \emph{and} prior knowledge about the area of study; the
+latter being particularly useful when there are a large number of
+variables in the network, or only a small number of observations are
+available in the data (Xiao-xuan et al., 2008). In such instances, a
+purely data driven approach to learning the network would be
+time-consuming due to the large parameter space, and inefficient at
+identifying an approximation of the true network structure. Prior
+knowledge about dependencies between variables can therefore be included
+in the network structure, while still allowing a data driven approach
+for unknown dependencies, to improve the overall computation of the
+network structure (Heckerman et al., 1995; Xu et al., 2015). The
+following sections detail the steps taken in the current study to
+firstly prepare the data for the network, and then obtain the structure
+of the network that was used for inference.
+
+\hypertarget{data-preparation}{%
+\subsubsection*{Data Preparation}\label{data-preparation}}
+\addcontentsline{toc}{subsubsection}{Data Preparation}
+
+Of the 351 participants that were initially recruited for the study, 94
+only completed the first time point, and therefore had to be removed
+from the study as no injury information was obtained for these
+participants following the first time point. To prepare the data for the
+BN, missing values in the dataset were first imputed. Out of the 650
+total measurements across all time points in the current study, there
+were 31 (4.64\%) missing muscle stiffness measurements and 70 (10.48\%)
+missing heart rate recordings. The missing data were due to technical
+faults in the data collection equipment and were considered to be
+missing completely at random. A missing rate of 15-20\% has been
+reported to be common in psychological studies, and several techniques
+are available to handle missing values (Enders, 2003; Lang et al.,
+2013). In the current study, the \emph{caret} package (Kuhn et al.,
+2008) was used to impute the missing values. A bagged tree model using
+all of the non-missing data was first generated and then used to predict
+each missing value in the dataset. The bagged tree method is a reliable
+and accurate method for imputing missing values in data and is superior
+to other commonly used methods such a median imputation (Kuhn et al.,
+2008).
+
+A median split technique was used to discretise the data used in the
+network into ``Low'' and ``High'' levels. All variables apart from
+negative and total life events were approximately normally distributed
+(based on visual inspection, see supplementary table \ldots) and
+required no further transformation prior to the median split. For the
+LESCA questionnaire data, a cumulative total of the current, and
+previous time points was calculated at each time point to account for
+the potential continuing effect of the life events experienced by
+athletes over time. Given the limited support for a relationship between
+positive life events and injury (Williams and Andersen, 2007), only
+negative and total life events were included in the network. Cumulative
+negative, and cumulative total life event scores at each time point were
+first log scaled so distributions were approximately normal, and then
+binarised using the median at each time point (nlelg and tlelg
+respectively). In addition to the log scaled cumulative values, an
+untransformed NLE score from the first time point was included as an
+additional variable based on previous literature that indicates this
+variable should have a strong relationship with injury outcome (Ivarsson
+et al., 2017).
+
+\hypertarget{network-structure}{%
+\subsubsection*{Network structure}\label{network-structure}}
+\addcontentsline{toc}{subsubsection}{Network structure}
+
+To obtain the network structure, several steps were taken to ensure that
+both a theoretically realistic network, and a network that was an
+appropriate fit to the collected data, was used for inference. Prior
+knowledge about the network structure was included by providing a list
+of arcs that are always \emph{restricted} from being in the network
+(blacklist), and a list of arcs that are always \emph{included} in the
+network (whitelist). Additionally, there are several scoring functions
+such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet
+equivalent uniform (BDeu) that can be used to compare network structures
+with certain nodes and arcs included or excluded (Scutari and Denis,
+2014). To account for the repeated measures design and to maximise the
+use of the data, pairs of complete cases (e.g., participants who
+completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network
+(2TBN) structure (Murphy, 2002). In the 2TBN, variables measured T2
+could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2)
+and variables measured at T3 could depend on variables measured at T2
+(e.g., T2 \(\rightarrow\) T3). However, arcs were blacklisted between T2
+\(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in
+which data was collected. Variables were separated into two groups;
+``explanatory'', for variables that did not change during the study
+(e.g., gender), or ``independent'', for variables that were measured at
+each time point and could vary during the study. Independent variable
+names were suffixed with \_1 for time point T, and \_2 for time point
+T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
+Formatting the data in this way meant participants who completed T1 and
+T2, but did not complete T3, could still be included in the analysis.
+Table \ref{tab:table2} provides an example of the formatted data.
+Participants 1 and 3 have complete data, and therefore have two rows of
+data each representing variables from T1 \(\rightarrow\) T2 and T2
+\(\rightarrow\) T3, respectively. Participant 2 did not complete the
+final data collection at T3 and therefore only has one row of data
+representing the variables collected at T1 and T2. In addition to the
+blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\)
+T2, the direction of arcs was restricted between independent variables
+and explanatory variables (e.g., independent \(\rightarrow\)
+explanatory); however, arcs were not restricted between explanatory
+\(\rightarrow\) independent variables. Finally, arc direction was
+restricted between specific nodes within the explanatory variables. Arcs
+from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and
+nlebase \(\rightarrow\) ind\_team were included in the blacklist, as
+arcs in these directions did not make logical sense. All subsequent
+models used the same blacklist.
+
+\begin{table}[H]
+
+\caption{\label{tab:table2}Example of the data arrangement used for the network.}
+\centering
+\begin{tabular}[t]{l|l|l|l}
+\hline
+\textbf{Participant} & \textbf{X\_1} & \textbf{ } & \textbf{X\_2}\\
+\hline
+1 & T1 & -> & T2\\
+\hline
+1 & T2 & -> & T3\\
+\hline
+2 & T1 & -> & T2\\
+\hline
+3 & T1 & -> & T2\\
+\hline
+3 & T2 & -> & T3\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{preliminary-network-structures}{%
+\subsubsection*{Preliminary network
+structures}\label{preliminary-network-structures}}
+\addcontentsline{toc}{subsubsection}{Preliminary network structures}
+
+Prior to the final network structure presented in the results section,
+several structures with different combinations of variables were
+explred. Networks were learned using a Tabu search algorithm (Russell
+and Norvig, 2009) and BIC was used to compare different models. A higher
+BIC value indicates the structure of a DAG is a better fit to the
+observed data (Scutari and Denis, 2014). BIC values for each combination
+of variables of interest are reported as the combination of variables
+with the highest BIC value, followed by the relative scores of the other
+variables in the model.
+
+Initially, both negative life events and total life events were included
+in the network structure, however, the network score was improved when
+only nlelg or tlelg was included (highest BIC value = nleleg, BIC values
+relative to nlelg; tlelg only = -85.13, tlelg and nlelg = -219.14).
+Additionally, despite strong evidence in the literature that both
+negative and total life event stress are related to injury occurrence
+(Williams and Andersen, 2007), network structures learned using the Tabu
+search algorithm failed to identify a relationship between NLE and
+injury or TLE and injury in the data. Given that nlelg provided the
+highest network score, and there is a stronger relationship between
+negative life events and injury in the literature, an arc was
+whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2
+in the final network structure. Total life event score was not included
+in the final structure.
+
+The subscales representing the BAS (RR, RI, GDP and I) showed limited
+connection to other variables in the network. Therefore, several models
+were run with each scale individually to find the scale that resulted in
+the highest BIC value (values are shown relative to the highest value).
+RI provided the highest BIC value, compared to RR (-13.13), GDP (-15.06)
+and I (-15.2). Including all the variables (RR, RI, GDP and I) resulted
+in a significantly lower score -894.37) indicating that including all
+the variables was not beneficial to the model structure and did not
+offset the cost of the additional parameters. Therefore, only RI was
+included in the final structure.
+
+Finally, both total score and asymmetry were included in the initial
+network. However, visual inspection of the network revealed no arcs
+between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
+Therefore, balance asymmetry was removed from the final network
+structure. To summarise, Table \ref{tab:table3} includes the variables
+that were included in the final network structure.
+
+\begin{table}[H]
+\begin{adjustwidth}{-2.25in}{0in}
+\caption{\label{tab:table3}Variables included in the final network structure.}
+\centering
+\begin{tabular}{l|l|l|l}
+\hline
+{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
+\hline
+clevel & Current competitive level & Club\_university\_county & National\_international\\
+\hline
+gender & Gender of the participant & Female & Male\\
+\hline
+hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
+\hline
+ind\_team & Participate in an individual or team based sport & Individual & Team\\
+\hline
+pi & Sustained previous injury & No Injury & Injury\\
+\hline
+nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
+\hline
+FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
+\hline
+BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
+\hline
+RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
+\hline
+stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
+\hline
+rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
+\hline
+balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
+\hline
+nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+\end{tabular}
+\end{adjustwidth}
+\end{table}
+
+Preliminary network structures also revealed strong dependencies between
+the same variables at subsequent time points. For example, the
+probability that stiffness\_1 and stiffness\_2 were both ``High'', or
+both ``Low'' was approximately 80\%. Including the arcs between the same
+variables from X\_1 \(\rightarrow\) X\_2 did not provide any
+theoretically meaningful information to the network structure as the
+majority of participants would be either be in a ``Low'' or ``High''
+state for each pair of variables in the network. To more appropriately
+assess changes \emph{within} variables over time, a second BN was
+investigated by modelling the differences between variables at different
+time points. The use of differential equations to model changes in
+variables over time is a common procedure in BN analysis when there are
+repeated measurements in the data (Scutari et al., 2017). To obtain the
+structure, variables suffixed with \_1 were subtracted from variables
+suffixed with \_2 to calculate the difference between variables measured
+at time points T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3.
+Independent variables were then standardized to allow relative changes
+between variables to be compared. The ``injured'' variable was also
+modified to represent whether a participant had sustained an injury at
+any point over the duration of the study or were healthy for the
+duration of the study. The result was a network that explicitly modelled
+the \emph{amount} of change within variables between time points, as
+opposed to the first network that would only have captured changes when
+the median threshold was crossed from ``Low'' to ``High''. Identical
+blacklists to the first network were used for arcs between independent
+and explanatory variables. The nlebase variable was also dropped from
+the list of explanatory variables to allow the \emph{changes} in
+negative life events to be the only life event variable in the network.
+
+To obtain the final networks, the appropriate blacklist and whitelists
+were provided and a Tabu search algorithm identified the remaining
+structure of the network. The final network structure was obtained by
+averaging 1000 bootstrapped models (Efron and Tibshirani, 1994) to
+reduce the impact of locally optimal, but globally suboptimal network
+learning, and to obtain a more robust model (Olmedilla et al., 2018).
+Arcs that were present in at least 30\% of the models were included in
+the averaged model. The strength of each arc was determined by the
+percentage of models that the arc was included in, independent of the
+arc's direction. An arc strength of 1 indicates that the arc is always
+present in the network, with the value decreasing as arcs are found in
+fewer networks. In the respective study arcs above 0.5 were considered
+``significant'' with arcs below 0.5 and above 0.3 ``non-significant''
+(Scutari and Nagarajan, 2013). Arcs below 0.3 were not included in the
+model. The full table of arc strengths for the first and second network
+are available in S2 Table and S3 Table respectively.
+
+\hypertarget{network-inference}{%
+\subsubsection*{Network Inference}\label{network-inference}}
+\addcontentsline{toc}{subsubsection}{Network Inference}
+
+Conditional probability queries (CPQ) were used to perform inference on
+both network structures. To conduct a CPQ, the joint probability
+distribution of the nodes was modified to include a new piece of
+evidence. The query allows the odds of a particular node state (e.g.,
+injured\_1 = ``injured'') to be calculated based on the new evidence.
+CPQs were performed using a likelihood weighting approach; a form of
+importance sampling where random observations are generated from the
+probability distribution in such a way that all observations match the
+evidence given in the query. The algorithm then re-weights each
+observation based on the evidence when computing the conditional
+probability for the query (Scutari and Denis, 2014). Inference was first
+performed on arcs that had a strength greater than 0.50 between the
+explanatory variables and independent variables and between different
+independent variables in the network. Of particular interest in the
+current study were the variables that were connected to ``injured''
+nodes. To examine the variables that were associated with injured nodes
+in the network, the Markov blanket of ``injured\_1'' and ``injured\_2''
+were examined. A Markov blanket contains all the nodes that make the
+node of interest conditionally independent from the rest of the network
+(Fuster-Parra et al., 2017). CPQ's were used to determine what effect
+the variables in the Markov blanket of injured nodes had on the
+probability of the injured node being in the ``injured'' state.
+
+The second network contained both continuous and discrete data. To
+examine dependencies between continuous variables with arc strengths
+above 0.5 in the second network, random samples were generated based on
+the conditional distribution of the nodes included as evidence in the
+query. The samples were then extracted and examined with Bayesian linear
+regression models using the \emph{brms} package (Bürkner, 2017) to
+determine the relationship between nodes in the network. Similar to the
+first network, the Markov blanket of the ``injured'' node was also
+investigated by determining the highest probability of injury with
+combinations of variables in the Markov blanket below the mean change,
+at the mean change and above the mean change.
+
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
+Agyapong-Badu, S., Warner, M., Samuel, D., and Stokes, M. (2016).
+Measurement of ageing effects on muscle tone and mechanical properties
+of rectus femoris and biceps brachii in healthy males and females using
+a novel hand-held myometric device. \emph{Archives of Gerontology and
+Geriatrics} 62, 59--67.
+doi:\href{https://doi.org/10.1016/j.archger.2015.09.011}{10.1016/j.archger.2015.09.011}.
+
+\leavevmode\hypertarget{ref-Aird2012}{}%
+Aird, L., Samuel, D., and Stokes, M. (2012). Quadriceps muscle tone,
+elasticity and stiffness in older males: Reliability and symmetry using
+the MyotonPRO. \emph{Archives of Gerontology and Geriatrics} 55,
+e31--e39.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.archger.2012.03.005}{https://doi.org/10.1016/j.archger.2012.03.005}.
+
 \leavevmode\hypertarget{ref-Andersen1988}{}%
@@
 doi:\href{https://doi.org/10.1080/026404199365597}{10.1080/026404199365597}.
+
+\leavevmode\hypertarget{ref-Appaneal2009}{}%
+Appaneal, R. N., Levine, B. R., Perna, F. M., and Roh, J. L. (2009).
+Measuring postinjury depression among male and female competitive
+athletes. \emph{Journal of Sport and Exercise Psychology} 31, 60--76.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/19325188}.
 
@@
 
+\leavevmode\hypertarget{ref-Burkner2017a}{}%
+Bürkner, P. C. (2017). brms: An R package for Bayesian multilevel models
+using Stan. \emph{Journal of Statistical Software} 80.
+doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}.
+
+\leavevmode\hypertarget{ref-Caminal2018}{}%
+Caminal, P., Sola, F., Gomis, P., Guasch, E., Perera, A., Soriano, N.,
+and Mont, L. (2018). Validity of the Polar V800 monitor for measuring
+heart rate variability in mountain running route conditions.
+\emph{European Journal of Applied Physiology} 118, 669--677.
+doi:\href{https://doi.org/10.1007/s00421-018-3808-0}{10.1007/s00421-018-3808-0}.
+
+\leavevmode\hypertarget{ref-Chen2012}{}%
+Chen, S. H., and Pollino, C. A. (2012). Good practice in Bayesian
+network modelling. \emph{Environmental Modelling and Software} 37,
+134--145.
+doi:\href{https://doi.org/10.1016/j.envsoft.2012.03.012}{10.1016/j.envsoft.2012.03.012}.
+
+\leavevmode\hypertarget{ref-Chuang2013}{}%
+Chuang, L. L., Lin, K. C., Wu, C. Y., Chang, C. W., Chen, H. C., Yin, H.
+P., and Wang, L. (2013). Relative and absolute reliabilities of the
+myotonometric measurements of hemiparetic arms in patients with stroke.
+\emph{Archives of Physical Medicine and Rehabilitation} 94, 459--466.
+doi:\href{https://doi.org/10.1016/j.apmr.2012.08.212}{10.1016/j.apmr.2012.08.212}.
+
+\leavevmode\hypertarget{ref-Chuang2012}{}%
+Chuang, L. L., Wu, C. Y., and Lin, K. C. (2012). Reliability, validity,
+and responsiveness of myotonometric measurement of muscle tone,
+elasticity, and stiffness in patients with stroke. \emph{Archives of
+Physical Medicine and Rehabilitation} 93, 532--540.
+doi:\href{https://doi.org/10.1016/j.apmr.2011.09.014}{10.1016/j.apmr.2011.09.014}.
+
+\leavevmode\hypertarget{ref-VanCingel2017}{}%
+Cingel, R. E. H. van, Hoogeboom, T. J., Melick, N. van, Meddeler, B. M.,
+and Nijhuis-van der Sanden, M. W. G. (2017). How to determine leg
+dominance: The agreement between self-reported and observed performance
+in healthy adults. \emph{Plos One} 12, 1--9.
+doi:\href{https://doi.org/10.1371/journal.pone.0189876}{10.1371/journal.pone.0189876}.
+
+\leavevmode\hypertarget{ref-Corr2016c}{}%
+Corr, P. J., and Cooper, A. J. (2016). The reinforcement sensitivity
+theory of personality questionnaire (RST-PQ): Development and
+validation. \emph{Psychological Assessment} 28, 1427--1440.
+doi:\href{https://doi.org/10.1037/pas0000273}{10.1037/pas0000273}.
+
+\leavevmode\hypertarget{ref-Denver2007}{}%
+Denver, J. W., Reed, S. F., and Porges, S. W. (2007). Methodological
+issues in the quantification of respiratory sinus arrhythmia.
+\emph{Biological Psychology} 74, 286--294.
+doi:\href{https://doi.org/10.1016/j.biopsycho.2005.09.005}{10.1016/j.biopsycho.2005.09.005}.
+
 \leavevmode\hypertarget{ref-Djaoui2017}{}%
@@
 doi:\href{https://doi.org/10.1016/j.physbeh.2017.09.004}{10.1016/j.physbeh.2017.09.004}.
+
+\leavevmode\hypertarget{ref-Efron1993}{}%
+Efron, B., and Tibshirani, R. J. (1994). \emph{An introduction to the
+bootstrap}. New York: Chapman \& Hall Available at:
+\url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}.
+
+\leavevmode\hypertarget{ref-Enders2003}{}%
+Enders, C. K. (2003). Using the expectation maximization algorithm to
+estimate coefficient alpha for scales with item-level missing data.
+\emph{Psychological Methods} 8, 322--337.
+doi:\href{https://doi.org/10.1037/1082-989X.8.3.322}{10.1037/1082-989X.8.3.322}.
+
+\leavevmode\hypertarget{ref-Fornell1981}{}%
+Fornell, C., and Larcker, D. F. (1981). Evaluating structural equation
+models with unobservable variables and measurement error. \emph{Journal
+of Marketing Research} 18, 39--50.
+doi:\href{https://doi.org/10.2307/3151312}{10.2307/3151312}.
+
+\leavevmode\hypertarget{ref-Fuller2006}{}%
+Fuller, C. W., Ekstrand, J., Junge, A., Andersen, T. E., Bahr, R.,
+Dvorak, J., Hägglund, M., McCrory, P., and Meeuwisse, W. H. (2006).
+Consensus statement on injury definitions and data collection procedures
+in studies of football (soccer) injuries. \emph{British journal of
+sports medicine} 40, 193--201.
+doi:\href{https://doi.org/10.1136/bjsm.2005.025270}{10.1136/bjsm.2005.025270}.
+
+\leavevmode\hypertarget{ref-Fuller2007b}{}%
+Fuller, C. W., Molloy, M. G., Bagate, C., Bahr, R., Brooks, J. H. M.,
+Donson, H., Kemp, S. P. T., McCrory, P., McIntosh, A. S., Meeuwisse, W.
+H., et al. (2007). Consensus statement on injury definitions and data
+collection procedures for studies of injuries in rugby union.
+\emph{British journal of sports medicine} 41, 328--31.
+doi:\href{https://doi.org/10.1136/bjsm.2006.033282}{10.1136/bjsm.2006.033282}.
+
+\leavevmode\hypertarget{ref-Fuster-Parra2017}{}%
+Fuster-Parra, P., Vidal-Conti, J., Borràs, P. A., and Palou, P. (2017).
+Bayesian networks to identify statistical dependencies. A case study of
+Spanish university students' habits. \emph{Informatics for Health and
+Social Care} 42, 166--179.
+doi:\href{https://doi.org/10.1080/17538157.2016.1178117}{10.1080/17538157.2016.1178117}.
 
@@
 
+\leavevmode\hypertarget{ref-Giles2016}{}%
+Giles, D., Draper, N., and Neil, W. (2016). Validity of the Polar V800
+heart rate monitor to measure RR intervals at rest. \emph{European
+Journal of Applied Physiology} 116, 563--571.
+doi:\href{https://doi.org/10.1007/s00421-015-3303-9}{10.1007/s00421-015-3303-9}.
+
+\leavevmode\hypertarget{ref-Heckerman1995}{}%
+Heckerman, D., Geiger, D., and Chickering, D. M. (1995). Learning
+Bayesian networks: The combination of knowledge and statistical data.
+\emph{Machine Learning} 20, 197--243.
+doi:\href{https://doi.org/10.1023/A:1022623210503}{10.1023/A:1022623210503}.
+
 \leavevmode\hypertarget{ref-Hughes2014}{}%
@@
 doi:\href{https://doi.org/10.1080/15438627.2014.881821}{10.1080/15438627.2014.881821}.
+
+\leavevmode\hypertarget{ref-Hunt2009}{}%
+Hunt, T. N., Ferrara, M. S., Bornstein, R. A., and Baumgartner, T. A.
+(2009). The reliability of the modified balance error scoring system.
+\emph{Clinical Journal of Sport Medicine} 19, 471--475.
+doi:\href{https://doi.org/10.1097/JSM.0b013e3181c12c7b}{10.1097/JSM.0b013e3181c12c7b}.
 
@@
 
+\leavevmode\hypertarget{ref-Koller2009}{}%
+Koller, D., and Friedman, N. (2009). \emph{Probabilistic graphical
+models: Principles and techniques - Adaptive computation and machine
+learning}. Cambridge, MA: The MIT Press
+doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.
+
+\leavevmode\hypertarget{ref-Koo2016}{}%
+Koo, T. K., and Li, M. Y. (2016). A guideline of selecting and reporting
+intraclass correlation coefficients for reliability research.
+\emph{Journal of Chiropractic Medicine} 15, 155--63.
+doi:\href{https://doi.org/10.1016/j.jcm.2016.02.012}{10.1016/j.jcm.2016.02.012}.
+
+\leavevmode\hypertarget{ref-Kuhn2008}{}%
+Kuhn, M. C. from J. W., Weston, S., Williams, A., Keefer, C., and
+Engelhardt, A. (2008). caret: Classification and Regression Training.
+\emph{Journal of Statistical Software} 28.
+doi:\href{https://doi.org/10.1053/j.sodo.2009.03.002}{10.1053/j.sodo.2009.03.002}.
+
 \leavevmode\hypertarget{ref-Kumar2001}{}%
@@
 doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}.
+
+\leavevmode\hypertarget{ref-Laborde2017}{}%
+Laborde, S., Mosley, E., and Thayer, J. F. (2017). Heart rate
+variability and cardiac vagal tone in psychophysiological research -
+Recommendations for experiment planning, data analysis, and data
+reporting. \emph{Frontiers in Psychology} 8, 1--18.
+doi:\href{https://doi.org/10.3389/fpsyg.2017.00213}{10.3389/fpsyg.2017.00213}.
+
+\leavevmode\hypertarget{ref-Lang2014}{}%
+Lang, K. M., Jorgensen, T. D., Moore, E. W. G., and Little, T. D.
+(2013). On the joys of missing data. \emph{Journal of Pediatric
+Psychology} 39, 151--162.
+doi:\href{https://doi.org/10.1093/jpepsy/jst048}{10.1093/jpepsy/jst048}.
 
@@
 
+\leavevmode\hypertarget{ref-Malik1996}{}%
+Malik, M., Camm, A. J., Bigger, J. T., Breithardt, G., Cerutti, S.,
+Cohen, R. J., Coumel, P., Fallen, E. L., Kennedy, H. L., Kleiger, R. E.,
+et al. (1996). Heart rate variability. Standards of measurement,
+physiological interpretation, and clinical use. \emph{European Heart
+Journal} 17, 354--381.
+doi:\href{https://doi.org/10.1093/oxfordjournals.eurheartj.a014868}{10.1093/oxfordjournals.eurheartj.a014868}.
+
+\leavevmode\hypertarget{ref-McCrory2013}{}%
+McCrory, P., Meeuwisse, W. H., Aubry, M., Cantu, R. C., Dvořák, J.,
+Echemendia, R. J., Engebretsen, L., Johnston, K., Kutcher, J. S.,
+Raftery, M., et al. (2013). Consensus statement on concussion in sport:
+The 4th international conference on concussion in sport, Zurich,
+November 2012. \emph{Journal of Athletic Training} 48, 554--575.
+doi:\href{https://doi.org/10.4085/1062-6050-48.4.05}{10.4085/1062-6050-48.4.05}.
+
 \leavevmode\hypertarget{ref-Meeuwisse2007}{}%
@@
 
+\leavevmode\hypertarget{ref-Morgan2018}{}%
+Morgan, G. E., Martin, R., Williams, L., Pearce, O., and Morris, K.
+(2018). Objective assessment of stiffness in Achilles tendinopathy: a
+novel approach using the MyotonPRO. \emph{BMJ Open Sport \& Exercise
+Medicine} 4, e000446.
+doi:\href{https://doi.org/10.1136/bmjsem-2018-000446}{10.1136/bmjsem-2018-000446}.
+
 \leavevmode\hypertarget{ref-Murphy2003}{}%
@@
 doi:\href{https://doi.org/10.1136/bjsm.37.1.13}{10.1136/bjsm.37.1.13}.
+
+\leavevmode\hypertarget{ref-Murphy2002}{}%
+Murphy, K. (2002). Dynamic Bayesian networks: Representation, inference
+and learning.
+
+\leavevmode\hypertarget{ref-Nair2014}{}%
+Nair, K., Dougherty, J., Schaefer, E., Kelly, J., and Masi, A. (2014).
+Repeatability, reproducibility, and calibration of the MyotonPRO on
+tissue mimicking phantoms. in \emph{ASME summer bioengineering
+conference}, 1--2.
+doi:\href{https://doi.org/10.1115/SBC2013-14622}{10.1115/SBC2013-14622}.
 
@@
 doi:\href{https://doi.org/10.1080/0097840X.1983.9935025}{10.1080/0097840X.1983.9935025}.
+
+\leavevmode\hypertarget{ref-Pearl1988}{}%
+Pearl, J. (1988). \emph{Probabilistic reasoning in intelligent systems:
+Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann
+Publishers.
 
@@
 
+\leavevmode\hypertarget{ref-Petrie1992}{}%
+Petrie, T. A. (1992). Psychosocial antecedents of athletic injury: The
+effects of life stress and social support on female collegiate gymnasts.
+\emph{Journal of Behavioral Medicine} 18, 127--138.
+doi:\href{https://doi.org/10.1080/08964289.1992.9936963}{10.1080/08964289.1992.9936963}.
+
+\leavevmode\hypertarget{ref-Plews2017}{}%
+Plews, D. J., Scott, B., Altini, M., Wood, M., Kilding, A. E., and
+Laursen, P. B. (2017). Comparison of heart-rate-variability recording
+with smartphone photoplethysmography, polar H7 chest strap, and
+electrocardiography. \emph{International Journal of Sports Physiology
+and Performance} 12, 1324--1328.
+doi:\href{https://doi.org/10.1123/ijspp.2016-0668}{10.1123/ijspp.2016-0668}.
+
 \leavevmode\hypertarget{ref-Pruyn2015}{}%
@@
 doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}.
+
+\leavevmode\hypertarget{ref-Pruyn2016}{}%
+Pruyn, E. C., Watsford, M. L., and Murphy, A. J. (2016). Validity and
+reliability of three methods of stiffness assessment. \emph{Journal of
+Sport and Health Science} 5, 476--483.
+doi:\href{https://doi.org/10.1016/j.jshs.2015.12.001}{10.1016/j.jshs.2015.12.001}.
+
+\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
+R Core Team (2019). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing
+Available at: \url{https://www.r-project.org/}.
+
+\leavevmode\hypertarget{ref-Riemann1999d}{}%
+Riemann, B. L., Guskiewicz, K. M., and Shields, E. W. (1999).
+Relationship between clinical and forceplate measures of postural
+stability. \emph{Journal of Sport Rehabilitation} 8, 71--82.
+doi:\href{https://doi.org/10.1123/jsr.8.2.71}{10.1123/jsr.8.2.71}.
 
@@
 
+\leavevmode\hypertarget{ref-Norvig2009}{}%
+Russell, S. J., and Norvig, P. (2009). \emph{Artificial Intelligence: A
+Modern Approach}. 3rd ed. Prentice Hall Available at:
+\url{http://aima.cs.berkeley.edu/}.
+
+\leavevmode\hypertarget{ref-Scutari2017}{}%
+Scutari, M., Auconi, P., Caldarelli, G., and Franchi, L. (2017).
+Bayesian networks analysis of malocclusion data. \emph{Scientific
+Reports} 7, 1--11.
+doi:\href{https://doi.org/10.1038/s41598-017-15293-w}{10.1038/s41598-017-15293-w}.
+
 \leavevmode\hypertarget{ref-Scutari2014}{}%
@@
 
+\leavevmode\hypertarget{ref-Scutari2013}{}%
+Scutari, M., and Nagarajan, R. (2013). Identifying significant edges in
+graphical models of molecular networks. \emph{Artificial Intelligence in
+Medicine} 57, 207--217.
+doi:\href{https://doi.org/10.1016/j.artmed.2012.12.006}{10.1016/j.artmed.2012.12.006}.
+
 \leavevmode\hypertarget{ref-Sheu2016}{}%
 Sheu, Y., Chen, L. H., and Hedegaard, H. (2016). Sports- and
-Recreation-related Injury Episodes in the United States, 2011-2014.
+recreation-related injury episodes in the United States, 2011-2014.
 \emph{National Health Statistics Reports}, 1--12. Available at:
@@
 doi:\href{https://doi.org/10.1177/0363546507299532}{10.1177/0363546507299532}.
+
+\leavevmode\hypertarget{ref-Timpka2014}{}%
+Timpka, T., Alonso, J. M., Jacobsson, J., Junge, A., Branco, P.,
+Clarsen, B., Kowalski, J., Mountjoy, M., Nilsson, S., Pluim, B., et al.
+(2014). Injury and illness definitions and data collection procedures
+for use in epidemiological studies in Athletics (track and field):
+Consensus statement. \emph{British Journal of Sports Medicine} 48,
+483--490.
+doi:\href{https://doi.org/10.1136/bjsports-2013-093241}{10.1136/bjsports-2013-093241}.
+
+\leavevmode\hypertarget{ref-Valovich2003}{}%
+Valovich, T. C., Perrin, D. H., and Gansneder, B. M. (2003). Repeat
+administration elicits a practice effect with the Balance Error Scoring
+System but not with the Standardized Assessment of Concussion in high
+school athletes. \emph{Journal of Athletic Training} 38, 51--56.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/12937472}.
 
@@
 
+\leavevmode\hypertarget{ref-Xiao-xuan2007}{}%
+Xiao-xuan, H., Hui, W., and Shuo, W. (2008). Using expert's knowledge to
+build Bayesian networks. in \emph{International conference on
+computational intelligence and security workshops} (IEEE), 220--223.
+doi:\href{https://doi.org/10.1109/cisw.2007.4425484}{10.1109/cisw.2007.4425484}.
+
+\leavevmode\hypertarget{ref-Xu2015}{}%
+Xu, J. G., Zhao, Y., Chen, J., and Han, C. (2015). A structure learning
+algorithm for Bayesian network using prior knowledge. \emph{Journal of
+Computer Science and Technology} 30, 713--724.
+doi:\href{https://doi.org/10.1007/s11390-015-1556-8}{10.1007/s11390-015-1556-8}.
+
 \end{document}

### 2020-05-30 full_paper_v2.Rmd
pre methods update

@@
 One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014].
-The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
+BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables [@Scutari2014].
+The learned structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C being value z?).
 Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
@@
 Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
 Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
@@
 Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-
 A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
-Each participant was asked to attend a total of four data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
+Each participant was asked to attend a total of 3 data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
 Participants provided informed consent before data collection commenced.
@@
 
+In order to use discrete networks, continuous variables must first be split into categorical levels.
+When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
+Typically, no more than three levels have been used in Bayesian networks in the sports injury literature [@Olmedilla2018; @Ruiz-Perez2019]
+
 Learning the structure of the network is an important step in BN modelling.
@@
 
+Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
 To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
@@
 
-Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
-Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
-All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.
-
-For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
+A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary table ...) and required no further transformation prior to the median split.
+For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
 Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
 In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
@@
 
-Prior to the final network structure presented in the results section, several structures were investigated.
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explred.
 Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
@@
 
-\begin{table}[H]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{\label{tab:table3}Variables included in the final network structure.}
-\centering
-\begin{tabular}{l|l|l|l}
-\hline
-{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
-\hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Sustained previous injury & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
-\hline
-FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
-\hline
-BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
-\hline
-RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
-\hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
-\hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
-\hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
+```{r finalvars, results='asis'}
+library(knitr)
+library(kableExtra)
+varcutoffs %>% 
+  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
+) %>% 
+  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
+  kable("latex",
+       booktabs = T,
+       align = c("llcc"),
+       caption = "Variables included in the final Bayesian network structure.",
+       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
+  kable_styling(latex_options = "HOLD_position") %>% 
+  column_spec(2, width = c("5.2cm")) 
+```
 
@@
 The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
-To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
+To obtain the structure, continuous variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
 Independent variables were then standardized to allow relative changes between variables to be compared. 

### 2020-05-30 wordmaker.Rmd
pre methods update

--- 2020-05-23 full_paper_v2.Rmd
@@
 ---
-title: "An interdisciplinary examination of stress and injury occurrence in athletes"
-runtitle: "Short Title"
-documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
-author:
-  - name: Dr Harry Fisher
-    affiliation: '1'
-    etal: Fisher # First author's last name. 
-    email: harryfisher21@gmail.com # Indicates corresponding Author
-    institution: Cardiff Metropolitan University
-    street: Cyncoed Road
-    zip: CF23 6XD
-    country: United Kingdom
-  - name: Dr Marianne Gittoes
-    affiliation: '1'
-  - name: Professor Lynne Evans
-    affiliation: '1'
-  - name: Miss Leah Bitchell
-    affiliation: '1'  
-  - name: Dr Richard Mullen  
-    affiliation: '2'
-  - name: Dr Marco Scutari
-    affiliation: '3'    
-affiliation:
-  - id: '1'
-    department: School of Sport and Health Science
-    institution: Cardiff Metropolitan University
-    city: Cardiff
-    country: United Kingdom
-  - id: '2' 
-    department: Division of Sport, Health & Exercise Sciences
-    institution: Brunel University
-    city: London
-    country: United Kingdom
-  - id: '3' 
-    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
-    city: Manno
-    country: Switzerland
+title: "wordmaker"
+output: word_document
+always_allow_html: true
 
-output: rticles::frontiers_article
-csl: frontiers.csl
-bibliography: library.bib
 ---
-
-\begin{abstract}
-
-This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
-Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
-Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury.
-Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury.
-In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury.
-Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.
-
-\tiny
- \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}
-
-\end{abstract}
 
@@
 source("r-files/methods_code.R")
-source("r-files/results_code.R")
+#source("r-files/results_code.R")
 ```
@@
 
-## Participants
+## Participants {-}
 
@@
 
-## Measures
+## Measures {-}
 
-### Major life events
+### Major life events  {-}
 
@@
 
-### Reinforcement Sensitivity Theory Personality Questionnaire
+### Reinforcement Sensitivity Theory Personality Questionnaire  {-}
 
@@
 
-### Heart rate variability
+### Heart rate variability {-}
 
@@
 
-### Musculoskeletal properties
+### Musculoskeletal properties {-}
 
@@
 
-### Postural stability
+### Postural stability {-}
 
@@
 
-### Injury
+### Injury {-}
 
 Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
 Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
@@
 
-## Procedure
+## Procedure {-}
 
@@
 
-### Questionnaires
+### Questionnaires {-}
 
@@
 
-### HRV
+### HRV {-}
 
@@
 
-### Muscle stiffness
+### Muscle stiffness {-}
 
@@
 
-## Data Analysis
+## Data Analysis {-}
 
@@
 \end{figure}
+
+In order to use discrete networks, continuous variables must first be split into categorical levels.
+When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
 
@@
 
-### Data Preparation
+### Data Preparation {-}
 
+Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
 To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
@@
 
-Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
-Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
-All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.
-
-For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
+A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary table ...) and required no further transformation prior to the median split.
+For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
 Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
 In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
 
-### Network structure
+### Network structure {-}
 
@@
 
-### Preliminary network structures
+### Preliminary network structures {-}
 
-Prior to the final network structure presented in the results section, several structures were investigated.
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explred.
 Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models. 
@@
 \hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
+HRV & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
 \hline
@@
 Identical blacklists to the first network were used for arcs between independent and explanatory variables.
-The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
+The "nlebase" variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
 
@@
 
-### Network Inference
+### Network Inference {-}
 
@@
 Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-
-# Results {-}
-
-# Results
-
-During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
-Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
-
-```{r table4, results='asis'}
-
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
-  kable(
-    .,
-    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
-    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
-    escape = F
-  ) %>%
-  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
-  kable_styling(latex_options = "HOLD_position")  %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r mainnetworkresults}
-set.seed(999)
-strengthfinder <- function(var1, var2){
-  str %>%
-    filter(from == var1 & to == var2)
-}
-
-from <- c("nlebase", "ind_team", "ind_team", "gender", "gender", "clevel", "pi")
-to <- c("RI_1", "hours", "nlebase", "stiffness_1", "stiffness_2", "balance_1", "stiffness_1")
-x <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-
-from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
-to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")
-
-y <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-```
-
-### First network structure
-
-The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
-Several of the explanatory variables showed strong connections with independent variables in the network.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
-The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
-Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
-Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
-The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
-The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
-In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
-For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
-### Markov blanket for injured_1
-
-```{r mb1}
-
-set.seed(1001)
-var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")
-
-query1 <- map_dfr(var, ~ newprobtable(
-  tallc,
-  .x,
-  "injured_1",
-  "injured",
-  avg30
-)) %>%
-  gather(variable, state, -prob, -outcome) %>%
-  drop_na() %>%
-  select(variable, state, prob) %>%
-  spread(state, prob) %>%
-  select(variable, Low, High) %>%
-  dendroTools::round_df(2)
-
-query2 <- newprobtable(
-  tallc,
-  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
-  "injured_1",
-  "injured",
-  avg30
-) %>%
-  toptail(n = 3) %>%
-  arrange(-prob) %>%
-  select(outcome, everything()) %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
-The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
-
-```{r}
-# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
-# q1p <- graphviz.plot(plotsg1,
-#                      shape = "rectangle",
-#                      render = F,
-#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
-# 
-# strq1 <- str %>% 
-#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") & 
-#            to %in% c("balance_1", "injured_1")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq1$strength)
-# labels[4] <- str_pad(labels[4], side = "left", width = 2)
-# 
-# edgenames <- edgeNames(q1p)
-# names(labels) <- edgenames
-# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
-# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6, 
-#                                     "stiffness_1~injured_1" = 3,
-#                                     "clevel~balance_1" = 3),
-#                             lty = c("injured_1~balance_1" = "dashed",
-#                                     "hours~injured_1" = "dashed"))
-# 
-# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 12)))
-# 
-# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
-# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
-# renderGraph(q1p)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-```{r table5, results='asis'}
-query1 %>%
-  select(c(1, 2, 3)) %>%
-  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
-Negative life events had a negligible effect when moving from the "Low" to "High" state.
-
-The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
-The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
-When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
-Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
-In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.
-
-```{r table6, results='asis'}
-
-query2 %>%
-  select(7, 2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
-The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}. 
-Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.
-
-```{r table7, results="asis"}
-injured1comp %>%
-  filter(variable != "outcome") %>% 
-  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(  .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
-    col.names = c("Variable", "High risk", "Low risk"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-### Markov blanket of injured_2
-
-```{r}
-set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2")
-query3 <- map_dfr(var2, ~newprobtable(tallc,
-                                      .x,
-                                      "injured_2",
-                                      "injured",
-                                      avg30)) %>% 
-  gather(variable, state, -prob, - outcome) %>% 
-  drop_na() %>% 
-  select(variable, state, prob) %>% 
-  spread(state, prob) %>% 
-  select(variable, Low, High) %>% 
-  dendroTools::round_df(2)
-
-query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
-                       "injured_2",
-                       "injured",
-                       avg30
-) %>% 
-  toptail(3) %>% 
-  arrange(-prob) %>% 
-  select(outcome, everything()) %>% 
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
-The arc between stiffness_2 and injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
-Results of the first query for injured_2 in the "injured" state are presented in table \ref{tab:table8}.
-Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
-FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
-"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
-
-```{r}
-# plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & 
-#            to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[6] <- str_pad(labels[6], side = "left", width = 2)
-# edgenames <- c("gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
-
-```{r table9, results="asis" }
-
-query4 %>%
-  select(7,2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-## Second network structure - changes within variables
-
-```{r include=FALSE, warning=FALSE}
-set.seed(1003)
-bmtable <- tidy(bm1, effects = "fixed") %>% 
-  select(term:conf.high) %>% 
-  dendroTools::round_df(2) %>% 
-  mutate_at(vars(conf.low, conf.high), list(~format(round(., 2), nsmall = 2))) %>%
-  mutate(conf.low = str_c("[", conf.low, sep = ""),
-         conf.high = str_c(conf.high, "]", sep = "")) %>% 
-  unite(., "95% CI", c("conf.low", "conf.high"), sep = ", ") %>% 
-  mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
-         "95% CI" = str_replace_all(`95% CI`, ",", ", ")) %>% 
-  slice(1:4) %>% 
-  mutate(term = c("Intercept", "RMSSD", "BIS", "RMSSD:BIS"))
-
-```
-
-The network for changes within variables is presented in Fig \ref{fig:fig9}.
-An arc between BIS $\rightarrow$ FFFS with strength 1 was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strength of 0.60.
-The arc between RMSSD $\rightarrow$ FFFS was 0.56.
-The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a linear regression model.
-While the continuous nodes in the second network are already parameterised using linear regressions, it was not possible to investigate the interaction between explanatory variables.
-The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
-
-```{r}
-# mb(avg30, "injured_2")
-# plotsg2 <- model2network('[nlelg_2][gender][pi][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|pi:gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("pi", "FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   slice(2:8) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[7] <- str_pad(labels[7], side = "left", width = 2)
-# edgenames <- c("pi~stiffness_2", "gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "pi~stiffness_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in RMSSD where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table10, results="asis"}
-
-bmtable %>%
-  kable(
-    .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
-    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r}
-set.seed(1005)
-test <- expand.grid(list(c(-1:1),
-                         c(-1:1)))
-query6 <- map2_dfr(.x = as.numeric(test$Var1),
-                   .y = as.numeric(test$Var2), ~tibble(prob= cpquery(fittedcont, 
-                                                                     event = (injured == "injured"),
-                                                                     evidence = list(nlec = .x,  stiffness = .y), method = "lw"),
-                                                       nle = .x,
-                                                       stiffness = .y))
-
-query6a <- query6 %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for the "injured" node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
-Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of `r query6a[9,1]` over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
-
-```{r}
-# plotsg3 <- model2network('[gender][pi][hours][injured|pi:hours][stiffness|gender:injured][nlec|injured]')
-# q3p <- graphviz.plot(plotsg3,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi", "hours"), "injured", c("stiffness", "nlec")))
-# strq3 <- strchange %>% 
-#   filter(from %in% c("pi", "hours", "injured", "gender") & 
-#            to %in% c("stiffness", "nlec", "injured")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq3$strength)
-# edgenames <- c("pi~injured", "gender~stiffness", "hours~injured",
-#                "injured~stiffness", "injured~nlec")
-# names(labels) <- edgenames
-# nodeRenderInfo(q3p)$fill[c("injured")] = "gold"
-# edgeRenderInfo(q3p) <- list(lwd = c("injured~stiffness" = 4, 
-#                                     "hours~injured" = 5,
-#                                     "gender~stiffness" = 4),
-#                             lty = c("pi~injured" = "dashed"))
-# 
-# q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q3p)$fill[c("nlec", "stiffness")] = "tomato"
-# nodeRenderInfo(q3p)$fill["injured"] = "gold"
-# renderGraph(q3p)
-# rm(plotsg3, strq3, labels)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-```{r table11, results="asis" }
-set.seed(1006)
-test <- expand.grid(list(
-  c(-1:1),
-  c(-1:1)
-))
-query6 <- map2_dfr(
-  .x = as.numeric(test$Var1),
-  .y = as.numeric(test$Var2),
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(nlec = .x, stiffness = .y),
-      method = "lw"
-    ),
-    nle = .x,
-    stiffness = .y
-  )
-)
-
-query6 %>%
-  select(nle, stiffness, prob) %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  arrange(-prob) %>%
-  mutate(prob = round(prob, digits = 2)) %>%
-  select(c(3, 1, 2)) %>%
-  kable(
-    .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
-
-```{r table12, results="asis"}
-set.seed(1007)
-testt <- expand.grid(list(
-  c(-1:1),
-  c(-1:1),
-  c("Low", "High"),
-  c("injury", "no injury")
-))
-
-query7 <- pmap_dfr(
-  .l = testt,
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(
-        nlec = as.numeric(..1),
-        stiffness = as.numeric(..2),
-        hours = ..3,
-        pi = ..4
-      ),
-      method = "lw"
-    ),
-    nle = ..1,
-    stiffness = ..2,
-    hours = ..3,
-    pi = ..4
-  )
-) %>%
-  toptail(3) %>%
-  arrange(-prob)
-
-query7 %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
-  ) %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-# Discussion
-
-# Disclosure/Conflict-of-Interest Statement {-}
-
-<!--  
-Frontiers follows the recommendations by the International Committee of Medical
-Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
-that all financial, commercial or other relationships that might be perceived by
-the academic community as representing a potential conflict of interest must be
-disclosed. If no such relationship exists, authors will be asked to declare that
-the research was conducted in the absence of any commercial or financial
-relationships that could be construed as a potential conflict of interest. When
-disclosing the potential conflict of interest, the authors need to address the
-following points:
-
- - Did you or your institution at any time receive payment or services from a
-   third party for any aspect of the submitted work?
- - Please declare financial relationships with entities that could be perceived
-   to influence, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- - Please declare patents and copyrights, whether pending, issued, licensed
-   and/or receiving royalties relevant to the work.
- - Please state other relationships or activities that readers could perceive to
-   have influenced, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- -->
-
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Author Contributions {-}
-
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-# Acknowledgments {-}
-
-Funding:
-
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# References
-
-A reference list should be automatically created here. However it won't. Pandoc 
-will place the list of references at the end of the document instead. There are 
-no convenient solution for now to force Pandoc to do otherwise. The easiest way 
-to get around this problem is to edit the LaTeX file created by Pandoc before 
-compiling it again using the traditional LaTeX commands.
-
-# Figures {-}
-
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```

### 2020-05-30 r-files/methods_code.R

@@ -25,10 +25,6 @@ x %>%
   group_by(time) %>% 
   count()
 
-datanet %>% 
-  group_by(time) %>% 
-  count()
-
 parchar <- read_csv("data/participant_chars.csv") %>% 
   filter(!id %in% id_remove)
 

### 2020-06-27 full_paper_v2.Rmd
before intro+methods edits

@@
 knitr::opts_chunk$set(echo = FALSE)
-source("r-files/functions.R")
-source("r-files/methods_code.R")
-source("r-files/results_code.R")
-```
-
-# Introduction {-}
+# source("r-files/functions.R")
+# source("r-files/methods_code.R")
+# source("r-files/results_code.R")
+load("all_data.RData")
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect, Rgraphviz)
+library(readxl)
+library(janitor)
+library(irr)
+library(lavaan)
+library(caret)
+library(parallel)
+library(zoo)
+library(sjmisc)
+
+```
+
+\section{Introduction}
 
@@
 
-# Methods {-}
+\section{Methods}
 
@@
 
-## Participants
+\subsection{Participants}
 
@@
 
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
-
 The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
@@
 Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
-The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.
-
-### Data Preparation
+The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
+
+### First network
+
+#### Data Preparation
 
@@
 A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
-All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary table ...) and required no further transformation prior to the median split.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
 For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
@@
 
-### Network structure
+#### Network structure
 
@@
 ```{r table2, results='asis'}
-exampledata <- tibble(
-  Participant = c("1", "1", "2", "3", "3"),
-  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-  " " = "->",
-  "X_2" = c("T2", "T3", "T2", "T2", "T3")
-)
-
-kable(
-  exampledata,
-  caption = "Example of the data arrangement used for the network.",
-  escape = TRUE
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-```
-
-### Preliminary network structures
+# exampledata <- tibble(
+#   Participant = c("1", "1", "2", "3", "3"),
+#   "X_1" = c("T1", "T2", "T1", "T1", "T2"),
+#   "link" = "->",
+#   "X_2" = c("T2", "T3", "T2", "T2", "T3")
+# )
+# 
+# kable(
+#   exampledata,
+#   caption = "Example of the data arrangement used for the network.",
+#   escape = FALSE
+# ) %>%
+#   kable_styling(latex_options = "HOLD_position") %>%
+#   row_spec(0, bold = TRUE)
+```
+
+#### Preliminary network structures
 
@@
 Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
-The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
-To obtain the structure, continuous variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
-Independent variables were then standardized to allow relative changes between variables to be compared. 
-The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
-Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
-The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
-
-To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
+Therefore, these arcs were blacked listed from the network.
+
+To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
 The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
@@
 
-### Network Inference
+#### Network Inference
 
@@
 
-The second network contained both continuous and discrete data. 
-To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
-Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-
-# Results {-}
+### Second network
+
+#### Data preparation
+
+For the second network, continuous variables suffixed with _1 were subtracted from variables suffixed with _2 to obtain the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
+Independent variables were then standardised (mean-centred at 0) to allow relative changes between variables to be compared. 
+The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
+Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
+
+#### Network structure
+
+Similar to the first network, blacklists were used to prevent arcs from independent variables $\rightarrow$ explanatory variables.
+In addition, "nlebase" variable was dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network. 
+The final network was obtained using the same approach as the first network.
+
+#### Network Inference
+
+Conditional probability queries (CPQ) were again used to perform inference the network structures.
+The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
+Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
+To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
+The samples were then examined using a Bayesian linear regression models with the _brms_ package [@Burkner2017a] to determine the relationship between these nodes. 
+Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
 
@@
 
-### First network structure
+## First network structure
 
@@
 
-The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
-
-```{r table9, results="asis" }
-
-query4 %>%
-  select(7,2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
+<!-- The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}. -->
+<!-- Again, only the three lowest and highest probabilities are shown (complete results in S5 Table). -->
+<!-- The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
+-->
+<!-- With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state. -->
+<!-- The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured". -->
+
+<!-- ```{r table9, results="asis" } -->
+
+<!-- query4 %>% -->
+<!--   select(7,2:6) %>% -->
+<!--   rename(Probability = "prob") %>% -->
+<!--   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>% -->
+<!--   mutate_all(~str_replace_all(., "_", "\\\\_"))%>% -->
+<!--   kable( -->
+<!--     ., -->
+<!--     caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.", -->
+<!--     escape = F -->
+<!--   ) %>% -->
+<!--   kable_styling(latex_options = "HOLD_position") %>% -->
+<!--   pack_rows("Highest", 1, 3) %>% -->
+<!--   pack_rows("Lowest", 4, 6) %>% -->
+<!--   row_spec(0,bold=TRUE) -->
+<!-- ``` -->
 
@@
 set.seed(1003)
-bmtable <- tidy(bm1, effects = "fixed") %>% 
-  select(term:conf.high) %>% 
+bmtable <- fixef(bm1) %>% 
+  as_tibble(rownames = "term") %>% 
+  `colnames<-`(c("Term", "Estimate", "Error", "conf.low", "conf.high")) %>% 
   dendroTools::round_df(2) %>%
@@
   mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
-         "95% CI" = str_replace_all(`95% CI`, ",", ", ")) %>% 
-  slice(1:4) %>% 
-  mutate(term = c("Intercept", "RMSSD", "BIS", "RMSSD:BIS"))
+         "95% CI" = str_replace_all(`95% CI`, ",", ", "))
 
@@
 
-# Disclosure/Conflict-of-Interest Statement {-}
-
-<!--  
-Frontiers follows the recommendations by the International Committee of Medical
-Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
-that all financial, commercial or other relationships that might be perceived by
-the academic community as representing a potential conflict of interest must be
-disclosed. If no such relationship exists, authors will be asked to declare that
-the research was conducted in the absence of any commercial or financial
-relationships that could be construed as a potential conflict of interest. When
-disclosing the potential conflict of interest, the authors need to address the
-following points:
-
- - Did you or your institution at any time receive payment or services from a
-   third party for any aspect of the submitted work?
- - Please declare financial relationships with entities that could be perceived
-   to influence, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- - Please declare patents and copyrights, whether pending, issued, licensed
-   and/or receiving royalties relevant to the work.
- - Please state other relationships or activities that readers could perceive to
-   have influenced, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- -->
+Informed by Appaneal and Perna’s (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature [@Johnson2014].
+Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables _across_ time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables _between_ time points (Fig \ref{fig:fig8}).
+
+The first network revealed several links between the injured nodes and other variables in the network.
+For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured_1 and injured_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
+The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for injured_1 and injured_2.
+The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury. 
+In the second network, the highest probability of injury was observed when changes in stiffness _and_ negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury [@Appaneal2014].
+Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
+Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury.
+The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
+[@Bittencourt2016], where injury is an emergent property from a web of determinants that interact in different ways.
+
+Of all the variables measured in the study, muscle stiffness appeared to be most strongly related to injury.
+Both "High" levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
+In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
+To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players [@PickeringRodriguez2017].
+The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
+However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness [@Pruyn2015; @Kalkhoven2018].
+Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury [@Butler2003].
+Additionally, high stiffness may only increase the risk of injury if other factors are also present.
+To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
+In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
+Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
+
+In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance $\rightarrow$ injured_1 and balance $\rightarrow$  injured_2 respectively.
+Despite the weak arc strength, a "High" balance score, indicating impaired postural stability, was found to increase the probability of injury.
+This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
+Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts [@Paillard2006].
+In the current study, athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
+These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition [@Hrysomallis2011].
+
+Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
+In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
+However, negative life event stress had almost no effect on the probability of injury in the first network.
+This finding suggests that the _relative change_ in life events may be more important than the _absolute score_ for life events, despite the latter being commonly used in sports injury research to date.
+For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
+Although the absolute score would be "Low", the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect [@Appaneal2014].
+Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed [cf. @Ivarsson2014].
+
+The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a _negative_ valence, as the strongest predictor of injury in Williams and Andersen's [@Williams1998] model [@Ivarsson2017].
+During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
+However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg_1 $\rightarrow$ injured_1 and nlelg_2 $\rightarrow$ injured_2 to allow this relationship to be examined more closely.
+When negative life events were "High" the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[3,3] - query3[3,2]` at the injured_2 node.
+One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
+In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
+However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred _since the previous data collection session_, to avoid inflated scores caused by reporting the same event on multiple occasions.
+The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
+For example, at the second and third time points, 26% of participants reported 0 negative life events for the preceding four-month period.
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events [cf.
+@Fawkner1999].
+
+Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
+Williams and Andersen's [@Williams1998] model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
+Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury [@Petrie1993; @Petrie2014; @Johnson2014].
+Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
+
+Of the remaining variables, both FFFS and RMSSD were also linked to injury.
+A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was weaker than muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
+Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, "Low" RMSSD increased the probability of injury from 0.17 (RMSSD = "High'') to 0.27 (RMSSD = "Low'').
+This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
+An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
+Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points; for example, the "Low" FFFS recorded at the first time point was related to injuries recorded at the second time point in the study.
+One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
+The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
+
+In the first network, "High" BIS was associated with "High" FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
+RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
+High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018], which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
+These findings align with the proposed actions of the RST theory [@Corr2016b]. 
+For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems. 
+The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict [@Corr2016b].
+To extend these findings, the BAS should also be considered. 
+Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
+However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
+In the first network, RI_1 was connected to both BIS_1 and BIS_2, and in both instances, the probability of "High" BIS was increased when RI was also "High". 
+However, the arcs between RI_1 and BIS were weak (< 0.50), and RI represents only one component of the BAS system. 
+Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury. 
+Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
+
+The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury [@Johnson2014].
+The longitudinal repeated measures design of the current study allowed _changes_ over time and between time points to be captured and explored.
+Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
+Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
+Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
+The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
+Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018].
+
+As with most research, there were several limitations with the present study.
+Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
+Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
+For example, the meaning of a "Low" and "High" value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
+A second limitation was the nature of the physiological measures used in the current study.
+In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
+For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis [@Ross2011].
+
+In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
+Although the current study used a range of measures to capture "stress" from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
+For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], and could help elucidate the relationship between stress and injury.
+Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
+For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
+Both responses would be scored a "-4"; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
+Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
+
+In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
+Specifically, muscle stiffness and _increases_ in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
+
+# Disclosure/Conflict-of-Interest Statement
 
@@
 
-# Author Contributions {-}
+# Author Contributions
 
@@
 
-# Acknowledgments {-}
-
-Funding:
-
 # Supplemental Data
@@
 
-# Figures {-}
+# Figures
 
@@
 ```
+
+\begin{figure}[!h]
+\caption{{\bf mBESS positions (A-F).}
+Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
+\label{fig:fig2}
+\end{figure}

### 2020-07-20 apa_style.Rmd
pre results edits

--- 2020-06-27 full_paper_v2.Rmd
@@
 ---
-title: "An interdisciplinary examination of stress and injury occurrence in athletes"
-runtitle: "Short Title"
-documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
+title             : "An interdisciplinary examination of stress and injury occurrence in athletes"
+shorttitle        : "Stress and sports injury"
+
 author: 
-  - name: Dr Harry Fisher
-    affiliation: '1'
-    etal: Fisher # First author's last name. 
-    email: harryfisher21@gmail.com # Indicates corresponding Author
-    institution: Cardiff Metropolitan University
-    street: Cyncoed Road
-    zip: CF23 6XD
-    country: United Kingdom
+  - name          : "Harry Fisher"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       :  "Cyncoed Road, Cardiff, CF23 6XD"
+    email         : "harryfisher21@gmail.com"
   - name: Dr Marianne Gittoes
@@
     affiliation: '3'
+
 affiliation:
@@
 
-output: rticles::frontiers_article
-csl: frontiers.csl
-bibliography: library.bib
----
-
-\begin{abstract}
-
+abstract: |
  This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
+At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 
-\tiny
- \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}
-
-\end{abstract}
+keywords          : "Sports Injury, Stress, Interdisciplinary, Bayesian Network"
+
+bibliography      : [library.bib, r-references.bib]
+
+floatsintext      : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_pdf
+csl: frontiers.csl
+always_allow_html: true
+---
+
+```{r setup, include = FALSE}
+library("papaja")
+options(kableExtra.auto_format = FALSE)
+```
 
@@
 options(scipen=999, digits = 2)
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 # source("r-files/functions.R")
@@
 
-Over the last four decades sport related injuries have received increased research attention [@Ivarsson2017].
-This attention is unsurprising given the high incidence [@Rosa2014; @Sheu2016],
-and undesirable physical and psychological effects of sports injuries [@Leddy1994; @Brewer2012].
-To mitigate against both the increasing incidence and	undesirable consequences of injury, research  has identified several psychological [@Slimani2018],
-anatomical [@Murphy2003],
-biomechanical [@Neely1998; @Hughes2014]
-and environmental [@Meeuwisse2007]
-factors associated with sports injury occurrence.
-Indeed, several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009],
-of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
-
-\begin{figure}[!h]
-\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
-\label{fig:fig1}
-\end{figure}
-
-Williams and Andersen's [@Williams1998] model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) will contribute to their response, either interactively or in isolation.
-Central to the model is the stress response, which reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of many stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater risk of injury.
-
-Within Williams and Andersen’s [@Williams1998] model, major life events, a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007];
-specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
-However, personality traits and coping resources have also been	found to predict injury, with for example, athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to athletes with the opposite profile.
-[@Smith1990; @Lavallee1996; @Ivarsson2010].
-However, the amount of variance explained by the psychosocial factors proposed by the model has been modest, typically between 5 - 30% [@Galambos2005; @Ivarsson2010]; suggesting other factors are also likely to contribute to injury occurrence.
-
-While the psychosocial factors proposed in Williams and Andersen’s [@Williams1998] model have received the most research attention, the mechanisms through which these factors are proposed to exert their effect have remained under-investigated in the literature.
+Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012].
+Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
+
+Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
+The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
+
+Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
+Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.
+
+While the psychosocial factors in Williams and Andersen’s [@Williams1998] model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
 To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
 However, to date, the research has largely focused on attentional deficits [@Andersen1999; @Rogers2005; @Wilkerson2012a; @Swanik2007].
-For example, Andersen and Williams [@Andersen1999]
-measured peripheral and central vision during high and low stress conditions and found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following season.
-Indeed, Rodgers and Landers [@Rogers2005] supported Andersen and Williams's [@Andersen1999] earlier findings reporting that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
-However, the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control, remains to be explored [cf.
+For example, Andersen and Williams [@Andersen1999] found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
+Rodgers and Landers [@Rogers2005] further supported Andersen and Williams’s [@Andersen1999] earlier findings by identifying that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
+
+Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse [cf.
 @Williams1998].
-
-One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
-[@Meeuwisse2007; @Wiese-Bjornstal2009].
-For example, a large body of research indicates that training-related stress is also likely to be related to the stress response and injury occurrence [@Lee2017; @Djaoui2017],
-and may account for the unexplained variance from the psychological predictors of injury.
-Appaneal and Perna [@Appaneal2014]
-proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model and to address some of these issues.
-To elaborate, the BMSAIH aimed to clarify the mediating pathways between the stress response and injury, consider other health outcomes and behavioural factors that impact sports participation,
-and integrate the impact of training on athletes’ health [@Appaneal2014].
+One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response [@Meeuwisse2007; @Wiese-Bjornstal2009].
+For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence [@Lee2017; @Djaoui2017] and may account for the unexplained variance from the psychological predictors.
+Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation [@Appaneal2014].
 The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p.
 74] to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides a framework for future research to build on Williams and Andersen's [@Williams1998] model, by including other physiological markers of training-related stress, which together may provide greater insight into the injury process.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other research has identified additional markers of training-related stress that are associated with an increased risk of injury; for example,
-heart rate variability [@Bellenger2016; @Williams2017],
-postural stability [@Romero-Franco2014]
-and muscle stiffness [@Pruyn2015].
-However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
-Recently, Bittencourt et al. [@Bittencourt2016]
-suggested that to better understand the multifactorial nature of sports injuries, research needs to move away from studying risk factors in isolation and instead adopt a complex systems approach to injury.
-Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p. 3],
-where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
-
-A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables.
-One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables [@Scutari2014].
-The learned structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C being value z?).
-Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
-
-To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
-Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
-Given the exploratory nature of the current study the following objectives were defined:
-
-- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
-- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
-- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.
+Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen’s [@Williams1998] model whilst including other physiological markers of training-related stress.
+
+Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury.
+However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
+Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
+Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
+
+Recently, Bittencourt et al.
+[@Bittencourt2016] advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p.
+3], where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
+Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables [@Scutari2014].
+The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
+Furthermore, unlike regression or structural equation models, BN’s do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known [@Olmedilla2018].
+BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen’s [@Williams1998] stress injury model and Appaneal’s BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
+A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
 
@@
 
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
+A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
 Participants self-rated competitive level ranged from recreational to international standard. 
+They were engaged in training for their respective sports for at least five hours per week.
 A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at least five hours per week.
+All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study. 
 Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
@@
 
+### Injury
+
+Participants self-reported any injuries they sustained at each data collection during the study period.
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
+Injury status (did / did not sustain an injury) served as the main outcome measure.
+
 ### Major life events
 
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
+The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table). 
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
-Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
-Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
-A score for total life events was calculated by summing the absolute values for both negative and positive events.
+Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”.
+Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4  ($extremely\;negative$) and +4 ($extremely\;positive$).
+Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
+A score for total life events was also calculated by summing the absolute values for both negative and positive events.
 Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively.
 Petrie also provided evidence of discriminant, convergent and predictive validity.
-The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
-The composite reliability for the LESCA in this study was 0.84.
+For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+The composite reliability for the LESCA in this study was .84.
 
@@
 Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
-The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total score that was subsequently used for further analysis.
+The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
 The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
@@
 
-A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure muscle stiffness.
+A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure passive muscle stiffness.
 The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
@@
 A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
-To calculate limb asymmetry, the DL and NDL leg score was calculated by summing the DL and NDL errors respectively.
 A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
@@
 
-### Injury
-
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
-Injury status (did / did not sustain an injury) served as the main outcome measure.
-
 ## Procedure
 
-At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
-With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
-To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week.
-Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
-Each participant was asked to attend a total of 3 data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
+At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study.
+With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
+Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
+A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
 Participants provided informed consent before data collection commenced.
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
 
@@
 
-\begin{figure}[!h]
-\caption{{\bf Session protocol.}
-Outline of the protocol for each data collection.}
-\label{fig:fig4}
-\end{figure}
-
 ### Questionnaires
@@
 For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis purposes.
+The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis.
 
@@
 Participants were seated and asked to remain as still as possible for the duration of the recording. 
-No attempt was made to control participants respiratory frequency or tidal volume [@Denver2007]. 
+No attempt was made to control participants' respiratory frequency or tidal volume [@Denver2007]. 
 Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
@@
 A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one second interval [@Morgan2018].
-For each impulse, the device computed stiffness values, with the median of the five values being saved by the device for further analysis. 
+The device was programmed to deliver five consecutive impulses, separated by a one-second interval [@Morgan2018]. 
+For each impulse, the device computed passive stiffness values, with the median of the five values being saved by the device for further analysis.
 In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted.
-Sets of measurements above 3 % were measured again to ensure the reliability of the data. 
-The CV was calculated in real time by the device after each set of measurements. 
-Measurements saved on the device were uploaded to a computer using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis. 
-For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis.
+Sets of measurements above 3% were measured again to ensure data reliability.
+Measurements were uploaded using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis.
+For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis
 
@@
 The video recording was started prior to the first stance position and stopped after all stances had been completed.
-Each completed mBESS protocol took approximately 4 minutes.
+Each completed mBESS protocol took approximately four-minutes.
 Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
@@
 In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
-
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
 
@@
 Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
-In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure.
+In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
 Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
@@
 Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
-To account for the repeated measures design and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
+To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
 In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
@@
 Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
-Table \ref{tab:table2} provides an example of the formatted data. 
-Participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
+Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively.
 Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
@@
 ```{r table2, results='asis'}
-# exampledata <- tibble(
-#   Participant = c("1", "1", "2", "3", "3"),
-#   "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-#   "link" = "->",
-#   "X_2" = c("T2", "T3", "T2", "T2", "T3")
-# )
-# 
-# kable(
-#   exampledata,
-#   caption = "Example of the data arrangement used for the network.",
-#   escape = FALSE
-# ) %>%
-#   kable_styling(latex_options = "HOLD_position") %>%
-#   row_spec(0, bold = TRUE)
+exampledata <- tibble(
+  Participant = c("1", "1", "2", "3", "3"),
+  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
+  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
+)
+
+kable(
+  exampledata,
+  caption = "Example of the data arrangement used for the network.",
+  escape = FALSE
+) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0, bold = TRUE)
 ```
@@
 
-Prior to the final network structure presented in the results section, several structures with different combinations of variables were explred.
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
 Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models.
@@
 Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
-Additionally, despite strong evidence in the literature that both negative and total life event stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
 Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
-Total life event score was not included in the final structure.
+TLE score was not included in the final structure.
 
@@
 
-Finally, both total score and asymmetry were included in the initial network.
+Finally, both total score and asymmetry for balance were included in the initial network.
 However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
 Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
-
-```{r finalvars, results='asis'}
-library(knitr)
-library(kableExtra)
+To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
+
+```{r table3, results='asis'}
 varcutoffs %>% 
@@
   kable("latex",
-       booktabs = T,
        align = c("llcc"),
@@
 
-Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
-For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
-Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-Therefore, these arcs were blacked listed from the network.
-
+Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
+For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%.
+Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network.
+Therefore, these arcs were blacklisted from the network.
 To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
@@
 Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
-An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
-In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
-Arcs below 0.3 were not included in the model.
-The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction.
+An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
+In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
+Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
 
@@
 To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
+The query allows the odds of a particular node state (e.g., injured_1 = “injured”) to be calculated based on the new evidence.
 CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
 The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
-Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
-Of particular interest in the current study were the variables that were connected to "injured" nodes.
-To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
+Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
+Of particular interest were the variables that were connected to “injured” nodes, which were examined in the network using the Markov blanket of “injured_1” and “injured_2”.
 A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
+CPQ’s were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the “injured” state.
 
@@
 
-For the second network, continuous variables suffixed with _1 were subtracted from variables suffixed with _2 to obtain the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
-Independent variables were then standardised (mean-centred at 0) to allow relative changes between variables to be compared. 
+Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
+
+For the second network, change scores for continuous variables between  T1 $\rightarrow$  T2 and  T2 $\rightarrow$  T3 were standardized to allow relative changes between variables to be compared.
 The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
+Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from "Low" to "High".
 
@@
 Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
-
-```{r table4, results='asis'}
-
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
-  kable(
-    .,
-    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
-    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
-    escape = F
-  ) %>%
-  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
-  kable_styling(latex_options = "HOLD_position")  %>%
-  row_spec(0,bold=TRUE)
-
-```
 
@@
   mutate(prob = format(round(prob, digits = 2), nsmall = 2))
+```
+
+```{r table4, results='asis'}
+
+cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
+  kable(
+    .,
+    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
+    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
+    escape = F
+  ) %>%
+  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
+  footnote(general = "Other included bone, skin and brain injuries.") %>%
+  kable_styling(latex_options = "HOLD_position")  %>%
+  row_spec(0,bold=TRUE)
+
 ```
@@
 For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
+The conditional probability query (CPQ) for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
+The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
+Negative life events had a negligible effect when moving from the "Low" to "High" state.
 
@@
 
-The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
-Negative life events had a negligible effect when moving from the "Low" to "High" state.
-
 The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
@@
 
-Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
-The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}. 
-Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.
-
-```{r table7, results="asis"}
-injured1comp %>%
-  filter(variable != "outcome") %>% 
-  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(  .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
-    col.names = c("Variable", "High risk", "Low risk"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
+<!-- Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
+-->
+<!-- The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}.  -->
+<!-- Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state. -->
+
+<!-- # ```{r table7, results="asis"} -->
+<!-- # injured1comp %>% -->
+<!-- #   filter(variable != "outcome") %>%  -->
+<!-- #   mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>%  -->
+<!-- #   mutate_all(~str_replace_all(., "_", "\\\\_"))%>% -->
+<!-- #   kable(  ., -->
+<!-- #     caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.", -->
+<!-- #     col.names = c("Variable", "High risk", "Low risk"), -->
+<!-- #     escape = F -->
+<!-- #   ) %>% -->
+<!-- #   kable_styling(latex_options = "HOLD_position") %>% -->
+<!-- #   row_spec(0,bold=TRUE) -->
+<!-- # ``` -->
 
@@
 The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
-The arc between stiffness_2 and injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
+The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
 Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
@@
 "High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
+
+```{r table8, results="asis"}
+
+query3 %>%
+  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
+  kable(
+    .,
+    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
+    col.names = c("Variable", "Low", "High"),
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+```
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
+The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
+Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
+The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
+With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
+The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
+
+```{r table9, results="asis" }
+
+query4 %>%
+  select(7,2:6) %>%
+  rename(Probability = "prob") %>%
+  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
   mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
@@
     .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
+    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
     escape = F
@@
   kable_styling(latex_options = "HOLD_position") %>%
+  pack_rows("Highest", 1, 3) %>%
+  pack_rows("Lowest", 4, 6) %>%
   row_spec(0,bold=TRUE)
 ```
-
-<!-- The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}. -->
-<!-- Again, only the three lowest and highest probabilities are shown (complete results in S5 Table). -->
-<!-- The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
--->
-<!-- With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state. -->
-<!-- The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured". -->
-
-<!-- ```{r table9, results="asis" } -->
-
-<!-- query4 %>% -->
-<!--   select(7,2:6) %>% -->
-<!--   rename(Probability = "prob") %>% -->
-<!--   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>% -->
-<!--   mutate_all(~str_replace_all(., "_", "\\\\_"))%>% -->
-<!--   kable( -->
-<!--     ., -->
-<!--     caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.", -->
-<!--     escape = F -->
-<!--   ) %>% -->
-<!--   kable_styling(latex_options = "HOLD_position") %>% -->
-<!--   pack_rows("Highest", 1, 3) %>% -->
-<!--   pack_rows("Lowest", 4, 6) %>% -->
-<!--   row_spec(0,bold=TRUE) -->
-<!-- ``` -->
 
@@
 The arc between RMSSD $\rightarrow$ FFFS was 0.56.
-The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a linear regression model.
-While the continuous nodes in the second network are already parameterised using linear regressions, it was not possible to investigate the interaction between explanatory variables.
+The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
 The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
@@
 
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
 Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
@@
 ```{r table10, results="asis"}
-
 bmtable %>%
@@
 With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
+
+```{r table11, results="asis" }
+set.seed(1006)
+test <- expand.grid(list(
+  c(-1:1),
+  c(-1:1)
+))
+# query6 <- map2_dfr(
+#   .x = as.numeric(test$Var1),
+#   .y = as.numeric(test$Var2),
+#   ~ tibble(
+#     prob = cpquery(
+#       fittedcont,
+#       event = (injured == "injured"),
+#       evidence = list(nlec = .x, stiffness = .y),
+#       method = "lw"
+#     ),
+#     nle = .x,
+#     stiffness = .y
+#   )
+# )
+
+query6 %>%
+  select(nle, stiffness, prob) %>%
+  mutate_at(vars(nle, stiffness), ~ case_when(
+    . == 1 ~ "+1SD",
+    . == 0 ~ "mean",
+    . == -1 ~ "-1SD"
+  )) %>%
+  arrange(-prob) %>%
+  mutate(prob = round(prob, digits = 2)) %>%
+  select(c(3, 1, 2)) %>%
+  kable(
+    .,
+    col.names = c("Probability", "nlec", "Stiffness"),
+    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+
+```
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-```{r table11, results="asis" }
-set.seed(1006)
-test <- expand.grid(list(
-  c(-1:1),
-  c(-1:1)
-))
-query6 <- map2_dfr(
-  .x = as.numeric(test$Var1),
-  .y = as.numeric(test$Var2),
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(nlec = .x, stiffness = .y),
-      method = "lw"
-    ),
-    nle = .x,
-    stiffness = .y
-  )
-)
-
-query6 %>%
-  select(nle, stiffness, prob) %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  arrange(-prob) %>%
-  mutate(prob = round(prob, digits = 2)) %>%
-  select(c(3, 1, 2)) %>%
-  kable(
-    .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
 Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
@@
 
+## Summary fo results
+
+(not sure where this would go /if needed at all?).
+The first BN revealed both high stiffness and poor balance resulted in an increased probability of injury.
+However, contrary to previous literature, negative life event stress had a negligible effect on the probability of injury.
+In the second network that modelled changes between time points in the study, increases in muscle stiffness was found to increase the risk of injury.
+In addition, the probability of injury was further increased when there were increases in NLE stress.
+
+(Main discussion starts here..)
+
 Informed by Appaneal and Perna’s (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature [@Johnson2014].
@@
 For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured_1 and injured_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
-The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for injured_1 and injured_2.
+The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for "injured" nodes.
 The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury. 
 In the second network, the highest probability of injury was observed when changes in stiffness _and_ negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury [@Appaneal2014].
-Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
-Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury.
-The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
+
+<!-- Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
+-->
+<!-- Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury. -->
+<!-- The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
 [@Bittencourt2016], where injury is an emergent property from a web of determinants that interact in different ways.
+-->
 
@@
 Of the remaining variables, both FFFS and RMSSD were also linked to injury.
-A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was weaker than muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
+A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was not as certain as muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
 Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, "Low" RMSSD increased the probability of injury from 0.17 (RMSSD = "High'') to 0.27 (RMSSD = "Low'').
@@
 An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
-Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points; for example, the "Low" FFFS recorded at the first time point was related to injuries recorded at the second time point in the study.
+Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points.
 One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
@@
 
-# References
-
-A reference list should be automatically created here. However it won't. Pandoc 
-will place the list of references at the end of the document instead. There are 
-no convenient solution for now to force Pandoc to do otherwise. The easiest way 
-to get around this problem is to edit the LaTeX file created by Pandoc before 
-compiling it again using the traditional LaTeX commands.
-
 # Figures
@@
 ```
+
+\begin{figure}[!h]
+\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
+\label{fig:fig1}
+\end{figure}
 
@@
 \end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Study design.}
+For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
+\label{fig:fig3}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Session protocol.}
+Outline of the protocol for each data collection.}
+\label{fig:fig4}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Example network.}
+A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
+\label{fig:fig5}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Network structure.}
+The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+\label{fig:fig6}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket of injured\_1.}
+Arc strengths are included as arc labels.}
+\label{fig:fig7}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for injured\_2.} }
+\label{fig:fig8}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Change network.}
+Network structure of the changes within variables between time points.}
+\label{fig:fig9}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
+\label{fig:fig10}
+\end{figure}
+
+# Tables
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup

### 2020-07-20 apa_style.tex
pre results edits

--- 2020-05-30 demo.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
-%                                              %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Version 3.4 Generated 2018/06/15 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-
-\documentclass[utf8]{frontiersHLTH}
-
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-
-\linenumbers
-
-% BELOW TAKEN FROM rticles plos template
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
 %
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% hyperref package, useful for hyperlinks
-\usepackage{hyperref}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
+\documentclass[
+  english,
+  man]{apa6}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={An interdisciplinary examination of stress and injury occurrence in athletes},
+  pdflang={en-EN},
+  pdfkeywords={Sports Injury, Stress, Interdisciplinary, Bayesian Network},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
-
-% Sweave(-like)
-\usepackage{fancyvrb}
-\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
-\newenvironment{Schunk}{}{}
-\DefineVerbatimEnvironment{Code}{Verbatim}{}
-\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-\newenvironment{CodeChunk}{}{}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% Below is from frontiers
-%
-\bibliographystyle{frontiersinSCNS}
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace}
-%\doublespacing
-
-% Leave a blank line between paragraphs instead of using \\
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
-
-\usepackage{booktabs}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+% Manuscript styling
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
+
+% Table formatting
 \usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\def\Authors{
-  Dr Harry Fisher\,\textsuperscript{1*},
-  Dr Marianne Gittoes\,\textsuperscript{1},
-  Professor Lynne Evans\,\textsuperscript{1},
-  Miss Leah Bitchell\,\textsuperscript{1},
-  Dr Richard Mullen\,\textsuperscript{2},
-  Dr Marco Scutari\,\textsuperscript{3}}
-
-\def\Address{
-
-  \textsuperscript{1} School of Sport and Health Science, Cardiff Metropolitan University,  Cardiff,   United Kingdom
-
-  \textsuperscript{2} Division of Sport, Health \& Exercise Sciences, Brunel University,  London,   United Kingdom
-
-  \textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA),  Manno,   Switzerland
-  }
-
-  \def\corrAuthor{Dr Harry Fisher}\def\corrAddress{Cardiff Metropolitan University\\Cyncoed Road\\, CF23 6XD United Kingdom}\def\corrEmail{\href{mailto:harryfisher21@gmail.com}{\nolinkurl{harryfisher21@gmail.com}}}
-  \def\firstAuthorLast{Fisher {et~al.}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Stress and sports injury]{An interdisciplinary examination of stress and injury occurrence in
-athletes}
-\author[\firstAuthorLast]{\Authors}
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
-
-\maketitle
-
-\begin{abstract}
-
+\usepackage{lscape}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
+
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
+
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\makeatother
+\shorttitle{Stress and sports injury}
+\author{Harry Fisher\textsuperscript{1}, Dr Marianne Gittoes\textsuperscript{1}, Professor Lynne Evans\textsuperscript{1}, Miss Leah Bitchell\textsuperscript{1}, Dr Richard Mullen\textsuperscript{2}, \& Dr Marco Scutari\textsuperscript{3}}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} Cardiff Metropolitan University, Cardiff, United Kingdom\\\textsuperscript{2} Brunel University, London, United Kingdom\\\textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA), Manno, Switzerland}
+\authornote{
+
+Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: harryfisher21@gmail.com}
+\keywords{Sports Injury, Stress, Interdisciplinary, Bayesian Network}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
+\usepackage{lineno}
+
+\linenumbers
+\usepackage{csquotes}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
+\title{An interdisciplinary examination of stress and injury occurrence in athletes}
+
+\date{}
+
+\abstract{
 This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, an
-d reported any injuries they had sustained since the last data collection.
+At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related markers and injury occurrence.
-
-\tiny
- \keyFont{ \section{Keywords:} Sports Injury, Stress, Interdisciplinary, Bayesian Network}
-
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-Over the last four decades sport related injuries have received
-increased research attention (Ivarsson et al., 2017). This attention is
-unsurprising given the high incidence (Rosa et al., 2014; Sheu et al.,
-2016), and undesirable physical and psychological effects of sports
-injuries (Leddy et al., 1994; Brewer, 2012). To mitigate against both
-the increasing incidence and undesirable consequences of injury,
-research has identified several psychological (Slimani et al., 2018),
-anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes,
-2014) and environmental (Meeuwisse et al., 2007) factors associated with
-sports injury occurrence. Indeed, several models of injury causation
-have been proposed that highlight the multifactorial nature of injury
-occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009),
-of which one of the most widely cited was developed by Williams and
-Anderson (Fig \ref{fig:fig1}; Andersen and Williams, 1988; Williams and
-Andersen, 1998).
+}
+
+\begin{document}
+\maketitle
+
+\section{Introduction}
+
+Over the last four decades sport-related injuries have received increased research attention (Ivarsson et al., 2017) in response to the high incidence (Rosa et al., 2014; Sheu et al., 2016) and associated undesirable physical and psychological effects (Leddy et al., 1994; Brewer, 2012).
+Multiple psychological (Slimani et al., 2018), anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes, 2014) and environmental (Meeuwisse et al., 2007) factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009), of which one of the most widely cited was developed by Williams and Anderson (Fig \ref{fig:fig1}; Andersen and Williams, 1988; Williams and Andersen, 1998).
+
+Williams and Andersen's (Williams and Andersen, 1998) stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete's personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
+The stress response is central to the model and reflects the bi-directional relationship between athletes' appraisal of, and response to, a stressful athletic situation.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
+
+Within Williams and Andersen's (Williams and Andersen, 1998) model, major life events, which is a component of an athlete's history of stressors, most consistently predicts injury occurrence (Williams and Andersen, 2007); specifically, major life events with a negative, as opposed to positive, valence (Passer and Seese, 1983; Maddison and Prapavessis, 2005).
+Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile (Smith et al., 1990; Lavallée and Flint, 1996; Ivarsson and Johnson, 2010).However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30\% (Galambos et al., 2005; Ivarsson and Johnson, 2010), which indicates a likely interaction with other factors.
+
+While the psychosocial factors in Williams and Andersen's (Williams and Andersen, 1998) model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
+To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
+However, to date, the research has largely focused on attentional deficits (Andersen and Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et al., 2007).
+For example, Andersen and Williams (Andersen and Williams, 1999) found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
+Rodgers and Landers (Rogers and Landers, 2005) further supported Andersen and Williams's (Andersen and Williams, 1999) earlier findings by identifying that peripheral narrowing under stress mediated 8.1\% of the relationship between negative life events and injury.
+
+Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse (cf.~Williams and Andersen, 1998).
+One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009).
+For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence (Lee et al., 2017; Djaoui et al., 2017) and may account for the unexplained variance from the psychological predictors.
+Appaneal and Perna (Appaneal and Perna, 2014) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (Williams and Andersen, 1998) model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation (Appaneal and Perna, 2014).
+The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and ``widen the window of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of undesirable health outcomes including illness and injury.
+Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen's (Williams and Andersen, 1998) model whilst including other physiological markers of training-related stress.
+
+Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other training-related stress markers including heart rate variability (Bellenger et al., 2016; Williams et al., 2017), muscle stiffness (Pruyn et al., 2015) and postural stability (Romero-Franco et al., 2014) have been reported to be associated with an increased risk of injury.
+However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries (Johnson et al., 2014).
+Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
+Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
+
+Recently, Bittencourt et al.~(Bittencourt et al., 2016) advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex ``web of determinants'' (Bittencourt et al., 2016, 3), where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
+Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables (Scutari and Denis, 2014).
+The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
+Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known (Olmedilla et al., 2018).
+BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen's (Williams and Andersen, 1998) stress injury model and Appaneal's BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
+A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
+
+\section{Methods}
+
+\subsection{Participants}
+
+A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants had an average age of 22.0 \(\pm\) 7.0 years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
+Participants self-rated competitive level ranged from recreational to international standard.
+They were engaged in training for their respective sports for at least five hours per week.
+A total of 126 (49.03\%) participants had sustained an injury in the 12 months prior to the start of the study (male: \emph{n} = 114 {[}49\%{]}, female; \emph{n} = 48 {[}40\%{]}).
+All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
+Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
+
+\begin{table}[H]
+
+\caption{\label{tab:table1}Participant characteristics.}
+\centering
+\begin{tabular}[t]{c|c|c}
+\hline
+\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
+\hline
+\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
+\hline
+\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
+\hline
+\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
+\hline
+\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
+\hline
+\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
+\hline
+\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
+\hline
+\hspace{1em}Recreational & 3 (4) & 7 (4)\\
+\hline
+\hspace{1em}University & 45 (56) & 141 (80)\\
+\hline
+\hspace{1em}National/International & 33 (41) & 28 (16)\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{measures}{%
+\subsection{Measures}\label{measures}}
+
+\hypertarget{injury}{%
+\subsubsection{Injury}\label{injury}}
+
+Participants self-reported any injuries they sustained at each data collection during the study period.
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours (Fuller et al., 2006, 2007; Timpka et al., 2014).
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries (cf.~Appaneal et al., 2009).
+Injury status (did / did not sustain an injury) served as the main outcome measure.
+
+\hypertarget{major-life-events}{%
+\subsubsection{Major life events}\label{major-life-events}}
+
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress (Petrie, 1992).
+The LESCA is the most widely used measure of major life events for athletes in the sports injury literature.
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table).
+The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
+Example items include, ``Major change in the frequency (increased or decreased) of social activities due to participation in sport'', ``Major change in the amount (more or less) of academic activity (homework, class time, etc)'' and ``Major change in level of athletic performance in actual competition (better or worse)''.
+Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4 (\(extremely\;negative\)) and +4 (\(extremely\;positive\)).
+Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
+A score for total life events was also calculated by summing the absolute values for both negative and positive events.
+Petrie (1992) (Petrie, 1992) reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001) respectively.
+Petrie also provided evidence of discriminant, convergent and predictive validity.
+For this study, composite reliability (Fornell and Larcker, 1981) was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+The composite reliability for the LESCA in this study was .84.
+
+\hypertarget{reinforcement-sensitivity-theory-personality-questionnaire}{%
+\subsubsection{Reinforcement Sensitivity Theory Personality Questionnaire}\label{reinforcement-sensitivity-theory-personality-questionnaire}}
+
+A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology (Corr and Cooper, 2016).
+The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I am the sort of person who easily freezes-up when scared''), Behavioural Inhibition System (BIS; e.g., ``When trying to make a decision, I find myself constantly chewing it over'') and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., ``I regularly try new activities just to see if I enjoy them''), Goal Drive Persistence (e.g., ``I am very persistent in achieving my goals''), Reward Reactivity (e.g., ``I get a special thrill when I am praised for something I've done well'') and Impulsivity (e.g., ``I find myself doing things on the spur of the moment'').
+Participants rated each item on a scale from 1 (\(not\; at\; all\)) to 4 (\(highly\)) to reflect how well each statement described their personality in general.
+The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
+The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in S1 Appendix.
+
+\hypertarget{heart-rate-variability}{%
+\subsubsection{Heart rate variability}\label{heart-rate-variability}}
+
+A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings (Giles et al., 2016), which are considered the gold standard for assessing HRV.
+In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes (Caminal et al., 2018) and under spontaneous breathing conditions (Plews et al., 2017).
+
+\hypertarget{musculoskeletal-properties}{%
+\subsubsection{Musculoskeletal properties}\label{musculoskeletal-properties}}
+
+A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure passive muscle stiffness.
+The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
+The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
+The raw oscillation signal is then processed, and the stiffness parameter is calculated (Agyapong-Badu et al., 2016).
+The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties (Chuang et al., 2013; Pruyn et al., 2016; Nair et al., 2014), and has demonstrated good internal consistency (coefficient of variation \textless{} 1.4\%) over sets of 10 repetitions (Aird et al., 2012).
+
+\hypertarget{postural-stability}{%
+\subsubsection{Postural stability}\label{postural-stability}}
+
+Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by Hunt et al. (2009).
+In total, each trial of the mBESS was performed without shoes (McCrory et al., 2013) and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
+non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion)
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}).
+To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (cf.~Cingel et al., 2017).
+Participants were asked to maintain each stance for a total of 20-seconds.
+Participants hands were placed on hips at the level of the iliac crests.
+A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
+
+The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials.
+One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests;
+b) opening eyes;
+c) stepping, stumbling, or falling;
+d) moving the thigh into more than 30 degrees of flexion or abduction;
+e) lifting the forefoot or heel; and
+f) remaining out of the testing position for more than 5-seconds (Riemann et al., 1999).
+A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
+A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
+A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces).
+To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation (ICC; Koo and Li, 2016) was used on a sample of 40 participants from the first time point.
+The test-retest scoring of BESS resulted in a ``good'' to ``excellent'' ICC score (ICC = 0.93, 95\% confidence interval = 0.88 - 0.96), indicating the scoring was reliable.
+
+\hypertarget{procedure}{%
+\subsection{Procedure}\label{procedure}}
+
+At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study.
+With the coaches' permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
+Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
+A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
+Participants provided informed consent before data collection commenced.
+
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}).
+To ensure all measures could be collected within a viable time-frame, participants were separated into two groups.
+The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures.
+Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
+
+\hypertarget{questionnaires}{%
+\subsubsection{Questionnaires}\label{questionnaires}}
+
+The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
+The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
+For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
+The data were downloaded from surverymonkey.com and imported into R (R Core Team, 2019) for analysis.
+
+\hypertarget{hrv}{%
+\subsubsection{HRV}\label{hrv}}
+
+To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
+Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
+Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines.
+Participants were seated and asked to remain as still as possible for the duration of the recording.
+No attempt was made to control participants' respiratory frequency or tidal volume (Denver et al., 2007).
+Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
+
+Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R (R Core Team, 2019) where the \emph{RHRV} package (Rodriguez-Linares et al., 2019) was used to calculate HRV indices.
+Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings (Laborde et al., 2017; Malik et al., 1996).
+RMSSD was calculated as:
+
+\begin{equation} 
+  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
+\end{equation}
+
+Where N is the length of the time series, and \(RR_i\) the RR interval between beats \(i\) and \(i-1\), where each beat position corresponds to the beat detection instant.
+
+\hypertarget{muscle-stiffness}{%
+\subsubsection{Muscle stiffness}\label{muscle-stiffness}}
+
+To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb.
+The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site (Chuang et al., 2012).
+The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle.
+The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
+A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.
+
+After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site.
+A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
+The device was programmed to deliver five consecutive impulses, separated by a one-second interval (Morgan et al., 2018).
+For each impulse, the device computed passive stiffness values, with the median of the five values being saved by the device for further analysis.
+In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3\% was accepted.
+Sets of measurements above 3\% were measured again to ensure data reliability.
+Measurements were uploaded using MyotonPRO software and imported in R (R Core Team, 2019) for further analysis.
+For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis
+
+\hypertarget{postural-stability-1}{%
+\subsubsection{Postural stability}\label{postural-stability-1}}
+
+Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
+For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds.
+Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
+The video recording was started prior to the first stance position and stopped after all stances had been completed.
+Each completed mBESS protocol took approximately four-minutes.
+Only one trial was performed to avoid familiarisation effects across the repeated measurement (cf.~Valovich et al., 2003).
+The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury.
+A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way (Scutari and Denis, 2014).
+A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter (Fig \ref{fig:fig5}; Pearl, 1988).
+However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual (Scutari and Denis, 2014).
+The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN (Chen and Pollino, 2012).
+In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
+
+In order to use discrete networks, continuous variables must first be split into categorical levels.
+When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100 (10\^{}2) possible parameter combinations, however the same network with three levels has 1000 (10\^{}3) possible parameter combinations, the latter being significantly more computationally expensive.
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio (Scutari and Denis, 2014).
+Typically, no more than three levels have been used in Bayesian networks in the sports injury literature (Olmedilla et al., 2018; Ruiz-Pérez et al., 2019)
+
+Learning the structure of the network is an important step in BN modelling.
+The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari \& Denis, 2014).
+The learned structure can then be used for inference by querying the network\footnote{The term ``query'' in relation to Bayesian Networks stems from Pearl's expert systems theory (1988).
+A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
+Widely used texts on Bayesian Network analysis (Koller and Friedman, 2009) have widely adopted the terminology in favour of that used in traditional statistics.} and obtaining the posterior probabilities of a particular node for a given query.
+The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (Scutari and Denis, 2014).
+In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to ``Low'') could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
+In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network.
+BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
+
+BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several benefits over traditional statistical analysis.
+For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (Olmedilla et al., 2018; Bittencourt et al., 2016).
+Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (Xiao-xuan et al., 2008).
+In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
+Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (Heckerman et al., 1995; Xu et al., 2015).
+The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
+
+\hypertarget{first-network}{%
+\subsubsection{First network}\label{first-network}}
+
+\hypertarget{data-preparation}{%
+\paragraph{Data Preparation}\label{data-preparation}}
+
+Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
+To prepare the data for the BN, missing values in the dataset were first imputed.
+Out of the 650 total measurements across all time points in the current study, there were 31 (4.64\%) missing muscle stiffness measurements and 70 (10.48\%) missing heart rate recordings.
+The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
+A missing rate of 15-20\% has been reported to be common in psychological studies, and several techniques are available to handle missing values (Enders, 2003; Lang et al., 2013).
+In the current study, the \emph{caret} package (Kuhn et al., 2008) was used to impute the missing values.
+A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset.
+The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation (Kuhn et al., 2008).
+
+A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
+For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
+Given the limited support for a relationship between positive life events and injury (Williams and Andersen, 2007), only negative and total life events were included in the network.
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
+In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (Ivarsson et al., 2017).
+
+\hypertarget{network-structure}{%
+\paragraph{Network structure}\label{network-structure}}
+
+To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
+Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
+Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (Scutari and Denis, 2014).
+To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (Murphy, 2002).
+In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
+However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
+Variables were separated into two groups; ``explanatory'', for variables that did not change during the study (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
+Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
+Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
+Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3, respectively.
+Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
+In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
+Finally, arc direction was restricted between specific nodes within the explanatory variables.
+Arcs from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and nlebase \(\rightarrow\) ind\_team were included in the blacklist, as arcs in these directions did not make logical sense.
+All subsequent models used the same blacklist.
+
+\begin{table}[H]
+
+\caption{\label{tab:table2}Example of the data arrangement used for the network.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+\textbf{Participant} & \textbf{X\_1} & \textbf{X\_2}\\
+\hline
+1 & T1 & T2\\
+\hline
+1 & T2 & T3\\
+\hline
+2 & T1 & T2\\
+\hline
+3 & T1 & T2\\
+\hline
+3 & T2 & T3\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{preliminary-network-structures}{%
+\paragraph{Preliminary network structures}\label{preliminary-network-structures}}
+
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
+Networks were learned using a Tabu search algorithm (Russell and Norvig, 2009) and BIC was used to compare different models.
+A higher BIC value indicates the structure of a DAG is a better fit to the observed data (Scutari and Denis, 2014).
+BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
+
+Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -218.31).
+Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (Williams and Andersen, 2007), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2 in the final network structure.
+TLE score was not included in the final structure.
+
+The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network.
+Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
+RI provided the highest BIC value, compared to RR (-9.07), GDP (-15.06) and I (-16.33).
+Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -894.33) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
+Therefore, only RI was included in the final structure.
+
+Finally, both total score and asymmetry for balance were included in the initial network.
+However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
+Therefore, balance asymmetry was removed from the final network structure.
+To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
+
+\begin{table}[H]
+
+\caption{\label{tab:table3}Variables included in the final Bayesian network structure.}
+\centering
+\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
+\hline
+Variable & Definition & State 1 & State 2\\
+\hline
+clevel & Current competitive level & Club\_university\_county & National\_international\\
+\hline
+gender & Gender of the participant & Female & Male\\
+\hline
+hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
+\hline
+ind\_team & Participate in an individual or team based sport & Individual & Team\\
+\hline
+pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
+\hline
+nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
+\hline
+FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
+\hline
+BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
+\hline
+RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
+\hline
+stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
+\hline
+HRV & Root mean squared difference of successive RR intervals & 2.03-4 (Low) & >4-5.94 (High)\\
+\hline
+balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
+\hline
+nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+\end{tabular}
+\end{table}
+
+Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
+For example, the probability that stiffness\_1 and stiffness\_2 were both ``High'', or both ``Low'' was approximately 80\%.
+Including the arcs between the same variables from X\_1 → X\_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a ``Low'' or ``High'' state for each pair of variables in the network.
+Therefore, these arcs were blacklisted from the network.
+To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network.
+The final network structure was obtained by averaging 1000 bootstrapped models (Efron and Tibshirani, 1994) to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model (Olmedilla et al., 2018).
+Arcs that were present in at least 30\% of the models were included in the averaged model.
+The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
+An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
+In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (Scutari and Nagarajan, 2013).
+Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+
+\hypertarget{network-inference}{%
+\paragraph{Network Inference}\label{network-inference}}
+
+Conditional probability queries (CPQ) were used to perform inference on both network structures.
+To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
+The query allows the odds of a particular node state (e.g., injured\_1 = ``injured'') to be calculated based on the new evidence.
+CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
+The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query (Scutari and Denis, 2014).
+Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
+Of particular interest were the variables that were connected to ``injured'' nodes, which were examined in the network using the Markov blanket of ``injured\_1'' and ``injured\_2''.
+A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network (Fuster-Parra et al., 2017).
+CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the ``injured'' state.
+
+\hypertarget{second-network}{%
+\subsubsection{Second network}\label{second-network}}
+
+\hypertarget{data-preparation-1}{%
+\paragraph{Data preparation}\label{data-preparation-1}}
+
+Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
+
+For the second network, change scores for continuous variables between T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3 were standardized to allow relative changes between variables to be compared.
+The ``injured'' variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
+Setting the data up in this way enabled the construction of a network that explicitly modelled the \emph{amount} of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from ``Low'' to ``High''.
+
+\hypertarget{network-structure-1}{%
+\paragraph{Network structure}\label{network-structure-1}}
+
+Similar to the first network, blacklists were used to prevent arcs from independent variables \(\rightarrow\) explanatory variables.
+In addition, ``nlebase'' variable was dropped from the list of explanatory variables to allow the \emph{changes} in negative life events to be the only life event variable in the network.
+The final network was obtained using the same approach as the first network.
+
+\hypertarget{network-inference-1}{%
+\paragraph{Network Inference}\label{network-inference-1}}
+
+Conditional probability queries (CPQ) were again used to perform inference the network structures.
+The Markov blanket of the ``injured'' node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
+Initial visual inspection of the network structure also revealed arcs between BIS \(\rightarrow\) FFFS and HRV \(\rightarrow\) FFFS.
+To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
+The samples were then examined using a Bayesian linear regression models with the \emph{brms} package (Bürkner, 2017) to determine the relationship between these nodes.
+Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+During the study, 26\% of participants reported at least one injury with an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days).
+Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39 {[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and non-contact injuries were more common than contact injuries (male, non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female, non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}).
+Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
+
+\begin{table}[H]
+
+\caption{\label{tab:table4}The number and percentage (\%) of types of injuries sustained by male and female participants.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l}
+\hline
+\multicolumn{1}{c|}{\textbf{ }} & \multicolumn{2}{c|}{\textbf{Female}} & \multicolumn{2}{c}{\textbf{Male}} \\
+\cline{2-3} \cline{4-5}
+\textbf{} & \textbf{Lower body} & \textbf{Upper body} & \textbf{Lower body} & \textbf{Upper body}\\
+\hline
+Joint / ligament & 14 (36) & 5 (36) & 37 (43) & 14 (38)\\
+\hline
+Muscle / tendon & 17 (44) & 6 (43) & 45 (52) & 12 (32)\\
+\hline
+Other & 8 (21) & 3 (21) & 5 (6) & 11 (30)\\
+\hline
+\multicolumn{5}{l}{\textit{Note: }}\\
+\multicolumn{5}{l}{Other included bone, skin and brain injuries.}\\
+\end{tabular}
+\end{table}
+
+\hypertarget{first-network-structure}{%
+\subsection{First network structure}\label{first-network-structure}}
+
+The first network structure obtained from the data is shown in Fig \ref{fig:fig6}.
+Several of the explanatory variables showed strong connections with independent variables in the network.
+The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.23 to 0.47 when nlebase increased from ``Low'' to ``High''.
+The ind\_team node had strong arcs to hours (0.90) and nlebase (0.84).
+Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
+Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = 0.65, team-based athletes = 0.41).
+The arcs from gender \(\rightarrow\) stiffness\_1 and gender \(\rightarrow\) stiffness\_2 were 0.76, and 0.65 respectively, with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
+The arc from pi \(\rightarrow\) stiffness\_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have ``High'' (0.65) compared to ``Low'' (0.35) stiffness.
+The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
+Arcs were also present between independent variables in the network. Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.68).
+In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.65 for \_1, 0.61 for \_2) compared to ``Low'' (0.32 for \_1, 0.36 for \_2).
+The arc between nlelg \(\rightarrow\) BIS had a strength of 0.62 for nlelg\_1 \(\rightarrow\) BIS\_1, however no arc was present between nlelg\_2 and BIS\_2.
+For nlelg\_1 \(\rightarrow\) BIS\_1, ``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55.
+
+\hypertarget{markov-blanket-for-injured_1}{%
+\subsubsection{Markov blanket for injured\_1}\label{markov-blanket-for-injured_1}}
+
+The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1), is shown in Fig \ref{fig:fig7}.
+The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
+
+The conditional probability query (CPQ) for injured\_1 in the ``injured'' state for all variables that were directly linked to injured\_1 is shown in Table \ref{tab:table5}.
+The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when stiffness was ``High'' compared to ``Low''.
+Negative life events had a negligible effect when moving from the ``Low'' to ``High'' state.
+
+\begin{table}[H]
+
+\caption{\label{tab:table5}Probability of injured\_1 being in the ``injured'' state, conditional on each variable.}
+\centering
+\begin{tabular}[t]{l|r|r}
+\hline
+\textbf{Variable} & \textbf{Low} & \textbf{High}\\
+\hline
+balance\_1 & 0.21 & 0.30\\
+\hline
+hours & 0.18 & 0.28\\
+\hline
+nlelg\_1 & 0.24 & 0.26\\
+\hline
+stiffness\_1 & 0.17 & 0.31\\
+\hline
+\end{tabular}
+\end{table}
+
+The second CPQ investigated the outcome of injured\_1 being ``injured'' conditional on all variables in the Markov blanket.
+The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
+The combination of lower competitive level, ``High'' hours per week, ``Low'' negative life events, ``High'' balance and ``High'' stiffness resulted in a probability of 0.53 for injured\_1 being in the ``injured'' state.
+When all variables were in the ``Low'' state the probability of ``injured'' was approximately 0.04
+Negative life events only had a substantial effect on injured\_1 when all other variable were fixed to ``Low''.
+In this instance the probability of injured\_1 being ``injured'' rose from 0.04 to 0.19, when negative life events was in the ``Low'' and ``High'' states respectively.
+
+\begin{table}[H]
+
+\caption{\label{tab:table6}Highest and lowest probability of injured\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_1.}
+\centering
+\begin{tabular}[t]{r|l|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{nlelg\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
+\hline
+\multicolumn{6}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.53 & club\_university\_county & High & Low & High & High\\
+\hline
+\hspace{1em}0.46 & national\_international & High & Low & High & Low\\
+\hline
+\hspace{1em}0.43 & national\_international & High & Low & High & High\\
+\hline
+\multicolumn{6}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.06 & national\_international & Low & Low & Low & Low\\
+\hline
+\hspace{1em}0.05 & national\_international & Low & Low & Low & High\\
+\hline
+\hspace{1em}0.04 & club\_university\_county & Low & Low & Low & Low\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{markov-blanket-of-injured_2}{%
+\subsubsection{Markov blanket of injured\_2}\label{markov-blanket-of-injured_2}}
+
+The Markov blanket for injured\_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and rmssd\_2.
+The arc between stiffness\_2 \(\rightarrow\) injured\_2 was comparable to the arc between stiffness\_1 \(\rightarrow\) injured\_1.
+Very weak arcs (0.3) between injured\_2 \(\rightarrow\) balance\_2 and injured\_2 \(\rightarrow\) rmssd\_2 were also present in the Markov blanket for injured\_2.
+Results of the first query for injured\_2 in the ``injured'' state are presented in table \ref{tab:table8}.
+Similar to injured\_1, stiffness\_2 doubled the probability of injured\_2 being ``injured'' from 0.13 in the ``Low'' state to 0.27 in the ``High'' state.
+FFFS\_1 in the ``Low'' state increased probability of injured\_2 being ``injured'' by 0.19 compared to the ``High'' state.
+``High'' negative life events decreased the probability of injury from 0.24 to 0.19.
+
+\begin{table}[H]
+
+\caption{\label{tab:table8}Probability of injured\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\_2.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+\textbf{Variable} & \textbf{Low} & \textbf{High}\\
+\hline
+balance\_2 & 0.17 & 0.27\\
+\hline
+FFFS\_1 & 0.3 & 0.11\\
+\hline
+nlelg\_2 & 0.24 & 0.19\\
+\hline
+rmssd\_2 & 0.24 & 0.19\\
+\hline
+stiffness\_2 & 0.13 & 0.27\\
+\hline
+\end{tabular}
+\end{table}
+
+The conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table9}.
+Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
+The combination ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.52.
+With all other variables held in the ``High'' state, the probability of injured\_2 being ``injured'' rose from 0.14 to 0.34 when FFFS\_1 was in the ``Low'' compared to ``High'' state.
+The combination of ``Low'' stiffness, ``Low'' balance and ``High'' FFFS resulted in the lowest probability of injured\_2 being ``injured''.
+
+\begin{table}[H]
+
+\caption{\label{tab:table9}Highest and lowest probability of injured\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_2.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{FFFS\_1} & \textbf{nlelg\_2} & \textbf{stiffness\_2} & \textbf{rmssd\_2} & \textbf{balance\_2}\\
+\hline
+\multicolumn{6}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.52 & Low & Low & High & Low & High\\
+\hline
+\hspace{1em}0.45 & Low & High & High & Low & High\\
+\hline
+\hspace{1em}0.43 & Low & Low & High & High & High\\
+\hline
+\multicolumn{6}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.06 & High & High & Low & Low & Low\\
+\hline
+\hspace{1em}0.05 & High & Low & Low & High & Low\\
+\hline
+\hspace{1em}0.04 & High & High & Low & High & Low\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{second-network-structure---changes-within-variables}{%
+\subsection{Second network structure - changes within variables}\label{second-network-structure---changes-within-variables}}
+
+The network for changes within variables is presented in Fig \ref{fig:fig9}.
+An arc between BIS \(\rightarrow\) FFFS with strength 1 was present in the network.
+Arcs between clevel \(\rightarrow\) BIS and gender \(\rightarrow\) stiffness had a strength of 0.60.
+The arc between RMSSD \(\rightarrow\) FFFS was 0.56.
+The arcs between BIS \(\rightarrow\) FFFS and RMSSD \(\rightarrow\) FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
+The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
+
+Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95\% credible intervals (CrI).
+Increases in BIS were associated with increases in FFFS (\(b\) = -0.14, 95\% CrI = {[}-0.20, -0.08{]}), whereas positive changes in RMSSD where associated with decreased changes in FFFS (\(b\) = 0.4, 95\% CrI = {[}0.34, 0.45{]}).
+There was no clear interaction between RMSSD and BIS (\(b\) = -0.04, 95\% CrI = {[}-0.10, 0.02{]}).
+
+\begin{table}[H]
+
+\caption{\label{tab:table10}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.}
+\centering
+\begin{tabular}[t]{l|r|r|l}
+\hline
+\textbf{Term} & \textbf{Estimate} & \textbf{Error} & \textbf{95\% CI}\\
+\hline
+Intercept & 0.00 & 0.03 & [-0.06, 0.05]\\
+\hline
+BIS & 0.40 & 0.03 & [0.34, 0.45]\\
+\hline
+rmssd & -0.14 & 0.03 & [-0.20, -0.08]\\
+\hline
+BIS:rmssd & -0.04 & 0.03 & [-0.10, 0.02]\\
+\hline
+\end{tabular}
+\end{table}
+
+The Markov blanket for the ``injured'' node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
+For stiffness and nlec the values in the nodes represent the standardised change between time point.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
+Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of 0.71 over the duration of the study.
+With stiffness held at the mean change, the probability of ``injured'' rose from 0.35 to 0.64 with nlec at 1 SD below an 1 SD above respectively.
+
+\begin{table}[H]
+
+\caption{\label{tab:table11}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
+\centering
+\begin{tabular}[t]{r|l|l}
+\hline
+\textbf{Probability} & \textbf{nlec} & \textbf{Stiffness}\\
+\hline
+0.71 & +1SD & +1SD\\
+\hline
+0.64 & +1SD & mean\\
+\hline
+0.62 & +1SD & -1SD\\
+\hline
+0.52 & mean & +1SD\\
+\hline
+0.44 & mean & mean\\
+\hline
+0.43 & mean & -1SD\\
+\hline
+0.42 & -1SD & +1SD\\
+\hline
+0.35 & -1SD & mean\\
+\hline
+0.35 & -1SD & -1SD\\
+\hline
+\end{tabular}
+\end{table}
+
+Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
+In contrast, below average changes in nlec and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
+
+\begin{table}[H]
+
+\caption{\label{tab:table12}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+\centering
+\begin{tabular}[t]{r|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
+\hline
+\multicolumn{5}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.77 & High & injury & +1SD & +1SD\\
+\hline
+\hspace{1em}0.74 & High & no injury & +1SD & +1SD\\
+\hline
+\hspace{1em}0.72 & Low & injury & +1SD & +1SD\\
+\hline
+\multicolumn{5}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.15 & Low & no injury & -1SD & +1SD\\
+\hline
+\hspace{1em}0.13 & Low & no injury & -1SD & mean\\
+\hline
+\hspace{1em}0.11 & Low & no injury & -1SD & -1SD\\
+\hline
+\end{tabular}
+\end{table}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{summary-fo-results}{%
+\subsection{Summary fo results}\label{summary-fo-results}}
+
+(not sure where this would go /if needed at all?).
+The first BN revealed both high stiffness and poor balance resulted in an increased probability of injury.
+However, contrary to previous literature, negative life event stress had a negligible effect on the probability of injury.
+In the second network that modelled changes between time points in the study, increases in muscle stiffness was found to increase the risk of injury.
+In addition, the probability of injury was further increased when there were increases in NLE stress.
+
+(Main discussion starts here..)
+
+Informed by Appaneal and Perna's (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature (Johnson et al., 2014).
+Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables \emph{across} time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables \emph{between} time points (Fig \ref{fig:fig8}).
+
+The first network revealed several links between the injured nodes and other variables in the network.
+For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured\_1 and injured\_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
+The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for ``injured'' nodes.
+The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury.
+In the second network, the highest probability of injury was observed when changes in stiffness \emph{and} negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury (Appaneal and Perna, 2014).
+
+Of all the variables measured in the study, muscle stiffness appeared to be most strongly related to injury.
+Both ``High'' levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
+In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
+To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players (Pickering-Rodriguez et al., 2017).
+The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
+However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness (Pruyn et al., 2015; Kalkhoven and Watsford, 2017).
+Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury (Butler et al., 2003).
+Additionally, high stiffness may only increase the risk of injury if other factors are also present.
+To elaborate, the combination of high stiffness \emph{and} poor balance was found to result in the greatest probability of injury.
+In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
+Several studies have identified how balance (Romero-Franco et al., 2014; Trojian and McKeag, 2006) and muscle stiffness (Butler et al., 2003; Pickering-Rodriguez et al., 2017) are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
+
+In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance \(\rightarrow\) injured\_1 and balance \(\rightarrow\) injured\_2 respectively.
+Despite the weak arc strength, a ``High'' balance score, indicating impaired postural stability, was found to increase the probability of injury.
+This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (Riemann et al., 1999; Romero-Franco et al., 2014; Trojian and McKeag, 2006).
+Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (Paillard et al., 2006).
+In the current study, athletes who competed at a higher level were also more likely to have good balance (``Low'' balance), compared to their lower level counterparts.
+These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (Hrysomallis, 2011).
+
+Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury (Williams and Andersen, 2007; Ivarsson et al., 2017).
+In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
+However, negative life event stress had almost no effect on the probability of injury in the first network.
+This finding suggests that the \emph{relative change} in life events may be more important than the \emph{absolute score} for life events, despite the latter being commonly used in sports injury research to date.
+For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400\% increase in their life event score.
+Although the absolute score would be ``Low'', the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect (Appaneal and Perna, 2014).
+Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed (cf.~Ivarsson et al., 2014).
+
+The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a \emph{negative} valence, as the strongest predictor of injury in Williams and Andersen's (Williams and Andersen, 1998) model (Ivarsson et al., 2017).
+During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
+However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg\_1 \(\rightarrow\) injured\_1 and nlelg\_2 \(\rightarrow\) injured\_2 to allow this relationship to be examined more closely.
+When negative life events were ``High'' the probability of injury showed a negligible change at the injured\_1 node and decreased by -0.05 at the injured\_2 node.
+One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
+In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months (Petrie, 1992).
+However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred \emph{since the previous data collection session}, to avoid inflated scores caused by reporting the same event on multiple occasions.
+The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
+For example, at the second and third time points, 26\% of participants reported 0 negative life events for the preceding four-month period.
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events (cf.~Fawkner et al., 1999).
+
+Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
+Williams and Andersen's (Williams and Andersen, 1998) model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
+Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (Petrie, 1993; Petrie et al., 2014; Johnson et al., 2014).
+Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
+
+Of the remaining variables, both FFFS and RMSSD were also linked to injury.
+A weak arc was observed between RMSSD\_2 \(\rightarrow\) injured\_2 (arc strength = 0.30), however no arc was present between RMSSD\_1 and injured\_1, suggesting the link between RMSSD and injury was not as certain as muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
+Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, ``Low'' RMSSD increased the probability of injury from 0.17 (RMSSD = ``High'\,') to 0.27 (RMSSD =''Low'\,').
+This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (Williams et al., 2017; Bellenger et al., 2016; Gisselman et al., 2016).
+An arc between FFFS\_1 and injured\_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the ``High'' and ``Low'' states respectively.
+Interestingly, the ``Low'' FFFS score was also related to injuries at subsequent time points.
+One possible explanation for this finding could be that those athletes who reported ``Low'' FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
+The RST theory proposes that higher levels of FFFS increase avoidance motivation (Corr et al., 2016), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
+
+In the first network, ``High'' BIS was associated with ``High'' FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
+RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (Corr, 2013).
+High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD (Chalmers et al., 2014; Pulopulos et al., 2018), which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
+These findings align with the proposed actions of the RST theory (Corr et al., 2016).
+For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems.
+The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (Corr et al., 2016).
+To extend these findings, the BAS should also be considered.
+Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
+However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
+In the first network, RI\_1 was connected to both BIS\_1 and BIS\_2, and in both instances, the probability of ``High'' BIS was increased when RI was also ``High''.
+However, the arcs between RI\_1 and BIS were weak (\textless{} 0.50), and RI represents only one component of the BAS system.
+Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury.
+Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
+
+The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury (Johnson et al., 2014).
+The longitudinal repeated measures design of the current study allowed \emph{changes} over time and between time points to be captured and explored.
+Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
+Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
+Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (Bittencourt et al., 2016) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
+The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
+Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (Olmedilla et al., 2018).
+
+As with most research, there were several limitations with the present study.
+Firstly, the choice was made to binarise variables in the first network so only ``Low'' and ``High'' states were observed.
+Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner Qian and Miltner (2015) highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
+For example, the meaning of a ``Low'' and ``High'' value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
+A second limitation was the nature of the physiological measures used in the current study.
+In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
+For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis (Ross et al., 2011).
+
+In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
+Although the current study used a range of measures to capture ``stress'' from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
+For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress (Appaneal and Perna, 2014; Perna and McDowell, 1995), and could help elucidate the relationship between stress and injury.
+Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
+For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
+Both responses would be scored a ``-4''; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
+Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
+
+In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
+Specifically, muscle stiffness and \emph{increases} in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
+
+\hypertarget{disclosureconflict-of-interest-statement}{%
+\section{Disclosure/Conflict-of-Interest Statement}\label{disclosureconflict-of-interest-statement}}
+
+The authors declare that the research was conducted in the absence of any
+commercial or financial relationships that could be construed as a potential
+conflict of interest.
+
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+The statement about the authors and contributors can be up to several sentences
+long, describing the tasks of individual authors referred to by their initials
+and should be included at the end of the manuscript before the References
+section.
+
+\hypertarget{supplemental-data}{%
+\section{Supplemental Data}\label{supplemental-data}}
+
+Supplementary Material should be uploaded separately on submission, if there are
+Supplementary Figures, please include the caption in the same file as the
+figure. LaTeX Supplementary Material templates can be found in the Frontiers
+LaTeX folder
+
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
 
@@
 \end{figure}
-
-Williams and Andersen's (Williams and Andersen, 1998) model proposed
-that when faced with a potentially stressful athletic situation, an
-athlete's personality traits (e.g., hardiness, locus of control and
-competitive trait anxiety), history of stressors (e.g., major life
-events and previous injuries) and coping resources (e.g., general coping
-behaviours) will contribute to their response, either interactively or
-in isolation. Central to the model is the stress response, which
-reflects the bi-directional relationship between athletes' appraisal of,
-and response to, a stressful athletic situation. The model predicts that
-athletes who have a history of many stressors, personality traits that
-intensify the stress response and few coping resources, will exhibit
-greater attentional (e.g., peripheral narrowing) and/or physiological
-(e.g., increased muscle tension) responses that put these individuals at
-greater risk of injury.
-
-Within Williams and Andersen's (Williams and Andersen, 1998) model,
-major life events, a component of an athlete's history of stressors,
-most consistently predicts injury occurrence (Williams and Andersen,
-2007); specifically, major life events with a negative, as opposed to
-positive, valence (Passer and Seese, 1983; Maddison and Prapavessis,
-2005). However, personality traits and coping resources have also been
-found to predict injury, with for example, athletes more likely to
-sustain an injury if they have poor social support and psychological
-coping skills, and high trait anxiety and elevated competitive state
-anxiety; compared to athletes with the opposite profile. (Smith et al.,
-1990; Lavallée and Flint, 1996; Ivarsson and Johnson, 2010). However,
-the amount of variance explained by the psychosocial factors proposed by
-the model has been modest, typically between 5 - 30\% (Galambos et al.,
-2005; Ivarsson and Johnson, 2010); suggesting other factors are also
-likely to contribute to injury occurrence.
-
-While the psychosocial factors proposed in Williams and Andersen's
-(Williams and Andersen, 1998) model have received the most research
-attention, the mechanisms through which these factors are proposed to
-exert their effect have remained under-investigated in the literature.
-To elaborate, the model suggests that injuries are likely to occur
-through either increased physiological arousal resulting in increased
-muscle tension and reduced flexibility or attentional deficits caused by
-increased distractibility and peripheral narrowing. However, to date,
-the research has largely focused on attentional deficits (Andersen and
-Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et
-al., 2007). For example, Andersen and Williams (Andersen and Williams,
-1999) measured peripheral and central vision during high and low stress
-conditions and found athletes with high life event stress coupled with
-low social support had greater peripheral narrowing under stressful
-conditions compared to athletes with the opposing profile; these
-athletes went on to sustain an increased number of injuries during the
-following season. Indeed, Rodgers and Landers (Rogers and Landers, 2005)
-supported Andersen and Williams's (Andersen and Williams, 1999) earlier
-findings reporting that peripheral narrowing under stress mediated 8.1\%
-of the relationship between negative life events and injury. However,
-the remaining variance between negative life events and athletic injury
-through the other proposed mechanisms, such as increased muscle tension
-and reduced motor control, remains to be explored (cf.~Williams and
-Andersen, 1998).
-
-One challenge faced by researchers addressing the sports injury problem
-with a psychological lens is the multifactorial nature of injury, and
-the possible contribution of other non-psychological factors to the
-stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009). For
-example, a large body of research indicates that training-related stress
-is also likely to be related to the stress response and injury
-occurrence (Lee et al., 2017; Djaoui et al., 2017), and may account for
-the unexplained variance from the psychological predictors of injury.
-Appaneal and Perna (Appaneal and Perna, 2014) proposed the
-biopsychosocial model of stress athletic injury and health (BMSAIH) to
-serve as an extension to Williams and Andersen's (Williams and Andersen,
-1998) model and to address some of these issues. To elaborate, the
-BMSAIH aimed to clarify the mediating pathways between the stress
-response and injury, consider other health outcomes and behavioural
-factors that impact sports participation, and integrate the impact of
-training on athletes' health (Appaneal and Perna, 2014). The central
-tenet of the BMSAIH is that psychosocial distress (e.g., negative life
-events) may act synergistically with training-related stress as a result
-of high-intensity and high-volume sports training, and ``widen the
-window of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of
-undesirable health outcomes including illness and injury. Consequently,
-the BMSAIH provides a framework for future research to build on Williams
-and Andersen's (Williams and Andersen, 1998) model, by including other
-physiological markers of training-related stress, which together may
-provide greater insight into the injury process.
-
-Although research supporting the BMSAIH has mainly focused on the
-relationship between hormonal responses to training and injury
-occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other
-research has identified additional markers of training-related stress
-that are associated with an increased risk of injury; for example, heart
-rate variability (Bellenger et al., 2016; Williams et al., 2017),
-postural stability (Romero-Franco et al., 2014) and muscle stiffness
-(Pruyn et al., 2015). However, these markers are often studied in
-isolation without an assessment of the psychosocial factors that are
-known to contribute to injury, thereby limiting our understanding of how
-psychosocially and physiologically derived stress may contribute
-synergistically to injury occurrence. Recently, Bittencourt et
-al.~(Bittencourt et al., 2016) suggested that to better understand the
-multifactorial nature of sports injuries, research needs to move away
-from studying risk factors in isolation and instead adopt a complex
-systems approach to injury. Such an approach posits that injury may
-arise from a complex ``web of determinants'' (Bittencourt et al., 2016,
-3), where different factors interact in unpredictable and unplanned
-ways, but result in a global outcome pattern of either adaptation or
-injury.
-
-A challenge when adopting a complex systems approach is using an
-appropriate analysis technique that is able to capture the uncertainty
-and complexity of the relationships between different variables. One
-technique that provides a solution to this challenge is Bayesian network
-(BN) modelling. BN's allow the construction of graphical probabilistic
-models using the underlying structure that connects different variables
-(nodes in the network) (Scutari and Denis, 2014). The learned structure
-can be used for inference by obtaining the posterior probabilities of a
-particular node for a given query (e.g., if the value of Node A is x and
-the value of Node B is y, what is the probability Node C being value
-z?). Furthermore, BN's do not distinguish between dependent and
-independent variables as they are a form of unsupervised learning, which
-is a strength over regression or structure equation models when the
-underlying relationship in the network may not be known (Olmedilla et
-al., 2018).
-
-To summarise, despite offering a possible framework to build on the
-research stemming from Williams and Andersen's (Williams and Andersen,
-1998) model, there remains and opportunity to explore other
-physiological stress-related markers proposed by the BMSAIH, in addition
-to the already well-established psychological characteristics known to
-be related to injury (Appaneal and Perna, 2014). Furthermore, research
-has typically not captured changes in both psychosocial factors and
-stress-related physiological markers that may occur between initial
-measurement and injury occurrence. Given the exploratory nature of the
-current study the following objectives were defined:
-
-\begin{itemize}
-\tightlist
-\item
-  Identify suitable markers of stress that can be easily captured in a
-  large cohort of athletes in a field based setting.
-\item
-  Capture the markers of stress and injury occurrence using a
-  prospective, repeated measures design.
-\item
-  Explore and evaluate the relationships between the markers of stress
-  and injury using Bayesian network modelling.
-\end{itemize}
-
-\hypertarget{methods}{%
-\section*{Methods}\label{methods}}
-\addcontentsline{toc}{section}{Methods}
-
-\hypertarget{participants}{%
-\subsection*{Participants}\label{participants}}
-\addcontentsline{toc}{subsection}{Participants}
-
-A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120)
-were initially recruited for the study. Participants had an average age
-of 22.0 \(\pm\) 7.0 years and represented a range of team (football,
-rugby, netball, cricket, lacrosse, basketball and field hockey) and
-individual sports (athletics, tennis, weightlifting, gymnastics, judo,
-swimming and golf from a British University and local sports clubs
-(Table \ref{tab:table1}). Participants self-rated competitive level
-ranged from recreational to international standard. A total of 126
-(49.03\%) participants had sustained an injury in the 12 months prior to
-the start of the study (male: \emph{n} = 114 {[}49\%{]}, female;
-\emph{n} = 48 {[}40\%{]}). At the start of the study, all participants
-were injury free (no modifications to their usual training routine due
-to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at
-least five hours per week. Ethical approval was obtained from the
-University ethics committee prior to the start of the study and all
-participants provided informed consent.
-
-\begin{table}[H]
-
-\caption{\label{tab:table1}Participant characteristics.}
-\centering
-\begin{tabular}[t]{c|c|c}
-\hline
-\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
-\hline
-\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
-\hline
-\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
-\hline
-\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
-\hline
-\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
-\hline
-\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
-\hline
-\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
-\hline
-\hspace{1em}Recreational & 3 (4) & 7 (4)\\
-\hline
-\hspace{1em}University & 45 (56) & 141 (80)\\
-\hline
-\hspace{1em}National/International & 33 (41) & 28 (16)\\
-\hline
-\end{tabular}
-\end{table}
-
-\hypertarget{measures}{%
-\subsection*{Measures}\label{measures}}
-\addcontentsline{toc}{subsection}{Measures}
-
-\hypertarget{major-life-events}{%
-\subsubsection*{Major life events}\label{major-life-events}}
-\addcontentsline{toc}{subsubsection}{Major life events}
-
-A modified version of the Life Events Survey for Collegiate Athletes
-(LESCA) was used to measure participants' history of life event stress
-(Petrie, 1992). Modifications were made to the LESCA to ensure the
-suitability of the items for the study population (S1 Table). The LESCA
-comprises 69 items that reflect possible life events that participants
-may have experienced. Example items include, ``Major change in the
-frequency (increased or decreased) of social activities due to
-participation in sport'', ``Major change in the amount (more or less) of
-academic activity (homework, class time, etc)'' and ``Major change in
-level of athletic performance in actual competition (better or worse)''.
-Participants were asked to rate the perceived impact of each life event
-they had experienced within the last 12 months on an 8-point Likert
-scale anchored at -4 (\(extremely\;negative\)) and +4
-(\(extremely\;positive\)). Negative and positive life event scores were
-calculated by summing the negative and positive scores respectively. A
-score for total life events was calculated by summing the absolute
-values for both negative and positive events. Petrie (1992) (Petrie,
-1992) reported test-retest reliabilities at 1-week and 8-weeks with
-values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p
-\textless{} .001) respectively. Petrie also provided evidence of
-discriminant, convergent and predictive validity. The LESCA is the most
-widely used measure of major life events for athletes in the sports
-injury literature. For this study, Composite Reliability (Fornell and
-Larcker, 1981) was preferred to Cronbach's alpha as it does not assume
-parallelity (i.e., all factor loadings are constrained to be equal, and
-all error variances are constrained to be equal) and instead takes into
-consideration the varying factor loadings of the items in the
-questionnaire. The composite reliability for the LESCA in this study was
-0.84.
-
-\hypertarget{reinforcement-sensitivity-theory-personality-questionnaire}{%
-\subsubsection*{Reinforcement Sensitivity Theory Personality
-Questionnaire}\label{reinforcement-sensitivity-theory-personality-questionnaire}}
-\addcontentsline{toc}{subsubsection}{Reinforcement Sensitivity Theory
-Personality Questionnaire}
-
-A revised version of the Reinforcement Sensitivity Theory Personality
-Questionnaire (RST-PQ) was used to measure motivation, emotion,
-personality and their relevance to psychopathology (Corr and Cooper,
-2016). The revised version of the RST-PQ comprises 51 statements that
-measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I
-am the sort of person who easily freezes-up when scared''), Behavioural
-Inhibition System (BIS; e.g., ``When trying to make a decision, I find
-myself constantly chewing it over'') and four Behavioural Approach
-System (BAS) factors; Reward Interest (e.g., ``I regularly try new
-activities just to see if I enjoy them''), Goal Drive Persistence (e.g.,
-``I am very persistent in achieving my goals''), Reward Reactivity
-(e.g., ``I get a special thrill when I am praised for something I've
-done well'') and Impulsivity (e.g., ``I find myself doing things on the
-spur of the moment''). Participants rated each item on a scale from 1
-(\(not\; at\; all\)) to 4 (\(highly\)) to reflect how well each
-statement described their personality in general. The responses to items
-associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed
-to give a total score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = 0.92, FFFS =
-0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details
-regarding the revised RST are in S1 Appendix.
-
-\hypertarget{heart-rate-variability}{%
-\subsubsection*{Heart rate variability}\label{heart-rate-variability}}
-\addcontentsline{toc}{subsubsection}{Heart rate variability}
-
-A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap
-(Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00)
-with ECG recordings (Giles et al., 2016), which are considered the gold
-standard for assessing HRV. In addition, HRV indices calculated from IBI
-and ECG data have shown a strong correlation (r = .99) in athletes
-(Caminal et al., 2018) and under spontaneous breathing conditions (Plews
-et al., 2017).
-
-\hypertarget{musculoskeletal-properties}{%
-\subsubsection*{Musculoskeletal
-properties}\label{musculoskeletal-properties}}
-\addcontentsline{toc}{subsubsection}{Musculoskeletal properties}
-
-A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to
-measure muscle stiffness. The MyotonPRO is a non-invasive, handheld
-device that applies a mechanical impulse of 0.40 N for 0.15 ms
-perpendicular to the surface of the skin. The impulse causes natural
-damped oscillations in the tissue, which are recorded by a three-axis
-digital accelerometer sensor in the device. The raw oscillation signal
-is then processed, and the stiffness parameter is calculated
-(Agyapong-Badu et al., 2016). The MyotonPRO has previously been reported
-to be a reliable and valid tool for the measurement of in-vivo tissue
-stiffness properties (Chuang et al., 2013; Pruyn et al., 2016; Nair et
-al., 2014), and has demonstrated good internal consistency (coefficient
-of variation \textless{} 1.4\%) over sets of 10 repetitions (Aird et
-al., 2012).
-
-\hypertarget{postural-stability}{%
-\subsubsection*{Postural stability}\label{postural-stability}}
-\addcontentsline{toc}{subsubsection}{Postural stability}
-
-Postural stability was assessed with a modified version of the balance
-error scoring system (mBESS) based on the protocol recommended by Hunt
-et al. (2009). In total, each trial of the mBESS was performed without
-shoes (McCrory et al., 2013) and included six stances in the following
-order; dominant leg (DL; standing on the dominant foot with the
-non-dominant foot at approximately 30-degrees of hip flexion and
-45-degrees of knee flexion), non-dominant leg (NDL; standing on the
-non-dominant foot with the dominant foot at approximately 30-degrees of
-hip flexion and 45-degrees of knee flexion) and tandem leg stance (TS;
-standing heel-to-toe with the non-dominant foot behind the dominant) on
-firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively
-(Fig \ref{fig:fig2}). To determine leg dominance, participants were
-asked their preferred leg to kick a ball to a target, and the chosen
-limb was labelled as dominant (cf.~Cingel et al., 2017). Participants
-were asked to maintain each stance for a total of 20-seconds.
-Participants hands were placed on hips at the level of the iliac crests.
-A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United
-Kingdom) was used to record each participants performance during the
-mBESS.
 
@@
 
-The error identification criteria from the original BESS protocol was
-used by the lead researcher who scored all the BESS trials. One error
-was recorded if any of the following movements were observed during each
-trial: a) lifting hands off iliac crests; b) opening eyes; c) stepping,
-stumbling, or falling; d) moving the thigh into more than 30 degrees of
-flexion or abduction; e) lifting the forefoot or heel; and f) remaining
-out of the testing position for more than 5-seconds (Riemann et al.,
-1999). A maximum score of 10 errors was possible for each stance.
-Multiple errors occurring simultaneously were recorded as one error. A
-participant was given the maximum score of 10 if they remained out of
-the stance position for more than 5-seconds. To calculate limb
-asymmetry, the DL and NDL leg score was calculated by summing the DL and
-NDL errors respectively. A total score was calculated by summing the
-total number of errors recorded on all stances (DL, NLD and TS, on foam
-and firm surfaces). To assess the intra-rater reliability, a single
-measurement, absolute agreement, two-way mixed effects model for the
-intraclass correlation (ICC; Koo and Li, 2016) was used on a sample of
-40 participants from the first time point. The test-retest scoring of
-BESS resulted in a ``good'' to ``excellent'' ICC score (ICC = 0.93, 95\%
-confidence interval = 0.88 - 0.96), indicating the scoring was reliable.
-
-\hypertarget{injury}{%
-\subsubsection*{Injury}\label{injury}}
-\addcontentsline{toc}{subsubsection}{Injury}
-
-Participants self-reported any injuries they sustained at each data
-collection during the study period. An injury was defined as any sports
-related medical problem causing the athlete to miss or modify their
-usual training routine for at least 24 hours (Fuller et al., 2006, 2007;
-Timpka et al., 2014). Minor scrapes and bruises that may require certain
-modifications (e.g., strapping or taping) but did not limit continued
-participation were not considered injuries (cf.~Appaneal et al., 2009).
-Injury status (did / did not sustain an injury) served as the main
-outcome measure.
-
-\hypertarget{procedure}{%
-\subsection*{Procedure}\label{procedure}}
-\addcontentsline{toc}{subsection}{Procedure}
-
-At the start of the UK academic year (September 2016 and 2017), coaches
-of sports teams at a British University and local sports clubs were
-contacted and informed about the study. With the coaches' permission,
-the lead researcher attended training sessions to inform the athletes
-about the overall purpose of the study and the requirements of
-participation. To be eligible for the study athletes had to be injury
-free (no modifications to their usual training routine due to a sport
-related medical problem for a minimum of four weeks) and training a
-minimum of five hours per week. Athletes who met the criteria and
-volunteered to take part in the study were invited to attend scheduled
-testing sessions. A repeated measures prospective cohort design was used
-to assess athletes' major life events, stress-related physiological
-markers and injury status over two consecutive twelve-month periods
-between September 2016 and September 2018. Each participant was asked to
-attend a total of 3 data collections over a twelve-month period, with
-each data collection separated by a four-month interval (Fig
-\ref{fig:fig3}). Participants provided informed consent before data
-collection commenced.
-
 \begin{figure}[!h]
@@
 
-For the first three data collections (T1, T2 and T3), participants
-followed the same protocol in a specific order (Fig \ref{fig:fig4}). To
-ensure all measures could be collected within a viable time-frame,
-participants were separated into two groups. The first group completed
-all computer-based measures followed by all physical measurements,
-whereas the second group completed all physical measurements followed by
-computer-based measures. Participants were randomly assigned to one of
-the two groups and remained in those groups across all time points.
-
 \begin{figure}[!h]
@@
 
-\hypertarget{questionnaires}{%
-\subsubsection*{Questionnaires}\label{questionnaires}}
-\addcontentsline{toc}{subsubsection}{Questionnaires}
-
-The questionnaires, which included demographic information, the LESCA,
-RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed
-on-line (SurveyMonkey Inc., USA, www.surveymonkey.com). The instructions
-for the LESCA were modified at T2 and T3 so that participants reported
-major life events that had occurred since the previous testing session.
-For injury reporting, participants were asked to record any injuries
-that they had sustained since the last data collection. The data were
-downloaded from surverymonkey.com and imported into R (R Core Team,
-2019) for analysis purposes.
-
-\hypertarget{hrv}{%
-\subsubsection*{HRV}\label{hrv}}
-\addcontentsline{toc}{subsubsection}{HRV}
-
-To minimise potential distractions, participants were directed to a
-designated quiet area in the laboratory where IBI data were recorded.
-Participants were instructed to turn off their mobile devices to avoid
-any interference with the Bluetooth sensor. Each chest strap was
-dampened with water and adjusted so it fitted tightly but comfortably,
-as outlined by Polar's guidelines. Participants were seated and asked to
-remain as still as possible for the duration of the recording. No
-attempt was made to control participants respiratory frequency or tidal
-volume (Denver et al., 2007). Inter-beat interval (IBI) data was
-collected for 10-minutes at a sampling frequency of 1000 Hz.
-
-Raw, unfiltered IBI recordings were exported from the Polar Flow web
-service as a space delimited .txt file and imported into R (R Core Team,
-2019) where the \emph{RHRV} package ({\textbf{???}}) was used to
-calculate HRV indices. Raw IBI data was filtered using an adaptive
-threshold filter, and the first 3-minutes and last 2-minutes of each
-recording were discarded, leaving a 5-minutes window that was used to
-calculate the root mean square of successive differences (RMSSD) in RR
-intervals following the recommendations for short term IBI recordings
-(Laborde et al., 2017; Malik et al., 1996). RMSSD was calculated as:
-
-\begin{equation} 
-  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
-\end{equation}
-
-Where N is the length of the time series, and \(RR_i\) the RR interval
-between beats \(i\) and \(i-1\), where each beat position corresponds to
-the beat detection instant.
-
-\hypertarget{muscle-stiffness}{%
-\subsubsection*{Muscle stiffness}\label{muscle-stiffness}}
-\addcontentsline{toc}{subsubsection}{Muscle stiffness}
-
-To assess muscle stiffness, participants lay horizontally on a massage
-bed and four testing sites were identified on each lower limb. The
-muscle belly of the rectus femoris (RF), biceps femoris (BF), medial
-gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified
-using a visual-palpatory technique to determine the exact location of
-each site (Chuang et al., 2012). The visual-palpatory technique required
-the participant to contract the target muscle to aid the lead researcher
-to visually identify the muscle. The participant was then asked to relax
-the muscle and the muscle was palpated to locate the muscle belly. A
-skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was
-used to mark the testing site in the centre of the muscle belly.
-
-After the eight testing sites had been identified, the testing end of
-the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin
-on the testing site. A constant pre-load of 0.18 N was applied for
-initial compression of subcutaneous tissues. The device was programmed
-to deliver five consecutive impulses, separated by a one second interval
-(Morgan et al., 2018). For each impulse, the device computed stiffness
-values, with the median of the five values being saved by the device for
-further analysis. In accordance with Myoton.com, a set of five
-measurements with a coefficient of variation (CV) of less than 3\% was
-accepted. Sets of measurements above 3 \% were measured again to ensure
-the reliability of the data. The CV was calculated in real time by the
-device after each set of measurements. Measurements saved on the device
-were uploaded to a computer using MyotonPRO software and imported in R
-(R Core Team, 2019) for further analysis. For each participant, the sum
-of all eight testing sites was calculated to provide a total lower
-extremity stiffness score and was used for further analysis.
-
-\hypertarget{postural-stability-1}{%
-\subsubsection*{Postural stability}\label{postural-stability-1}}
-\addcontentsline{toc}{subsubsection}{Postural stability}
-
-Instructions for the mBESS were read to each participant and a
-demonstration of the positions was provided by the research assistant.
-For each position, participants were instructed to close their eyes,
-rest their hands on their iliac crests and remain as still as possible
-for 20-seconds. Participants were instructed to return to the testing
-position as quickly as possible if they lost their balance. The video
-recording was started prior to the first stance position and stopped
-after all stances had been completed. Each completed mBESS protocol took
-approximately 4 minutes. Only one trial was performed to avoid
-familiarisation effects across the repeated measurement (cf.~Valovich et
-al., 2003). The video recordings for each participant were imported from
-the recording equipment (Sony DSC-RX10) and the lead researcher scored
-each trial using the error identification criteria.
-
-\hypertarget{data-analysis}{%
-\subsection*{Data Analysis}\label{data-analysis}}
-\addcontentsline{toc}{subsection}{Data Analysis}
-
-Two Bayesian Networks (BN) were used to explore the relationships
-between the psychological measures, physiological markers of stress and
-sports injury. A BN is a graphical representation of a joint probability
-distribution among a set of random variables, and provides a statistical
-model describing the dependencies and conditional independences from
-empirical data in a visually appealing way (Scutari and Denis, 2014). A
-BN consists of arcs and nodes that together are formally known as a
-directed acyclic graph (DAG), where a node is termed a parent of a child
-if there is an arc directed from the former to the latter (Fig
-\ref{fig:fig5}; Pearl, 1988). However, the direction of the arc does not
-necessarily imply causation, and the relationship between variables are
-often described as probabilistic instead of casual (Scutari and Denis,
-2014). The information within a node can be either continuous or
-discrete, and a complete network can contain both continuous and
-discrete nodes; however, discrete networks are the most commonly used
-form of BN (Chen and Pollino, 2012). In discrete networks, conditional
-probabilities for each child node are allocated for each combination of
-the possible states in their parent nodes and can be used to assess the
-strength of a dependency in the network.
-
 \begin{figure}[H]
@@
 
-In order to use discrete networks, continuous variables must first be
-split into categorical levels. When there are a large number of
-variables in the network, limiting the number of levels has the benefit
-of producing a network that is more parsimonious in terms of parameters.
-For example, a network with 10 variables each with two levels has 100
-(10\^{}2) possible parameter combinations, however the same network with
-three levels has 1000 (10\^{}3) possible parameter combinations, the
-latter being significantly more computationally expensive. Using a
-larger number of splits in the data also comes at a cost of reducing the
-statistical power in detecting probabilistic associations, and reduces
-the precision of parameter estimates for the probabilistic associations
-that are detected because it reduces the sample-size-to-parameters ratio
-(Scutari and Denis, 2014).
-
-Learning the structure of the network is an important step in BN
-modelling. The structure of a network can be constructed using expert
-knowledge and/or data-driven algorithm techniques (e.g., search and
-score, such as hill climbing and gradient descent algorithms; Scutari \&
-Denis, 2014). The learned structure can then be used for inference by
-querying the network\footnote{The term ``query'' in relation to Bayesian
-  Networks stems from Pearl's expert systems theory (1988). A query can
-  be submitted to an expert (in this case, the network is the expert) to
-  get an opinion, the expert then updates the querier's beliefs
-  accordingly. Widely used texts on Bayesian Network analysis (Koller
-  and Friedman, 2009) have widely adopted the terminology in favour of
-  that used in traditional statistics.} and obtaining the posterior
-probabilities of a particular node for a given query. The posterior
-distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where
-the learned network \(B\) with structure \(G\) and parameters
-\(\Theta\), are investigated with new evidence \(E\) using the
-information in \(B\) (Scutari and Denis, 2014). In the example network
-presented in Fig \ref{fig:fig5}, new values assigned to each of the
-parent nodes (e.g., both set to ``Low'') could be used to investigate
-what effect the new information has on the state of the child node
-(conditional probability of a particular state of the child node). In a
-more complex network containing many nodes, the outcome of a particular
-node can be assessed conditional on the states of any subset of nodes in
-the network. BNs therefore provide a unique and versatile approach to
-modelling a set of variables to uncover dependency structures within the
-data.
-
-BNs have recently been used in the sport psychology literature
-(Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several
-benefits over traditional statistical analysis. For example, predictions
-can be made about any variable in the network, rather than there being a
-distinction between dependent and independent variables in the data,
-such as in linear regression models that are often used within the sport
-psychology literature (Olmedilla et al., 2018; Bittencourt et al.,
-2016). Furthermore, the structure of a network can be obtained from both
-empirical data \emph{and} prior knowledge about the area of study; the
-latter being particularly useful when there are a large number of
-variables in the network, or only a small number of observations are
-available in the data (Xiao-xuan et al., 2008). In such instances, a
-purely data driven approach to learning the network would be
-time-consuming due to the large parameter space, and inefficient at
-identifying an approximation of the true network structure. Prior
-knowledge about dependencies between variables can therefore be included
-in the network structure, while still allowing a data driven approach
-for unknown dependencies, to improve the overall computation of the
-network structure (Heckerman et al., 1995; Xu et al., 2015). The
-following sections detail the steps taken in the current study to
-firstly prepare the data for the network, and then obtain the structure
-of the network that was used for inference.
-
-\hypertarget{data-preparation}{%
-\subsubsection*{Data Preparation}\label{data-preparation}}
-\addcontentsline{toc}{subsubsection}{Data Preparation}
-
-Of the 351 participants that were initially recruited for the study, 94
-only completed the first time point, and therefore had to be removed
-from the study as no injury information was obtained for these
-participants following the first time point. To prepare the data for the
-BN, missing values in the dataset were first imputed. Out of the 650
-total measurements across all time points in the current study, there
-were 31 (4.64\%) missing muscle stiffness measurements and 70 (10.48\%)
-missing heart rate recordings. The missing data were due to technical
-faults in the data collection equipment and were considered to be
-missing completely at random. A missing rate of 15-20\% has been
-reported to be common in psychological studies, and several techniques
-are available to handle missing values (Enders, 2003; Lang et al.,
-2013). In the current study, the \emph{caret} package (Kuhn et al.,
-2008) was used to impute the missing values. A bagged tree model using
-all of the non-missing data was first generated and then used to predict
-each missing value in the dataset. The bagged tree method is a reliable
-and accurate method for imputing missing values in data and is superior
-to other commonly used methods such a median imputation (Kuhn et al.,
-2008).
-
-A median split technique was used to discretise the data used in the
-network into ``Low'' and ``High'' levels. All variables apart from
-negative and total life events were approximately normally distributed
-(based on visual inspection, see supplementary table \ldots) and
-required no further transformation prior to the median split. For the
-LESCA questionnaire data, a cumulative total of the current, and
-previous time points was calculated at each time point to account for
-the potential continuing effect of the life events experienced by
-athletes over time. Given the limited support for a relationship between
-positive life events and injury (Williams and Andersen, 2007), only
-negative and total life events were included in the network. Cumulative
-negative, and cumulative total life event scores at each time point were
-first log scaled so distributions were approximately normal, and then
-binarised using the median at each time point (nlelg and tlelg
-respectively). In addition to the log scaled cumulative values, an
-untransformed NLE score from the first time point was included as an
-additional variable based on previous literature that indicates this
-variable should have a strong relationship with injury outcome (Ivarsson
-et al., 2017).
-
-\hypertarget{network-structure}{%
-\subsubsection*{Network structure}\label{network-structure}}
-\addcontentsline{toc}{subsubsection}{Network structure}
-
-To obtain the network structure, several steps were taken to ensure that
-both a theoretically realistic network, and a network that was an
-appropriate fit to the collected data, was used for inference. Prior
-knowledge about the network structure was included by providing a list
-of arcs that are always \emph{restricted} from being in the network
-(blacklist), and a list of arcs that are always \emph{included} in the
-network (whitelist). Additionally, there are several scoring functions
-such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet
-equivalent uniform (BDeu) that can be used to compare network structures
-with certain nodes and arcs included or excluded (Scutari and Denis,
-2014). To account for the repeated measures design and to maximise the
-use of the data, pairs of complete cases (e.g., participants who
-completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network
-(2TBN) structure (Murphy, 2002). In the 2TBN, variables measured T2
-could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2)
-and variables measured at T3 could depend on variables measured at T2
-(e.g., T2 \(\rightarrow\) T3). However, arcs were blacklisted between T2
-\(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in
-which data was collected. Variables were separated into two groups;
-``explanatory'', for variables that did not change during the study
-(e.g., gender), or ``independent'', for variables that were measured at
-each time point and could vary during the study. Independent variable
-names were suffixed with \_1 for time point T, and \_2 for time point
-T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
-Formatting the data in this way meant participants who completed T1 and
-T2, but did not complete T3, could still be included in the analysis.
-Table \ref{tab:table2} provides an example of the formatted data.
-Participants 1 and 3 have complete data, and therefore have two rows of
-data each representing variables from T1 \(\rightarrow\) T2 and T2
-\(\rightarrow\) T3, respectively. Participant 2 did not complete the
-final data collection at T3 and therefore only has one row of data
-representing the variables collected at T1 and T2. In addition to the
-blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\)
-T2, the direction of arcs was restricted between independent variables
-and explanatory variables (e.g., independent \(\rightarrow\)
-explanatory); however, arcs were not restricted between explanatory
-\(\rightarrow\) independent variables. Finally, arc direction was
-restricted between specific nodes within the explanatory variables. Arcs
-from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and
-nlebase \(\rightarrow\) ind\_team were included in the blacklist, as
-arcs in these directions did not make logical sense. All subsequent
-models used the same blacklist.
-
-\begin{table}[H]
-
-\caption{\label{tab:table2}Example of the data arrangement used for the network.}
-\centering
-\begin{tabular}[t]{l|l|l|l}
-\hline
-\textbf{Participant} & \textbf{X\_1} & \textbf{ } & \textbf{X\_2}\\
-\hline
-1 & T1 & -> & T2\\
-\hline
-1 & T2 & -> & T3\\
-\hline
-2 & T1 & -> & T2\\
-\hline
-3 & T1 & -> & T2\\
-\hline
-3 & T2 & -> & T3\\
-\hline
-\end{tabular}
-\end{table}
-
-\hypertarget{preliminary-network-structures}{%
-\subsubsection*{Preliminary network
-structures}\label{preliminary-network-structures}}
-\addcontentsline{toc}{subsubsection}{Preliminary network structures}
-
-Prior to the final network structure presented in the results section,
-several structures with different combinations of variables were
-explred. Networks were learned using a Tabu search algorithm (Russell
-and Norvig, 2009) and BIC was used to compare different models. A higher
-BIC value indicates the structure of a DAG is a better fit to the
-observed data (Scutari and Denis, 2014). BIC values for each combination
-of variables of interest are reported as the combination of variables
-with the highest BIC value, followed by the relative scores of the other
-variables in the model.
-
-Initially, both negative life events and total life events were included
-in the network structure, however, the network score was improved when
-only nlelg or tlelg was included (highest BIC value = nleleg, BIC values
-relative to nlelg; tlelg only = -85.13, tlelg and nlelg = -219.14).
-Additionally, despite strong evidence in the literature that both
-negative and total life event stress are related to injury occurrence
-(Williams and Andersen, 2007), network structures learned using the Tabu
-search algorithm failed to identify a relationship between NLE and
-injury or TLE and injury in the data. Given that nlelg provided the
-highest network score, and there is a stronger relationship between
-negative life events and injury in the literature, an arc was
-whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2
-in the final network structure. Total life event score was not included
-in the final structure.
-
-The subscales representing the BAS (RR, RI, GDP and I) showed limited
-connection to other variables in the network. Therefore, several models
-were run with each scale individually to find the scale that resulted in
-the highest BIC value (values are shown relative to the highest value).
-RI provided the highest BIC value, compared to RR (-13.13), GDP (-15.06)
-and I (-15.2). Including all the variables (RR, RI, GDP and I) resulted
-in a significantly lower score -894.37) indicating that including all
-the variables was not beneficial to the model structure and did not
-offset the cost of the additional parameters. Therefore, only RI was
-included in the final structure.
-
-Finally, both total score and asymmetry were included in the initial
-network. However, visual inspection of the network revealed no arcs
-between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
-Therefore, balance asymmetry was removed from the final network
-structure. To summarise, Table \ref{tab:table3} includes the variables
-that were included in the final network structure.
-
-\begin{table}[H]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{\label{tab:table3}Variables included in the final network structure.}
-\centering
-\begin{tabular}{l|l|l|l}
-\hline
-{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
-\hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Sustained previous injury & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
-\hline
-FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
-\hline
-BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
-\hline
-RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
-\hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
-\hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
-\hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-Preliminary network structures also revealed strong dependencies between
-the same variables at subsequent time points. For example, the
-probability that stiffness\_1 and stiffness\_2 were both ``High'', or
-both ``Low'' was approximately 80\%. Including the arcs between the same
-variables from X\_1 \(\rightarrow\) X\_2 did not provide any
-theoretically meaningful information to the network structure as the
-majority of participants would be either be in a ``Low'' or ``High''
-state for each pair of variables in the network. To more appropriately
-assess changes \emph{within} variables over time, a second BN was
-investigated by modelling the differences between variables at different
-time points. The use of differential equations to model changes in
-variables over time is a common procedure in BN analysis when there are
-repeated measurements in the data (Scutari et al., 2017). To obtain the
-structure, variables suffixed with \_1 were subtracted from variables
-suffixed with \_2 to calculate the difference between variables measured
-at time points T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3.
-Independent variables were then standardized to allow relative changes
-between variables to be compared. The ``injured'' variable was also
-modified to represent whether a participant had sustained an injury at
-any point over the duration of the study or were healthy for the
-duration of the study. The result was a network that explicitly modelled
-the \emph{amount} of change within variables between time points, as
-opposed to the first network that would only have captured changes when
-the median threshold was crossed from ``Low'' to ``High''. Identical
-blacklists to the first network were used for arcs between independent
-and explanatory variables. The nlebase variable was also dropped from
-the list of explanatory variables to allow the \emph{changes} in
-negative life events to be the only life event variable in the network.
-
-To obtain the final networks, the appropriate blacklist and whitelists
-were provided and a Tabu search algorithm identified the remaining
-structure of the network. The final network structure was obtained by
-averaging 1000 bootstrapped models (Efron and Tibshirani, 1994) to
-reduce the impact of locally optimal, but globally suboptimal network
-learning, and to obtain a more robust model (Olmedilla et al., 2018).
-Arcs that were present in at least 30\% of the models were included in
-the averaged model. The strength of each arc was determined by the
-percentage of models that the arc was included in, independent of the
-arc's direction. An arc strength of 1 indicates that the arc is always
-present in the network, with the value decreasing as arcs are found in
-fewer networks. In the respective study arcs above 0.5 were considered
-``significant'' with arcs below 0.5 and above 0.3 ``non-significant''
-(Scutari and Nagarajan, 2013). Arcs below 0.3 were not included in the
-model. The full table of arc strengths for the first and second network
-are available in S2 Table and S3 Table respectively.
-
-\hypertarget{network-inference}{%
-\subsubsection*{Network Inference}\label{network-inference}}
-\addcontentsline{toc}{subsubsection}{Network Inference}
-
-Conditional probability queries (CPQ) were used to perform inference on
-both network structures. To conduct a CPQ, the joint probability
-distribution of the nodes was modified to include a new piece of
-evidence. The query allows the odds of a particular node state (e.g.,
-injured\_1 = ``injured'') to be calculated based on the new evidence.
-CPQs were performed using a likelihood weighting approach; a form of
-importance sampling where random observations are generated from the
-probability distribution in such a way that all observations match the
-evidence given in the query. The algorithm then re-weights each
-observation based on the evidence when computing the conditional
-probability for the query (Scutari and Denis, 2014). Inference was first
-performed on arcs that had a strength greater than 0.50 between the
-explanatory variables and independent variables and between different
-independent variables in the network. Of particular interest in the
-current study were the variables that were connected to ``injured''
-nodes. To examine the variables that were associated with injured nodes
-in the network, the Markov blanket of ``injured\_1'' and ``injured\_2''
-were examined. A Markov blanket contains all the nodes that make the
-node of interest conditionally independent from the rest of the network
-(Fuster-Parra et al., 2017). CPQ's were used to determine what effect
-the variables in the Markov blanket of injured nodes had on the
-probability of the injured node being in the ``injured'' state.
-
-The second network contained both continuous and discrete data. To
-examine dependencies between continuous variables with arc strengths
-above 0.5 in the second network, random samples were generated based on
-the conditional distribution of the nodes included as evidence in the
-query. The samples were then extracted and examined with Bayesian linear
-regression models using the \emph{brms} package (Bürkner, 2017) to
-determine the relationship between nodes in the network. Similar to the
-first network, the Markov blanket of the ``injured'' node was also
-investigated by determining the highest probability of injury with
-combinations of variables in the Markov blanket below the mean change,
-at the mean change and above the mean change.
+\begin{figure}[H]
+\caption{{\bf Network structure.}
+The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+\label{fig:fig6}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket of injured\_1.}
+Arc strengths are included as arc labels.}
+\label{fig:fig7}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for injured\_2.} }
+\label{fig:fig8}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Change network.}
+Network structure of the changes within variables between time points.}
+\label{fig:fig9}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
+\label{fig:fig10}
+\end{figure}
+
+\hypertarget{tables}{%
+\section{Tables}\label{tables}}
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
 
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
-Agyapong-Badu, S., Warner, M., Samuel, D., and Stokes, M. (2016).
-Measurement of ageing effects on muscle tone and mechanical properties
-of rectus femoris and biceps brachii in healthy males and females using
-a novel hand-held myometric device. \emph{Archives of Gerontology and
-Geriatrics} 62, 59--67.
+Agyapong-Badu, S., Warner, M., Samuel, D., and Stokes, M.
+(2016).
+Measurement of ageing effects on muscle tone and mechanical properties of rectus femoris and biceps brachii in healthy males and females using a novel hand-held myometric device.
+\emph{Archives of Gerontology and Geriatrics} 62, 59--67.
 doi:\href{https://doi.org/10.1016/j.archger.2015.09.011}{10.1016/j.archger.2015.09.011}.
@@
 \leavevmode\hypertarget{ref-Aird2012}{}%
-Aird, L., Samuel, D., and Stokes, M. (2012). Quadriceps muscle tone,
-elasticity and stiffness in older males: Reliability and symmetry using
-the MyotonPRO. \emph{Archives of Gerontology and Geriatrics} 55,
-e31--e39.
+Aird, L., Samuel, D., and Stokes, M.
+(2012).
+Quadriceps muscle tone, elasticity and stiffness in older males: Reliability and symmetry using the MyotonPRO.
+\emph{Archives of Gerontology and Geriatrics} 55, e31--e39.
 doi:\href{https://doi.org/https://doi.org/10.1016/j.archger.2012.03.005}{https://doi.org/10.1016/j.archger.2012.03.005}.
@@
 \leavevmode\hypertarget{ref-Andersen1988}{}%
-Andersen, M. B., and Williams, J. M. (1988). A model of stress and
-athletic injury: Prediction and prevention. \emph{Journal of Sport and
-Exercise Psychology} 10, 294--306.
+Andersen, M.
+B., and Williams, J.
+M.
+(1988).
+A model of stress and athletic injury: Prediction and prevention.
+\emph{Journal of Sport and Exercise Psychology} 10, 294--306.
 doi:\href{https://doi.org/10.1123/jsep.10.3.294}{10.1123/jsep.10.3.294}.
@@
 \leavevmode\hypertarget{ref-Andersen1999}{}%
-Andersen, M. B., and Williams, J. M. (1999). Athletic injury,
-psychosocial factors and perceptual changes during stress. \emph{Journal
-of Sports Sciences} 17, 735--741.
+Andersen, M.
+B., and Williams, J.
+M.
+(1999).
+Athletic injury, psychosocial factors and perceptual changes during stress.
+\emph{Journal of Sports Sciences} 17, 735--741.
 doi:\href{https://doi.org/10.1080/026404199365597}{10.1080/026404199365597}.
@@
 \leavevmode\hypertarget{ref-Appaneal2009}{}%
-Appaneal, R. N., Levine, B. R., Perna, F. M., and Roh, J. L. (2009).
-Measuring postinjury depression among male and female competitive
-athletes. \emph{Journal of Sport and Exercise Psychology} 31, 60--76.
+Appaneal, R.
+N., Levine, B.
+R., Perna, F.
+M., and Roh, J.
+L.
+(2009).
+Measuring postinjury depression among male and female competitive athletes.
+\emph{Journal of Sport and Exercise Psychology} 31, 60--76.
 Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/19325188}.
@@
 \leavevmode\hypertarget{ref-Appaneal2014}{}%
-Appaneal, R. N., and Perna, F. M. (2014). ``Biopsychosocial model of
-injury,'' in \emph{Encyclopedia of sport and exercise psychology}, eds.
-R. C. Eklund and G. Tenenbaum (Thousand Oaks, CA: Sage), 74--77.
+Appaneal, R.
+N., and Perna, F.
+M.
+(2014).
+``Biopsychosocial model of injury,'' in \emph{Encyclopedia of sport and exercise psychology}, eds.
+R.
+C.
+Eklund and G.
+Tenenbaum (Thousand Oaks, CA: Sage), 74--77.
 
 \leavevmode\hypertarget{ref-Bellenger2016}{}%
-Bellenger, C. R., Fuller, J. T., Thomson, R. L., Davison, K., Robertson,
-E. Y., and Buckley, J. D. (2016). Monitoring athletic training status
-through autonomic heart rate regulation: A systematic review and
-meta-analysis. \emph{Sports Medicine} 46, 1461--1486.
+Bellenger, C.
+R., Fuller, J.
+T., Thomson, R.
+L., Davison, K., Robertson, E.
+Y., and Buckley, J.
+D.
+(2016).
+Monitoring athletic training status through autonomic heart rate regulation: A systematic review and meta-analysis.
+\emph{Sports Medicine} 46, 1461--1486.
 doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}.
@@
 \leavevmode\hypertarget{ref-Bittencourt2016}{}%
-Bittencourt, N. F. N., Meeuwisse, W. H., Mendonça, L. D.,
-Nettel-Aguirre, A., Ocarino, J. M., and Fonseca, S. T. (2016). Complex
-systems approach for sports injuries: Moving from risk factor
-identification to injury pattern recognition - narrative review and new
-concept. \emph{British Journal of Sports Medicine} 50, 1309--1314.
+Bittencourt, N.
+F.
+N., Meeuwisse, W.
+H., Mendonça, L.
+D., Nettel-Aguirre, A., Ocarino, J.
+M., and Fonseca, S.
+T.
+(2016).
+Complex systems approach for sports injuries: Moving from risk factor identification to injury pattern recognition - narrative review and new concept.
+\emph{British Journal of Sports Medicine} 50, 1309--1314.
 doi:\href{https://doi.org/10.1136/bjsports-2015-095850}{10.1136/bjsports-2015-095850}.
@@
 \leavevmode\hypertarget{ref-Brewer2012}{}%
-Brewer, B. W. (2012). ``Psychology of sport injury rehabilitation,'' in
-\emph{Handbook of sport psychology}, eds. G. Tenenbaum and R. C. Eklund
-(Hoboken, NJ, USA: Wiley), 404--424.
+Brewer, B.
+W.
+(2012).
+``Psychology of sport injury rehabilitation,'' in \emph{Handbook of sport psychology}, eds.
+G.
+Tenenbaum and R.
+C.
+Eklund (Hoboken, NJ, USA: Wiley), 404--424.
 doi:\href{https://doi.org/10.1002/9781118270011.ch18}{10.1002/9781118270011.ch18}.
 
+\leavevmode\hypertarget{ref-Butler2003}{}%
+Butler, R.
+J., Crowell, H.
+P., and Davis, I.
+M.
+C.
+(2003).
+Lower extremity stiffness: Implications for performance and injury.
+\emph{Clinical Biomechanics} 18, 511--517.
+doi:\href{https://doi.org/10.1016/S0268-0033(03)00071-8}{10.1016/S0268-0033(03)00071-8}.
+
 \leavevmode\hypertarget{ref-Burkner2017a}{}%
-Bürkner, P. C. (2017). brms: An R package for Bayesian multilevel models
-using Stan. \emph{Journal of Statistical Software} 80.
-doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}.
+Bürkner, P. C. (2017). brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} 80. doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}.
 
 \leavevmode\hypertarget{ref-Caminal2018}{}%
-Caminal, P., Sola, F., Gomis, P., Guasch, E., Perera, A., Soriano, N.,
-and Mont, L. (2018). Validity of the Polar V800 monitor for measuring
-heart rate variability in mountain running route conditions.
+Caminal, P., Sola, F., Gomis, P., Guasch, E., Perera, A., Soriano, N., and Mont, L.
+(2018).
+Validity of the Polar V800 monitor for measuring heart rate variability in mountain running route conditions.
 \emph{European Journal of Applied Physiology} 118, 669--677.
@@
 
+\leavevmode\hypertarget{ref-Chalmers2014}{}%
+Chalmers, J.
+A., Quintana, D.
+S., Abbott, M.
+J.-A., and Kemp, A.
+H.
+(2014).
+Anxiety disorders are associated with reduced heart rate variability: A meta-analysis.
+\emph{Frontiers in psychiatry} 5, 80.
+doi:\href{https://doi.org/10.3389/fpsyt.2014.00080}{10.3389/fpsyt.2014.00080}.
+
 \leavevmode\hypertarget{ref-Chen2012}{}%
-Chen, S. H., and Pollino, C. A. (2012). Good practice in Bayesian
-network modelling. \emph{Environmental Modelling and Software} 37,
-134--145.
+Chen, S.
+H., and Pollino, C.
+A.
+(2012).
+Good practice in Bayesian network modelling.
+\emph{Environmental Modelling and Software} 37, 134--145.
 doi:\href{https://doi.org/10.1016/j.envsoft.2012.03.012}{10.1016/j.envsoft.2012.03.012}.
@@
 \leavevmode\hypertarget{ref-Chuang2013}{}%
-Chuang, L. L., Lin, K. C., Wu, C. Y., Chang, C. W., Chen, H. C., Yin, H.
-P., and Wang, L. (2013). Relative and absolute reliabilities of the
-myotonometric measurements of hemiparetic arms in patients with stroke.
+Chuang, L.
+L., Lin, K.
+C., Wu, C.
+Y., Chang, C.
+W., Chen, H.
+C., Yin, H.
+P., and Wang, L.
+(2013).
+Relative and absolute reliabilities of the myotonometric measurements of hemiparetic arms in patients with stroke.
 \emph{Archives of Physical Medicine and Rehabilitation} 94, 459--466.
@@
 \leavevmode\hypertarget{ref-Chuang2012}{}%
-Chuang, L. L., Wu, C. Y., and Lin, K. C. (2012). Reliability, validity,
-and responsiveness of myotonometric measurement of muscle tone,
-elasticity, and stiffness in patients with stroke. \emph{Archives of
-Physical Medicine and Rehabilitation} 93, 532--540.
+Chuang, L.
+L., Wu, C.
+Y., and Lin, K.
+C.
+(2012).
+Reliability, validity, and responsiveness of myotonometric measurement of muscle tone, elasticity, and stiffness in patients with stroke.
+\emph{Archives of Physical Medicine and Rehabilitation} 93, 532--540.
 doi:\href{https://doi.org/10.1016/j.apmr.2011.09.014}{10.1016/j.apmr.2011.09.014}.
@@
 \leavevmode\hypertarget{ref-VanCingel2017}{}%
-Cingel, R. E. H. van, Hoogeboom, T. J., Melick, N. van, Meddeler, B. M.,
-and Nijhuis-van der Sanden, M. W. G. (2017). How to determine leg
-dominance: The agreement between self-reported and observed performance
-in healthy adults. \emph{Plos One} 12, 1--9.
+Cingel, R.
+E.
+H.
+van, Hoogeboom, T.
+J., Melick, N.
+van, Meddeler, B.
+M., and Nijhuis-van der Sanden, M.
+W.
+G.
+(2017).
+How to determine leg dominance: The agreement between self-reported and observed performance in healthy adults.
+\emph{Plos One} 12, 1--9.
 doi:\href{https://doi.org/10.1371/journal.pone.0189876}{10.1371/journal.pone.0189876}.
 
+\leavevmode\hypertarget{ref-Corr2013}{}%
+Corr, P.
+J.
+(2013).
+Approach and avoidance behaviour: Multiple systems and their interactions.
+\emph{Emotion Review} 5, 285--290.
+doi:\href{https://doi.org/10.1177/1754073913477507}{10.1177/1754073913477507}.
+
 \leavevmode\hypertarget{ref-Corr2016c}{}%
-Corr, P. J., and Cooper, A. J. (2016). The reinforcement sensitivity
-theory of personality questionnaire (RST-PQ): Development and
-validation. \emph{Psychological Assessment} 28, 1427--1440.
+Corr, P.
+J., and Cooper, A.
+J.
+(2016).
+The reinforcement sensitivity theory of personality questionnaire (RST-PQ): Development and validation.
+\emph{Psychological Assessment} 28, 1427--1440.
 doi:\href{https://doi.org/10.1037/pas0000273}{10.1037/pas0000273}.
 
+\leavevmode\hypertarget{ref-Corr2016b}{}%
+Corr, P.
+J., McNaughton, N., Wilson, M.
+R., Hutchison, A., Burch, G., and Poropat, A.
+(2016).
+Neuroscience of motivation and organizational behavior: Putting the reinforcement sensitivity theory (RST) to work.
+\emph{Advances in Motivation and Achievement} 19, 65--92.
+doi:\href{https://doi.org/10.1108/S0749-742320160000019010}{10.1108/S0749-742320160000019010}.
+
 \leavevmode\hypertarget{ref-Denver2007}{}%
-Denver, J. W., Reed, S. F., and Porges, S. W. (2007). Methodological
-issues in the quantification of respiratory sinus arrhythmia.
+Denver, J.
+W., Reed, S.
+F., and Porges, S.
+W.
+(2007).
+Methodological issues in the quantification of respiratory sinus arrhythmia.
 \emph{Biological Psychology} 74, 286--294.
@@
 \leavevmode\hypertarget{ref-Djaoui2017}{}%
-Djaoui, L., Haddad, M., Chamari, K., and Dellal, A. (2017). Monitoring
-training load and fatigue in soccer players with physiological markers.
+Djaoui, L., Haddad, M., Chamari, K., and Dellal, A.
+(2017).
+Monitoring training load and fatigue in soccer players with physiological markers.
 \emph{Physiology and Behavior} 181, 86--94.
@@
 \leavevmode\hypertarget{ref-Efron1993}{}%
-Efron, B., and Tibshirani, R. J. (1994). \emph{An introduction to the
-bootstrap}. New York: Chapman \& Hall Available at:
-\url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}.
+Efron, B., and Tibshirani, R.
+J.
+(1994).
+\emph{An introduction to the bootstrap}.
+New York: Chapman \& Hall Available at: \url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}.
 
 \leavevmode\hypertarget{ref-Enders2003}{}%
-Enders, C. K. (2003). Using the expectation maximization algorithm to
-estimate coefficient alpha for scales with item-level missing data.
+Enders, C.
+K.
+(2003).
+Using the expectation maximization algorithm to estimate coefficient alpha for scales with item-level missing data.
 \emph{Psychological Methods} 8, 322--337.
@@
 
+\leavevmode\hypertarget{ref-Fawkner1999}{}%
+Fawkner, H.
+J., McMurrary, N.
+E., and Summers, J.
+J.
+(1999).
+Athletic injury and minor life events: A prospective study.
+\emph{Journal of Science and Medicine in Sport} 2, 117--124.
+doi:\href{https://doi.org/10.1016/S1440-2440(99)80191-1}{10.1016/S1440-2440(99)80191-1}.
+
 \leavevmode\hypertarget{ref-Fornell1981}{}%
-Fornell, C., and Larcker, D. F. (1981). Evaluating structural equation
-models with unobservable variables and measurement error. \emph{Journal
-of Marketing Research} 18, 39--50.
+Fornell, C., and Larcker, D.
+F.
+(1981).
+Evaluating structural equation models with unobservable variables and measurement error.
+\emph{Journal of Marketing Research} 18, 39--50.
 doi:\href{https://doi.org/10.2307/3151312}{10.2307/3151312}.
@@
 \leavevmode\hypertarget{ref-Fuller2006}{}%
-Fuller, C. W., Ekstrand, J., Junge, A., Andersen, T. E., Bahr, R.,
-Dvorak, J., Hägglund, M., McCrory, P., and Meeuwisse, W. H. (2006).
-Consensus statement on injury definitions and data collection procedures
-in studies of football (soccer) injuries. \emph{British journal of
-sports medicine} 40, 193--201.
+Fuller, C.
+W., Ekstrand, J., Junge, A., Andersen, T.
+E., Bahr, R., Dvorak, J., Hägglund, M., McCrory, P., and Meeuwisse, W.
+H.
+(2006).
+Consensus statement on injury definitions and data collection procedures in studies of football (soccer) injuries.
+\emph{British journal of sports medicine} 40, 193--201.
 doi:\href{https://doi.org/10.1136/bjsm.2005.025270}{10.1136/bjsm.2005.025270}.
@@
 \leavevmode\hypertarget{ref-Fuller2007b}{}%
-Fuller, C. W., Molloy, M. G., Bagate, C., Bahr, R., Brooks, J. H. M.,
-Donson, H., Kemp, S. P. T., McCrory, P., McIntosh, A. S., Meeuwisse, W.
-H., et al. (2007). Consensus statement on injury definitions and data
-collection procedures for studies of injuries in rugby union.
+Fuller, C.
+W., Molloy, M.
+G., Bagate, C., Bahr, R., Brooks, J.
+H.
+M., Donson, H., Kemp, S.
+P.
+T., McCrory, P., McIntosh, A.
+S., Meeuwisse, W.
+H., et al.
+(2007).
+Consensus statement on injury definitions and data collection procedures for studies of injuries in rugby union.
 \emph{British journal of sports medicine} 41, 328--31.
@@
 \leavevmode\hypertarget{ref-Fuster-Parra2017}{}%
-Fuster-Parra, P., Vidal-Conti, J., Borràs, P. A., and Palou, P. (2017).
-Bayesian networks to identify statistical dependencies. A case study of
-Spanish university students' habits. \emph{Informatics for Health and
-Social Care} 42, 166--179.
+Fuster-Parra, P., Vidal-Conti, J., Borràs, P.
+A., and Palou, P.
+(2017).
+Bayesian networks to identify statistical dependencies.
+A case study of Spanish university students' habits.
+\emph{Informatics for Health and Social Care} 42, 166--179.
 doi:\href{https://doi.org/10.1080/17538157.2016.1178117}{10.1080/17538157.2016.1178117}.
@@
 \leavevmode\hypertarget{ref-Galambos2005}{}%
-Galambos, S. A., Terry, P. C., Moyle, G. M., and Locke, S. A. (2005).
-Psychological predictors of injury among elite athletes. \emph{British
-Journal of Sports Medicine} 39, 351--354.
+Galambos, S.
+A., Terry, P.
+C., Moyle, G.
+M., and Locke, S.
+A.
+(2005).
+Psychological predictors of injury among elite athletes.
+\emph{British Journal of Sports Medicine} 39, 351--354.
 doi:\href{https://doi.org/10.1136/bjsm.2005.018440}{10.1136/bjsm.2005.018440}.
@@
 \leavevmode\hypertarget{ref-Giles2016}{}%
-Giles, D., Draper, N., and Neil, W. (2016). Validity of the Polar V800
-heart rate monitor to measure RR intervals at rest. \emph{European
-Journal of Applied Physiology} 116, 563--571.
+Giles, D., Draper, N., and Neil, W.
+(2016).
+Validity of the Polar V800 heart rate monitor to measure RR intervals at rest.
+\emph{European Journal of Applied Physiology} 116, 563--571.
 doi:\href{https://doi.org/10.1007/s00421-015-3303-9}{10.1007/s00421-015-3303-9}.
 
+\leavevmode\hypertarget{ref-Gisselman2016}{}%
+Gisselman, A.
+S., Baxter, G.
+D., Wright, A., Hegedus, E., and Tumilty, S.
+(2016).
+Musculoskeletal overuse injuries and heart rate variability: Is there a link?
+\emph{Medical Hypotheses} 87, 1--7.
+doi:\href{https://doi.org/10.1016/j.mehy.2015.12.003}{10.1016/j.mehy.2015.12.003}.
+
 \leavevmode\hypertarget{ref-Heckerman1995}{}%
-Heckerman, D., Geiger, D., and Chickering, D. M. (1995). Learning
-Bayesian networks: The combination of knowledge and statistical data.
+Heckerman, D., Geiger, D., and Chickering, D.
+M.
+(1995).
+Learning Bayesian networks: The combination of knowledge and statistical data.
 \emph{Machine Learning} 20, 197--243.
@@
 
+\leavevmode\hypertarget{ref-Hrysomallis2011}{}%
+Hrysomallis, C. (2011). Balance ability and athletic performance. \emph{Sports Medicine} 41, 221--232. doi:\href{https://doi.org/10.2165/11538560-000000000-00000}{10.2165/11538560-000000000-00000}.
+
 \leavevmode\hypertarget{ref-Hughes2014}{}%
-Hughes, G. (2014). A review of recent perspectives on biomechanical risk
-factors associated with anterior cruciate ligament injury.
+Hughes, G.
+(2014).
+A review of recent perspectives on biomechanical risk factors associated with anterior cruciate ligament injury.
 \emph{Research in Sports Medicine} 22, 193--212.
@@
 \leavevmode\hypertarget{ref-Hunt2009}{}%
-Hunt, T. N., Ferrara, M. S., Bornstein, R. A., and Baumgartner, T. A.
-(2009). The reliability of the modified balance error scoring system.
+Hunt, T.
+N., Ferrara, M.
+S., Bornstein, R.
+A., and Baumgartner, T.
+A.
+(2009).
+The reliability of the modified balance error scoring system.
 \emph{Clinical Journal of Sport Medicine} 19, 471--475.
@@
 \leavevmode\hypertarget{ref-Ivarsson2010}{}%
-Ivarsson, A., and Johnson, U. (2010). Psychological factors as
-predictors of injuries among senior soccer players. A prospective study.
-\emph{Journal of Sports Science and Medicine} 9, 347--352. Available at:
-\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3761721/}.
+Ivarsson, A., and Johnson, U.
+(2010).
+Psychological factors as predictors of injuries among senior soccer players.
+A prospective study.
+\emph{Journal of Sports Science and Medicine} 9, 347--352.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3761721/}.
 
 \leavevmode\hypertarget{ref-Ivarsson2017}{}%
-Ivarsson, A., Johnson, U., Andersen, M. B., Tranaeus, U., Stenling, A.,
-and Lindwall, M. (2017). Psychosocial factors and sport injuries:
-Meta-analyses for prediction and prevention. \emph{Sports Medicine} 47,
-353--365.
+Ivarsson, A., Johnson, U., Andersen, M.
+B., Tranaeus, U., Stenling, A., and Lindwall, M.
+(2017).
+Psychosocial factors and sport injuries: Meta-analyses for prediction and prevention.
+\emph{Sports Medicine} 47, 353--365.
 doi:\href{https://doi.org/10.1007/s40279-016-0578-x}{10.1007/s40279-016-0578-x}.
 
+\leavevmode\hypertarget{ref-Ivarsson2014}{}%
+Ivarsson, A., Johnson, U., Lindwall, M., Gustafsson, H., and Altemyr, M.
+(2014).
+Psychosocial stress as a predictor of injury in elite junior soccer: A latent growth curve analysis.
+\emph{Journal of Science and Medicine in Sport} 17, 366--370.
+doi:\href{https://doi.org/10.1016/j.jsams.2013.10.242}{10.1016/j.jsams.2013.10.242}.
+
+\leavevmode\hypertarget{ref-Johnson2014}{}%
+Johnson, U., Tranaeus, U., and Ivarsson, A.
+(2014).
+Current status and future challenges in psychological research of sport injury prediction and prevention: A methodological perspective.
+\emph{Revista de Psicologia del Deporte} 23, 401--409.
+Available at: \url{https://www.rpd-online.com/article/view/v23-n2-johnson-tranaeus-ivarsson}.
+
+\leavevmode\hypertarget{ref-Kalkhoven2018}{}%
+Kalkhoven, J.
+T., and Watsford, M.
+L.
+(2017).
+The relationship between mechanical stiffness and athletic performance markers in sub-elite footballers.
+\emph{Journal of Sports Sciences} 36, 1022--1029.
+doi:\href{https://doi.org/10.1080/02640414.2017.1349921}{10.1080/02640414.2017.1349921}.
+
 \leavevmode\hypertarget{ref-Koller2009}{}%
-Koller, D., and Friedman, N. (2009). \emph{Probabilistic graphical
-models: Principles and techniques - Adaptive computation and machine
-learning}. Cambridge, MA: The MIT Press
-doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.
+Koller, D., and Friedman, N.
+(2009).
+\emph{Probabilistic graphical models: Principles and techniques - Adaptive computation and machine learning}.
+Cambridge, MA: The MIT Press doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.
 
 \leavevmode\hypertarget{ref-Koo2016}{}%
-Koo, T. K., and Li, M. Y. (2016). A guideline of selecting and reporting
-intraclass correlation coefficients for reliability research.
+Koo, T.
+K., and Li, M.
+Y.
+(2016).
+A guideline of selecting and reporting intraclass correlation coefficients for reliability research.
 \emph{Journal of Chiropractic Medicine} 15, 155--63.
@@
 \leavevmode\hypertarget{ref-Kuhn2008}{}%
-Kuhn, M. C. from J. W., Weston, S., Williams, A., Keefer, C., and
-Engelhardt, A. (2008). caret: Classification and Regression Training.
+Kuhn, M.
+C.
+from J.
+W., Weston, S., Williams, A., Keefer, C., and Engelhardt, A.
+(2008).
+caret: Classification and Regression Training.
 \emph{Journal of Statistical Software} 28.
@@
 \leavevmode\hypertarget{ref-Kumar2001}{}%
-Kumar, S. (2001). Theories of musculoskeletal injury causation.
-\emph{Ergonomics} 44, 17--47.
-doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}.
+Kumar, S. (2001). Theories of musculoskeletal injury causation. \emph{Ergonomics} 44, 17--47. doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}.
 
 \leavevmode\hypertarget{ref-Laborde2017}{}%
-Laborde, S., Mosley, E., and Thayer, J. F. (2017). Heart rate
-variability and cardiac vagal tone in psychophysiological research -
-Recommendations for experiment planning, data analysis, and data
-reporting. \emph{Frontiers in Psychology} 8, 1--18.
+Laborde, S., Mosley, E., and Thayer, J.
+F.
+(2017).
+Heart rate variability and cardiac vagal tone in psychophysiological research - Recommendations for experiment planning, data analysis, and data reporting.
+\emph{Frontiers in Psychology} 8, 1--18.
 doi:\href{https://doi.org/10.3389/fpsyg.2017.00213}{10.3389/fpsyg.2017.00213}.
@@
 \leavevmode\hypertarget{ref-Lang2014}{}%
-Lang, K. M., Jorgensen, T. D., Moore, E. W. G., and Little, T. D.
-(2013). On the joys of missing data. \emph{Journal of Pediatric
-Psychology} 39, 151--162.
+Lang, K.
+M., Jorgensen, T.
+D., Moore, E.
+W.
+G., and Little, T.
+D.
+(2013).
+On the joys of missing data.
+\emph{Journal of Pediatric Psychology} 39, 151--162.
 doi:\href{https://doi.org/10.1093/jpepsy/jst048}{10.1093/jpepsy/jst048}.
@@
 \leavevmode\hypertarget{ref-Lavallee1996}{}%
-Lavallée, L., and Flint, F. (1996). The relationship of stress,
-competitive anxiety, mood state, and social support to athletic injury.
-\emph{Journal of Athletic Training} 31, 296--299. Available at:
-\url{http://www.ncbi.nlm.nih.gov/pubmed/16558413}.
+Lavallée, L., and Flint, F.
+(1996).
+The relationship of stress, competitive anxiety, mood state, and social support to athletic injury.
+\emph{Journal of Athletic Training} 31, 296--299.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16558413}.
 
 \leavevmode\hypertarget{ref-Leddy1994}{}%
-Leddy, M. H., Lambert, M. J., and Ogles, B. M. (1994). Psychological
-consequences of athletic injury among high-level competitors.
+Leddy, M.
+H., Lambert, M.
+J., and Ogles, B.
+M.
+(1994).
+Psychological consequences of athletic injury among high-level competitors.
 \emph{Research Quarterly for Exercise and Sport} 65, 347--354.
@@
 \leavevmode\hypertarget{ref-Lee2017}{}%
-Lee, E. C., Fragala, M. S., Kavouras, S. A., Queen, R. M., Pryor, J. L.,
-and Casa, D. J. (2017). Biomarkers in sports and exercise: Tracking
-health, performance, and recovery in athletes. \emph{Journal of Strength
-and Conditioning Research} 31, 2920--2937.
+Lee, E.
+C., Fragala, M.
+S., Kavouras, S.
+A., Queen, R.
+M., Pryor, J.
+L., and Casa, D.
+J.
+(2017).
+Biomarkers in sports and exercise: Tracking health, performance, and recovery in athletes.
+\emph{Journal of Strength and Conditioning Research} 31, 2920--2937.
 doi:\href{https://doi.org/10.1519/JSC.0000000000002122}{10.1519/JSC.0000000000002122}.
@@
 \leavevmode\hypertarget{ref-Maddison2005}{}%
-Maddison, R., and Prapavessis, H. (2005). A psychological approach to
-the prediction and prevention of athletic injury. \emph{Journal of Sport
-and Exercise Psychology} 27, 289--310.
+Maddison, R., and Prapavessis, H.
+(2005).
+A psychological approach to the prediction and prevention of athletic injury.
+\emph{Journal of Sport and Exercise Psychology} 27, 289--310.
 doi:\href{https://doi.org/10.1123/jsep.27.3.289}{10.1123/jsep.27.3.289}.
@@
 \leavevmode\hypertarget{ref-Malik1996}{}%
-Malik, M., Camm, A. J., Bigger, J. T., Breithardt, G., Cerutti, S.,
-Cohen, R. J., Coumel, P., Fallen, E. L., Kennedy, H. L., Kleiger, R. E.,
-et al. (1996). Heart rate variability. Standards of measurement,
-physiological interpretation, and clinical use. \emph{European Heart
-Journal} 17, 354--381.
+Malik, M., Camm, A.
+J., Bigger, J.
+T., Breithardt, G., Cerutti, S., Cohen, R.
+J., Coumel, P., Fallen, E.
+L., Kennedy, H.
+L., Kleiger, R.
+E., et al.
+(1996).
+Heart rate variability.
+Standards of measurement, physiological interpretation, and clinical use.
+\emph{European Heart Journal} 17, 354--381.
 doi:\href{https://doi.org/10.1093/oxfordjournals.eurheartj.a014868}{10.1093/oxfordjournals.eurheartj.a014868}.
@@
 \leavevmode\hypertarget{ref-McCrory2013}{}%
-McCrory, P., Meeuwisse, W. H., Aubry, M., Cantu, R. C., Dvořák, J.,
-Echemendia, R. J., Engebretsen, L., Johnston, K., Kutcher, J. S.,
-Raftery, M., et al. (2013). Consensus statement on concussion in sport:
-The 4th international conference on concussion in sport, Zurich,
-November 2012. \emph{Journal of Athletic Training} 48, 554--575.
+McCrory, P., Meeuwisse, W.
+H., Aubry, M., Cantu, R.
+C., Dvořák, J., Echemendia, R.
+J., Engebretsen, L., Johnston, K., Kutcher, J.
+S., Raftery, M., et al.
+(2013).
+Consensus statement on concussion in sport: The 4th international conference on concussion in sport, Zurich, November 2012.
+\emph{Journal of Athletic Training} 48, 554--575.
 doi:\href{https://doi.org/10.4085/1062-6050-48.4.05}{10.4085/1062-6050-48.4.05}.
@@
 \leavevmode\hypertarget{ref-Meeuwisse2007}{}%
-Meeuwisse, W. H., Tyreman, H., Hagel, B., and Emery, C. (2007). A
-dynamic model of etiology in sport injury: The recursive nature of risk
-and causation. \emph{Clinical Journal of Sport Medicine} 17, 215--219.
+Meeuwisse, W.
+H., Tyreman, H., Hagel, B., and Emery, C.
+(2007).
+A dynamic model of etiology in sport injury: The recursive nature of risk and causation.
+\emph{Clinical Journal of Sport Medicine} 17, 215--219.
 doi:\href{https://doi.org/10.1097/JSM.0b013e3180592a48}{10.1097/JSM.0b013e3180592a48}.
@@
 \leavevmode\hypertarget{ref-Morgan2018}{}%
-Morgan, G. E., Martin, R., Williams, L., Pearce, O., and Morris, K.
-(2018). Objective assessment of stiffness in Achilles tendinopathy: a
-novel approach using the MyotonPRO. \emph{BMJ Open Sport \& Exercise
-Medicine} 4, e000446.
+Morgan, G.
+E., Martin, R., Williams, L., Pearce, O., and Morris, K.
+(2018).
+Objective assessment of stiffness in Achilles tendinopathy: a novel approach using the MyotonPRO.
+\emph{BMJ Open Sport \& Exercise Medicine} 4, e000446.
 doi:\href{https://doi.org/10.1136/bmjsem-2018-000446}{10.1136/bmjsem-2018-000446}.
@@
 \leavevmode\hypertarget{ref-Murphy2003}{}%
-Murphy, D. F., Connolly, D. A. J., and Beynnon, B. D. (2003). Risk
-factors for lower extremity injury: A review of the literature.
+Murphy, D.
+F., Connolly, D.
+A.
+J., and Beynnon, B.
+D.
+(2003).
+Risk factors for lower extremity injury: A review of the literature.
 \emph{British Journal of Sports Medicine} 37, 13--29.
@@
 \leavevmode\hypertarget{ref-Murphy2002}{}%
-Murphy, K. (2002). Dynamic Bayesian networks: Representation, inference
-and learning.
+Murphy, K. (2002). Dynamic Bayesian networks: Representation, inference and learning.
 
 \leavevmode\hypertarget{ref-Nair2014}{}%
-Nair, K., Dougherty, J., Schaefer, E., Kelly, J., and Masi, A. (2014).
-Repeatability, reproducibility, and calibration of the MyotonPRO on
-tissue mimicking phantoms. in \emph{ASME summer bioengineering
-conference}, 1--2.
+Nair, K., Dougherty, J., Schaefer, E., Kelly, J., and Masi, A.
+(2014).
+Repeatability, reproducibility, and calibration of the MyotonPRO on tissue mimicking phantoms.
+in \emph{ASME summer bioengineering conference}, 1--2.
 doi:\href{https://doi.org/10.1115/SBC2013-14622}{10.1115/SBC2013-14622}.
@@
 \leavevmode\hypertarget{ref-Neely1998}{}%
-Neely, F. G. (1998). Biomechanical risk factors for exercise-related
-lower limb injuries. \emph{Sports Medicine} 26, 395--413.
+Neely, F.
+G.
+(1998).
+Biomechanical risk factors for exercise-related lower limb injuries.
+\emph{Sports Medicine} 26, 395--413.
 doi:\href{https://doi.org/10.2165/00007256-199826060-00003}{10.2165/00007256-199826060-00003}.
@@
 \leavevmode\hypertarget{ref-Olmedilla2018}{}%
-Olmedilla, A., Rubio, V. J., Fuster-Parra, P., Pujals, C., and
-García-Mas, A. (2018). A Bayesian approach to sport injuries likelihood:
-Does player's self-efficacy and environmental factors plays the main
-role? \emph{Frontiers in Psychology} 9, 1--10.
+Olmedilla, A., Rubio, V.
+J., Fuster-Parra, P., Pujals, C., and García-Mas, A.
+(2018).
+A Bayesian approach to sport injuries likelihood: Does player's self-efficacy and environmental factors plays the main role?
+\emph{Frontiers in Psychology} 9, 1--10.
 doi:\href{https://doi.org/10.3389/fpsyg.2018.01174}{10.3389/fpsyg.2018.01174}.
 
+\leavevmode\hypertarget{ref-Paillard2006}{}%
+Paillard, T., Noé, F., Rivière, T., Marion, V., Montoya, R., and Dupui, P.
+(2006).
+Postural performance and strategy in the unipedal stance of soccer players at different levels of competition.
+\emph{Journal of Athletic Training} 41, 172--176.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16791302}.
+
 \leavevmode\hypertarget{ref-Passer1983a}{}%
-Passer, M. W., and Seese, M. D. (1983). Life stress and athletic injury:
-Examination of positive versus negative events and three moderator
-variables. \emph{Journal of Human Stress} 9, 11--16.
+Passer, M.
+W., and Seese, M.
+D.
+(1983).
+Life stress and athletic injury: Examination of positive versus negative events and three moderator variables.
+\emph{Journal of Human Stress} 9, 11--16.
 doi:\href{https://doi.org/10.1080/0097840X.1983.9935025}{10.1080/0097840X.1983.9935025}.
@@
 \leavevmode\hypertarget{ref-Pearl1988}{}%
-Pearl, J. (1988). \emph{Probabilistic reasoning in intelligent systems:
-Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann
-Publishers.
+Pearl, J. (1988). \emph{Probabilistic reasoning in intelligent systems: Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann Publishers.
 
 \leavevmode\hypertarget{ref-Perna2003}{}%
-Perna, F. M., Antoni, M. H., Baum, A., Gordon, P., and Schneiderman, N.
-(2003). Cognitive behavioral stress management effects on injury and
-illness among competitive athletes: A randomized clinical trial.
-\emph{Annals of Behavioral Medicine} 25, 66--73. Available at:
-\url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}.
+Perna, F.
+M., Antoni, M.
+H., Baum, A., Gordon, P., and Schneiderman, N.
+(2003).
+Cognitive behavioral stress management effects on injury and illness among competitive athletes: A randomized clinical trial.
+\emph{Annals of Behavioral Medicine} 25, 66--73.
+Available at: \url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}.
 
 \leavevmode\hypertarget{ref-Perna1995}{}%
-Perna, F. M., and McDowell, S. L. (1995). Role of psychological stress
-in cortisol recovery from exhaustive exercise among elite athletes.
+Perna, F.
+M., and McDowell, S.
+L.
+(1995).
+Role of psychological stress in cortisol recovery from exhaustive exercise among elite athletes.
 \emph{International Journal of Behavioral Medicine} 2, 13--26.
@@
 \leavevmode\hypertarget{ref-Perna1997}{}%
-Perna, F., Schneiderman, N., and LaPerriere, A. (1997). Psychological
-stress, exercise and immunity. \emph{International Journal of Sports
-Medicine} 18, 78--83.
+Perna, F., Schneiderman, N., and LaPerriere, A.
+(1997).
+Psychological stress, exercise and immunity.
+\emph{International Journal of Sports Medicine} 18, 78--83.
 doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}.
 
+\leavevmode\hypertarget{ref-Petrie1993}{}%
+Petrie, T.
+A.
+(1993).
+Coping skills, competitive trait anxiety, and playing status: Moderating effects on the life stress-injury relationship.
+\emph{Journal of Sport and Exercise Psychology} 15, 261--274.
+doi:\href{https://doi.org/10.1123/jsep.15.3.261}{10.1123/jsep.15.3.261}.
+
 \leavevmode\hypertarget{ref-Petrie1992}{}%
-Petrie, T. A. (1992). Psychosocial antecedents of athletic injury: The
-effects of life stress and social support on female collegiate gymnasts.
+Petrie, T.
+A.
+(1992).
+Psychosocial antecedents of athletic injury: The effects of life stress and social support on female collegiate gymnasts.
 \emph{Journal of Behavioral Medicine} 18, 127--138.
@@
 
+\leavevmode\hypertarget{ref-Petrie2014}{}%
+Petrie, T.
+A., Deiters, J., and Harmison, R.
+J.
+(2014).
+Mental toughness, social support, and athletic identity: Moderators of the life stress-injury relationship in collegiate football players.
+\emph{Sport, Exercise, and Performance Psychology} 3, 13--27.
+doi:\href{https://doi.org/10.1037/a0032698}{10.1037/a0032698}.
+
+\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
+Pickering-Rodriguez, E.
+C., Watsford, M.
+L., Bower, R.
+G., and Murphy, A.
+J.
+(2017).
+The relationship between lower body stiffness and injury incidence in female netballers.
+\emph{Sports Biomechanics} 16, 361--373.
+doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}.
+
 \leavevmode\hypertarget{ref-Plews2017}{}%
-Plews, D. J., Scott, B., Altini, M., Wood, M., Kilding, A. E., and
-Laursen, P. B. (2017). Comparison of heart-rate-variability recording
-with smartphone photoplethysmography, polar H7 chest strap, and
-electrocardiography. \emph{International Journal of Sports Physiology
-and Performance} 12, 1324--1328.
+Plews, D.
+J., Scott, B., Altini, M., Wood, M., Kilding, A.
+E., and Laursen, P.
+B.
+(2017).
+Comparison of heart-rate-variability recording with smartphone photoplethysmography, polar H7 chest strap, and electrocardiography.
+\emph{International Journal of Sports Physiology and Performance} 12, 1324--1328.
 doi:\href{https://doi.org/10.1123/ijspp.2016-0668}{10.1123/ijspp.2016-0668}.
@@
 \leavevmode\hypertarget{ref-Pruyn2015}{}%
-Pruyn, E. C., Watsford, M. L., and Murphy, A. J. (2015). Differences in
-lower-body stiffness between levels of netball competition.
+Pruyn, E.
+C., Watsford, M.
+L., and Murphy, A.
+J.
+(2015).
+Differences in lower-body stiffness between levels of netball competition.
 \emph{Journal of Strength and Conditioning Research} 29, 1197--1202.
@@
 \leavevmode\hypertarget{ref-Pruyn2016}{}%
-Pruyn, E. C., Watsford, M. L., and Murphy, A. J. (2016). Validity and
-reliability of three methods of stiffness assessment. \emph{Journal of
-Sport and Health Science} 5, 476--483.
+Pruyn, E.
+C., Watsford, M.
+L., and Murphy, A.
+J.
+(2016).
+Validity and reliability of three methods of stiffness assessment.
+\emph{Journal of Sport and Health Science} 5, 476--483.
 doi:\href{https://doi.org/10.1016/j.jshs.2015.12.001}{10.1016/j.jshs.2015.12.001}.
 
+\leavevmode\hypertarget{ref-Pulopulos2018}{}%
+Pulopulos, M.
+M., Vanderhasselt, M.
+A., and De Raedt, R.
+(2018).
+Association between changes in heart rate variability during the anticipation of a stressful situation and the stress-induced cortisol response.
+\emph{Psychoneuroendocrinology} 94, 63--71.
+doi:\href{https://doi.org/10.1016/j.psyneuen.2018.05.004}{10.1016/j.psyneuen.2018.05.004}.
+
+\leavevmode\hypertarget{ref-Qian2015}{}%
+Qian, S.
+S., and Miltner, R.
+J.
+(2015).
+A continuous variable Bayesian networks model for water quality modeling: A case study of setting nitrogen criterion for small rivers and streams in Ohio, USA.
+\emph{Environmental Modelling and Software} 69, 14--22.
+doi:\href{https://doi.org/10.1016/j.envsoft.2015.03.001}{10.1016/j.envsoft.2015.03.001}.
+
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing
-Available at: \url{https://www.r-project.org/}.
+R Core Team (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing Available at: \url{https://www.r-project.org/}.
 
 \leavevmode\hypertarget{ref-Riemann1999d}{}%
-Riemann, B. L., Guskiewicz, K. M., and Shields, E. W. (1999).
-Relationship between clinical and forceplate measures of postural
-stability. \emph{Journal of Sport Rehabilitation} 8, 71--82.
+Riemann, B.
+L., Guskiewicz, K.
+M., and Shields, E.
+W.
+(1999).
+Relationship between clinical and forceplate measures of postural stability.
+\emph{Journal of Sport Rehabilitation} 8, 71--82.
 doi:\href{https://doi.org/10.1123/jsr.8.2.71}{10.1123/jsr.8.2.71}.
 
+\leavevmode\hypertarget{ref-Rodriguez-Linares2017}{}%
+Rodriguez-Linares, L., Vila, X., Lado, M.
+J., Mendez, A., Otero, A., and Garcia, C.
+A.
+(2019).
+\emph{RHRV: Heart rate variability analysis of ecg data}.
+Available at: \url{https://CRAN.R-project.org/package=RHRV}.
+
 \leavevmode\hypertarget{ref-Rogers2005}{}%
-Rogers, T., and Landers, D. M. (2005). Mediating effects of peripheral
-vision in the life event stress/athletic injury relationship.
+Rogers, T., and Landers, D.
+M.
+(2005).
+Mediating effects of peripheral vision in the life event stress/athletic injury relationship.
 \emph{Journal of Sport and Exercise Psychology} 27, 271--288.
@@
 \leavevmode\hypertarget{ref-Romero-Franco2014}{}%
-Romero-Franco, N., Gallego-Izquierdo, T., Martínez-López, E. J.,
-Hita-Contreras, F., Osuna-Pére, Catalina, M., and Martínez-Amat, A.
-(2014). Postural stability and subsequent sports injuries during indoor
-season of athletes. \emph{Journal of Physical Therapy Science} 26,
-683--687.
+Romero-Franco, N., Gallego-Izquierdo, T., Martínez-López, E.
+J., Hita-Contreras, F., Osuna-Pére, Catalina, M., and Martínez-Amat, A.
+(2014).
+Postural stability and subsequent sports injuries during indoor season of athletes.
+\emph{Journal of Physical Therapy Science} 26, 683--687.
 doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}.
@@
 \leavevmode\hypertarget{ref-Rosa2014}{}%
-Rosa, B. B., Asperti, A. M., Helito, C. P., Demange, M. K., Fernandes,
-T. L., and Hernandez, A. J. (2014). Epidemiology of sports injuries on
-collegiate athletes at a single center. \emph{Acta Ortopédica
-Brasileira} 22, 321--324.
+Rosa, B.
+B., Asperti, A.
+M., Helito, C.
+P., Demange, M.
+K., Fernandes, T.
+L., and Hernandez, A.
+J.
+(2014).
+Epidemiology of sports injuries on collegiate athletes at a single center.
+\emph{Acta Ortopédica Brasileira} 22, 321--324.
 doi:\href{https://doi.org/10.1590/1413-78522014220601007}{10.1590/1413-78522014220601007}.
 
+\leavevmode\hypertarget{ref-Ross2011}{}%
+Ross, S.
+E., Linens, S.
+W., Wright, C.
+J., and Arnold, B.
+L.
+(2011).
+Balance assessments for predicting functional ankle instability and stable ankles.
+\emph{Gait and Posture} 34, 539--542.
+doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}.
+
+\leavevmode\hypertarget{ref-Ruiz-Perez2019}{}%
+Ruiz-Pérez, I., Ayala, F., Miguel Puerta, J., Elvira, J.
+L.
+L., De Ste Croix, M., Hernández-Sánchez, S., and Vera-Garcia, F.
+J.
+(2019).
+A Bayesian Network approach to study the relationships between several neuromuscular performance measures and dynamic postural control in futsal players.
+\emph{PLoS ONE} 14, 1--18.
+doi:\href{https://doi.org/10.1371/journal.pone.0220065}{10.1371/journal.pone.0220065}.
+
 \leavevmode\hypertarget{ref-Norvig2009}{}%
-Russell, S. J., and Norvig, P. (2009). \emph{Artificial Intelligence: A
-Modern Approach}. 3rd ed. Prentice Hall Available at:
-\url{http://aima.cs.berkeley.edu/}.
-
-\leavevmode\hypertarget{ref-Scutari2017}{}%
-Scutari, M., Auconi, P., Caldarelli, G., and Franchi, L. (2017).
-Bayesian networks analysis of malocclusion data. \emph{Scientific
-Reports} 7, 1--11.
-doi:\href{https://doi.org/10.1038/s41598-017-15293-w}{10.1038/s41598-017-15293-w}.
+Russell, S. J., and Norvig, P. (2009). \emph{Artificial Intelligence: A Modern Approach}. 3rd ed. Prentice Hall Available at: \url{http://aima.cs.berkeley.edu/}.
 
 \leavevmode\hypertarget{ref-Scutari2014}{}%
-Scutari, M., and Denis, J.-B. (2014). \emph{Bayesian networks: with
-examples in R}. 1st ed. Chapman \& Hall/CRC Available at:
-\url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}.
+Scutari, M., and Denis, J.-B.
+(2014).
+\emph{Bayesian networks: with examples in R}.
+1st ed.
+Chapman \& Hall/CRC Available at: \url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}.
 
 \leavevmode\hypertarget{ref-Scutari2013}{}%
-Scutari, M., and Nagarajan, R. (2013). Identifying significant edges in
-graphical models of molecular networks. \emph{Artificial Intelligence in
-Medicine} 57, 207--217.
+Scutari, M., and Nagarajan, R.
+(2013).
+Identifying significant edges in graphical models of molecular networks.
+\emph{Artificial Intelligence in Medicine} 57, 207--217.
 doi:\href{https://doi.org/10.1016/j.artmed.2012.12.006}{10.1016/j.artmed.2012.12.006}.
@@
 \leavevmode\hypertarget{ref-Sheu2016}{}%
-Sheu, Y., Chen, L. H., and Hedegaard, H. (2016). Sports- and
-recreation-related injury episodes in the United States, 2011-2014.
-\emph{National Health Statistics Reports}, 1--12. Available at:
-\url{https://www.ncbi.nlm.nih.gov/pubmed/27906643}.
+Sheu, Y., Chen, L.
+H., and Hedegaard, H.
+(2016).
+Sports- and recreation-related injury episodes in the United States, 2011-2014.
+\emph{National Health Statistics Reports}, 1--12.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/27906643}.
 
 \leavevmode\hypertarget{ref-Slimani2018}{}%
-Slimani, M., Bragazzi, N. L., Znazen, H., Paravlic, A., Azaiez, F., and
-Tod, D. (2018). Psychosocial predictors and psychological prevention of
-soccer injuries: A systematic review and meta-analysis of the
-literature. \emph{Physical Therapy in Sport} 32, 293--300.
+Slimani, M., Bragazzi, N.
+L., Znazen, H., Paravlic, A., Azaiez, F., and Tod, D.
+(2018).
+Psychosocial predictors and psychological prevention of soccer injuries: A systematic review and meta-analysis of the literature.
+\emph{Physical Therapy in Sport} 32, 293--300.
 doi:\href{https://doi.org/10.1016/j.ptsp.2018.05.006}{10.1016/j.ptsp.2018.05.006}.
@@
 \leavevmode\hypertarget{ref-Smith1990}{}%
-Smith, R. E., Smoll, F. L., and Ptacek, J. T. (1990). Conjunctive
-moderator variables in vulnerability and resiliency research: Life
-stress, social support and coping skills, and adolescent sport injuries.
+Smith, R.
+E., Smoll, F.
+L., and Ptacek, J.
+T.
+(1990).
+Conjunctive moderator variables in vulnerability and resiliency research: Life stress, social support and coping skills, and adolescent sport injuries.
 \emph{Journal of Personality and Social Psychology} 58, 360--370.
@@
 \leavevmode\hypertarget{ref-Swanik2007}{}%
-Swanik, C. B., Covassin, T., Stearne, D. J., and Schatz, P. (2007). The
-relationship between neurocognitive function and noncontact anterior
-cruciate ligament injuries. \emph{American Journal of Sports Medicine}
-35, 943--948.
+Swanik, C.
+B., Covassin, T., Stearne, D.
+J., and Schatz, P.
+(2007).
+The relationship between neurocognitive function and noncontact anterior cruciate ligament injuries.
+\emph{American Journal of Sports Medicine} 35, 943--948.
 doi:\href{https://doi.org/10.1177/0363546507299532}{10.1177/0363546507299532}.
@@
 \leavevmode\hypertarget{ref-Timpka2014}{}%
-Timpka, T., Alonso, J. M., Jacobsson, J., Junge, A., Branco, P.,
-Clarsen, B., Kowalski, J., Mountjoy, M., Nilsson, S., Pluim, B., et al.
-(2014). Injury and illness definitions and data collection procedures
-for use in epidemiological studies in Athletics (track and field):
-Consensus statement. \emph{British Journal of Sports Medicine} 48,
-483--490.
+Timpka, T., Alonso, J.
+M., Jacobsson, J., Junge, A., Branco, P., Clarsen, B., Kowalski, J., Mountjoy, M., Nilsson, S., Pluim, B., et al.
+(2014).
+Injury and illness definitions and data collection procedures for use in epidemiological studies in Athletics (track and field): Consensus statement.
+\emph{British Journal of Sports Medicine} 48, 483--490.
 doi:\href{https://doi.org/10.1136/bjsports-2013-093241}{10.1136/bjsports-2013-093241}.
 
+\leavevmode\hypertarget{ref-Trojian2006}{}%
+Trojian, T.
+H., and McKeag, D.
+B.
+(2006).
+Single leg balance test to identify risk of ankle sprains.
+\emph{British Journal of Sports Medicine} 40, 610--613.
+doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}.
+
 \leavevmode\hypertarget{ref-Valovich2003}{}%
-Valovich, T. C., Perrin, D. H., and Gansneder, B. M. (2003). Repeat
-administration elicits a practice effect with the Balance Error Scoring
-System but not with the Standardized Assessment of Concussion in high
-school athletes. \emph{Journal of Athletic Training} 38, 51--56.
+Valovich, T.
+C., Perrin, D.
+H., and Gansneder, B.
+M.
+(2003).
+Repeat administration elicits a practice effect with the Balance Error Scoring System but not with the Standardized Assessment of Concussion in high school athletes.
+\emph{Journal of Athletic Training} 38, 51--56.
 Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/12937472}.
@@
 \leavevmode\hypertarget{ref-Wiese-Bjornstal2009}{}%
-Wiese-Bjornstal, D. M. (2009). Sport injury and college athlete health
-across the lifespan. \emph{Journal of Intercollegiate Sport} 2, 64--80.
+Wiese-Bjornstal, D.
+M.
+(2009).
+Sport injury and college athlete health across the lifespan.
+\emph{Journal of Intercollegiate Sport} 2, 64--80.
 doi:\href{https://doi.org/10.1123/jis.2.1.64}{10.1123/jis.2.1.64}.
@@
 \leavevmode\hypertarget{ref-Wilkerson2012a}{}%
-Wilkerson, G. B. (2012). Neurocognitive reaction time predicts lower
-extremity sprains and strains. \emph{International Journal of Athletic
-Therapy and Training} 17, 4--9.
+Wilkerson, G.
+B.
+(2012).
+Neurocognitive reaction time predicts lower extremity sprains and strains.
+\emph{International Journal of Athletic Therapy and Training} 17, 4--9.
 doi:\href{https://doi.org/10.1123/ijatt.17.6.4}{10.1123/ijatt.17.6.4}.
@@
 \leavevmode\hypertarget{ref-Williams2007}{}%
-Williams, J. M., and Andersen, M. B. (2007). ``Psychosocial antecedents
-of sport injury and interventions for risk reduction,'' in
-\emph{Handbook of sport psychology}, eds. G. Tenenbaum and R. C. Eklund
-(Hoboken, NJ, USA: Wiley), 379--403.
+Williams, J.
+M., and Andersen, M.
+B.
+(2007).
+``Psychosocial antecedents of sport injury and interventions for risk reduction,'' in \emph{Handbook of sport psychology}, eds.
+G.
+Tenenbaum and R.
+C.
+Eklund (Hoboken, NJ, USA: Wiley), 379--403.
 doi:\href{https://doi.org/10.1002/9781118270011.ch17}{10.1002/9781118270011.ch17}.
@@
 \leavevmode\hypertarget{ref-Williams1998}{}%
-Williams, J. M., and Andersen, M. B. (1998). Psychosocial antecedents of
-sport injury: review and critique of the stress and injury model.
+Williams, J.
+M., and Andersen, M.
+B.
+(1998).
+Psychosocial antecedents of sport injury: review and critique of the stress and injury model.
 \emph{Journal of Applied Sport Psychology} 10, 5--25.
@@
 Williams, S., Booton, T., Watson, M., Rowland, D., and Altini, M.
-(2017). Heart rate variability is a moderating factor in the
-workload-injury relationship of competitive crossfit™ athletes.
+(2017).
+Heart rate variability is a moderating factor in the workload-injury relationship of competitive crossfit™ athletes.
 \emph{Journal of Sports Science and Medicine} 16, 443--449.
@@
 \leavevmode\hypertarget{ref-Xiao-xuan2007}{}%
-Xiao-xuan, H., Hui, W., and Shuo, W. (2008). Using expert's knowledge to
-build Bayesian networks. in \emph{International conference on
-computational intelligence and security workshops} (IEEE), 220--223.
+Xiao-xuan, H., Hui, W., and Shuo, W.
+(2008).
+Using expert's knowledge to build Bayesian networks.
+in \emph{International conference on computational intelligence and security workshops} (IEEE), 220--223.
 doi:\href{https://doi.org/10.1109/cisw.2007.4425484}{10.1109/cisw.2007.4425484}.
@@
 \leavevmode\hypertarget{ref-Xu2015}{}%
-Xu, J. G., Zhao, Y., Chen, J., and Han, C. (2015). A structure learning
-algorithm for Bayesian network using prior knowledge. \emph{Journal of
-Computer Science and Technology} 30, 713--724.
+Xu, J.
+G., Zhao, Y., Chen, J., and Han, C.
+(2015).
+A structure learning algorithm for Bayesian network using prior knowledge.
+\emph{Journal of Computer Science and Technology} 30, 713--724.
 doi:\href{https://doi.org/10.1007/s11390-015-1556-8}{10.1007/s11390-015-1556-8}.
+\end{cslreferences}
+
+\endgroup
 

### 2020-07-20 full_paper_v2.Rmd
pre results edits

@@
 
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
+A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
 Participants self-rated competitive level ranged from recreational to international standard. 
+They were engaged in training for their respective sports for at least five hours per week.
 A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at least five hours per week.
+All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study. 
 Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
 
-```{r table1, results='asis'}
-
-kable(
-  npars %>% 
-    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
-    slice(c(1,3,8, 4, 9:11)) %>% 
-  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
-                          name == "Height" ~ "Height (cm)",
-                          name == "Weight" ~ "Body mass (kg)",
-                          TRUE ~ name)),
-  caption = "Participant characteristics.",
-  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
-  align = "c",
-  escape = F
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Demographics M (SD)", 1, 4) %>%
-  pack_rows("Current competitive level n (%)", 5, 7) %>%
-  row_spec(0, bold = TRUE)
-```
-
 ## Measures
 
+### Injury
+
+Participants self-reported any injuries they sustained at each data collection during the study period.
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
+Injury status (did / did not sustain an injury) served as the main outcome measure.
+
 ### Major life events
 
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992]. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study population (S1 Table).
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
+The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table). 
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
-Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
-Participants were asked to rate the perceived impact of each life event they had experienced within the last 12 months on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores respectively. 
-A score for total life events was calculated by summing the absolute values for both negative and positive events.
+Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”.
+Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4  ($extremely\;negative$) and +4 ($extremely\;positive$).
+Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
+A score for total life events was also calculated by summing the absolute values for both negative and positive events.
 Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively.
 Petrie also provided evidence of discriminant, convergent and predictive validity.
-The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-For this study, Composite Reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
-The composite reliability for the LESCA in this study was 0.84.
+For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+The composite reliability for the LESCA in this study was .84.
 
@@
 
-```{r table2, results='asis'}
-# exampledata <- tibble(
-#   Participant = c("1", "1", "2", "3", "3"),
-#   "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-#   "link" = "->",
-#   "X_2" = c("T2", "T3", "T2", "T2", "T3")
-# )
-# 
-# kable(
-#   exampledata,
-#   caption = "Example of the data arrangement used for the network.",
-#   escape = FALSE
-# ) %>%
-#   kable_styling(latex_options = "HOLD_position") %>%
-#   row_spec(0, bold = TRUE)
-```
-
 #### Preliminary network structures
 
-Prior to the final network structure presented in the results section, several structures with different combinations of variables were explred.
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
 Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models.
@@
 
-Finally, both total score and asymmetry were included in the initial network.
+Finally, both total score and asymmetry for balance were included in the initial network.
 However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
 Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
-
-```{r finalvars, results='asis'}
-library(knitr)
-library(kableExtra)
-varcutoffs %>% 
-  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
-) %>% 
-  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
-  kable("latex",
-       booktabs = T,
-       align = c("llcc"),
-       caption = "Variables included in the final Bayesian network structure.",
-       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
-  kable_styling(latex_options = "HOLD_position") %>% 
-  column_spec(2, width = c("5.2cm")) 
-```
-
-Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
-For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
-Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-Therefore, these arcs were blacked listed from the network.
-
+To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
+
+Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
+For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%.
+Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network.
+Therefore, these arcs were blacklisted from the network.
 To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-```{r table5, results='asis'}
-query1 %>%
-  select(c(1, 2, 3)) %>%
-  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
 The CPQ for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
@@
 
-```{r table6, results='asis'}
-
-query2 %>%
-  select(7, 2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
 Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
@@
 Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state.
-
-```{r table7, results="asis"}
-injured1comp %>%
-  filter(variable != "outcome") %>% 
-  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(  .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
-    col.names = c("Variable", "High risk", "Low risk"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-<!-- The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}. -->
-<!-- Again, only the three lowest and highest probabilities are shown (complete results in S5 Table). -->
-<!-- The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
--->
-<!-- With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state. -->
-<!-- The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured". -->
-
-<!-- ```{r table9, results="asis" } -->
-
-<!-- query4 %>% -->
-<!--   select(7,2:6) %>% -->
-<!--   rename(Probability = "prob") %>% -->
-<!--   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>% -->
-<!--   mutate_all(~str_replace_all(., "_", "\\\\_"))%>% -->
-<!--   kable( -->
-<!--     ., -->
-<!--     caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.", -->
-<!--     escape = F -->
-<!--   ) %>% -->
-<!--   kable_styling(latex_options = "HOLD_position") %>% -->
-<!--   pack_rows("Highest", 1, 3) %>% -->
-<!--   pack_rows("Lowest", 4, 6) %>% -->
-<!--   row_spec(0,bold=TRUE) -->
-<!-- ``` -->
+The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
+Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
+The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
+With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
+The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
 
@@
 There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table10, results="asis"}
-
-bmtable %>%
-  kable(
-    .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
-    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
 
@@
 
+Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
+In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
+
+# Discussion
+
+Informed by Appaneal and Perna’s (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature [@Johnson2014].
+Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables _across_ time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables _between_ time points (Fig \ref{fig:fig8}).
+
+The first network revealed several links between the injured nodes and other variables in the network.
+For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured_1 and injured_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
+The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for injured_1 and injured_2.
+The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury. 
+In the second network, the highest probability of injury was observed when changes in stiffness _and_ negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury [@Appaneal2014].
+Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
+Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury.
+The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
+[@Bittencourt2016], where injury is an emergent property from a web of determinants that interact in different ways.
+
+Of all the variables measured in the study, muscle stiffness appeared to be most strongly related to injury.
+Both "High" levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
+In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
+To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players [@PickeringRodriguez2017].
+The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
+However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness [@Pruyn2015; @Kalkhoven2018].
+Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury [@Butler2003].
+Additionally, high stiffness may only increase the risk of injury if other factors are also present.
+To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
+In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
+Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
+
+In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance $\rightarrow$ injured_1 and balance $\rightarrow$  injured_2 respectively.
+Despite the weak arc strength, a "High" balance score, indicating impaired postural stability, was found to increase the probability of injury.
+This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
+Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts [@Paillard2006].
+In the current study, athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
+These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition [@Hrysomallis2011].
+
+Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
+In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
+However, negative life event stress had almost no effect on the probability of injury in the first network.
+This finding suggests that the _relative change_ in life events may be more important than the _absolute score_ for life events, despite the latter being commonly used in sports injury research to date.
+For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
+Although the absolute score would be "Low", the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect [@Appaneal2014].
+Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed [cf. @Ivarsson2014].
+
+The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a _negative_ valence, as the strongest predictor of injury in Williams and Andersen's [@Williams1998] model [@Ivarsson2017].
+During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
+However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg_1 $\rightarrow$ injured_1 and nlelg_2 $\rightarrow$ injured_2 to allow this relationship to be examined more closely.
+When negative life events were "High" the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[3,3] - query3[3,2]` at the injured_2 node.
+One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
+In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
+However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred _since the previous data collection session_, to avoid inflated scores caused by reporting the same event on multiple occasions.
+The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
+For example, at the second and third time points, 26% of participants reported 0 negative life events for the preceding four-month period.
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events [cf.
+@Fawkner1999].
+
+Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
+Williams and Andersen's [@Williams1998] model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
+Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury [@Petrie1993; @Petrie2014; @Johnson2014].
+Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
+
+Of the remaining variables, both FFFS and RMSSD were also linked to injury.
+A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was weaker than muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
+Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, "Low" RMSSD increased the probability of injury from 0.17 (RMSSD = "High'') to 0.27 (RMSSD = "Low'').
+This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
+An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
+Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points; for example, the "Low" FFFS recorded at the first time point was related to injuries recorded at the second time point in the study.
+One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
+The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
+
+In the first network, "High" BIS was associated with "High" FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
+RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
+High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018], which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
+These findings align with the proposed actions of the RST theory [@Corr2016b]. 
+For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems. 
+The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict [@Corr2016b].
+To extend these findings, the BAS should also be considered. 
+Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
+However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
+In the first network, RI_1 was connected to both BIS_1 and BIS_2, and in both instances, the probability of "High" BIS was increased when RI was also "High". 
+However, the arcs between RI_1 and BIS were weak (< 0.50), and RI represents only one component of the BAS system. 
+Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury. 
+Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
+
+The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury [@Johnson2014].
+The longitudinal repeated measures design of the current study allowed _changes_ over time and between time points to be captured and explored.
+Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
+Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
+Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
+The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
+Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018].
+
+As with most research, there were several limitations with the present study.
+Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
+Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
+For example, the meaning of a "Low" and "High" value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
+A second limitation was the nature of the physiological measures used in the current study.
+In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
+For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis [@Ross2011].
+
+In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
+Although the current study used a range of measures to capture "stress" from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
+For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], and could help elucidate the relationship between stress and injury.
+Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
+For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
+Both responses would be scored a "-4"; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
+Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
+
+In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
+Specifically, muscle stiffness and _increases_ in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
+
+# Disclosure/Conflict-of-Interest Statement
+
+The authors declare that the research was conducted in the absence of any
+commercial or financial relationships that could be construed as a potential
+conflict of interest.
+
+# Author Contributions
+
+<!--
+
+When determining authorship the following criteria should be observed:
+
+ - Substantial contributions to the conception or design of the work; or the
+   acquisition, analysis, or interpretation of data for the work; AND
+ - Drafting the work or revising it critically for important intellectual
+   content; AND
+ - Final approval of the version to be published ; AND
+ - Agreement to be accountable for all aspects of the work in ensuring that
+   questions related to the accuracy or integrity of any part of the work are
+   appropriately investigated and resolved.
+
+Contributors who meet fewer than all 4 of the above criteria for authorship
+should not be listed as authors, but they should be acknowledged.
+(http://www.icmje.org/roles_a.html)
+
+-->
+
+The statement about the authors and contributors can be up to several sentences
+long, describing the tasks of individual authors referred to by their initials
+and should be included at the end of the manuscript before the References
+section.
+
+# Supplemental Data
+
+Supplementary Material should be uploaded separately on submission, if there are
+Supplementary Figures, please include the caption in the same file as the
+figure. LaTeX Supplementary Material templates can be found in the Frontiers
+LaTeX folder
+
+# References
+
+A reference list should be automatically created here. However it won't. Pandoc 
+will place the list of references at the end of the document instead. There are 
+no convenient solution for now to force Pandoc to do otherwise. The easiest way 
+to get around this problem is to edit the LaTeX file created by Pandoc before 
+compiling it again using the traditional LaTeX commands.
+
+# Figures
+
+```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```
+
+\begin{figure}[!h]
+\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
+\label{fig:fig1}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf mBESS positions (A-F).}
+Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
+\label{fig:fig2}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Study design.}
+For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
+\label{fig:fig3}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Session protocol.}
+Outline of the protocol for each data collection.}
+\label{fig:fig4}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Example network.}
+A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
+\label{fig:fig5}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Network structure.}
+The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+\label{fig:fig6}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket of injured\_1.}
+Arc strengths are included as arc labels.}
+\label{fig:fig7}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for injured\_2.} }
+\label{fig:fig8}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Change network.}
+Network structure of the changes within variables between time points.}
+\label{fig:fig9}
+\end{figure}
+
 \begin{figure}[H]
@@
 \end{figure}
+
+# Tables
+
+```{r table1, results='asis'}
+
+kable(
+  npars %>% 
+    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
+    slice(c(1,3,8, 4, 9:11)) %>% 
+  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
+                          name == "Height" ~ "Height (cm)",
+                          name == "Weight" ~ "Body mass (kg)",
+                          TRUE ~ name)),
+  caption = "Participant characteristics.",
+  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
+  align = "c",
+  escape = F
+) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  pack_rows("Demographics M (SD)", 1, 4) %>%
+  pack_rows("Current competitive level n (%)", 5, 7) %>%
+  row_spec(0, bold = TRUE)
+```
+
+```{r table2, results='asis'}
+exampledata <- tibble(
+  Participant = c("1", "1", "2", "3", "3"),
+  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
+  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
+)
+
+kable(
+  exampledata,
+  caption = "Example of the data arrangement used for the network.",
+  escape = FALSE
+) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0, bold = TRUE)
+```
+
+```{r table3, results='asis'}
+varcutoffs %>% 
+  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
+) %>% 
+  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
+  kable("latex",
+       align = c("llcc"),
+       caption = "Variables included in the final Bayesian network structure.",
+       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
+  kable_styling(latex_options = "HOLD_position") %>% 
+  column_spec(2, width = c("5.2cm")) 
+```
+
+```{r table4, results='asis'}
+
+cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
+  kable(
+    .,
+    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
+    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
+    escape = F
+  ) %>%
+  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
+  footnote(general = "Other included bone, skin and brain injuries.") %>%
+  kable_styling(latex_options = "HOLD_position")  %>%
+  row_spec(0,bold=TRUE)
+
+```
+
+```{r table5, results='asis'}
+query1 %>%
+  select(c(1, 2, 3)) %>%
+  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
+  kable(
+    .,
+    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
+    col.names = c("Variable", "Low", "High"),
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+```
+
+```{r table6, results='asis'}
+
+query2 %>%
+  select(7, 2:6) %>%
+  rename(Probability = "prob") %>%
+  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
+  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
+  kable(
+    .,
+    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  pack_rows("Highest", 1, 3) %>%
+  pack_rows("Lowest", 4, 6) %>%
+  row_spec(0,bold=TRUE)
+```
+
+```{r table7, results="asis"}
+injured1comp %>%
+  filter(variable != "outcome") %>% 
+  mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>% 
+  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
+  kable(  .,
+    caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.",
+    col.names = c("Variable", "High risk", "Low risk"),
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+```
+
+```{r table8, results="asis"}
+
+query3 %>%
+  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
+  kable(
+    .,
+    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
+    col.names = c("Variable", "Low", "High"),
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+```
+
+```{r table9, results="asis" }
+
+query4 %>%
+  select(7,2:6) %>%
+  rename(Probability = "prob") %>%
+  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
+  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
+  kable(
+    .,
+    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  pack_rows("Highest", 1, 3) %>%
+  pack_rows("Lowest", 4, 6) %>%
+  row_spec(0,bold=TRUE)
+```
+
+```{r table10, results="asis"}
+
+bmtable %>%
+  kable(
+    .,
+    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
+    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
+    escape = F
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+  row_spec(0,bold=TRUE)
+
+```
 
@@
 
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
-
 ```{r table12, results="asis"}
@@
 ```
-
-# Discussion
-
-Informed by Appaneal and Perna’s (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature [@Johnson2014].
-Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables _across_ time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables _between_ time points (Fig \ref{fig:fig8}).
-
-The first network revealed several links between the injured nodes and other variables in the network.
-For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured_1 and injured_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
-The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for injured_1 and injured_2.
-The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury. 
-In the second network, the highest probability of injury was observed when changes in stiffness _and_ negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury [@Appaneal2014].
-Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
-Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury.
-The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
-[@Bittencourt2016], where injury is an emergent property from a web of determinants that interact in different ways.
-
-Of all the variables measured in the study, muscle stiffness appeared to be most strongly related to injury.
-Both "High" levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
-In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
-To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players [@PickeringRodriguez2017].
-The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
-However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness [@Pruyn2015; @Kalkhoven2018].
-Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury [@Butler2003].
-Additionally, high stiffness may only increase the risk of injury if other factors are also present.
-To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
-In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
-
-In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance $\rightarrow$ injured_1 and balance $\rightarrow$  injured_2 respectively.
-Despite the weak arc strength, a "High" balance score, indicating impaired postural stability, was found to increase the probability of injury.
-This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
-Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts [@Paillard2006].
-In the current study, athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
-These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition [@Hrysomallis2011].
-
-Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
-In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
-However, negative life event stress had almost no effect on the probability of injury in the first network.
-This finding suggests that the _relative change_ in life events may be more important than the _absolute score_ for life events, despite the latter being commonly used in sports injury research to date.
-For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
-Although the absolute score would be "Low", the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect [@Appaneal2014].
-Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed [cf. @Ivarsson2014].
-
-The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a _negative_ valence, as the strongest predictor of injury in Williams and Andersen's [@Williams1998] model [@Ivarsson2017].
-During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg_1 $\rightarrow$ injured_1 and nlelg_2 $\rightarrow$ injured_2 to allow this relationship to be examined more closely.
-When negative life events were "High" the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[3,3] - query3[3,2]` at the injured_2 node.
-One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
-In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
-However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred _since the previous data collection session_, to avoid inflated scores caused by reporting the same event on multiple occasions.
-The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
-For example, at the second and third time points, 26% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events [cf.
-@Fawkner1999].
-
-Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
-Williams and Andersen's [@Williams1998] model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
-Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury [@Petrie1993; @Petrie2014; @Johnson2014].
-Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
-
-Of the remaining variables, both FFFS and RMSSD were also linked to injury.
-A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was weaker than muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
-Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, "Low" RMSSD increased the probability of injury from 0.17 (RMSSD = "High'') to 0.27 (RMSSD = "Low'').
-This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
-An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
-Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points; for example, the "Low" FFFS recorded at the first time point was related to injuries recorded at the second time point in the study.
-One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
-The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-
-In the first network, "High" BIS was associated with "High" FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
-RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
-High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018], which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
-These findings align with the proposed actions of the RST theory [@Corr2016b]. 
-For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems. 
-The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict [@Corr2016b].
-To extend these findings, the BAS should also be considered. 
-Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
-However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
-In the first network, RI_1 was connected to both BIS_1 and BIS_2, and in both instances, the probability of "High" BIS was increased when RI was also "High". 
-However, the arcs between RI_1 and BIS were weak (< 0.50), and RI represents only one component of the BAS system. 
-Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury. 
-Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
-
-The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury [@Johnson2014].
-The longitudinal repeated measures design of the current study allowed _changes_ over time and between time points to be captured and explored.
-Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
-Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
-To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
-The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
-Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018].
-
-As with most research, there were several limitations with the present study.
-Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
-Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
-For example, the meaning of a "Low" and "High" value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
-A second limitation was the nature of the physiological measures used in the current study.
-In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
-For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis [@Ross2011].
-
-In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
-Although the current study used a range of measures to capture "stress" from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
-For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], and could help elucidate the relationship between stress and injury.
-Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
-For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
-Both responses would be scored a "-4"; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
-Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
-
-In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
-Specifically, muscle stiffness and _increases_ in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
-Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
-
-# Disclosure/Conflict-of-Interest Statement
-
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Author Contributions
-
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# References
-
-A reference list should be automatically created here. However it won't. Pandoc 
-will place the list of references at the end of the document instead. There are 
-no convenient solution for now to force Pandoc to do otherwise. The easiest way 
-to get around this problem is to edit the LaTeX file created by Pandoc before 
-compiling it again using the traditional LaTeX commands.
-
-# Figures
-
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}

### 2020-07-20 full_paper_v2.tex
pre results edits

--- 2020-05-23 demo.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
-%                                              %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Version 3.4 Generated 2018/06/15 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-
-\documentclass[utf8]{frontiersHLTH}
-
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-
-\linenumbers
-
-% BELOW TAKEN FROM rticles plos template
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
 %
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% hyperref package, useful for hyperlinks
-\usepackage{hyperref}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
+\documentclass[
+]{frontiersHLTH}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={An interdisciplinary examination of stress and injury occurrence in athletes},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
-
-% Sweave(-like)
-\usepackage{fancyvrb}
-\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
-\newenvironment{Schunk}{}{}
-\DefineVerbatimEnvironment{Code}{Verbatim}{}
-\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-\newenvironment{CodeChunk}{}{}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% Below is from frontiers
-%
-\bibliographystyle{frontiersinSCNS}
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace}
-%\doublespacing
-
-% Leave a blank line between paragraphs instead of using \\
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
-
-\def\Authors{
-  Dr Harry Fisher\,\textsuperscript{1*},
-  Dr Marianne Gittoes\,\textsuperscript{1},
-  Professor Lynne Evans\,\textsuperscript{1},
-  Miss Leah Bitchell\,\textsuperscript{1},
-  Dr Richard Mullen\,\textsuperscript{2},
-  Dr Marco Scutari\,\textsuperscript{3}}
-
-\def\Address{
-
-  \textsuperscript{1} School of Sport and Health Science, Cardiff Metropolitan University,  Cardiff,   United Kingdom
-
-  \textsuperscript{2} Division of Sport, Health \& Exercise Sciences, Brunel University,  London,   United Kingdom
-
-  \textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA),  Manno,   Switzerland
-  }
-
-  \def\corrAuthor{Dr Harry Fisher}\def\corrAddress{Cardiff Metropolitan University\\Cyncoed Road\\, CF23 6XD United Kingdom}\def\corrEmail{\href{mailto:harryfisher21@gmail.com}{\nolinkurl{harryfisher21@gmail.com}}}
-  \def\firstAuthorLast{Fisher {et~al.}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
+\title{An interdisciplinary examination of stress and injury occurrence
+in athletes}
+\author{true \and true \and true \and true \and true \and true}
+\date{}
 
 \begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Short Title]{An interdisciplinary examination of stress and injury occurrence in
-athletes}
-\author[\firstAuthorLast]{\Authors}
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
-
 \maketitle
@@
 
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-Over the last four decades sport related injuries have received
-increased research attention (Ivarsson et al., 2017). This attention is
-unsurprising given the high incidence (Rosa et al., 2014; Sheu et al.,
-2016), and undesirable physical and psychological effects of sports
-injuries (Leddy et al., 1994; Brewer, 2012). To mitigate against both
-the increasing incidence and undesirable consequences of injury,
-research has identified several psychological (Slimani et al., 2018),
-anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes,
-2014) and environmental (Meeuwisse et al., 2007) factors associated with
-sports injury occurrence. Indeed, several models of injury causation
-have been proposed that highlight the multifactorial nature of injury
+\section{Introduction}
+
+Over the last four decades sport-related injuries have received
+increased research attention (Ivarsson et al., 2017) in response to the
+high incidence (Rosa et al., 2014; Sheu et al., 2016) and associated
+undesirable physical and psychological effects (Leddy et al., 1994;
+Brewer, 2012). Multiple psychological (Slimani et al., 2018), anatomical
+(Murphy et al., 2003), biomechanical (Neely, 1998; Hughes, 2014) and
+environmental (Meeuwisse et al., 2007) factors have been associated with
+sports injury occurrence and several models of injury causation have
+been proposed that highlight the multifactorial nature of injury
 occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009),
@@
 
+Williams and Andersen's (Williams and Andersen, 1998) stress-injury
+model proposed that when faced with a potentially stressful athletic
+situation, an athlete's personality traits (e.g., hardiness, locus of
+control and competitive trait anxiety), history of stressors (e.g.,
+major life events and previous injuries) and coping resources (e.g.,
+general coping behaviours) contribute to the injury response, either
+interactively or in isolation. The stress response is central to the
+model and reflects the bi-directional relationship between athletes'
+appraisal of, and response to, a stressful athletic situation. The model
+predicts that athletes who have a history of stressors, personality
+traits that intensify the stress response and few coping resources, will
+exhibit greater attentional (e.g., peripheral narrowing) and/or
+physiological (e.g., increased muscle tension) responses that put these
+individuals at greater injury risk.
+
+Within Williams and Andersen's (Williams and Andersen, 1998) model,
+major life events, which is a component of an athlete's history of
+stressors, most consistently predicts injury occurrence (Williams and
+Andersen, 2007); specifically, major life events with a negative, as
+opposed to positive, valence (Passer and Seese, 1983; Maddison and
+Prapavessis, 2005). Personality traits and coping resources have also
+been found to predict injury with athletes more likely to sustain an
+injury if they have poor social support and psychological coping skills,
+and high trait anxiety and elevated competitive state anxiety; compared
+to those with the opposing profile (Smith et al., 1990; Lavallée and
+Flint, 1996; Ivarsson and Johnson, 2010).However, the amount of variance
+explained by these psychosocial factors has been modest and typically
+between 5 - 30\% (Galambos et al., 2005; Ivarsson and Johnson, 2010),
+which indicates a likely interaction with other factors.
+
+While the psychosocial factors in Williams and Andersen's (Williams and
+Andersen, 1998) model have received the most research attention, less
+insight into the mechanisms through which these factors are proposed to
+exert their effect exists. To elaborate, the model suggests that
+injuries are likely to occur through either increased physiological
+arousal resulting in increased muscle tension and reduced flexibility or
+attentional deficits caused by increased distractibility and peripheral
+narrowing. However, to date, the research has largely focused on
+attentional deficits (Andersen and Williams, 1999; Rogers and Landers,
+2005; Wilkerson, 2012; Swanik et al., 2007). For example, Andersen and
+Williams (Andersen and Williams, 1999) found athletes with high life
+event stress coupled with low social support had greater peripheral
+narrowing under stressful conditions compared to athletes with the
+opposing profile; these athletes went on to sustain an increased number
+of injuries during the following athletic season. Rodgers and Landers
+(Rogers and Landers, 2005) further supported Andersen and Williams's
+(Andersen and Williams, 1999) earlier findings by identifying that
+peripheral narrowing under stress mediated 8.1\% of the relationship
+between negative life events and injury.
+
+Knowledge of the physiological factors (e.g., increased muscle tension
+and reduced motor control) contributing to the remaining variance
+between negative life events and athletic injury remains sparse
+(cf.~Williams and Andersen, 1998). One challenge faced by researchers
+addressing the sports injury problem through a psychological lens is the
+multifactorial nature of injury, and the possible interaction with
+physiological factors in the stress response (Meeuwisse et al., 2007;
+Wiese-Bjornstal, 2009). For example, a large body of research has
+suggested that training-related stress is also likely to contribute to
+the stress response and injury occurrence (Lee et al., 2017; Djaoui et
+al., 2017) and may account for the unexplained variance from the
+psychological predictors. Appaneal and Perna (Appaneal and Perna, 2014)
+proposed the biopsychosocial model of stress athletic injury and health
+(BMSAIH) to serve as an extension to Williams and Andersen's (Williams
+and Andersen, 1998) model. The BMSAIH enhances our understanding of the
+mediating pathways between the stress response and injury alongside
+other health outcomes and behavioral factors that impact sports
+participation (Appaneal and Perna, 2014). The central tenet of the
+BMSAIH is that psychosocial distress (e.g., negative life events) may
+act synergistically with training-related stress as a result of
+high-intensity and high-volume sports training, and ``widen the window
+of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of
+undesirable health outcomes including illness and injury. Consequently,
+the BMSAIH provides an important framework that has enhanced insight
+into the multi-faceted nature of the injury process by building on
+Williams and Andersen's (Williams and Andersen, 1998) model whilst
+including other physiological markers of training-related stress.
+
+Although research supporting the BMSAIH has mainly focused on the
+relationship between hormonal responses to training and injury
+occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other
+training-related stress markers including heart rate variability
+(Bellenger et al., 2016; Williams et al., 2017), muscle stiffness (Pruyn
+et al., 2015) and postural stability (Romero-Franco et al., 2014) have
+been reported to be associated with an increased risk of injury.
+However, the largely mono-disciplinary approach to examining the role of
+these markers has provided limited insight into the multifaceted
+interactions with psychological factors that may contribute to injury
+occurrence. Furthermore, a reliance on designs that capture a single
+point of measurement precludes the assessment of intra as well as
+inter-individual changes and the effect of the time interval between
+measurement and injury occurrence on subsequent injuries (Johnson et
+al., 2014). Such an approach fails to capture changes in both
+psychological factors and physiological markers that may occur preceding
+an injury. Importantly, a repeated measurement approach would enable an
+assessing of how variables and their interactions change with respect to
+time provide greater insight into the effect that repeated exposure to
+major life events and other stress-related factors has on injury
+occurrence.
+
+Recently, Bittencourt et al.~(Bittencourt et al., 2016) advocated a move
+away from studying isolated risk factors and instead, adopt a complex
+systems approach in order to understand injury occurrence. Such an
+approach posits that injury may arise from a complex ``web of
+determinants'' (Bittencourt et al., 2016, 3), where different factors
+interact in unpredictable and unplanned ways, but result in a global
+outcome pattern of either adaptation or injury. Capturing the
+uncertainty and complexity of the relationships between different
+variables using an appropriate interdisciplinary analysis within the
+framework of a complex systems approach is challenging. Bayesian network
+(BN) modelling provides one solution by allowing the construction of
+graphical probabilistic models using the underlying structure that
+connects different variables (Scutari and Denis, 2014). The learned BN
+structure can be used for inference by obtaining the posterior
+probabilities of a particular variable for a given query (e.g., if the
+value of variable A is x and the value of variable B is y, what is the
+probability variable C of being value z?). Furthermore, unlike
+regression or structural equation models, BN's do not distinguish
+between dependent and independent variables when the underlying
+relationship in the network may not be known (Olmedilla et al., 2018).
+BN modelling subsequently provides a valuable but underused
+interdisciplinary approach to investigating the complex and
+unpredictable interactions of psychological and physiological factors
+implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen's (Williams and
+Andersen, 1998) stress injury model and Appaneal's BMSAIH model, the aim
+of this interdisciplinary study was to develop new understanding of the
+multifaceted interactions of psychological and physiological
+stress-related factors with injury occurrence. A prospective, repeated
+measures design incorporating field-based physiological and
+psychological measures combined with a BN modelling analysis was used to
+address the study aim.
+
+\section{Methods}
+
+\subsection{Participants}
+
+A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120)
+were initially recruited for the study from a British University and
+local sports clubs (Table \ref{tab:table1}). Participants had an average
+age of 22.0 \(\pm\) 7.0 years and represented a range of team (football,
+rugby, netball, cricket, lacrosse, basketball and field hockey) and
+individual sports (athletics, tennis, weightlifting, gymnastics, judo,
+swimming and golf. Participants self-rated competitive level ranged from
+recreational to international standard. They were engaged in training
+for their respective sports for at least five hours per week. A total of
+126 (49.03\%) participants had sustained an injury in the 12 months
+prior to the start of the study (male: \emph{n} = 114 {[}49\%{]},
+female; \emph{n} = 48 {[}40\%{]}). All participants were injury free (no
+modifications to their usual training routine due to a sport related
+medical problem for a minimum of four weeks) at the start of the study.
+Ethical approval was obtained from the University ethics committee prior
+to the start of the study and all participants provided informed
+consent.
+
+\hypertarget{measures}{%
+\subsection{Measures}\label{measures}}
+
+\hypertarget{injury}{%
+\subsubsection{Injury}\label{injury}}
+
+Participants self-reported any injuries they sustained at each data
+collection during the study period. An injury was defined as any sports
+related medical problem causing the athlete to miss or modify their
+usual training routine for at least 24 hours (Fuller et al., 2006, 2007;
+Timpka et al., 2014). Minor scrapes and bruises that may require certain
+modifications (e.g., strapping or taping) but did not limit continued
+participation were not considered injuries (cf.~Appaneal et al., 2009).
+Injury status (did / did not sustain an injury) served as the main
+outcome measure.
+
+\hypertarget{major-life-events}{%
+\subsubsection{Major life events}\label{major-life-events}}
+
+A modified version of the Life Events Survey for Collegiate Athletes
+(LESCA) was used to measure participants' history of life event stress
+(Petrie, 1992). The LESCA is the most widely used measure of major life
+events for athletes in the sports injury literature. Modifications were
+made to the LESCA to ensure the suitability of the items for the study
+cohort (S1 Table). The LESCA comprises 69 items that reflect possible
+life events that participants may have experienced. Example items
+include, ``Major change in the frequency (increased or decreased) of
+social activities due to participation in sport'', ``Major change in the
+amount (more or less) of academic activity (homework, class time, etc)''
+and ``Major change in level of athletic performance in actual
+competition (better or worse)''. Participants were asked to rate the
+perceived impact of each life event they had experienced within
+12-months preceding the study onset on an 8-point Likert scale anchored
+at -4 (\(extremely\;negative\)) and +4 (\(extremely\;positive\)).
+Negative and positive life event scores were calculated by summing the
+negative and positive scores, respectively. A score for total life
+events was also calculated by summing the absolute values for both
+negative and positive events. Petrie (1992) (Petrie, 1992) reported
+test-retest reliabilities at 1-week and 8-weeks with values ranging from
+.76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001)
+respectively. Petrie also provided evidence of discriminant, convergent
+and predictive validity. For this study, composite reliability (Fornell
+and Larcker, 1981) was preferred to Cronbach's alpha as it does not
+assume parallelity (i.e., all factor loadings are constrained to be
+equal, and all error variances are constrained to be equal) and instead
+takes into consideration the varying factor loadings of the items in the
+questionnaire. The composite reliability for the LESCA in this study was
+.84.
+
+\hypertarget{reinforcement-sensitivity-theory-personality-questionnaire}{%
+\subsubsection{Reinforcement Sensitivity Theory Personality
+Questionnaire}\label{reinforcement-sensitivity-theory-personality-questionnaire}}
+
+A revised version of the Reinforcement Sensitivity Theory Personality
+Questionnaire (RST-PQ) was used to measure motivation, emotion,
+personality and their relevance to psychopathology (Corr and Cooper,
+2016). The revised version of the RST-PQ comprises 51 statements that
+measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I
+am the sort of person who easily freezes-up when scared''), Behavioural
+Inhibition System (BIS; e.g., ``When trying to make a decision, I find
+myself constantly chewing it over'') and four Behavioural Approach
+System (BAS) factors; Reward Interest (e.g., ``I regularly try new
+activities just to see if I enjoy them''), Goal Drive Persistence (e.g.,
+``I am very persistent in achieving my goals''), Reward Reactivity
+(e.g., ``I get a special thrill when I am praised for something I've
+done well'') and Impulsivity (e.g., ``I find myself doing things on the
+spur of the moment''). Participants rated each item on a scale from 1
+(\(not\; at\; all\)) to 4 (\(highly\)) to reflect how well each
+statement described their personality in general. The responses to items
+associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed
+to give a total personality score that was subsequently used for further
+analysis. The composite reliabilities for each subscale were; BIS =
+0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further
+details regarding the revised RST are in S1 Appendix.
+
+\hypertarget{heart-rate-variability}{%
+\subsubsection{Heart rate variability}\label{heart-rate-variability}}
+
+A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap
+(Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00)
+with ECG recordings (Giles et al., 2016), which are considered the gold
+standard for assessing HRV. In addition, HRV indices calculated from IBI
+and ECG data have shown a strong correlation (r = .99) in athletes
+(Caminal et al., 2018) and under spontaneous breathing conditions (Plews
+et al., 2017).
+
+\hypertarget{musculoskeletal-properties}{%
+\subsubsection{Musculoskeletal
+properties}\label{musculoskeletal-properties}}
+
+A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to
+measure passive muscle stiffness. The MyotonPRO is a non-invasive,
+handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms
+perpendicular to the surface of the skin. The impulse causes natural
+damped oscillations in the tissue, which are recorded by a three-axis
+digital accelerometer sensor in the device. The raw oscillation signal
+is then processed, and the stiffness parameter is calculated
+(Agyapong-Badu et al., 2016). The MyotonPRO has previously been reported
+to be a reliable and valid tool for the measurement of in-vivo tissue
+stiffness properties (Chuang et al., 2013; Pruyn et al., 2016; Nair et
+al., 2014), and has demonstrated good internal consistency (coefficient
+of variation \textless{} 1.4\%) over sets of 10 repetitions (Aird et
+al., 2012).
+
+\hypertarget{postural-stability}{%
+\subsubsection{Postural stability}\label{postural-stability}}
+
+Postural stability was assessed with a modified version of the balance
+error scoring system (mBESS) based on the protocol recommended by Hunt
+et al. (2009). In total, each trial of the mBESS was performed without
+shoes (McCrory et al., 2013) and included six stances in the following
+order; dominant leg (DL; standing on the dominant foot with the
+non-dominant foot at approximately 30-degrees of hip flexion and
+45-degrees of knee flexion), non-dominant leg (NDL; standing on the
+non-dominant foot with the dominant foot at approximately 30-degrees of
+hip flexion and 45-degrees of knee flexion) and tandem leg stance (TS;
+standing heel-to-toe with the non-dominant foot behind the dominant) on
+firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively
+(Fig \ref{fig:fig2}). To determine leg dominance, participants were
+asked their preferred leg to kick a ball to a target, and the chosen
+limb was labelled as dominant (cf.~Cingel et al., 2017). Participants
+were asked to maintain each stance for a total of 20-seconds.
+Participants hands were placed on hips at the level of the iliac crests.
+A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United
+Kingdom) was used to record each participants performance during the
+mBESS.
+
+The error identification criteria from the original BESS protocol was
+used by the lead researcher who scored all the BESS trials. One error
+was recorded if any of the following movements were observed during each
+trial: a) lifting hands off iliac crests; b) opening eyes; c) stepping,
+stumbling, or falling; d) moving the thigh into more than 30 degrees of
+flexion or abduction; e) lifting the forefoot or heel; and f) remaining
+out of the testing position for more than 5-seconds (Riemann et al.,
+1999). A maximum score of 10 errors was possible for each stance.
+Multiple errors occurring simultaneously were recorded as one error. A
+participant was given the maximum score of 10 if they remained out of
+the stance position for more than 5-seconds. A total score was
+calculated by summing the total number of errors recorded on all stances
+(DL, NLD and TS, on foam and firm surfaces). To assess the intra-rater
+reliability, a single measurement, absolute agreement, two-way mixed
+effects model for the intraclass correlation (ICC; Koo and Li, 2016) was
+used on a sample of 40 participants from the first time point. The
+test-retest scoring of BESS resulted in a ``good'' to ``excellent'' ICC
+score (ICC = 0.93, 95\% confidence interval = 0.88 - 0.96), indicating
+the scoring was reliable.
+
+\hypertarget{procedure}{%
+\subsection{Procedure}\label{procedure}}
+
+At the start of the academic year (September), coaches of sports teams
+at a British University and local sports clubs were contacted and
+informed about the study. With the coaches' permission, the lead
+researcher attended training sessions to inform athletes about the
+overall purpose of the study and the requirements of participation.
+Athletes who met the participation criteria and volunteered to take part
+in the study were invited to attend scheduled testing sessions. A
+repeated measures prospective cohort design was used to assess athletes'
+major life events, stress-related physiological markers and injury
+status over two consecutive 12-month periods. Each participant was asked
+to attend a total of three data collections over a 12-month period, with
+each data collection separated by a four-month interval (Fig
+\ref{fig:fig3}). Participants provided informed consent before data
+collection commenced.
+
+For the first three data collections (T1, T2 and T3), participants
+followed the same protocol in a specific order (Fig \ref{fig:fig4}). To
+ensure all measures could be collected within a viable time-frame,
+participants were separated into two groups. The first group completed
+all computer-based measures followed by all physical measurements,
+whereas the second group completed all physical measurements followed by
+computer-based measures. Participants were randomly assigned to one of
+the two groups and remained in those groups across all time points.
+
+\hypertarget{questionnaires}{%
+\subsubsection{Questionnaires}\label{questionnaires}}
+
+The questionnaires, which included demographic information, the LESCA,
+RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed
+on-line (SurveyMonkey Inc., USA, www.surveymonkey.com). The instructions
+for the LESCA were modified at T2 and T3 so that participants reported
+major life events that had occurred since the previous testing session.
+For injury reporting, participants were asked to record any injuries
+that they had sustained since the last data collection. The data were
+downloaded from surverymonkey.com and imported into R (R Core Team,
+2019) for analysis.
+
+\hypertarget{hrv}{%
+\subsubsection{HRV}\label{hrv}}
+
+To minimise potential distractions, participants were directed to a
+designated quiet area in the laboratory where IBI data were recorded.
+Participants were instructed to turn off their mobile devices to avoid
+any interference with the Bluetooth sensor. Each chest strap was
+dampened with water and adjusted so it fitted tightly but comfortably,
+as outlined by Polar's guidelines. Participants were seated and asked to
+remain as still as possible for the duration of the recording. No
+attempt was made to control participants' respiratory frequency or tidal
+volume (Denver et al., 2007). Inter-beat interval (IBI) data was
+collected for 10-minutes at a sampling frequency of 1000 Hz.
+
+Raw, unfiltered IBI recordings were exported from the Polar Flow web
+service as a space delimited .txt file and imported into R (R Core Team,
+2019) where the \emph{RHRV} package ({\textbf{???}}) was used to
+calculate HRV indices. Raw IBI data was filtered using an adaptive
+threshold filter, and the first 3-minutes and last 2-minutes of each
+recording were discarded, leaving a 5-minutes window that was used to
+calculate the root mean square of successive differences (RMSSD) in RR
+intervals following the recommendations for short term IBI recordings
+(Laborde et al., 2017; Malik et al., 1996). RMSSD was calculated as:
+
+\begin{equation} 
+  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
+\end{equation}
+
+Where N is the length of the time series, and \(RR_i\) the RR interval
+between beats \(i\) and \(i-1\), where each beat position corresponds to
+the beat detection instant.
+
+\hypertarget{muscle-stiffness}{%
+\subsubsection{Muscle stiffness}\label{muscle-stiffness}}
+
+To assess muscle stiffness, participants lay horizontally on a massage
+bed and four testing sites were identified on each lower limb. The
+muscle belly of the rectus femoris (RF), biceps femoris (BF), medial
+gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified
+using a visual-palpatory technique to determine the exact location of
+each site (Chuang et al., 2012). The visual-palpatory technique required
+the participant to contract the target muscle to aid the lead researcher
+to visually identify the muscle. The participant was then asked to relax
+the muscle and the muscle was palpated to locate the muscle belly. A
+skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was
+used to mark the testing site in the centre of the muscle belly.
+
+After the eight testing sites had been identified, the testing end of
+the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin
+on the testing site. A constant pre-load of 0.18 N was applied for
+initial compression of subcutaneous tissues. The device was programmed
+to deliver five consecutive impulses, separated by a one-second interval
+(Morgan et al., 2018). For each impulse, the device computed passive
+stiffness values, with the median of the five values being saved by the
+device for further analysis. In accordance with Myoton.com, a set of
+five measurements with a coefficient of variation (CV) of less than 3\%
+was accepted. Sets of measurements above 3\% were measured again to
+ensure data reliability. Measurements were uploaded using MyotonPRO
+software and imported in R (R Core Team, 2019) for further analysis. For
+each participant, the sum of all eight testing sites was calculated to
+provide a total lower extremity stiffness score and was used for further
+analysis
+
+\hypertarget{postural-stability-1}{%
+\subsubsection{Postural stability}\label{postural-stability-1}}
+
+Instructions for the mBESS were read to each participant and a
+demonstration of the positions was provided by the research assistant.
+For each position, participants were instructed to close their eyes,
+rest their hands on their iliac crests and remain as still as possible
+for 20-seconds. Participants were instructed to return to the testing
+position as quickly as possible if they lost their balance. The video
+recording was started prior to the first stance position and stopped
+after all stances had been completed. Each completed mBESS protocol took
+approximately four-minutes. Only one trial was performed to avoid
+familiarisation effects across the repeated measurement (cf.~Valovich et
+al., 2003). The video recordings for each participant were imported from
+the recording equipment (Sony DSC-RX10) and the lead researcher scored
+each trial using the error identification criteria.
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+Two Bayesian Networks (BN) were used to explore the relationships
+between the psychological measures, physiological markers of stress and
+sports injury. A BN is a graphical representation of a joint probability
+distribution among a set of random variables, and provides a statistical
+model describing the dependencies and conditional independences from
+empirical data in a visually appealing way (Scutari and Denis, 2014). A
+BN consists of arcs and nodes that together are formally known as a
+directed acyclic graph (DAG), where a node is termed a parent of a child
+if there is an arc directed from the former to the latter (Fig
+\ref{fig:fig5}; Pearl, 1988). However, the direction of the arc does not
+necessarily imply causation, and the relationship between variables are
+often described as probabilistic instead of casual (Scutari and Denis,
+2014). The information within a node can be either continuous or
+discrete, and a complete network can contain both continuous and
+discrete nodes; however, discrete networks are the most commonly used
+form of BN (Chen and Pollino, 2012). In discrete networks, conditional
+probabilities for each child node are allocated for each combination of
+the possible states in their parent nodes and can be used to assess the
+strength of a dependency in the network.
+
+In order to use discrete networks, continuous variables must first be
+split into categorical levels. When there are a large number of
+variables in the network, limiting the number of levels has the benefit
+of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100
+(10\^{}2) possible parameter combinations, however the same network with
+three levels has 1000 (10\^{}3) possible parameter combinations, the
+latter being significantly more computationally expensive. Using a
+larger number of splits in the data also comes at a cost of reducing the
+statistical power in detecting probabilistic associations, and reduces
+the precision of parameter estimates for the probabilistic associations
+that are detected because it reduces the sample-size-to-parameters ratio
+(Scutari and Denis, 2014). Typically, no more than three levels have
+been used in Bayesian networks in the sports injury literature
+(Olmedilla et al., 2018; Ruiz-Pérez et al., 2019)
+
+Learning the structure of the network is an important step in BN
+modelling. The structure of a network can be constructed using expert
+knowledge and/or data-driven algorithm techniques (e.g., search and
+score, such as hill climbing and gradient descent algorithms; Scutari \&
+Denis, 2014). The learned structure can then be used for inference by
+querying the network\footnote{The term ``query'' in relation to Bayesian
+  Networks stems from Pearl's expert systems theory (1988). A query can
+  be submitted to an expert (in this case, the network is the expert) to
+  get an opinion, the expert then updates the querier's beliefs
+  accordingly. Widely used texts on Bayesian Network analysis (Koller
+  and Friedman, 2009) have widely adopted the terminology in favour of
+  that used in traditional statistics.} and obtaining the posterior
+probabilities of a particular node for a given query. The posterior
+distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where
+the learned network \(B\) with structure \(G\) and parameters
+\(\Theta\), are investigated with new evidence \(E\) using the
+information in \(B\) (Scutari and Denis, 2014). In the example network
+presented in Fig \ref{fig:fig5}, new values assigned to each of the
+parent nodes (e.g., both set to ``Low'') could be used to investigate
+what effect the new information has on the state of the child node
+(conditional probability of a particular state of the child node). In a
+more complex network containing many nodes, the outcome of a particular
+node can be assessed conditional on the states of any subset of nodes in
+the network. BNs therefore provide a unique and versatile approach to
+modelling a set of variables to uncover dependency structures within the
+data.
+
+BNs have recently been used in the sport psychology literature
+(Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several
+benefits over traditional statistical analysis. For example, predictions
+can be made about any variable in the network, rather than there being a
+distinction between dependent and independent variables in the data,
+such as in linear regression models that are often used within the sport
+psychology literature (Olmedilla et al., 2018; Bittencourt et al.,
+2016). Furthermore, the structure of a network can be obtained from both
+empirical data \emph{and} prior knowledge about the area of study; the
+latter being particularly useful when there are a large number of
+variables in the network, or only a small number of observations are
+available in the data (Xiao-xuan et al., 2008). In such instances, a
+purely data driven approach to learning the network would be
+time-consuming due to the large parameter space, and inefficiency at
+identifying an approximation of the true network structure. Prior
+knowledge about dependencies between variables can therefore be included
+in the network structure, while still allowing a data driven approach
+for unknown dependencies, to improve the overall computation of the
+network structure (Heckerman et al., 1995; Xu et al., 2015). The
+following sections detail the steps taken in the current study to
+firstly prepare the data for each network, and then obtain the structure
+of each network that was used for further inference.
+
+\hypertarget{first-network}{%
+\subsubsection{First network}\label{first-network}}
+
+\hypertarget{data-preparation}{%
+\paragraph{Data Preparation}\label{data-preparation}}
+
+Of the 351 participants that were initially recruited for the study, 94
+only completed the first time point, and therefore had to be removed
+from the study as no injury information was obtained for these
+participants following the first time point. To prepare the data for the
+BN, missing values in the dataset were first imputed. Out of the 650
+total measurements across all time points in the current study, there
+were 31 (4.64\%) missing muscle stiffness measurements and 70 (10.48\%)
+missing heart rate recordings. The missing data were due to technical
+faults in the data collection equipment and were considered to be
+missing completely at random. A missing rate of 15-20\% has been
+reported to be common in psychological studies, and several techniques
+are available to handle missing values (Enders, 2003; Lang et al.,
+2013). In the current study, the \emph{caret} package (Kuhn et al.,
+2008) was used to impute the missing values. A bagged tree model using
+all of the non-missing data was first generated and then used to predict
+each missing value in the dataset. The bagged tree method is a reliable
+and accurate method for imputing missing values in data and is superior
+to other commonly used methods such a median imputation (Kuhn et al.,
+2008).
+
+A median split technique was used to discretise the data used in the
+network into ``Low'' and ``High'' levels. All variables apart from
+negative and total life events were approximately normally distributed
+(based on visual inspection, see supplementary material Sx) and required
+no further transformation prior to the median split. For the LESCA
+questionnaire data, a cumulative total of the current, and previous time
+points was calculated at each time point to account for the potential
+continuing effect of the life events experienced by athletes over time.
+Given the limited support for a relationship between positive life
+events and injury (Williams and Andersen, 2007), only negative and total
+life events were included in the network. Cumulative negative, and
+cumulative total life event scores at each time point were first log
+scaled so distributions were approximately normal, and then binarised
+using the median at each time point (nlelg and tlelg respectively). In
+addition to the log scaled cumulative values, an untransformed NLE score
+from the first time point was included as an additional variable based
+on previous literature that indicates this variable should have a strong
+relationship with injury outcome (Ivarsson et al., 2017).
+
+\hypertarget{network-structure}{%
+\paragraph{Network structure}\label{network-structure}}
+
+To obtain the network structure, several steps were taken to ensure that
+both a theoretically realistic network, and a network that was an
+appropriate fit to the collected data, was used for inference. Prior
+knowledge about the network structure was included by providing a list
+of arcs that are always \emph{restricted} from being in the network
+(blacklist), and a list of arcs that are always \emph{included} in the
+network (whitelist). Additionally, there are several scoring functions
+such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet
+equivalent uniform (BDeu) that can be used to compare network structures
+with certain nodes and arcs included or excluded (Scutari and Denis,
+2014). To account for the repeated measures design employed in this
+study and to maximise the use of the data, pairs of complete cases
+(e.g., participants who completed T1 + T2, and T2 + T3) were used in a
+two-time Bayesian network (2TBN) structure (Murphy, 2002). In the 2TBN,
+variables measured T2 could depend on variables measured at T1 (e.g., T1
+\(\rightarrow\) T2) and variables measured at T3 could depend on
+variables measured at T2 (e.g., T2 \(\rightarrow\) T3). However, arcs
+were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2
+to preserve the order in which data was collected. Variables were
+separated into two groups; ``explanatory'', for variables that did not
+change during the study (e.g., gender), or ``independent'', for
+variables that were measured at each time point and could vary during
+the study. Independent variable names were suffixed with \_1 for time
+point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2
+and T2\_1 \(\rightarrow\) T3\_2). Formatting the data in this way meant
+participants who completed T1 and T2, but did not complete T3, could
+still be included in the analysis. Table \ref{tab:table2} provides an
+example of the formatted data and demonstrates that participants 1 and 3
+have complete data, and therefore have two rows of data each
+representing variables from T1 \(\rightarrow\) T2 and T2 \(\rightarrow\)
+T3, respectively. Participant 2 did not complete the final data
+collection at T3 and therefore only has one row of data representing the
+variables collected at T1 and T2. In addition to the blacklisted arcs
+between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction
+of arcs was restricted between independent variables and explanatory
+variables (e.g., independent \(\rightarrow\) explanatory); however, arcs
+were not restricted between explanatory \(\rightarrow\) independent
+variables. Finally, arc direction was restricted between specific nodes
+within the explanatory variables. Arcs from clevel \(\rightarrow\)
+gender, nlebase \(\rightarrow\) gender and nlebase \(\rightarrow\)
+ind\_team were included in the blacklist, as arcs in these directions
+did not make logical sense. All subsequent models used the same
+blacklist.
+
+\hypertarget{preliminary-network-structures}{%
+\paragraph{Preliminary network
+structures}\label{preliminary-network-structures}}
+
+Prior to the final network structure presented in the results section,
+several structures with different combinations of variables were
+explored. Networks were learned using a Tabu search algorithm (Russell
+and Norvig, 2009) and BIC was used to compare different models. A higher
+BIC value indicates the structure of a DAG is a better fit to the
+observed data (Scutari and Denis, 2014). BIC values for each combination
+of variables of interest are reported as the combination of variables
+with the highest BIC value, followed by the relative scores of the other
+variables in the model.
+
+Initially, both negative life events and total life events were included
+in the network structure, however, the network score was improved when
+only nlelg or tlelg was included (highest BIC value = nleleg, BIC values
+relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -218.31).
+Additionally, despite strong evidence in the literature that both
+negative (NLE) and total life event (TLE) stress are related to injury
+occurrence (Williams and Andersen, 2007), network structures learned
+using the Tabu search algorithm failed to identify a relationship
+between NLE and injury or TLE and injury in the data. Given that nlelg
+provided the highest network score, and there is a stronger relationship
+between negative life events and injury in the literature, an arc was
+whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2
+in the final network structure. TLE score was not included in the final
+structure.
+
+The subscales representing the BAS (RR, RI, GDP and I) showed limited
+connection to other variables in the network. Therefore, several models
+were run with each scale individually to find the scale that resulted in
+the highest BIC value (values are shown relative to the highest value).
+RI provided the highest BIC value, compared to RR (-9.07), GDP (-15.06)
+and I (-16.33). Including all the variables (RR, RI, GDP and I) resulted
+in a significantly lower score -894.33) indicating that including all
+the variables was not beneficial to the model structure and did not
+offset the cost of the additional parameters. Therefore, only RI was
+included in the final structure.
+
+Finally, both total score and asymmetry for balance were included in the
+initial network. However, visual inspection of the network revealed no
+arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the
+network. Therefore, balance asymmetry was removed from the final network
+structure. To summarise, Table \ref{tab:table3} reports the variables
+that were included in the final network structure.
+
+Preliminary network structures also revealed strong dependencies between
+the same variables at sequential time points. For example, the
+probability that stiffness\_1 and stiffness\_2 were both ``High'', or
+both ``Low'' was approximately 80\%. Including the arcs between the same
+variables from X\_1 → X\_2 did not provide any theoretically meaningful
+information to the network structure as the majority of participants
+would either be in a ``Low'' or ``High'' state for each pair of
+variables in the network. Therefore, these arcs were blacklisted from
+the network. To obtain the final network, the appropriate blacklist and
+whitelists were provided and a Tabu search algorithm identified the
+remaining structure of the network. The final network structure was
+obtained by averaging 1000 bootstrapped models (Efron and Tibshirani,
+1994) to reduce the impact of locally optimal, but globally suboptimal
+network learning, and to obtain a more robust model (Olmedilla et al.,
+2018). Arcs that were present in at least 30\% of the models were
+included in the averaged model. The strength of each arc was determined
+by the percentage of models that the arc was included in, independent of
+the arc's direction. An arc strength of 1 indicated that the arc was
+always present in the network, with the value decreasing as arcs were
+found in fewer networks. In the respective study arcs above 0.5 were
+considered ``significant'' with arcs below 0.5 and above 0.3
+``non-significant'' (Scutari and Nagarajan, 2013). Arcs below 0.3 were
+not included in the model. The full table of arc strengths for the first
+and second network are available in S2 Table and S3 Table respectively.
+
+\hypertarget{network-inference}{%
+\paragraph{Network Inference}\label{network-inference}}
+
+Conditional probability queries (CPQ) were used to perform inference on
+both network structures. To conduct a CPQ, the joint probability
+distribution of the nodes was modified to include a new piece of
+evidence. The query allows the odds of a particular node state (e.g.,
+injured\_1 = ``injured'') to be calculated based on the new evidence.
+CPQs were performed using a likelihood weighting approach; a form of
+importance sampling where random observations are generated from the
+probability distribution in such a way that all observations match the
+evidence given in the query. The algorithm then re-weights each
+observation based on the evidence when computing the conditional
+probability for the query (Scutari and Denis, 2014). Inference was first
+performed on arcs that had a strength greater than 0.5 between the
+explanatory variables and independent variables and between different
+independent variables in the network. Of particular interest were the
+variables that were connected to ``injured'' nodes, which were examined
+in the network using the Markov blanket of ``injured\_1'' and
+``injured\_2''. A Markov blanket contains all the nodes that make the
+node of interest conditionally independent from the rest of the network
+(Fuster-Parra et al., 2017). CPQ's were used to determine what effect
+the variables in the Markov blanket of injured nodes had on the
+probability of the injured node being in the ``injured'' state.
+
+\hypertarget{second-network}{%
+\subsubsection{Second network}\label{second-network}}
+
+\hypertarget{data-preparation-1}{%
+\paragraph{Data preparation}\label{data-preparation-1}}
+
+Change scores between timepoint 1 and timepoint 2 were standardized to
+allow relative changes between variables to be compared.
+
+For the second network, change scores for continuous variables between
+T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3 were standardized to
+allow relative changes between variables to be compared. The ``injured''
+variable was also modified to represent whether a participant had
+sustained an injury at any point over the duration of the study or were
+healthy for the duration of the study. Setting the data up in this way
+enabled the construction of a network that explicitly modelled the
+\emph{amount} of change within variables between time points, as opposed
+to the first network that only captured changes when the median
+threshold was crossed from ``Low'' to ``High''.
+
+\hypertarget{network-structure-1}{%
+\paragraph{Network structure}\label{network-structure-1}}
+
+Similar to the first network, blacklists were used to prevent arcs from
+independent variables \(\rightarrow\) explanatory variables. In
+addition, ``nlebase'' variable was dropped from the list of explanatory
+variables to allow the \emph{changes} in negative life events to be the
+only life event variable in the network. The final network was obtained
+using the same approach as the first network.
+
+\hypertarget{network-inference-1}{%
+\paragraph{Network Inference}\label{network-inference-1}}
+
+Conditional probability queries (CPQ) were again used to perform
+inference the network structures. The Markov blanket of the ``injured''
+node was of particular interest, and the probability of injury was
+investigated with combinations of variables in the Markov blanket below
+the mean change, at the mean change and above the mean change. Initial
+visual inspection of the network structure also revealed arcs between
+BIS \(\rightarrow\) FFFS and HRV \(\rightarrow\) FFFS. To investigate
+this relationship further, random samples were generated for BIS, HRV
+and FFFS based on the conditional distribution of the nodes included as
+evidence in the query. The samples were then examined using a Bayesian
+linear regression models with the \emph{brms} package (Bürkner, 2017) to
+determine the relationship between these nodes. Weakly regularising
+priors (normal prior with mean of 0 and standard deviation of 5) were
+used for all parameters in the model.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+During the study, 26\% of participants reported at least one injury with
+an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days). Both
+male and female participants reported a greater number of acute compared
+to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39
+{[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and
+non-contact injuries were more common than contact injuries (male,
+non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female,
+non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}). Table
+\ref{tab:table4} shows the number and percentage of injury types
+sustained by both male and female participants.
+
+\begin{verbatim}
+## Note: Using an external vector in selections is ambiguous.
+## i Use `all_of(vars)` instead of `vars` to silence this message.
+## i See <https://tidyselect.r-lib.org/reference/faq-external-vector.html>.
+## This message is displayed once per session.
+\end{verbatim}
+
+\begin{verbatim}
+## Note: Using an external vector in selections is ambiguous.
+## i Use `all_of(outcome)` instead of `outcome` to silence this message.
+## i See <https://tidyselect.r-lib.org/reference/faq-external-vector.html>.
+## This message is displayed once per session.
+\end{verbatim}
+
+\hypertarget{first-network-structure}{%
+\subsection{First network structure}\label{first-network-structure}}
+
+The first network structure obtained from the data is shown in Fig
+\ref{fig:fig6}. Several of the explanatory variables showed strong
+connections with independent variables in the network. The arc from
+nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the
+probability of RI\_1 being in the ``High'' state increased from 0.23 to
+0.47 when nlebase increased from ``Low'' to ``High''. The ind\_team node
+had strong arcs to hours (0.90) and nlebase (0.84). Individual athletes
+were more likely to have ``High'' hours per week (0.84) compared to
+team-based athletes (0.60). Individual athletes were also more likely to
+have ``High'' negative life events in the 12 months preceding the start
+of the study compared to team based athletes (individual athletes =
+0.65, team-based athletes = 0.41). The arcs from gender \(\rightarrow\)
+stiffness\_1 and gender \(\rightarrow\) stiffness\_2 were 0.76, and 0.65
+respectively, with males more likely to have ``High'' stiffness compared
+to females (males = 0.62, females = 0.43). The arc from pi
+\(\rightarrow\) stiffness\_1 was 0.55 with athletes who reported an
+injury in the preceding 12 months more likely to have ``High'' (0.65)
+compared to ``Low'' (0.35) stiffness. The arc from clevel
+\(\rightarrow\) balance\_1 had a strength of 0.51, with lower level
+performers more likely to have decreased balance ability (0.48),
+compared to national level athletes (0.29). Arcs were also present
+between independent variables in the network. Strong arcs were present
+between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\)
+FFFS\_2 (0.68). In both instances, ``High'' FFFS was more likely when
+BIS was ``High'' (0.65 for \_1, 0.61 for \_2) compared to ``Low'' (0.32
+for \_1, 0.36 for \_2). The arc between nlelg \(\rightarrow\) BIS had a
+strength of 0.62 for nlelg\_1 \(\rightarrow\) BIS\_1, however no arc was
+present between nlelg\_2 and BIS\_2. For nlelg\_1 \(\rightarrow\)
+BIS\_1, ``Low'' negative life events increased the probability of BIS
+being in the ``High'' state from 0.33 to 0.55.
+
+\hypertarget{markov-blanket-for-injured_1}{%
+\subsubsection{Markov blanket for
+injured\_1}\label{markov-blanket-for-injured_1}}
+
+The Markov blanket for injured\_1, which contained hours spent training
+per week (hours), negative life events (nlelg\_1), muscle stiffness
+(stiffness\_1), current competitive level (clevel) and balance
+(balance\_1), is shown in Fig \ref{fig:fig7}. The arc between nlelg\_1
+and injured\_1 was fixed in the network, so has the maximum strength of
+1.
+
+The CPQ for injured\_1 in the ``injured'' state for all variables that
+were directly linked to injured\_1 is shown in Table \ref{tab:table5}.
+The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when
+stiffness was ``High'' compared to ``Low''. Negative life events had a
+negligible effect when moving from the ``Low'' to ``High'' state.
+
+The second CPQ investigated the outcome of injured\_1 being ``injured''
+conditional on all variables in the Markov blanket. The Markov blanket
+contained five nodes, each with two possible states resulting in \(2^5\)
+combinations of variables, therefore only the three lowest and highest
+probabilities are shown in Table \ref{tab:table6} (complete results in
+S4 Table). The combination of lower competitive level, ``High'' hours
+per week, ``Low'' negative life events, ``High'' balance and ``High''
+stiffness resulted in a probability of 0.53 for injured\_1 being in the
+``injured'' state. When all variables were in the ``Low'' state the
+probability of ``injured'' was approximately 0.04 Negative life events
+only had a substantial effect on injured\_1 when all other variable were
+fixed to ``Low''. In this instance the probability of injured\_1 being
+``injured'' rose from 0.04 to 0.19, when negative life events was in the
+``Low'' and ``High'' states respectively.
+
+Table \ref{tab:table7} shows the states of all explanatory and
+independent variables suffixed with \_1 that resulted in the highest and
+lowest probability of injured\_1 being in the ``injured'' state. The
+probability of injured\_1 being ``injured'' rose from 0.03 to 0.54 with
+the combination of variable states in Table \ref{tab:table7}. Notably,
+hours per week, stiffness\_1 and balance\_1 had the greatest effect on
+raising the probability of injured\_1 being in the ``injured'' state.
+
+\hypertarget{markov-blanket-of-injured_2}{%
+\subsubsection{Markov blanket of
+injured\_2}\label{markov-blanket-of-injured_2}}
+
+The Markov blanket for injured\_2 is shown in Fig \ref{fig:fig8} and
+contained gender, FFFS\_1, stiffness\_2, balance\_2 and rmssd\_2. The
+arc between stiffness\_2 and injured\_2 was comparable to the arc
+between stiffness\_1 \(\rightarrow\) injured\_1. Very weak arcs (0.3)
+between injured\_2 \(\rightarrow\) balance\_2 and injured\_2
+\(\rightarrow\) rmssd\_2 were also present in the Markov blanket for
+injured\_2. Results of the first query for injured\_2 in the ``injured''
+state are presented in table \ref{tab:table8}. Similar to injured\_1,
+stiffness\_2 doubled the probability of injured\_2 being ``injured''
+from 0.13 in the ``Low'' state to 0.27 in the ``High'' state. FFFS\_1 in
+the ``Low'' state increased probability of injured\_2 being ``injured''
+by 0.19 compared to the ``High'' state. ``High'' negative life events
+decreased the probability of injury from 0.24 to 0.19.
+
+The conditional probabilities based on all the variables in injured\_2
+Markov blanket are presented in Table \ref{tab:table9}. Again, only the
+three lowest and highest probabilities are shown (complete results in S5
+Table). The combination ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High''
+balance resulted in the greatest probability of injured\_2 being
+``injured'', with the highest probability of injury being 0.52. With all
+other variables held in the ``High'' state, the probability of
+injured\_2 being ``injured'' rose from 0.14 to 0.34 when FFFS\_1 was in
+the ``Low'' compared to ``High'' state. The combination of ``Low''
+stiffness, ``Low'' balance and ``High'' FFFS resulted in the lowest
+probability of injured\_2 being ``injured''.
+
+\hypertarget{second-network-structure---changes-within-variables}{%
+\subsection{Second network structure - changes within
+variables}\label{second-network-structure---changes-within-variables}}
+
+The network for changes within variables is presented in Fig
+\ref{fig:fig9}. An arc between BIS \(\rightarrow\) FFFS with strength 1
+was present in the network. Arcs between clevel \(\rightarrow\) BIS and
+gender \(\rightarrow\) stiffness had a strength of 0.60. The arc between
+RMSSD \(\rightarrow\) FFFS was 0.56. The arcs between BIS
+\(\rightarrow\) FFFS and RMSSD \(\rightarrow\) FFFS were examined
+further by drawing random observations from the conditional probability
+distribution and examining the relationship in a linear regression
+model. While the continuous nodes in the second network are already
+parameterised using linear regressions, it was not possible to
+investigate the interaction between explanatory variables. The use of a
+separate linear regression enabled the interaction between BIS and RMSSD
+to be examined.
+
+Results from the Bayesian linear regression model are presented in Table
+\ref{tab:table10} and include 95\% credible intervals (CrI). Increases
+in BIS were associated with increases in FFFS (\(b\) = -0.14, 95\% CrI =
+{[}-0.20, -0.08{]}), whereas positive changes in RMSSD where associated
+with decreased changes in FFFS (\(b\) = 0.4, 95\% CrI = {[}0.34,
+0.45{]}). There was no clear interaction between RMSSD and BIS (\(b\) =
+-0.04, 95\% CrI = {[}-0.10, 0.02{]}).
+
+The Markov blanket for the ``injured'' node contained previous injury,
+gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}). For
+stiffness and nlec the values in the nodes represent the standardised
+change between time point. Combinations of nlec and stiffness at one SD
+below the mean change, at the mean change, and 1 SD above the mean
+change are presented in table \ref{tab:table11}. Changes in both nlec
+and stiffness 1SD above the mean change resulted in a probability of
+being injured of 0.71 over the duration of the study. With stiffness
+held at the mean change, the probability of ``injured'' rose from 0.35
+to 0.64 with nlec at 1 SD below an 1 SD above respectively.
+
+Table \ref{tab:table12} shows the three highest and lowest probabilities
+for injury for all variables in the Markov blanket. The combination of 1
+SD above the mean change for nlec and stiffness and ``High'' hours per
+week and previous injury resulted in the highest probability that an
+injury would be sustained during the study (0.77). In contrast, below
+average changes in nlec and stiffness combined with ``Low'' hours per
+week and no previous injury resulted in the lowest probability of an
+injury (0.12).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Informed by Appaneal and Perna's (2014) extension to the widely cited
+Williams and Anderson's (1998) stress-injury model, the purpose of this
+study was to examine how psychosocial stress-related factors, and
+physiological stress-related markers may interact and act
+synergistically to increase the risk of injury; therein addressing
+several of the limitations of the sport injury literature (Johnson et
+al., 2014). Potential relationships between psychosocial stress-related
+factors, physiological stress-related markers of stress and injury were
+investigated using two BN structures; the first was a two-time Bayesian
+Network that investigated the relationships between variables
+\emph{across} time points in the study (Fig \ref{fig:fig5}), and the
+second network used differential equations to model the changes in
+variables \emph{between} time points (Fig \ref{fig:fig8}).
+
+The first network revealed several links between the injured nodes and
+other variables in the network. For example, Fig \ref{fig:fig6} and Fig
+\ref{fig:fig7} show the Markov blankets for the injured\_1 and
+injured\_2 nodes in the first network and include all the variables that
+had a direct effect on the probability of injury. The combination of
+high stiffness and poor balance resulted in the highest probability of
+injury in the Markov blankets for injured\_1 and injured\_2. The
+presence of these factors at both injured nodes indicates that the
+combined action of these variables is important for determining an
+athlete's risk of injury. In the second network, the highest probability
+of injury was observed when changes in stiffness \emph{and} negative
+life events were both greater than average (Table \ref{tab:table11}),
+indicating that the combination of changes in psychological and
+physiological measures of stress may combine additively to increase the
+risk of injury (Appaneal and Perna, 2014). Table \ref{tab:table7} shows
+the combinations of variables that resulted in the highest and lowest
+probabilities of injury conditional on all explanatory variables and
+variables suffixed with \_1. Although only the highest and lowest values
+are shown, there were over 16000 possible combinations of variables
+states that could influence the risk of injury. The adopted approach
+attempts to take the complex systems view of injury recommended by
+Bittencourt et al.~(Bittencourt et al., 2016), where injury is an
+emergent property from a web of determinants that interact in different
+ways.
+
+Of all the variables measured in the study, muscle stiffness appeared to
+be most strongly related to injury. Both ``High'' levels of stiffness in
+the first network, and greater than average increases in stiffness in
+the second network were found to increase the risk of injury. In this
+study, a novel hand-held device (MyotonPRO) was used to measure muscle
+stiffness. To date, one of only a small number of studies that have used
+the MyotonPRO to explore the relationship between muscle stiffness and
+sports injury, found that increased muscle stiffness in the soleus and
+Achilles tendon was related to increased injury incidence in elite level
+netball players (Pickering-Rodriguez et al., 2017). The results of the
+current study build on these findings, with a larger sample of athletes
+from a range of different sports, strengthening the evidence for a
+relationship between higher levels of muscle stiffness and injury.
+However, high levels of muscle stiffness, as measured by the MyotonPRO,
+have also been found to be related to improved performance, with elite
+level athletes having increased lower extremity stiffness (Pruyn et al.,
+2015; Kalkhoven and Watsford, 2017). Collectively, therefore, these
+findings suggest that while muscle stiffness plays a vital role in
+performance, increased levels of stiffness also increase the risk of
+injury, with each athlete likely to have an optimum level of stiffness
+that maximises performance while minimising the risk of injury (Butler
+et al., 2003). Additionally, high stiffness may only increase the risk
+of injury if other factors are also present. To elaborate, the
+combination of high stiffness \emph{and} poor balance was found to
+result in the greatest probability of injury. In contrast, athletes with
+high stiffness and good balance were less likely to be injured,
+suggesting that improved postural stability may counteract the potential
+harmful effects of high levels of muscle stiffness. Several studies have
+identified how balance (Romero-Franco et al., 2014; Trojian and McKeag,
+2006) and muscle stiffness (Butler et al., 2003; Pickering-Rodriguez et
+al., 2017) are related to injury individually; however this study has
+demonstrated how these two factors may have an additive effect in
+relation to injury occurrence.
+
+In addition to stiffness, balance is also linked to injury at both
+injured nodes in the first network, however the strength of the arc was
+only 0.35 and 0.30 from balance \(\rightarrow\) injured\_1 and balance
+\(\rightarrow\) injured\_2 respectively. Despite the weak arc strength,
+a ``High'' balance score, indicating impaired postural stability, was
+found to increase the probability of injury. This finding is consistent
+with previous research that has reported an association between
+decrements in postural stability and increased injury risk (Riemann et
+al., 1999; Romero-Franco et al., 2014; Trojian and McKeag, 2006).
+Postural stability is often used as an indicator of athlete performance
+level, with higher level athletes demonstrating better postural
+stability over their lower level counterparts (Paillard et al., 2006).
+In the current study, athletes who competed at a higher level were also
+more likely to have good balance (``Low'' balance), compared to their
+lower level counterparts. These findings suggest that better postural
+stability is associated with both a higher level of performance and a
+lower probability of sustaining an injury, reinforcing the importance of
+postural stability as a feature of athletic training programmes designed
+to prepare athletes for the demands of high intensity training and
+competition (Hrysomallis, 2011).
+
+Of the psychological variables in the study, negative life events have
+previously been reported to be most strongly associated with injury
+(Williams and Andersen, 2007; Ivarsson et al., 2017). In the current
+study, the second network revealed that greater than average increases
+in negative life event stress increased the probability of being injured
+during the study period. However, negative life event stress had almost
+no effect on the probability of injury in the first network. This
+finding suggests that the \emph{relative change} in life events may be
+more important than the \emph{absolute score} for life events, despite
+the latter being commonly used in sports injury research to date. For
+example, an athlete who reports a negative life event score of 1 during
+the first time point, but then a score of 5 at the second time point
+will have a 400\% increase in their life event score. Although the
+absolute score would be ``Low'', the relative increase could have been
+caused by a significant event in the athlete's life, that could have a
+considerable psychological and physiological effect (Appaneal and Perna,
+2014). Future research should therefore consider study designs and
+appropriate analysis methods that enable relative changes in an
+individual athlete's life events to be assessed (cf.~Ivarsson et al.,
+2014).
+
+The finding that negative life events had almost no impact on the
+probability of injury in the first network was surprising given that the
+majority of research has consistently identified major life events,
+particularly those events with a \emph{negative} valence, as the
+strongest predictor of injury in Williams and Andersen's (Williams and
+Andersen, 1998) model (Ivarsson et al., 2017). During the initial
+network structure development, no arcs between the negative life event
+nodes and injured nodes were found by the Tabu search algorithm.
+However, given the strength of the literature indicating that negative
+life events are related to injury, an arc was fixed between nlelg\_1
+\(\rightarrow\) injured\_1 and nlelg\_2 \(\rightarrow\) injured\_2 to
+allow this relationship to be examined more closely. When negative life
+events were ``High'' the probability of injury showed a negligible
+change at the injured\_1 node and decreased by -0.05 at the injured\_2
+node. One possible explanation for these findings may be due to the use
+of the LESCA questionnaire in a repeated measures design. In the
+original LESCA instructions, participants are asked to report major life
+events that have occurred over the previous 12-months (Petrie, 1992).
+However, in the current study, participants completed the LESCA at
+approximately 4-month intervals after baseline and were therefore asked
+to report any events which had occurred \emph{since the previous data
+collection session}, to avoid inflated scores caused by reporting the
+same event on multiple occasions. The reduced 4-month time interval
+between data collections may have reduced the likelihood for life events
+listed in the LESCA to have taken place. For example, at the second and
+third time points, 26\% of participants reported 0 negative life events
+for the preceding four-month period. Simply, it may be that the items on
+the LESCA are less suitable for repeated measurements with durations
+shorter than the original 12-months than a measure that captures minor
+life events (cf.~Fawkner et al., 1999).
+
+Another possible explanation for the findings for major/negative life
+events is that participants in the study may have had access to the
+necessary coping resources to mitigate against the effects of any
+negative life event stress they experienced. Williams and Andersen's
+(Williams and Andersen, 1998) model proposed a number of coping
+resources that were either directly related to injury or moderated the
+relationship between life stress and injury occurrence; for example,
+general coping strategies (e.g., good sleeping habits and self-care),
+social support systems and stress management skills. Although coping was
+not measured in the current study, several studies have found high
+levels of social support can reduce the risk of injury (Petrie, 1993;
+Petrie et al., 2014; Johnson et al., 2014). Therefore, future research
+should consider including a measure of coping alongside that of life
+event stress to help explain the possible moderating effect.
+
+Of the remaining variables, both FFFS and RMSSD were also linked to
+injury. A weak arc was observed between RMSSD\_2 \(\rightarrow\)
+injured\_2 (arc strength = 0.30), however no arc was present between
+RMSSD\_1 and injured\_1, suggesting the link between RMSSD and injury
+was weaker than muscle stiffness and balance, where stronger arcs
+existed at both of the injured nodes. Despite the uncertainty regarding
+the relationship between injury and RMSSD in the first network, ``Low''
+RMSSD increased the probability of injury from 0.17 (RMSSD = ``High'\,')
+to 0.27 (RMSSD =''Low'\,'). This finding is consistent with previous
+research that has found reduced RMSSD to be indicative of illness or
+maladaptation to training due to decreased parasymapthic activity, which
+often precedes injury (Williams et al., 2017; Bellenger et al., 2016;
+Gisselman et al., 2016). An arc between FFFS\_1 and injured\_2 (arc
+strength = 0.40) was also observed in the first network, where the risk
+of injury was increased from 0.13 to 0.29 with FFFS in the ``High'' and
+``Low'' states respectively. Interestingly, the ``Low'' FFFS score was
+also related to injuries at subsequent time points; for example, the
+``Low'' FFFS recorded at the first time point was related to injuries
+recorded at the second time point in the study. One possible explanation
+for this finding could be that those athletes who reported ``Low'' FFFS
+score were less fearful, and may therefore engage in more risk taking
+behaviours, increasing the probability of injury. The RST theory
+proposes that higher levels of FFFS increase avoidance motivation (Corr
+et al., 2016), and therefore ``High'' FFFS may have acted as a deterrent
+from taking risks while training and competing, reducing exposure to
+situations that could have resulted in injury.
+
+In the first network, ``High'' BIS was associated with ``High'' FFFS,
+while in the second network, increases in BIS were associated with
+increases in FFFS. RST proposes that the combination of high BIS and
+high FFFS is likely to result in a more anxious disposition due to high
+levels of avoidance and high goal conflict characterised by high levels
+of FFFS and BIS (Corr, 2013). High levels of anxiety and anticipation of
+stressful situations have been associated with reductions in HRV indices
+including RMSSD (Chalmers et al., 2014; Pulopulos et al., 2018), which
+is supported by the negative relationship between FFFS and RMSSD in the
+second network (Table \ref{tab:table10}). These findings align with the
+proposed actions of the RST theory (Corr et al., 2016). For example,
+high levels of BIS are proposed to be the result of goal conflict
+between the FFFS (avoidance) and BAS (approach) systems. The goal
+conflict is likely to elicit a physiological response (e.g., decreased
+HRV) in preparation to engage in the required behaviour to resolve the
+goal conflict (Corr et al., 2016). To extend these findings, the BAS
+should also be considered. Specifically, to establish how the BAS and
+FFFS interact, and how these two systems affect the BIS. However, in the
+current study, initial network structures revealed the BAS sub-scales to
+have limited connectivity with other measures in the network, therefore
+only one of the BAS sub-scales (RI) was included in the final network
+structure. In the first network, RI\_1 was connected to both BIS\_1 and
+BIS\_2, and in both instances, the probability of ``High'' BIS was
+increased when RI was also ``High''. However, the arcs between RI\_1 and
+BIS were weak (\textless{} 0.50), and RI represents only one component
+of the BAS system. Other BAS factors such as impulsivity may be more
+closely related to risk-taking behaviours and may reveal additional
+links to sports injury. Therefore, a more detailed examination of the
+different elements of RST, and specifically the BAS in relation to
+injury occurrence is warranted.
+
+The current research had several strengths, including the longitudinal
+repeated measures design and modelling approach. A major critique of the
+sport injury literature has been the use of only one wave of measurement
+that may not be reflective of and capture the dynamic nature of the
+variables that are associated with injury (Johnson et al., 2014). The
+longitudinal repeated measures design of the current study allowed
+\emph{changes} over time and between time points to be captured and
+explored. Another significant strength of the current research was the
+interdisciplinary approach, which enabled an examination of the complex
+interplay between psychosocial and physiological markers of stress.
+Although there are unique and significant challenges with research
+employing longitudinal repeated measures designs, they provide for far
+more fine-grained understanding of the dynamic relationships between
+stress-related factors and injury occurrence in athletes. Sport injury
+research has been criticised for adopting analytic approaches that are
+reductionist in nature (Bittencourt et al., 2016) that fail to account
+for the complex, emergent behaviour that is characteristic of injury
+occurrence. To address this issue, Bayesian networks (BN) were used to
+more closely align with the complex, multifactorial nature of injury.
+The networks allowed several markers of stress that were free to
+interact with each other, as well as injury, to be explored.
+Consequently, BN's provided a contemporary approach that improved upon
+traditional methods such as logistic regression (Olmedilla et al.,
+2018).
+
+As with most research, there were several limitations with the present
+study. Firstly, the choice was made to binarise variables in the first
+network so only ``Low'' and ``High'' states were observed. Although
+binarising variables is a common procedure in Bayesian network analysis
+and has several advantages, Qian and Miltner Qian and Miltner (2015)
+highlighted that both a loss of statistical accuracy and potential
+difficulty in subsequent interpretation of the model may arise when
+following a binarising procedure. For example, the meaning of a ``Low''
+and ``High'' value in the current study is only meaningful for the
+population that was studied, and there could be additional levels within
+each category that were not investigated. A second limitation was the
+nature of the physiological measures used in the current study. In order
+to collect data on a large sample of participants, suitable measures
+were required to ensure the viability of the data collection; however,
+some of these measures may not have been sensitive enough to detect more
+subtle variation in athletes. For example, postural stability could have
+been assessed with the use of a force plate, which is considered the
+gold standard, to provide detailed data and enable a more fine-grained
+analysis (Ross et al., 2011).
+
+In addition to the future directions already outlined, the findings from
+the current study offer several avenues for future research. Although
+the current study used a range of measures to capture ``stress'' from
+both a psychological and physiological perspective, there may be
+additional measures available that could provide further insight into
+the relationship between stress-related factors and sports injury. For
+example, stress hormones such as cortisol have been found to be a marker
+of both psychological and training-related stress (Appaneal and Perna,
+2014; Perna and McDowell, 1995), and could help elucidate the
+relationship between stress and injury. Additionally, although the LESCA
+is the most widely used measures of major life events in sports injury
+research, the current study found several limitations with using the
+LESCA in a repeated measures design, including how the items were
+scored. For example, there is no way to differentiate between an athlete
+who has answered four items as moderately negative, and one item as
+extremely negative. Both responses would be scored a ``-4''; however,
+there could be vastly different psychological and physiological effects
+between moderately negative and extremely negative events. Therefore,
+future research could develop a modified version of the LESCA that could
+distinguish between these types of responses and their effects.
+
+In conclusion, the purpose of this research was to explore the
+multifaceted nature of the stress-injury relationship, and several
+psychosocial and physiological markers were found to combine and
+exacerbate the risk of injury. Specifically, muscle stiffness and
+\emph{increases} in negative life event stress were identified as strong
+predictors of injury, while other factors including personality
+characteristics and postural stability were also found to contribute to
+the probability of injury occurrence. Taken together, the
+interdisciplinary approach coupled with the advanced analytical
+techniques used and complex systems framework has provided a novel
+examination of the stress-injury relationship that has addressed many of
+the limitations identified in previous research.
+
+\hypertarget{disclosureconflict-of-interest-statement}{%
+\section{Disclosure/Conflict-of-Interest
+Statement}\label{disclosureconflict-of-interest-statement}}
+
+The authors declare that the research was conducted in the absence of
+any commercial or financial relationships that could be construed as a
+potential conflict of interest.
+
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+The statement about the authors and contributors can be up to several
+sentences long, describing the tasks of individual authors referred to
+by their initials and should be included at the end of the manuscript
+before the References section.
+
+\hypertarget{supplemental-data}{%
+\section{Supplemental Data}\label{supplemental-data}}
+
+Supplementary Material should be uploaded separately on submission, if
+there are Supplementary Figures, please include the caption in the same
+file as the figure. LaTeX Supplementary Material templates can be found
+in the Frontiers LaTeX folder
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+A reference list should be automatically created here. However it won't.
+Pandoc will place the list of references at the end of the document
+instead. There are no convenient solution for now to force Pandoc to do
+otherwise. The easiest way to get around this problem is to edit the
+LaTeX file created by Pandoc before compiling it again using the
+traditional LaTeX commands.
+
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
+
 \begin{figure}[!h]
@@
 
-Williams and Andersen's (Williams and Andersen, 1998) model proposed
-that when faced with a potentially stressful athletic situation, an
-athlete's personality traits (e.g., hardiness, locus of control and
-competitive trait anxiety), history of stressors (e.g., major life
-events and previous injuries) and coping resources (e.g., general coping
-behaviours) will contribute to their response, either interactively or
-in isolation. Central to the model is the stress response, which
-reflects the bi-directional relationship between athletes' appraisal of,
-and response to, a stressful athletic situation. The model predicts that
-athletes who have a history of many stressors, personality traits that
-intensify the stress response and few coping resources, will exhibit
-greater attentional (e.g., peripheral narrowing) and/or physiological
-(e.g., increased muscle tension) responses that put these individuals at
-greater risk of injury.
-
-Within Williams and Andersen's (Williams and Andersen, 1998) model,
-major life events, a component of an athlete's history of stressors,
-most consistently predicts injury occurrence (Williams and Andersen,
-2007); specifically, major life events with a negative, as opposed to
-positive, valence (Passer and Seese, 1983; Maddison and Prapavessis,
-2005). However, personality traits and coping resources have also been
-found to predict injury, with for example, athletes more likely to
-sustain an injury if they have poor social support and psychological
-coping skills, and high trait anxiety and elevated competitive state
-anxiety; compared to athletes with the opposite profile. (Smith et al.,
-1990; Lavallée and Flint, 1996; Ivarsson and Johnson, 2010). However,
-the amount of variance explained by the psychosocial factors proposed by
-the model has been modest, typically between 5 - 30\% (Galambos et al.,
-2005; Ivarsson and Johnson, 2010); suggesting other factors are also
-likely to contribute to injury occurrence.
-
-While the psychosocial factors proposed in Williams and Andersen's
-(Williams and Andersen, 1998) model have received the most research
-attention, the mechanisms through which these factors are proposed to
-exert their effect have remained under-investigated in the literature.
-To elaborate, the model suggests that injuries are likely to occur
-through either increased physiological arousal resulting in increased
-muscle tension and reduced flexibility or attentional deficits caused by
-increased distractibility and peripheral narrowing. However, to date,
-the research has largely focused on attentional deficits (Andersen and
-Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et
-al., 2007). For example, Andersen and Williams (Andersen and Williams,
-1999) measured peripheral and central vision during high and low stress
-conditions and found athletes with high life event stress coupled with
-low social support had greater peripheral narrowing under stressful
-conditions compared to athletes with the opposing profile; these
-athletes went on to sustain an increased number of injuries during the
-following season. Indeed, Rodgers and Landers (Rogers and Landers, 2005)
-supported Andersen and Williams's (Andersen and Williams, 1999) earlier
-findings reporting that peripheral narrowing under stress mediated 8.1\%
-of the relationship between negative life events and injury. However,
-the remaining variance between negative life events and athletic injury
-through the other proposed mechanisms, such as increased muscle tension
-and reduced motor control, remains to be explored (cf.~Williams and
-Andersen, 1998).
-
-One challenge faced by researchers addressing the sports injury problem
-with a psychological lens is the multifactorial nature of injury, and
-the possible contribution of other non-psychological factors to the
-stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009). For
-example, a large body of research indicates that training-related stress
-is also likely to be related to the stress response and injury
-occurrence (Lee et al., 2017; Djaoui et al., 2017), and may account for
-the unexplained variance from the psychological predictors of injury.
-Appaneal and Perna (Appaneal and Perna, 2014) proposed the
-biopsychosocial model of stress athletic injury and health (BMSAIH) to
-serve as an extension to Williams and Andersen's (Williams and Andersen,
-1998) model and to address some of these issues. To elaborate, the
-BMSAIH aimed to clarify the mediating pathways between the stress
-response and injury, consider other health outcomes and behavioural
-factors that impact sports participation, and integrate the impact of
-training on athletes' health (Appaneal and Perna, 2014). The central
-tenet of the BMSAIH is that psychosocial distress (e.g., negative life
-events) may act synergistically with training-related stress as a result
-of high-intensity and high-volume sports training, and ``widen the
-window of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of
-undesirable health outcomes including illness and injury. Consequently,
-the BMSAIH provides a framework for future research to build on Williams
-and Andersen's (Williams and Andersen, 1998) model, by including other
-physiological markers of training-related stress, which together may
-provide greater insight into the injury process.
-
-Although research supporting the BMSAIH has mainly focused on the
-relationship between hormonal responses to training and injury
-occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other
-research has identified additional markers of training-related stress
-that are associated with an increased risk of injury; for example, heart
-rate variability (Bellenger et al., 2016; Williams et al., 2017),
-postural stability (Romero-Franco et al., 2014) and muscle stiffness
-(Pruyn et al., 2015). However, these markers are often studied in
-isolation without an assessment of the psychosocial factors that are
-known to contribute to injury, thereby limiting our understanding of how
-psychosocially and physiologically derived stress may contribute
-synergistically to injury occurrence. Recently, Bittencourt et
-al.~(Bittencourt et al., 2016) suggested that to better understand the
-multifactorial nature of sports injuries, research needs to move away
-from studying risk factors in isolation and instead adopt a complex
-systems approach to injury. Such an approach posits that injury may
-arise from a complex ``web of determinants'' (Bittencourt et al., 2016,
-3), where different factors interact in unpredictable and unplanned
-ways, but result in a global outcome pattern of either adaptation or
-injury.
-
-A challenge when adopting a complex systems approach is using an
-appropriate analysis technique that is able to capture the uncertainty
-and complexity of the relationships between different factors. One
-technique that provides a solution to this challenge is Bayesian network
-(BN) modelling. BN's allow the construction of graphical probabilistic
-models using the underlying structure that connects different variables
-(nodes in the network) of interest (Scutari and Denis, 2014). Once
-constructed, the learned structure can be used for inference by
-obtaining the posterior probabilities of a particular node for a given
-query (e.g., if the value of Node A is x and the value of Node B is y,
-what is the probability Node C being value z?). Furthermore, BN's do not
-distinguish between dependent and independent variables as they are a
-form of unsupervised learning, which is a strength over regression or
-structure equation models when the underlying relationship in the
-network may not be known (Olmedilla et al., 2018).
-
-To summarise, despite offering a possible framework to build on the
-research stemming from Williams and Andersen's (Williams and Andersen,
-1998) model, there remains and opportunity to explore other
-physiological stress-related markers proposed by the BMSAIH, in addition
-to the already well-established psychological characteristics known to
-be related to injury (Appaneal and Perna, 2014). Furthermore, research
-has typically not captured changes in both psychosocial factors and
-stress-related physiological markers that may occur between initial
-measurement and injury occurrence. Given the exploratory the current
-study had the following objectives:
-
-\begin{itemize}
-\tightlist
-\item
-  Identify suitable markers of stress that can be easily captured in a
-  large cohort of athletes in a field based setting.
-\item
-  Examine the relationships between the markers of stress and injury
-  using a prospective, repeated measures design.
-\item
-  Evaluate the relationships between the markers of stress and injury
-  using Bayesian network modelling that captures the complex nature of
-  injury occurrence.
-\end{itemize}
+\begin{figure}[!h]
+\caption{{\bf mBESS positions (A-F).}
+Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
+\label{fig:fig2}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Study design.}
+For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
+\label{fig:fig3}
+\end{figure}
+
+\begin{figure}[!h]
+\caption{{\bf Session protocol.}
+Outline of the protocol for each data collection.}
+\label{fig:fig4}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Example network.}
+A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
+\label{fig:fig5}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Network structure.}
+The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+\label{fig:fig6}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket of injured\_1.}
+Arc strengths are included as arc labels.}
+\label{fig:fig7}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for injured\_2.} }
+\label{fig:fig8}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Change network.}
+Network structure of the changes within variables between time points.}
+\label{fig:fig9}
+\end{figure}
+
+\begin{figure}[H]
+\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
+\label{fig:fig10}
+\end{figure}
+
+\hypertarget{tables}{%
+\section{Tables}\label{tables}}
+
+\begin{table}[H]
+
+\caption{\label{tab:table1}Participant characteristics.}
+\centering
+\begin{tabular}[t]{c|c|c}
+\hline
+\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
+\hline
+\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
+\hline
+\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
+\hline
+\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
+\hline
+\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
+\hline
+\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
+\hline
+\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
+\hline
+\hspace{1em}Recreational & 3 (4) & 7 (4)\\
+\hline
+\hspace{1em}University & 45 (56) & 141 (80)\\
+\hline
+\hspace{1em}National/International & 33 (41) & 28 (16)\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table2}Example of the data arrangement used for the network.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+\textbf{Participant} & \textbf{X\_1} & \textbf{X\_2}\\
+\hline
+1 & T1 & T2\\
+\hline
+1 & T2 & T3\\
+\hline
+2 & T1 & T2\\
+\hline
+3 & T1 & T2\\
+\hline
+3 & T2 & T3\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table3}Variables included in the final Bayesian network structure.}
+\centering
+\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
+\hline
+Variable & Definition & State 1 & State 2\\
+\hline
+clevel & Current competitive level & Club\_university\_county & National\_international\\
+\hline
+gender & Gender of the participant & Female & Male\\
+\hline
+hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
+\hline
+ind\_team & Participate in an individual or team based sport & Individual & Team\\
+\hline
+pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
+\hline
+nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
+\hline
+FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
+\hline
+BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
+\hline
+RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
+\hline
+stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
+\hline
+HRV & Root mean squared difference of successive RR intervals & 2.03-4 (Low) & >4-5.94 (High)\\
+\hline
+balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
+\hline
+nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table4}The number and percentage (\%) of types of injuries sustained by male and female participants.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l}
+\hline
+\multicolumn{1}{c|}{\textbf{ }} & \multicolumn{2}{c|}{\textbf{Female}} & \multicolumn{2}{c}{\textbf{Male}} \\
+\cline{2-3} \cline{4-5}
+\textbf{} & \textbf{Lower body} & \textbf{Upper body} & \textbf{Lower body} & \textbf{Upper body}\\
+\hline
+Joint / ligament & 14 (36) & 5 (36) & 37 (43) & 14 (38)\\
+\hline
+Muscle / tendon & 17 (44) & 6 (43) & 45 (52) & 12 (32)\\
+\hline
+Other & 8 (21) & 3 (21) & 5 (6) & 11 (30)\\
+\hline
+\multicolumn{5}{l}{\textit{Note: }}\\
+\multicolumn{5}{l}{Other included bone, skin and brain injuries.}\\
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table5}Probability of injured\_1 being in the ``injured'' state, conditional on each variable.}
+\centering
+\begin{tabular}[t]{l|r|r}
+\hline
+\textbf{Variable} & \textbf{Low} & \textbf{High}\\
+\hline
+balance\_1 & 0.21 & 0.30\\
+\hline
+hours & 0.18 & 0.28\\
+\hline
+nlelg\_1 & 0.24 & 0.26\\
+\hline
+stiffness\_1 & 0.17 & 0.31\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table6}Highest and lowest probability of injured\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_1.}
+\centering
+\begin{tabular}[t]{r|l|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{nlelg\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
+\hline
+\multicolumn{6}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.53 & club\_university\_county & High & Low & High & High\\
+\hline
+\hspace{1em}0.46 & national\_international & High & Low & High & Low\\
+\hline
+\hspace{1em}0.43 & national\_international & High & Low & High & High\\
+\hline
+\multicolumn{6}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.06 & national\_international & Low & Low & Low & Low\\
+\hline
+\hspace{1em}0.05 & national\_international & Low & Low & Low & High\\
+\hline
+\hspace{1em}0.04 & club\_university\_county & Low & Low & Low & Low\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table7}Probability of injured\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \_1.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+\textbf{Variable} & \textbf{High risk} & \textbf{Low risk}\\
+\hline
+Probability & 0.54 & 0.03\\
+\hline
+pi & injury & no injury\\
+\hline
+clevel & club\_university\_county & club\_university\_county\\
+\hline
+hours & High & Low\\
+\hline
+ind\_team & individual & individual\\
+\hline
+nlebase & High & High\\
+\hline
+stiffness\_1 & High & Low\\
+\hline
+balance\_1 & High & Low\\
+\hline
+RI\_1 & Low & High\\
+\hline
+BIS\_1 & Low & High\\
+\hline
+FFFS\_1 & High & Low\\
+\hline
+nlelg\_1 & Low & Low\\
+\hline
+rmssd\_1 & High & High\\
+\hline
+gender & male & female\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table8}Probability of injured\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\_2.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+\textbf{Variable} & \textbf{Low} & \textbf{High}\\
+\hline
+balance\_2 & 0.17 & 0.27\\
+\hline
+FFFS\_1 & 0.3 & 0.11\\
+\hline
+nlelg\_2 & 0.24 & 0.19\\
+\hline
+rmssd\_2 & 0.24 & 0.19\\
+\hline
+stiffness\_2 & 0.13 & 0.27\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table9}Highest and lowest probability of injured\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_2.}
+\centering
+\begin{tabular}[t]{l|l|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{FFFS\_1} & \textbf{nlelg\_2} & \textbf{stiffness\_2} & \textbf{rmssd\_2} & \textbf{balance\_2}\\
+\hline
+\multicolumn{6}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.52 & Low & Low & High & Low & High\\
+\hline
+\hspace{1em}0.45 & Low & High & High & Low & High\\
+\hline
+\hspace{1em}0.43 & Low & Low & High & High & High\\
+\hline
+\multicolumn{6}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.06 & High & High & Low & Low & Low\\
+\hline
+\hspace{1em}0.05 & High & Low & Low & High & Low\\
+\hline
+\hspace{1em}0.04 & High & High & Low & High & Low\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table10}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.}
+\centering
+\begin{tabular}[t]{l|r|r|l}
+\hline
+\textbf{Term} & \textbf{Estimate} & \textbf{Error} & \textbf{95\% CI}\\
+\hline
+Intercept & 0.00 & 0.03 & [-0.06, 0.05]\\
+\hline
+BIS & 0.40 & 0.03 & [0.34, 0.45]\\
+\hline
+rmssd & -0.14 & 0.03 & [-0.20, -0.08]\\
+\hline
+BIS:rmssd & -0.04 & 0.03 & [-0.10, 0.02]\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table11}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
+\centering
+\begin{tabular}[t]{r|l|l}
+\hline
+\textbf{Probability} & \textbf{nlec} & \textbf{Stiffness}\\
+\hline
+0.70 & +1SD & +1SD\\
+\hline
+0.64 & +1SD & mean\\
+\hline
+0.62 & +1SD & -1SD\\
+\hline
+0.53 & mean & +1SD\\
+\hline
+0.45 & mean & mean\\
+\hline
+0.44 & mean & -1SD\\
+\hline
+0.42 & -1SD & +1SD\\
+\hline
+0.35 & -1SD & mean\\
+\hline
+0.35 & -1SD & -1SD\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:table12}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+\centering
+\begin{tabular}[t]{r|l|l|l|l}
+\hline
+\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
+\hline
+\multicolumn{5}{l}{\textbf{Highest}}\\
+\hline
+\hspace{1em}0.77 & High & injury & +1SD & +1SD\\
+\hline
+\hspace{1em}0.74 & High & no injury & +1SD & +1SD\\
+\hline
+\hspace{1em}0.72 & Low & injury & +1SD & +1SD\\
+\hline
+\multicolumn{5}{l}{\textbf{Lowest}}\\
+\hline
+\hspace{1em}0.15 & Low & no injury & -1SD & +1SD\\
+\hline
+\hspace{1em}0.13 & Low & no injury & -1SD & mean\\
+\hline
+\hspace{1em}0.11 & Low & no injury & -1SD & -1SD\\
+\hline
+\end{tabular}
+\end{table}
 
 \hypertarget{refs}{}
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
+Agyapong-Badu, S., Warner, M., Samuel, D., and Stokes, M. (2016).
+Measurement of ageing effects on muscle tone and mechanical properties
+of rectus femoris and biceps brachii in healthy males and females using
+a novel hand-held myometric device. \emph{Archives of Gerontology and
+Geriatrics} 62, 59--67.
+doi:\href{https://doi.org/10.1016/j.archger.2015.09.011}{10.1016/j.archger.2015.09.011}.
+
+\leavevmode\hypertarget{ref-Aird2012}{}%
+Aird, L., Samuel, D., and Stokes, M. (2012). Quadriceps muscle tone,
+elasticity and stiffness in older males: Reliability and symmetry using
+the MyotonPRO. \emph{Archives of Gerontology and Geriatrics} 55,
+e31--e39.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.archger.2012.03.005}{https://doi.org/10.1016/j.archger.2012.03.005}.
+
 \leavevmode\hypertarget{ref-Andersen1988}{}%
@@
 
+\leavevmode\hypertarget{ref-Butler2003}{}%
+Butler, R. J., Crowell, H. P., and Davis, I. M. C. (2003). Lower
+extremity stiffness: Implications for performance and injury.
+\emph{Clinical Biomechanics} 18, 511--517.
+doi:\href{https://doi.org/10.1016/S0268-0033(03)00071-8}{10.1016/S0268-0033(03)00071-8}.
+
+\leavevmode\hypertarget{ref-Burkner2017a}{}%
+Bürkner, P. C. (2017). brms: An R package for Bayesian multilevel models
+using Stan. \emph{Journal of Statistical Software} 80.
+doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}.
+
+\leavevmode\hypertarget{ref-Caminal2018}{}%
+Caminal, P., Sola, F., Gomis, P., Guasch, E., Perera, A., Soriano, N.,
+and Mont, L. (2018). Validity of the Polar V800 monitor for measuring
+heart rate variability in mountain running route conditions.
+\emph{European Journal of Applied Physiology} 118, 669--677.
+doi:\href{https://doi.org/10.1007/s00421-018-3808-0}{10.1007/s00421-018-3808-0}.
+
+\leavevmode\hypertarget{ref-Chalmers2014}{}%
+Chalmers, J. A., Quintana, D. S., Abbott, M. J.-A., and Kemp, A. H.
+(2014). Anxiety disorders are associated with reduced heart rate
+variability: A meta-analysis. \emph{Frontiers in psychiatry} 5, 80.
+doi:\href{https://doi.org/10.3389/fpsyt.2014.00080}{10.3389/fpsyt.2014.00080}.
+
+\leavevmode\hypertarget{ref-Chen2012}{}%
+Chen, S. H., and Pollino, C. A. (2012). Good practice in Bayesian
+network modelling. \emph{Environmental Modelling and Software} 37,
+134--145.
+doi:\href{https://doi.org/10.1016/j.envsoft.2012.03.012}{10.1016/j.envsoft.2012.03.012}.
+
+\leavevmode\hypertarget{ref-Chuang2013}{}%
+Chuang, L. L., Lin, K. C., Wu, C. Y., Chang, C. W., Chen, H. C., Yin, H.
+P., and Wang, L. (2013). Relative and absolute reliabilities of the
+myotonometric measurements of hemiparetic arms in patients with stroke.
+\emph{Archives of Physical Medicine and Rehabilitation} 94, 459--466.
+doi:\href{https://doi.org/10.1016/j.apmr.2012.08.212}{10.1016/j.apmr.2012.08.212}.
+
+\leavevmode\hypertarget{ref-Chuang2012}{}%
+Chuang, L. L., Wu, C. Y., and Lin, K. C. (2012). Reliability, validity,
+and responsiveness of myotonometric measurement of muscle tone,
+elasticity, and stiffness in patients with stroke. \emph{Archives of
+Physical Medicine and Rehabilitation} 93, 532--540.
+doi:\href{https://doi.org/10.1016/j.apmr.2011.09.014}{10.1016/j.apmr.2011.09.014}.
+
+\leavevmode\hypertarget{ref-VanCingel2017}{}%
+Cingel, R. E. H. van, Hoogeboom, T. J., Melick, N. van, Meddeler, B. M.,
+and Nijhuis-van der Sanden, M. W. G. (2017). How to determine leg
+dominance: The agreement between self-reported and observed performance
+in healthy adults. \emph{Plos One} 12, 1--9.
+doi:\href{https://doi.org/10.1371/journal.pone.0189876}{10.1371/journal.pone.0189876}.
+
+\leavevmode\hypertarget{ref-Corr2013}{}%
+Corr, P. J. (2013). Approach and avoidance behaviour: Multiple systems
+and their interactions. \emph{Emotion Review} 5, 285--290.
+doi:\href{https://doi.org/10.1177/1754073913477507}{10.1177/1754073913477507}.
+
+\leavevmode\hypertarget{ref-Corr2016c}{}%
+Corr, P. J., and Cooper, A. J. (2016). The reinforcement sensitivity
+theory of personality questionnaire (RST-PQ): Development and
+validation. \emph{Psychological Assessment} 28, 1427--1440.
+doi:\href{https://doi.org/10.1037/pas0000273}{10.1037/pas0000273}.
+
+\leavevmode\hypertarget{ref-Corr2016b}{}%
+Corr, P. J., McNaughton, N., Wilson, M. R., Hutchison, A., Burch, G.,
+and Poropat, A. (2016). Neuroscience of motivation and organizational
+behavior: Putting the reinforcement sensitivity theory (RST) to work.
+\emph{Advances in Motivation and Achievement} 19, 65--92.
+doi:\href{https://doi.org/10.1108/S0749-742320160000019010}{10.1108/S0749-742320160000019010}.
+
+\leavevmode\hypertarget{ref-Denver2007}{}%
+Denver, J. W., Reed, S. F., and Porges, S. W. (2007). Methodological
+issues in the quantification of respiratory sinus arrhythmia.
+\emph{Biological Psychology} 74, 286--294.
+doi:\href{https://doi.org/10.1016/j.biopsycho.2005.09.005}{10.1016/j.biopsycho.2005.09.005}.
+
 \leavevmode\hypertarget{ref-Djaoui2017}{}%
@@
 
+\leavevmode\hypertarget{ref-Efron1993}{}%
+Efron, B., and Tibshirani, R. J. (1994). \emph{An introduction to the
+bootstrap}. New York: Chapman \& Hall Available at:
+\url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}.
+
+\leavevmode\hypertarget{ref-Enders2003}{}%
+Enders, C. K. (2003). Using the expectation maximization algorithm to
+estimate coefficient alpha for scales with item-level missing data.
+\emph{Psychological Methods} 8, 322--337.
+doi:\href{https://doi.org/10.1037/1082-989X.8.3.322}{10.1037/1082-989X.8.3.322}.
+
+\leavevmode\hypertarget{ref-Fawkner1999}{}%
+Fawkner, H. J., McMurrary, N. E., and Summers, J. J. (1999). Athletic
+injury and minor life events: A prospective study. \emph{Journal of
+Science and Medicine in Sport} 2, 117--124.
+doi:\href{https://doi.org/10.1016/S1440-2440(99)80191-1}{10.1016/S1440-2440(99)80191-1}.
+
+\leavevmode\hypertarget{ref-Fornell1981}{}%
+Fornell, C., and Larcker, D. F. (1981). Evaluating structural equation
+models with unobservable variables and measurement error. \emph{Journal
+of Marketing Research} 18, 39--50.
+doi:\href{https://doi.org/10.2307/3151312}{10.2307/3151312}.
+
+\leavevmode\hypertarget{ref-Fuller2006}{}%
+Fuller, C. W., Ekstrand, J., Junge, A., Andersen, T. E., Bahr, R.,
+Dvorak, J., Hägglund, M., McCrory, P., and Meeuwisse, W. H. (2006).
+Consensus statement on injury definitions and data collection procedures
+in studies of football (soccer) injuries. \emph{British journal of
+sports medicine} 40, 193--201.
+doi:\href{https://doi.org/10.1136/bjsm.2005.025270}{10.1136/bjsm.2005.025270}.
+
+\leavevmode\hypertarget{ref-Fuller2007b}{}%
+Fuller, C. W., Molloy, M. G., Bagate, C., Bahr, R., Brooks, J. H. M.,
+Donson, H., Kemp, S. P. T., McCrory, P., McIntosh, A. S., Meeuwisse, W.
+H., et al. (2007). Consensus statement on injury definitions and data
+collection procedures for studies of injuries in rugby union.
+\emph{British journal of sports medicine} 41, 328--31.
+doi:\href{https://doi.org/10.1136/bjsm.2006.033282}{10.1136/bjsm.2006.033282}.
+
+\leavevmode\hypertarget{ref-Fuster-Parra2017}{}%
+Fuster-Parra, P., Vidal-Conti, J., Borràs, P. A., and Palou, P. (2017).
+Bayesian networks to identify statistical dependencies. A case study of
+Spanish university students' habits. \emph{Informatics for Health and
+Social Care} 42, 166--179.
+doi:\href{https://doi.org/10.1080/17538157.2016.1178117}{10.1080/17538157.2016.1178117}.
+
 \leavevmode\hypertarget{ref-Galambos2005}{}%
@@
 
+\leavevmode\hypertarget{ref-Giles2016}{}%
+Giles, D., Draper, N., and Neil, W. (2016). Validity of the Polar V800
+heart rate monitor to measure RR intervals at rest. \emph{European
+Journal of Applied Physiology} 116, 563--571.
+doi:\href{https://doi.org/10.1007/s00421-015-3303-9}{10.1007/s00421-015-3303-9}.
+
+\leavevmode\hypertarget{ref-Gisselman2016}{}%
+Gisselman, A. S., Baxter, G. D., Wright, A., Hegedus, E., and Tumilty,
+S. (2016). Musculoskeletal overuse injuries and heart rate variability:
+Is there a link? \emph{Medical Hypotheses} 87, 1--7.
+doi:\href{https://doi.org/10.1016/j.mehy.2015.12.003}{10.1016/j.mehy.2015.12.003}.
+
+\leavevmode\hypertarget{ref-Heckerman1995}{}%
+Heckerman, D., Geiger, D., and Chickering, D. M. (1995). Learning
+Bayesian networks: The combination of knowledge and statistical data.
+\emph{Machine Learning} 20, 197--243.
+doi:\href{https://doi.org/10.1023/A:1022623210503}{10.1023/A:1022623210503}.
+
+\leavevmode\hypertarget{ref-Hrysomallis2011}{}%
+Hrysomallis, C. (2011). Balance ability and athletic performance.
+\emph{Sports Medicine} 41, 221--232.
+doi:\href{https://doi.org/10.2165/11538560-000000000-00000}{10.2165/11538560-000000000-00000}.
+
 \leavevmode\hypertarget{ref-Hughes2014}{}%
@@
 
+\leavevmode\hypertarget{ref-Ivarsson2014}{}%
+Ivarsson, A., Johnson, U., Lindwall, M., Gustafsson, H., and Altemyr, M.
+(2014). Psychosocial stress as a predictor of injury in elite junior
+soccer: A latent growth curve analysis. \emph{Journal of Science and
+Medicine in Sport} 17, 366--370.
+doi:\href{https://doi.org/10.1016/j.jsams.2013.10.242}{10.1016/j.jsams.2013.10.242}.
+
+\leavevmode\hypertarget{ref-Johnson2014}{}%
+Johnson, U., Tranaeus, U., and Ivarsson, A. (2014). Current status and
+future challenges in psychological research of sport injury prediction
+and prevention: A methodological perspective. \emph{Revista de
+Psicologia del Deporte} 23, 401--409. Available at:
+\url{https://www.rpd-online.com/article/view/v23-n2-johnson-tranaeus-ivarsson}.
+
+\leavevmode\hypertarget{ref-Kalkhoven2018}{}%
+Kalkhoven, J. T., and Watsford, M. L. (2017). The relationship between
+mechanical stiffness and athletic performance markers in sub-elite
+footballers. \emph{Journal of Sports Sciences} 36, 1022--1029.
+doi:\href{https://doi.org/10.1080/02640414.2017.1349921}{10.1080/02640414.2017.1349921}.
+
+\leavevmode\hypertarget{ref-Koller2009}{}%
+Koller, D., and Friedman, N. (2009). \emph{Probabilistic graphical
+models: Principles and techniques - Adaptive computation and machine
+learning}. Cambridge, MA: The MIT Press
+doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.
+
+\leavevmode\hypertarget{ref-Koo2016}{}%
+Koo, T. K., and Li, M. Y. (2016). A guideline of selecting and reporting
+intraclass correlation coefficients for reliability research.
+\emph{Journal of Chiropractic Medicine} 15, 155--63.
+doi:\href{https://doi.org/10.1016/j.jcm.2016.02.012}{10.1016/j.jcm.2016.02.012}.
+
+\leavevmode\hypertarget{ref-Kuhn2008}{}%
+Kuhn, M. C. from J. W., Weston, S., Williams, A., Keefer, C., and
+Engelhardt, A. (2008). caret: Classification and Regression Training.
+\emph{Journal of Statistical Software} 28.
+doi:\href{https://doi.org/10.1053/j.sodo.2009.03.002}{10.1053/j.sodo.2009.03.002}.
+
 \leavevmode\hypertarget{ref-Kumar2001}{}%
@@
 
+\leavevmode\hypertarget{ref-Paillard2006}{}%
+Paillard, T., Noé, F., Rivière, T., Marion, V., Montoya, R., and Dupui,
+P. (2006). Postural performance and strategy in the unipedal stance of
+soccer players at different levels of competition. \emph{Journal of
+Athletic Training} 41, 172--176. Available at:
+\url{http://www.ncbi.nlm.nih.gov/pubmed/16791302}.
+
 \leavevmode\hypertarget{ref-Passer1983a}{}%
@@
 
+\leavevmode\hypertarget{ref-Petrie1993}{}%
+Petrie, T. A. (1993). Coping skills, competitive trait anxiety, and
+playing status: Moderating effects on the life stress-injury
+relationship. \emph{Journal of Sport and Exercise Psychology} 15,
+261--274.
+doi:\href{https://doi.org/10.1123/jsep.15.3.261}{10.1123/jsep.15.3.261}.
+
+\leavevmode\hypertarget{ref-Petrie1992}{}%
+Petrie, T. A. (1992). Psychosocial antecedents of athletic injury: The
+effects of life stress and social support on female collegiate gymnasts.
+\emph{Journal of Behavioral Medicine} 18, 127--138.
+doi:\href{https://doi.org/10.1080/08964289.1992.9936963}{10.1080/08964289.1992.9936963}.
+
+\leavevmode\hypertarget{ref-Petrie2014}{}%
+Petrie, T. A., Deiters, J., and Harmison, R. J. (2014). Mental
+toughness, social support, and athletic identity: Moderators of the life
+stress-injury relationship in collegiate football players. \emph{Sport,
+Exercise, and Performance Psychology} 3, 13--27.
+doi:\href{https://doi.org/10.1037/a0032698}{10.1037/a0032698}.
+
+\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
+Pickering-Rodriguez, E. C., Watsford, M. L., Bower, R. G., and Murphy,
+A. J. (2017). The relationship between lower body stiffness and injury
+incidence in female netballers. \emph{Sports Biomechanics} 16, 361--373.
+doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}.
+
+\leavevmode\hypertarget{ref-Plews2017}{}%
+Plews, D. J., Scott, B., Altini, M., Wood, M., Kilding, A. E., and
+Laursen, P. B. (2017). Comparison of heart-rate-variability recording
+with smartphone photoplethysmography, polar H7 chest strap, and
+electrocardiography. \emph{International Journal of Sports Physiology
+and Performance} 12, 1324--1328.
+doi:\href{https://doi.org/10.1123/ijspp.2016-0668}{10.1123/ijspp.2016-0668}.
+
 \leavevmode\hypertarget{ref-Pruyn2015}{}%
@@
 doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}.
+
+\leavevmode\hypertarget{ref-Pruyn2016}{}%
+Pruyn, E. C., Watsford, M. L., and Murphy, A. J. (2016). Validity and
+reliability of three methods of stiffness assessment. \emph{Journal of
+Sport and Health Science} 5, 476--483.
+doi:\href{https://doi.org/10.1016/j.jshs.2015.12.001}{10.1016/j.jshs.2015.12.001}.
+
+\leavevmode\hypertarget{ref-Pulopulos2018}{}%
+Pulopulos, M. M., Vanderhasselt, M. A., and De Raedt, R. (2018).
+Association between changes in heart rate variability during the
+anticipation of a stressful situation and the stress-induced cortisol
+response. \emph{Psychoneuroendocrinology} 94, 63--71.
+doi:\href{https://doi.org/10.1016/j.psyneuen.2018.05.004}{10.1016/j.psyneuen.2018.05.004}.
+
+\leavevmode\hypertarget{ref-Qian2015}{}%
+Qian, S. S., and Miltner, R. J. (2015). A continuous variable Bayesian
+networks model for water quality modeling: A case study of setting
+nitrogen criterion for small rivers and streams in Ohio, USA.
+\emph{Environmental Modelling and Software} 69, 14--22.
+doi:\href{https://doi.org/10.1016/j.envsoft.2015.03.001}{10.1016/j.envsoft.2015.03.001}.
+
+\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
+R Core Team (2019). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing
+Available at: \url{https://www.r-project.org/}.
+
+\leavevmode\hypertarget{ref-Riemann1999d}{}%
+Riemann, B. L., Guskiewicz, K. M., and Shields, E. W. (1999).
+Relationship between clinical and forceplate measures of postural
+stability. \emph{Journal of Sport Rehabilitation} 8, 71--82.
+doi:\href{https://doi.org/10.1123/jsr.8.2.71}{10.1123/jsr.8.2.71}.
 
@@
 
+\leavevmode\hypertarget{ref-Ross2011}{}%
+Ross, S. E., Linens, S. W., Wright, C. J., and Arnold, B. L. (2011).
+Balance assessments for predicting functional ankle instability and
+stable ankles. \emph{Gait and Posture} 34, 539--542.
+doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}.
+
+\leavevmode\hypertarget{ref-Ruiz-Perez2019}{}%
+Ruiz-Pérez, I., Ayala, F., Miguel Puerta, J., Elvira, J. L. L., De Ste
+Croix, M., Hernández-Sánchez, S., and Vera-Garcia, F. J. (2019). A
+Bayesian Network approach to study the relationships between several
+neuromuscular performance measures and dynamic postural control in
+futsal players. \emph{PLoS ONE} 14, 1--18.
+doi:\href{https://doi.org/10.1371/journal.pone.0220065}{10.1371/journal.pone.0220065}.
+
+\leavevmode\hypertarget{ref-Norvig2009}{}%
+Russell, S. J., and Norvig, P. (2009). \emph{Artificial Intelligence: A
+Modern Approach}. 3rd ed. Prentice Hall Available at:
+\url{http://aima.cs.berkeley.edu/}.
+
 \leavevmode\hypertarget{ref-Scutari2014}{}%
@@
 
+\leavevmode\hypertarget{ref-Timpka2014}{}%
+Timpka, T., Alonso, J. M., Jacobsson, J., Junge, A., Branco, P.,
+Clarsen, B., Kowalski, J., Mountjoy, M., Nilsson, S., Pluim, B., et al.
+(2014). Injury and illness definitions and data collection procedures
+for use in epidemiological studies in Athletics (track and field):
+Consensus statement. \emph{British Journal of Sports Medicine} 48,
+483--490.
+doi:\href{https://doi.org/10.1136/bjsports-2013-093241}{10.1136/bjsports-2013-093241}.
+
+\leavevmode\hypertarget{ref-Trojian2006}{}%
+Trojian, T. H., and McKeag, D. B. (2006). Single leg balance test to
+identify risk of ankle sprains. \emph{British Journal of Sports
+Medicine} 40, 610--613.
+doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}.
+
+\leavevmode\hypertarget{ref-Valovich2003}{}%
+Valovich, T. C., Perrin, D. H., and Gansneder, B. M. (2003). Repeat
+administration elicits a practice effect with the Balance Error Scoring
+System but not with the Standardized Assessment of Concussion in high
+school athletes. \emph{Journal of Athletic Training} 38, 51--56.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/12937472}.
+
 \leavevmode\hypertarget{ref-Wiese-Bjornstal2009}{}%
@@
 
+\leavevmode\hypertarget{ref-Xiao-xuan2007}{}%
+Xiao-xuan, H., Hui, W., and Shuo, W. (2008). Using expert's knowledge to
+build Bayesian networks. in \emph{International conference on
+computational intelligence and security workshops} (IEEE), 220--223.
+doi:\href{https://doi.org/10.1109/cisw.2007.4425484}{10.1109/cisw.2007.4425484}.
+
+\leavevmode\hypertarget{ref-Xu2015}{}%
+Xu, J. G., Zhao, Y., Chen, J., and Han, C. (2015). A structure learning
+algorithm for Bayesian network using prior knowledge. \emph{Journal of
+Computer Science and Technology} 30, 713--724.
+doi:\href{https://doi.org/10.1007/s11390-015-1556-8}{10.1007/s11390-015-1556-8}.
+\end{cslreferences}
+
 \end{document}

### 2020-07-20 wordmaker.Rmd
pre results edits

--- 2020-05-23 full_paper_v2.Rmd
@@
 ---
-title: "An interdisciplinary examination of stress and injury occurrence in athletes"
-runtitle: "Short Title"
-documentclass: frontiersHLTH  # or frontiersHLTH, or frontiersFPHY
-author:
-  - name: Dr Harry Fisher
-    affiliation: '1'
-    etal: Fisher # First author's last name. 
-    email: harryfisher21@gmail.com # Indicates corresponding Author
-    institution: Cardiff Metropolitan University
-    street: Cyncoed Road
-    zip: CF23 6XD
-    country: United Kingdom
-  - name: Dr Marianne Gittoes
-    affiliation: '1'
-  - name: Professor Lynne Evans
-    affiliation: '1'
-  - name: Miss Leah Bitchell
-    affiliation: '1'  
-  - name: Dr Richard Mullen  
-    affiliation: '2'
-  - name: Dr Marco Scutari
-    affiliation: '3'    
-affiliation:
-  - id: '1'
-    department: School of Sport and Health Science
-    institution: Cardiff Metropolitan University
-    city: Cardiff
-    country: United Kingdom
-  - id: '2' 
-    department: Division of Sport, Health & Exercise Sciences
-    institution: Brunel University
-    city: London
-    country: United Kingdom
-  - id: '3' 
-    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
-    city: Manno
-    country: Switzerland
-
-output: rticles::frontiers_article
-csl: frontiers.csl
-bibliography: library.bib
+title: "wordmaker"
+output: bookdown::html_document2
+always_allow_html: true
+
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
-source("r-files/functions.R")
-source("r-files/methods_code.R")
-source("r-files/results_code.R")
-```
-
-# Introduction {-}
-
-Over the last four decades sport related injuries have received increased research attention [@Ivarsson2017].
-This attention is unsurprising given the high incidence [@Rosa2014; @Sheu2016],
-and undesirable physical and psychological effects of sports injuries [@Leddy1994; @Brewer2012].
-To mitigate against both the increasing incidence and	undesirable consequences of injury, research  has identified several psychological [@Slimani2018],
-anatomical [@Murphy2003],
-biomechanical [@Neely1998; @Hughes2014]
-and environmental [@Meeuwisse2007]
-factors associated with sports injury occurrence.
-Indeed, several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009],
-of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
-
-\begin{figure}[!h]
-\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
-\label{fig:fig1}
-\end{figure}
-
-Williams and Andersen's [@Williams1998] model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) will contribute to their response, either interactively or in isolation.
-Central to the model is the stress response, which reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of many stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater risk of injury.
-
-Within Williams and Andersen’s [@Williams1998] model, major life events, a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007];
-specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
-However, personality traits and coping resources have also been	found to predict injury, with for example, athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to athletes with the opposite profile.
-[@Smith1990; @Lavallee1996; @Ivarsson2010].
-However, the amount of variance explained by the psychosocial factors proposed by the model has been modest, typically between 5 - 30% [@Galambos2005; @Ivarsson2010]; suggesting other factors are also likely to contribute to injury occurrence.
-
-While the psychosocial factors proposed in Williams and Andersen’s [@Williams1998] model have received the most research attention, the mechanisms through which these factors are proposed to exert their effect have remained under-investigated in the literature.
+# source("r-files/functions.R")
+# source("r-files/methods_code.R")
+# source("r-files/results_code.R")
+load("all_data.RData")
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect, Rgraphviz)
+library(readxl)
+library(janitor)
+library(irr)
+library(lavaan)
+library(caret)
+library(parallel)
+library(zoo)
+library(sjmisc)
+
+```
+
+\section{Introduction}
+
+Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012].
+Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
+
+Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
+The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
+
+Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
+Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.
+
+While the psychosocial factors in Williams and Andersen’s [@Williams1998] model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
 To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
 However, to date, the research has largely focused on attentional deficits [@Andersen1999; @Rogers2005; @Wilkerson2012a; @Swanik2007].
-For example, Andersen and Williams [@Andersen1999]
-measured peripheral and central vision during high and low stress conditions and found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following season.
-Indeed, Rodgers and Landers [@Rogers2005] supported Andersen and Williams's [@Andersen1999] earlier findings reporting that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
-However, the remaining variance between negative life events and athletic injury through the other proposed mechanisms, such as increased muscle tension and reduced motor control, remains to be explored [cf.
+For example, Andersen and Williams [@Andersen1999] found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
+Rodgers and Landers [@Rogers2005] further supported Andersen and Williams’s [@Andersen1999] earlier findings by identifying that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
+
+Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse [cf.
 @Williams1998].
-
-One challenge faced by researchers addressing the sports injury problem with a psychological lens is the multifactorial nature of injury, and the possible contribution of other non-psychological factors to the stress response
-[@Meeuwisse2007; @Wiese-Bjornstal2009].
-For example, a large body of research indicates that training-related stress is also likely to be related to the stress response and injury occurrence [@Lee2017; @Djaoui2017],
-and may account for the unexplained variance from the psychological predictors of injury.
-Appaneal and Perna [@Appaneal2014]
-proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model and to address some of these issues.
-To elaborate, the BMSAIH aimed to clarify the mediating pathways between the stress response and injury, consider other health outcomes and behavioural factors that impact sports participation,
-and integrate the impact of training on athletes’ health [@Appaneal2014].
+One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response [@Meeuwisse2007; @Wiese-Bjornstal2009].
+For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence [@Lee2017; @Djaoui2017] and may account for the unexplained variance from the psychological predictors.
+Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation [@Appaneal2014].
 The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p.
 74] to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides a framework for future research to build on Williams and Andersen's [@Williams1998] model, by including other physiological markers of training-related stress, which together may provide greater insight into the injury process.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other research has identified additional markers of training-related stress that are associated with an increased risk of injury; for example,
-heart rate variability [@Bellenger2016; @Williams2017],
-postural stability [@Romero-Franco2014]
-and muscle stiffness [@Pruyn2015].
-However, these markers are often studied in isolation without an assessment of the psychosocial factors that are known to contribute to injury, thereby limiting our understanding of how psychosocially and physiologically derived stress may contribute synergistically to injury occurrence.
-Recently, Bittencourt et al. [@Bittencourt2016]
-suggested that to better understand the multifactorial nature of sports injuries, research needs to move away from studying risk factors in isolation and instead adopt a complex systems approach to injury.
-Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p. 3],
-where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
-
-A challenge when adopting a complex systems approach is using an appropriate analysis technique that is able to capture the uncertainty and complexity of the relationships between different variables.
-One technique that provides a solution to this challenge is Bayesian network (BN) modelling.
-BN's allow the construction of graphical probabilistic models using the underlying structure that connects different variables (nodes in the network)  [@Scutari2014].
-The learned structure can be used for inference by obtaining the posterior probabilities of a particular node for a given query (e.g., if the value of Node A is x and the value of Node B is y, what is the probability Node C being value z?).
-Furthermore, BN's do not distinguish between dependent and independent variables as they are a form of unsupervised learning, which is a strength over regression or structure equation models when the underlying relationship in the network may not be known [@Olmedilla2018].
-
-To summarise, despite offering a possible framework to build on the research stemming from Williams and Andersen's [@Williams1998] model, there remains and opportunity to explore other physiological stress-related markers proposed by the BMSAIH, in addition to the already well-established psychological characteristics known to be related to injury [@Appaneal2014].
-Furthermore, research has typically not captured changes in both psychosocial factors and stress-related physiological markers that may occur between initial measurement and injury occurrence.
-Given the exploratory nature of the current study the following objectives were defined:
-
-- Identify suitable markers of stress that can be easily captured in a large cohort of athletes in a field based setting. 
-- Capture the markers of stress and injury occurrence using a prospective, repeated measures design.
-- Explore and evaluate the relationships between the markers of stress and injury using Bayesian network modelling.
-
-# Methods {-}
+Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen’s [@Williams1998] model whilst including other physiological markers of training-related stress.
+
+Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury.
+However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
+Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
+Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
+
+Recently, Bittencourt et al.
+[@Bittencourt2016] advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p.
+3], where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
+Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables [@Scutari2014].
+The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
+Furthermore, unlike regression or structural equation models, BN’s do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known [@Olmedilla2018].
+BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen’s [@Williams1998] stress injury model and Appaneal’s BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
+A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
+
+\section{Methods}
 
@@
 
-## Participants
-
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study.
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf from a British University and local sports clubs (Table \ref{tab:table1}).
+\subsection{Participants}
+
+A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
 Participants self-rated competitive level ranged from recreational to international standard. 
+They were engaged in training for their respective sports for at least five hours per week.
 A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-At the start of the study, all participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks).
-Participants were engaged in training for their respective sports for at least five hours per week.
+All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study. 
 Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
@@
 A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
 
@@
 
-### Injury
-
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine. 
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
-Injury status (did / did not sustain an injury) served as the main outcome measure.
-
 ## Procedure
 
-At the start of the UK academic year (September 2016 and 2017), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. 
-With the coaches' permission, the lead researcher attended training sessions to inform the athletes about the overall purpose of the study and the requirements of participation.
-To be eligible for the study athletes had to be injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) and training a minimum of five hours per week.
-Athletes who met the criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-
-A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive twelve-month periods between September 2016 and September 2018.
-Each participant was asked to attend a total of four data collections over a twelve-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). 
+At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study.
+With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
+Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
+A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
 Participants provided informed consent before data collection commenced.
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
+In order to use discrete networks, continuous variables must first be split into categorical levels.
+When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
+For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
+Typically, no more than three levels have been used in Bayesian networks in the sports injury literature [@Olmedilla2018; @Ruiz-Perez2019]
 
@@
 Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
-In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficient at identifying an approximation of the true network structure.
+In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
 Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
-The following sections detail the steps taken in the current study to firstly prepare the data for the network, and then obtain the structure of the network that was used for inference.
-
-### Data Preparation
-
+The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
+
+### First network
+
+#### Data Preparation
+
+Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
 To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 668 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
@@
 
-Preliminary correlation analysis of the data revealed no strong linear dependencies in the data, and therefore the data were binarised to create "Low" and "High" levels within each variable.
-Binarising variables is a common procedure for BN modelling and can help improve the fit of the model by reducing the number of possible states for each node [@Beuzen2018].
-All variables were approximately normally distributed; therefore, the median of each variable was used as the cut-off point to create the "Low" and "High" levels.
-
-For the LESCA, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
+A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
+All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
+For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
 Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative and cumulative total life event scores were first log scaled and then binarised based on the median at each time point (nlelg and tlelg respectively).
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
 In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
 
-### Network structure
+#### Network structure
 
@@
   Participant = c("1", "1", "2", "3", "3"),
-  "X_1" = c("T1", "T2", "T1", "T1", "T2"),
-  " " = "->",
-  "X_2" = c("T2", "T3", "T2", "T2", "T3")
+  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
+  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
 )
@@
   caption = "Example of the data arrangement used for the network.",
-  escape = TRUE
+  escape = FALSE
 ) %>%
@@
 
-### Preliminary network structures
-
-Prior to the final network structure presented in the results section, several structures were investigated.
+#### Preliminary network structures
+
+Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
 Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models.
@@
 
-Finally, both total score and asymmetry were included in the initial network.
+Finally, both total score and asymmetry for balance were included in the initial network.
 However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
 Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} includes the variables that were included in the final network structure.
-
-\begin{table}[H]
-\begin{adjustwidth}{-2.25in}{0in}
-\caption{\label{tab:table3}Variables included in the final network structure.}
-\centering
-\begin{tabular}{l|l|l|l}
-\hline
-{\bf Variable} & {\bf Definition} & {\bf State} 1 & {\bf State 2}\\
-\hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Sustained previous injury & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
-\hline
-FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
-\hline
-BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
-\hline
-RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
-\hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
-\hline
-rmssd & Root mean squared difference of successive RR intervals & 2.03-4.02 (Low) & >4.02-5.94 (High)\\
-\hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-\end{tabular}
-\end{adjustwidth}
-\end{table}
-
-Preliminary network structures also revealed strong dependencies between the same variables at subsequent time points. 
-For example, the probability that stiffness_1 and stiffness_2 were both "High", or both "Low" was approximately 80%. 
-Including the arcs between the same variables from X_1 $\rightarrow$ X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would be either be in a "Low" or "High" state for each pair of variables in the network.
-To more appropriately assess changes _within_ variables over time, a second BN was investigated by modelling the differences between variables at different time points.
-The use of differential equations to model changes in variables over time is a common procedure in BN analysis when there are repeated measurements in the data [@Scutari2017].
-To obtain the structure, variables suffixed with _1 were subtracted from variables suffixed with _2 to calculate the difference between variables measured at time points T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3.
-Independent variables were then standardized to allow relative changes between variables to be compared. 
-The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-The result was a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that would only have captured changes when the median threshold was crossed from "Low" to "High".
-Identical blacklists to the first network were used for arcs between independent and explanatory variables. 
-The nlebase variable was also dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network.
-
-To obtain the final networks, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
+To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
+
+```{r table3, results='asis'}
+varcutoffs %>% 
+  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
+) %>% 
+  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
+  kable("latex",
+       align = c("llcc"),
+       caption = "Variables included in the final Bayesian network structure.",
+       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
+  kable_styling(latex_options = "HOLD_position") %>% 
+  column_spec(2, width = c("5.2cm")) 
+```
+
+Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
+For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%.
+Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network.
+Therefore, these arcs were blacklisted from the network.
+To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
 The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
 Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
-An arc strength of 1 indicates that the arc is always present in the network, with the value decreasing as arcs are found in fewer networks. 
-In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013]. 
-Arcs below 0.3 were not included in the model.
-The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
-
-### Network Inference
+The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction.
+An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
+In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
+Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+
+#### Network Inference
 
@@
 To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = "injured") to be calculated based on the new evidence. 
+The query allows the odds of a particular node state (e.g., injured_1 = “injured”) to be calculated based on the new evidence.
 CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
 The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
-Inference was first performed on arcs that had a strength greater than 0.50 between the explanatory variables and independent variables and between different independent variables in the network. 
-Of particular interest in the current study were the variables that were connected to "injured" nodes.
-To examine the variables that were associated with injured nodes in the network, the Markov blanket of "injured_1" and "injured_2" were examined. 
+Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
+Of particular interest were the variables that were connected to “injured” nodes, which were examined in the network using the Markov blanket of “injured_1” and “injured_2”.
 A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
-
-The second network contained both continuous and discrete data. 
-To examine dependencies between continuous variables with arc strengths above 0.5 in the second network, random samples were generated based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then extracted and examined with Bayesian linear regression models using the _brms_ package [@Burkner2017a] to determine the relationship between nodes in the network. 
-Similar to the first network, the Markov blanket of the "injured" node was also investigated by determining the highest probability of injury with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-
-# Results {-}
+CPQ’s were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the “injured” state.
+
+### Second network
+
+#### Data preparation
+
+Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
+
+For the second network, change scores for continuous variables between  T1 $\rightarrow$  T2 and  T2 $\rightarrow$  T3 were standardized to allow relative changes between variables to be compared.
+The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
+Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from "Low" to "High".
+
+#### Network structure
+
+Similar to the first network, blacklists were used to prevent arcs from independent variables $\rightarrow$ explanatory variables.
+In addition, "nlebase" variable was dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network. 
+The final network was obtained using the same approach as the first network.
+
+#### Network Inference
+
+Conditional probability queries (CPQ) were again used to perform inference the network structures.
+The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
+Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
+To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
+The samples were then examined using a Bayesian linear regression models with the _brms_ package [@Burkner2017a] to determine the relationship between these nodes. 
+Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
 
@@
 
-### First network structure
+```{r table4, results='asis'}
+
+cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
+  kable(
+    .,
+    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
+    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
+    escape = F
+  ) %>%
+  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
+  footnote(general = "Other included bone, skin and brain injuries.") %>%
+  kable_styling(latex_options = "HOLD_position")  %>%
+  row_spec(0,bold=TRUE)
+
+```
+
+## First network structure
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
 ### Markov blanket for injured_1
@@
 ```{r}
-# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
-# q1p <- graphviz.plot(plotsg1,
-#                      shape = "rectangle",
-#                      render = F,
-#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
-# 
-# strq1 <- str %>% 
-#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") & 
-#            to %in% c("balance_1", "injured_1")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq1$strength)
-# labels[4] <- str_pad(labels[4], side = "left", width = 2)
-# 
-# edgenames <- edgeNames(q1p)
-# names(labels) <- edgenames
-# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
-# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6, 
-#                                     "stiffness_1~injured_1" = 3,
-#                                     "clevel~balance_1" = 3),
-#                             lty = c("injured_1~balance_1" = "dashed",
-#                                     "hours~injured_1" = "dashed"))
-# 
-# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 12)))
-# 
-# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
-# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
-# renderGraph(q1p)
-
-```
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
+plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
+q1p <- graphviz.plot(plotsg1,
+                     shape = "rectangle",
+                     render = F,
+                     groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
+
+strq1 <- str %>%
+  filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") &
+           to %in% c("balance_1", "injured_1")) %>%
+  filter(strength > 0.30) %>%
+  mutate(strength = format(round(strength, digits = 2), nsmall = 2))
+
+labels <- as.character(strq1$strength)
+labels[4] <- str_pad(labels[4], side = "left", width = 2)
+
+edgenames <- edgeNames(q1p)
+names(labels) <- edgenames
+nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
+edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6,
+                                    "stiffness_1~injured_1" = 3,
+                                    "clevel~balance_1" = 3),
+                            lty = c("injured_1~balance_1" = "dashed",
+                                    "hours~injured_1" = "dashed"))
+
+q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
+graph.par(list(edges=list(fontsize = 12)))
+
+nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
+nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
+renderGraph(q1p)
+
+```
+
+The conditional probability query (CPQ) for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
+The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
+Negative life events had a negligible effect when moving from the "Low" to "High" state.
 
@@
 
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-```{r table8, results="asis"}
-
-query3 %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
 The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
@@
 ```
-
-# Discussion
-
-# Disclosure/Conflict-of-Interest Statement {-}
-
-<!--  
-Frontiers follows the recommendations by the International Committee of Medical
-Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
-that all financial, commercial or other relationships that might be perceived by
-the academic community as representing a potential conflict of interest must be
-disclosed. If no such relationship exists, authors will be asked to declare that
-the research was conducted in the absence of any commercial or financial
-relationships that could be construed as a potential conflict of interest. When
-disclosing the potential conflict of interest, the authors need to address the
-following points:
-
- - Did you or your institution at any time receive payment or services from a
-   third party for any aspect of the submitted work?
- - Please declare financial relationships with entities that could be perceived
-   to influence, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- - Please declare patents and copyrights, whether pending, issued, licensed
-   and/or receiving royalties relevant to the work.
- - Please state other relationships or activities that readers could perceive to
-   have influenced, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- -->
-
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Author Contributions {-}
-
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-# Acknowledgments {-}
-
-Funding:
-
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# References
-
-A reference list should be automatically created here. However it won't. Pandoc 
-will place the list of references at the end of the document instead. There are 
-no convenient solution for now to force Pandoc to do otherwise. The easiest way 
-to get around this problem is to edit the LaTeX file created by Pandoc before 
-compiling it again using the traditional LaTeX commands.
-
-# Figures {-}
-
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```

### 2020-07-20 r-files/results_code.R

@@ -235,9 +235,9 @@ fittedcont <- bn.fit(avgcont, diffg)
 
 testdataa <- cpdist(fittedcont, nodes = c("FFFS", "rmssd", "BIS"), evidence = TRUE, n = 1000) 
 
-bm1 <- brm(FFFS ~ BIS*rmssd, data = testdataa)
+prior1 <- prior(normal(0, 5), class = "b")
 
-
+bm1 <- brm(FFFS ~ BIS*rmssd, data = testdataa, prior = prior1)
 
 
 

### 2020-07-29 apa_style.Rmd
progress

@@
 
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no
@@
 always_allow_html: true
+header_includes:
+  - \usepackage{pdflscape}
+  - \newcommand{\blandscape}{\begin{landscape}}
+  - \newcommand{\elandscape}{\end{landscape}}
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
-# source("r-files/functions.R")
-# source("r-files/methods_code.R")
-# source("r-files/results_code.R")
-load("all_data.RData")
-pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect, Rgraphviz)
-library(readxl)
-library(janitor)
-library(irr)
-library(lavaan)
-library(caret)
-library(parallel)
-library(zoo)
-library(sjmisc)
-
-```
-
-\section{Introduction}
+ source("r-files/functions.R")
+ source("r-files/methods_code.R")
+ source("r-files/results_code.R")
+#load("all_data.RData", envir = myEnv)
+#myEnv <- new.env()
+#attach(myEnv, name="sourced_scripts")
+
+pacman::p_load(tidyverse, 
+               knitr,
+               kableExtra,
+               qwraps2, 
+               brms,
+               broom.mixed, 
+               bnlearn, 
+               tidyselect, 
+               Rgraphviz, 
+               readxl, 
+               janitor,
+               irr, 
+               lavaan,
+               caret,
+               parallel,
+               zoo, 
+               sjmisc)
+
+```
+
+10422 words in text body
+3142 words in reference section
+
+# Introduction
 
@@
 Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
+
+```{r fig1, fig.cap="Stress and injury model (Williams and Andersen, 1998).", fig.align='center',fig.pos = "!h"}
+include_graphics("figures_doc/Fig1.jpeg")
+```
 
@@
 
-\section{Methods}
+# Material and Methods
 
@@
 
-\subsection{Participants}
+## Ethics Statement
+
+Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
+
+## Participants
 
@@
 All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
-Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
 
@@
 
+```{r fig2, fig.cap="mBESS positions (A-F).
+Top row, firm surface.
+Bottom row, foam surface.
+Left column, dominant leg stance.
+Middle column, non-dominant leg stance.
+Right column, Tandem leg stance.", out.width = "100%", fig.align='center',fig.pos = "!h"}
+include_graphics("figures_doc/Fig2.jpeg")
+```
+
 The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
@@
 
+```{r fig3, fig.cap = 'Study design.
+For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width="100%"}
+
+include_graphics("figures_doc/studydesign_new.png")
+```
+
 For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
@@
 Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
+
+```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
+include_graphics("figures_doc/sessionprotocol.png")
+```
 
@@
 
+```{r fig5, fig.cap="A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.", fig.pos="H", out.width="100%", fig.align='center'}
+
+include_graphics("figures_doc/examplenetwork.png")
+
+```
+
 In order to use discrete networks, continuous variables must first be split into categorical levels.
@@
 
-BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
+BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017; @Ruiz-Perez2019] and offer several benefits over traditional statistical analysis. 
 For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
@@
 To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings.
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings.
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
@@
 A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
-All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
+All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split. 
 For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
@@
 To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
-In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
+In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
 However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
-Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
+Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
 Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
@@
 The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
-Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
-RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
+Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
+RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
 Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
 Therefore, only RI was included in the final structure.
-
 Finally, both total score and asymmetry for balance were included in the initial network.
@@
 ) %>% 
-  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
+  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
   kable("latex",
@@
 
-During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
+```{r include=FALSE}
+datanet %>% 
+  group_by(id) %>% 
+  mutate(count = sum(ifelse(injured == "injured", 1, 0))) %>% 
+  select(id, count) %>% 
+  mutate(count = ifelse(count == 0, 0, 1)) %>%
+  distinct(id, .keep_all = TRUE) %>%
+  ungroup() %>% 
+  janitor::tabyl(count)
+```
+
+During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
 Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
@@
   mutate(prob = format(round(prob, digits = 2), nsmall = 2))
+
 ```
@@
   kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
+  row_spec(0, bold = TRUE)
+
+str_finder <- function(from, to) {
+  format(round(str[which(str$from == from & str$to == to), 3], digits = 2), nsmall = 2)
+}
+
+str_cfinder <- function(from, to) {
+  format(
+    round(
+      strchange[which(strchange$from == from &
+        strchange$to == to), 3],
+      digits = 2
+    ),
+    nsmall = 2
+  )
+}
 ```
@@
 
-The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
-Several of the explanatory variables showed strong connections with independent variables in the network.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
-The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
+The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
+Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
+The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
 Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
-Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
-The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
-The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
+Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
+The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
+High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
+Irrespective of sport or gender, strong connections were found between explanatory variables.
+The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.
+
+The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability. 
+The arc from pi $\rightarrow$ stiffness_1 was `r str_finder("pi", "stiffness_1")` with athletes who reported an injury in the preceding 12 months being more likely to have “High” (`r x[26,1]`) compared to “Low” (`r x[28,1]`).
+stiffness.
+Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (`r str_finder("BIS_1", "FFFS_1")`) and BIS_2 $\rightarrow$ FFFS_2 (`r str_finder("BIS_2", "FFFS_2")`) were identified.
 In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
-For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.
+The arc between nlelg $\rightarrow$  BIS had a strength of `r str_finder("nlelg_1", "BIS_1")` for nlelg_1  $\rightarrow$ BIS_1, and `r str_finder("nlelg_2", "BIS_2")` nlelg_2 and BIS_2.
+"Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]` for nlelg_1  $\rightarrow$ BIS_1, and `r y[17,1]` to `r y[18,1]` for nlelg_2  $\rightarrow$ BIS_2.
+
+\newpage
+
+```{r fig6, fig.align='default', fig.height=8, fig.width=12, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables.
+The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
+renderGraph(gR)
+```
 
@@
 
-The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
 The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
 
-```{r}
-# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
-# q1p <- graphviz.plot(plotsg1,
-#                      shape = "rectangle",
-#                      render = F,
-#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
-# 
-# strq1 <- str %>%
-#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") &
-#            to %in% c("balance_1", "injured_1")) %>%
-#   filter(strength > 0.30) %>%
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq1$strength)
-# labels[4] <- str_pad(labels[4], side = "left", width = 2)
-# 
-# edgenames <- edgeNames(q1p)
-# names(labels) <- edgenames
-# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
-# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6,
-#                                     "stiffness_1~injured_1" = 3,
-#                                     "clevel~balance_1" = 3),
-#                             lty = c("injured_1~balance_1" = "dashed",
-#                                     "hours~injured_1" = "dashed"))
-# 
-# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 12)))
-# 
-# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
-# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
-# renderGraph(q1p)
-
-```
-
-The conditional probability query (CPQ) for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
+```{r fig7, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig7.jpeg")
+```
+
+The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table5}.
 The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
@@
 When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
-Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
-In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.
+
+Negative life event stress had a negligible effect on the probability of injury, only influencing  injured_1 when all other variable were fixed to "Low".
+In this instance, the probability of injured_1 being "injured" rose marginally from 0.04 to 0.19, when negative life events was in the "Low" and "High" states, respectively.
 
@@
     caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F
+    escape = F,
+    align = "ccccc"
   ) %>%
@@
 
-<!-- Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state.
--->
-<!-- The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}.  -->
-<!-- Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state. -->
-
-<!-- # ```{r table7, results="asis"} -->
-<!-- # injured1comp %>% -->
-<!-- #   filter(variable != "outcome") %>%  -->
-<!-- #   mutate(variable = ifelse(variable == "prob", "Probability", as.character(variable))) %>%  -->
-<!-- #   mutate_all(~str_replace_all(., "_", "\\\\_"))%>% -->
-<!-- #   kable(  ., -->
-<!-- #     caption = "Probability of injured\\_1 being in the ``injured'' state conditional on explanatory variables and independent variables suffixed with \\_1.", -->
-<!-- #     col.names = c("Variable", "High risk", "Low risk"), -->
-<!-- #     escape = F -->
-<!-- #   ) %>% -->
-<!-- #   kable_styling(latex_options = "HOLD_position") %>% -->
-<!-- #   row_spec(0,bold=TRUE) -->
-<!-- # ``` -->
-
 ### Markov blanket of injured_2
@@
 set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2")
+var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2")
 query3 <- map_dfr(var2, ~newprobtable(tallc,
@@
 query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
+                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2"),
                        "injured_2",
@@
 ) %>% 
-  toptail(3) %>% 
   arrange(-prob) %>% 
@@
 
-The Markov blanket for injured_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and rmssd_2.
+The Markov blanket for injured_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2.
 The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ rmssd_2 were also present in the Markov blanket for injured_2.
-Results of the first query for injured_2 in the "injured" state are presented in table \ref{tab:table8}.
+Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ hrv_2 were also present in the Markov blanket for injured_2.
 Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
@@
 
-```{r table8, results="asis"}
+```{r table7, results="asis"}
 
 query3 %>%
+  mutate(Low = format(Low, nsmall = 2)) %>% 
   mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
@@
     col.names = c("Variable", "Low", "High"),
-    escape = F
+    escape = F,
+    align = "cccccc"
   ) %>%
@@
 
+```{r fig8, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig8.jpeg")
+```
+
 ```{r}
-# plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & 
-#            to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[6] <- str_pad(labels[6], side = "left", width = 2)
-# edgenames <- c("gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-The conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from 0.14 to 0.34 when FFFS_1 was in the "Low" compared to "High" state.
+fffs_query <- query4 %>% 
+  filter(stiffness_2 == "High" & nlelg_2 == "High" & balance_2 == "High" & hrv_2 == "High")
+```
+
+The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
+The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
+With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
 The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
 
-```{r table9, results="asis" }
+```{r table8, results="asis" }
 
 query4 %>% 
+  toptail(3) %>%
   select(7,2:6) %>%
@@
     caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F
+    escape = F,
+    align = "cccccc"
   ) %>%
@@
 
-The network for changes within variables is presented in Fig \ref{fig:fig9}.
-An arc between BIS $\rightarrow$ FFFS with strength 1 was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strength of 0.60.
-The arc between RMSSD $\rightarrow$ FFFS was 0.56.
-The arcs between BIS $\rightarrow$ FFFS and RMSSD $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
-The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
-
-```{r}
-# mb(avg30, "injured_2")
-# plotsg2 <- model2network('[nlelg_2][gender][pi][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|pi:gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
-# q2p <- graphviz.plot(plotsg2,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi"), "injured_2", c("stiffness_2", "nlelg_2", "balance_2", "rmssd_2")))
-# 
-# strq2 <- str %>% 
-#   filter(from %in% c("pi", "FFFS_1", "nlelg_2", "stiffness_1", "injured_2", "gender") & to %in% c("balance_2", "injured_2", "nlelg_2", "stiffness_2", "rmssd_2")) %>% 
-#   filter(strength > 0.30) %>% 
-#   filter(direction != 0) %>% 
-#   slice(2:8) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq2$strength)
-# labels[7] <- str_pad(labels[7], side = "left", width = 2)
-# edgenames <- c("pi~stiffness_2", "gender~stiffness_2", "FFFS_1~injured_2",
-#                "injured_2~stiffness_2", "injured_2~balance_2", "injured_2~rmssd_2", "nlelg_2~injured_2")
-# names(labels) <- edgenames
-# nodeRenderInfo(q2p)$fill[c("injured_2", "nlelg_2")] = "gold"
-# edgeRenderInfo(q2p) <- list(lwd = c("injured_2~stiffness_2" = 3, 
-#                                     "nlelg_2~injured_2" = 6,
-#                                     "clevel~balance_1" = 3,
-#                                     "gender~stiffness_2" = 4),
-#                             lty = c("FFFS_1~injured_2" = "dashed",
-#                                     "injured_2~balance_2" = "dashed",
-#                                     "pi~stiffness_2" = "dashed",
-#                                     "injured_2~rmssd_2" = "dashed"))
-# 
-# q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "rmssd_2")] = "tomato"
-# nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-# renderGraph(q2p)
-```
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in RMSSD where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between RMSSD and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table10, results="asis"}
+The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
+An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
+Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively
+Similarly, the arc between hrv $\rightarrow$ FFFS was `r str_cfinder("hrv","FFFS")`.
+The arcs between BIS $\rightarrow$ FFFS and hrv $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
+A separate linear regression examined the interaction between BIS and hrv to be examined.
+
+```{r fig9, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}
+
+include_graphics("figures_doc/Fig9.jpeg")
+
+```
+
+Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95% credible intervals (CrI).
+Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
+There was no clear interaction between hrv and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
+
+```{r table9, results="asis"}
 bmtable %>%
@@
     .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.",
+    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.",
     col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F
+    escape = F,
+    align = "lccc"
   ) %>%
@@
 
-The Markov blanket for the "injured" node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
+The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
-Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of `r query6a[9,1]` over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
-
-```{r table11, results="asis" }
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
+Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
+Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
+With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
+
+```{r fig10, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig10.jpeg")
+```
+
+```{r table10, results="asis" }
 set.seed(1006)
@@
     col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change."
+    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.",
+    align = "ccc"
   ) %>%
@@
 ```{r}
-# plotsg3 <- model2network('[gender][pi][hours][injured|pi:hours][stiffness|gender:injured][nlec|injured]')
-# q3p <- graphviz.plot(plotsg3,
-#                      render = F,
-#                      shape = "rectangle",
-#                      groups = list(c("gender", "pi", "hours"), "injured", c("stiffness", "nlec")))
-# strq3 <- strchange %>% 
-#   filter(from %in% c("pi", "hours", "injured", "gender") & 
-#            to %in% c("stiffness", "nlec", "injured")) %>% 
-#   filter(strength > 0.30) %>% 
-#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
-# 
-# labels <- as.character(strq3$strength)
-# edgenames <- c("pi~injured", "gender~stiffness", "hours~injured",
-#                "injured~stiffness", "injured~nlec")
-# names(labels) <- edgenames
-# nodeRenderInfo(q3p)$fill[c("injured")] = "gold"
-# edgeRenderInfo(q3p) <- list(lwd = c("injured~stiffness" = 4, 
-#                                     "hours~injured" = 5,
-#                                     "gender~stiffness" = 4),
-#                             lty = c("pi~injured" = "dashed"))
-# 
-# q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
-# graph.par(list(edges=list(fontsize = 13)))
-# nodeRenderInfo(q3p)$fill[c("nlec", "stiffness")] = "tomato"
-# nodeRenderInfo(q3p)$fill["injured"] = "gold"
-# renderGraph(q3p)
-# rm(plotsg3, strq3, labels)
-
-```
-
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
-
-```{r table12, results="asis"}
 set.seed(1007)
@@
   arrange(-prob)
+```
+
+Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
+In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).
+
+```{r table11, results="asis"}
 
@@
     caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
+    align = "ccccc"
   ) %>%
@@
 
-(not sure where this would go /if needed at all?).
-The first BN revealed both high stiffness and poor balance resulted in an increased probability of injury.
-However, contrary to previous literature, negative life event stress had a negligible effect on the probability of injury.
-In the second network that modelled changes between time points in the study, increases in muscle stiffness was found to increase the risk of injury.
-In addition, the probability of injury was further increased when there were increases in NLE stress.
-
-(Main discussion starts here..)
-
-Informed by Appaneal and Perna’s (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature [@Johnson2014].
-Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables _across_ time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables _between_ time points (Fig \ref{fig:fig8}).
+The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
+Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
+The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
 
 The first network revealed several links between the injured nodes and other variables in the network.
-For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured_1 and injured_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
-The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for "injured" nodes.
-The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury. 
-In the second network, the highest probability of injury was observed when changes in stiffness _and_ negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury [@Appaneal2014].
-
-<!-- Table \ref{tab:table7} shows the combinations of variables that resulted in the highest and lowest probabilities of injury conditional on all explanatory variables and variables suffixed with _1.
--->
-<!-- Although only the highest and lowest values are shown, there were over 16000 possible combinations of variables states that could influence the risk of injury. -->
-<!-- The adopted approach attempts to take the complex systems view of injury recommended by Bittencourt et al.
-[@Bittencourt2016], where injury is an emergent property from a web of determinants that interact in different ways.
--->
+A combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for “injured” nodes.
+The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete’s risk of injury.
+In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury [@Appaneal2014].
 
@@
 Both "High" levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
-In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
-To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players [@PickeringRodriguez2017].
-The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
-However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness [@Pruyn2015; @Kalkhoven2018].
-Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury [@Butler2003].
-Additionally, high stiffness may only increase the risk of injury if other factors are also present.
+However, high stiffness may only increase the risk of injury if other factors are also present.
 To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
 In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
-
-In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance $\rightarrow$ injured_1 and balance $\rightarrow$  injured_2 respectively.
-Despite the weak arc strength, a "High" balance score, indicating impaired postural stability, was found to increase the probability of injury.
+Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are separately related to injury.
+The BN structures examined in this study provided insight into these two stress-related factors and their relation to injury occurrence.
+
+The findings also facilitated an understanding of the interaction between balance and injury.
+At both injured nodes in the first network, balance was linked to injury.
+Despite the weak arc strength at both injured nodes, a “High” balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014;], was found to increase the probability of injury.
 This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
 Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts [@Paillard2006].
-In the current study, athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
+Athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
 These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition [@Hrysomallis2011].
 
-Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
-In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
-However, negative life event stress had almost no effect on the probability of injury in the first network.
-This finding suggests that the _relative change_ in life events may be more important than the _absolute score_ for life events, despite the latter being commonly used in sports injury research to date.
+Negative life events captured at a single time point have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
+The repeated measures approach combined with the second network analysis employed in this study demonstrated that greater than average increases in negative life event stress between time points increased the probability of being injured during the study period.
+However, negative life event stress had almost no effect on the probability of injury in the first network, which indicated that the relative change in life events may be more important than the absolute score for life events, despite the latter being commonly used in sports injury research to date.
 For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
-Although the absolute score would be "Low", the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect [@Appaneal2014].
-Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed [cf. @Ivarsson2014].
-
-The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a _negative_ valence, as the strongest predictor of injury in Williams and Andersen's [@Williams1998] model [@Ivarsson2017].
+Although the absolute score would be “Low”, the relative increase between time points may have been caused by a significant event in the athlete’s life that had a considerable psychological and physiological effect [@Appaneal2014].
+Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete’s life events to be assessed [cf. @Ivarsson2014].
+
+During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
+The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model [@Ivarsson2017].
 During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg_1 $\rightarrow$ injured_1 and nlelg_2 $\rightarrow$ injured_2 to allow this relationship to be examined more closely.
-When negative life events were "High" the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[3,3] - query3[3,2]` at the injured_2 node.
+Given the reported reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
+When negative life events were “High” the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[which(query3$variable == "nlelg_2"), 3] -  query3[which(query3$variable == "nlelg_2"), 2]` at the injured_2 node.
 One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design. 
-In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
-However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred _since the previous data collection session_, to avoid inflated scores caused by reporting the same event on multiple occasions.
-The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
+In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
+In this study, athletes completed the LESCA at three time points with an approximate four-month interval after baseline.
+Athletes were asked to report any events which had occurred since the previous data collection session in order to avoid inflated scores caused by reporting the same event on multiple occasions.
+While modifications were made to the LESCA to tailor the items to the study cohort, the use of a shorter four-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
 For example, at the second and third time points, 26% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events [cf.
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events  [cf.
 @Fawkner1999].
 
-Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
 Williams and Andersen's [@Williams1998] model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
 Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury [@Petrie1993; @Petrie2014; @Johnson2014].
+The lack of association between major/negative life events and injury reported here may be attributed to athletes in this study having the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
 Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
 
-Of the remaining variables, both FFFS and RMSSD were also linked to injury.
-A weak arc was observed between RMSSD_2 $\rightarrow$ injured_2 (arc strength = 0.30), however no arc was present between RMSSD_1 and injured_1, suggesting the link between RMSSD and injury was not as certain as muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
-Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, "Low" RMSSD increased the probability of injury from 0.17 (RMSSD = "High'') to 0.27 (RMSSD = "Low'').
-This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
+Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, "Low" HRV increased the probability of injury from `r query3[which(query3$variable == "hrv_2"), 3]` ("High" HRV) to `r query3[which(query3$variable == "hrv_2"), 2]` ("Low" HRV).
+This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
 An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
-Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points.
-One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
+Interestingly, the “Low” FFFS score was also related to injuries at subsequent time points.
+One possible explanation for this finding could be that those athletes who reported “Low” FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
 The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-
-In the first network, "High" BIS was associated with "High" FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
-RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
-High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018], which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
-These findings align with the proposed actions of the RST theory [@Corr2016b]. 
-For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems. 
+The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
+The first network reported an association between “High” FFFS and “High” BIS, while the second network found an association between increases in FFFS and increases in BIS.
+High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018].
+This association along with the proposed actions of the RST theory [@Corr2016b] provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network. 
+To elaborate, high levels of BIS are proposed to be the result of goal conflict, an example of which would be simultaneous triggering of the FFFS (avoidance) and BAS (approach) systems.
 The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict [@Corr2016b].
-To extend these findings, the BAS should also be considered. 
-Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
-However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
-In the first network, RI_1 was connected to both BIS_1 and BIS_2, and in both instances, the probability of "High" BIS was increased when RI was also "High". 
-However, the arcs between RI_1 and BIS were weak (< 0.50), and RI represents only one component of the BAS system. 
-Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury. 
-Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
-
-The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury [@Johnson2014].
-The longitudinal repeated measures design of the current study allowed _changes_ over time and between time points to be captured and explored.
-Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
-Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
+In the present study, however, the role of BAS was limited, as evidenced by the initial network structures in which the BAS had limited connectivity with other components of the network. 
+Consequently, a more detailed examination of the role of RST in the injury process is warranted.
+
+This study had a number of strengths.
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury [@Johnson2014].
+The longitudinal repeated measures design of the current study allowed changes over time and between time points to be captured and explored. 
+Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
 Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
-To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
-The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
-Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018].
-
-As with most research, there were several limitations with the present study.
-Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
-Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
-For example, the meaning of a "Low" and "High" value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
-A second limitation was the nature of the physiological measures used in the current study.
-In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
-For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis [@Ross2011].
-
-In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
-Although the current study used a range of measures to capture "stress" from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
-For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], and could help elucidate the relationship between stress and injury.
-Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
-For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
-Both responses would be scored a "-4"; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
+The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
+The BN networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
+
+While the BN’s provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
+Firstly, the choice was made to binarise variables in the first network so only “Low” and “High” states were observed.
+Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+For example, the meaning of a “Low” and “High” value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
+Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
+However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes’ responses.
+For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis [@Ross2011].
+However, the video-capture approach employed in the study ensured an accessible, non-invasive and readily applied method of capturing the respective measure.
+
+The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychological and physiological perspective.
+The variables used in the present study were not definitive.
+Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], could help to further elucidate the relationship between stress and injury.
+Future developments in the capture of life event stress using the LESCA are also warranted.
+Although the LESCA is the most widely used measures of major life events in sports injury research, modifications, including adjustment to the scoring of items are potentially justified to facilitate extended insight into the reported responses.
+For example, the LESCA may negate vastly different psychological and physiological effects between moderately negative and extremely negative events since there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
 Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
 
-In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
-Specifically, muscle stiffness and _increases_ in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
-Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
+To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Baysian Network analytical techniques,  Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Future research combining a repeated measures approach and complex analyses of the interactions between multifaceted stress-related measures are advocated to enhance understanding of the injury occurrence in sport.
 
@@
 
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# Figures
-
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```
-
-\begin{figure}[!h]
-\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
-\label{fig:fig1}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf Session protocol.}
-Outline of the protocol for each data collection.}
-\label{fig:fig4}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-# Tables
+HF, MG, LE and RM developed the concept for the paper.
+HF was the lead author for the paper and designed, collected and analysed that data.
+MG, LE and RM were co-authors on the paper.
+MS provided assistance with data analysis.
+LB provided assistance with data collection.
 

### 2020-07-29 apa_style.tex
progress

@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \keywords{Sports Injury, Stress, Interdisciplinary, Bayesian Network}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{lineno}
@@
 
-\section{Introduction}
+10422 words in text body
+3142 words in reference section
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 Multiple psychological (Slimani et al., 2018), anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes, 2014) and environmental (Meeuwisse et al., 2007) factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009), of which one of the most widely cited was developed by Williams and Anderson (Fig \ref{fig:fig1}; Andersen and Williams, 1988; Williams and Andersen, 1998).
+
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=5.67in]{figures_doc/Fig1}
+
+}
+
+\caption{Stress and injury model (Williams and Andersen, 1998).}\label{fig:fig1}
+\end{figure}
 
@@
 
-\section{Methods}
-
-\subsection{Participants}
+\hypertarget{material-and-methods}{%
+\section{Material and Methods}\label{material-and-methods}}
+
+\hypertarget{ethics-statement}{%
+\subsection{Ethics Statement}\label{ethics-statement}}
+
+Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
+
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
 
@@
 
+\begin{figure}[!h]
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig2}
+
+}
+
+\caption{mBESS positions (A-F).
+Top row, firm surface.
+Bottom row, foam surface.
+Left column, dominant leg stance.
+Middle column, non-dominant leg stance.
+Right column, Tandem leg stance.}\label{fig:fig2}
+\end{figure}
+
 The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials.
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{figures_doc/studydesign_new} \caption{Study design.
+For each time point (T), each box contains the number of participants who completed the data collection (\textit{n}), the measures used for data collection and the approximate date of the data collection.}\label{fig:fig3}
+\end{figure}
+
 For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}).
@@
 Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/sessionprotocol}
+
+}
+
+\caption{Outline of the protocol for each data collection.}\label{fig:fig4}
+\end{figure}
 
@@
 In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/examplenetwork}
+
+}
+
+\caption{A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.}\label{fig:fig5}
+\end{figure}
 
@@
 
-BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017) and offer several benefits over traditional statistical analysis.
+BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017; Ruiz-Pérez et al., 2019) and offer several benefits over traditional statistical analysis.
 For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (Olmedilla et al., 2018; Bittencourt et al., 2016).
@@
 A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
-All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split.
+All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split.
 For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
@@
 To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (Murphy, 2002).
-In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
+In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
 However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
-Variables were separated into two groups; ``explanatory'', for variables that did not change during the study (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
+Variables were separated into two groups; ``explanatory'', for variables that were fixed (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
 Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
@@
 
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -218.31).
+Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -216.47).
 Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (Williams and Andersen, 2007), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
@@
 The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network.
-Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
-RI provided the highest BIC value, compared to RR (-9.07), GDP (-15.06) and I (-16.33).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -894.33) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
+Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
+RI provided the highest BIC value (-3,563.13), compared to RR (-15.98), GDP (-19.26) and I (-19.76).
+Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -900.12) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
 Therefore, only RI was included in the final structure.
-
 Finally, both total score and asymmetry for balance were included in the initial network.
@@
 \hline
-HRV & Root mean squared difference of successive RR intervals & 2.03-4 (Low) & >4-5.94 (High)\\
+HRV & Root mean squared difference of successive RR intervals & 2.03-4.01 (Low) & >4.01-5.94 (High)\\
 \hline
@@
 
-During the study, 26\% of participants reported at least one injury with an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days).
+During the study, 46\% (\emph{n} = 117) of participants reported at least one injury with an average severity of 11 \(\pm\) 31, days (range = 2 - 365 days).
 Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39 {[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and non-contact injuries were more common than contact injuries (male, non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female, non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}).
@@
 
-The first network structure obtained from the data is shown in Fig \ref{fig:fig6}.
-Several of the explanatory variables showed strong connections with independent variables in the network.
-The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.23 to 0.47 when nlebase increased from ``Low'' to ``High''.
-The ind\_team node had strong arcs to hours (0.90) and nlebase (0.84).
+The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
+Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
+The ``ind\_team'' node had strong arcs to hours (0.94) and nlebase (0.78).
 Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
-Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = 0.65, team-based athletes = 0.41).
-The arcs from gender \(\rightarrow\) stiffness\_1 and gender \(\rightarrow\) stiffness\_2 were 0.76, and 0.65 respectively, with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
-The arc from pi \(\rightarrow\) stiffness\_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have ``High'' (0.65) compared to ``Low'' (0.35) stiffness.
-The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
-Arcs were also present between independent variables in the network. Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.68).
-In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.65 for \_1, 0.61 for \_2) compared to ``Low'' (0.32 for \_1, 0.36 for \_2).
-The arc between nlelg \(\rightarrow\) BIS had a strength of 0.62 for nlelg\_1 \(\rightarrow\) BIS\_1, however no arc was present between nlelg\_2 and BIS\_2.
-For nlelg\_1 \(\rightarrow\) BIS\_1, ``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55.
+Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
+The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
+High gender-based connections were reported for the arcs from gender \(\rightarrow\) stiffness\_1 (0.71) and gender \(\rightarrow\) stiffness\_2 (0.43), with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
+Irrespective of sport or gender, strong connections were found between explanatory variables.
+The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.
+
+The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability.
+The arc from pi \(\rightarrow\) stiffness\_1 was 0.57 with athletes who reported an injury in the preceding 12 months being more likely to have ``High'' (0.65) compared to ``Low'' (0.35). stiffness.
+Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.74) were identified.
+In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.64 for \_1, 0.61 for \_2) compared to ``Low'' (0.33 for \_1, 0.37 for \_2).
+The arc between nlelg \(\rightarrow\) BIS had a strength of 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.37 nlelg\_2 and BIS\_2.
+``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.38 to 0.58 for nlelg\_2 \(\rightarrow\) BIS\_2.
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{apa_style_files/figure-latex/fig6-1.pdf}
+\caption{\label{fig:fig6}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
+The strength of each arc is shown graphically by the style of the arc.
+Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+\end{figure}
 
@@
 
-The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1), is shown in Fig \ref{fig:fig7}.
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
 The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
 
-The conditional probability query (CPQ) for injured\_1 in the ``injured'' state for all variables that were directly linked to injured\_1 is shown in Table \ref{tab:table5}.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig7}
+
+}
+
+\caption{Markov blanket of injured\_1. Arc strengths are included as arc labels.}\label{fig:fig7}
+\end{figure}
+
+The CPQ for injured\_1 in the ``injured'' state for all variables that were linked to injured\_1 is shown in Table \ref{tab:table5}.
 The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when stiffness was ``High'' compared to ``Low''.
@@
 When all variables were in the ``Low'' state the probability of ``injured'' was approximately 0.04
-Negative life events only had a substantial effect on injured\_1 when all other variable were fixed to ``Low''.
-In this instance the probability of injured\_1 being ``injured'' rose from 0.04 to 0.19, when negative life events was in the ``Low'' and ``High'' states respectively.
+
+Negative life event stress had a negligible effect on the probability of injury, only influencing injured\_1 when all other variable were fixed to ``Low''.
+In this instance, the probability of injured\_1 being ``injured'' rose marginally from 0.04 to 0.19, when negative life events was in the ``Low'' and ``High'' states, respectively.
 
@@
 \centering
-\begin{tabular}[t]{r|l|l|l|l|l}
+\begin{tabular}[t]{c|c|c|c|c|c}
 \hline
@@
 \hline
-\hspace{1em}0.43 & national\_international & High & Low & High & High\\
+\hspace{1em}0.44 & national\_international & High & Low & High & High\\
 \hline
@@
 
-The Markov blanket for injured\_2 is shown in Fig \ref{fig:fig8} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and rmssd\_2.
+The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2.
 The arc between stiffness\_2 \(\rightarrow\) injured\_2 was comparable to the arc between stiffness\_1 \(\rightarrow\) injured\_1.
-Very weak arcs (0.3) between injured\_2 \(\rightarrow\) balance\_2 and injured\_2 \(\rightarrow\) rmssd\_2 were also present in the Markov blanket for injured\_2.
-Results of the first query for injured\_2 in the ``injured'' state are presented in table \ref{tab:table8}.
+Very weak arcs (0.3) between injured\_2 \(\rightarrow\) balance\_2 and injured\_2 \(\rightarrow\) hrv\_2 were also present in the Markov blanket for injured\_2.
 Similar to injured\_1, stiffness\_2 doubled the probability of injured\_2 being ``injured'' from 0.13 in the ``Low'' state to 0.27 in the ``High'' state.
 FFFS\_1 in the ``Low'' state increased probability of injured\_2 being ``injured'' by 0.19 compared to the ``High'' state.
-``High'' negative life events decreased the probability of injury from 0.24 to 0.19.
+``High'' negative life events decreased the probability of injury from 0.26 to 0.17.
 
@@
 
-\caption{\label{tab:table8}Probability of injured\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\_2.}
+\caption{\label{tab:table7}Probability of injured\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\_2.}
 \centering
-\begin{tabular}[t]{l|l|l}
+\begin{tabular}[t]{c|c|c}
 \hline
@@
 \hline
-FFFS\_1 & 0.3 & 0.11\\
-\hline
-nlelg\_2 & 0.24 & 0.19\\
-\hline
-rmssd\_2 & 0.24 & 0.19\\
+FFFS\_1 & 0.30 & 0.11\\
+\hline
+hrv\_2 & 0.26 & 0.17\\
+\hline
+nlelg\_2 & 0.23 & 0.19\\
 \hline
@@
 
-The conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table9}.
-Again, only the three lowest and highest probabilities are shown (complete results in S5 Table).
-The combination ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.52.
-With all other variables held in the ``High'' state, the probability of injured\_2 being ``injured'' rose from 0.14 to 0.34 when FFFS\_1 was in the ``Low'' compared to ``High'' state.
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig8}
+
+}
+
+\caption{Markov blanket for injured\_2.}\label{fig:fig8}
+\end{figure}
+
+The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
+The combination of ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.53.
+With all other variables held in the ``High'' state, the probability of injured\_2 being ``injured'' rose from 0.15 to 0.35 when FFFS\_1 was in the ``Low'' compared to ``High'' state.
 The combination of ``Low'' stiffness, ``Low'' balance and ``High'' FFFS resulted in the lowest probability of injured\_2 being ``injured''.
@@
 
-\caption{\label{tab:table9}Highest and lowest probability of injured\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_2.}
+\caption{\label{tab:table8}Highest and lowest probability of injured\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_2.}
 \centering
-\begin{tabular}[t]{l|l|l|l|l|l}
-\hline
-\textbf{Probability} & \textbf{FFFS\_1} & \textbf{nlelg\_2} & \textbf{stiffness\_2} & \textbf{rmssd\_2} & \textbf{balance\_2}\\
+\begin{tabular}[t]{c|c|c|c|c|c}
+\hline
+\textbf{Probability} & \textbf{FFFS\_1} & \textbf{nlelg\_2} & \textbf{stiffness\_2} & \textbf{hrv\_2} & \textbf{balance\_2}\\
 \hline
@@
 \hline
-\hspace{1em}0.52 & Low & Low & High & Low & High\\
-\hline
-\hspace{1em}0.45 & Low & High & High & Low & High\\
-\hline
-\hspace{1em}0.43 & Low & Low & High & High & High\\
+\hspace{1em}0.53 & Low & Low & High & Low & High\\
+\hline
+\hspace{1em}0.46 & Low & High & High & Low & High\\
+\hline
+\hspace{1em}0.41 & Low & Low & High & High & High\\
 \hline
@@
 
-The network for changes within variables is presented in Fig \ref{fig:fig9}.
-An arc between BIS \(\rightarrow\) FFFS with strength 1 was present in the network.
-Arcs between clevel \(\rightarrow\) BIS and gender \(\rightarrow\) stiffness had a strength of 0.60.
-The arc between RMSSD \(\rightarrow\) FFFS was 0.56.
-The arcs between BIS \(\rightarrow\) FFFS and RMSSD \(\rightarrow\) FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
-The use of a separate linear regression enabled the interaction between BIS and RMSSD to be examined.
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table10} and include 95\% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS (\(b\) = -0.14, 95\% CrI = {[}-0.20, -0.08{]}), whereas positive changes in RMSSD where associated with decreased changes in FFFS (\(b\) = 0.4, 95\% CrI = {[}0.34, 0.45{]}).
-There was no clear interaction between RMSSD and BIS (\(b\) = -0.04, 95\% CrI = {[}-0.10, 0.02{]}).
+The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
+An arc between BIS \(\rightarrow\) FFFS with strength 1.00 was present in the network.
+Arcs between clevel \(\rightarrow\) BIS and gender \(\rightarrow\) stiffness had a strengths of 0.60 and 0.56 respectively
+Similarly, the arc between hrv \(\rightarrow\) FFFS was 0.58.
+The arcs between BIS \(\rightarrow\) FFFS and hrv \(\rightarrow\) FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
+A separate linear regression examined the interaction between BIS and hrv to be examined.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig9}
+
+}
+
+\caption{Network structure of the changes within variables between time points.}\label{fig:fig9}
+\end{figure}
+
+Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95\% credible intervals (CrI).
+Increases in BIS were associated with increases in FFFS (\(b\) = -0.19, 95\% CrI = {[}-0.25, -0.13{]}), whereas positive changes in hrv where associated with decreased changes in FFFS (\(b\) = 0.41, 95\% CrI = {[}0.36, 0.47{]}).
+There was no clear interaction between hrv and BIS (\(b\) = -0.02, 95\% CrI = {[}-0.08, 0.03{]}).
 
@@
 
-\caption{\label{tab:table10}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and RMSSD.}
+\caption{\label{tab:table9}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.}
 \centering
-\begin{tabular}[t]{l|r|r|l}
+\begin{tabular}[t]{l|c|c|c}
 \hline
@@
 \hline
-Intercept & 0.00 & 0.03 & [-0.06, 0.05]\\
-\hline
-BIS & 0.40 & 0.03 & [0.34, 0.45]\\
-\hline
-rmssd & -0.14 & 0.03 & [-0.20, -0.08]\\
-\hline
-BIS:rmssd & -0.04 & 0.03 & [-0.10, 0.02]\\
+Intercept & 0.00 & 0.03 & [-0.05, 0.06]\\
+\hline
+BIS & 0.41 & 0.03 & [0.36, 0.47]\\
+\hline
+hrv & -0.19 & 0.03 & [-0.25, -0.13]\\
+\hline
+BIS:hrv & -0.02 & 0.03 & [-0.08, 0.03]\\
 \hline
@@
 
-The Markov blanket for the ``injured'' node contained previous injury, gender, hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
+The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table11}.
-Changes in both nlec and stiffness 1SD above the mean change resulted in a probability of being injured of 0.71 over the duration of the study.
-With stiffness held at the mean change, the probability of ``injured'' rose from 0.35 to 0.64 with nlec at 1 SD below an 1 SD above respectively.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
+Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress.
+Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (0.71) over the duration of the study.
+With stiffness held at the mean change, the probability of ``injured'' rose notably from 0.35 to 0.64 with nlec at 1 SD below an 1 SD above respectively.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig10}
+
+}
+
+\caption{Markov blanket for the injured node in the network reflecting changes within variables between time points.}\label{fig:fig10}
+\end{figure}
 
@@
 
-\caption{\label{tab:table11}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
+\caption{\label{tab:table10}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
 \centering
-\begin{tabular}[t]{r|l|l}
+\begin{tabular}[t]{c|c|c}
 \hline
@@
 
-Table \ref{tab:table12} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
 The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.12).
+In contrast, below average changes in nlec and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.11).
 
@@
 
-\caption{\label{tab:table12}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+\caption{\label{tab:table11}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
 \centering
-\begin{tabular}[t]{r|l|l|l|l}
+\begin{tabular}[t]{c|c|c|c|c}
 \hline
@@
 
-(not sure where this would go /if needed at all?).
-The first BN revealed both high stiffness and poor balance resulted in an increased probability of injury.
-However, contrary to previous literature, negative life event stress had a negligible effect on the probability of injury.
-In the second network that modelled changes between time points in the study, increases in muscle stiffness was found to increase the risk of injury.
-In addition, the probability of injury was further increased when there were increases in NLE stress.
-
-(Main discussion starts here..)
-
-Informed by Appaneal and Perna's (2014) extension to the widely cited Williams and Anderson's (1998) stress-injury model, the purpose of this study was to examine how psychosocial stress-related factors, and physiological stress-related markers may interact and act synergistically to increase the risk of injury; therein addressing several of the limitations of the sport injury literature (Johnson et al., 2014).
-Potential relationships between psychosocial stress-related factors, physiological stress-related markers of stress and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables \emph{across} time points in the study (Fig \ref{fig:fig5}), and the second network used differential equations to model the changes in variables \emph{between} time points (Fig \ref{fig:fig8}).
+The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
+Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
+The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
 
 The first network revealed several links between the injured nodes and other variables in the network.
-For example, Fig \ref{fig:fig6} and Fig \ref{fig:fig7} show the Markov blankets for the injured\_1 and injured\_2 nodes in the first network and include all the variables that had a direct effect on the probability of injury.
-The combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for ``injured'' nodes.
-The presence of these factors at both injured nodes indicates that the combined action of these variables is important for determining an athlete's risk of injury.
-In the second network, the highest probability of injury was observed when changes in stiffness \emph{and} negative life events were both greater than average (Table \ref{tab:table11}), indicating that the combination of changes in psychological and physiological measures of stress may combine additively to increase the risk of injury (Appaneal and Perna, 2014).
+A combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for ``injured'' nodes.
+The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete's risk of injury.
+In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury (Appaneal and Perna, 2014).
 
@@
 Both ``High'' levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
-In this study, a novel hand-held device (MyotonPRO) was used to measure muscle stiffness.
-To date, one of only a small number of studies that have used the MyotonPRO to explore the relationship between muscle stiffness and sports injury, found that increased muscle stiffness in the soleus and Achilles tendon was related to increased injury incidence in elite level netball players (Pickering-Rodriguez et al., 2017).
-The results of the current study build on these findings, with a larger sample of athletes from a range of different sports, strengthening the evidence for a relationship between higher levels of muscle stiffness and injury.
-However, high levels of muscle stiffness, as measured by the MyotonPRO, have also been found to be related to improved performance, with elite level athletes having increased lower extremity stiffness (Pruyn et al., 2015; Kalkhoven and Watsford, 2017).
-Collectively, therefore, these findings suggest that while muscle stiffness plays a vital role in performance, increased levels of stiffness also increase the risk of injury, with each athlete likely to have an optimum level of stiffness that maximises performance while minimising the risk of injury (Butler et al., 2003).
-Additionally, high stiffness may only increase the risk of injury if other factors are also present.
+However, high stiffness may only increase the risk of injury if other factors are also present.
 To elaborate, the combination of high stiffness \emph{and} poor balance was found to result in the greatest probability of injury.
 In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance (Romero-Franco et al., 2014; Trojian and McKeag, 2006) and muscle stiffness (Butler et al., 2003; Pickering-Rodriguez et al., 2017) are related to injury individually; however this study has demonstrated how these two factors may have an additive effect in relation to injury occurrence.
-
-In addition to stiffness, balance is also linked to injury at both injured nodes in the first network, however the strength of the arc was only 0.35 and 0.30 from balance \(\rightarrow\) injured\_1 and balance \(\rightarrow\) injured\_2 respectively.
-Despite the weak arc strength, a ``High'' balance score, indicating impaired postural stability, was found to increase the probability of injury.
+Several studies have identified how balance (Romero-Franco et al., 2014; Trojian and McKeag, 2006) and muscle stiffness (Butler et al., 2003; Pickering-Rodriguez et al., 2017) are separately related to injury.
+The BN structures examined in this study provided insight into these two stress-related factors and their relation to injury occurrence.
+
+The findings also facilitated an understanding of the interaction between balance and injury.
+At both injured nodes in the first network, balance was linked to injury.
+Despite the weak arc strength at both injured nodes, a ``High'' balance score, which is considered indicative of impaired postural stability {[}Romero-Franco et al.
+(2014);{]}, was found to increase the probability of injury.
 This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (Riemann et al., 1999; Romero-Franco et al., 2014; Trojian and McKeag, 2006).
 Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (Paillard et al., 2006).
-In the current study, athletes who competed at a higher level were also more likely to have good balance (``Low'' balance), compared to their lower level counterparts.
+Athletes who competed at a higher level were also more likely to have good balance (``Low'' balance), compared to their lower level counterparts.
 These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (Hrysomallis, 2011).
 
-Of the psychological variables in the study, negative life events have previously been reported to be most strongly associated with injury (Williams and Andersen, 2007; Ivarsson et al., 2017).
-In the current study, the second network revealed that greater than average increases in negative life event stress increased the probability of being injured during the study period.
-However, negative life event stress had almost no effect on the probability of injury in the first network.
-This finding suggests that the \emph{relative change} in life events may be more important than the \emph{absolute score} for life events, despite the latter being commonly used in sports injury research to date.
+Negative life events captured at a single time point have previously been reported to be most strongly associated with injury (Williams and Andersen, 2007; Ivarsson et al., 2017).
+The repeated measures approach combined with the second network analysis employed in this study demonstrated that greater than average increases in negative life event stress between time points increased the probability of being injured during the study period.
+However, negative life event stress had almost no effect on the probability of injury in the first network, which indicated that the relative change in life events may be more important than the absolute score for life events, despite the latter being commonly used in sports injury research to date.
 For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400\% increase in their life event score.
-Although the absolute score would be ``Low'', the relative increase could have been caused by a significant event in the athlete's life, that could have a considerable psychological and physiological effect (Appaneal and Perna, 2014).
-Future research should therefore consider study designs and appropriate analysis methods that enable relative changes in an individual athlete's life events to be assessed (cf.~Ivarsson et al., 2014).
-
-The finding that negative life events had almost no impact on the probability of injury in the first network was surprising given that the majority of research has consistently identified major life events, particularly those events with a \emph{negative} valence, as the strongest predictor of injury in Williams and Andersen's (Williams and Andersen, 1998) model (Ivarsson et al., 2017).
+Although the absolute score would be ``Low'', the relative increase between time points may have been caused by a significant event in the athlete's life that had a considerable psychological and physiological effect (Appaneal and Perna, 2014).
+Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed (cf.~Ivarsson et al., 2014).
+
+During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
+The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model (Ivarsson et al., 2017).
 During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-However, given the strength of the literature indicating that negative life events are related to injury, an arc was fixed between nlelg\_1 \(\rightarrow\) injured\_1 and nlelg\_2 \(\rightarrow\) injured\_2 to allow this relationship to be examined more closely.
-When negative life events were ``High'' the probability of injury showed a negligible change at the injured\_1 node and decreased by -0.05 at the injured\_2 node.
+Given the reported reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
+When negative life events were ``High'' the probability of injury showed a negligible change at the injured\_1 node and decreased by -0.04 at the injured\_2 node.
 One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
-In the original LESCA instructions, participants are asked to report major life events that have occurred over the previous 12-months (Petrie, 1992).
-However, in the current study, participants completed the LESCA at approximately 4-month intervals after baseline and were therefore asked to report any events which had occurred \emph{since the previous data collection session}, to avoid inflated scores caused by reporting the same event on multiple occasions.
-The reduced 4-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
+In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months (Petrie, 1992).
+In this study, athletes completed the LESCA at three time points with an approximate four-month interval after baseline.
+Athletes were asked to report any events which had occurred since the previous data collection session in order to avoid inflated scores caused by reporting the same event on multiple occasions.
+While modifications were made to the LESCA to tailor the items to the study cohort, the use of a shorter four-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
 For example, at the second and third time points, 26\% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the original 12-months than a measure that captures minor life events (cf.~Fawkner et al., 1999).
-
-Another possible explanation for the findings for major/negative life events is that participants in the study may have had access to the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events (cf.~Fawkner et al., 1999).
+
 Williams and Andersen's (Williams and Andersen, 1998) model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
 Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (Petrie, 1993; Petrie et al., 2014; Johnson et al., 2014).
+The lack of association between major/negative life events and injury reported here may be attributed to athletes in this study having the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
 Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
 
-Of the remaining variables, both FFFS and RMSSD were also linked to injury.
-A weak arc was observed between RMSSD\_2 \(\rightarrow\) injured\_2 (arc strength = 0.30), however no arc was present between RMSSD\_1 and injured\_1, suggesting the link between RMSSD and injury was not as certain as muscle stiffness and balance, where stronger arcs existed at both of the injured nodes.
-Despite the uncertainty regarding the relationship between injury and RMSSD in the first network, ``Low'' RMSSD increased the probability of injury from 0.17 (RMSSD = ``High'\,') to 0.27 (RMSSD =''Low'\,').
-This finding is consistent with previous research that has found reduced RMSSD to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (Williams et al., 2017; Bellenger et al., 2016; Gisselman et al., 2016).
+Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, ``Low'' HRV increased the probability of injury from 0.17 (``High'' HRV) to 0.26 (``Low'' HRV).
+This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (Williams et al., 2017; Bellenger et al., 2016; Gisselman et al., 2016).
 An arc between FFFS\_1 and injured\_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the ``High'' and ``Low'' states respectively.
@@
 The RST theory proposes that higher levels of FFFS increase avoidance motivation (Corr et al., 2016), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-
-In the first network, ``High'' BIS was associated with ``High'' FFFS, while in the second network, increases in BIS were associated with increases in FFFS.
-RST proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (Corr, 2013).
-High levels of anxiety and anticipation of stressful situations have been associated with reductions in HRV indices including RMSSD (Chalmers et al., 2014; Pulopulos et al., 2018), which is supported by the negative relationship between FFFS and RMSSD in the second network (Table \ref{tab:table10}).
-These findings align with the proposed actions of the RST theory (Corr et al., 2016).
-For example, high levels of BIS are proposed to be the result of goal conflict between the FFFS (avoidance) and BAS (approach) systems.
+The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (Corr, 2013).
+The first network reported an association between ``High'' FFFS and ``High'' BIS, while the second network found an association between increases in FFFS and increases in BIS.
+High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD (Chalmers et al., 2014; Pulopulos et al., 2018).
+This association along with the proposed actions of the RST theory (Corr et al., 2016) provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network.
+To elaborate, high levels of BIS are proposed to be the result of goal conflict, an example of which would be simultaneous triggering of the FFFS (avoidance) and BAS (approach) systems.
 The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (Corr et al., 2016).
-To extend these findings, the BAS should also be considered.
-Specifically, to establish how the BAS and FFFS interact, and how these two systems affect the BIS.
-However, in the current study, initial network structures revealed the BAS sub-scales to have limited connectivity with other measures in the network, therefore only one of the BAS sub-scales (RI) was included in the final network structure.
-In the first network, RI\_1 was connected to both BIS\_1 and BIS\_2, and in both instances, the probability of ``High'' BIS was increased when RI was also ``High''.
-However, the arcs between RI\_1 and BIS were weak (\textless{} 0.50), and RI represents only one component of the BAS system.
-Other BAS factors such as impulsivity may be more closely related to risk-taking behaviours and may reveal additional links to sports injury.
-Therefore, a more detailed examination of the different elements of RST, and specifically the BAS in relation to injury occurrence is warranted.
-
-The current research had several strengths, including the longitudinal repeated measures design and modelling approach.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of and capture the dynamic nature of the variables that are associated with injury (Johnson et al., 2014).
-The longitudinal repeated measures design of the current study allowed \emph{changes} over time and between time points to be captured and explored.
-Another significant strength of the current research was the interdisciplinary approach, which enabled an examination of the complex interplay between psychosocial and physiological markers of stress.
-Although there are unique and significant challenges with research employing longitudinal repeated measures designs, they provide for far more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes.
+In the present study, however, the role of BAS was limited, as evidenced by the initial network structures in which the BAS had limited connectivity with other components of the network.
+Consequently, a more detailed examination of the role of RST in the injury process is warranted.
+
+This study had a number of strengths.
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury (Johnson et al., 2014).
+The longitudinal repeated measures design of the current study allowed changes over time and between time points to be captured and explored.
+Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
 Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (Bittencourt et al., 2016) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
-To address this issue, Bayesian networks (BN) were used to more closely align with the complex, multifactorial nature of injury.
-The networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
-Consequently, BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (Olmedilla et al., 2018).
-
-As with most research, there were several limitations with the present study.
+The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
+The BN networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
+
+While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (Olmedilla et al., 2018), a number of assumptions were made that potentially limited the approach employed in the study.
 Firstly, the choice was made to binarise variables in the first network so only ``Low'' and ``High'' states were observed.
-Although binarising variables is a common procedure in Bayesian network analysis and has several advantages, Qian and Miltner Qian and Miltner (2015) highlighted that both a loss of statistical accuracy and potential difficulty in subsequent interpretation of the model may arise when following a binarising procedure.
-For example, the meaning of a ``Low'' and ``High'' value in the current study is only meaningful for the population that was studied, and there could be additional levels within each category that were not investigated.
-A second limitation was the nature of the physiological measures used in the current study.
-In order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection; however, some of these measures may not have been sensitive enough to detect more subtle variation in athletes.
-For example, postural stability could have been assessed with the use of a force plate, which is considered the gold standard, to provide detailed data and enable a more fine-grained analysis (Ross et al., 2011).
-
-In addition to the future directions already outlined, the findings from the current study offer several avenues for future research.
-Although the current study used a range of measures to capture ``stress'' from both a psychological and physiological perspective, there may be additional measures available that could provide further insight into the relationship between stress-related factors and sports injury.
-For example, stress hormones such as cortisol have been found to be a marker of both psychological and training-related stress (Appaneal and Perna, 2014; Perna and McDowell, 1995), and could help elucidate the relationship between stress and injury.
-Additionally, although the LESCA is the most widely used measures of major life events in sports injury research, the current study found several limitations with using the LESCA in a repeated measures design, including how the items were scored.
-For example, there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
-Both responses would be scored a ``-4''; however, there could be vastly different psychological and physiological effects between moderately negative and extremely negative events.
+Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner Qian and Miltner (2015) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+For example, the meaning of a ``Low'' and ``High'' value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
+Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
+However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes' responses.
+For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis (Ross et al., 2011).
+However, the video-capture approach employed in the study ensured an accessible, non-invasive and readily applied method of capturing the respective measure.
+
+The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychological and physiological perspective.
+The variables used in the present study were not definitive.
+Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress (Appaneal and Perna, 2014; Perna and McDowell, 1995), could help to further elucidate the relationship between stress and injury.
+Future developments in the capture of life event stress using the LESCA are also warranted.
+Although the LESCA is the most widely used measures of major life events in sports injury research, modifications, including adjustment to the scoring of items are potentially justified to facilitate extended insight into the reported responses.
+For example, the LESCA may negate vastly different psychological and physiological effects between moderately negative and extremely negative events since there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
 Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
 
-In conclusion, the purpose of this research was to explore the multifaceted nature of the stress-injury relationship, and several psychosocial and physiological markers were found to combine and exacerbate the risk of injury.
-Specifically, muscle stiffness and \emph{increases} in negative life event stress were identified as strong predictors of injury, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
-Taken together, the interdisciplinary approach coupled with the advanced analytical techniques used and complex systems framework has provided a novel examination of the stress-injury relationship that has addressed many of the limitations identified in previous research.
+To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Baysian Network analytical techniques, Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Future research combining a repeated measures approach and complex analyses of the interactions between multifaceted stress-related measures are advocated to enhance understanding of the injury occurrence in sport.
 
@@
 
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
-
-\hypertarget{supplemental-data}{%
-\section{Supplemental Data}\label{supplemental-data}}
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
-
-\begin{figure}[!h]
-\caption{{\bf Stress and injury model (Williams and Andersen, 1998).}}
-\label{fig:fig1}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf mBESS positions (A-F).}
-Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.}
-\label{fig:fig2}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf Study design.}
-For each time point (T), each box contains the number of participants who completed the data collection (N), the measures used for data collection and the approximate date of the data collection.}
-\label{fig:fig3}
-\end{figure}
-
-\begin{figure}[!h]
-\caption{{\bf Session protocol.}
-Outline of the protocol for each data collection.}
-\label{fig:fig4}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Example network.}
-A simple discrete network contain nodes, possible states of the nodes and the arcs connecting nodes.}
-\label{fig:fig5}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Network structure.}
-The full Bayesian network structure was plotted using the strength.plot function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
-\label{fig:fig6}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket of injured\_1.}
-Arc strengths are included as arc labels.}
-\label{fig:fig7}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for injured\_2.} }
-\label{fig:fig8}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Change network.}
-Network structure of the changes within variables between time points.}
-\label{fig:fig9}
-\end{figure}
-
-\begin{figure}[H]
-\caption{{\bf Markov blanket for the injured node in the network reflecting changes within variables between time points.}}
-\label{fig:fig10}
-\end{figure}
-
-\hypertarget{tables}{%
-\section{Tables}\label{tables}}
+HF, MG, LE and RM developed the concept for the paper.
+HF was the lead author for the paper and designed, collected and analysed that data.
+MG, LE and RM were co-authors on the paper.
+MS provided assistance with data analysis.
+LB provided assistance with data collection.
 
@@
 
-\leavevmode\hypertarget{ref-Kalkhoven2018}{}%
-Kalkhoven, J.
-T., and Watsford, M.
-L.
-(2017).
-The relationship between mechanical stiffness and athletic performance markers in sub-elite footballers.
-\emph{Journal of Sports Sciences} 36, 1022--1029.
-doi:\href{https://doi.org/10.1080/02640414.2017.1349921}{10.1080/02640414.2017.1349921}.
-
 \leavevmode\hypertarget{ref-Koller2009}{}%

### 2020-07-29 r-files/functions.R

@@ -46,6 +46,7 @@ networkmaker <- function(names){
   cl = makeCluster(4, type = "SOCK")
   clusterSetRNGStream(cl, 40)
   
+  set.seed(155)
   str = boot.strength(tallc, algorithm = "tabu",
                       algorithm.args = list(blacklist = bl, 
                                             #whitelist = wl,
@@ -155,7 +156,7 @@ newprobtable <- function(tallc, vars, outcome, state, model, repeats = 500000) {
     lapply(tallc[, (names(tallc) %in% vars)], levels) 
   } else {
     all.levels <- tallc %>% 
-      select(vars) %>% 
+      select(all_of(vars)) %>% 
       sapply(levels) %>% 
       as_tibble() 
   } # makes the code work for when only one variable is used as evidence

### 2020-07-29 r-files/methods_code.R

@@ -1,13 +1,13 @@
 ## Methods
-pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect)
-library(readxl)
-library(janitor)
-library(irr)
-library(lavaan)
-library(caret)
-library(parallel)
-library(zoo)
-library(sjmisc)
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect,
+               readxl,
+               janitor,
+               irr,
+               lavaan,
+               caret,
+               parallel,
+               zoo,
+               sjmisc)
 ## Table 1
 
 x <- read_csv("data/new_cleandata_inj.csv")
@@ -63,7 +63,7 @@ npars <- parchar %>%
   ) %>%
   ungroup()
 
-rm(clevel)
+rm(clevel,x)
 
 ## RST reliability
 
@@ -97,6 +97,7 @@ i=~V22_IMP+V27_IMP+V28_IMP+V38_IMP+V40_IMP+V44_IMP+V51_IMP"
 fullmodel <- str_remove_all(fullmodel, "_")
 
 # example of cfa model
+set.seed(123)
 fit <- cfa(fullmodel, data = d)
 sl <- standardizedSolution(fit)
 sl <- sl$est.std[sl$op == "=~"]
@@ -142,7 +143,8 @@ icc <- irr::icc(d, model = "twoway", type = "agreement")
 
 
 data <- read_csv("data/fulldatajoined.csv") %>%
-  filter(time != 4) # remove last time point as only questionnaire data is available
+  filter(time != 4) %>% 
+  rename(hrv= rmssd)# remove last time point as only questionnaire data is available
 
 ####################### Cleaning up variables + impute #################################
 
@@ -181,7 +183,7 @@ voi <- data %>%
     stiffness,
     frequency,
     decrement,
-    rmssd,
+    hrv,
     sdnn,
     meanHR,
     bis,
@@ -202,20 +204,20 @@ ids <- data %>%
 
 missingcount <- tibble(
   var = c("hrv", "myoton"),
-  missing = c(map(voi, ~ sum(is.na(.)))$rmssd, map(voi, ~ sum(is.na(.)))$stiffness),
+  missing = c(map(voi, ~ sum(is.na(.)))$hrv, map(voi, ~ sum(is.na(.)))$stiffness),
   total = c(668, 668)
 ) %>%
[…]

### 2020-07-29 r-files/results_code.R

@@ -80,7 +80,7 @@ library(parallel)
 
 cl = makeCluster(8, type = "SOCK")
 clusterSetRNGStream(cl, 30)
-
+set.seed(44)
 str <- boot.strength(
   tallc,
   algorithm = "tabu",
@@ -94,11 +94,9 @@ str <- boot.strength(
 
 
 stopCluster(cl)
-
 ## get average network with 30% cut off value
 avg30 <- averaged.network(str, threshold = 0.30)
 fitted <- bn.fit(avg30, tallc)
-
 ## plot the network
 gR <- strength.plot(
   avg30,
@@ -113,8 +111,8 @@ nodeRenderInfo(gR)$fill[g1] = "deepskyblue"
 nodeRenderInfo(gR)$fill[g2] = "tomato"
 nodeRenderInfo(gR)$fill["injured_1"] = "gold"
 nodeRenderInfo(gR)$fill["injured_2"] = "gold"
-#gR <- renderGraph(gR)
-
+gR <- renderGraph(gR)
+renderGraph(gR)
 # tiff("main_network.tif", res = 300, width = 15, height = 10, units = "in")
 #renderGraph(gR)
 # dev.off()
@@ -137,7 +135,7 @@ networkdatacont <- datanet %>%
     nlec,
     RI,
     BIS,
-    rmssd,
+    hrv,
     balance,
     FFFS
   ) %>%
@@ -200,7 +198,7 @@ bl <- expand.grid(
 cl = makeCluster(4, type = "SOCK")
 clusterSetRNGStream(cl, 44)
 
-
+set.seed(55)
 strchange <- boot.strength(
   diffg,
   algorithm = "tabu",
@@ -223,74 +221,175 @@ change <- strength.plot(
   layout = "dot"
 )
 nodeRenderInfo(change)$fill[indivars] = "tomato"
+
 nodeRenderInfo(change)$fill["injured"] = "gold"
-# 
-# tiff("change_network.tif", res = 300, width = 15, height = 10, units = "in")
-# renderGraph(change)
-# dev.off()
+
+jpeg("./figures_doc/Fig9.jpeg", res = 300, width = 15, height = 10, units = "in")
+renderGraph(change)
+dev.off()
 fittedcont <- bn.fit(avgcont, diffg)
 
 
 ## Bayesian lm
 
[…]

### 2020-08-14 manuscript.Rmd
pre figure removeal

--- 2020-07-29 apa_style.Rmd
@@
     affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
+    corresponding : yes    
     address       :  "Cyncoed Road, Cardiff, CF23 6XD"
@@
 
+authornote: |
+    12961 words in text body and refence section
+
+    Number of tables: 11
+
+    Number of figures: 10
+
+    Format: British English
+
 abstract: |
@@
 
-keywords          : "Sports Injury, Stress, Interdisciplinary, Bayesian Network"
+keywords          : "Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology"
 
@@
 
-floatsintext      : yes
+floatsintext      : no
 figurelist        : no
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
- source("r-files/functions.R")
- source("r-files/methods_code.R")
- source("r-files/results_code.R")
-#load("all_data.RData", envir = myEnv)
+ # source("r-files/functions.R")
+ # source("r-files/methods_code.R")
+ # source("r-files/results_code.R")
+
+load("all_data.RData")
 #myEnv <- new.env()
@@
 
-10422 words in text body
-3142 words in reference section
-
 # Introduction
@@
 Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
-Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.
+Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].
+However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.
 
@@
 However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
-Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
 Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
@@
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table). 
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (supplementary table 1). 
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
@@
 The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
-Further details regarding the revised RST are in S1 Appendix.
+Further details regarding the revised RST are in supplementary Appendix 1.
 
@@
 ```{r fig3, fig.cap = 'Study design.
-For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width="100%"}
-
-include_graphics("figures_doc/studydesign_new.png")
+For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width = "100%"}
+
+include_graphics("figures_doc/Fig3.jpeg")
 ```
@@
 ```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
-include_graphics("figures_doc/sessionprotocol.png")
+include_graphics("figures_doc/Fig4.jpeg")
 ```
@@
 A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988].
+A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [@Pearl1988].
 However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
@@
 In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
-
-```{r fig5, fig.cap="A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.", fig.pos="H", out.width="100%", fig.align='center'}
-
-include_graphics("figures_doc/examplenetwork.png")
-
-```
 
@@
 A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
-In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
+In the current study, the _caret_ R package [@Kuhn2008] was used to impute the missing values. 
 A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
@@
 Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
-Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively.
-Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
 In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
@@
 
-```{r table2, results='asis'}
-exampledata <- tibble(
-  Participant = c("1", "1", "2", "3", "3"),
-  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
-  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
-)
-
-kable(
-  exampledata,
-  caption = "Example of the data arrangement used for the network.",
-  escape = FALSE
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-```
-
 #### Preliminary network structures
@@
 
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
+Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (nleleg = `r round(le[1,2], 2)`, tlelg only = `r round(le[2,2],2)`, tlelg and nlelg = `r round(le[3,2],2)`).
 Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
@@
 Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
-RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
+RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]`), GDP (`r rstmodel[4,2]`) and I (`r rstmodel[5,2]`).
+Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
 Therefore, only RI was included in the final structure.
@@
 In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
-Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in supplementary table 1 and supplementary table 3 respectively.
 
@@
 
-Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
-
 For the second network, change scores for continuous variables between  T1 $\rightarrow$  T2 and  T2 $\rightarrow$  T3 were standardized to allow relative changes between variables to be compared.
@@
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then examined using a Bayesian linear regression models with the _brms_ package [@Burkner2017a] to determine the relationship between these nodes. 
+The samples were then examined using a Bayesian linear regression models with the _brms_ R package [@Burkner2017a] to determine the relationship between these nodes. 
 Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
@@
     col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
+    align = "lcccc",
     escape = F
@@
 
-\newpage
-
-```{r fig6, fig.align='default', fig.height=8, fig.width=12, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
+```{r fig6, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
 The strength of each arc is shown graphically by the style of the arc.
@@
     .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
+    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable",
     col.names = c("Variable", "Low", "High"),
@@
 The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
+The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
 The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
@@
 
-The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
+The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
 The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
@@
 
-Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
@@
 
-## Summary fo results
-
 The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
@@
 At both injured nodes in the first network, balance was linked to injury.
-Despite the weak arc strength at both injured nodes, a “High” balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014;], was found to increase the probability of injury.
+Despite the weak arc strength at both injured nodes, a “High” balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014], was found to increase the probability of injury.
 This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].

### 2020-08-14 manuscript.tex
pre figure removeal

--- 2020-07-29 apa_style.tex
@@
   english,
-  man,floatsintext]{apa6}
+  man]{apa6}
 \usepackage{lmodern}
@@
   pdflang={en-EN},
-  pdfkeywords={Sports Injury, Stress, Interdisciplinary, Bayesian Network},
+  pdfkeywords={Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology},
   hidelinks,
@@
 \textsuperscript{1} Cardiff Metropolitan University, Cardiff, United Kingdom\\\textsuperscript{2} Brunel University, London, United Kingdom\\\textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA), Manno, Switzerland}
-\authornote{
+\authornote{12961 words in text body and refence section
+
+Number of tables: 11
+
+Number of figures: 10
+
+Format: British English
 
 Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: harryfisher21@gmail.com}
-\keywords{Sports Injury, Stress, Interdisciplinary, Bayesian Network}
+\keywords{Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{lineno}
@@
 \newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {}%
   {\par}
@@
 
-10422 words in text body
-3142 words in reference section
-
 \hypertarget{introduction}{%
@@
 
-Over the last four decades sport-related injuries have received increased research attention (Ivarsson et al., 2017) in response to the high incidence (Rosa et al., 2014; Sheu et al., 2016) and associated undesirable physical and psychological effects (Leddy et al., 1994; Brewer, 2012).
-Multiple psychological (Slimani et al., 2018), anatomical (Murphy et al., 2003), biomechanical (Neely, 1998; Hughes, 2014) and environmental (Meeuwisse et al., 2007) factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence (Kumar, 2001; Meeuwisse et al., 2007; Wiese-Bjornstal, 2009), of which one of the most widely cited was developed by Williams and Anderson (Fig \ref{fig:fig1}; Andersen and Williams, 1988; Williams and Andersen, 1998).
+Over the last four decades sport-related injuries have received increased research attention (1) in response to the high incidence (2,3) and associated undesirable physical and psychological effects (4,5).
+Multiple psychological (6), anatomical (7), biomechanical (8,9) and environmental (10) factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence (10--12), of which one of the most widely cited was developed by Williams and Anderson (13,14).
 
@@
 
-Williams and Andersen's (Williams and Andersen, 1998) stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete's personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
+Williams and Andersen's (14) stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete's personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
 The stress response is central to the model and reflects the bi-directional relationship between athletes' appraisal of, and response to, a stressful athletic situation.
@@
 
-Within Williams and Andersen's (Williams and Andersen, 1998) model, major life events, which is a component of an athlete's history of stressors, most consistently predicts injury occurrence (Williams and Andersen, 2007); specifically, major life events with a negative, as opposed to positive, valence (Passer and Seese, 1983; Maddison and Prapavessis, 2005).
-Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile (Smith et al., 1990; Lavallée and Flint, 1996; Ivarsson and Johnson, 2010).However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30\% (Galambos et al., 2005; Ivarsson and Johnson, 2010), which indicates a likely interaction with other factors.
-
-While the psychosocial factors in Williams and Andersen's (Williams and Andersen, 1998) model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
+Within Williams and Andersen's (14) model, major life events, which is a component of an athlete's history of stressors, most consistently predicts injury occurrence (15); specifically, major life events with a negative, as opposed to positive, valence (16,17).
+Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile (18--20).
+However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30\% (20,21), which indicates a likely interaction with other factors.
+
+While the psychosocial factors in Williams and Andersen's (14) model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
 To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
-However, to date, the research has largely focused on attentional deficits (Andersen and Williams, 1999; Rogers and Landers, 2005; Wilkerson, 2012; Swanik et al., 2007).
-For example, Andersen and Williams (Andersen and Williams, 1999) found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
-Rodgers and Landers (Rogers and Landers, 2005) further supported Andersen and Williams's (Andersen and Williams, 1999) earlier findings by identifying that peripheral narrowing under stress mediated 8.1\% of the relationship between negative life events and injury.
-
-Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse (cf.~Williams and Andersen, 1998).
-One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response (Meeuwisse et al., 2007; Wiese-Bjornstal, 2009).
-For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence (Lee et al., 2017; Djaoui et al., 2017) and may account for the unexplained variance from the psychological predictors.
-Appaneal and Perna (Appaneal and Perna, 2014) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (Williams and Andersen, 1998) model.
-The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation (Appaneal and Perna, 2014).
-The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and ``widen the window of susceptibility'' (Appaneal and Perna, 2014, 74) to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen's (Williams and Andersen, 1998) model whilst including other physiological markers of training-related stress.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence (Perna and McDowell, 1995; Perna et al., 1997, 2003), other training-related stress markers including heart rate variability (Bellenger et al., 2016; Williams et al., 2017), muscle stiffness (Pruyn et al., 2015) and postural stability (Romero-Franco et al., 2014) have been reported to be associated with an increased risk of injury.
+However, to date, the research has largely focused on attentional deficits (22--25).
+For example, Andersen and Williams (22) found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
+Rodgers and Landers (23) further supported Andersen and Williams's (22) earlier findings by identifying that peripheral narrowing under stress mediated 8.1\% of the relationship between negative life events and injury.
+
+Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse (14).
+One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response (10,12).
+For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence (26,27) and may account for the unexplained variance from the psychological predictors.
+Appaneal and Perna (28) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (14) model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation (28).
+The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and ``widen the window of susceptibility'' (28) to a range of undesirable health outcomes including illness and injury.
+Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen's (14) model whilst including other physiological markers of training-related stress.
+
+Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence (29--31), other training-related stress markers including heart rate variability (32,33), muscle stiffness (34) and postural stability (35) have been reported to be associated with an increased risk of injury.
 However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
-Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries (Johnson et al., 2014).
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries (36).
 Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
@@
 
-Recently, Bittencourt et al.~(Bittencourt et al., 2016) advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
-Such an approach posits that injury may arise from a complex ``web of determinants'' (Bittencourt et al., 2016, 3), where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+Recently, Bittencourt et al.~(37) advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex ``web of determinants'' (37), where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
 Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
-Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables (Scutari and Denis, 2014).
+Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables (38).
 The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
-Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known (Olmedilla et al., 2018).
+Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known (39).
 BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
 
-Using the frameworks provided by Williams and Andersen's (Williams and Andersen, 1998) stress injury model and Appaneal's BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
+Using the frameworks provided by Williams and Andersen's (14) stress injury model and Appaneal's BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
 A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
@@
 Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours (Fuller et al., 2006, 2007; Timpka et al., 2014).
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries (cf.~Appaneal et al., 2009).
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours (40--42).
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries (43).
 Injury status (did / did not sustain an injury) served as the main outcome measure.
@@
 
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress (Petrie, 1992).
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress (44).
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature.
-Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table).
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (supplementary table 1).
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
@@
 A score for total life events was also calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) (Petrie, 1992) reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001) respectively.
+Petrie (1992) (44) reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001) respectively.
 Petrie also provided evidence of discriminant, convergent and predictive validity.
-For this study, composite reliability (Fornell and Larcker, 1981) was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+For this study, composite reliability (45) was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
 The composite reliability for the LESCA in this study was .84.
@@
 
-A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology (Corr and Cooper, 2016).
+A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology (46).
 The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I am the sort of person who easily freezes-up when scared''), Behavioural Inhibition System (BIS; e.g., ``When trying to make a decision, I find myself constantly chewing it over'') and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., ``I regularly try new activities just to see if I enjoy them''), Goal Drive Persistence (e.g., ``I am very persistent in achieving my goals''), Reward Reactivity (e.g., ``I get a special thrill when I am praised for something I've done well'') and Impulsivity (e.g., ``I find myself doing things on the spur of the moment'').
@@
 The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in S1 Appendix.
+The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in supplementary Appendix 1.
 
@@
 A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings (Giles et al., 2016), which are considered the gold standard for assessing HRV.
-In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes (Caminal et al., 2018) and under spontaneous breathing conditions (Plews et al., 2017).
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings (47), which are considered the gold standard for assessing HRV.
+In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes (48) and under spontaneous breathing conditions (49).
 
@@
 The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
-The raw oscillation signal is then processed, and the stiffness parameter is calculated (Agyapong-Badu et al., 2016).
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties (Chuang et al., 2013; Pruyn et al., 2016; Nair et al., 2014), and has demonstrated good internal consistency (coefficient of variation \textless{} 1.4\%) over sets of 10 repetitions (Aird et al., 2012).
+The raw oscillation signal is then processed, and the stiffness parameter is calculated (50).
+The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties (51--53), and has demonstrated good internal consistency (coefficient of variation \textless{} 1.4\%) over sets of 10 repetitions (54).
 
@@
 
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by Hunt et al. (2009).
-In total, each trial of the mBESS was performed without shoes (McCrory et al., 2013) and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
+Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by (55).
+In total, each trial of the mBESS was performed without shoes (56) and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
 non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion)
 and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}).
-To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (cf.~Cingel et al., 2017).
+To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (57).
 Participants were asked to maintain each stance for a total of 20-seconds.
@@
 e) lifting the forefoot or heel; and
-f) remaining out of the testing position for more than 5-seconds (Riemann et al., 1999).
+f) remaining out of the testing position for more than 5-seconds (58).
 A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
@@
 A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces).
-To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation (ICC; Koo and Li, 2016) was used on a sample of 40 participants from the first time point.
+To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation (59) was used on a sample of 40 participants from the first time point.
 The test-retest scoring of BESS resulted in a ``good'' to ``excellent'' ICC score (ICC = 0.93, 95\% confidence interval = 0.88 - 0.96), indicating the scoring was reliable.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{figures_doc/studydesign_new} \caption{Study design.
+\includegraphics[width=1\linewidth]{figures_doc/Fig3} \caption{Study design.
 For each time point (T), each box contains the number of participants who completed the data collection (\textit{n}), the measures used for data collection and the approximate date of the data collection.}\label{fig:fig3}
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/sessionprotocol}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig4}
 
@@
 For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R (R Core Team, 2019) for analysis.
+The data were downloaded from surverymonkey.com and imported into R (60) for analysis.
 
@@
 Participants were seated and asked to remain as still as possible for the duration of the recording.
-No attempt was made to control participants' respiratory frequency or tidal volume (Denver et al., 2007).
+No attempt was made to control participants' respiratory frequency or tidal volume (61).
 Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
 
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R (R Core Team, 2019) where the \emph{RHRV} package (Rodriguez-Linares et al., 2019) was used to calculate HRV indices.
-Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings (Laborde et al., 2017; Malik et al., 1996).
+Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R (60) where the \emph{RHRV} package (62) was used to calculate HRV indices.
+Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings (63,64).
 RMSSD was calculated as:
@@
 To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb.
-The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site (Chuang et al., 2012).
+The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site (65).
 The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle.
@@
 A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one-second interval (Morgan et al., 2018).
+The device was programmed to deliver five consecutive impulses, separated by a one-second interval (66).
 For each impulse, the device computed passive stiffness values, with the median of the five values being saved by the device for further analysis.
@@
 Sets of measurements above 3\% were measured again to ensure data reliability.
-Measurements were uploaded using MyotonPRO software and imported in R (R Core Team, 2019) for further analysis.
+Measurements were uploaded using MyotonPRO software and imported in R (60) for further analysis.
 For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis
@@
 Each completed mBESS protocol took approximately four-minutes.
-Only one trial was performed to avoid familiarisation effects across the repeated measurement (cf.~Valovich et al., 2003).
+Only one trial was performed to avoid familiarisation effects across the repeated measurement (67).
 The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
@@
 Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury.
-A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way (Scutari and Denis, 2014).
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter (Fig \ref{fig:fig5}; Pearl, 1988).
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual (Scutari and Denis, 2014).
-The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN (Chen and Pollino, 2012).
+A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way (38).
+A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter (68).
+However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual (38).
+The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN (69).
 In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/examplenetwork}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig5}
 
@@
 For example, a network with 10 variables each with two levels has 100 (10\^{}2) possible parameter combinations, however the same network with three levels has 1000 (10\^{}3) possible parameter combinations, the latter being significantly more computationally expensive.
-Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio (Scutari and Denis, 2014).
-Typically, no more than three levels have been used in Bayesian networks in the sports injury literature (Olmedilla et al., 2018; Ruiz-Pérez et al., 2019)
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio (38).
+Typically, no more than three levels have been used in Bayesian networks in the sports injury literature (39,70)
 
@@
 A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
-Widely used texts on Bayesian Network analysis (Koller and Friedman, 2009) have widely adopted the terminology in favour of that used in traditional statistics.} and obtaining the posterior probabilities of a particular node for a given query.
-The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (Scutari and Denis, 2014).
+Widely used texts on Bayesian Network analysis (71) have widely adopted the terminology in favour of that used in traditional statistics.} and obtaining the posterior probabilities of a particular node for a given query.
+The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (38).
 In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to ``Low'') could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
@@
 
-BNs have recently been used in the sport psychology literature (Olmedilla et al., 2018; Fuster-Parra et al., 2017; Ruiz-Pérez et al., 2019) and offer several benefits over traditional statistical analysis.
-For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (Olmedilla et al., 2018; Bittencourt et al., 2016).
-Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (Xiao-xuan et al., 2008).
+BNs have recently been used in the sport psychology literature (39,70,72) and offer several benefits over traditional statistical analysis.
+For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (37,39).
+Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (73).
 In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
-Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (Heckerman et al., 1995; Xu et al., 2015).
+Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (74,75).
 The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
@@
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
-A missing rate of 15-20\% has been reported to be common in psychological studies, and several techniques are available to handle missing values (Enders, 2003; Lang et al., 2013).
-In the current study, the \emph{caret} package (Kuhn et al., 2008) was used to impute the missing values.
+A missing rate of 15-20\% has been reported to be common in psychological studies, and several techniques are available to handle missing values (76,77).
+In the current study, the \emph{caret} R package (78) was used to impute the missing values.
 A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset.
-The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation (Kuhn et al., 2008).
+The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation (78).
 
@@
 For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
-Given the limited support for a relationship between positive life events and injury (Williams and Andersen, 2007), only negative and total life events were included in the network.
+Given the limited support for a relationship between positive life events and injury (15), only negative and total life events were included in the network.
 Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (Ivarsson et al., 2017).
+In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).
 
@@
 Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
-Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (Scutari and Denis, 2014).
-To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (Murphy, 2002).
+Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (38).
+To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (79).
 In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
@@
 \centering
-\begin{tabular}[t]{l|l|l}
+\begin{tabular}[t]{c|c|c}
 \hline
@@
 Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
-Networks were learned using a Tabu search algorithm (Russell and Norvig, 2009) and BIC was used to compare different models.
-A higher BIC value indicates the structure of a DAG is a better fit to the observed data (Scutari and Denis, 2014).
+Networks were learned using a Tabu search algorithm (80) and BIC was used to compare different models.
+A higher BIC value indicates the structure of a DAG is a better fit to the observed data (38).
 BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
 
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = -83.64, tlelg and nlelg = -216.47).
-Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (Williams and Andersen, 2007), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (nleleg = -4,242.76, tlelg only = -4,326.39, tlelg and nlelg = -4,459.23).
+Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
 Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2 in the final network structure.
@@
 Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
-RI provided the highest BIC value (-3,563.13), compared to RR (-15.98), GDP (-19.26) and I (-19.76).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -900.12) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
+RI provided the highest BIC value (-3,563.13), compared to RR (-3,579.10), GDP (-3,582.39) and I (-3,582.89).
+Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -4,463.25) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
 Therefore, only RI was included in the final structure.
@@
 To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network.
-The final network structure was obtained by averaging 1000 bootstrapped models (Efron and Tibshirani, 1994) to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model (Olmedilla et al., 2018).
+The final network structure was obtained by averaging 1000 bootstrapped models (81) to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model (39).
 Arcs that were present in at least 30\% of the models were included in the averaged model.
@@
 An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
-In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (Scutari and Nagarajan, 2013).
-Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
+In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (82).
+Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in supplementary table 1 and supplementary table 3 respectively.
 
@@
 CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
-The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query (Scutari and Denis, 2014).
+The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query (38).
 Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
 Of particular interest were the variables that were connected to ``injured'' nodes, which were examined in the network using the Markov blanket of ``injured\_1'' and ``injured\_2''.
-A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network (Fuster-Parra et al., 2017).
+A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network (72).
 CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the ``injured'' state.
@@
 \paragraph{Data preparation}\label{data-preparation-1}}
-
-Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
 
@@
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then examined using a Bayesian linear regression models with the \emph{brms} package (Bürkner, 2017) to determine the relationship between these nodes.
+The samples were then examined using a Bayesian linear regression models with the \emph{brms} R package (83) to determine the relationship between these nodes.
 Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
@@
 \centering
-\begin{tabular}[t]{l|l|l|l|l}
+\begin{tabular}[t]{l|c|c|c|c}
 \hline
@@
 
-\newpage
-
 \begin{figure}
 \centering
-\includegraphics{apa_style_files/figure-latex/fig6-1.pdf}
+\includegraphics{manuscript_files/figure-latex/fig6-1.pdf}
 \caption{\label{fig:fig6}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
@@
 
-\caption{\label{tab:table5}Probability of injured\_1 being in the ``injured'' state, conditional on each variable.}
+\caption{\label{tab:table5}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
 \centering
@@
 The second CPQ investigated the outcome of injured\_1 being ``injured'' conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
+The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
 The combination of lower competitive level, ``High'' hours per week, ``Low'' negative life events, ``High'' balance and ``High'' stiffness resulted in a probability of 0.53 for injured\_1 being in the ``injured'' state.
@@
 
-The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
+The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
 The combination of ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.53.
@@
 
-Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
+Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
@@
 
-\hypertarget{summary-fo-results}{%
-\subsection{Summary fo results}\label{summary-fo-results}}
-
-The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
+The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.\\
 Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
@@
 The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete's risk of injury.
-In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury (Appaneal and Perna, 2014).
+In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury (28).
 
@@
 In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance (Romero-Franco et al., 2014; Trojian and McKeag, 2006) and muscle stiffness (Butler et al., 2003; Pickering-Rodriguez et al., 2017) are separately related to injury.
+Several studies have identified how balance (35,84) and muscle stiffness (85,86) are separately related to injury.
 The BN structures examined in this study provided insight into these two stress-related factors and their relation to injury occurrence.
@@
 At both injured nodes in the first network, balance was linked to injury.
-Despite the weak arc strength at both injured nodes, a ``High'' balance score, which is considered indicative of impaired postural stability {[}Romero-Franco et al.
-(2014);{]}, was found to increase the probability of injury.
-This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (Riemann et al., 1999; Romero-Franco et al., 2014; Trojian and McKeag, 2006).
-Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (Paillard et al., 2006).
+Despite the weak arc strength at both injured nodes, a ``High'' balance score, which is considered indicative of impaired postural stability (35), was found to increase the probability of injury.
+This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (35,58,84).
+Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (87).
 Athletes who competed at a higher level were also more likely to have good balance (``Low'' balance), compared to their lower level counterparts.
-These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (Hrysomallis, 2011).
-
-Negative life events captured at a single time point have previously been reported to be most strongly associated with injury (Williams and Andersen, 2007; Ivarsson et al., 2017).
+These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (88).
+
+Negative life events captured at a single time point have previously been reported to be most strongly associated with injury (1,15).
 The repeated measures approach combined with the second network analysis employed in this study demonstrated that greater than average increases in negative life event stress between time points increased the probability of being injured during the study period.
@@
 For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400\% increase in their life event score.
-Although the absolute score would be ``Low'', the relative increase between time points may have been caused by a significant event in the athlete's life that had a considerable psychological and physiological effect (Appaneal and Perna, 2014).
-Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed (cf.~Ivarsson et al., 2014).
+Although the absolute score would be ``Low'', the relative increase between time points may have been caused by a significant event in the athlete's life that had a considerable psychological and physiological effect (28).
+Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed (89).
 
 During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
-The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model (Ivarsson et al., 2017).
+The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model (1).
 During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
@@
 One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
-In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months (Petrie, 1992).
+In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months (44).
 In this study, athletes completed the LESCA at three time points with an approximate four-month interval after baseline.
@@
 For example, at the second and third time points, 26\% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events (cf.~Fawkner et al., 1999).
-
-Williams and Andersen's (Williams and Andersen, 1998) model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
-Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (Petrie, 1993; Petrie et al., 2014; Johnson et al., 2014).
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events (90).
+
+Williams and Andersen's (14) model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
+Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (36,91,92).
 The lack of association between major/negative life events and injury reported here may be attributed to athletes in this study having the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
@@
 Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, ``Low'' HRV increased the probability of injury from 0.17 (``High'' HRV) to 0.26 (``Low'' HRV).
-This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (Williams et al., 2017; Bellenger et al., 2016; Gisselman et al., 2016).
+This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (32,33,93).
 An arc between FFFS\_1 and injured\_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the ``High'' and ``Low'' states respectively.
@@
 One possible explanation for this finding could be that those athletes who reported ``Low'' FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
-The RST theory proposes that higher levels of FFFS increase avoidance motivation (Corr et al., 2016), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (Corr, 2013).
+The RST theory proposes that higher levels of FFFS increase avoidance motivation (94), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
+The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (95).
 The first network reported an association between ``High'' FFFS and ``High'' BIS, while the second network found an association between increases in FFFS and increases in BIS.
-High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD (Chalmers et al., 2014; Pulopulos et al., 2018).
-This association along with the proposed actions of the RST theory (Corr et al., 2016) provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network.
+High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD (96,97).
+This association along with the proposed actions of the RST theory (94) provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network.
 To elaborate, high levels of BIS are proposed to be the result of goal conflict, an example of which would be simultaneous triggering of the FFFS (avoidance) and BAS (approach) systems.
-The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (Corr et al., 2016).
+The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (94).
 In the present study, however, the role of BAS was limited, as evidenced by the initial network structures in which the BAS had limited connectivity with other components of the network.
@@
 This study had a number of strengths.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury (Johnson et al., 2014).
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury (36).
 The longitudinal repeated measures design of the current study allowed changes over time and between time points to be captured and explored.
 Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (Bittencourt et al., 2016) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (37) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
 The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
@@
 
-While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (Olmedilla et al., 2018), a number of assumptions were made that potentially limited the approach employed in the study.
+While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (39), a number of assumptions were made that potentially limited the approach employed in the study.
 Firstly, the choice was made to binarise variables in the first network so only ``Low'' and ``High'' states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner Qian and Miltner (2015) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner (98) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
 For example, the meaning of a ``Low'' and ``High'' value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
@@
 However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes' responses.
-For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis (Ross et al., 2011).
+For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis (99).
 However, the video-capture approach employed in the study ensured an accessible, non-invasive and readily applied method of capturing the respective measure.
@@
 The variables used in the present study were not definitive.
-Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress (Appaneal and Perna, 2014; Perna and McDowell, 1995), could help to further elucidate the relationship between stress and injury.
+Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress (28,29), could help to further elucidate the relationship between stress and injury.
 Future developments in the capture of life event stress using the LESCA are also warranted.
@@
 \begin{cslreferences}
-\leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
-Agyapong-Badu, S., Warner, M., Samuel, D., and Stokes, M.
-(2016).
-Measurement of ageing effects on muscle tone and mechanical properties of rectus femoris and biceps brachii in healthy males and females using a novel hand-held myometric device.
-\emph{Archives of Gerontology and Geriatrics} 62, 59--67.
-doi:\href{https://doi.org/10.1016/j.archger.2015.09.011}{10.1016/j.archger.2015.09.011}.
-
-\leavevmode\hypertarget{ref-Aird2012}{}%
-Aird, L., Samuel, D., and Stokes, M.
-(2012).
-Quadriceps muscle tone, elasticity and stiffness in older males: Reliability and symmetry using the MyotonPRO.
-\emph{Archives of Gerontology and Geriatrics} 55, e31--e39.
-doi:\href{https://doi.org/https://doi.org/10.1016/j.archger.2012.03.005}{https://doi.org/10.1016/j.archger.2012.03.005}.
-
-\leavevmode\hypertarget{ref-Andersen1988}{}%
-Andersen, M.
-B., and Williams, J.
-M.
-(1988).
-A model of stress and athletic injury: Prediction and prevention.
-\emph{Journal of Sport and Exercise Psychology} 10, 294--306.
-doi:\href{https://doi.org/10.1123/jsep.10.3.294}{10.1123/jsep.10.3.294}.
-
-\leavevmode\hypertarget{ref-Andersen1999}{}%
-Andersen, M.
-B., and Williams, J.
-M.
-(1999).
-Athletic injury, psychosocial factors and perceptual changes during stress.
-\emph{Journal of Sports Sciences} 17, 735--741.
-doi:\href{https://doi.org/10.1080/026404199365597}{10.1080/026404199365597}.
-
-\leavevmode\hypertarget{ref-Appaneal2009}{}%
-Appaneal, R.
-N., Levine, B.
-R., Perna, F.
-M., and Roh, J.
-L.
-(2009).
-Measuring postinjury depression among male and female competitive athletes.
-\emph{Journal of Sport and Exercise Psychology} 31, 60--76.
-Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/19325188}.
-
-\leavevmode\hypertarget{ref-Appaneal2014}{}%
-Appaneal, R.
-N., and Perna, F.
-M.
-(2014).
-``Biopsychosocial model of injury,'' in \emph{Encyclopedia of sport and exercise psychology}, eds.
-R.
-C.
-Eklund and G.
-Tenenbaum (Thousand Oaks, CA: Sage), 74--77.
-
-\leavevmode\hypertarget{ref-Bellenger2016}{}%
-Bellenger, C.
-R., Fuller, J.
-T., Thomson, R.
-L., Davison, K., Robertson, E.
-Y., and Buckley, J.
-D.
-(2016).
-Monitoring athletic training status through autonomic heart rate regulation: A systematic review and meta-analysis.
-\emph{Sports Medicine} 46, 1461--1486.
-doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}.
-
-\leavevmode\hypertarget{ref-Bittencourt2016}{}%
-Bittencourt, N.
-F.
-N., Meeuwisse, W.
-H., Mendonça, L.
-D., Nettel-Aguirre, A., Ocarino, J.
-M., and Fonseca, S.
-T.
-(2016).
-Complex systems approach for sports injuries: Moving from risk factor identification to injury pattern recognition - narrative review and new concept.
-\emph{British Journal of Sports Medicine} 50, 1309--1314.
-doi:\href{https://doi.org/10.1136/bjsports-2015-095850}{10.1136/bjsports-2015-095850}.
+\leavevmode\hypertarget{ref-Ivarsson2017}{}%
+1.
+Ivarsson A, Johnson U, Andersen MB, Tranaeus U, Stenling A, Lindwall M.
+Psychosocial factors and sport injuries: Meta-analyses for prediction and prevention.
+\emph{Sports Medicine} (2017) \textbf{47}:353--365.
+doi:\href{https://doi.org/10.1007/s40279-016-0578-x}{10.1007/s40279-016-0578-x}
+
+\leavevmode\hypertarget{ref-Rosa2014}{}%
+2.
+Rosa BB, Asperti AM, Helito CP, Demange MK, Fernandes TL, Hernandez AJ.
+Epidemiology of sports injuries on collegiate athletes at a single center.
+\emph{Acta Ortopédica Brasileira} (2014) \textbf{22}:321--324.
+doi:\href{https://doi.org/10.1590/1413-78522014220601007}{10.1590/1413-78522014220601007}
+
+\leavevmode\hypertarget{ref-Sheu2016}{}%
+3.
+Sheu Y, Chen LH, Hedegaard H.
+Sports- and recreation-related injury episodes in the United States, 2011-2014.
+\emph{National Health Statistics Reports} (2016)1--12.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/27906643}
+
+\leavevmode\hypertarget{ref-Leddy1994}{}%
+4.
+Leddy MH, Lambert MJ, Ogles BM.
+Psychological consequences of athletic injury among high-level competitors.
+\emph{Research Quarterly for Exercise and Sport} (1994) \textbf{65}:347--354.
+doi:\href{https://doi.org/10.1080/02701367.1994.10607639}{10.1080/02701367.1994.10607639}
 
 \leavevmode\hypertarget{ref-Brewer2012}{}%
-Brewer, B.
-W.
-(2012).
+5.
+Brewer BW.
 ``Psychology of sport injury rehabilitation,'' in \emph{Handbook of sport psychology}, eds.
 G.
-Tenenbaum and R.
+Tenenbaum, R.
 C.
 Eklund (Hoboken, NJ, USA: Wiley), 404--424.
-doi:\href{https://doi.org/10.1002/9781118270011.ch18}{10.1002/9781118270011.ch18}.
-
-\leavevmode\hypertarget{ref-Butler2003}{}%
-Butler, R.
-J., Crowell, H.
-P., and Davis, I.
-M.
+doi:\href{https://doi.org/10.1002/9781118270011.ch18}{10.1002/9781118270011.ch18}
+
+\leavevmode\hypertarget{ref-Slimani2018}{}%
+6.
+Slimani M, Bragazzi NL, Znazen H, Paravlic A, Azaiez F, Tod D.
+Psychosocial predictors and psychological prevention of soccer injuries: A systematic review and meta-analysis of the literature.
+\emph{Physical Therapy in Sport} (2018) \textbf{32}:293--300.
+doi:\href{https://doi.org/10.1016/j.ptsp.2018.05.006}{10.1016/j.ptsp.2018.05.006}
+
+\leavevmode\hypertarget{ref-Murphy2003}{}%
+7.
+Murphy DF, Connolly DAJ, Beynnon BD.
+Risk factors for lower extremity injury: A review of the literature.
+\emph{British Journal of Sports Medicine} (2003) \textbf{37}:13--29.
+doi:\href{https://doi.org/10.1136/bjsm.37.1.13}{10.1136/bjsm.37.1.13}
+
+\leavevmode\hypertarget{ref-Neely1998}{}%
+8.
+Neely FG.
+Biomechanical risk factors for exercise-related lower limb injuries.
+\emph{Sports Medicine} (1998) \textbf{26}:395--413.
+doi:\href{https://doi.org/10.2165/00007256-199826060-00003}{10.2165/00007256-199826060-00003}
+
+\leavevmode\hypertarget{ref-Hughes2014}{}%
+9.
+Hughes G.
+A review of recent perspectives on biomechanical risk factors associated with anterior cruciate ligament injury.
+\emph{Research in Sports Medicine} (2014) \textbf{22}:193--212.
+doi:\href{https://doi.org/10.1080/15438627.2014.881821}{10.1080/15438627.2014.881821}
+
+\leavevmode\hypertarget{ref-Meeuwisse2007}{}%
+10.
+Meeuwisse WH, Tyreman H, Hagel B, Emery C.
+A dynamic model of etiology in sport injury: The recursive nature of risk and causation.
+\emph{Clinical Journal of Sport Medicine} (2007) \textbf{17}:215--219.
+doi:\href{https://doi.org/10.1097/JSM.0b013e3180592a48}{10.1097/JSM.0b013e3180592a48}
+
+\leavevmode\hypertarget{ref-Kumar2001}{}%
+11. Kumar S. Theories of musculoskeletal injury causation. \emph{Ergonomics} (2001) \textbf{44}:17--47. doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}
+
+\leavevmode\hypertarget{ref-Wiese-Bjornstal2009}{}%
+12.
+Wiese-Bjornstal DM.
+Sport injury and college athlete health across the lifespan.
+\emph{Journal of Intercollegiate Sport} (2009) \textbf{2}:64--80.
+doi:\href{https://doi.org/10.1123/jis.2.1.64}{10.1123/jis.2.1.64}
+
+\leavevmode\hypertarget{ref-Andersen1988}{}%
+13.
+Andersen MB, Williams JM.
+A model of stress and athletic injury: Prediction and prevention.
+\emph{Journal of Sport and Exercise Psychology} (1988) \textbf{10}:294--306.
+doi:\href{https://doi.org/10.1123/jsep.10.3.294}{10.1123/jsep.10.3.294}
+
+\leavevmode\hypertarget{ref-Williams1998}{}%
+14.
+Williams JM, Andersen MB.
+Psychosocial antecedents of sport injury: review and critique of the stress and injury model.
+\emph{Journal of Applied Sport Psychology} (1998) \textbf{10}:5--25.
+doi:\href{https://doi.org/10.1080/10413209808406375}{10.1080/10413209808406375}
+
+\leavevmode\hypertarget{ref-Williams2007}{}%
+15.
+Williams JM, Andersen MB.
+``Psychosocial antecedents of sport injury and interventions for risk reduction,'' in \emph{Handbook of sport psychology}, eds.
+G.
+Tenenbaum, R.
 C.
-(2003).
-Lower extremity stiffness: Implications for performance and injury.
-\emph{Clinical Biomechanics} 18, 511--517.
-doi:\href{https://doi.org/10.1016/S0268-0033(03)00071-8}{10.1016/S0268-0033(03)00071-8}.
-
-\leavevmode\hypertarget{ref-Burkner2017a}{}%
-Bürkner, P. C. (2017). brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} 80. doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}.
+Eklund (Hoboken, NJ, USA: Wiley), 379--403.
+doi:\href{https://doi.org/10.1002/9781118270011.ch17}{10.1002/9781118270011.ch17}
+
+\leavevmode\hypertarget{ref-Passer1983a}{}%
+16.
+Passer MW, Seese MD.
+Life stress and athletic injury: Examination of positive versus negative events and three moderator variables.
+\emph{Journal of Human Stress} (1983) \textbf{9}:11--16.
+doi:\href{https://doi.org/10.1080/0097840X.1983.9935025}{10.1080/0097840X.1983.9935025}
+
+\leavevmode\hypertarget{ref-Maddison2005}{}%
+17.
+Maddison R, Prapavessis H.
+A psychological approach to the prediction and prevention of athletic injury.
+\emph{Journal of Sport and Exercise Psychology} (2005) \textbf{27}:289--310.
+doi:\href{https://doi.org/10.1123/jsep.27.3.289}{10.1123/jsep.27.3.289}
+
+\leavevmode\hypertarget{ref-Smith1990}{}%
+18.
+Smith RE, Smoll FL, Ptacek JT.
+Conjunctive moderator variables in vulnerability and resiliency research: Life stress, social support and coping skills, and adolescent sport injuries.
+\emph{Journal of Personality and Social Psychology} (1990) \textbf{58}:360--370.
+doi:\href{https://doi.org/10.1037/0022-3514.58.2.360}{10.1037/0022-3514.58.2.360}
+
+\leavevmode\hypertarget{ref-Lavallee1996}{}%
+19.
+Lavallée L, Flint F.
+The relationship of stress, competitive anxiety, mood state, and social support to athletic injury.
+\emph{Journal of Athletic Training} (1996) \textbf{31}:296--299.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16558413}
+
+\leavevmode\hypertarget{ref-Ivarsson2010}{}%
+20.
+Ivarsson A, Johnson U.
+Psychological factors as predictors of injuries among senior soccer players.
+A prospective study.
+\emph{Journal of Sports Science and Medicine} (2010) \textbf{9}:347--352.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3761721/}
+
+\leavevmode\hypertarget{ref-Galambos2005}{}%
+21.
+Galambos SA, Terry PC, Moyle GM, Locke SA.
+Psychological predictors of injury among elite athletes.
+\emph{British Journal of Sports Medicine} (2005) \textbf{39}:351--354.
+doi:\href{https://doi.org/10.1136/bjsm.2005.018440}{10.1136/bjsm.2005.018440}
+
+\leavevmode\hypertarget{ref-Andersen1999}{}%
+22.
+Andersen MB, Williams JM.
+Athletic injury, psychosocial factors and perceptual changes during stress.
+\emph{Journal of Sports Sciences} (1999) \textbf{17}:735--741.
+doi:\href{https://doi.org/10.1080/026404199365597}{10.1080/026404199365597}
+
+\leavevmode\hypertarget{ref-Rogers2005}{}%
+23.
+Rogers T, Landers DM.
+Mediating effects of peripheral vision in the life event stress/athletic injury relationship.
+\emph{Journal of Sport and Exercise Psychology} (2005) \textbf{27}:271--288.
+doi:\href{https://doi.org/10.1002/9781444303650}{10.1002/9781444303650}
+
+\leavevmode\hypertarget{ref-Wilkerson2012a}{}%
+24.
+Wilkerson GB.
+Neurocognitive reaction time predicts lower extremity sprains and strains.
+\emph{International Journal of Athletic Therapy and Training} (2012) \textbf{17}:4--9.
+doi:\href{https://doi.org/10.1123/ijatt.17.6.4}{10.1123/ijatt.17.6.4}
+
+\leavevmode\hypertarget{ref-Swanik2007}{}%
+25.
+Swanik CB, Covassin T, Stearne DJ, Schatz P.
+The relationship between neurocognitive function and noncontact anterior cruciate ligament injuries.
+\emph{American Journal of Sports Medicine} (2007) \textbf{35}:943--948.
+doi:\href{https://doi.org/10.1177/0363546507299532}{10.1177/0363546507299532}
+
+\leavevmode\hypertarget{ref-Lee2017}{}%
+26.
+Lee EC, Fragala MS, Kavouras SA, Queen RM, Pryor JL, Casa DJ.
+Biomarkers in sports and exercise: Tracking health, performance, and recovery in athletes.
+\emph{Journal of Strength and Conditioning Research} (2017) \textbf{31}:2920--2937.
+doi:\href{https://doi.org/10.1519/JSC.0000000000002122}{10.1519/JSC.0000000000002122}
+
+\leavevmode\hypertarget{ref-Djaoui2017}{}%
+27.
+Djaoui L, Haddad M, Chamari K, Dellal A.
+Monitoring training load and fatigue in soccer players with physiological markers.
+\emph{Physiology and Behavior} (2017) \textbf{181}:86--94.
+doi:\href{https://doi.org/10.1016/j.physbeh.2017.09.004}{10.1016/j.physbeh.2017.09.004}
+
+\leavevmode\hypertarget{ref-Appaneal2014}{}%
+28. Appaneal RN, Perna FM. ``Biopsychosocial model of injury,'' in \emph{Encyclopedia of sport and exercise psychology}, eds. R. C. Eklund, G. Tenenbaum (Thousand Oaks, CA: Sage), 74--77.
+
+\leavevmode\hypertarget{ref-Perna1995}{}%
+29.
+Perna FM, McDowell SL.
+Role of psychological stress in cortisol recovery from exhaustive exercise among elite athletes.
+\emph{International Journal of Behavioral Medicine} (1995) \textbf{2}:13--26.
+doi:\href{https://doi.org/10.1207/s15327558ijbm0201_2}{10.1207/s15327558ijbm0201\_2}
+
+\leavevmode\hypertarget{ref-Perna1997}{}%
+30.
+Perna F, Schneiderman N, LaPerriere A.
+Psychological stress, exercise and immunity.
+\emph{International Journal of Sports Medicine} (1997) \textbf{18}:78--83.
+doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}
+
+\leavevmode\hypertarget{ref-Perna2003}{}%
+31.
+Perna FM, Antoni MH, Baum A, Gordon P, Schneiderman N.
+Cognitive behavioral stress management effects on injury and illness among competitive athletes: A randomized clinical trial.
+\emph{Annals of Behavioral Medicine} (2003) \textbf{25}:66--73.
+Available at: \url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}
+
+\leavevmode\hypertarget{ref-Bellenger2016}{}%
+32.
+Bellenger CR, Fuller JT, Thomson RL, Davison K, Robertson EY, Buckley JD.
+Monitoring athletic training status through autonomic heart rate regulation: A systematic review and meta-analysis.
+\emph{Sports Medicine} (2016) \textbf{46}:1461--1486.
+doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}
+
+\leavevmode\hypertarget{ref-Williams2017}{}%
+33.
+Williams S, Booton T, Watson M, Rowland D, Altini M.
+Heart rate variability is a moderating factor in the workload-injury relationship of competitive crossfit™ athletes.
+\emph{Journal of Sports Science and Medicine} (2017) \textbf{16}:443--449.
+
+\leavevmode\hypertarget{ref-Pruyn2015}{}%
+34.
+Pruyn EC, Watsford ML, Murphy AJ.
+Differences in lower-body stiffness between levels of netball competition.
+\emph{Journal of Strength and Conditioning Research} (2015) \textbf{29}:1197--1202.
+doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}
+
+\leavevmode\hypertarget{ref-Romero-Franco2014}{}%
+35.
+Romero-Franco N, Gallego-Izquierdo T, Martínez-López EJ, Hita-Contreras F, Osuna-Pére, Catalina M, Martínez-Amat A.
+Postural stability and subsequent sports injuries during indoor season of athletes.
+\emph{Journal of Physical Therapy Science} (2014) \textbf{26}:683--687.
+doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}
+
+\leavevmode\hypertarget{ref-Johnson2014}{}%
+36.
+Johnson U, Tranaeus U, Ivarsson A.
+Current status and future challenges in psychological research of sport injury prediction and prevention: A methodological perspective.
+\emph{Revista de Psicologia del Deporte} (2014) \textbf{23}:401--409.
+Available at: \url{https://www.rpd-online.com/article/view/v23-n2-johnson-tranaeus-ivarsson}
+
+\leavevmode\hypertarget{ref-Bittencourt2016}{}%
+37.
+Bittencourt NFN, Meeuwisse WH, Mendonça LD, Nettel-Aguirre A, Ocarino JM, Fonseca ST.
+Complex systems approach for sports injuries: Moving from risk factor identification to injury pattern recognition - narrative review and new concept.
+\emph{British Journal of Sports Medicine} (2016) \textbf{50}:1309--1314.
+doi:\href{https://doi.org/10.1136/bjsports-2015-095850}{10.1136/bjsports-2015-095850}
+
+\leavevmode\hypertarget{ref-Scutari2014}{}%
+38.
+Scutari M, Denis J-B.
+\emph{Bayesian networks: with examples in R}.
+1st ed.
+Chapman \& Hall/CRC (2014).
+Available at: \url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}
+
+\leavevmode\hypertarget{ref-Olmedilla2018}{}%
+39.
+Olmedilla A, Rubio VJ, Fuster-Parra P, Pujals C, García-Mas A.
+A Bayesian approach to sport injuries likelihood: Does player's self-efficacy and environmental factors plays the main role?
+\emph{Frontiers in Psychology} (2018) \textbf{9}:1--10.
+doi:\href{https://doi.org/10.3389/fpsyg.2018.01174}{10.3389/fpsyg.2018.01174}
+
+\leavevmode\hypertarget{ref-Fuller2006}{}%
+40.
+Fuller CW, Ekstrand J, Junge A, Andersen TE, Bahr R, Dvorak J, et al.
+Consensus statement on injury definitions and data collection procedures in studies of football (soccer) injuries.
+\emph{British journal of sports medicine} (2006) \textbf{40}:193--201.
+doi:\href{https://doi.org/10.1136/bjsm.2005.025270}{10.1136/bjsm.2005.025270}
+
+\leavevmode\hypertarget{ref-Fuller2007b}{}%
+41.
+Fuller CW, Molloy MG, Bagate C, Bahr R, Brooks JHM, Donson H, et al.
+Consensus statement on injury definitions and data collection procedures for studies of injuries in rugby union.
+\emph{British journal of sports medicine} (2007) \textbf{41}:328--31.
+doi:\href{https://doi.org/10.1136/bjsm.2006.033282}{10.1136/bjsm.2006.033282}
+
+\leavevmode\hypertarget{ref-Timpka2014}{}%
+42.
+Timpka T, Alonso JM, Jacobsson J, Junge A, Branco P, Clarsen B, et al.
+Injury and illness definitions and data collection procedures for use in epidemiological studies in Athletics (track and field): Consensus statement.
+\emph{British Journal of Sports Medicine} (2014) \textbf{48}:483--490.
+doi:\href{https://doi.org/10.1136/bjsports-2013-093241}{10.1136/bjsports-2013-093241}
+
+\leavevmode\hypertarget{ref-Appaneal2009}{}%
+43.
+Appaneal RN, Levine BR, Perna FM, Roh JL.
+Measuring postinjury depression among male and female competitive athletes.
+\emph{Journal of Sport and Exercise Psychology} (2009) \textbf{31}:60--76.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/19325188}
+
+\leavevmode\hypertarget{ref-Petrie1992}{}%
+44.
+Petrie TA.
+Psychosocial antecedents of athletic injury: The effects of life stress and social support on female collegiate gymnasts.
+\emph{Journal of Behavioral Medicine} (1992) \textbf{18}:127--138.
+doi:\href{https://doi.org/10.1080/08964289.1992.9936963}{10.1080/08964289.1992.9936963}
+
+\leavevmode\hypertarget{ref-Fornell1981}{}%
+45.
+Fornell C, Larcker DF.
+Evaluating structural equation models with unobservable variables and measurement error.
+\emph{Journal of Marketing Research} (1981) \textbf{18}:39--50.
+doi:\href{https://doi.org/10.2307/3151312}{10.2307/3151312}
+
+\leavevmode\hypertarget{ref-Corr2016c}{}%
+46.
+Corr PJ, Cooper AJ.
+The reinforcement sensitivity theory of personality questionnaire (RST-PQ): Development and validation.
+\emph{Psychological Assessment} (2016) \textbf{28}:1427--1440.
+doi:\href{https://doi.org/10.1037/pas0000273}{10.1037/pas0000273}
+
+\leavevmode\hypertarget{ref-Giles2016}{}%
+47.
+Giles D, Draper N, Neil W.
+Validity of the Polar V800 heart rate monitor to measure RR intervals at rest.
+\emph{European Journal of Applied Physiology} (2016) \textbf{116}:563--571.
+doi:\href{https://doi.org/10.1007/s00421-015-3303-9}{10.1007/s00421-015-3303-9}
 
 \leavevmode\hypertarget{ref-Caminal2018}{}%
-Caminal, P., Sola, F., Gomis, P., Guasch, E., Perera, A., Soriano, N., and Mont, L.
-(2018).
+48.
+Caminal P, Sola F, Gomis P, Guasch E, Perera A, Soriano N, et al.
 Validity of the Polar V800 monitor for measuring heart rate variability in mountain running route conditions.
-\emph{European Journal of Applied Physiology} 118, 669--677.
-doi:\href{https://doi.org/10.1007/s00421-018-3808-0}{10.1007/s00421-018-3808-0}.
-
-\leavevmode\hypertarget{ref-Chalmers2014}{}%
-Chalmers, J.
-A., Quintana, D.
-S., Abbott, M.
-J.-A., and Kemp, A.
-H.
-(2014).
-Anxiety disorders are associated with reduced heart rate variability: A meta-analysis.
-\emph{Frontiers in psychiatry} 5, 80.
-doi:\href{https://doi.org/10.3389/fpsyt.2014.00080}{10.3389/fpsyt.2014.00080}.
+\emph{European Journal of Applied Physiology} (2018) \textbf{118}:669--677.
+doi:\href{https://doi.org/10.1007/s00421-018-3808-0}{10.1007/s00421-018-3808-0}
+
+\leavevmode\hypertarget{ref-Plews2017}{}%
+49.
+Plews DJ, Scott B, Altini M, Wood M, Kilding AE, Laursen PB.
+Comparison of heart-rate-variability recording with smartphone photoplethysmography, polar H7 chest strap, and electrocardiography.
+\emph{International Journal of Sports Physiology and Performance} (2017) \textbf{12}:1324--1328.
+doi:\href{https://doi.org/10.1123/ijspp.2016-0668}{10.1123/ijspp.2016-0668}
+
+\leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
+50.
+Agyapong-Badu S, Warner M, Samuel D, Stokes M.
+Measurement of ageing effects on muscle tone and mechanical properties of rectus femoris and biceps brachii in healthy males and females using a novel hand-held myometric device.
+\emph{Archives of Gerontology and Geriatrics} (2016) \textbf{62}:59--67.
+doi:\href{https://doi.org/10.1016/j.archger.2015.09.011}{10.1016/j.archger.2015.09.011}
+
+\leavevmode\hypertarget{ref-Chuang2013}{}%
+51.
+Chuang LL, Lin KC, Wu CY, Chang CW, Chen HC, Yin HP, et al.
+Relative and absolute reliabilities of the myotonometric measurements of hemiparetic arms in patients with stroke.
+\emph{Archives of Physical Medicine and Rehabilitation} (2013) \textbf{94}:459--466.
+doi:\href{https://doi.org/10.1016/j.apmr.2012.08.212}{10.1016/j.apmr.2012.08.212}
+
+\leavevmode\hypertarget{ref-Pruyn2016}{}%
+52.
+Pruyn EC, Watsford ML, Murphy AJ.
+Validity and reliability of three methods of stiffness assessment.
+\emph{Journal of Sport and Health Science} (2016) \textbf{5}:476--483.
+doi:\href{https://doi.org/10.1016/j.jshs.2015.12.001}{10.1016/j.jshs.2015.12.001}
+
+\leavevmode\hypertarget{ref-Nair2014}{}%
+53.
+Nair K, Dougherty J, Schaefer E, Kelly J, Masi A.
+Repeatability, reproducibility, and calibration of the MyotonPRO on tissue mimicking phantoms.
+in \emph{ASME summer bioengineering conference}, 1--2.
+doi:\href{https://doi.org/10.1115/SBC2013-14622}{10.1115/SBC2013-14622}
+
+\leavevmode\hypertarget{ref-Aird2012}{}%
+54.
+Aird L, Samuel D, Stokes M.
+Quadriceps muscle tone, elasticity and stiffness in older males: Reliability and symmetry using the MyotonPRO.
+\emph{Archives of Gerontology and Geriatrics} (2012) \textbf{55}:e31--e39.
+doi:\href{https://doi.org/https://doi.org/10.1016/j.archger.2012.03.005}{https://doi.org/10.1016/j.archger.2012.03.005}
+
+\leavevmode\hypertarget{ref-Hunt2009}{}%
+55.
+Hunt TN, Ferrara MS, Bornstein RA, Baumgartner TA.
+The reliability of the modified balance error scoring system.
+\emph{Clinical Journal of Sport Medicine} (2009) \textbf{19}:471--475.
+doi:\href{https://doi.org/10.1097/JSM.0b013e3181c12c7b}{10.1097/JSM.0b013e3181c12c7b}
+
+\leavevmode\hypertarget{ref-McCrory2013}{}%
+56.
+McCrory P, Meeuwisse WH, Aubry M, Cantu RC, Dvořák J, Echemendia RJ, et al.
+Consensus statement on concussion in sport: The 4th international conference on concussion in sport, Zurich, November 2012.
+\emph{Journal of Athletic Training} (2013) \textbf{48}:554--575.
+doi:\href{https://doi.org/10.4085/1062-6050-48.4.05}{10.4085/1062-6050-48.4.05}
+
+\leavevmode\hypertarget{ref-VanCingel2017}{}%
+57.
+Cingel REH van, Hoogeboom TJ, Melick N van, Meddeler BM, Nijhuis-van der Sanden MWG.
+How to determine leg dominance: The agreement between self-reported and observed performance in healthy adults.
+\emph{Plos One} (2017) \textbf{12}:1--9.
+doi:\href{https://doi.org/10.1371/journal.pone.0189876}{10.1371/journal.pone.0189876}
+
+\leavevmode\hypertarget{ref-Riemann1999d}{}%
+58.
+Riemann BL, Guskiewicz KM, Shields EW.
+Relationship between clinical and forceplate measures of postural stability.
+\emph{Journal of Sport Rehabilitation} (1999) \textbf{8}:71--82.
+doi:\href{https://doi.org/10.1123/jsr.8.2.71}{10.1123/jsr.8.2.71}
+
+\leavevmode\hypertarget{ref-Koo2016}{}%
+59.
+Koo TK, Li MY.
+A guideline of selecting and reporting intraclass correlation coefficients for reliability research.
+\emph{Journal of Chiropractic Medicine} (2016) \textbf{15}:155--63.
+doi:\href{https://doi.org/10.1016/j.jcm.2016.02.012}{10.1016/j.jcm.2016.02.012}
+
+\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
+60. R Core Team. \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing (2019). Available at: \url{https://www.r-project.org/}
+
+\leavevmode\hypertarget{ref-Denver2007}{}%
+61.
+Denver JW, Reed SF, Porges SW.
+Methodological issues in the quantification of respiratory sinus arrhythmia.
+\emph{Biological Psychology} (2007) \textbf{74}:286--294.
+doi:\href{https://doi.org/10.1016/j.biopsycho.2005.09.005}{10.1016/j.biopsycho.2005.09.005}
+
+\leavevmode\hypertarget{ref-Rodriguez-Linares2017}{}%
+62. Rodriguez-Linares L, Vila X, Lado MJ, Mendez A, Otero A, Garcia CA. \emph{RHRV: Heart rate variability analysis of ecg data}. (2019). Available at: \url{https://CRAN.R-project.org/package=RHRV}
+
+\leavevmode\hypertarget{ref-Laborde2017}{}%
+63.
+Laborde S, Mosley E, Thayer JF.
+Heart rate variability and cardiac vagal tone in psychophysiological research - Recommendations for experiment planning, data analysis, and data reporting.
+\emph{Frontiers in Psychology} (2017) \textbf{8}:1--18.
+doi:\href{https://doi.org/10.3389/fpsyg.2017.00213}{10.3389/fpsyg.2017.00213}
+
+\leavevmode\hypertarget{ref-Malik1996}{}%
+64.
+Malik M, Camm AJ, Bigger JT, Breithardt G, Cerutti S, Cohen RJ, et al.
+Heart rate variability.
+Standards of measurement, physiological interpretation, and clinical use.
+\emph{European Heart Journal} (1996) \textbf{17}:354--381.
+doi:\href{https://doi.org/10.1093/oxfordjournals.eurheartj.a014868}{10.1093/oxfordjournals.eurheartj.a014868}
+
+\leavevmode\hypertarget{ref-Chuang2012}{}%
+65.
+Chuang LL, Wu CY, Lin KC.
+Reliability, validity, and responsiveness of myotonometric measurement of muscle tone, elasticity, and stiffness in patients with stroke.
+\emph{Archives of Physical Medicine and Rehabilitation} (2012) \textbf{93}:532--540.
+doi:\href{https://doi.org/10.1016/j.apmr.2011.09.014}{10.1016/j.apmr.2011.09.014}
+
+\leavevmode\hypertarget{ref-Morgan2018}{}%
+66.
+Morgan GE, Martin R, Williams L, Pearce O, Morris K.
+Objective assessment of stiffness in Achilles tendinopathy: a novel approach using the MyotonPRO.
+\emph{BMJ Open Sport \& Exercise Medicine} (2018) \textbf{4}:e000446.
+doi:\href{https://doi.org/10.1136/bmjsem-2018-000446}{10.1136/bmjsem-2018-000446}
+
+\leavevmode\hypertarget{ref-Valovich2003}{}%
+67.
+Valovich TC, Perrin DH, Gansneder BM.
+Repeat administration elicits a practice effect with the Balance Error Scoring System but not with the Standardized Assessment of Concussion in high school athletes.
+\emph{Journal of Athletic Training} (2003) \textbf{38}:51--56.
+Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/12937472}
+
+\leavevmode\hypertarget{ref-Pearl1988}{}%
+68. Pearl J. \emph{Probabilistic reasoning in intelligent systems: Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann Publishers (1988).
 
 \leavevmode\hypertarget{ref-Chen2012}{}%
-Chen, S.
-H., and Pollino, C.
-A.
-(2012).
+69.
+Chen SH, Pollino CA.
 Good practice in Bayesian network modelling.
-\emph{Environmental Modelling and Software} 37, 134--145.
-doi:\href{https://doi.org/10.1016/j.envsoft.2012.03.012}{10.1016/j.envsoft.2012.03.012}.
-
-\leavevmode\hypertarget{ref-Chuang2013}{}%
-Chuang, L.
-L., Lin, K.
-C., Wu, C.
-Y., Chang, C.
-W., Chen, H.
-C., Yin, H.
-P., and Wang, L.
-(2013).
-Relative and absolute reliabilities of the myotonometric measurements of hemiparetic arms in patients with stroke.
-\emph{Archives of Physical Medicine and Rehabilitation} 94, 459--466.
-doi:\href{https://doi.org/10.1016/j.apmr.2012.08.212}{10.1016/j.apmr.2012.08.212}.
-
-\leavevmode\hypertarget{ref-Chuang2012}{}%
-Chuang, L.
-L., Wu, C.
-Y., and Lin, K.
-C.
-(2012).
-Reliability, validity, and responsiveness of myotonometric measurement of muscle tone, elasticity, and stiffness in patients with stroke.
-\emph{Archives of Physical Medicine and Rehabilitation} 93, 532--540.
-doi:\href{https://doi.org/10.1016/j.apmr.2011.09.014}{10.1016/j.apmr.2011.09.014}.
-
-\leavevmode\hypertarget{ref-VanCingel2017}{}%
-Cingel, R.
-E.
-H.
-van, Hoogeboom, T.
-J., Melick, N.
-van, Meddeler, B.
-M., and Nijhuis-van der Sanden, M.
-W.
-G.
-(2017).
-How to determine leg dominance: The agreement between self-reported and observed performance in healthy adults.
-\emph{Plos One} 12, 1--9.
-doi:\href{https://doi.org/10.1371/journal.pone.0189876}{10.1371/journal.pone.0189876}.
-
-\leavevmode\hypertarget{ref-Corr2013}{}%
-Corr, P.
-J.
-(2013).
-Approach and avoidance behaviour: Multiple systems and their interactions.
-\emph{Emotion Review} 5, 285--290.
-doi:\href{https://doi.org/10.1177/1754073913477507}{10.1177/1754073913477507}.
-
-\leavevmode\hypertarget{ref-Corr2016c}{}%
-Corr, P.
-J., and Cooper, A.
-J.
-(2016).
-The reinforcement sensitivity theory of personality questionnaire (RST-PQ): Development and validation.
-\emph{Psychological Assessment} 28, 1427--1440.
-doi:\href{https://doi.org/10.1037/pas0000273}{10.1037/pas0000273}.
-
-\leavevmode\hypertarget{ref-Corr2016b}{}%
-Corr, P.
-J., McNaughton, N., Wilson, M.
-R., Hutchison, A., Burch, G., and Poropat, A.
-(2016).
-Neuroscience of motivation and organizational behavior: Putting the reinforcement sensitivity theory (RST) to work.
-\emph{Advances in Motivation and Achievement} 19, 65--92.
-doi:\href{https://doi.org/10.1108/S0749-742320160000019010}{10.1108/S0749-742320160000019010}.
-
-\leavevmode\hypertarget{ref-Denver2007}{}%
-Denver, J.
-W., Reed, S.
-F., and Porges, S.
-W.
-(2007).
-Methodological issues in the quantification of respiratory sinus arrhythmia.
-\emph{Biological Psychology} 74, 286--294.
-doi:\href{https://doi.org/10.1016/j.biopsycho.2005.09.005}{10.1016/j.biopsycho.2005.09.005}.
-
-\leavevmode\hypertarget{ref-Djaoui2017}{}%
-Djaoui, L., Haddad, M., Chamari, K., and Dellal, A.
-(2017).
-Monitoring training load and fatigue in soccer players with physiological markers.
-\emph{Physiology and Behavior} 181, 86--94.
-doi:\href{https://doi.org/10.1016/j.physbeh.2017.09.004}{10.1016/j.physbeh.2017.09.004}.
-
-\leavevmode\hypertarget{ref-Efron1993}{}%
-Efron, B., and Tibshirani, R.
-J.
-(1994).
-\emph{An introduction to the bootstrap}.
-New York: Chapman \& Hall Available at: \url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}.
-
-\leavevmode\hypertarget{ref-Enders2003}{}%
-Enders, C.
-K.
-(2003).
-Using the expectation maximization algorithm to estimate coefficient alpha for scales with item-level missing data.
-\emph{Psychological Methods} 8, 322--337.
-doi:\href{https://doi.org/10.1037/1082-989X.8.3.322}{10.1037/1082-989X.8.3.322}.
-
-\leavevmode\hypertarget{ref-Fawkner1999}{}%
-Fawkner, H.
-J., McMurrary, N.
-E., and Summers, J.
-J.
-(1999).
-Athletic injury and minor life events: A prospective study.
-\emph{Journal of Science and Medicine in Sport} 2, 117--124.
-doi:\href{https://doi.org/10.1016/S1440-2440(99)80191-1}{10.1016/S1440-2440(99)80191-1}.
-
-\leavevmode\hypertarget{ref-Fornell1981}{}%
-Fornell, C., and Larcker, D.
-F.
-(1981).
-Evaluating structural equation models with unobservable variables and measurement error.
-\emph{Journal of Marketing Research} 18, 39--50.
-doi:\href{https://doi.org/10.2307/3151312}{10.2307/3151312}.
-
-\leavevmode\hypertarget{ref-Fuller2006}{}%
-Fuller, C.
-W., Ekstrand, J., Junge, A., Andersen, T.
-E., Bahr, R., Dvorak, J., Hägglund, M., McCrory, P., and Meeuwisse, W.
-H.
-(2006).
-Consensus statement on injury definitions and data collection procedures in studies of football (soccer) injuries.
-\emph{British journal of sports medicine} 40, 193--201.
-doi:\href{https://doi.org/10.1136/bjsm.2005.025270}{10.1136/bjsm.2005.025270}.
-
-\leavevmode\hypertarget{ref-Fuller2007b}{}%
-Fuller, C.
-W., Molloy, M.
-G., Bagate, C., Bahr, R., Brooks, J.
-H.
-M., Donson, H., Kemp, S.
-P.
-T., McCrory, P., McIntosh, A.
-S., Meeuwisse, W.
-H., et al.
-(2007).
-Consensus statement on injury definitions and data collection procedures for studies of injuries in rugby union.
-\emph{British journal of sports medicine} 41, 328--31.
-doi:\href{https://doi.org/10.1136/bjsm.2006.033282}{10.1136/bjsm.2006.033282}.
+\emph{Environmental Modelling and Software} (2012) \textbf{37}:134--145.
+doi:\href{https://doi.org/10.1016/j.envsoft.2012.03.012}{10.1016/j.envsoft.2012.03.012}
+
+\leavevmode\hypertarget{ref-Ruiz-Perez2019}{}%
+70.
+Ruiz-Pérez I, Ayala F, Miguel Puerta J, Elvira JLL, De Ste Croix M, Hernández-Sánchez S, et al.
+A Bayesian Network approach to study the relationships between several neuromuscular performance measures and dynamic postural control in futsal players.
+\emph{PLoS ONE} (2019) \textbf{14}:1--18.
+doi:\href{https://doi.org/10.1371/journal.pone.0220065}{10.1371/journal.pone.0220065}
+
+\leavevmode\hypertarget{ref-Koller2009}{}%
+71.
+Koller D, Friedman N.
+\emph{Probabilistic graphical models: Principles and techniques - Adaptive computation and machine learning}.
+Cambridge, MA: The MIT Press (2009).
+doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}
 
 \leavevmode\hypertarget{ref-Fuster-Parra2017}{}%
-Fuster-Parra, P., Vidal-Conti, J., Borràs, P.
-A., and Palou, P.
-(2017).
+72.
+Fuster-Parra P, Vidal-Conti J, Borràs PA, Palou P.
 Bayesian networks to identify statistical dependencies.
 A case study of Spanish university students' habits.
-\emph{Informatics for Health and Social Care} 42, 166--179.
-doi:\href{https://doi.org/10.1080/17538157.2016.1178117}{10.1080/17538157.2016.1178117}.
-
-\leavevmode\hypertarget{ref-Galambos2005}{}%
-Galambos, S.
-A., Terry, P.
-C., Moyle, G.
-M., and Locke, S.
-A.
-(2005).
-Psychological predictors of injury among elite athletes.
-\emph{British Journal of Sports Medicine} 39, 351--354.
-doi:\href{https://doi.org/10.1136/bjsm.2005.018440}{10.1136/bjsm.2005.018440}.
-
-\leavevmode\hypertarget{ref-Giles2016}{}%
-Giles, D., Draper, N., and Neil, W.
-(2016).
-Validity of the Polar V800 heart rate monitor to measure RR intervals at rest.
-\emph{European Journal of Applied Physiology} 116, 563--571.
-doi:\href{https://doi.org/10.1007/s00421-015-3303-9}{10.1007/s00421-015-3303-9}.
-
-\leavevmode\hypertarget{ref-Gisselman2016}{}%
-Gisselman, A.
-S., Baxter, G.
-D., Wright, A., Hegedus, E., and Tumilty, S.
-(2016).
-Musculoskeletal overuse injuries and heart rate variability: Is there a link?
-\emph{Medical Hypotheses} 87, 1--7.
-doi:\href{https://doi.org/10.1016/j.mehy.2015.12.003}{10.1016/j.mehy.2015.12.003}.
-
-\leavevmode\hypertarget{ref-Heckerman1995}{}%
-Heckerman, D., Geiger, D., and Chickering, D.
-M.
-(1995).
-Learning Bayesian networks: The combination of knowledge and statistical data.
-\emph{Machine Learning} 20, 197--243.
-doi:\href{https://doi.org/10.1023/A:1022623210503}{10.1023/A:1022623210503}.
-
-\leavevmode\hypertarget{ref-Hrysomallis2011}{}%
-Hrysomallis, C. (2011). Balance ability and athletic performance. \emph{Sports Medicine} 41, 221--232. doi:\href{https://doi.org/10.2165/11538560-000000000-00000}{10.2165/11538560-000000000-00000}.
-
-\leavevmode\hypertarget{ref-Hughes2014}{}%
-Hughes, G.
-(2014).
-A review of recent perspectives on biomechanical risk factors associated with anterior cruciate ligament injury.
-\emph{Research in Sports Medicine} 22, 193--212.
-doi:\href{https://doi.org/10.1080/15438627.2014.881821}{10.1080/15438627.2014.881821}.
-
-\leavevmode\hypertarget{ref-Hunt2009}{}%
-Hunt, T.
-N., Ferrara, M.
-S., Bornstein, R.
-A., and Baumgartner, T.
-A.
-(2009).
-The reliability of the modified balance error scoring system.
-\emph{Clinical Journal of Sport Medicine} 19, 471--475.
-doi:\href{https://doi.org/10.1097/JSM.0b013e3181c12c7b}{10.1097/JSM.0b013e3181c12c7b}.
-
-\leavevmode\hypertarget{ref-Ivarsson2010}{}%
-Ivarsson, A., and Johnson, U.
-(2010).
-Psychological factors as predictors of injuries among senior soccer players.
-A prospective study.
-\emph{Journal of Sports Science and Medicine} 9, 347--352.
-Available at: \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3761721/}.
-
-\leavevmode\hypertarget{ref-Ivarsson2017}{}%
-Ivarsson, A., Johnson, U., Andersen, M.
-B., Tranaeus, U., Stenling, A., and Lindwall, M.
-(2017).
-Psychosocial factors and sport injuries: Meta-analyses for prediction and prevention.
-\emph{Sports Medicine} 47, 353--365.
-doi:\href{https://doi.org/10.1007/s40279-016-0578-x}{10.1007/s40279-016-0578-x}.
-
-\leavevmode\hypertarget{ref-Ivarsson2014}{}%
-Ivarsson, A., Johnson, U., Lindwall, M., Gustafsson, H., and Altemyr, M.
-(2014).
-Psychosocial stress as a predictor of injury in elite junior soccer: A latent growth curve analysis.
-\emph{Journal of Science and Medicine in Sport} 17, 366--370.
-doi:\href{https://doi.org/10.1016/j.jsams.2013.10.242}{10.1016/j.jsams.2013.10.242}.
-
-\leavevmode\hypertarget{ref-Johnson2014}{}%
-Johnson, U., Tranaeus, U., and Ivarsson, A.
-(2014).
-Current status and future challenges in psychological research of sport injury prediction and prevention: A methodological perspective.
-\emph{Revista de Psicologia del Deporte} 23, 401--409.
-Available at: \url{https://www.rpd-online.com/article/view/v23-n2-johnson-tranaeus-ivarsson}.
-
-\leavevmode\hypertarget{ref-Koller2009}{}%
-Koller, D., and Friedman, N.
-(2009).
-\emph{Probabilistic graphical models: Principles and techniques - Adaptive computation and machine learning}.
-Cambridge, MA: The MIT Press doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.
-
-\leavevmode\hypertarget{ref-Koo2016}{}%
-Koo, T.
-K., and Li, M.
-Y.
-(2016).
-A guideline of selecting and reporting intraclass correlation coefficients for reliability research.
-\emph{Journal of Chiropractic Medicine} 15, 155--63.
-doi:\href{https://doi.org/10.1016/j.jcm.2016.02.012}{10.1016/j.jcm.2016.02.012}.
-
-\leavevmode\hypertarget{ref-Kuhn2008}{}%
-Kuhn, M.
-C.
-from J.
-W., Weston, S., Williams, A., Keefer, C., and Engelhardt, A.
-(2008).
-caret: Classification and Regression Training.
-\emph{Journal of Statistical Software} 28.
-doi:\href{https://doi.org/10.1053/j.sodo.2009.03.002}{10.1053/j.sodo.2009.03.002}.
-
-\leavevmode\hypertarget{ref-Kumar2001}{}%
-Kumar, S. (2001). Theories of musculoskeletal injury causation. \emph{Ergonomics} 44, 17--47. doi:\href{https://doi.org/10.1080/00140130120716}{10.1080/00140130120716}.
-
-\leavevmode\hypertarget{ref-Laborde2017}{}%
-Laborde, S., Mosley, E., and Thayer, J.
-F.
-(2017).
-Heart rate variability and cardiac vagal tone in psychophysiological research - Recommendations for experiment planning, data analysis, and data reporting.
-\emph{Frontiers in Psychology} 8, 1--18.
-doi:\href{https://doi.org/10.3389/fpsyg.2017.00213}{10.3389/fpsyg.2017.00213}.
-
-\leavevmode\hypertarget{ref-Lang2014}{}%
-Lang, K.
-M., Jorgensen, T.
-D., Moore, E.
-W.
-G., and Little, T.
-D.
-(2013).
-On the joys of missing data.
-\emph{Journal of Pediatric Psychology} 39, 151--162.
-doi:\href{https://doi.org/10.1093/jpepsy/jst048}{10.1093/jpepsy/jst048}.
-
-\leavevmode\hypertarget{ref-Lavallee1996}{}%
-Lavallée, L., and Flint, F.
-(1996).
-The relationship of stress, competitive anxiety, mood state, and social support to athletic injury.
-\emph{Journal of Athletic Training} 31, 296--299.
-Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16558413}.
-
-\leavevmode\hypertarget{ref-Leddy1994}{}%
-Leddy, M.
-H., Lambert, M.
-J., and Ogles, B.
-M.
-(1994).
-Psychological consequences of athletic injury among high-level competitors.
-\emph{Research Quarterly for Exercise and Sport} 65, 347--354.
-doi:\href{https://doi.org/10.1080/02701367.1994.10607639}{10.1080/02701367.1994.10607639}.
-
-\leavevmode\hypertarget{ref-Lee2017}{}%
-Lee, E.
-C., Fragala, M.
-S., Kavouras, S.
-A., Queen, R.
-M., Pryor, J.
-L., and Casa, D.
-J.
-(2017).
-Biomarkers in sports and exercise: Tracking health, performance, and recovery in athletes.
-\emph{Journal of Strength and Conditioning Research} 31, 2920--2937.
-doi:\href{https://doi.org/10.1519/JSC.0000000000002122}{10.1519/JSC.0000000000002122}.
-
-\leavevmode\hypertarget{ref-Maddison2005}{}%
-Maddison, R., and Prapavessis, H.
-(2005).
-A psychological approach to the prediction and prevention of athletic injury.
-\emph{Journal of Sport and Exercise Psychology} 27, 289--310.
-doi:\href{https://doi.org/10.1123/jsep.27.3.289}{10.1123/jsep.27.3.289}.
-
-\leavevmode\hypertarget{ref-Malik1996}{}%
-Malik, M., Camm, A.
-J., Bigger, J.
-T., Breithardt, G., Cerutti, S., Cohen, R.
-J., Coumel, P., Fallen, E.
-L., Kennedy, H.
-L., Kleiger, R.
-E., et al.
-(1996).
-Heart rate variability.
-Standards of measurement, physiological interpretation, and clinical use.
-\emph{European Heart Journal} 17, 354--381.
-doi:\href{https://doi.org/10.1093/oxfordjournals.eurheartj.a014868}{10.1093/oxfordjournals.eurheartj.a014868}.
-
-\leavevmode\hypertarget{ref-McCrory2013}{}%
-McCrory, P., Meeuwisse, W.
-H., Aubry, M., Cantu, R.
-C., Dvořák, J., Echemendia, R.
-J., Engebretsen, L., Johnston, K., Kutcher, J.
-S., Raftery, M., et al.
-(2013).
-Consensus statement on concussion in sport: The 4th international conference on concussion in sport, Zurich, November 2012.
-\emph{Journal of Athletic Training} 48, 554--575.
-doi:\href{https://doi.org/10.4085/1062-6050-48.4.05}{10.4085/1062-6050-48.4.05}.
-
-\leavevmode\hypertarget{ref-Meeuwisse2007}{}%
-Meeuwisse, W.
-H., Tyreman, H., Hagel, B., and Emery, C.
-(2007).
-A dynamic model of etiology in sport injury: The recursive nature of risk and causation.
-\emph{Clinical Journal of Sport Medicine} 17, 215--219.
-doi:\href{https://doi.org/10.1097/JSM.0b013e3180592a48}{10.1097/JSM.0b013e3180592a48}.
-
-\leavevmode\hypertarget{ref-Morgan2018}{}%
-Morgan, G.
-E., Martin, R., Williams, L., Pearce, O., and Morris, K.
-(2018).
-Objective assessment of stiffness in Achilles tendinopathy: a novel approach using the MyotonPRO.
-\emph{BMJ Open Sport \& Exercise Medicine} 4, e000446.
-doi:\href{https://doi.org/10.1136/bmjsem-2018-000446}{10.1136/bmjsem-2018-000446}.
-
-\leavevmode\hypertarget{ref-Murphy2003}{}%
-Murphy, D.
-F., Connolly, D.
-A.
-J., and Beynnon, B.
-D.
-(2003).
-Risk factors for lower extremity injury: A review of the literature.
-\emph{British Journal of Sports Medicine} 37, 13--29.
-doi:\href{https://doi.org/10.1136/bjsm.37.1.13}{10.1136/bjsm.37.1.13}.
-
-\leavevmode\hypertarget{ref-Murphy2002}{}%
-Murphy, K. (2002). Dynamic Bayesian networks: Representation, inference and learning.
-
-\leavevmode\hypertarget{ref-Nair2014}{}%
-Nair, K., Dougherty, J., Schaefer, E., Kelly, J., and Masi, A.
-(2014).
-Repeatability, reproducibility, and calibration of the MyotonPRO on tissue mimicking phantoms.
-in \emph{ASME summer bioengineering conference}, 1--2.
-doi:\href{https://doi.org/10.1115/SBC2013-14622}{10.1115/SBC2013-14622}.
-
-\leavevmode\hypertarget{ref-Neely1998}{}%
-Neely, F.
-G.
-(1998).
-Biomechanical risk factors for exercise-related lower limb injuries.
-\emph{Sports Medicine} 26, 395--413.
-doi:\href{https://doi.org/10.2165/00007256-199826060-00003}{10.2165/00007256-199826060-00003}.
-
-\leavevmode\hypertarget{ref-Olmedilla2018}{}%
-Olmedilla, A., Rubio, V.
-J., Fuster-Parra, P., Pujals, C., and García-Mas, A.
-(2018).
-A Bayesian approach to sport injuries likelihood: Does player's self-efficacy and environmental factors plays the main role?
-\emph{Frontiers in Psychology} 9, 1--10.
-doi:\href{https://doi.org/10.3389/fpsyg.2018.01174}{10.3389/fpsyg.2018.01174}.
-
-\leavevmode\hypertarget{ref-Paillard2006}{}%
-Paillard, T., Noé, F., Rivière, T., Marion, V., Montoya, R., and Dupui, P.
-(2006).
-Postural performance and strategy in the unipedal stance of soccer players at different levels of competition.
-\emph{Journal of Athletic Training} 41, 172--176.
-Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16791302}.
-
-\leavevmode\hypertarget{ref-Passer1983a}{}%
-Passer, M.
-W., and Seese, M.
-D.
-(1983).
-Life stress and athletic injury: Examination of positive versus negative events and three moderator variables.
-\emph{Journal of Human Stress} 9, 11--16.
-doi:\href{https://doi.org/10.1080/0097840X.1983.9935025}{10.1080/0097840X.1983.9935025}.
-
-\leavevmode\hypertarget{ref-Pearl1988}{}%
-Pearl, J. (1988). \emph{Probabilistic reasoning in intelligent systems: Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann Publishers.
-
-\leavevmode\hypertarget{ref-Perna2003}{}%
-Perna, F.
-M., Antoni, M.
-H., Baum, A., Gordon, P., and Schneiderman, N.
-(2003).
-Cognitive behavioral stress management effects on injury and illness among competitive athletes: A randomized clinical trial.
-\emph{Annals of Behavioral Medicine} 25, 66--73.
-Available at: \url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}.
-
-\leavevmode\hypertarget{ref-Perna1995}{}%
-Perna, F.
-M., and McDowell, S.
-L.
-(1995).
-Role of psychological stress in cortisol recovery from exhaustive exercise among elite athletes.
-\emph{International Journal of Behavioral Medicine} 2, 13--26.
-doi:\href{https://doi.org/10.1207/s15327558ijbm0201_2}{10.1207/s15327558ijbm0201\_2}.
-
-\leavevmode\hypertarget{ref-Perna1997}{}%
-Perna, F., Schneiderman, N., and LaPerriere, A.
-(1997).
-Psychological stress, exercise and immunity.
-\emph{International Journal of Sports Medicine} 18, 78--83.
-doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}.
-
-\leavevmode\hypertarget{ref-Petrie1993}{}%
-Petrie, T.
-A.
-(1993).
-Coping skills, competitive trait anxiety, and playing status: Moderating effects on the life stress-injury relationship.
-\emph{Journal of Sport and Exercise Psychology} 15, 261--274.
-doi:\href{https://doi.org/10.1123/jsep.15.3.261}{10.1123/jsep.15.3.261}.
-
-\leavevmode\hypertarget{ref-Petrie1992}{}%
-Petrie, T.
-A.
-(1992).
-Psychosocial antecedents of athletic injury: The effects of life stress and social support on female collegiate gymnasts.
-\emph{Journal of Behavioral Medicine} 18, 127--138.
-doi:\href{https://doi.org/10.1080/08964289.1992.9936963}{10.1080/08964289.1992.9936963}.
-
-\leavevmode\hypertarget{ref-Petrie2014}{}%
-Petrie, T.
-A., Deiters, J., and Harmison, R.
-J.
-(2014).
-Mental toughness, social support, and athletic identity: Moderators of the life stress-injury relationship in collegiate football players.
-\emph{Sport, Exercise, and Performance Psychology} 3, 13--27.
-doi:\href{https://doi.org/10.1037/a0032698}{10.1037/a0032698}.
-
-\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
-Pickering-Rodriguez, E.
-C., Watsford, M.
-L., Bower, R.
-G., and Murphy, A.
-J.
-(2017).
-The relationship between lower body stiffness and injury incidence in female netballers.
-\emph{Sports Biomechanics} 16, 361--373.
-doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}.
-
-\leavevmode\hypertarget{ref-Plews2017}{}%
-Plews, D.
-J., Scott, B., Altini, M., Wood, M., Kilding, A.
-E., and Laursen, P.
-B.
-(2017).
-Comparison of heart-rate-variability recording with smartphone photoplethysmography, polar H7 chest strap, and electrocardiography.
-\emph{International Journal of Sports Physiology and Performance} 12, 1324--1328.
-doi:\href{https://doi.org/10.1123/ijspp.2016-0668}{10.1123/ijspp.2016-0668}.
-
-\leavevmode\hypertarget{ref-Pruyn2015}{}%
-Pruyn, E.
-C., Watsford, M.
-L., and Murphy, A.
-J.
-(2015).
-Differences in lower-body stiffness between levels of netball competition.
-\emph{Journal of Strength and Conditioning Research} 29, 1197--1202.
-doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}.
-
-\leavevmode\hypertarget{ref-Pruyn2016}{}%
-Pruyn, E.
-C., Watsford, M.
-L., and Murphy, A.
-J.
-(2016).
-Validity and reliability of three methods of stiffness assessment.
-\emph{Journal of Sport and Health Science} 5, 476--483.
-doi:\href{https://doi.org/10.1016/j.jshs.2015.12.001}{10.1016/j.jshs.2015.12.001}.
-
-\leavevmode\hypertarget{ref-Pulopulos2018}{}%
-Pulopulos, M.
-M., Vanderhasselt, M.
-A., and De Raedt, R.
-(2018).
-Association between changes in heart rate variability during the anticipation of a stressful situation and the stress-induced cortisol response.
-\emph{Psychoneuroendocrinology} 94, 63--71.
-doi:\href{https://doi.org/10.1016/j.psyneuen.2018.05.004}{10.1016/j.psyneuen.2018.05.004}.
-
-\leavevmode\hypertarget{ref-Qian2015}{}%
-Qian, S.
-S., and Miltner, R.
-J.
-(2015).
-A continuous variable Bayesian networks model for water quality modeling: A case study of setting nitrogen criterion for small rivers and streams in Ohio, USA.
-\emph{Environmental Modelling and Software} 69, 14--22.
-doi:\href{https://doi.org/10.1016/j.envsoft.2015.03.001}{10.1016/j.envsoft.2015.03.001}.
-
-\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing Available at: \url{https://www.r-project.org/}.
-
-\leavevmode\hypertarget{ref-Riemann1999d}{}%
-Riemann, B.
-L., Guskiewicz, K.
-M., and Shields, E.
-W.
-(1999).
-Relationship between clinical and forceplate measures of postural stability.
-\emph{Journal of Sport Rehabilitation} 8, 71--82.
-doi:\href{https://doi.org/10.1123/jsr.8.2.71}{10.1123/jsr.8.2.71}.
-
-\leavevmode\hypertarget{ref-Rodriguez-Linares2017}{}%
-Rodriguez-Linares, L., Vila, X., Lado, M.
-J., Mendez, A., Otero, A., and Garcia, C.
-A.
-(2019).
-\emph{RHRV: Heart rate variability analysis of ecg data}.
-Available at: \url{https://CRAN.R-project.org/package=RHRV}.
-
-\leavevmode\hypertarget{ref-Rogers2005}{}%
-Rogers, T., and Landers, D.
-M.
-(2005).
-Mediating effects of peripheral vision in the life event stress/athletic injury relationship.
-\emph{Journal of Sport and Exercise Psychology} 27, 271--288.
-doi:\href{https://doi.org/10.1002/9781444303650}{10.1002/9781444303650}.
-
-\leavevmode\hypertarget{ref-Romero-Franco2014}{}%
-Romero-Franco, N., Gallego-Izquierdo, T., Martínez-López, E.
-J., Hita-Contreras, F., Osuna-Pére, Catalina, M., and Martínez-Amat, A.
-(2014).
-Postural stability and subsequent sports injuries during indoor season of athletes.
-\emph{Journal of Physical Therapy Science} 26, 683--687.
-doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}.
-
-\leavevmode\hypertarget{ref-Rosa2014}{}%
-Rosa, B.
-B., Asperti, A.
-M., Helito, C.
-P., Demange, M.
-K., Fernandes, T.
-L., and Hernandez, A.
-J.
-(2014).
-Epidemiology of sports injuries on collegiate athletes at a single center.
-\emph{Acta Ortopédica Brasileira} 22, 321--324.
-doi:\href{https://doi.org/10.1590/1413-78522014220601007}{10.1590/1413-78522014220601007}.
-
-\leavevmode\hypertarget{ref-Ross2011}{}%
-Ross, S.
-E., Linens, S.
-W., Wright, C.
-J., and Arnold, B.
-L.
-(2011).
-Balance assessments for predicting functional ankle instability and stable ankles.
-\emph{Gait and Posture} 34, 539--542.
-doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}.
-
-\leavevmode\hypertarget{ref-Ruiz-Perez2019}{}%
-Ruiz-Pérez, I., Ayala, F., Miguel Puerta, J., Elvira, J.
-L.
-L., De Ste Croix, M., Hernández-Sánchez, S., and Vera-Garcia, F.
-J.
-(2019).
-A Bayesian Network approach to study the relationships between several neuromuscular performance measures and dynamic postural control in futsal players.
-\emph{PLoS ONE} 14, 1--18.
-doi:\href{https://doi.org/10.1371/journal.pone.0220065}{10.1371/journal.pone.0220065}.
-
-\leavevmode\hypertarget{ref-Norvig2009}{}%
-Russell, S. J., and Norvig, P. (2009). \emph{Artificial Intelligence: A Modern Approach}. 3rd ed. Prentice Hall Available at: \url{http://aima.cs.berkeley.edu/}.
-
-\leavevmode\hypertarget{ref-Scutari2014}{}%
-Scutari, M., and Denis, J.-B.
-(2014).
-\emph{Bayesian networks: with examples in R}.
-1st ed.
-Chapman \& Hall/CRC Available at: \url{https://www.crcpress.com/Bayesian-Networks-With-Examples-in-R/Scutari-Denis/p/book/9781482225587}.
-
-\leavevmode\hypertarget{ref-Scutari2013}{}%
-Scutari, M., and Nagarajan, R.
-(2013).
-Identifying significant edges in graphical models of molecular networks.
-\emph{Artificial Intelligence in Medicine} 57, 207--217.
-doi:\href{https://doi.org/10.1016/j.artmed.2012.12.006}{10.1016/j.artmed.2012.12.006}.
-
-\leavevmode\hypertarget{ref-Sheu2016}{}%
-Sheu, Y., Chen, L.
-H., and Hedegaard, H.
-(2016).
-Sports- and recreation-related injury episodes in the United States, 2011-2014.
-\emph{National Health Statistics Reports}, 1--12.
-Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/27906643}.
-
-\leavevmode\hypertarget{ref-Slimani2018}{}%
-Slimani, M., Bragazzi, N.
-L., Znazen, H., Paravlic, A., Azaiez, F., and Tod, D.
-(2018).
-Psychosocial predictors and psychological prevention of soccer injuries: A systematic review and meta-analysis of the literature.
-\emph{Physical Therapy in Sport} 32, 293--300.
-doi:\href{https://doi.org/10.1016/j.ptsp.2018.05.006}{10.1016/j.ptsp.2018.05.006}.
-
-\leavevmode\hypertarget{ref-Smith1990}{}%
-Smith, R.
-E., Smoll, F.
-L., and Ptacek, J.
-T.
-(1990).
-Conjunctive moderator variables in vulnerability and resiliency research: Life stress, social support and coping skills, and adolescent sport injuries.
-\emph{Journal of Personality and Social Psychology} 58, 360--370.
-doi:\href{https://doi.org/10.1037/0022-3514.58.2.360}{10.1037/0022-3514.58.2.360}.
-
-\leavevmode\hypertarget{ref-Swanik2007}{}%
-Swanik, C.
-B., Covassin, T., Stearne, D.
-J., and Schatz, P.
-(2007).
-The relationship between neurocognitive function and noncontact anterior cruciate ligament injuries.
-\emph{American Journal of Sports Medicine} 35, 943--948.
-doi:\href{https://doi.org/10.1177/0363546507299532}{10.1177/0363546507299532}.
-
-\leavevmode\hypertarget{ref-Timpka2014}{}%
-Timpka, T., Alonso, J.
-M., Jacobsson, J., Junge, A., Branco, P., Clarsen, B., Kowalski, J., Mountjoy, M., Nilsson, S., Pluim, B., et al.
-(2014).
-Injury and illness definitions and data collection procedures for use in epidemiological studies in Athletics (track and field): Consensus statement.
-\emph{British Journal of Sports Medicine} 48, 483--490.
-doi:\href{https://doi.org/10.1136/bjsports-2013-093241}{10.1136/bjsports-2013-093241}.
-
-\leavevmode\hypertarget{ref-Trojian2006}{}%
-Trojian, T.
-H., and McKeag, D.
-B.
-(2006).
-Single leg balance test to identify risk of ankle sprains.
-\emph{British Journal of Sports Medicine} 40, 610--613.
-doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}.
-
-\leavevmode\hypertarget{ref-Valovich2003}{}%
-Valovich, T.
-C., Perrin, D.
-H., and Gansneder, B.
-M.
-(2003).
-Repeat administration elicits a practice effect with the Balance Error Scoring System but not with the Standardized Assessment of Concussion in high school athletes.
-\emph{Journal of Athletic Training} 38, 51--56.
-Available at: \url{https://www.ncbi.nlm.nih.gov/pubmed/12937472}.
-
-\leavevmode\hypertarget{ref-Wiese-Bjornstal2009}{}%
-Wiese-Bjornstal, D.
-M.
-(2009).
-Sport injury and college athlete health across the lifespan.
-\emph{Journal of Intercollegiate Sport} 2, 64--80.
-doi:\href{https://doi.org/10.1123/jis.2.1.64}{10.1123/jis.2.1.64}.
-
-\leavevmode\hypertarget{ref-Wilkerson2012a}{}%
-Wilkerson, G.
-B.
-(2012).
-Neurocognitive reaction time predicts lower extremity sprains and strains.
-\emph{International Journal of Athletic Therapy and Training} 17, 4--9.
-doi:\href{https://doi.org/10.1123/ijatt.17.6.4}{10.1123/ijatt.17.6.4}.
-
-\leavevmode\hypertarget{ref-Williams2007}{}%
-Williams, J.
-M., and Andersen, M.
-B.
-(2007).
-``Psychosocial antecedents of sport injury and interventions for risk reduction,'' in \emph{Handbook of sport psychology}, eds.
-G.
-Tenenbaum and R.
-C.
-Eklund (Hoboken, NJ, USA: Wiley), 379--403.
-doi:\href{https://doi.org/10.1002/9781118270011.ch17}{10.1002/9781118270011.ch17}.
-
-\leavevmode\hypertarget{ref-Williams1998}{}%
-Williams, J.
-M., and Andersen, M.
-B.
-(1998).
-Psychosocial antecedents of sport injury: review and critique of the stress and injury model.
-\emph{Journal of Applied Sport Psychology} 10, 5--25.
-doi:\href{https://doi.org/10.1080/10413209808406375}{10.1080/10413209808406375}.
-
-\leavevmode\hypertarget{ref-Williams2017}{}%
-Williams, S., Booton, T., Watson, M., Rowland, D., and Altini, M.
-(2017).
-Heart rate variability is a moderating factor in the workload-injury relationship of competitive crossfit™ athletes.
-\emph{Journal of Sports Science and Medicine} 16, 443--449.
+\emph{Informatics for Health and Social Care} (2017) \textbf{42}:166--179.
+doi:\href{https://doi.org/10.1080/17538157.2016.1178117}{10.1080/17538157.2016.1178117}
 
 \leavevmode\hypertarget{ref-Xiao-xuan2007}{}%
-Xiao-xuan, H., Hui, W., and Shuo, W.
-(2008).
+73.
+Xiao-xuan H, Hui W, Shuo W.
 Using expert's knowledge to build Bayesian networks.
 in \emph{International conference on computational intelligence and security workshops} (IEEE), 220--223.
-doi:\href{https://doi.org/10.1109/cisw.2007.4425484}{10.1109/cisw.2007.4425484}.
+doi:\href{https://doi.org/10.1109/cisw.2007.4425484}{10.1109/cisw.2007.4425484}
+
+\leavevmode\hypertarget{ref-Heckerman1995}{}%
+74.
+Heckerman D, Geiger D, Chickering DM.
+Learning Bayesian networks: The combination of knowledge and statistical data.
+\emph{Machine Learning} (1995) \textbf{20}:197--243.
+doi:\href{https://doi.org/10.1023/A:1022623210503}{10.1023/A:1022623210503}
 
 \leavevmode\hypertarget{ref-Xu2015}{}%
-Xu, J.
-G., Zhao, Y., Chen, J., and Han, C.
-(2015).
+75.
+Xu JG, Zhao Y, Chen J, Han C.
 A structure learning algorithm for Bayesian network using prior knowledge.
-\emph{Journal of Computer Science and Technology} 30, 713--724.
-doi:\href{https://doi.org/10.1007/s11390-015-1556-8}{10.1007/s11390-015-1556-8}.
+\emph{Journal of Computer Science and Technology} (2015) \textbf{30}:713--724.
+doi:\href{https://doi.org/10.1007/s11390-015-1556-8}{10.1007/s11390-015-1556-8}
+
+\leavevmode\hypertarget{ref-Enders2003}{}%
+76.
+Enders CK.
+Using the expectation maximization algorithm to estimate coefficient alpha for scales with item-level missing data.
+\emph{Psychological Methods} (2003) \textbf{8}:322--337.
+doi:\href{https://doi.org/10.1037/1082-989X.8.3.322}{10.1037/1082-989X.8.3.322}
+
+\leavevmode\hypertarget{ref-Lang2014}{}%
+77.
+Lang KM, Jorgensen TD, Moore EWG, Little TD.
+On the joys of missing data.
+\emph{Journal of Pediatric Psychology} (2013) \textbf{39}:151--162.
+doi:\href{https://doi.org/10.1093/jpepsy/jst048}{10.1093/jpepsy/jst048}
+
+\leavevmode\hypertarget{ref-Kuhn2008}{}%
+78.
+Kuhn MC from JW, Weston S, Williams A, Keefer C, Engelhardt A.
+caret: Classification and Regression Training.
+\emph{Journal of Statistical Software} (2008) \textbf{28}: doi:\href{https://doi.org/10.1053/j.sodo.2009.03.002}{10.1053/j.sodo.2009.03.002}
+
+\leavevmode\hypertarget{ref-Murphy2002}{}%
+79. Murphy K. Dynamic Bayesian networks: Representation, inference and learning. (2002)
+
+\leavevmode\hypertarget{ref-Norvig2009}{}%
+80. Russell SJ, Norvig P. \emph{Artificial Intelligence: A Modern Approach}. 3rd ed. Prentice Hall (2009). Available at: \url{http://aima.cs.berkeley.edu/}
+
+\leavevmode\hypertarget{ref-Efron1993}{}%
+81.
+Efron B, Tibshirani RJ.
+\emph{An introduction to the bootstrap}.
+New York: Chapman \& Hall (1994).
+Available at: \url{https://www.crcpress.com/An-Introduction-to-the-Bootstrap/Efron-Tibshirani/p/book/9780412042317}
+
+\leavevmode\hypertarget{ref-Scutari2013}{}%
+82.
+Scutari M, Nagarajan R.
+Identifying significant edges in graphical models of molecular networks.
+\emph{Artificial Intelligence in Medicine} (2013) \textbf{57}:207--217.
+doi:\href{https://doi.org/10.1016/j.artmed.2012.12.006}{10.1016/j.artmed.2012.12.006}
+
+\leavevmode\hypertarget{ref-Burkner2017a}{}%
+83.
+Bürkner PC.
+brms: An R package for Bayesian multilevel models using Stan.
+\emph{Journal of Statistical Software} (2017) \textbf{80}: doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}
+
+\leavevmode\hypertarget{ref-Trojian2006}{}%
+84.
+Trojian TH, McKeag DB.
+Single leg balance test to identify risk of ankle sprains.
+\emph{British Journal of Sports Medicine} (2006) \textbf{40}:610--613.
+doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}
+
+\leavevmode\hypertarget{ref-Butler2003}{}%
+85.
+Butler RJ, Crowell HP, Davis IMC.
+Lower extremity stiffness: Implications for performance and injury.
+\emph{Clinical Biomechanics} (2003) \textbf{18}:511--517.
+doi:\href{https://doi.org/10.1016/S0268-0033(03)00071-8}{10.1016/S0268-0033(03)00071-8}
+
+\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
+86.
+Pickering-Rodriguez EC, Watsford ML, Bower RG, Murphy AJ.
+The relationship between lower body stiffness and injury incidence in female netballers.
+\emph{Sports Biomechanics} (2017) \textbf{16}:361--373.
+doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}
+
+\leavevmode\hypertarget{ref-Paillard2006}{}%
+87.
+Paillard T, Noé F, Rivière T, Marion V, Montoya R, Dupui P.
+Postural performance and strategy in the unipedal stance of soccer players at different levels of competition.
+\emph{Journal of Athletic Training} (2006) \textbf{41}:172--176.
+Available at: \url{http://www.ncbi.nlm.nih.gov/pubmed/16791302}
+
+\leavevmode\hypertarget{ref-Hrysomallis2011}{}%
+88.
+Hrysomallis C.
+Balance ability and athletic performance.
+\emph{Sports Medicine} (2011) \textbf{41}:221--232.
+doi:\href{https://doi.org/10.2165/11538560-000000000-00000}{10.2165/11538560-000000000-00000}
+
+\leavevmode\hypertarget{ref-Ivarsson2014}{}%
+89.
+Ivarsson A, Johnson U, Lindwall M, Gustafsson H, Altemyr M.
+Psychosocial stress as a predictor of injury in elite junior soccer: A latent growth curve analysis.
+\emph{Journal of Science and Medicine in Sport} (2014) \textbf{17}:366--370.
+doi:\href{https://doi.org/10.1016/j.jsams.2013.10.242}{10.1016/j.jsams.2013.10.242}
+
+\leavevmode\hypertarget{ref-Fawkner1999}{}%
+90.
+Fawkner HJ, McMurrary NE, Summers JJ.
+Athletic injury and minor life events: A prospective study.
+\emph{Journal of Science and Medicine in Sport} (1999) \textbf{2}:117--124.
+doi:\href{https://doi.org/10.1016/S1440-2440(99)80191-1}{10.1016/S1440-2440(99)80191-1}
+
+\leavevmode\hypertarget{ref-Petrie1993}{}%
+91.
+Petrie TA.
+Coping skills, competitive trait anxiety, and playing status: Moderating effects on the life stress-injury relationship.
+\emph{Journal of Sport and Exercise Psychology} (1993) \textbf{15}:261--274.
+doi:\href{https://doi.org/10.1123/jsep.15.3.261}{10.1123/jsep.15.3.261}
+
+\leavevmode\hypertarget{ref-Petrie2014}{}%
+92.
+Petrie TA, Deiters J, Harmison RJ.
+Mental toughness, social support, and athletic identity: Moderators of the life stress-injury relationship in collegiate football players.
+\emph{Sport, Exercise, and Performance Psychology} (2014) \textbf{3}:13--27.
+doi:\href{https://doi.org/10.1037/a0032698}{10.1037/a0032698}
+
+\leavevmode\hypertarget{ref-Gisselman2016}{}%
+93.
+Gisselman AS, Baxter GD, Wright A, Hegedus E, Tumilty S.
+Musculoskeletal overuse injuries and heart rate variability: Is there a link?
+\emph{Medical Hypotheses} (2016) \textbf{87}:1--7.
+doi:\href{https://doi.org/10.1016/j.mehy.2015.12.003}{10.1016/j.mehy.2015.12.003}
+
+\leavevmode\hypertarget{ref-Corr2016b}{}%
+94.
+Corr PJ, McNaughton N, Wilson MR, Hutchison A, Burch G, Poropat A.
+Neuroscience of motivation and organizational behavior: Putting the reinforcement sensitivity theory (RST) to work.
+\emph{Advances in Motivation and Achievement} (2016) \textbf{19}:65--92.
+doi:\href{https://doi.org/10.1108/S0749-742320160000019010}{10.1108/S0749-742320160000019010}
+
+\leavevmode\hypertarget{ref-Corr2013}{}%
+95.
+Corr PJ.
+Approach and avoidance behaviour: Multiple systems and their interactions.
+\emph{Emotion Review} (2013) \textbf{5}:285--290.
+doi:\href{https://doi.org/10.1177/1754073913477507}{10.1177/1754073913477507}
+
+\leavevmode\hypertarget{ref-Chalmers2014}{}%
+96.
+Chalmers JA, Quintana DS, Abbott MJ-A, Kemp AH.
+Anxiety disorders are associated with reduced heart rate variability: A meta-analysis.
+\emph{Frontiers in psychiatry} (2014) \textbf{5}:80.
+doi:\href{https://doi.org/10.3389/fpsyt.2014.00080}{10.3389/fpsyt.2014.00080}
+
+\leavevmode\hypertarget{ref-Pulopulos2018}{}%
+97.
+Pulopulos MM, Vanderhasselt MA, De Raedt R.
+Association between changes in heart rate variability during the anticipation of a stressful situation and the stress-induced cortisol response.
+\emph{Psychoneuroendocrinology} (2018) \textbf{94}:63--71.
+doi:\href{https://doi.org/10.1016/j.psyneuen.2018.05.004}{10.1016/j.psyneuen.2018.05.004}
+
+\leavevmode\hypertarget{ref-Qian2015}{}%
+98.
+Qian SS, Miltner RJ.
+A continuous variable Bayesian networks model for water quality modeling: A case study of setting nitrogen criterion for small rivers and streams in Ohio, USA.
+\emph{Environmental Modelling and Software} (2015) \textbf{69}:14--22.
+doi:\href{https://doi.org/10.1016/j.envsoft.2015.03.001}{10.1016/j.envsoft.2015.03.001}
+
+\leavevmode\hypertarget{ref-Ross2011}{}%
+99.
+Ross SE, Linens SW, Wright CJ, Arnold BL.
+Balance assessments for predicting functional ankle instability and stable ankles.
+\emph{Gait and Posture} (2011) \textbf{34}:539--542.
+doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}
 \end{cslreferences}

### 2020-08-14 mission_statement.Rmd
pre figure removeal

---
title: "Untitled"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Despite the numerous physical, mental and social benefits of participating in sport, one of the more negative and arguably inevitable consequences, is injury. One important factor that has been linked to injury is stress. However sports science research has addressed the stress-injury relationship from a somewhat narrow perspective. The sub-disciplines of psychology, physiology and biomechanics have employed discipline specific approaches to examine factors associated with the stress-injury relationship – but rarely has this research gone beyond their disciplinary boundaries or included the multiple waves of measurement required to capture the dynamic nature of the stress-injury relationship. This study addresses those shortcomings by adopting an interdisciplinary approach and repeated measure design to investigate the relationships between psychological (major life events) and physiological measures of stress (muscle stiffness, heart rate variability and balance) to capture the multifaceted and dynamic nature of the stress- sports injury relationship. Findings revealed: the probability of injury was greatest when there was a combination of high muscle stiffness and poor balance  and increases in negative life event stress and muscle stiffness over time increased the probability of injury. These findings provide important insights that can enhance our understanding of the complex relationship between stress and injury.

### 2020-08-14 supplimentary/supplimentary_tables.tex
pre figure removeal

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \captionsetup[table]{labelformat=empty}
-% https://github.com/rstudio/rmarkdown/issues/337
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-% https://github.com/rstudio/rmarkdown/pull/252
-\usepackage{titling}
-\setlength{\droptitle}{-2em}
-
-\pretitle{\vspace{\droptitle}\centering\huge}
-\posttitle{\par}
-
-\preauthor{\centering\large\emph}
-\postauthor{\par}
-
-\predate{\centering\large\emph}
-\postdate{\par}
 \usepackage{booktabs}
@@
 \title{Supplimentary materials}
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 \endfirsthead
-\caption[]{ S3 Table continued}\\
+\caption[]{ S2 Table continued}\\
 \hline
@@
 \hline
-BIS\_1 & FFFS\_1 & 0.98 & 0.68\\
+BIS\_1 & FFFS\_1 & 0.98 & 0.70\\
 \hline
-FFFS\_1 & BIS\_1 & 0.98 & 0.32\\
+FFFS\_1 & BIS\_1 & 0.98 & 0.30\\
 \hline
-ind\_team & hours & 0.95 & 0.50\\
+ind\_team & hours & 0.94 & 0.50\\
 \hline
-hours & ind\_team & 0.95 & 0.50\\
+hours & ind\_team & 0.94 & 0.50\\
 \hline
-nlebase & RI\_1 & 0.83 & 1.00\\
+nlebase & RI\_1 & 0.84 & 1.00\\
 \hline
@@
 \hline
-BIS\_2 & FFFS\_2 & 0.73 & 0.71\\
+BIS\_2 & FFFS\_2 & 0.74 & 0.71\\
 \hline
-FFFS\_2 & BIS\_2 & 0.73 & 0.29\\
+FFFS\_2 & BIS\_2 & 0.74 & 0.29\\
 \hline
-gender & stiffness\_1 & 0.72 & 1.00\\
+gender & stiffness\_1 & 0.71 & 1.00\\
 \hline
-hours & rmssd\_1 & 0.66 & 1.00\\
+pi & stiffness\_1 & 0.57 & 1.00\\
 \hline
-gender & stiffness\_2 & 0.60 & 1.00\\
+nlelg\_1 & BIS\_1 & 0.55 & 0.97\\
 \hline
-nlelg\_1 & BIS\_1 & 0.52 & 0.97\\
+BIS\_1 & nlelg\_1 & 0.55 & 0.03\\
 \hline
-BIS\_1 & nlelg\_1 & 0.52 & 0.03\\
+FFFS\_1 & BIS\_2 & 0.52 & 1.00\\
 \hline
-pi & stiffness\_1 & 0.51 & 1.00\\
+clevel & balance\_1 & 0.47 & 1.00\\
 \hline
-injured\_2 & stiffness\_2 & 0.51 & 0.65\\
+injured\_2 & stiffness\_2 & 0.46 & 0.66\\
 \hline
-stiffness\_2 & injured\_2 & 0.51 & 0.35\\
+stiffness\_2 & injured\_2 & 0.46 & 0.34\\
 \hline
-clevel & balance\_1 & 0.49 & 1.00\\
+injured\_1 & stiffness\_1 & 0.46 & 0.36\\
 \hline
-FFFS\_1 & BIS\_2 & 0.49 & 1.00\\
+stiffness\_1 & injured\_1 & 0.46 & 0.64\\
 \hline
-nlebase & rmssd\_1 & 0.49 & 1.00\\
+gender & stiffness\_2 & 0.43 & 1.00\\
 \hline
-injured\_1 & stiffness\_1 & 0.46 & 0.34\\
+FFFS\_1 & injured\_2 & 0.42 & 1.00\\
 \hline
-stiffness\_1 & injured\_1 & 0.46 & 0.66\\
+hours & injured\_1 & 0.42 & 1.00\\
 \hline
-clevel & RI\_2 & 0.46 & 1.00\\
+clevel & RI\_2 & 0.42 & 1.00\\
 \hline
-FFFS\_1 & injured\_2 & 0.44 & 1.00\\
+nlebase & hrv\_1 & 0.41 & 1.00\\
 \hline
-hours & injured\_1 & 0.40 & 1.00\\
+hours & hrv\_1 & 0.40 & 1.00\\
 \hline
-nlelg\_2 & BIS\_2 & 0.38 & 0.98\\
-\hline
-BIS\_2 & nlelg\_2 & 0.38 & 0.02\\
+hrv\_1 & nlelg\_2 & 0.39 & 1.00\\
 \hline
@@
 \hline
-injured\_2 & rmssd\_2 & 0.37 & 0.84\\
+injured\_2 & hrv\_2 & 0.37 & 0.89\\
 \hline
-rmssd\_2 & injured\_2 & 0.37 & 0.16\\
+nlelg\_2 & BIS\_2 & 0.37 & 0.96\\
 \hline
-pi & nlelg\_2 & 0.37 & 1.00\\
+BIS\_2 & nlelg\_2 & 0.37 & 0.04\\
+\hline
+hrv\_2 & injured\_2 & 0.37 & 0.11\\
+\hline
+pi & nlelg\_2 & 0.36 & 1.00\\
+\hline
+RI\_1 & BIS\_1 & 0.36 & 0.41\\
+\hline
+BIS\_1 & RI\_1 & 0.36 & 0.59\\
 \hline
@@
 \hline
-RI\_1 & BIS\_1 & 0.34 & 0.44\\
+injured\_1 & balance\_1 & 0.36 & 0.70\\
 \hline
-BIS\_1 & RI\_1 & 0.34 & 0.56\\
+balance\_1 & injured\_1 & 0.36 & 0.30\\
 \hline
-injured\_2 & balance\_2 & 0.34 & 0.95\\
+injured\_2 & balance\_2 & 0.34 & 0.94\\
 \hline
-balance\_2 & injured\_2 & 0.34 & 0.05\\
+balance\_2 & injured\_2 & 0.34 & 0.06\\
 \hline
-injured\_1 & balance\_1 & 0.34 & 0.68\\
+RI\_1 & balance\_1 & 0.33 & 0.22\\
 \hline
-balance\_1 & injured\_1 & 0.34 & 0.32\\
+balance\_1 & RI\_1 & 0.33 & 0.78\\
 \hline
 gender & FFFS\_2 & 0.33 & 1.00\\
-\hline
-RI\_1 & balance\_1 & 0.32 & 0.21\\
-\hline
-balance\_1 & RI\_1 & 0.32 & 0.79\\
 \hline
@@
 \hline
-BIS\_1 & FFFS\_2 & 0.32 & 1.00\\
+ind\_team & FFFS\_1 & 0.31 & 1.00\\
 \hline
-pi & ind\_team & 0.31 & 0.44\\
+BIS\_1 & FFFS\_2 & 0.31 & 1.00\\
 \hline
-ind\_team & pi & 0.31 & 0.56\\
+pi & ind\_team & 0.30 & 0.46\\
 \hline
-pi & stiffness\_2 & 0.31 & 1.00\\
-\hline
-ind\_team & FFFS\_1 & 0.30 & 1.00\\
+ind\_team & pi & 0.30 & 0.54\\
 \hline
@@
 
+\clearpage
 \newpage
@@
 \hline
-BIS & FFFS & 1.00 & 0.75\\
+BIS & FFFS & 1.00 & 0.76\\
 \hline
@@
 \hline
-injured & stiffness & 0.66 & 1.00\\
+injured & stiffness & 0.67 & 1.00\\
 \hline
-rmssd & FFFS & 0.64 & 0.67\\
+clevel & BIS & 0.60 & 1.00\\
 \hline
-FFFS & rmssd & 0.64 & 0.33\\
+hrv & FFFS & 0.58 & 0.69\\
 \hline
-clevel & BIS & 0.61 & 1.00\\
+FFFS & hrv & 0.58 & 0.31\\
 \hline
-gender & stiffness & 0.57 & 1.00\\
+gender & stiffness & 0.56 & 1.00\\
 \hline
@@
 \hline
-pi & balance & 0.42 & 1.00\\
+pi & balance & 0.43 & 1.00\\
 \hline
@@
 \hline
-RI & BIS & 0.36 & 0.35\\
+stiffness & BIS & 0.38 & 0.93\\
 \hline
-BIS & RI & 0.36 & 0.65\\
+BIS & stiffness & 0.38 & 0.07\\
 \hline
-stiffness & BIS & 0.34 & 0.92\\
+RI & BIS & 0.35 & 0.29\\
 \hline
-BIS & stiffness & 0.34 & 0.08\\
+BIS & RI & 0.35 & 0.71\\
+\hline
+pi & hrv & 0.33 & 1.00\\
 \hline
@@
 \hline
-0.34 & club\_university\_county & High & High & High & High\\
+0.32 & national\_international & Low & High & High & Low\\
 \hline
-0.32 & national\_international & Low & High & High & Low\\
+0.32 & club\_university\_county & High & High & High & High\\
+\hline
+0.31 & club\_university\_county & High & High & Low & High\\
 \hline
@@
 \hline
-0.29 & club\_university\_county & High & High & Low & High\\
+0.25 & national\_international & High & High & High & Low\\
 \hline
-0.27 & national\_international & High & High & High & Low\\
-\hline
-0.26 & national\_international & High & High & High & High\\
+0.25 & national\_international & High & High & Low & Low\\
 \hline
@@
 \hline
+0.24 & national\_international & High & High & Low & High\\
+\hline
+0.24 & national\_international & High & High & High & High\\
+\hline
 0.24 & national\_international & Low & High & Low & Low\\
-\hline
-0.23 & national\_international & High & High & Low & Low\\
 \hline
@@
 \hline
-0.22 & national\_international & High & High & Low & High\\
+0.19 & club\_university\_county & High & High & Low & Low\\
 \hline
-0.21 & club\_university\_county & High & High & High & Low\\
+0.19 & club\_university\_county & High & High & High & Low\\
 \hline
 0.18 & club\_university\_county & Low & High & Low & Low\\
-\hline
-0.18 & club\_university\_county & High & High & Low & Low\\
 \hline
@@
 \hline
-prob & FFFS\_1 & nlelg\_2 & stiffness\_2 & rmssd\_2 & balance\_2\\
+prob & FFFS\_1 & nlelg\_2 & stiffness\_2 & hrv\_2 & balance\_2\\
 \hline
@@
 \hline
-0.45 & Low & High & High & Low & High\\
+0.46 & Low & High & High & Low & High\\
 \hline
@@
 \hline
-0.34 & Low & High & High & High & High\\
+0.35 & Low & High & High & High & High\\
 \hline
-0.32 & Low & High & High & Low & Low\\
+0.33 & Low & High & High & Low & Low\\
 \hline
-0.31 & Low & Low & Low & Low & High\\
+0.32 & Low & Low & Low & Low & High\\
 \hline
-0.29 & Low & Low & High & High & Low\\
+0.28 & Low & Low & High & High & Low\\
 \hline
-0.28 & Low & High & Low & Low & High\\
+0.26 & Low & High & Low & Low & High\\
 \hline
@@
 \hline
-0.22 & Low & Low & Low & High & High\\
+0.23 & Low & Low & Low & High & High\\
 \hline
-0.21 & High & High & High & Low & High\\
+0.22 & High & High & High & Low & High\\
 \hline
@@
 \hline
-0.19 & Low & High & Low & High & High\\
-\hline
-0.18 & Low & High & Low & Low & Low\\
+0.18 & Low & High & Low & High & High\\
 \hline
@@
 \hline
+0.17 & Low & High & Low & Low & Low\\
+\hline
 0.16 & High & Low & High & Low & Low\\
 \hline
-0.14 & High & High & High & High & High\\
+0.15 & High & High & High & High & High\\
 \hline
@@
 \hline
-0.12 & Low & High & Low & High & Low\\
+0.11 & Low & High & Low & High & Low\\
 \hline
-0.12 & High & High & Low & Low & High\\
+0.11 & High & High & Low & Low & High\\
 \hline
@@
 \hline
-0.08 & High & High & High & High & Low\\
+0.09 & High & High & High & High & Low\\
 \hline
@@
 \hline
-0.07 & High & High & Low & Low & Low\\
+0.06 & High & High & Low & Low & Low\\
 \hline
@@
 
+\newpage
+
+\textbf{S6 Table. Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{c|c|c|c|c}
+\hline
+\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
+\hline
+0.77 & High & injury & +1SD & +1SD\\
+\hline
+0.74 & High & no injury & +1SD & +1SD\\
+\hline
+0.72 & Low & injury & +1SD & +1SD\\
+\hline
+0.70 & High & injury & +1SD & mean\\
+\hline
+0.69 & High & injury & +1SD & -1SD\\
+\hline
+0.68 & High & no injury & +1SD & mean\\
+\hline
+0.66 & High & no injury & +1SD & -1SD\\
+\hline
+0.65 & Low & injury & +1SD & mean\\
+\hline
+0.64 & Low & injury & +1SD & -1SD\\
+\hline
+0.61 & High & injury & mean & +1SD\\
+\hline
+0.58 & High & no injury & mean & +1SD\\
+\hline
+0.54 & Low & injury & mean & +1SD\\
+\hline
+0.54 & High & injury & mean & mean\\
+\hline
+0.51 & High & injury & -1SD & +1SD\\
+\hline
+0.51 & High & injury & mean & -1SD\\
+\hline
+0.50 & High & no injury & mean & mean\\
+\hline
+0.48 & High & no injury & mean & -1SD\\
+\hline
+0.48 & Low & injury & mean & mean\\
+\hline
+0.47 & High & no injury & -1SD & +1SD\\
+\hline
+0.46 & Low & injury & mean & -1SD\\
+\hline
+0.44 & Low & injury & -1SD & +1SD\\
+\hline
+0.44 & High & injury & -1SD & mean\\
+\hline
+0.42 & High & injury & -1SD & -1SD\\
+\hline
+0.41 & High & no injury & -1SD & mean\\
+\hline
+0.38 & High & no injury & -1SD & -1SD\\
+\hline
+0.37 & Low & no injury & +1SD & +1SD\\
+\hline
+0.37 & Low & injury & -1SD & mean\\
+\hline
+0.35 & Low & injury & -1SD & -1SD\\
+\hline
+0.32 & Low & no injury & +1SD & mean\\
+\hline
+0.29 & Low & no injury & +1SD & -1SD\\
+\hline
+0.20 & Low & no injury & mean & +1SD\\
+\hline
+0.18 & Low & no injury & mean & mean\\
+\hline
+0.16 & Low & no injury & mean & -1SD\\
+\hline
+0.15 & Low & no injury & -1SD & +1SD\\
+\hline
+0.13 & Low & no injury & -1SD & mean\\
+\hline
+0.11 & Low & no injury & -1SD & -1SD\\
+\hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2020-08-14 supplimentary_tables.Rmd
pre figure removeal

@@
 ```{r include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn)
@@
 
+\clearpage
+\newpage
+
 \textbf{S3 Table. All arc strengths greater than 0.3 included in the second network.}
@@
   tallc,
-  c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "rmssd_2"),
+  c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2"),
   "injured_2",
@@
 ```
+
+\newpage
+
+\textbf{S6 Table. Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+
+```{r results='asis'}
+set.seed(1007)
+testt <- expand.grid(list(
+  c(-1:1),
+  c(-1:1),
+  c("Low", "High"),
+  c("injury", "no injury")
+))
+
+query7 <- pmap_dfr(
+  .l = testt,
+  ~ tibble(
+    prob = cpquery(
+      fittedcont,
+      event = (injured == "injured"),
+      evidence = list(
+        nlec = as.numeric(..1),
+        stiffness = as.numeric(..2),
+        hours = ..3,
+        pi = ..4
+      ),
+      method = "lw"
+    ),
+    nle = ..1,
+    stiffness = ..2,
+    hours = ..3,
+    pi = ..4
+  )
+) %>%
+  arrange(-prob)
+
+query7 %>%
+  mutate_at(vars(nle, stiffness), ~ case_when(
+    . == 1 ~ "+1SD",
+    . == 0 ~ "mean",
+    . == -1 ~ "-1SD"
+  )) %>%
+  select(prob, hours, pi, nle, stiffness) %>%
+  rename(Probability = "prob") %>%
+  dendroTools::round_df(2) %>%
+  kable(
+    .,
+    align = "ccccc"
+  ) %>%
+  kable_styling(latex_options = "HOLD_position") %>%
+
+  row_spec(0,bold=TRUE)
+```

### 2020-08-14 r-files/results_code.R

@@ -112,7 +112,11 @@ nodeRenderInfo(gR)$fill[g2] = "tomato"
 nodeRenderInfo(gR)$fill["injured_1"] = "gold"
 nodeRenderInfo(gR)$fill["injured_2"] = "gold"
 gR <- renderGraph(gR)
+
+
+jpeg("./figures_doc/Fig6.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(gR)
+dev.off()
 # tiff("main_network.tif", res = 300, width = 15, height = 10, units = "in")
 #renderGraph(gR)
 # dev.off()
@@ -224,7 +228,7 @@ nodeRenderInfo(change)$fill[indivars] = "tomato"
 
 nodeRenderInfo(change)$fill["injured"] = "gold"
 
-jpeg("./figures_doc/Fig9.jpeg", res = 300, width = 15, height = 10, units = "in")
+jpeg("./figures_doc/Fig9.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(change)
 dev.off()
 fittedcont <- bn.fit(avgcont, diffg)
@@ -318,33 +322,37 @@ edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6,
                                     "hours~injured_1" = "dashed"))
 
 q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
-graph.par(list(edges=list(fontsize = 25)))
+graph.par(list(edges=list(fontsize = 15)))
 
 nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
 nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
 
-jpeg("figures_doc/Fig7.jpeg", res = 300, width = 15, height = 10, units = "in")
+jpeg("figures_doc/Fig7.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(q1p)
 dev.off()
 
 
-plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][rmssd_2|injured_2]')
+plotsg2 <- model2network('[nlelg_2][gender][FFFS_1][injured_2|FFFS_1:nlelg_2][stiffness_2|gender:injured_2][balance_2|injured_2][hrv_2|injured_2]')
[…]

### 2020-08-14 manuscript.Rmd
for mac

@@
 authornote: |
-    12961 words in text body and refence section
-
-    Number of tables: 11
-
-    Number of figures: 10
+    12856 words in text body and refence section
+
+    Number of tables: 10
+
+    Number of figures: 5
 
@@
 Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012].
-Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
-
-```{r fig1, fig.cap="Stress and injury model (Williams and Andersen, 1998).", fig.align='center',fig.pos = "!h"}
-include_graphics("figures_doc/Fig1.jpeg")
-```
+Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [@Andersen1988; @Williams1998].
 
@@
 The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (supplementary figure 1).
 
@@
 non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (supplementary figure 2). 
 To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
@@
 A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
-
-```{r fig2, fig.cap="mBESS positions (A-F).
-Top row, firm surface.
-Bottom row, foam surface.
-Left column, dominant leg stance.
-Middle column, non-dominant leg stance.
-Right column, Tandem leg stance.", out.width = "100%", fig.align='center',fig.pos = "!h"}
-include_graphics("figures_doc/Fig2.jpeg")
-```
 
@@
 A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (supplementary figure 3).
 Participants provided informed consent before data collection commenced.
 
-```{r fig3, fig.cap = 'Study design.
-For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width = "100%"}
-
-include_graphics("figures_doc/Fig3.jpeg")
-```
-
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (supplementary figure 4). 
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
@@
 Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
-
-```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
-include_graphics("figures_doc/Fig4.jpeg")
-```
 
@@
 The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014].
-In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
-In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
+When a network contains many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
 BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
@@
 Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
-
-```{r table3, results='asis'}
+To summarise, Table \ref{tab:table2} reports the variables that were included in the final network structure.
+
+```{r table2, results='asis'}
 varcutoffs %>% 
@@
 Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
+Table \ref{tab:table3} shows the number and percentage of injury types sustained by both male and female participants.
 
@@
 
-```{r table4, results='asis'}
+```{r table3, results='asis'}
 
@@
 
-The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
+The first network structure obtained from the data (Fig \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
@@
 
-```{r fig6, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
+```{r fig1, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
 The strength of each arc is shown graphically by the style of the arc.
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
 The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
 
-```{r fig7, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig7.jpeg")
-```
-
-The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table5}.
+```{r fig2, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig2.jpeg")
+```
+
+The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table4}.
 The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
@@
 
-```{r table5, results='asis'}
+```{r table4, results='asis'}
 query1 %>%
@@
 The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
+The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in supplementary table 4).
 The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
@@
 
-```{r table6, results='asis'}
+```{r table5, results='asis'}
 
@@
     .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
+    caption = "Highest and lowest probability of injured\\_1 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\\_1.",
     escape = F,
@@
 
-The Markov blanket for injured_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2.
+The Markov blanket for injured_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2 (table \ref{tab:table6}).
 The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
@@
 
-```{r table7, results="asis"}
+```{r table6, results="asis"}
 
@@
     .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
+    caption = "Probability of injured\\_2 being in the ``injured'' state, conditional on each variable in the Markov blanket for injured\\_2.",
     col.names = c("Variable", "Low", "High"),
@@
 
-```{r fig8, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig8.jpeg")
+```{r fig3, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig3.jpeg")
 ```
@@
 
-The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
+The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in supplementary table 5).
 The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
@@
 
-```{r table8, results="asis" }
+```{r table7, results="asis" }
 
@@
     .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
+    caption = "Highest and lowest probability of injured\\_2 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\\_2.",
     escape = F,
@@
 
-The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
+The second network structure (Fig \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
 An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
@@
 
-```{r fig9, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}
-
-include_graphics("figures_doc/Fig9.jpeg")
-
-```
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95% credible intervals (CrI).
+```{r fig4, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}
+
+include_graphics("figures_doc/Fig4.jpeg")
+
+```
+
+Results from the Bayesian linear regression model are presented in Table \ref{tab:table8} and include 95% credible intervals (CrI).
 Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
@@
 
-```{r table9, results="asis"}
+```{r table8, results="asis"}
 bmtable %>%
@@
 
-The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
+The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig5}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
@@
 
-```{r fig10, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig10.jpeg")
-```
-
-```{r table10, results="asis" }
+```{r fig5, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
+include_graphics("figures_doc/Fig5.jpeg")
+```
+
+```{r table9, results="asis" }
 set.seed(1006)
@@
 
-Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
+Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
@@
 
-```{r table11, results="asis"}
+```{r table10, results="asis"}
 

### 2020-08-14 manuscript.tex
for mac

@@
 \textsuperscript{1} Cardiff Metropolitan University, Cardiff, United Kingdom\\\textsuperscript{2} Brunel University, London, United Kingdom\\\textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA), Manno, Switzerland}
-\authornote{12961 words in text body and refence section
-
-Number of tables: 11
-
-Number of figures: 10
+\authornote{12856 words in text body and refence section
+
+Number of tables: 10
+
+Number of figures: 5
 
@@
 
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=5.67in]{figures_doc/Fig1}
-
-}
-
-\caption{Stress and injury model (Williams and Andersen, 1998).}\label{fig:fig1}
-\end{figure}
-
 Williams and Andersen's (14) stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete's personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
@@
 non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion)
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}).
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (supplementary figure 2).
 To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (57).
@@
 A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
-
-\begin{figure}[!h]
-
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig2}
-
-}
-
-\caption{mBESS positions (A-F).
-Top row, firm surface.
-Bottom row, foam surface.
-Left column, dominant leg stance.
-Middle column, non-dominant leg stance.
-Right column, Tandem leg stance.}\label{fig:fig2}
-\end{figure}
 
@@
 A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (supplementary figure 3).
 Participants provided informed consent before data collection commenced.
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{figures_doc/Fig3} \caption{Study design.
-For each time point (T), each box contains the number of participants who completed the data collection (\textit{n}), the measures used for data collection and the approximate date of the data collection.}\label{fig:fig3}
-\end{figure}
-
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}).
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (supplementary figure 4).
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups.
@@
 Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig4}
-
-}
-
-\caption{Outline of the protocol for each data collection.}\label{fig:fig4}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig5}
-
-}
-
-\caption{A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.}\label{fig:fig5}
-\end{figure}
-
 In order to use discrete networks, continuous variables must first be split into categorical levels.
@@
 The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (38).
-In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to ``Low'') could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
-In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network.
+When a network contains many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network.
 BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
@@
 Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
-Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 \(\rightarrow\) T2 and T2 \(\rightarrow\) T3, respectively.
-Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
 In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
@@
 All subsequent models used the same blacklist.
-
-\begin{table}[H]
-
-\caption{\label{tab:table2}Example of the data arrangement used for the network.}
-\centering
-\begin{tabular}[t]{c|c|c}
-\hline
-\textbf{Participant} & \textbf{X\_1} & \textbf{X\_2}\\
-\hline
-1 & T1 & T2\\
-\hline
-1 & T2 & T3\\
-\hline
-2 & T1 & T2\\
-\hline
-3 & T1 & T2\\
-\hline
-3 & T2 & T3\\
-\hline
-\end{tabular}
-\end{table}
 
@@
 Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
+To summarise, Table \ref{tab:table2} reports the variables that were included in the final network structure.
 
@@
 
-\caption{\label{tab:table3}Variables included in the final Bayesian network structure.}
+\caption{\label{tab:table2}Variables included in the final Bayesian network structure.}
 \centering
@@
 Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = 85 {[}69\%{]}, chronic = 39 {[}31\%{]}; female, acute = 38 {[}72\%{]} chronic = 15 {[}28\%{]}), and non-contact injuries were more common than contact injuries (male, non-contact = 83 {[}67\%{]}, contact = 39 {[}31\%{]}; female, non-contact = 35 {[}66\%{]} contact = 18 {[}34\%{]}).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
+Table \ref{tab:table3} shows the number and percentage of injury types sustained by both male and female participants.
 
@@
 
-\caption{\label{tab:table4}The number and percentage (\%) of types of injuries sustained by male and female participants.}
+\caption{\label{tab:table3}The number and percentage (\%) of types of injuries sustained by male and female participants.}
 \centering
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/fig6-1.pdf}
-\caption{\label{fig:fig6}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
+\includegraphics{manuscript_files/figure-latex/fig1-1.pdf}
+\caption{\label{fig:fig1}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
 The strength of each arc is shown graphically by the style of the arc.
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
 The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig7}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig2}
 
@@
 
-\caption{Markov blanket of injured\_1. Arc strengths are included as arc labels.}\label{fig:fig7}
+\caption{Markov blanket of injured\_1. Arc strengths are included as arc labels.}\label{fig:fig2}
 \end{figure}
 
-The CPQ for injured\_1 in the ``injured'' state for all variables that were linked to injured\_1 is shown in Table \ref{tab:table5}.
+The CPQ for injured\_1 in the ``injured'' state for all variables that were linked to injured\_1 is shown in Table \ref{tab:table4}.
 The probability of injured\_1 = ``injured'' rose from 0.17 to 0.31 when stiffness was ``High'' compared to ``Low''.
@@
 
-\caption{\label{tab:table5}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
+\caption{\label{tab:table4}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
 \centering
@@
 The second CPQ investigated the outcome of injured\_1 being ``injured'' conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
+The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in supplementary table 4).
 The combination of lower competitive level, ``High'' hours per week, ``Low'' negative life events, ``High'' balance and ``High'' stiffness resulted in a probability of 0.53 for injured\_1 being in the ``injured'' state.
@@
 
-\caption{\label{tab:table6}Highest and lowest probability of injured\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_1.}
+\caption{\label{tab:table5}Highest and lowest probability of injured\_1 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_1.}
 \centering
@@
 
-The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2.
+The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2 (table \ref{tab:table6}).
 The arc between stiffness\_2 \(\rightarrow\) injured\_2 was comparable to the arc between stiffness\_1 \(\rightarrow\) injured\_1.
@@
 
-\caption{\label{tab:table7}Probability of injured\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\_2.}
+\caption{\label{tab:table6}Probability of injured\_2 being in the ``injured'' state, conditional on each variable in the Markov blanket for injured\_2.}
 \centering
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig8}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig3}
 
@@
 
-\caption{Markov blanket for injured\_2.}\label{fig:fig8}
+\caption{Markov blanket for injured\_2.}\label{fig:fig3}
 \end{figure}
 
-The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
+The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in supplementary table 5).
 The combination of ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.53.
@@
 
-\caption{\label{tab:table8}Highest and lowest probability of injured\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\_2.}
+\caption{\label{tab:table7}Highest and lowest probability of injured\_2 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_2.}
 \centering
@@
 
-The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
+The second network structure (Fig \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
 An arc between BIS \(\rightarrow\) FFFS with strength 1.00 was present in the network.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig9}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig4}
 
@@
 
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95\% credible intervals (CrI).
+Results from the Bayesian linear regression model are presented in Table \ref{tab:table8} and include 95\% credible intervals (CrI).
 Increases in BIS were associated with increases in FFFS (\(b\) = -0.19, 95\% CrI = {[}-0.25, -0.13{]}), whereas positive changes in hrv where associated with decreased changes in FFFS (\(b\) = 0.41, 95\% CrI = {[}0.36, 0.47{]}).
@@
 
-\caption{\label{tab:table9}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.}
+\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.}
 \centering
@@
 
-The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
+The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig5}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig10}
+{\centering \includegraphics[width=1\linewidth]{figures_doc/Fig5}
 
@@
 
-\caption{Markov blanket for the injured node in the network reflecting changes within variables between time points.}\label{fig:fig10}
+\caption{Markov blanket for the injured node in the network reflecting changes within variables between time points.}\label{fig:fig5}
 \end{figure}
@@
 
-\caption{\label{tab:table10}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
+\caption{\label{tab:table9}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
 \centering
@@
 
-Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
+Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
@@
 
-\caption{\label{tab:table11}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+\caption{\label{tab:table10}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
 \centering

### 2020-08-14 supplementary_figures.Rmd
for mac

--- 2020-07-29 apa_style.Rmd
@@
 ---
-title             : "An interdisciplinary examination of stress and injury occurrence in athletes"
-shorttitle        : "Stress and sports injury"
-
-author: 
-  - name          : "Harry Fisher"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       :  "Cyncoed Road, Cardiff, CF23 6XD"
-    email         : "harryfisher21@gmail.com"
-  - name: Dr Marianne Gittoes
-    affiliation: '1'
-  - name: Professor Lynne Evans
-    affiliation: '1'
-  - name: Miss Leah Bitchell
-    affiliation: '1'  
-  - name: Dr Richard Mullen  
-    affiliation: '2'
-  - name: Dr Marco Scutari
-    affiliation: '3'
-
-affiliation:
-  - id: '1'
-    department: School of Sport and Health Science
-    institution: Cardiff Metropolitan University
-    city: Cardiff
-    country: United Kingdom
-  - id: '2' 
-    department: Division of Sport, Health & Exercise Sciences
-    institution: Brunel University
-    city: London
-    country: United Kingdom
-  - id: '3' 
-    institution: Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA) 
-    city: Manno
-    country: Switzerland
-
-abstract: |
- This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
-Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
-Findings revealed muscle stiffness to have the strongest relationship with injury occurrence, with high levels of stiffness increasing the probability of sustaining an injury.
-Negative life events did not increase the probability of injury occurrence at any single time-point; however, when examining changes between time points, increases in negative life events did increase the probability of injury.
-In addition, the combination of increases in negative life events and muscle stiffness resulted in the greatest probability of sustaining an injury.
-Findings demonstrated the importance of both an interdisciplinary approach and a repeated measures design to furthering our understanding of the relationship between stress-related  markers and injury occurrence.
-
-keywords          : "Sports Injury, Stress, Interdisciplinary, Bayesian Network"
-
-bibliography      : [library.bib, r-references.bib]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
-csl: frontiers.csl
-always_allow_html: true
-header_includes:
-  - \usepackage{pdflscape}
-  - \newcommand{\blandscape}{\begin{landscape}}
-  - \newcommand{\elandscape}{\end{landscape}}
+title: "Supplimentary figures"
+output: pdf_document
 ---
 
-```{r setup, include = FALSE}
-library("papaja")
-options(kableExtra.auto_format = FALSE)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+pacman::p_load(knitr)
 ```
-
-```{r include=FALSE}
-options(scipen=999, digits = 2)
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
- source("r-files/functions.R")
- source("r-files/methods_code.R")
- source("r-files/results_code.R")
-#load("all_data.RData", envir = myEnv)
-#myEnv <- new.env()
-#attach(myEnv, name="sourced_scripts")
-
-pacman::p_load(tidyverse, 
-               knitr,
-               kableExtra,
-               qwraps2, 
-               brms,
-               broom.mixed, 
-               bnlearn, 
-               tidyselect, 
-               Rgraphviz, 
-               readxl, 
-               janitor,
-               irr, 
-               lavaan,
-               caret,
-               parallel,
-               zoo, 
-               sjmisc)
-
-```
-
-10422 words in text body
-3142 words in reference section
-
-# Introduction
-
-Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012].
-Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].
 
@@
 ```
-
-Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
-The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
-
-Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
-Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.
-
-While the psychosocial factors in Williams and Andersen’s [@Williams1998] model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
-To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
-However, to date, the research has largely focused on attentional deficits [@Andersen1999; @Rogers2005; @Wilkerson2012a; @Swanik2007].
-For example, Andersen and Williams [@Andersen1999] found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
-Rodgers and Landers [@Rogers2005] further supported Andersen and Williams’s [@Andersen1999] earlier findings by identifying that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
-
-Knowledge of the physiological factors (e.g., increased muscle tension and reduced motor control) contributing to the remaining variance between negative life events and athletic injury remains sparse [cf.
-@Williams1998].
-One challenge faced by researchers addressing the sports injury problem through a psychological lens is the multifactorial nature of injury, and the possible interaction with physiological factors in the stress response [@Meeuwisse2007; @Wiese-Bjornstal2009].
-For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence [@Lee2017; @Djaoui2017] and may account for the unexplained variance from the psychological predictors.
-Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model.
-The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation [@Appaneal2014].
-The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p.
-74] to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen’s [@Williams1998] model whilst including other physiological markers of training-related stress.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury.
-However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
-Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
-Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
-Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
-
-Recently, Bittencourt et al.
-[@Bittencourt2016] advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
-Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p.
-3], where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
-Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
-Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables [@Scutari2014].
-The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
-Furthermore, unlike regression or structural equation models, BN’s do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known [@Olmedilla2018].
-BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
-
-Using the frameworks provided by Williams and Andersen’s [@Williams1998] stress injury model and Appaneal’s BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
-A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
-
-# Material and Methods
-
-```{r methodssetup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
-
-## Ethics Statement
-
-Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.
-
-## Participants
-
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
-Participants self-rated competitive level ranged from recreational to international standard. 
-They were engaged in training for their respective sports for at least five hours per week.
-A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
-All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
-
-```{r table1, results='asis'}
-
-kable(
-  npars %>% 
-    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
-    slice(c(1,3,8, 4, 9:11)) %>% 
-  mutate(name = case_when(name == "Age" ~ "Age (yrs)",
-                          name == "Height" ~ "Height (cm)",
-                          name == "Weight" ~ "Body mass (kg)",
-                          TRUE ~ name)),
-  caption = "Participant characteristics.",
-  col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
-  align = "c",
-  escape = F
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Demographics M (SD)", 1, 4) %>%
-  pack_rows("Current competitive level n (%)", 5, 7) %>%
-  row_spec(0, bold = TRUE)
-```
-
-## Measures
-
-### Injury
-
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014]. 
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries [cf. @Appaneal2009].
-Injury status (did / did not sustain an injury) served as the main outcome measure.
-
-### Major life events
-
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
-The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table). 
-The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
-Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”.
-Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4  ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
-A score for total life events was also calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively.
-Petrie also provided evidence of discriminant, convergent and predictive validity.
-For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
-The composite reliability for the LESCA in this study was .84.
-
-### Reinforcement Sensitivity Theory Personality Questionnaire
-
-A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
-The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
-Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
-The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
-Further details regarding the revised RST are in S1 Appendix.
-
-### Heart rate variability
-
-A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
-In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].
-
-### Musculoskeletal properties
-
-A handheld myometer (MyotonPRO, Myoton AS, Tallinn, Estonia) was used to measure passive muscle stiffness.
-The MyotonPRO is a non-invasive, handheld device that applies a mechanical impulse of 0.40 N for 0.15 ms perpendicular to the surface of the skin.
-The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
-The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016]. 
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].
-
-### Postural stability
-
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
-In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
-non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
-To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
-Participants were asked to maintain each stance for a total of 20-seconds.
-Participants hands were placed on hips at the level of the iliac crests. 
-A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
 
@@
 
-The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials. 
-One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
-b) opening eyes; 
-c) stepping, stumbling, or falling; 
-d) moving the thigh into more than 30 degrees of flexion or abduction; 
-e) lifting the forefoot or heel; and 
-f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
-A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
-A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
-A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces). 
-To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
-The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.
+```{r fig3, fig.cap = 'Study design.
+For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width = "100%"}
 
-## Procedure
-
-At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study.
-With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
-Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}).
-Participants provided informed consent before data collection commenced.
-
-```{r fig3, fig.cap = 'Study design.
-For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width="100%"}
-
-include_graphics("figures_doc/studydesign_new.png")
+include_graphics("figures_doc/Fig3.jpeg")
 ```
 
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
-To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
-The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
-Participants were randomly assigned to one of the two groups and remained in those groups across all time points.
-
 ```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
-include_graphics("figures_doc/sessionprotocol.png")
+include_graphics("figures_doc/Fig4.jpeg")
 ```
-
-### Questionnaires
-
-The questionnaires, which included demographic information, the LESCA, RST-PQ (T1, T2, T3) and injury status (T2, T3, T4) were completed  on-line (SurveyMonkey Inc., USA, www.surveymonkey.com).
-The instructions for the LESCA were modified at T2 and T3 so that participants reported major life events that had occurred since the previous testing session.
-For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis.
-
-### HRV
-
-To minimise potential distractions, participants were directed to a designated quiet area in the laboratory where IBI data were recorded.
-Participants were instructed to turn off their mobile devices to avoid any interference with the Bluetooth sensor.
-Each chest strap was dampened with water and adjusted so it fitted tightly but comfortably, as outlined by Polar's guidelines. 
-Participants were seated and asked to remain as still as possible for the duration of the recording. 
-No attempt was made to control participants' respiratory frequency or tidal volume [@Denver2007]. 
-Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
-
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices.
-Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996].
-RMSSD was calculated as:
-
-\begin{equation} 
-  \overline{RR} = \frac{1}{N} \sum_{i=1}^{n} RR_i
-\end{equation}
-
-Where N is the length of the time series, and $RR_i$ the RR interval between beats $i$ and $i-1$, where each beat position corresponds to the beat detection instant.
-
-### Muscle stiffness
-
-To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb. 
-The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site [@Chuang2012].
-The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle. 
-The participant was then asked to relax the muscle and the muscle was palpated to locate the muscle belly.
-A skin safe pen (Viscot all skin marker pen, Viscot Medical LLC, NJ) was used to mark the testing site in the centre of the muscle belly.
-
-After the eight testing sites had been identified, the testing end of the MyotonPRO (diameter = 3 mm) was positioned perpendicular to the skin on the testing site. 
-A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one-second interval [@Morgan2018]. 
-For each impulse, the device computed passive stiffness values, with the median of the five values being saved by the device for further analysis.
-In accordance with Myoton.com, a set of five measurements with a coefficient of variation (CV) of less than 3% was accepted.
-Sets of measurements above 3% were measured again to ensure data reliability.
-Measurements were uploaded using MyotonPRO software and imported in R [@RCoreTeam2019] for further analysis.
-For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis
-
-### Postural stability
-
-Instructions for the mBESS were read to each participant and a demonstration of the positions was provided by the research assistant.
-For each position, participants were instructed to close their eyes, rest their hands on their iliac crests and remain as still as possible for 20-seconds. 
-Participants were instructed to return to the testing position as quickly as possible if they lost their balance.
-The video recording was started prior to the first stance position and stopped after all stances had been completed.
-Each completed mBESS protocol took approximately four-minutes.
-Only one trial was performed to avoid familiarisation effects across the repeated measurement [cf. @Valovich2003]. 
-The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
-
-## Data Analysis
-
-Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
-A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988].
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
-The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012].
-In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
-
-```{r fig5, fig.cap="A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.", fig.pos="H", out.width="100%", fig.align='center'}
-
-include_graphics("figures_doc/examplenetwork.png")
-
-```
-
-In order to use discrete networks, continuous variables must first be split into categorical levels.
-When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
-For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
-Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
-Typically, no more than three levels have been used in Bayesian networks in the sports injury literature [@Olmedilla2018; @Ruiz-Perez2019]
-
-Learning the structure of the network is an important step in BN modelling.
-The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
-The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
-The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014].
-In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
-In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
-BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.
-
-[^1]: The term "query" in relation to Bayesian Networks stems from Pearl's expert systems theory (1988).
-A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
-Widely used texts on Bayesian Network analysis [@Koller2009] have widely adopted the terminology in favour of that used in traditional statistics.
-
-BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017; @Ruiz-Perez2019] and offer several benefits over traditional statistical analysis. 
-For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
-Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
-In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
-Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
-The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
-
-### First network
-
-#### Data Preparation
-
-Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point.
-To prepare the data for the BN, missing values in the dataset were first imputed. 
-Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings.
-The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
-A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
-In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
-A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
-The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].
-
-A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
-All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split. 
-For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
-Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
-
-#### Network structure
-
-To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
-Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist).
-Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
-To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
-In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
-However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
-Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
-Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
-Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
-Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively.
-Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
-In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
-Finally, arc direction was restricted between specific nodes within the explanatory variables. 
-Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
-All subsequent models used the same blacklist.
-
-```{r table2, results='asis'}
-exampledata <- tibble(
-  Participant = c("1", "1", "2", "3", "3"),
-  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
-  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
-)
-
-kable(
-  exampledata,
-  caption = "Example of the data arrangement used for the network.",
-  escape = FALSE
-) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-```
-
-#### Preliminary network structures
-
-Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
-Networks were learned using a Tabu search algorithm [@Norvig2009] and BIC was used to compare different models.
-A higher BIC value indicates the structure of a DAG is a better fit to the observed data [@Scutari2014].
-BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
-
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`).
-Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
-TLE score was not included in the final structure.
-
-The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
-Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
-RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
-Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
-Therefore, only RI was included in the final structure.
-Finally, both total score and asymmetry for balance were included in the initial network.
-However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
-Therefore, balance asymmetry was removed from the final network structure.
-To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.
-
-```{r table3, results='asis'}
-varcutoffs %>% 
-  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
-) %>% 
-  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
-  kable("latex",
-       align = c("llcc"),
-       caption = "Variables included in the final Bayesian network structure.",
-       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
-  kable_styling(latex_options = "HOLD_position") %>% 
-  column_spec(2, width = c("5.2cm")) 
-```
-
-Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
-For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%.
-Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network.
-Therefore, these arcs were blacklisted from the network.
-To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
-The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018].
-Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction.
-An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
-In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
-Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.
-
-#### Network Inference
-
-Conditional probability queries (CPQ) were used to perform inference on both network structures.
-To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = “injured”) to be calculated based on the new evidence.
-CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
-The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
-Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
-Of particular interest were the variables that were connected to “injured” nodes, which were examined in the network using the Markov blanket of “injured_1” and “injured_2”.
-A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ’s were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the “injured” state.
-
-### Second network
-
-#### Data preparation
-
-Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.
-
-For the second network, change scores for continuous variables between  T1 $\rightarrow$  T2 and  T2 $\rightarrow$  T3 were standardized to allow relative changes between variables to be compared.
-The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from "Low" to "High".
-
-#### Network structure
-
-Similar to the first network, blacklists were used to prevent arcs from independent variables $\rightarrow$ explanatory variables.
-In addition, "nlebase" variable was dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network. 
-The final network was obtained using the same approach as the first network.
-
-#### Network Inference
-
-Conditional probability queries (CPQ) were again used to perform inference the network structures.
-The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
-To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then examined using a Bayesian linear regression models with the _brms_ package [@Burkner2017a] to determine the relationship between these nodes. 
-Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
-
-# Results
-
-```{r include=FALSE}
-datanet %>% 
-  group_by(id) %>% 
-  mutate(count = sum(ifelse(injured == "injured", 1, 0))) %>% 
-  select(id, count) %>% 
-  mutate(count = ifelse(count == 0, 0, 1)) %>%
-  distinct(id, .keep_all = TRUE) %>%
-  ungroup() %>% 
-  janitor::tabyl(count)
-```
-
-During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
-Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
-Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.
-
-```{r mainnetworkresults}
-set.seed(999)
-strengthfinder <- function(var1, var2){
-  str %>%
-    filter(from == var1 & to == var2)
-}
-
-from <- c("nlebase", "ind_team", "ind_team", "gender", "gender", "clevel", "pi")
-to <- c("RI_1", "hours", "nlebase", "stiffness_1", "stiffness_2", "balance_1", "stiffness_1")
-x <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-
-from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
-to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")
-
-y <- pmap_dfr(.l = list(from = rep(from, 2),
-     to = rep(to, 2),
-     state = rep(c("High", "Low"), each = length(from))),
-         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
-  gather(var, state, -prob, -outcome) %>%
-  drop_na(state) %>%
-  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
-
-```
-
-```{r table4, results='asis'}
-
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
-  kable(
-    .,
-    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
-    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
-    escape = F
-  ) %>%
-  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0, bold = TRUE)
-
-str_finder <- function(from, to) {
-  format(round(str[which(str$from == from & str$to == to), 3], digits = 2), nsmall = 2)
-}
-
-str_cfinder <- function(from, to) {
-  format(
-    round(
-      strchange[which(strchange$from == from &
-        strchange$to == to), 3],
-      digits = 2
-    ),
-    nsmall = 2
-  )
-}
-```
-
-## First network structure
-
-The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
-Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
-The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
-Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
-Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
-Irrespective of sport or gender, strong connections were found between explanatory variables.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.
-
-The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability. 
-The arc from pi $\rightarrow$ stiffness_1 was `r str_finder("pi", "stiffness_1")` with athletes who reported an injury in the preceding 12 months being more likely to have “High” (`r x[26,1]`) compared to “Low” (`r x[28,1]`).
-stiffness.
-Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (`r str_finder("BIS_1", "FFFS_1")`) and BIS_2 $\rightarrow$ FFFS_2 (`r str_finder("BIS_2", "FFFS_2")`) were identified.
-In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of `r str_finder("nlelg_1", "BIS_1")` for nlelg_1  $\rightarrow$ BIS_1, and `r str_finder("nlelg_2", "BIS_2")` nlelg_2 and BIS_2.
-"Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]` for nlelg_1  $\rightarrow$ BIS_1, and `r y[17,1]` to `r y[18,1]` for nlelg_2  $\rightarrow$ BIS_2.
-
-\newpage
-
-```{r fig6, fig.align='default', fig.height=8, fig.width=12, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
-The strength of each arc is shown graphically by the style of the arc.
-Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables.
-The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
-renderGraph(gR)
-```
-
-### Markov blanket for injured_1
-
-```{r mb1}
-
-set.seed(1001)
-var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")
-
-query1 <- map_dfr(var, ~ newprobtable(
-  tallc,
-  .x,
-  "injured_1",
-  "injured",
-  avg30
-)) %>%
-  gather(variable, state, -prob, -outcome) %>%
-  drop_na() %>%
-  select(variable, state, prob) %>%
-  spread(state, prob) %>%
-  select(variable, Low, High) %>%
-  dendroTools::round_df(2)
-
-query2 <- newprobtable(
-  tallc,
-  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
-  "injured_1",
-  "injured",
-  avg30
-) %>%
-  toptail(n = 3) %>%
-  arrange(-prob) %>%
-  select(outcome, everything()) %>%
-  dendroTools::round_df(2)
-```
-
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
-The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
-
-```{r fig7, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig7.jpeg")
-```
-
-The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table5}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
-Negative life events had a negligible effect when moving from the "Low" to "High" state.
-
-```{r table5, results='asis'}
-query1 %>%
-  select(c(1, 2, 3)) %>%
-  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
-The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
-When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
-
-Negative life event stress had a negligible effect on the probability of injury, only influencing  injured_1 when all other variable were fixed to "Low".
-In this instance, the probability of injured_1 being "injured" rose marginally from 0.04 to 0.19, when negative life events was in the "Low" and "High" states, respectively.
-
-```{r table6, results='asis'}
-
-query2 %>%
-  select(7, 2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
-    escape = F,
-    align = "ccccc"
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-### Markov blanket of injured_2
-
-```{r}
-set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2")
-query3 <- map_dfr(var2, ~newprobtable(tallc,
-                                      .x,
-                                      "injured_2",
-                                      "injured",
-                                      avg30)) %>% 
-  gather(variable, state, -prob, - outcome) %>% 
-  drop_na() %>% 
-  select(variable, state, prob) %>% 
-  spread(state, prob) %>% 
-  select(variable, Low, High) %>% 
-  dendroTools::round_df(2)
-
-query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2"),
-                       "injured_2",
-                       "injured",
-                       avg30
-) %>% 
-  arrange(-prob) %>% 
-  select(outcome, everything()) %>% 
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for injured_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2.
-The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ hrv_2 were also present in the Markov blanket for injured_2.
-Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
-FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
-"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
-
-```{r table7, results="asis"}
-
-query3 %>%
-  mutate(Low = format(Low, nsmall = 2)) %>% 
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
-    col.names = c("Variable", "Low", "High"),
-    escape = F,
-    align = "cccccc"
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-```{r fig8, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig8.jpeg")
-```
-
-```{r}
-fffs_query <- query4 %>% 
-  filter(stiffness_2 == "High" & nlelg_2 == "High" & balance_2 == "High" & hrv_2 == "High")
-```
-
-The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
-The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
-
-```{r table8, results="asis" }
-
-query4 %>% 
-  toptail(3) %>%
-  select(7,2:6) %>%
-  rename(Probability = "prob") %>%
-  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
-    escape = F,
-    align = "cccccc"
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
-```
-
-## Second network structure - changes within variables
-
-```{r include=FALSE, warning=FALSE}
-set.seed(1003)
-bmtable <- fixef(bm1) %>% 
-  as_tibble(rownames = "term") %>% 
-  `colnames<-`(c("Term", "Estimate", "Error", "conf.low", "conf.high")) %>% 
-  dendroTools::round_df(2) %>%
-  mutate_at(vars(conf.low, conf.high), list(~format(round(., 2), nsmall = 2))) %>%
-  mutate(conf.low = str_c("[", conf.low, sep = ""),
-         conf.high = str_c(conf.high, "]", sep = "")) %>%
-  unite(., "95% CI", c("conf.low", "conf.high"), sep = ", ") %>%
-  mutate("95% CI" = str_replace_all(`95% CI`, " ", ""),
-         "95% CI" = str_replace_all(`95% CI`, ",", ", "))
-
-```
-
-The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
-An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively
-Similarly, the arc between hrv $\rightarrow$ FFFS was `r str_cfinder("hrv","FFFS")`.
-The arcs between BIS $\rightarrow$ FFFS and hrv $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
-A separate linear regression examined the interaction between BIS and hrv to be examined.
-
-```{r fig9, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}
-
-include_graphics("figures_doc/Fig9.jpeg")
-
-```
-
-Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between hrv and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
-
-```{r table9, results="asis"}
-bmtable %>%
-  kable(
-    .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.",
-    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
-    escape = F,
-    align = "lccc"
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r}
-set.seed(1005)
-test <- expand.grid(list(c(-1:1),
-                         c(-1:1)))
-query6 <- map2_dfr(.x = as.numeric(test$Var1),
-                   .y = as.numeric(test$Var2), ~tibble(prob= cpquery(fittedcont, 
-                                                                     event = (injured == "injured"),
-                                                                     evidence = list(nlec = .x,  stiffness = .y), method = "lw"),
-                                                       nle = .x,
-                                                       stiffness = .y))
-
-query6a <- query6 %>%
-  dendroTools::round_df(2)
-```
-
-The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
-Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
-Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
-
-```{r fig10, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
-include_graphics("figures_doc/Fig10.jpeg")
-```
-
-```{r table10, results="asis" }
-set.seed(1006)
-test <- expand.grid(list(
-  c(-1:1),
-  c(-1:1)
-))
-# query6 <- map2_dfr(
-#   .x = as.numeric(test$Var1),
-#   .y = as.numeric(test$Var2),
-#   ~ tibble(
-#     prob = cpquery(
-#       fittedcont,
-#       event = (injured == "injured"),
-#       evidence = list(nlec = .x, stiffness = .y),
-#       method = "lw"
-#     ),
-#     nle = .x,
-#     stiffness = .y
-#   )
-# )
-
-query6 %>%
-  select(nle, stiffness, prob) %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  arrange(-prob) %>%
-  mutate(prob = round(prob, digits = 2)) %>%
-  select(c(3, 1, 2)) %>%
-  kable(
-    .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.",
-    align = "ccc"
-  ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-
-```
-
-```{r}
-set.seed(1007)
-testt <- expand.grid(list(
-  c(-1:1),
-  c(-1:1),
-  c("Low", "High"),
-  c("injury", "no injury")
-))
-
-query7 <- pmap_dfr(
-  .l = testt,
-  ~ tibble(
-    prob = cpquery(
-      fittedcont,
-      event = (injured == "injured"),
-      evidence = list(
-        nlec = as.numeric(..1),
-        stiffness = as.numeric(..2),
-        hours = ..3,
-        pi = ..4
-      ),
-      method = "lw"
-    ),
-    nle = ..1,
-    stiffness = ..2,
-    hours = ..3,
-    pi = ..4
-  )
-) %>%
-  toptail(3) %>%
-  arrange(-prob)
-```
-
-Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).
-
-```{r table11, results="asis"}
-
-query7 %>%
-  mutate_at(vars(nle, stiffness), ~ case_when(
-    . == 1 ~ "+1SD",
-    . == 0 ~ "mean",
-    . == -1 ~ "-1SD"
-  )) %>%
-  select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
-  kable(
-    .,
-    caption = "Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.",
-    align = "ccccc"
-  ) %>%
-  pack_rows("Highest", 1, 3) %>%
-  pack_rows("Lowest", 4, 6) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
-```
-
-# Discussion
-
-## Summary fo results
-
-The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
-Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
-The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
-
-The first network revealed several links between the injured nodes and other variables in the network.
-A combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for “injured” nodes.
-The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete’s risk of injury.
-In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury [@Appaneal2014].
-
-Of all the variables measured in the study, muscle stiffness appeared to be most strongly related to injury.
-Both "High" levels of stiffness in the first network, and greater than average increases in stiffness in the second network were found to increase the risk of injury.
-However, high stiffness may only increase the risk of injury if other factors are also present.
-To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
-In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance [@Romero-Franco2014; @Trojian2006] and muscle stiffness [@Butler2003; @PickeringRodriguez2017] are separately related to injury.
-The BN structures examined in this study provided insight into these two stress-related factors and their relation to injury occurrence.
-
-The findings also facilitated an understanding of the interaction between balance and injury.
-At both injured nodes in the first network, balance was linked to injury.
-Despite the weak arc strength at both injured nodes, a “High” balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014;], was found to increase the probability of injury.
-This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
-Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts [@Paillard2006].
-Athletes who competed at a higher level were also more likely to have good balance ("Low" balance), compared to their lower level counterparts.
-These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition [@Hrysomallis2011].
-
-Negative life events captured at a single time point have previously been reported to be most strongly associated with injury [@Williams2007; @Ivarsson2017].
-The repeated measures approach combined with the second network analysis employed in this study demonstrated that greater than average increases in negative life event stress between time points increased the probability of being injured during the study period.
-However, negative life event stress had almost no effect on the probability of injury in the first network, which indicated that the relative change in life events may be more important than the absolute score for life events, despite the latter being commonly used in sports injury research to date.
-For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
-Although the absolute score would be “Low”, the relative increase between time points may have been caused by a significant event in the athlete’s life that had a considerable psychological and physiological effect [@Appaneal2014].
-Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete’s life events to be assessed [cf. @Ivarsson2014].
-
-During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
-The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model [@Ivarsson2017].
-During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-Given the reported reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
-When negative life events were “High” the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[which(query3$variable == "nlelg_2"), 3] -  query3[which(query3$variable == "nlelg_2"), 2]` at the injured_2 node.
-One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design. 
-In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months [@Petrie1992].
-In this study, athletes completed the LESCA at three time points with an approximate four-month interval after baseline.
-Athletes were asked to report any events which had occurred since the previous data collection session in order to avoid inflated scores caused by reporting the same event on multiple occasions.
-While modifications were made to the LESCA to tailor the items to the study cohort, the use of a shorter four-month time interval between data collections may have reduced the likelihood for life events listed in the LESCA to have taken place.
-For example, at the second and third time points, 26% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events  [cf.
-@Fawkner1999].
-
-Williams and Andersen's [@Williams1998] model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
-Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury [@Petrie1993; @Petrie2014; @Johnson2014].
-The lack of association between major/negative life events and injury reported here may be attributed to athletes in this study having the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
-Therefore, future research should consider including a measure of coping alongside that of life event stress to help explain the possible moderating effect.
-
-Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, "Low" HRV increased the probability of injury from `r query3[which(query3$variable == "hrv_2"), 3]` ("High" HRV) to `r query3[which(query3$variable == "hrv_2"), 2]` ("Low" HRV).
-This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
-An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
-Interestingly, the “Low” FFFS score was also related to injuries at subsequent time points.
-One possible explanation for this finding could be that those athletes who reported “Low” FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
-The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
-The first network reported an association between “High” FFFS and “High” BIS, while the second network found an association between increases in FFFS and increases in BIS.
-High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018].
-This association along with the proposed actions of the RST theory [@Corr2016b] provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network. 
-To elaborate, high levels of BIS are proposed to be the result of goal conflict, an example of which would be simultaneous triggering of the FFFS (avoidance) and BAS (approach) systems.
-The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict [@Corr2016b].
-In the present study, however, the role of BAS was limited, as evidenced by the initial network structures in which the BAS had limited connectivity with other components of the network. 
-Consequently, a more detailed examination of the role of RST in the injury process is warranted.
-
-This study had a number of strengths.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury [@Johnson2014].
-The longitudinal repeated measures design of the current study allowed changes over time and between time points to be captured and explored. 
-Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
-The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
-The BN networks allowed several markers of stress that were free to interact with each other, as well as injury, to be explored.
-
-While the BN’s provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
-Firstly, the choice was made to binarise variables in the first network so only “Low” and “High” states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
-For example, the meaning of a “Low” and “High” value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
-Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
-However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes’ responses.
-For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis [@Ross2011].
-However, the video-capture approach employed in the study ensured an accessible, non-invasive and readily applied method of capturing the respective measure.
-
-The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychological and physiological perspective.
-The variables used in the present study were not definitive.
-Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], could help to further elucidate the relationship between stress and injury.
-Future developments in the capture of life event stress using the LESCA are also warranted.
-Although the LESCA is the most widely used measures of major life events in sports injury research, modifications, including adjustment to the scoring of items are potentially justified to facilitate extended insight into the reported responses.
-For example, the LESCA may negate vastly different psychological and physiological effects between moderately negative and extremely negative events since there is no way to differentiate between an athlete who has answered four items as moderately negative, and one item as extremely negative.
-Therefore, future research could develop a modified version of the LESCA that could distinguish between these types of responses and their effects.
-
-To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Baysian Network analytical techniques,  Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
-Future research combining a repeated measures approach and complex analyses of the interactions between multifaceted stress-related measures are advocated to enhance understanding of the injury occurrence in sport.
-
-# Disclosure/Conflict-of-Interest Statement
-
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Author Contributions
-
-HF, MG, LE and RM developed the concept for the paper.
-HF was the lead author for the paper and designed, collected and analysed that data.
-MG, LE and RM were co-authors on the paper.
-MS provided assistance with data analysis.
-LB provided assistance with data collection.
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup

### 2020-08-14 r-files/results_code.R

@@ -114,7 +114,7 @@ nodeRenderInfo(gR)$fill["injured_2"] = "gold"
 gR <- renderGraph(gR)
 
 
-jpeg("./figures_doc/Fig6.jpeg", res = 300, width = 20, height = 15, units = "cm")
+jpeg("./figures_doc/Fig1.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(gR)
 dev.off()
 # tiff("main_network.tif", res = 300, width = 15, height = 10, units = "in")
@@ -228,7 +228,7 @@ nodeRenderInfo(change)$fill[indivars] = "tomato"
 
 nodeRenderInfo(change)$fill["injured"] = "gold"
 
-jpeg("./figures_doc/Fig9.jpeg", res = 300, width = 20, height = 15, units = "cm")
+jpeg("./figures_doc/Fig4.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(change)
 dev.off()
 fittedcont <- bn.fit(avgcont, diffg)
@@ -327,7 +327,7 @@ graph.par(list(edges=list(fontsize = 15)))
 nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
 nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
 
-jpeg("figures_doc/Fig7.jpeg", res = 300, width = 20, height = 15, units = "cm")
+jpeg("figures_doc/Fig2.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(q1p)
 dev.off()
 
@@ -367,7 +367,7 @@ q2p <- layoutGraph(q2p, edgeAttrs=list(label = labels))
 graph.par(list(edges=list(fontsize = 15)))
 nodeRenderInfo(q2p)$fill[c("nlelg_2", "stiffness_2", "balance_2", "FFFS_1", "hrv_2")] = "tomato"
 nodeRenderInfo(q2p)$fill["injured_2"] = "gold"
-jpeg("figures_doc/Fig8.jpeg", res = 300, width = 20, height = 15, units = "cm")
+jpeg("figures_doc/Fig3.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(q2p)
 dev.off()
 
@@ -397,7 +397,7 @@ q3p <- layoutGraph(q3p, edgeAttrs=list(label = labels))
[…]

### 2020-08-17 manuscript.Rmd
For submission

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
- # source("r-files/functions.R")
- # source("r-files/methods_code.R")
- # source("r-files/results_code.R")
-
-load("all_data.RData")
+ source("r-files/functions.R")
+ source("r-files/methods_code.R")
+ source("r-files/results_code.R")
+
+#load("all_data.RData")
 #myEnv <- new.env()
@@
 The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (supplementary figure 1).
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary figure 1).
 
@@
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
-Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (supplementary table 1). 
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (Supplementary Table 1). 
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
@@
 The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`.
-Further details regarding the revised RST are in supplementary Appendix 1.
+Further details regarding the revised RST are in Supplementary Appendix 1.
 
@@
 non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (supplementary figure 2). 
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary figure 2). 
 To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
@@
 A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (supplementary figure 3).
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Supplementary figure 3).
 Participants provided informed consent before data collection commenced.
 
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (supplementary figure 4). 
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary figure 4). 
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
@@
 In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
-Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in supplementary table 1 and supplementary table 3 respectively.
+Arcs below 0.3 were not included in the model. The full Table of arc strengths for the first and second network are available in Supplementary Table 1 and Supplementary Table 3 respectively.
 
@@
 
-The first network structure obtained from the data (Fig \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
+The first network structure obtained from the data (Figure \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
 The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
@@
 The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in supplementary table 4).
+The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in Supplementary Table 4).
 The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
@@
 
-The Markov blanket for injured_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2 (table \ref{tab:table6}).
+The Markov blanket for injured_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2 (Table \ref{tab:table6}).
 The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
@@
 
-The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in supplementary table 5).
+The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in Supplementary Table 5).
 The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
@@
 
-The second network structure (Fig \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
+The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
 An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
@@
 
-The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig5}).
+The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Figure \ref{fig:fig5}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table9}.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
@@
 
-Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
+Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
@@
 \endgroup
+
+This paper adopts a novel, interdisciplinary approach to explore the relationship between 
+psychosocial factors, physiological stress-related markers and occurrence of injury in athletes 
+using a repeated measures design across a 2-year data collection period. At three data 
+collection time-points, athletes completed measures of major life events, the reinforcement 
+sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and 
+postural stability, and reported any injuries they had sustained since the last data collection. 
+Two Bayesian networks were used to examine the relationships between variables and model 
+the changes between data collection points in the study. Findings revealed muscle stiffness to 
+have the strongest relationship with injury occurrence, with high levels of stiffness increasing 
+the probability of sustaining an injury. Negative life events did not increase the probability 
+of injury occurrence at any single time-point; however, when examining changes between 
+time points, increases in negative life events did increase the probability of injury. In 
+addition, the combination of increases in negative life events and muscle stiffness resulted in 
+the greatest probability of sustaining an injury. Findings demonstrated the importance of 
+both an interdisciplinary approach and a repeated measures design to furthering our
+understanding of the relationship between stress-related markers and injury occurrence.

### 2020-08-17 manuscript.tex
For submission

@@
 The stress response is central to the model and reflects the bi-directional relationship between athletes' appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary figure 1).
 
@@
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature.
-Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (supplementary table 1).
+Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (Supplementary Table 1).
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
@@
 The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in supplementary Appendix 1.
+The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in Supplementary Appendix 1.
 
@@
 non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion)
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (supplementary figure 2).
+and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary figure 2).
 To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (57).
@@
 A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (supplementary figure 3).
+Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Supplementary figure 3).
 Participants provided informed consent before data collection commenced.
 
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (supplementary figure 4).
+For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary figure 4).
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups.
@@
 In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (82).
-Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in supplementary table 1 and supplementary table 3 respectively.
+Arcs below 0.3 were not included in the model. The full Table of arc strengths for the first and second network are available in Supplementary Table 1 and Supplementary Table 3 respectively.
 
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
 The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
@@
 The second CPQ investigated the outcome of injured\_1 being ``injured'' conditional on all variables in the Markov blanket.
-The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in supplementary table 4).
+The Markov blanket contained five nodes, each with two possible states resulting in \(2^5\) combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in Supplementary Table 4).
 The combination of lower competitive level, ``High'' hours per week, ``Low'' negative life events, ``High'' balance and ``High'' stiffness resulted in a probability of 0.53 for injured\_1 being in the ``injured'' state.
@@
 
-The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2 (table \ref{tab:table6}).
+The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2 (Table \ref{tab:table6}).
 The arc between stiffness\_2 \(\rightarrow\) injured\_2 was comparable to the arc between stiffness\_1 \(\rightarrow\) injured\_1.
@@
 
-The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in supplementary table 5).
+The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in Supplementary Table 5).
 The combination of ``Low'' FFFS\_1, ``High'' stiffness\_2, ``High'' balance resulted in the greatest probability of injured\_2 being ``injured'', with the highest probability of injury being 0.53.
@@
 
-The second network structure (Fig \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
+The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
 An arc between BIS \(\rightarrow\) FFFS with strength 1.00 was present in the network.
@@
 
-\caption{Network structure of the changes within variables between time points.}\label{fig:fig9}
+\caption{Network structure of the changes within variables between time points.}\label{fig:fig4}
 \end{figure}
@@
 
-The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig5}).
+The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Figure \ref{fig:fig5}).
 For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table9}.
+Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress.
@@
 
-Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
+Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
 The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).

### 2020-08-17 supplimentary_tables.Rmd
For submission

@@
 ---
-title: Supplimentary materials
+title: 
 output: 

### 2020-08-17 r-files/methods_code.R

@@ -142,7 +142,7 @@ icc <- irr::icc(d, model = "twoway", type = "agreement")
 
 
 
-data <- read_csv("data/fulldatajoined.csv") %>%
+data <- read_csv("data/full_data_raw.csv") %>%
   filter(time != 4) %>% 
   rename(hrv= rmssd)# remove last time point as only questionnaire data is available
 

### 2020-08-19 manuscript.Rmd
For submission

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
- source("r-files/functions.R")
- source("r-files/methods_code.R")
- source("r-files/results_code.R")
-
-#load("all_data.RData")
+ # source("r-files/functions.R")
+ # source("r-files/methods_code.R")
+ # source("r-files/results_code.R")
+# 
+load("all_data.RData")
+
 #myEnv <- new.env()
@@
 
-# Author Contributions
+# Acknowledgements
 
@@
 
+The content of this manuscript has been published as part of the thesis of Harry Fisher, @Fisher2019.
+
 # References
@@
 \endgroup
-
-This paper adopts a novel, interdisciplinary approach to explore the relationship between 
-psychosocial factors, physiological stress-related markers and occurrence of injury in athletes 
-using a repeated measures design across a 2-year data collection period. At three data 
-collection time-points, athletes completed measures of major life events, the reinforcement 
-sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and 
-postural stability, and reported any injuries they had sustained since the last data collection. 
-Two Bayesian networks were used to examine the relationships between variables and model 
-the changes between data collection points in the study. Findings revealed muscle stiffness to 
-have the strongest relationship with injury occurrence, with high levels of stiffness increasing 
-the probability of sustaining an injury. Negative life events did not increase the probability 
-of injury occurrence at any single time-point; however, when examining changes between 
-time points, increases in negative life events did increase the probability of injury. In 
-addition, the combination of increases in negative life events and muscle stiffness resulted in 
-the greatest probability of sustaining an injury. Findings demonstrated the importance of 
-both an interdisciplinary approach and a repeated measures design to furthering our
-understanding of the relationship between stress-related markers and injury occurrence.

### 2020-08-19 manuscript.tex
For submission

@@
 
-\hypertarget{author-contributions}{%
-\section{Author Contributions}\label{author-contributions}}
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
@@
 LB provided assistance with data collection.
+
+The content of this manuscript has been published as part of the thesis of Harry Fisher, (100).
 
@@
 78.
-Kuhn MC from JW, Weston S, Williams A, Keefer C, Engelhardt A.
-caret: Classification and Regression Training.
-\emph{Journal of Statistical Software} (2008) \textbf{28}: doi:\href{https://doi.org/10.1053/j.sodo.2009.03.002}{10.1053/j.sodo.2009.03.002}
+Kuhn M.
+Building Predictive Models in R Using the caret Package.
+\emph{Journal of Statistical Software, Articles} (2008) \textbf{28}:1--26.
+doi:\href{https://doi.org/10.18637/jss.v028.i05}{10.18637/jss.v028.i05}
 
@@
 doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}
+
+\leavevmode\hypertarget{ref-Fisher2019}{}%
+100.
+Fisher H.
+An interdisciplinary examination of stress and injury occurence in athletes.
+(2019) Cardiff Metroploitan Univeristy, Cardiff.
+Available at: \url{https://repository.cardiffmet.ac.uk/handle/10369/11021}
 \end{cslreferences}

### 2020-10-15 manuscript.Rmd
Update following reviewers comments

@@
 shorttitle        : "Stress and sports injury"
-
 author: 
@@
     affiliation: '3'  
-
 affiliation:
+
   - id: '1'
@@
     country: Switzerland
-
 authornote: |
-    12856 words in text body and refence section
+    13867 words in text body and refence section
 
@@
     Format: British English
-
 abstract: |
- This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
+ This paper adopts a novel, interdisciplinary approach to explore the relationship between stress-related psychosocial factors, physiological markers and occurrence of injury in athletes using a repeated measures prospective design.
+At four data collection time-points, across 1-year of a total 2-year data collection period, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 keywords          : "Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology"
-
 bibliography      : [library.bib, r-references.bib]
-
 floatsintext      : no
@@
 draft             : no
-
 documentclass     : "apa6"
@@
   - \newcommand{\elandscape}{\end{landscape}}
+editor_options: 
+  markdown: 
+    wrap: sentence
 ---
@@
 # source("r-files/results_code.R")
-# 
-load("all_data.RData")
-
+
+load("all_data2.RData")
 #myEnv <- new.env()
@@
 
-Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
-The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary figure 1).
-
-Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
+Williams and Andersen's [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete's personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation.
+The stress response is central to the model and reflects the bi-directional relationship between athletes' appraisal of, and response to, a stressful athletic situation.
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary Figure 1).
+
+Within Williams and Andersen's [@Williams1998] model, major life events, which is a component of an athlete's history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
 Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].
@@
 
-While the psychosocial factors in Williams and Andersen’s [@Williams1998] model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
+While the psychosocial factors in Williams and Andersen's [@Williams1998] model have received the most research attention, less insight into the mechanisms through which these factors are proposed to exert their effect exists.
 To elaborate, the model suggests that injuries are likely to occur through either increased physiological arousal resulting in increased muscle tension and reduced flexibility or attentional deficits caused by increased distractibility and peripheral narrowing.
@@
 For example, Andersen and Williams [@Andersen1999] found athletes with high life event stress coupled with low social support had greater peripheral narrowing under stressful conditions compared to athletes with the opposing profile; these athletes went on to sustain an increased number of injuries during the following athletic season.
-Rodgers and Landers [@Rogers2005] further supported Andersen and Williams’s [@Andersen1999] earlier findings by identifying that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
+Rodgers and Landers [@Rogers2005] further supported Andersen and Williams's [@Andersen1999] earlier findings by identifying that peripheral narrowing under stress mediated 8.1% of the relationship between negative life events and injury.
 
@@
 For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence [@Lee2017; @Djaoui2017] and may account for the unexplained variance from the psychological predictors.
-Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen’s [@Williams1998] model.
-The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation [@Appaneal2014].
-The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and “widen the window of susceptibility” [@Appaneal2014, p.
+In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's [@Williams1998] model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioural factors that impact sports participation [@Appaneal2014].
+The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and "widen the window of susceptibility" [@Appaneal2014, p.
 74] to a range of undesirable health outcomes including illness and injury.
-Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen’s [@Williams1998] model whilst including other physiological markers of training-related stress.
-
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury.
-However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
+Consequently, the BMSAIH provides an important framework that has enhanced insight into the multi-faceted nature of the injury process by building on Williams and Andersen's [@Williams1998] model whilst including other physiological markers of training-related stress.
+
+The BMSAIH offers researchers several potential avenues for research.
+To date, the physiological mechanisms explored using the BMSAIH have focused on the hormonal response to high intensity training and injury.
+For example, Perna and McDowell [@Perna1995] examined life event stress and cortisol response in athletes following an exhaustive graded exercise test.
+Participants were split into high and low life event stress groups, and the high life event stress group were found to have both higher cortisol in response to the graded exercise test, and increased symptomatology, including muscle complaints and viral illness, over the 30 days following the graded exercise test; however, Perna and McDowell [@Perna1995] did not explicitly examine the relationship between cortisol response to high intensity training and sports injury.
+Research from a wider physiological perspective has identified other important training-related stress markers that have been associated with an increased risk of injury.
+Postural stability, skeletal muscle characteristics and heart rate variability are of particular interest here as they have been examined in a variety of different stress-related disciplines including psychopathology, lifestyle and geriatric research, [@VanOckenburg2015a; @Gervasi2017; @Bailey2013; @Rath2017].
+Importantly, these variables have also been linked to injury occurrence in athletes [@PickeringRodriguez2017; @Plews2012; @Trojian2006; @Williams2017] but have typically been studied in isolation without an assessment of the interaction with the psychosocial factors that are known to contribute to injury.
 Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014].
-Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
-Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
-
-Recently, Bittencourt et al.
-[@Bittencourt2016] advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
-Such an approach posits that injury may arise from a complex “web of determinants” [@Bittencourt2016, p.
+Such an approach fails to capture changes in both psychosocial factors and physiological markers that may occur preceding an injury.
+Importantly, a repeated measurement approach would enable an assessment of how variables and their interactions change with respect to time would provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
+
+Recently, Bittencourt et al. [@Bittencourt2016] advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex "web of determinants" [@Bittencourt2016, p.
 3], where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
@@
 The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
-Furthermore, unlike regression or structural equation models, BN’s do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known [@Olmedilla2018].
-BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
-
-Using the frameworks provided by Williams and Andersen’s [@Williams1998] stress injury model and Appaneal’s BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
-A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
+Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known [@Olmedilla2018].
+BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychosocial and physiological factors implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen's [@Williams1998] stress injury model and Appaneal's BMSAIH model [@Appaneal2014], the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial and physiological stress-related factors with injury occurrence.
+A prospective, repeated measures design incorporating field-based physiological (heart rate variability, muscle stiffness and postural control) and psychosocial measures (Life Event Survey for Collegiate Athletes and Reinforcement Sensitivity Questionnaire Personality Questionnaire) combined with a BN modelling analysis was used to address the study aim.
 
@@
 knitr::opts_chunk$set(echo = FALSE)
+
 ```
@@
 
-A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
-Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
-Participants self-rated competitive level ranged from recreational to international standard. 
+A total of 351 athletes (male: *n* = 231, female: *n* = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
+Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (running, tennis, weightlifting, gymnastics, judo, swimming and golf).
+Participants' self-rated competitive level ranged from recreational to international standard.
 They were engaged in training for their respective sports for at least five hours per week.
-A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
+A total of 162 (46%) participants had sustained an injury in the 12 months prior to the start of the study (male: *n* = 114 [49%], female; *n* = 48 [40%]).
 All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
@@
 
+## Study design
+
+The study adopted a prospective, repeated measures design that included a total of four data collection sessions over a 12-month period.
+The first three data collections were conducted in person, with the final collection (injury reporting) being performed remotely via email (Supplementary Figure 3).
+The primary dependant variable was injury outcome (injured or non-injured), with major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, balance ability and heart rate variability serving as independent variables.
+The following sections describe the specific measures used and outline the procedure for both the data collection and analysis.
+
 ## Measures
@@
 
-Participants self-reported any injuries they sustained at each data collection during the study period.
+At each data collection, participants self-reported all injuries they sustained during the study period via a questionnaire.
 An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours [@Fuller2006; @Fuller2007b; @Timpka2014].
@@
 
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress [@Petrie1992].
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature.
@@
 The LESCA comprises 69 items that reflect possible life events that participants may have experienced.
-Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”.
+Example items include, "Major change in the frequency (increased or decreased) of social activities due to participation in sport", "Major change in the amount (more or less) of academic activity (homework, class time, etc)" and "Major change in level of athletic performance in actual competition (better or worse)".
 Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4 ($extremely\;negative$) and +4 ($extremely\;positive$).
-Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
-A score for total life events was also calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively.
+Negative life event (NLE) and positive life event (PLE) scores were calculated by summing the negative and positive scores, respectively.
+A score for total life events (TLE) was also calculated by summing the absolute values for both negative and positive events.
+Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \< .001) and .48 to .72 (p \< .001) respectively.
 Petrie also provided evidence of discriminant, convergent and predictive validity.
-For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+For this study, composite reliability [@Fornell1981] was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
 The composite reliability for the LESCA in this study was .84.
@@
 A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c].
-The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
+The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., "I regularly try new activities just to see if I enjoy them"), Goal Drive Persistence (e.g., "I am very persistent in achieving my goals"), Reward Reactivity (e.g., "I get a special thrill when I am praised for something I've done well") and Impulsivity (e.g., "I find myself doing things on the spur of the moment").
 Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general.
@@
 A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing HRV.
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings [@Giles2016], which are considered the gold standard for assessing heart rate variability (HRV).
 In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes [@Caminal2018] and under spontaneous breathing conditions [@Plews2017].
@@
 The raw oscillation signal is then processed, and the stiffness parameter is calculated [@Agyapong-Badu2016].
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation < 1.4%) over sets of 10 repetitions [@Aird2012].
+The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties [@Chuang2013; @Pruyn2016; @Nair2014], and has demonstrated good internal consistency (coefficient of variation \< 1.4%) over sets of 10 repetitions [@Aird2012].
 
@@
 
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
-In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
-non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary figure 2). 
+Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by [@Hunt2009].
+In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion), non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary Figure 2).
 To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017].
@@
 The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials.
-One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; 
-b) opening eyes; 
-c) stepping, stumbling, or falling; 
-d) moving the thigh into more than 30 degrees of flexion or abduction; 
-e) lifting the forefoot or heel; and 
-f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
-A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
+One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; b) opening eyes; c) stepping, stumbling, or falling; d) moving the thigh into more than 30 degrees of flexion or abduction; e) lifting the forefoot or heel; and f) remaining out of the testing position for more than 5-seconds [@Riemann1999d].
+A maximum score of 10 errors was possible for each stance.
+Multiple errors occurring simultaneously were recorded as one error.
 A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
@@
 To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation [ICC; @Koo2016] was used on a sample of 40 participants from the first time point.
-The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2)) `), indicating the scoring was reliable.
+The test-retest scoring of BESS resulted in a "good" to "excellent" ICC score (ICC = `r round(icc[["value"]], digits = 2)`, 95% confidence interval = `r paste(round(icc[["lbound"]], digits = 2), "-", round(icc[["ubound"]], digits = 2))`), indicating the scoring was reliable.
 
@@
 At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study.
-With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
+With the coaches' permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation.
 Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Supplementary figure 3).
+A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over a two-year period.
+Within the study period, each participant was asked to complete a total of four data collections, with each data collection separated by a four-month interval (Supplementary Figure 3).
 Participants provided informed consent before data collection commenced.
 
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary figure 4). 
+For each of the data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary Figure 4).
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups.
@@
 For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R [@RCoreTeam2019] for analysis.
-
-### HRV
+The data were downloaded from surveymonkey.com and imported into R [@RCoreTeam2019] for analysis.
+
+### Heart rate variability
 
@@
 
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the _RHRV_ package [@Rodriguez-Linares2017] was used to calculate HRV indices.
+Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R [@RCoreTeam2019] where the *RHRV* package [@Rodriguez-Linares2017] was used to calculate HRV indices.
 Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings [@Laborde2017; @Malik1996].
@@
 
-Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
+Two Bayesian Networks (BN) were used to explore the relationships between the psychosocial measures, physiological markers of stress and sports injury.
 A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014].
 A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [@Pearl1988].
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
+However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of causal [@Scutari2014].
 The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012].
@@
 When there are a large number of variables in the network, limiting the number of levels has the benefit of producing a network that is more parsimonious in terms of parameters.
-For example, a network with 10 variables each with two levels has 100 (10^2) possible parameter combinations, however the same network with three levels has 1000 (10^3) possible  parameter combinations, the latter being significantly more computationally expensive.
+For example, a network with 10 variables each with two levels has 100 (10\^2) possible parameter combinations, however the same network with three levels has 1000 (10\^3) possible parameter combinations, the latter being significantly more computationally expensive.
 Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio [@Scutari2014].
@@
 For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
-Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
+Furthermore, the structure of a network can be obtained from both empirical data *and* prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
 In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
@@
 To prepare the data for the BN, missing values in the dataset were first imputed.
-Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings.
+Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$HRV == 0)` missing heart rate recordings.
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
 A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
-In the current study, the _caret_ R package [@Kuhn2008] was used to impute the missing values. 
+In the current study, the *caret* R package [@Kuhn2008] was used to impute the missing values.
 A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset.
@@
 Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
+In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
 
@@
 To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
-Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist).
+Prior knowledge about the network structure was included by providing a list of arcs that are always *restricted* from being in the network (blacklist), and a list of arcs that are always *included* in the network (whitelist).
 Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014].
@@
 Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
-Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
+Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 $\rightarrow$ T2\_2 and T2\_1 $\rightarrow$ T3\_2).
 Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
@@
 Finally, arc direction was restricted between specific nodes within the explanatory variables.
-Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
+Arcs from competitive level $\rightarrow$ gender, baseline negative life events (NLE) $\rightarrow$ gender and baseline NLE $\rightarrow$ sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
 All subsequent models used the same blacklist.
@@
 
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (nleleg = `r round(le[1,2], 2)`, tlelg only = `r round(le[2,2],2)`, tlelg and nlelg = `r round(le[3,2],2)`).
-Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure.
+Initially, both negative life events (NLE) and total life events (TLE) were included in the network structure, however, the network score was improved when only NLE or TLE was included (NLE = `r round(le[1,2], 2)`, TLE only = `r round(le[2,2],2)`, TLE and NLE = `r round(le[3,2],2)`).
+Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
 TLE score was not included in the final structure.
 
-The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
+The subscales representing the Behavioural Activation System (Reward Reactivity [RR], Reward Interest [RI], Goal Drive Persistence [GDP] and Impulsivity [I]) showed limited connection to other variables in the network.
 Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
@@
 Finally, both total score and asymmetry for balance were included in the initial network.
-However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
+However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
 Therefore, balance asymmetry was removed from the final network structure.
@@
 ) %>% 
-  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
+  mutate(var = ifelse(var == "HRV", "HRV", var)) %>% 
   kable("latex",
@@
 Preliminary network structures also revealed strong dependencies between the same variables at sequential time points.
-For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%.
-Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network.
+For example, the probability that stiffness\_1 and stiffness\_2 were both "High", or both "Low" was approximately 80%.
+Including the arcs between the same variables from X\_1 → X\_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a "Low" or "High" state for each pair of variables in the network.
 Therefore, these arcs were blacklisted from the network.
@@
 Arcs that were present in at least 30% of the models were included in the averaged model.
-The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction.
+The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc's direction.
 An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
-In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013].
-Arcs below 0.3 were not included in the model. The full Table of arc strengths for the first and second network are available in Supplementary Table 1 and Supplementary Table 3 respectively.
+In the respective study arcs above 0.5 were considered "significant" with arcs below 0.5 and above 0.3 "non-significant" [@Scutari2013].
+Arcs below 0.3 were not included in the model.
+The full table of arc strengths for the first and second network are available in Supplementary Table 2 and Supplementary Table 3 respectively.
 
@@
 To conduct a CPQ, the joint probability distribution of the nodes was modified to include a new piece of evidence.
-The query allows the odds of a particular node state (e.g., injured_1 = “injured”) to be calculated based on the new evidence.
-CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
+The query allows the odds of a particular node state (e.g., injured\_1 = "injured") to be calculated based on the new evidence.
+CPQ were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
 The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query [@Scutari2014].
 Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
-Of particular interest were the variables that were connected to “injured” nodes, which were examined in the network using the Markov blanket of “injured_1” and “injured_2”.
+Of particular interest were the variables that were connected to "injured" nodes, which were examined in the network using the Markov blanket of "injured\_1" and "injured\_2".
 A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network [@Fuster-Parra2017].
-CPQ’s were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the “injured” state.
+CPQ were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the "injured" state.
 
@@
 The "injured" variable was also modified to represent whether a participant had sustained an injury at any point over the duration of the study or were healthy for the duration of the study.
-Setting the data up in this way enabled the construction of a network that explicitly modelled the _amount_ of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from "Low" to "High".
+Setting the data up in this way enabled the construction of a network that explicitly modelled the *amount* of change within variables between time points, as opposed to the first network that only captured changes when the median threshold was crossed from "Low" to "High".
 
@@
 Similar to the first network, blacklists were used to prevent arcs from independent variables $\rightarrow$ explanatory variables.
-In addition, "nlebase" variable was dropped from the list of explanatory variables to allow the _changes_ in negative life events to be the only life event variable in the network. 
-The final network was obtained using the same approach as the first network.
-
-#### Network Inference
+In addition, baseline negative life events was dropped from the list of explanatory variables to allow the *changes* in negative life events to be the only life event variable in the network.
+The final network was obtained using the same approach as the first network.Network Inference
 
@@
 The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
+Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) $\rightarrow$ Fight-Flight-Freeze System (FFFS) and HRV $\rightarrow$ FFFS.
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then examined using a Bayesian linear regression models with the _brms_ R package [@Burkner2017a] to determine the relationship between these nodes. 
+The samples were then examined using a Bayesian linear regression models with the *brms* R package [@Burkner2017a] to determine the relationship between these nodes.
 Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
@@
   janitor::tabyl(count)
-```
-
-During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
+
+```
+
+During the study, 46% (*n* = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
 Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
 Table \ref{tab:table3} shows the number and percentage of injury types sustained by both male and female participants.
+An additional breakdown of injury by sport, type (acute or chronic) and injury location is available in Supplementary tables 4 and 5.
 
@@
 
-from <- c("BIS_1", "BIS_2", "FFFS_1", "nlelg_1", "nlelg_2")
+from <- c("BIS_1", "BIS_2", "FFFS_1", "NLE_1", "NLE_2")
 to <- c("FFFS_1", "FFFS_2", "BIS_2", "BIS_1", "BIS_2")
@@
 
-cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
+cbind(inj_summary[c(1:3), c(2:4)] %>% 
+        mutate(variable = ifelse(variable == "Other", "Other (bone, brain and skin)", variable)), inj_summary[c(4:6), c(3:4)]) %>% 
   kable(
@@
     align = "lcccc",
+    format = "latex",
     escape = F
@@
   add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
-  footnote(general = "Other included bone, skin and brain injuries.") %>%
   kable_styling(latex_options = "HOLD_position") %>%
   row_spec(0, bold = TRUE)
+
+  #add_footnote("Other included bone, skin and brain injuries.") %>%
 
@@
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
-The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
+The "individual or team sport" node had strong arcs to hours per week (`r str_finder("ind_team", "hours")`) and baseline negative life events (NLE) (`r str_finder("ind_team", "nlebase")`).
 Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
 Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
-The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
-High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
+The arc from competitive level $\rightarrow$ balance\_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
+High gender-based connections were reported for the arcs from gender $\rightarrow$ stiffness\_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness\_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
 Irrespective of sport or gender, strong connections were found between explanatory variables.
-The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.
+The arc from baseline NLE (nlebase) $\rightarrow$ Reward Interest (RI\_1) had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI\_1 being in the "High" state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from "Low" to "High".
 
 The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability.
-The arc from pi $\rightarrow$ stiffness_1 was `r str_finder("pi", "stiffness_1")` with athletes who reported an injury in the preceding 12 months being more likely to have “High” (`r x[26,1]`) compared to “Low” (`r x[28,1]`).
+The arc from previous injury $\rightarrow$ stiffness\_1 was `r str_finder("pi", "stiffness_1")` with athletes who reported an injury in the preceding 12 months being more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`).
 stiffness.
-Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (`r str_finder("BIS_1", "FFFS_1")`) and BIS_2 $\rightarrow$ FFFS_2 (`r str_finder("BIS_2", "FFFS_2")`) were identified.
-In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
-The arc between nlelg $\rightarrow$  BIS had a strength of `r str_finder("nlelg_1", "BIS_1")` for nlelg_1  $\rightarrow$ BIS_1, and `r str_finder("nlelg_2", "BIS_2")` nlelg_2 and BIS_2.
-"Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]` for nlelg_1  $\rightarrow$ BIS_1, and `r y[17,1]` to `r y[18,1]` for nlelg_2  $\rightarrow$ BIS_2.
+Strong arcs were present between Behavioural Inhibition System (BIS) $\rightarrow$ Fight-Flight-Freeze System (FFFS; BIS\_1 $\rightarrow$ FFFS\_1 = `r str_finder("BIS_1", "FFFS_1")`, BIS\_2 $\rightarrow$ FFFS\_2 = `r str_finder("BIS_2", "FFFS_2")`).
+In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for \_1, `r y[6,1]` for \_2) compared to "Low" (`r y[1,1]` for \_1, `r y[5,1]` for \_2).
+The arc between NLE $\rightarrow$ BIS had a strength of `r str_finder("NLE_1", "BIS_1")` for NLE\_1 $\rightarrow$ BIS\_1, and `r str_finder("NLE_2", "BIS_2")` NLE\_2 and BIS\_2.
+"Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]` for NLE\_1 $\rightarrow$ BIS\_1, and `r y[17,1]` to `r y[18,1]` for NLE\_2 $\rightarrow$ BIS\_2.
 
@@
 
-### Markov blanket for injured_1
+### Markov blanket for injured\_1
 
@@
 set.seed(1001)
-var <- c("balance_1", "stiffness_1", "nlelg_1", "hours")
+var <- c("balance_1", "stiffness_1", "NLE_1", "hours")
 
@@
   tallc,
-  c("balance_1", "stiffness_1", "nlelg_1", "hours", "clevel"),
+  c("balance_1", "stiffness_1", "NLE_1", "hours", "clevel"),
   "injured_1",
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1).
-The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week, negative life events (NLE\_1), muscle stiffness (stiffness\_1), current competitive level and balance (balance\_1).
+The arc between NLE\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
 
@@
 
-The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table4}.
-The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
+The CPQ for injured\_1 in the "injured" state for all variables that were linked to injured\_1 is shown in Table \ref{tab:table4}.
+The probability of injured\_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
 Negative life events had a negligible effect when moving from the "Low" to "High" state.
@@
 
-The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
+The second CPQ investigated the outcome of injured\_1 being "injured" conditional on all variables in the Markov blanket.
 The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in Supplementary Table 4).
-The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
+The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured\_1 being in the "injured" state.
 When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
 
-Negative life event stress had a negligible effect on the probability of injury, only influencing  injured_1 when all other variable were fixed to "Low".
-In this instance, the probability of injured_1 being "injured" rose marginally from 0.04 to 0.19, when negative life events was in the "Low" and "High" states, respectively.
+Negative life event stress had a negligible effect on the probability of injury, only influencing injured\_1 when all other variable were fixed to "Low".
+In this instance, the probability of injured\_1 being "injured" rose marginally from 0.04 to 0.19, when negative life events was in the "Low" and "High" states, respectively.
 
@@
 
-### Markov blanket of injured_2
+### Markov blanket of injured\_2
 
@@
 set.seed(1002)
-var2 <- c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2")
+var2 <- c("balance_2", "stiffness_2", "NLE_2", "FFFS_1", "HRV_2")
 query3 <- map_dfr(var2, ~newprobtable(tallc,
@@
 query4 <- newprobtable(tallc, 
-                       c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2"),
+                       c("balance_2", "stiffness_2", "NLE_2", "FFFS_1", "HRV_2"),
                        "injured_2",
@@
 
-The Markov blanket for injured_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2 (Table \ref{tab:table6}).
-The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
-Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ hrv_2 were also present in the Markov blanket for injured_2.
-Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
-FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
+The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, Fight-Flight-Freeze System (FFFS\_1), stiffness\_2, balance\_2 and heart rate variability (HRV\_2; Table \ref{tab:table6}).
+The arc between stiffness\_2 $\rightarrow$ injured\_2 was comparable to the arc between stiffness\_1 $\rightarrow$ injured\_1.
+Very weak arcs (0.3) between injured\_2 $\rightarrow$ balance\_2 and injured\_2 $\rightarrow$ HRV\_2 were also present in the Markov blanket for injured\_2.
+Similar to injured\_1, stiffness\_2 doubled the probability of injured\_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
+FFFS\_1 in the "Low" state increased probability of injured\_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
 "High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.
@@
 fffs_query <- query4 %>% 
-  filter(stiffness_2 == "High" & nlelg_2 == "High" & balance_2 == "High" & hrv_2 == "High")
-```
-
-The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in Supplementary Table 5).
-The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
-With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
-The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".
+  filter(stiffness_2 == "High" & NLE_2 == "High" & balance_2 == "High" & HRV_2 == "High")
+```
+
+The three lowest and highest conditional probabilities based on all the variables in injured\_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in Supplementary Table 5).
+The combination of "Low" FFFS\_1, "High" stiffness\_2, "High" balance resulted in the greatest probability of injured\_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
+With all other variables held in the "High" state, the probability of injured\_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS\_1 was in the "Low" compared to "High" state.
+The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured\_2 being "injured".
 
@@
 
-The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
-An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
-Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively
-Similarly, the arc between hrv $\rightarrow$ FFFS was `r str_cfinder("hrv","FFFS")`.
-The arcs between BIS $\rightarrow$ FFFS and hrv $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
-A separate linear regression examined the interaction between BIS and hrv to be examined.
+The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/ An arc between Behavioural Inhibition System (BIS) $\rightarrow$ Fight-Flight-Freeze System (FFFS) with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
+Arcs between competitive level $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively Similarly, the arc between HRV $\rightarrow$ FFFS was `r str_cfinder("HRV","FFFS")`.
+The arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
+A separate linear regression examined the interaction between BIS and HRV to be examined.
 
@@
 Results from the Bayesian linear regression model are presented in Table \ref{tab:table8} and include 95% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
-There was no clear interaction between hrv and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
+Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in HRV where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
+There was no clear interaction between HRV and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).
 
@@
     .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.",
+    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and HRV.",
     col.names = c("Term", "Estimate", "Error", "95\\% CI"),
@@
 
-The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Figure \ref{fig:fig5}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
+The Markov blanket for the "injured" node contained previous injury, gender, training hours per week and stiffness and negative life events (NLE; Figure \ref{fig:fig5}).
+For stiffness and NLE, the values in the nodes represent the standardised change between time point.
+Combinations of NLE and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress.
-Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
-With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.
+Changes in both NLE and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
+With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with NLE at 1 SD below an 1 SD above respectively.
 
@@
 Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
-The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
+The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychosocial and physiological factors that may occur preceding an injury.
 
 The first network revealed several links between the injured nodes and other variables in the network.
-A combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for “injured” nodes.
-The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete’s risk of injury.
-In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury [@Appaneal2014].
+A combination of high stiffness and poor balance resulted in the highest probability of injury in the Markov blankets for "injured" nodes.
+The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete's risk of injury.
+In the second network, the highest probability of injury was observed when increases in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychosocial and physiological stress-related factors may combine additively to increase the risk of injury [@Appaneal2014].
 
@@
 However, high stiffness may only increase the risk of injury if other factors are also present.
-To elaborate, the combination of high stiffness _and_ poor balance was found to result in the greatest probability of injury.
+To elaborate, the combination of high stiffness *and* poor balance was found to result in the greatest probability of injury.
 In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
@@
 At both injured nodes in the first network, balance was linked to injury.
-Despite the weak arc strength at both injured nodes, a “High” balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014], was found to increase the probability of injury.
+Despite the weak arc strength at both injured nodes, a "High" balance score, which is considered indicative of impaired postural stability [@Romero-Franco2014], was found to increase the probability of injury.
 This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk [@Riemann1999d; @Romero-Franco2014; @Trojian2006].
@@
 For example, an athlete who reports a negative life event score of 1 during the first time point, but then a score of 5 at the second time point will have a 400% increase in their life event score.
-Although the absolute score would be “Low”, the relative increase between time points may have been caused by a significant event in the athlete’s life that had a considerable psychological and physiological effect [@Appaneal2014].
-Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete’s life events to be assessed [cf. @Ivarsson2014].
-
-During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
+Although the absolute score would be "Low", the relative increase between time points may have been caused by a significant event in the athlete's life that had a considerable psychological and physiological effect [@Appaneal2014].
+Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed [cf. @Ivarsson2014].
+
 The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model [@Ivarsson2017].
 During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-Given the reported reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
-When negative life events were “High” the probability of injury showed a negligible change at the injured_1 node and decreased by `r query3[which(query3$variable == "nlelg_2"), 3] -  query3[which(query3$variable == "nlelg_2"), 2]` at the injured_2 node.
+Given the reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
+When negative life events were "High" the probability of injury showed a negligible change at the injured\_1 node and decreased by `r query3[which(query3$variable == "NLE_2"), 3] -  query3[which(query3$variable == "NLE_2"), 2]` at the injured\_2 node.
 One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
@@
 
-Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, "Low" HRV increased the probability of injury from `r query3[which(query3$variable == "hrv_2"), 3]` ("High" HRV) to `r query3[which(query3$variable == "hrv_2"), 2]` ("Low" HRV).
+Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, "Low" HRV increased the probability of injury from `r query3[which(query3$variable == "HRV_2"), 3]` ("High" HRV) to `r query3[which(query3$variable == "HRV_2"), 2]` ("Low" HRV).
 This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury [@Williams2017; @Bellenger2016; @Gisselman2016].
-An arc between FFFS_1 and injured_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
-Interestingly, the “Low” FFFS score was also related to injuries at subsequent time points.
-One possible explanation for this finding could be that those athletes who reported “Low” FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
+An arc between FFFS\_1 and injured\_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the "High" and "Low" states respectively.
+Interestingly, the "Low" FFFS score was also related to injuries at subsequent time points.
+One possible explanation for this finding could be that those athletes who reported "Low" FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
 The RST theory proposes that higher levels of FFFS increase avoidance motivation [@Corr2016b], and therefore "High" FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
 The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS [@Corr2013].
-The first network reported an association between “High” FFFS and “High” BIS, while the second network found an association between increases in FFFS and increases in BIS.
+The first network reported an association between "High" FFFS and "High" BIS, while the second network found an association between increases in FFFS and increases in BIS.
 High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD [@Chalmers2014; @Pulopulos2018].
@@
 
-While the BN’s provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
-Firstly, the choice was made to binarise variables in the first network so only “Low” and “High” states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner @Qian2015 highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
-For example, the meaning of a “Low” and “High” value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
+While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
+Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
+Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner [@Qian2015] highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+For example, the meaning of a "Low" and "High" value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
 Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
-However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes’ responses.
+However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes' responses.
 For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis [@Ross2011].
@@
 
-The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychological and physiological perspective.
+The ability to capture large sample sizes of injured athletes has been recognised as a significant challenge in sports injury research [@Ruddy2019] and affects the success to which stress-injury interactions can be identified and understood.
+This is further exacerbated in studies that employ longitudinal prospective repeated measures designs where high levels of participant retention are required.
+Consistent with Williams and Andersen's (1998) stress-injury model, we therefore employed a global definition of injury as the primary outcome measure, to optimize sample size of injury occurrences across the repeated measures prospective study design.
+As suggested by Ruddy et al. (2019), a larger number of observations and injury events is needed to improve the ability to identify and to make more meaningful predictions.
+Further conventional injury exposure measures (sport type, training load) were subsequently discretised or reduced for the data capture, which caused some loss of information and has been suggested to limit the ability to capture risk profiles [@Carey2018].
+Using pooled or discretised definitions for some measures hindered a detailed causal-effect insight into injury-specific risk factors.
+However, the more holistic, interdisciplinary approach adopted in this study was benefited by the ability to gain and substantiate a more complete picture of the complex, multifaceted stress-injury interactions that exist in sport.
+
+The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychosocial and physiological perspective.
 The variables used in the present study were not definitive.
@@
 
-To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Baysian Network analytical techniques,  Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Finally, the findings of this study have important practical implications for athletes, coaches, and clinicians in relation to the additive and interactive effects of multiple sources of stress on injury occurrence.
+Specifically, the study evidenced a combined effect of psychosocial and physiological stress-related factors that could increase the probability of injury to a greater extent than any isolated factor.
+When assessing an athlete's training plan, readiness to engage in, and recovery from, training, coaches and clinicians should employ a risk profile that integrates multifaceted sources of stress.
+For example, in addition to monitoring training loads and using tools to determine an athlete's physiological status, coaches need to also consider an athlete's psychological state.
+In particular, when an athlete is facing significant life event stress, adjusted training intensity and volume may be necessitated to support athlete coping with the additional duress and to subsequently safeguard optimal health and well-being.
+In essence, injury risk is exacerbated when an athlete is experiencing psychological stress due to exposure to negative life events and exhibiting physiological responses associated with an increased injury potential.
+The identification of such a "high risk" profile is subsequently important in helping to monitor and reduce injury risk for athletes.
+For example, while high muscle stiffness is important for optimal performance [@Pruyn2015], this study demonstrated it can heighten injury risk, which is likely to be exacerbated when accompanied by the experience of negative life events by an athlete.
+In order to understand how an athlete's injury risk may increase over time, it is important to acknowledge the breadth and interaction of stress-related factors that could heighten susceptibility and be receptive to training and life experience changes.
+
+To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Bayesian Network analytical techniques.
+Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
 Future research combining a repeated measures approach and complex analyses of the interactions between multifaceted stress-related measures are advocated to enhance understanding of the injury occurrence in sport.
@@
 
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
-
-# Acknowledgements
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+# Author Contributions
 
@@
 
-The content of this manuscript has been published as part of the thesis of Harry Fisher, @Fisher2019.
-
 # References
 
+```{=tex}
 \begingroup
@@
 \setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
+```
+::: {#refs custom-style="Bibliography"}
+:::
+
 \endgroup

### 2020-10-15 manuscript.tex
Update following reviewers comments

--- 2020-08-17 manuscript.tex
@@
   pdftitle={An interdisciplinary examination of stress and injury occurrence in athletes},
+  pdfauthor={Harry Fisher1, Dr Marianne Gittoes1, Professor Lynne Evans1, Miss Leah Bitchell1, Dr Richard Mullen2, \& Dr Marco Scutari3},
   pdflang={en-EN},
@@
   {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
 \shorttitle{Stress and sports injury}
-\author{Harry Fisher\textsuperscript{1}, Dr Marianne Gittoes\textsuperscript{1}, Professor Lynne Evans\textsuperscript{1}, Miss Leah Bitchell\textsuperscript{1}, Dr Richard Mullen\textsuperscript{2}, \& Dr Marco Scutari\textsuperscript{3}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Cardiff Metropolitan University, Cardiff, United Kingdom\\\textsuperscript{2} Brunel University, London, United Kingdom\\\textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA), Manno, Switzerland}
-\authornote{12856 words in text body and refence section
-
-Number of tables: 10
-
-Number of figures: 5
-
-Format: British English
-
-Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: harryfisher21@gmail.com}
 \keywords{Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology}
@@
 \title{An interdisciplinary examination of stress and injury occurrence in athletes}
-
+\author{Harry Fisher\textsuperscript{1}, Dr Marianne Gittoes\textsuperscript{1}, Professor Lynne Evans\textsuperscript{1}, Miss Leah Bitchell\textsuperscript{1}, Dr Richard Mullen\textsuperscript{2}, \& Dr Marco Scutari\textsuperscript{3}}
 \date{}
 
+\authornote{
+
+13867 words in text body and refence section
+
+Number of tables: 10
+
+Number of figures: 5
+
+Format: British English
+
+Correspondence concerning this article should be addressed to Harry Fisher, Cyncoed Road, Cardiff, CF23 6XD. E-mail: \href{mailto:harryfisher21@gmail.com}{\nolinkurl{harryfisher21@gmail.com}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Cardiff Metropolitan University\\\textsuperscript{2} Brunel University\\\textsuperscript{3} Istituto Dalle Molle di Studi sull'Intelligenza Artificiale (IDSIA)}
+
 \abstract{
-This paper adopts a novel, interdisciplinary approach to explore the relationship between psychosocial factors, physiological stress-related markers and occurrence of injury in athletes using a repeated measures design across a 2-year data collection period.
-At three data collection time-points, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
+This paper adopts a novel, interdisciplinary approach to explore the relationship between stress-related psychosocial factors, physiological markers and occurrence of injury in athletes using a repeated measures prospective design.
+At four data collection time-points, across 1-year of a total 2-year data collection period, athletes completed measures of major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, heart rate variability and postural stability, and reported any injuries they had sustained since the last data collection.
 Two Bayesian networks were used to examine the relationships between variables and model the changes between data collection points in the study.
@@
 The stress response is central to the model and reflects the bi-directional relationship between athletes' appraisal of, and response to, a stressful athletic situation.
-The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary figure 1).
+The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (Supplementary Figure 1).
 
@@
 For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence (26,27) and may account for the unexplained variance from the psychological predictors.
-Appaneal and Perna (28) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (14) model.
-The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioral factors that impact sports participation (28).
+In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna (28) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (14) model.
+The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioural factors that impact sports participation (28).
 The central tenet of the BMSAIH is that psychosocial distress (e.g., negative life events) may act synergistically with training-related stress as a result of high-intensity and high-volume sports training, and ``widen the window of susceptibility'' (28) to a range of undesirable health outcomes including illness and injury.
@@
 
-Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence (29--31), other training-related stress markers including heart rate variability (32,33), muscle stiffness (34) and postural stability (35) have been reported to be associated with an increased risk of injury.
-However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence.
-Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries (36).
-Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury.
-Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
-
-Recently, Bittencourt et al.~(37) advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
-Such an approach posits that injury may arise from a complex ``web of determinants'' (37), where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
+The BMSAIH offers researchers several potential avenues for research.
+To date, the physiological mechanisms explored using the BMSAIH have focused on the hormonal response to high intensity training and injury.
+For example, Perna and McDowell (29) examined life event stress and cortisol response in athletes following an exhaustive graded exercise test.
+Participants were split into high and low life event stress groups, and the high life event stress group were found to have both higher cortisol in response to the graded exercise test, and increased symptomatology, including muscle complaints and viral illness, over the 30 days following the graded exercise test; however, Perna and McDowell (29) did not explicitly examine the relationship between cortisol response to high intensity training and sports injury.
+Research from a wider physiological perspective has identified other important training-related stress markers that have been associated with an increased risk of injury.
+Postural stability, skeletal muscle characteristics and heart rate variability are of particular interest here as they have been examined in a variety of different stress-related disciplines including psychopathology, lifestyle and geriatric research, (30--33).
+Importantly, these variables have also been linked to injury occurrence in athletes (34--37) but have typically been studied in isolation without an assessment of the interaction with the psychosocial factors that are known to contribute to injury.
+Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries (38).
+Such an approach fails to capture changes in both psychosocial factors and physiological markers that may occur preceding an injury.
+Importantly, a repeated measurement approach would enable an assessment of how variables and their interactions change with respect to time would provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.
+
+Recently, Bittencourt et al.~(39) advocated a move away from studying isolated risk factors and instead, adopt a complex systems approach in order to understand injury occurrence.
+Such an approach posits that injury may arise from a complex ``web of determinants'' (39), where different factors interact in unpredictable and unplanned ways, but result in a global outcome pattern of either adaptation or injury.
 Capturing the uncertainty and complexity of the relationships between different variables using an appropriate interdisciplinary analysis within the framework of a complex systems approach is challenging.
-Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables (38).
+Bayesian network (BN) modelling provides one solution by allowing the construction of graphical probabilistic models using the underlying structure that connects different variables (40).
 The learned BN structure can be used for inference by obtaining the posterior probabilities of a particular variable for a given query (e.g., if the value of variable A is x and the value of variable B is y, what is the probability variable C of being value z?).
-Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known (39).
-BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychological and physiological factors implicated in the injury process.
-
-Using the frameworks provided by Williams and Andersen's (14) stress injury model and Appaneal's BMSAIH model, the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychological and physiological stress-related factors with injury occurrence.
-A prospective, repeated measures design incorporating field-based physiological and psychological measures combined with a BN modelling analysis was used to address the study aim.
+Furthermore, unlike regression or structural equation models, BN's do not distinguish between dependent and independent variables when the underlying relationship in the network may not be known (41).
+BN modelling subsequently provides a valuable but underused interdisciplinary approach to investigating the complex and unpredictable interactions of psychosocial and physiological factors implicated in the injury process.
+
+Using the frameworks provided by Williams and Andersen's (14) stress injury model and Appaneal's BMSAIH model (28), the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial and physiological stress-related factors with injury occurrence.
+A prospective, repeated measures design incorporating field-based physiological (heart rate variability, muscle stiffness and postural control) and psychosocial measures (Life Event Survey for Collegiate Athletes and Reinforcement Sensitivity Questionnaire Personality Questionnaire) combined with a BN modelling analysis was used to address the study aim.
 
@@
 A total of 351 athletes (male: \emph{n} = 231, female: \emph{n} = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}).
-Participants had an average age of 22.0 \(\pm\) 7.0 years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf.
-Participants self-rated competitive level ranged from recreational to international standard.
+Participants had an average age of 22.0 \(\pm\) 7.0 years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (running, tennis, weightlifting, gymnastics, judo, swimming and golf).
+Participants' self-rated competitive level ranged from recreational to international standard.
 They were engaged in training for their respective sports for at least five hours per week.
-A total of 126 (49.03\%) participants had sustained an injury in the 12 months prior to the start of the study (male: \emph{n} = 114 {[}49\%{]}, female; \emph{n} = 48 {[}40\%{]}).
+A total of 162 (46\%) participants had sustained an injury in the 12 months prior to the start of the study (male: \emph{n} = 114 {[}49\%{]}, female; \emph{n} = 48 {[}40\%{]}).
 All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.
@@
 
+\hypertarget{study-design}{%
+\subsection{Study design}\label{study-design}}
+
+The study adopted a prospective, repeated measures design that included a total of four data collection sessions over a 12-month period.
+The first three data collections were conducted in person, with the final collection (injury reporting) being performed remotely via email (Supplementary Figure 3).
+The primary dependant variable was injury outcome (injured or non-injured), with major life events, the reinforcement sensitivity theory personality questionnaire, muscle stiffness, balance ability and heart rate variability serving as independent variables.
+The following sections describe the specific measures used and outline the procedure for both the data collection and analysis.
+
 \hypertarget{measures}{%
@@
 
-Participants self-reported any injuries they sustained at each data collection during the study period.
-An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours (40--42).
-Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries (43).
+At each data collection, participants self-reported all injuries they sustained during the study period via a questionnaire.
+An injury was defined as any sports related medical problem causing the athlete to miss or modify their usual training routine for at least 24 hours (42--44).
+Minor scrapes and bruises that may require certain modifications (e.g., strapping or taping) but did not limit continued participation were not considered injuries (45).
 Injury status (did / did not sustain an injury) served as the main outcome measure.
@@
 
-A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress (44).
+A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants' history of life event stress (46).
 The LESCA is the most widely used measure of major life events for athletes in the sports injury literature.
@@
 Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4 (\(extremely\;negative\)) and +4 (\(extremely\;positive\)).
-Negative and positive life event scores were calculated by summing the negative and positive scores, respectively.
-A score for total life events was also calculated by summing the absolute values for both negative and positive events.
-Petrie (1992) (44) reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001) respectively.
+Negative life event (NLE) and positive life event (PLE) scores were calculated by summing the negative and positive scores, respectively.
+A score for total life events (TLE) was also calculated by summing the absolute values for both negative and positive events.
+Petrie (1992) (46) reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p \textless{} .001) and .48 to .72 (p \textless{} .001) respectively.
 Petrie also provided evidence of discriminant, convergent and predictive validity.
-For this study, composite reliability (45) was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
+For this study, composite reliability (47) was preferred to Cronbach's alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire.
 The composite reliability for the LESCA in this study was .84.
@@
 
-A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology (46).
+A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology (48).
 The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., ``I am the sort of person who easily freezes-up when scared''), Behavioural Inhibition System (BIS; e.g., ``When trying to make a decision, I find myself constantly chewing it over'') and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., ``I regularly try new activities just to see if I enjoy them''), Goal Drive Persistence (e.g., ``I am very persistent in achieving my goals''), Reward Reactivity (e.g., ``I get a special thrill when I am praised for something I've done well'') and Impulsivity (e.g., ``I find myself doing things on the spur of the moment'').
@@
 The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
-The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81. Further details regarding the revised RST are in Supplementary Appendix 1.
+The composite reliabilities for each subscale were; BIS = 0.92, FFFS = 0.77, GDP = 0.87, I = 0.71, RI = 0.77, RR = 0.81.
+Further details regarding the revised RST are in Supplementary Appendix 1.
 
@@
 A Polar V800 heart rate monitor (HRM) and Polar H7 Bluetooth chest strap (Polar OY, Finland) was used to collect inter-beat interval (IBI) data.
-IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings (47), which are considered the gold standard for assessing HRV.
-In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes (48) and under spontaneous breathing conditions (49).
+IBI recordings using the Polar V800 are highly comparable (ICC = 1.00) with ECG recordings (49), which are considered the gold standard for assessing heart rate variability (HRV).
+In addition, HRV indices calculated from IBI and ECG data have shown a strong correlation (r = .99) in athletes (50) and under spontaneous breathing conditions (51).
 
@@
 The impulse causes natural damped oscillations in the tissue, which are recorded by a three-axis digital accelerometer sensor in the device.
-The raw oscillation signal is then processed, and the stiffness parameter is calculated (50).
-The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties (51--53), and has demonstrated good internal consistency (coefficient of variation \textless{} 1.4\%) over sets of 10 repetitions (54).
+The raw oscillation signal is then processed, and the stiffness parameter is calculated (52).
+The MyotonPRO has previously been reported to be a reliable and valid tool for the measurement of in-vivo tissue stiffness properties (53--55), and has demonstrated good internal consistency (coefficient of variation \textless{} 1.4\%) over sets of 10 repetitions (56).
 
@@
 
-Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by (55).
-In total, each trial of the mBESS was performed without shoes (56) and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
-non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion)
-and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary figure 2).
-To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (57).
+Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by (57).
+In total, each trial of the mBESS was performed without shoes (58) and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion), non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Supplementary Figure 2).
+To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant (59).
 Participants were asked to maintain each stance for a total of 20-seconds.
@@
 The error identification criteria from the original BESS protocol was used by the lead researcher who scored all the BESS trials.
-One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests;
-b) opening eyes;
-c) stepping, stumbling, or falling;
-d) moving the thigh into more than 30 degrees of flexion or abduction;
-e) lifting the forefoot or heel; and
-f) remaining out of the testing position for more than 5-seconds (58).
-A maximum score of 10 errors was possible for each stance. Multiple errors occurring simultaneously were recorded as one error.
+One error was recorded if any of the following movements were observed during each trial: a) lifting hands off iliac crests; b) opening eyes; c) stepping, stumbling, or falling; d) moving the thigh into more than 30 degrees of flexion or abduction; e) lifting the forefoot or heel; and f) remaining out of the testing position for more than 5-seconds (60).
+A maximum score of 10 errors was possible for each stance.
+Multiple errors occurring simultaneously were recorded as one error.
 A participant was given the maximum score of 10 if they remained out of the stance position for more than 5-seconds.
 A total score was calculated by summing the total number of errors recorded on all stances (DL, NLD and TS, on foam and firm surfaces).
-To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation (59) was used on a sample of 40 participants from the first time point.
+To assess the intra-rater reliability, a single measurement, absolute agreement, two-way mixed effects model for the intraclass correlation (61) was used on a sample of 40 participants from the first time point.
 The test-retest scoring of BESS resulted in a ``good'' to ``excellent'' ICC score (ICC = 0.93, 95\% confidence interval = 0.88 - 0.96), indicating the scoring was reliable.
@@
 Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions.
-A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over two consecutive 12-month periods.
-Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Supplementary figure 3).
+A repeated measures prospective cohort design was used to assess athletes' major life events, stress-related physiological markers and injury status over a two-year period.
+Within the study period, each participant was asked to complete a total of four data collections, with each data collection separated by a four-month interval (Supplementary Figure 3).
 Participants provided informed consent before data collection commenced.
 
-For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary figure 4).
+For each of the data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Supplementary Figure 4).
 To ensure all measures could be collected within a viable time-frame, participants were separated into two groups.
@@
 For injury reporting, participants were asked to record any injuries that they had sustained since the last data collection.
-The data were downloaded from surverymonkey.com and imported into R (60) for analysis.
-
-\hypertarget{hrv}{%
-\subsubsection{HRV}\label{hrv}}
+The data were downloaded from surveymonkey.com and imported into R (62) for analysis.
+
+\hypertarget{heart-rate-variability-1}{%
+\subsubsection{Heart rate variability}\label{heart-rate-variability-1}}
 
@@
 Participants were seated and asked to remain as still as possible for the duration of the recording.
-No attempt was made to control participants' respiratory frequency or tidal volume (61).
+No attempt was made to control participants' respiratory frequency or tidal volume (63).
 Inter-beat interval (IBI) data was collected for 10-minutes at a sampling frequency of 1000 Hz.
 
-Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R (60) where the \emph{RHRV} package (62) was used to calculate HRV indices.
-Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings (63,64).
+Raw, unfiltered IBI recordings were exported from the Polar Flow web service as a space delimited .txt file and imported into R (62) where the \emph{RHRV} package (64) was used to calculate HRV indices.
+Raw IBI data was filtered using an adaptive threshold filter, and the first 3-minutes and last 2-minutes of each recording were discarded, leaving a 5-minutes window that was used to calculate the root mean square of successive differences (RMSSD) in RR intervals following the recommendations for short term IBI recordings (65,66).
 RMSSD was calculated as:
@@
 To assess muscle stiffness, participants lay horizontally on a massage bed and four testing sites were identified on each lower limb.
-The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site (65).
+The muscle belly of the rectus femoris (RF), biceps femoris (BF), medial gastrocnemius (MG) and lateral gastrocnemius (LG) sites were identified using a visual-palpatory technique to determine the exact location of each site (67).
 The visual-palpatory technique required the participant to contract the target muscle to aid the lead researcher to visually identify the muscle.
@@
 A constant pre-load of 0.18 N was applied for initial compression of subcutaneous tissues.
-The device was programmed to deliver five consecutive impulses, separated by a one-second interval (66).
+The device was programmed to deliver five consecutive impulses, separated by a one-second interval (68).
 For each impulse, the device computed passive stiffness values, with the median of the five values being saved by the device for further analysis.
@@
 Sets of measurements above 3\% were measured again to ensure data reliability.
-Measurements were uploaded using MyotonPRO software and imported in R (60) for further analysis.
+Measurements were uploaded using MyotonPRO software and imported in R (62) for further analysis.
 For each participant, the sum of all eight testing sites was calculated to provide a total lower extremity stiffness score and was used for further analysis
@@
 Each completed mBESS protocol took approximately four-minutes.
-Only one trial was performed to avoid familiarisation effects across the repeated measurement (67).
+Only one trial was performed to avoid familiarisation effects across the repeated measurement (69).
 The video recordings for each participant were imported from the recording equipment (Sony DSC-RX10) and the lead researcher scored each trial using the error identification criteria.
@@
 
-Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury.
-A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way (38).
-A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter (68).
-However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual (38).
-The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN (69).
+Two Bayesian Networks (BN) were used to explore the relationships between the psychosocial measures, physiological markers of stress and sports injury.
+A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way (40).
+A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter (70).
+However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of causal (40).
+The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN (71).
 In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.
@@
 For example, a network with 10 variables each with two levels has 100 (10\^{}2) possible parameter combinations, however the same network with three levels has 1000 (10\^{}3) possible parameter combinations, the latter being significantly more computationally expensive.
-Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio (38).
-Typically, no more than three levels have been used in Bayesian networks in the sports injury literature (39,70)
+Using a larger number of splits in the data also comes at a cost of reducing the statistical power in detecting probabilistic associations, and reduces the precision of parameter estimates for the probabilistic associations that are detected because it reduces the sample-size-to-parameters ratio (40).
+Typically, no more than three levels have been used in Bayesian networks in the sports injury literature (41,72)
 
@@
   A query can be submitted to an expert (in this case, the network is the expert) to get an opinion, the expert then updates the querier's beliefs accordingly.
-Widely used texts on Bayesian Network analysis (71) have widely adopted the terminology in favour of that used in traditional statistics.} and obtaining the posterior probabilities of a particular node for a given query.
-The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (38).
+  Widely used texts on Bayesian Network analysis (73) have widely adopted the terminology in favour of that used in traditional statistics.} and obtaining the posterior probabilities of a particular node for a given query.
+The posterior distribution can be obtained by \(Pr(X|E,B) = Pr(X|E,G,\Theta)\), where the learned network \(B\) with structure \(G\) and parameters \(\Theta\), are investigated with new evidence \(E\) using the information in \(B\) (40).
 When a network contains many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network.
@@
 
-BNs have recently been used in the sport psychology literature (39,70,72) and offer several benefits over traditional statistical analysis.
-For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (37,39).
-Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (73).
+BNs have recently been used in the sport psychology literature (41,72,74) and offer several benefits over traditional statistical analysis.
+For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature (39,41).
+Furthermore, the structure of a network can be obtained from both empirical data \emph{and} prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data (75).
 In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure.
-Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (74,75).
+Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure (76,77).
 The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.
@@
 The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random.
-A missing rate of 15-20\% has been reported to be common in psychological studies, and several techniques are available to handle missing values (76,77).
-In the current study, the \emph{caret} R package (78) was used to impute the missing values.
+A missing rate of 15-20\% has been reported to be common in psychological studies, and several techniques are available to handle missing values (78,79).
+In the current study, the \emph{caret} R package (80) was used to impute the missing values.
 A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset.
-The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation (78).
+The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation (80).
 
@@
 Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).
+In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).
 
@@
 Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
-Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (38).
-To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (79).
+Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (40).
+To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (81).
 In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
@@
 Finally, arc direction was restricted between specific nodes within the explanatory variables.
-Arcs from clevel \(\rightarrow\) gender, nlebase \(\rightarrow\) gender and nlebase \(\rightarrow\) ind\_team were included in the blacklist, as arcs in these directions did not make logical sense.
+Arcs from competitive level \(\rightarrow\) gender, baseline negative life events (NLE) \(\rightarrow\) gender and baseline NLE \(\rightarrow\) sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
 All subsequent models used the same blacklist.
@@
 Prior to the final network structure presented in the results section, several structures with different combinations of variables were explored.
-Networks were learned using a Tabu search algorithm (80) and BIC was used to compare different models.
-A higher BIC value indicates the structure of a DAG is a better fit to the observed data (38).
+Networks were learned using a Tabu search algorithm (82) and BIC was used to compare different models.
+A higher BIC value indicates the structure of a DAG is a better fit to the observed data (40).
 BIC values for each combination of variables of interest are reported as the combination of variables with the highest BIC value, followed by the relative scores of the other variables in the model.
 
-Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (nleleg = -4,242.76, tlelg only = -4,326.39, tlelg and nlelg = -4,459.23).
-Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg\_1 and injured\_1 and nlelg\_2 and injured\_2 in the final network structure.
+Initially, both negative life events (NLE) and total life events (TLE) were included in the network structure, however, the network score was improved when only NLE or TLE was included (NLE = -4,242.76, TLE only = -4,326.39, TLE and NLE = -4,459.23).
+Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
+Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
 TLE score was not included in the final structure.
 
-The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network.
+The subscales representing the Behavioural Activation System (Reward Reactivity {[}RR{]}, Reward Interest {[}RI{]}, Goal Drive Persistence {[}GDP{]} and Impulsivity {[}I{]}) showed limited connection to other variables in the network.
 Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
@@
 \hline
-nlelg\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-nlelg\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-nlelg\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+NLE\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+NLE\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+NLE\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+TLE\_1 & Log TLE at TP 1 & 1.79-3.4 (Low) & >3.4-4.88 (High)\\
+\hline
+TLE\_2 & Log TLE at TP 2 & 1.79-3.74 (Low) & >3.74-5.42 (High)\\
+\hline
+TLE\_3 & Log TLE at TP 3 & 1.79-3.81 (Low) & >3.81-5.18 (High)\\
 \hline
@@
 To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network.
-The final network structure was obtained by averaging 1000 bootstrapped models (81) to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model (39).
+The final network structure was obtained by averaging 1000 bootstrapped models (83) to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model (41).
 Arcs that were present in at least 30\% of the models were included in the averaged model.
@@
 An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks.
-In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (82).
-Arcs below 0.3 were not included in the model. The full Table of arc strengths for the first and second network are available in Supplementary Table 1 and Supplementary Table 3 respectively.
+In the respective study arcs above 0.5 were considered ``significant'' with arcs below 0.5 and above 0.3 ``non-significant'' (84).
+Arcs below 0.3 were not included in the model.
+The full table of arc strengths for the first and second network are available in Supplementary Table 2 and Supplementary Table 3 respectively.
 
@@
 The query allows the odds of a particular node state (e.g., injured\_1 = ``injured'') to be calculated based on the new evidence.
-CPQs were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
-The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query (38).
+CPQ were performed using a likelihood weighting approach; a form of importance sampling where random observations are generated from the probability distribution in such a way that all observations match the evidence given in the query.
+The algorithm then re-weights each observation based on the evidence when computing the conditional probability for the query (40).
 Inference was first performed on arcs that had a strength greater than 0.5 between the explanatory variables and independent variables and between different independent variables in the network.
 Of particular interest were the variables that were connected to ``injured'' nodes, which were examined in the network using the Markov blanket of ``injured\_1'' and ``injured\_2''.
-A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network (72).
-CPQ's were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the ``injured'' state.
+A Markov blanket contains all the nodes that make the node of interest conditionally independent from the rest of the network (74).
+CPQ were used to determine what effect the variables in the Markov blanket of injured nodes had on the probability of the injured node being in the ``injured'' state.
 
@@
 Similar to the first network, blacklists were used to prevent arcs from independent variables \(\rightarrow\) explanatory variables.
-In addition, ``nlebase'' variable was dropped from the list of explanatory variables to allow the \emph{changes} in negative life events to be the only life event variable in the network.
-The final network was obtained using the same approach as the first network.
-
-\hypertarget{network-inference-1}{%
-\paragraph{Network Inference}\label{network-inference-1}}
+In addition, baseline negative life events was dropped from the list of explanatory variables to allow the \emph{changes} in negative life events to be the only life event variable in the network.
+The final network was obtained using the same approach as the first network.Network Inference
 
@@
 The Markov blanket of the ``injured'' node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-Initial visual inspection of the network structure also revealed arcs between BIS \(\rightarrow\) FFFS and HRV \(\rightarrow\) FFFS.
+Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) and HRV \(\rightarrow\) FFFS.
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
-The samples were then examined using a Bayesian linear regression models with the \emph{brms} R package (83) to determine the relationship between these nodes.
+The samples were then examined using a Bayesian linear regression models with the \emph{brms} R package (85) to determine the relationship between these nodes.
 Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.
@@
 Table \ref{tab:table3} shows the number and percentage of injury types sustained by both male and female participants.
+An additional breakdown of injury by sport, type (acute or chronic) and injury location is available in Supplementary tables 4 and 5.
 
@@
 \hline
-Other & 8 (21) & 3 (21) & 5 (6) & 11 (30)\\
-\hline
-\multicolumn{5}{l}{\textit{Note: }}\\
-\multicolumn{5}{l}{Other included bone, skin and brain injuries.}\\
+Other (bone, brain and skin) & 8 (21) & 3 (21) & 5 (6) & 11 (30)\\
+\hline
 \end{tabular}
@@
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
-The ``ind\_team'' node had strong arcs to hours (0.94) and nlebase (0.78).
+The ``individual or team sport'' node had strong arcs to hours per week (0.94) and baseline negative life events (NLE) (0.78).
 Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
 Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
-The arc from clevel \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
+The arc from competitive level \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
 High gender-based connections were reported for the arcs from gender \(\rightarrow\) stiffness\_1 (0.71) and gender \(\rightarrow\) stiffness\_2 (0.43), with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
 Irrespective of sport or gender, strong connections were found between explanatory variables.
-The arc from nlebase \(\rightarrow\) RI\_1 had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.
+The arc from baseline NLE (nlebase) \(\rightarrow\) Reward Interest (RI\_1) had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.
 
 The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability.
-The arc from pi \(\rightarrow\) stiffness\_1 was 0.57 with athletes who reported an injury in the preceding 12 months being more likely to have ``High'' (0.65) compared to ``Low'' (0.35). stiffness.
-Strong arcs were present between BIS\_1 \(\rightarrow\) FFFS\_1 (0.98) and BIS\_2 \(\rightarrow\) FFFS\_2 (0.74) were identified.
+The arc from previous injury \(\rightarrow\) stiffness\_1 was 0.57 with athletes who reported an injury in the preceding 12 months being more likely to have ``High'' (0.65) compared to ``Low'' (0.35).
+stiffness.
+Strong arcs were present between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS; BIS\_1 \(\rightarrow\) FFFS\_1 = 0.98, BIS\_2 \(\rightarrow\) FFFS\_2 = 0.74).
 In both instances, ``High'' FFFS was more likely when BIS was ``High'' (0.64 for \_1, 0.61 for \_2) compared to ``Low'' (0.33 for \_1, 0.37 for \_2).
-The arc between nlelg \(\rightarrow\) BIS had a strength of 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.37 nlelg\_2 and BIS\_2.
-``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55 for nlelg\_1 \(\rightarrow\) BIS\_1, and 0.38 to 0.58 for nlelg\_2 \(\rightarrow\) BIS\_2.
+The arc between NLE \(\rightarrow\) BIS had a strength of 0.55 for NLE\_1 \(\rightarrow\) BIS\_1, and 0.37 NLE\_2 and BIS\_2.
+``Low'' negative life events increased the probability of BIS being in the ``High'' state from 0.33 to 0.55 for NLE\_1 \(\rightarrow\) BIS\_1, and 0.38 to 0.58 for NLE\_2 \(\rightarrow\) BIS\_2.
 
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg\_1), muscle stiffness (stiffness\_1), current competitive level (clevel) and balance (balance\_1).
-The arc between nlelg\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week, negative life events (NLE\_1), muscle stiffness (stiffness\_1), current competitive level and balance (balance\_1).
+The arc between NLE\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
 
@@
 \hline
-nlelg\_1 & 0.24 & 0.26\\
+NLE\_1 & 0.24 & 0.26\\
 \hline
@@
 \hline
-\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{nlelg\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
+\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{NLE\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
 \hline
@@
 
-The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS\_1, stiffness\_2, balance\_2 and hrv\_2 (Table \ref{tab:table6}).
+The Markov blanket for injured\_2 is shown in Figure \ref{fig:fig3} and contained gender, Fight-Flight-Freeze System (FFFS\_1), stiffness\_2, balance\_2 and heart rate variability (HRV\_2; Table \ref{tab:table6}).
 The arc between stiffness\_2 \(\rightarrow\) injured\_2 was comparable to the arc between stiffness\_1 \(\rightarrow\) injured\_1.
-Very weak arcs (0.3) between injured\_2 \(\rightarrow\) balance\_2 and injured\_2 \(\rightarrow\) hrv\_2 were also present in the Markov blanket for injured\_2.
+Very weak arcs (0.3) between injured\_2 \(\rightarrow\) balance\_2 and injured\_2 \(\rightarrow\) HRV\_2 were also present in the Markov blanket for injured\_2.
 Similar to injured\_1, stiffness\_2 doubled the probability of injured\_2 being ``injured'' from 0.13 in the ``Low'' state to 0.27 in the ``High'' state.
@@
 \hline
-hrv\_2 & 0.26 & 0.17\\
-\hline
-nlelg\_2 & 0.23 & 0.19\\
+HRV\_2 & 0.26 & 0.17\\
+\hline
+NLE\_2 & 0.23 & 0.19\\
 \hline
@@
 \hline
-\textbf{Probability} & \textbf{FFFS\_1} & \textbf{nlelg\_2} & \textbf{stiffness\_2} & \textbf{hrv\_2} & \textbf{balance\_2}\\
+\textbf{Probability} & \textbf{FFFS\_1} & \textbf{NLE\_2} & \textbf{stiffness\_2} & \textbf{HRV\_2} & \textbf{balance\_2}\\
 \hline
@@
 
-The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
-An arc between BIS \(\rightarrow\) FFFS with strength 1.00 was present in the network.
-Arcs between clevel \(\rightarrow\) BIS and gender \(\rightarrow\) stiffness had a strengths of 0.60 and 0.56 respectively
-Similarly, the arc between hrv \(\rightarrow\) FFFS was 0.58.
-The arcs between BIS \(\rightarrow\) FFFS and hrv \(\rightarrow\) FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
-A separate linear regression examined the interaction between BIS and hrv to be examined.
+The second network structure (Figure \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/ An arc between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) with strength 1.00 was present in the network.
+Arcs between competitive level \(\rightarrow\) BIS and gender \(\rightarrow\) stiffness had a strengths of 0.60 and 0.56 respectively Similarly, the arc between HRV \(\rightarrow\) FFFS was 0.58.
+The arcs between BIS \(\rightarrow\) FFFS and HRV \(\rightarrow\) FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
+A separate linear regression examined the interaction between BIS and HRV to be examined.
 
@@
 Results from the Bayesian linear regression model are presented in Table \ref{tab:table8} and include 95\% credible intervals (CrI).
-Increases in BIS were associated with increases in FFFS (\(b\) = -0.19, 95\% CrI = {[}-0.25, -0.13{]}), whereas positive changes in hrv where associated with decreased changes in FFFS (\(b\) = 0.41, 95\% CrI = {[}0.36, 0.47{]}).
-There was no clear interaction between hrv and BIS (\(b\) = -0.02, 95\% CrI = {[}-0.08, 0.03{]}).
+Increases in BIS were associated with increases in FFFS (\(b\) = -0.19, 95\% CrI = {[}-0.25, -0.13{]}), whereas positive changes in HRV where associated with decreased changes in FFFS (\(b\) = 0.41, 95\% CrI = {[}0.36, 0.47{]}).
+There was no clear interaction between HRV and BIS (\(b\) = -0.02, 95\% CrI = {[}-0.08, 0.03{]}).
 
@@
 
-\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.}
+\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and HRV.}
 \centering
@@
 \hline
-hrv & -0.19 & 0.03 & [-0.25, -0.13]\\
-\hline
-BIS:hrv & -0.02 & 0.03 & [-0.08, 0.03]\\
+HRV & -0.19 & 0.03 & [-0.25, -0.13]\\
+\hline
+BIS:HRV & -0.02 & 0.03 & [-0.08, 0.03]\\
 \hline
@@
 
-The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week, stiffness and nlec (Figure \ref{fig:fig5}).
-For stiffness and nlec the values in the nodes represent the standardised change between time point.
-Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
+The Markov blanket for the ``injured'' node contained previous injury, gender, training hours per week and stiffness and negative life events (NLE; Figure \ref{fig:fig5}).
+For stiffness and NLE, the values in the nodes represent the standardised change between time point.
+Combinations of NLE and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in Table \ref{tab:table9}.
 Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress.
-Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (0.71) over the duration of the study.
-With stiffness held at the mean change, the probability of ``injured'' rose notably from 0.35 to 0.64 with nlec at 1 SD below an 1 SD above respectively.
+Changes in both NLE and stiffness of 1SD above the mean change resulted in a high probability of being injured (0.71) over the duration of the study.
+With stiffness held at the mean change, the probability of ``injured'' rose notably from 0.35 to 0.64 with NLE at 1 SD below an 1 SD above respectively.
 
@@
 
-The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.\\
+The study investigated the multifaceted interactions of stress-related variables and injury occurrence using the stress-injury frameworks presented by Williams and Andersen (1998) and Appaneal and Perna (2014) and a prospective, repeated measures design applied to a large cohort of athletes.
 Relationships between stress-related psychosocial and physiological factors and injury were investigated using two BN structures; the first was a two-time Bayesian Network that investigated the relationships between variables across time points in the study, and the second network used differential equations to model the changes in variables between time points.
-The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychological and physiological factors that may occur preceding an injury.
+The latter facilitated the development of new insights into the interactions of stress-related factors with injury occurrence, exploring changes in both psychosocial and physiological factors that may occur preceding an injury.
 
@@
 The presence of these factors at both injured nodes indicated that the interaction of these variables is important for determining an athlete's risk of injury.
-In the second network, the highest probability of injury was observed when changes in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychological and physiological stress-related factors may combine additively to increase the risk of injury (28).
+In the second network, the highest probability of injury was observed when increases in stiffness and negative life events were both greater than average, indicating that the combination of changes in psychosocial and physiological stress-related factors may combine additively to increase the risk of injury (28).
 
@@
 In contrast, athletes with high stiffness and good balance were less likely to be injured, suggesting that improved postural stability may counteract the potential harmful effects of high levels of muscle stiffness.
-Several studies have identified how balance (35,84) and muscle stiffness (85,86) are separately related to injury.
+Several studies have identified how balance (36,86) and muscle stiffness (34,87) are separately related to injury.
 The BN structures examined in this study provided insight into these two stress-related factors and their relation to injury occurrence.
@@
 At both injured nodes in the first network, balance was linked to injury.
-Despite the weak arc strength at both injured nodes, a ``High'' balance score, which is considered indicative of impaired postural stability (35), was found to increase the probability of injury.
-This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (35,58,84).
-Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (87).
+Despite the weak arc strength at both injured nodes, a ``High'' balance score, which is considered indicative of impaired postural stability (86), was found to increase the probability of injury.
+This finding is consistent with previous research that has reported an association between decrements in postural stability and increased injury risk (36,60,86).
+Postural stability is often used as an indicator of athlete performance level, with higher level athletes demonstrating better postural stability over their lower level counterparts (88).
 Athletes who competed at a higher level were also more likely to have good balance (``Low'' balance), compared to their lower level counterparts.
-These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (88).
+These findings suggest that better postural stability is associated with both a higher level of performance and a lower probability of sustaining an injury, reinforcing the importance of postural stability as a feature of athletic training programmes designed to prepare athletes for the demands of high intensity training and competition (89).
 
@@
 Although the absolute score would be ``Low'', the relative increase between time points may have been caused by a significant event in the athlete's life that had a considerable psychological and physiological effect (28).
-Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed (89).
-
-During the initial network structure development, no arcs between the negative life event and injured nodes were found by the Tabu search algorithm.
+Future research should therefore consider study designs and analyses that enable relative changes in an individual athlete's life events to be assessed (90).
+
 The majority of research has however, consistently identified major life events, particularly those events with a negative valence, as the strongest predictor of injury in Williams and Andersen's model (1).
 During the initial network structure development, no arcs between the negative life event nodes and injured nodes were found by the Tabu search algorithm.
-Given the reported reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
+Given the reported association between negative life events and injury, an arc was fixed between these variables to allow this relationship to be examined more closely.
 When negative life events were ``High'' the probability of injury showed a negligible change at the injured\_1 node and decreased by -0.04 at the injured\_2 node.
 One possible explanation for these findings may be due to the use of the LESCA questionnaire in a repeated measures design.
-In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months (44).
+In the original LESCA, participants are asked to report major life events that have occurred over the previous 12-months (46).
 In this study, athletes completed the LESCA at three time points with an approximate four-month interval after baseline.
@@
 For example, at the second and third time points, 26\% of participants reported 0 negative life events for the preceding four-month period.
-Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events (90).
+Simply, it may be that the items on the LESCA are less suitable for repeated measurements with durations shorter than the advocated 12-months than a measure that captures minor life events (91).
 
 Williams and Andersen's (14) model proposed a number of coping resources that were either directly related to injury or moderated the relationship between life stress and injury occurrence; for example, general coping strategies (e.g., good sleeping habits and self-care), social support systems and stress management skills.
-Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (36,91,92).
+Although coping was not measured in the current study, several studies have found high levels of social support can reduce the risk of injury (38,92,93).
 The lack of association between major/negative life events and injury reported here may be attributed to athletes in this study having the necessary coping resources to mitigate against the effects of any negative life event stress they experienced.
@@
 Despite the uncertainty regarding the relationship between injury and heart rate variability (HRV) in the first network, ``Low'' HRV increased the probability of injury from 0.17 (``High'' HRV) to 0.26 (``Low'' HRV).
-This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (32,33,93).
+This finding is consistent with previous research that has found reduced HRV indices to be indicative of illness or maladaptation to training due to decreased parasymapthic activity, which often precedes injury (37,94,95).
 An arc between FFFS\_1 and injured\_2 (arc strength = 0.40) was also observed in the first network, where the risk of injury was increased from 0.13 to 0.29 with FFFS in the ``High'' and ``Low'' states respectively.
@@
 One possible explanation for this finding could be that those athletes who reported ``Low'' FFFS score were less fearful, and may therefore engage in more risk taking behaviours, increasing the probability of injury.
-The RST theory proposes that higher levels of FFFS increase avoidance motivation (94), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
-The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (95).
+The RST theory proposes that higher levels of FFFS increase avoidance motivation (96), and therefore ``High'' FFFS may have acted as a deterrent from taking risks while training and competing, reducing exposure to situations that could have resulted in injury.
+The RST theory further proposes that the combination of high BIS and high FFFS is likely to result in a more anxious disposition due to high levels of avoidance and high goal conflict characterised by high levels of FFFS and BIS (97).
 The first network reported an association between ``High'' FFFS and ``High'' BIS, while the second network found an association between increases in FFFS and increases in BIS.
-High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD (96,97).
-This association along with the proposed actions of the RST theory (94) provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network.
+High levels of anxiety and anticipation of stressful situations have been linked to reductions in HRV indices including RMSSD (98,99).
+This association along with the proposed actions of the RST theory (96) provides a potential explanation for the negative relationship between FFFS and HRV identified in the second network.
 To elaborate, high levels of BIS are proposed to be the result of goal conflict, an example of which would be simultaneous triggering of the FFFS (avoidance) and BAS (approach) systems.
-The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (94).
+The goal conflict is likely to elicit a physiological response (e.g., decreased HRV) in preparation to engage in the required behaviour to resolve the goal conflict (96).
 In the present study, however, the role of BAS was limited, as evidenced by the initial network structures in which the BAS had limited connectivity with other components of the network.
@@
 This study had a number of strengths.
-A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury (36).
+A major critique of the sport injury literature has been the use of only one wave of measurement that may not be reflective of the dynamic nature of the variables that are associated with injury (38).
 The longitudinal repeated measures design of the current study allowed changes over time and between time points to be captured and explored.
 Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (37) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (39) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
 The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
@@
 
-While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (39), a number of assumptions were made that potentially limited the approach employed in the study.
+While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (41), a number of assumptions were made that potentially limited the approach employed in the study.
 Firstly, the choice was made to binarise variables in the first network so only ``Low'' and ``High'' states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner (98) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner (100) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
 For example, the meaning of a ``Low'' and ``High'' value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
@@
 However, a reduction in the sensitivity of some of these measures may have inhibited our ability to detect more subtle variation in the athletes' responses.
-For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis (99).
+For example, a more sensitive measure of postural stability may have been achieved with the use of a force plate, which is considered the gold standard to provide detailed data and enable a more fine-grained analysis (101).
 However, the video-capture approach employed in the study ensured an accessible, non-invasive and readily applied method of capturing the respective measure.
 
-The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychological and physiological perspective.
+The ability to capture large sample sizes of injured athletes has been recognised as a significant challenge in sports injury research (102) and affects the success to which stress-injury interactions can be identified and understood.
+This is further exacerbated in studies that employ longitudinal prospective repeated measures designs where high levels of participant retention are required.
+Consistent with Williams and Andersen's (1998) stress-injury model, we therefore employed a global definition of injury as the primary outcome measure, to optimize sample size of injury occurrences across the repeated measures prospective study design.
+As suggested by Ruddy et al.~(2019), a larger number of observations and injury events is needed to improve the ability to identify and to make more meaningful predictions.
+Further conventional injury exposure measures (sport type, training load) were subsequently discretised or reduced for the data capture, which caused some loss of information and has been suggested to limit the ability to capture risk profiles (103).
+Using pooled or discretised definitions for some measures hindered a detailed causal-effect insight into injury-specific risk factors.
+However, the more holistic, interdisciplinary approach adopted in this study was benefited by the ability to gain and substantiate a more complete picture of the complex, multifaceted stress-injury interactions that exist in sport.
+
+The multifactorial, interdisciplinary approach employed in the study required the selection of stress-related measures derived from a psychosocial and physiological perspective.
 The variables used in the present study were not definitive.
@@
 
-To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Baysian Network analytical techniques, Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
+Finally, the findings of this study have important practical implications for athletes, coaches, and clinicians in relation to the additive and interactive effects of multiple sources of stress on injury occurrence.
+Specifically, the study evidenced a combined effect of psychosocial and physiological stress-related factors that could increase the probability of injury to a greater extent than any isolated factor.
+When assessing an athlete's training plan, readiness to engage in, and recovery from, training, coaches and clinicians should employ a risk profile that integrates multifaceted sources of stress.
+For example, in addition to monitoring training loads and using tools to determine an athlete's physiological status, coaches need to also consider an athlete's psychological state.
+In particular, when an athlete is facing significant life event stress, adjusted training intensity and volume may be necessitated to support athlete coping with the additional duress and to subsequently safeguard optimal health and well-being.
+In essence, injury risk is exacerbated when an athlete is experiencing psychological stress due to exposure to negative life events and exhibiting physiological responses associated with an increased injury potential.
+The identification of such a ``high risk'' profile is subsequently important in helping to monitor and reduce injury risk for athletes.
+For example, while high muscle stiffness is important for optimal performance (104), this study demonstrated it can heighten injury risk, which is likely to be exacerbated when accompanied by the experience of negative life events by an athlete.
+In order to understand how an athlete's injury risk may increase over time, it is important to acknowledge the breadth and interaction of stress-related factors that could heighten susceptibility and be receptive to training and life experience changes.
+
+To summarise, this study provided novel insights into the multifaceted nature of the stress-injury relationship using a novel interdisciplinary approach coupled with an advanced Bayesian Network analytical techniques.
+Muscle stiffness and increases in negative life event stress were identified as strong predictors of injury within the multifaceted athlete cohort, while other factors including personality characteristics and postural stability were also found to contribute to the probability of injury occurrence.
 Future research combining a repeated measures approach and complex analyses of the interactions between multifaceted stress-related measures are advocated to enhance understanding of the injury occurrence in sport.
@@
 
-The authors declare that the research was conducted in the absence of any
-commercial or financial relationships that could be construed as a potential
-conflict of interest.
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
 
@@
 
-\leavevmode\hypertarget{ref-Perna1997}{}%
+\leavevmode\hypertarget{ref-VanOckenburg2015a}{}%
 30.
-Perna F, Schneiderman N, LaPerriere A.
-Psychological stress, exercise and immunity.
-\emph{International Journal of Sports Medicine} (1997) \textbf{18}:78--83.
-doi:\href{https://doi.org/10.1055/s-2007-972703}{10.1055/s-2007-972703}
-
-\leavevmode\hypertarget{ref-Perna2003}{}%
+Ockenburg SL van, Tak LM, Bakker SJL, Gans ROB, Jonge P de, Rosmalen JGM.
+Effects of adverse life events on heart rate variability, cortisol, and C-reactive protein.
+\emph{Acta Psychiatrica Scandinavica} (2015) \textbf{131}:40--50.
+doi:\href{https://doi.org/10.1111/acps.12286}{10.1111/acps.12286}
+
+\leavevmode\hypertarget{ref-Gervasi2017}{}%
 31.
-Perna FM, Antoni MH, Baum A, Gordon P, Schneiderman N.
-Cognitive behavioral stress management effects on injury and illness among competitive athletes: A randomized clinical trial.
-\emph{Annals of Behavioral Medicine} (2003) \textbf{25}:66--73.
-Available at: \url{https://www.scopus.com/inward/record.uri?eid=2-s2.0-0037262738\%7B/\&\%7DpartnerID=40\%7B/\&\%7Dmd5=37332ebe6a962abf5178e51d6ebaedb7}
-
-\leavevmode\hypertarget{ref-Bellenger2016}{}%
+Gervasi M, Sisti D, Amatori S, Andreazza M, Benelli P, Sestili P, et al.
+Muscular viscoelastic characteristics of athletes participating in the European Master Indoor Athletics Championship.
+\emph{European Journal of Applied Physiology} (2017) \textbf{117}:1739--1746.
+doi:\href{https://doi.org/10.1007/s00421-017-3668-z}{10.1007/s00421-017-3668-z}
+
+\leavevmode\hypertarget{ref-Bailey2013}{}%
 32.
-Bellenger CR, Fuller JT, Thomson RL, Davison K, Robertson EY, Buckley JD.
-Monitoring athletic training status through autonomic heart rate regulation: A systematic review and meta-analysis.
-\emph{Sports Medicine} (2016) \textbf{46}:1461--1486.
-doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}
+Bailey L, Dinesh S, Warner M, Stokes M, Samuel D.
+Parameters representing muscle tone, elasticity and stiffness of biceps brachii in healthy older males: Symmetry and within-session reliability using the MyotonPRO.
+\emph{Journal of Neurological Disorders} (2013) \textbf{1}:1--7.
+doi:\href{https://doi.org/10.4172/2329-6895.1000116}{10.4172/2329-6895.1000116}
+
+\leavevmode\hypertarget{ref-Rath2017}{}%
+33.
+Rath R, Wade MG.
+The two faces of postural control in older adults: Stability and function.
+\emph{EBioMedicine} (2017) \textbf{21}:5--6.
+doi:\href{https://doi.org/10.1016/j.ebiom.2017.03.030}{10.1016/j.ebiom.2017.03.030}
+
+\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
+34.
+Pickering-Rodriguez EC, Watsford ML, Bower RG, Murphy AJ.
+The relationship between lower body stiffness and injury incidence in female netballers.
+\emph{Sports Biomechanics} (2017) \textbf{16}:361--373.
+doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}
+
+\leavevmode\hypertarget{ref-Plews2012}{}%
+35.
+Plews DJ, Laursen PB, Kilding AE, Buchheit M.
+Heart rate variability in elite triathletes, is variation in variability the key to effective training?
+A case comparison.
+\emph{European Journal of Applied Physiology} (2012) \textbf{112}:3729--3741.
+doi:\href{https://doi.org/10.1007/s00421-012-2354-4}{10.1007/s00421-012-2354-4}
+
+\leavevmode\hypertarget{ref-Trojian2006}{}%
+36.
+Trojian TH, McKeag DB.
+Single leg balance test to identify risk of ankle sprains.
+\emph{British Journal of Sports Medicine} (2006) \textbf{40}:610--613.
+doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}
 
 \leavevmode\hypertarget{ref-Williams2017}{}%
-33.
+37.
 Williams S, Booton T, Watson M, Rowland D, Altini M.
@@
 
-\leavevmode\hypertarget{ref-Pruyn2015}{}%
-34.
-Pruyn EC, Watsford ML, Murphy AJ.
-Differences in lower-body stiffness between levels of netball competition.
-\emph{Journal of Strength and Conditioning Research} (2015) \textbf{29}:1197--1202.
-doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}
-
-\leavevmode\hypertarget{ref-Romero-Franco2014}{}%
-35.
-Romero-Franco N, Gallego-Izquierdo T, Martínez-López EJ, Hita-Contreras F, Osuna-Pére, Catalina M, Martínez-Amat A.
-Postural stability and subsequent sports injuries during indoor season of athletes.
-\emph{Journal of Physical Therapy Science} (2014) \textbf{26}:683--687.
-doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}
-
 \leavevmode\hypertarget{ref-Johnson2014}{}%
-36.
+38.
 Johnson U, Tranaeus U, Ivarsson A.
@@
 \leavevmode\hypertarget{ref-Bittencourt2016}{}%
-37.
+39.
 Bittencourt NFN, Meeuwisse WH, Mendonça LD, Nettel-Aguirre A, Ocarino JM, Fonseca ST.
@@
 \leavevmode\hypertarget{ref-Scutari2014}{}%
-38.
+40.
 Scutari M, Denis J-B.
@@
 \leavevmode\hypertarget{ref-Olmedilla2018}{}%
-39.
+41.
 Olmedilla A, Rubio VJ, Fuster-Parra P, Pujals C, García-Mas A.
@@
 \leavevmode\hypertarget{ref-Fuller2006}{}%
-40.
+42.
 Fuller CW, Ekstrand J, Junge A, Andersen TE, Bahr R, Dvorak J, et al.
@@
 \leavevmode\hypertarget{ref-Fuller2007b}{}%
-41.
+43.
 Fuller CW, Molloy MG, Bagate C, Bahr R, Brooks JHM, Donson H, et al.
@@
 \leavevmode\hypertarget{ref-Timpka2014}{}%
-42.
+44.
 Timpka T, Alonso JM, Jacobsson J, Junge A, Branco P, Clarsen B, et al.
@@
 \leavevmode\hypertarget{ref-Appaneal2009}{}%
-43.
+45.
 Appaneal RN, Levine BR, Perna FM, Roh JL.
@@
 \leavevmode\hypertarget{ref-Petrie1992}{}%
-44.
+46.
 Petrie TA.
@@
 \leavevmode\hypertarget{ref-Fornell1981}{}%
-45.
+47.
 Fornell C, Larcker DF.
@@
 \leavevmode\hypertarget{ref-Corr2016c}{}%
-46.
+48.
 Corr PJ, Cooper AJ.
@@
 \leavevmode\hypertarget{ref-Giles2016}{}%
-47.
+49.
 Giles D, Draper N, Neil W.
@@
 \leavevmode\hypertarget{ref-Caminal2018}{}%
-48.
+50.
 Caminal P, Sola F, Gomis P, Guasch E, Perera A, Soriano N, et al.
@@
 \leavevmode\hypertarget{ref-Plews2017}{}%
-49.
+51.
 Plews DJ, Scott B, Altini M, Wood M, Kilding AE, Laursen PB.
@@
 \leavevmode\hypertarget{ref-Agyapong-Badu2016}{}%
-50.
+52.
 Agyapong-Badu S, Warner M, Samuel D, Stokes M.
@@
 \leavevmode\hypertarget{ref-Chuang2013}{}%
-51.
+53.
 Chuang LL, Lin KC, Wu CY, Chang CW, Chen HC, Yin HP, et al.
@@
 \leavevmode\hypertarget{ref-Pruyn2016}{}%
-52.
+54.
 Pruyn EC, Watsford ML, Murphy AJ.
@@
 \leavevmode\hypertarget{ref-Nair2014}{}%
-53.
+55.
 Nair K, Dougherty J, Schaefer E, Kelly J, Masi A.
@@
 \leavevmode\hypertarget{ref-Aird2012}{}%
-54.
+56.
 Aird L, Samuel D, Stokes M.
@@
 \leavevmode\hypertarget{ref-Hunt2009}{}%
-55.
+57.
 Hunt TN, Ferrara MS, Bornstein RA, Baumgartner TA.
@@
 \leavevmode\hypertarget{ref-McCrory2013}{}%
-56.
+58.
 McCrory P, Meeuwisse WH, Aubry M, Cantu RC, Dvořák J, Echemendia RJ, et al.
@@
 \leavevmode\hypertarget{ref-VanCingel2017}{}%
-57.
+59.
 Cingel REH van, Hoogeboom TJ, Melick N van, Meddeler BM, Nijhuis-van der Sanden MWG.
@@
 \leavevmode\hypertarget{ref-Riemann1999d}{}%
-58.
+60.
 Riemann BL, Guskiewicz KM, Shields EW.
@@
 \leavevmode\hypertarget{ref-Koo2016}{}%
-59.
+61.
 Koo TK, Li MY.
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-60. R Core Team. \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing (2019). Available at: \url{https://www.r-project.org/}
+62. R Core Team. \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing (2019). Available at: \url{https://www.r-project.org/}
 
 \leavevmode\hypertarget{ref-Denver2007}{}%
-61.
+63.
 Denver JW, Reed SF, Porges SW.
@@
 \leavevmode\hypertarget{ref-Rodriguez-Linares2017}{}%
-62. Rodriguez-Linares L, Vila X, Lado MJ, Mendez A, Otero A, Garcia CA. \emph{RHRV: Heart rate variability analysis of ecg data}. (2019). Available at: \url{https://CRAN.R-project.org/package=RHRV}
+64. Rodriguez-Linares L, Vila X, Lado MJ, Mendez A, Otero A, Garcia CA. \emph{RHRV: Heart rate variability analysis of ecg data}. (2019). Available at: \url{https://CRAN.R-project.org/package=RHRV}
 
 \leavevmode\hypertarget{ref-Laborde2017}{}%
-63.
+65.
 Laborde S, Mosley E, Thayer JF.
@@
 \leavevmode\hypertarget{ref-Malik1996}{}%
-64.
+66.
 Malik M, Camm AJ, Bigger JT, Breithardt G, Cerutti S, Cohen RJ, et al.
@@
 \leavevmode\hypertarget{ref-Chuang2012}{}%
-65.
+67.
 Chuang LL, Wu CY, Lin KC.
@@
 \leavevmode\hypertarget{ref-Morgan2018}{}%
-66.
+68.
 Morgan GE, Martin R, Williams L, Pearce O, Morris K.
@@
 \leavevmode\hypertarget{ref-Valovich2003}{}%
-67.
+69.
 Valovich TC, Perrin DH, Gansneder BM.
@@
 \leavevmode\hypertarget{ref-Pearl1988}{}%
-68. Pearl J. \emph{Probabilistic reasoning in intelligent systems: Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann Publishers (1988).
+70. Pearl J. \emph{Probabilistic reasoning in intelligent systems: Networks of plausible inference}. San Francisco, CA: Morgan Kaufmann Publishers (1988).
 
 \leavevmode\hypertarget{ref-Chen2012}{}%
-69.
+71.
 Chen SH, Pollino CA.
@@
 \leavevmode\hypertarget{ref-Ruiz-Perez2019}{}%
-70.
+72.
 Ruiz-Pérez I, Ayala F, Miguel Puerta J, Elvira JLL, De Ste Croix M, Hernández-Sánchez S, et al.
@@
 \leavevmode\hypertarget{ref-Koller2009}{}%
-71.
+73.
 Koller D, Friedman N.
@@
 \leavevmode\hypertarget{ref-Fuster-Parra2017}{}%
-72.
+74.
 Fuster-Parra P, Vidal-Conti J, Borràs PA, Palou P.
@@
 \leavevmode\hypertarget{ref-Xiao-xuan2007}{}%
-73.
+75.
 Xiao-xuan H, Hui W, Shuo W.
@@
 \leavevmode\hypertarget{ref-Heckerman1995}{}%
-74.
+76.
 Heckerman D, Geiger D, Chickering DM.
@@
 \leavevmode\hypertarget{ref-Xu2015}{}%
-75.
+77.
 Xu JG, Zhao Y, Chen J, Han C.
@@
 \leavevmode\hypertarget{ref-Enders2003}{}%
-76.
+78.
 Enders CK.
@@
 \leavevmode\hypertarget{ref-Lang2014}{}%
-77.
+79.
 Lang KM, Jorgensen TD, Moore EWG, Little TD.
@@
 \leavevmode\hypertarget{ref-Kuhn2008}{}%
-78.
+80.
 Kuhn MC from JW, Weston S, Williams A, Keefer C, Engelhardt A.
@@
 \leavevmode\hypertarget{ref-Murphy2002}{}%
-79. Murphy K. Dynamic Bayesian networks: Representation, inference and learning. (2002)
+81. Murphy K. Dynamic Bayesian networks: Representation, inference and learning. (2002)
 
 \leavevmode\hypertarget{ref-Norvig2009}{}%
-80. Russell SJ, Norvig P. \emph{Artificial Intelligence: A Modern Approach}. 3rd ed. Prentice Hall (2009). Available at: \url{http://aima.cs.berkeley.edu/}
+82. Russell SJ, Norvig P. \emph{Artificial Intelligence: A Modern Approach}. 3rd ed. Prentice Hall (2009). Available at: \url{http://aima.cs.berkeley.edu/}
 
 \leavevmode\hypertarget{ref-Efron1993}{}%
-81.
+83.
 Efron B, Tibshirani RJ.
@@
 \leavevmode\hypertarget{ref-Scutari2013}{}%
-82.
+84.
 Scutari M, Nagarajan R.
@@
 \leavevmode\hypertarget{ref-Burkner2017a}{}%
-83.
+85.
 Bürkner PC.
@@
 
-\leavevmode\hypertarget{ref-Trojian2006}{}%
-84.
-Trojian TH, McKeag DB.
-Single leg balance test to identify risk of ankle sprains.
-\emph{British Journal of Sports Medicine} (2006) \textbf{40}:610--613.
-doi:\href{https://doi.org/10.1136/bjsm.2005.024356}{10.1136/bjsm.2005.024356}
+\leavevmode\hypertarget{ref-Romero-Franco2014}{}%
+86.
+Romero-Franco N, Gallego-Izquierdo T, Martínez-López EJ, Hita-Contreras F, Osuna-Pére, Catalina M, Martínez-Amat A.
+Postural stability and subsequent sports injuries during indoor season of athletes.
+\emph{Journal of Physical Therapy Science} (2014) \textbf{26}:683--687.
+doi:\href{https://doi.org/10.1589/jpts.26.683}{10.1589/jpts.26.683}
 
 \leavevmode\hypertarget{ref-Butler2003}{}%
-85.
+87.
 Butler RJ, Crowell HP, Davis IMC.
@@
 
-\leavevmode\hypertarget{ref-PickeringRodriguez2017}{}%
-86.
-Pickering-Rodriguez EC, Watsford ML, Bower RG, Murphy AJ.
-The relationship between lower body stiffness and injury incidence in female netballers.
-\emph{Sports Biomechanics} (2017) \textbf{16}:361--373.
-doi:\href{https://doi.org/10.1080/14763141.2017.1319970}{10.1080/14763141.2017.1319970}
-
 \leavevmode\hypertarget{ref-Paillard2006}{}%
-87.
+88.
 Paillard T, Noé F, Rivière T, Marion V, Montoya R, Dupui P.
@@
 \leavevmode\hypertarget{ref-Hrysomallis2011}{}%
-88.
+89.
 Hrysomallis C.
@@
 \leavevmode\hypertarget{ref-Ivarsson2014}{}%
-89.
+90.
 Ivarsson A, Johnson U, Lindwall M, Gustafsson H, Altemyr M.
@@
 \leavevmode\hypertarget{ref-Fawkner1999}{}%
-90.
+91.
 Fawkner HJ, McMurrary NE, Summers JJ.
@@
 \leavevmode\hypertarget{ref-Petrie1993}{}%
-91.
+92.
 Petrie TA.
@@
 \leavevmode\hypertarget{ref-Petrie2014}{}%
-92.
+93.
 Petrie TA, Deiters J, Harmison RJ.
@@
 
+\leavevmode\hypertarget{ref-Bellenger2016}{}%
+94.
+Bellenger CR, Fuller JT, Thomson RL, Davison K, Robertson EY, Buckley JD.
+Monitoring athletic training status through autonomic heart rate regulation: A systematic review and meta-analysis.
+\emph{Sports Medicine} (2016) \textbf{46}:1461--1486.
+doi:\href{https://doi.org/10.1007/s40279-016-0484-2}{10.1007/s40279-016-0484-2}
+
 \leavevmode\hypertarget{ref-Gisselman2016}{}%
-93.
+95.
 Gisselman AS, Baxter GD, Wright A, Hegedus E, Tumilty S.
@@
 \leavevmode\hypertarget{ref-Corr2016b}{}%
-94.
+96.
 Corr PJ, McNaughton N, Wilson MR, Hutchison A, Burch G, Poropat A.
@@
 \leavevmode\hypertarget{ref-Corr2013}{}%
-95.
+97.
 Corr PJ.
@@
 \leavevmode\hypertarget{ref-Chalmers2014}{}%
-96.
+98.
 Chalmers JA, Quintana DS, Abbott MJ-A, Kemp AH.
@@
 \leavevmode\hypertarget{ref-Pulopulos2018}{}%
-97.
+99.
 Pulopulos MM, Vanderhasselt MA, De Raedt R.
@@
 \leavevmode\hypertarget{ref-Qian2015}{}%
-98.
+100.
 Qian SS, Miltner RJ.
@@
 \leavevmode\hypertarget{ref-Ross2011}{}%
-99.
+101.
 Ross SE, Linens SW, Wright CJ, Arnold BL.
@@
 doi:\href{https://doi.org/10.1016/j.gaitpost.2011.07.011}{10.1016/j.gaitpost.2011.07.011}
+
+\leavevmode\hypertarget{ref-Ruddy2019}{}%
+102.
+Ruddy JD, Cormack SJ, Whiteley R, Williams MD, Timmins RG, Opar DA.
+Modeling the risk of team sport injuries: A narrative review of different statistical approaches.
+(2019) \textbf{10}:829.
+doi:\href{https://doi.org/10.3389/fphys.2019.00829}{10.3389/fphys.2019.00829}
+
+\leavevmode\hypertarget{ref-Carey2018}{}%
+103.
+Carey DL, Crossley KM, Whiteley R, Mosler A, Ong KL, Crow J, et al.
+Modeling Training Loads and Injuries: The Dangers of Discretization.
+\emph{Medicine and Science in Sports and Exercise} (2018) \textbf{50}:2267--2276.
+doi:\href{https://doi.org/10.1249/MSS.0000000000001685}{10.1249/MSS.0000000000001685}
+
+\leavevmode\hypertarget{ref-Pruyn2015}{}%
+104.
+Pruyn EC, Watsford ML, Murphy AJ.
+Differences in lower-body stiffness between levels of netball competition.
+\emph{Journal of Strength and Conditioning Research} (2015) \textbf{29}:1197--1202.
+doi:\href{https://doi.org/10.1519/JSC.0000000000000418}{10.1519/JSC.0000000000000418}
 \end{cslreferences}

### 2020-10-15 supplimentary/supplimentary_tables.tex
For submission
Update following reviewers comments

@@
 \hypersetup{
-  pdftitle={Supplimentary materials},
   hidelinks,
@@
 \captionsetup[table]{labelformat=empty}
+\usepackage{pdflscape}
+\newcommand{\blandscape}{\begin{landscape}}
+\newcommand{\elandscape}{\end{landscape}}
 \usepackage{booktabs}
@@
 
-\title{Supplimentary materials}
 \author{}
@@
 \begin{document}
-\maketitle
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}}
+\caption{\label{tab:unnamed-chunk-1}}
 \centering
@@
 \begin{longtable}[t]{l|l|r|r}
-\caption{\label{tab:unnamed-chunk-3}}\\
+\caption{\label{tab:unnamed-chunk-2}}\\
 \hline
@@
 \endfirsthead
-\caption[]{ S2 Table continued}\\
+\caption[]{ S3 Table continued}\\
 \hline
@@
 \endhead
-nlebase & nlelg\_1 & 1.00 & 1.00\\
+nlebase & NLE\_1 & 1.00 & 1.00\\
 \hline
-nlebase & nlelg\_2 & 1.00 & 1.00\\
+nlebase & NLE\_2 & 1.00 & 1.00\\
 \hline
-nlelg\_1 & injured\_1 & 1.00 & 1.00\\
+NLE\_1 & injured\_1 & 1.00 & 1.00\\
 \hline
-nlelg\_2 & injured\_2 & 1.00 & 1.00\\
+NLE\_2 & injured\_2 & 1.00 & 1.00\\
 \hline
@@
 \hline
-nlelg\_1 & BIS\_1 & 0.55 & 0.97\\
+NLE\_1 & BIS\_1 & 0.55 & 0.97\\
 \hline
-BIS\_1 & nlelg\_1 & 0.55 & 0.03\\
+BIS\_1 & NLE\_1 & 0.55 & 0.03\\
 \hline
@@
 \hline
-nlebase & hrv\_1 & 0.41 & 1.00\\
+nlebase & HRV\_1 & 0.41 & 1.00\\
 \hline
-hours & hrv\_1 & 0.40 & 1.00\\
+hours & HRV\_1 & 0.40 & 1.00\\
 \hline
-hrv\_1 & nlelg\_2 & 0.39 & 1.00\\
+HRV\_1 & NLE\_2 & 0.39 & 1.00\\
 \hline
@@
 \hline
-injured\_2 & hrv\_2 & 0.37 & 0.89\\
+injured\_2 & HRV\_2 & 0.37 & 0.89\\
 \hline
-nlelg\_2 & BIS\_2 & 0.37 & 0.96\\
+NLE\_2 & BIS\_2 & 0.37 & 0.96\\
 \hline
-BIS\_2 & nlelg\_2 & 0.37 & 0.04\\
+BIS\_2 & NLE\_2 & 0.37 & 0.04\\
 \hline
-hrv\_2 & injured\_2 & 0.37 & 0.11\\
+HRV\_2 & injured\_2 & 0.37 & 0.11\\
 \hline
-pi & nlelg\_2 & 0.36 & 1.00\\
+pi & NLE\_2 & 0.36 & 1.00\\
 \hline
@@
 \begin{longtable}[t]{l|l|r|r}
-\caption{\label{tab:unnamed-chunk-4}}\\
+\caption{\label{tab:unnamed-chunk-3}}\\
 \hline
@@
 \hline
-hrv & FFFS & 0.58 & 0.69\\
+HRV & FFFS & 0.58 & 0.69\\
 \hline
-FFFS & hrv & 0.58 & 0.31\\
+FFFS & HRV & 0.58 & 0.31\\
 \hline
@@
 \hline
-pi & hrv & 0.33 & 1.00\\
+pi & HRV & 0.33 & 1.00\\
 \hline
@@
 
-\textbf{S4 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}
+\textbf{S4 Table. Breakdown of injury count by sport and injury type (chronic or acute)}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-4}}
+\centering
+\begin{tabular}[t]{l|c|c}
+\hline
+sport & acute & chronic\\
+\hline
+athletics & 43.9\% (18) & 56.1\% (23)\\
+\hline
+basketball & 75.0\%  (6) & 25.0\%  (2)\\
+\hline
+cricket & 40.0\%  (4) & 60.0\%  (6)\\
+\hline
+football & 76.9\% (20) & 23.1\%  (6)\\
+\hline
+gym & 25.0\%  (1) & 75.0\%  (3)\\
+\hline
+hockey & 88.9\% (16) & 11.1\%  (2)\\
+\hline
+netball & 72.7\%  (8) & 27.3\%  (3)\\
+\hline
+other & 50.0\%  (3) & 50.0\%  (3)\\
+\hline
+rugby & 88.7\% (47) & 11.3\%  (6)\\
+\hline
+\end{tabular}
+\end{table}
+
+\newpage
+\begin{landscape}
+
+\textbf{S5 Table. Breakdown of injury count by sport and injury location}
 
@@
 \centering
+\begin{tabular}[t]{l|c|c|c|c|c|c|c|c|c}
+\hline
+sport & foot / ankle & lower leg & knee & upper leg & hip & torso & upper arm / shoulder & elbow / lower arm & head / neck\\
+\hline
+athletics & 36.6\% (15) & 14.6\%  (6) & 9.8\%  (4) & 17.1\%  (7) & 7.3\% (3) & 9.8\%  (4) & 4.9\%  (2) & 0.0\% (0) & 0.0\%  (0)\\
+\hline
+basketball & 25.0\%  (2) & 0.0\%  (0) & 12.5\%  (1) & 0.0\%  (0) & 0.0\% (0) & 37.5\%  (3) & 25.0\%  (2) & 0.0\% (0) & 0.0\%  (0)\\
+\hline
+cricket & 0.0\%  (0) & 10.0\%  (1) & 20.0\%  (2) & 30.0\%  (3) & 0.0\% (0) & 30.0\%  (3) & 0.0\%  (0) & 10.0\% (1) & 0.0\%  (0)\\
+\hline
+football & 15.4\%  (4) & 7.7\%  (2) & 15.4\%  (4) & 26.9\%  (7) & 0.0\% (0) & 0.0\%  (0) & 15.4\%  (4) & 3.8\% (1) & 15.4\%  (4)\\
+\hline
+gym & 0.0\%  (0) & 0.0\%  (0) & 25.0\%  (1) & 0.0\%  (0) & 25.0\% (1) & 25.0\%  (1) & 25.0\%  (1) & 0.0\% (0) & 0.0\%  (0)\\
+\hline
+hockey & 27.8\%  (5) & 16.7\%  (3) & 11.1\%  (2) & 22.2\%  (4) & 0.0\% (0) & 5.6\%  (1) & 0.0\%  (0) & 11.1\% (2) & 5.6\%  (1)\\
+\hline
+netball & 36.4\%  (4) & 0.0\%  (0) & 27.3\%  (3) & 9.1\%  (1) & 9.1\% (1) & 9.1\%  (1) & 9.1\%  (1) & 0.0\% (0) & 0.0\%  (0)\\
+\hline
+other & 16.7\%  (1) & 0.0\%  (0) & 16.7\%  (1) & 0.0\%  (0) & 0.0\% (0) & 50.0\%  (3) & 16.7\%  (1) & 0.0\% (0) & 0.0\%  (0)\\
+\hline
+rugby & 18.9\% (10) & 3.8\%  (2) & 26.4\% (14) & 17.0\%  (9) & 5.7\% (3) & 1.9\%  (1) & 17.0\%  (9) & 0.0\% (0) & 9.4\%  (5)\\
+\hline
+Total & 23.2\% (41) & 7.9\% (14) & 18.1\% (32) & 17.5\% (31) & 4.5\% (8) & 9.6\% (17) & 11.3\% (20) & 2.3\% (4) & 5.6\% (10)\\
+\hline
+\end{tabular}
+\end{table}
+
+\end{landscape}
+
+\newpage
+
+\textbf{S6 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-6}}
+\centering
 \begin{tabular}[t]{l|l|l|l|l|l}
 \hline
-prob & clevel & hours & nlelg\_1 & stiffness\_1 & balance\_1\\
+prob & clevel & hours & NLE\_1 & stiffness\_1 & balance\_1\\
 \hline
@@
 
-\textbf{S5 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_2.}
+\textbf{S7 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_2.}
 
@@
 
-\caption{\label{tab:unnamed-chunk-6}}
+\caption{\label{tab:unnamed-chunk-7}}
 \centering
@@
 \hline
-prob & FFFS\_1 & nlelg\_2 & stiffness\_2 & hrv\_2 & balance\_2\\
+prob & FFFS\_1 & NLE\_2 & stiffness\_2 & HRV\_2 & balance\_2\\
 \hline
@@
 
-\textbf{S6 Table. Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
+\textbf{S8 Table. Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
 

### 2020-10-15 supplimentary_tables.Rmd
Update following reviewers comments

@@
   - \captionsetup[table]{labelformat=empty}
+  - \usepackage{pdflscape}
+  - \newcommand{\blandscape}{\begin{landscape}}
+  - \newcommand{\elandscape}{\end{landscape}}
 knit: (function(inputFile, encoding) {
@@
 
-```{r include=FALSE}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn)
+pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, janitor)
 
-source("r-files/functions.R")
-source("r-files/methods_code.R")
-source("r-files/results_code.R")
+# source("r-files/functions.R")
+# source("r-files/methods_code.R")
+# source("r-files/results_code.R")
+load("all_data2.RData")
 ```
@@
 
-\textbf{S4 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}
+\textbf{S4 Table. Breakdown of injury count by sport and injury type (chronic or acute)}
+
+```{r results = 'asis'}
+
+injtable %>% 
+  left_join(datanet %>% 
+              select(id, sportg, time), c("id", "time")) %>% 
+  tabyl(sportg, acu_over) %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns() %>% 
+  select(sport = sportg, everything()) %>% 
+  mutate(across(.fns = ~str_replace(.x, "%", "\\\\%"))) %>% 
+  kable(
+    caption = "",
+    escape = F,
+    format = "latex",
+    align = "lcc"
+  ) %>% 
+  kable_styling(latex_options = ("HOLD_position"))
+```
+
+\newpage
+\blandscape
+
+\textbf{S5 Table. Breakdown of injury count by sport and injury location}
+
+```{r results = 'asis'}
+
+injtable %>% 
+  left_join(datanet %>% 
+              select(id, sportg, time), c("id", "time")) %>% 
+  mutate(bodypart = ifelse(bodypart %in% c("foot_toe", "ankle_achilles"), "foot_ankle", bodypart)) %>% 
+  tabyl(sportg, bodypart) %>% 
+  adorn_totals() %>% 
+  adorn_percentages() %>% 
+  adorn_pct_formatting() %>% 
+  adorn_ns() %>% 
+  select(sport = sportg,  
+         `foot / ankle` = foot_ankle, 
+         `lower leg` = lowerleg, 
+         knee,
+         `upper leg` = upperleg, 
+         hip, 
+         torso, 
+         `upper arm / shoulder` = upperarm_shoulder,
+         `elbow / lower arm` = elbow_lowerarm_hand, 
+         `head / neck` = head_neck) %>% 
+  mutate(across(.fns = ~str_replace(.x, "%", "\\\\%"))) %>% 
+  kable(
+    caption = "",
+    escape = F,
+    format = "latex",
+    align = "lccccccccc"
+  ) %>% 
+  kable_styling(latex_options = ("HOLD_position"))
+```
+
+\elandscape
+
+\newpage
+
+\textbf{S6 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}
 
@@
   tallc,
-  c("balance_2", "stiffness_2", "nlelg_2", "FFFS_1", "hrv_2"),
+  c("balance_2", "stiffness_2", "NLE_2", "FFFS_1", "HRV_2"),
   "injured_2",

### 2020-10-20 manuscript.Rmd
post editor comments

@@
 
-load("all_data2.RData")
+load("all_data2x.RData")
 #myEnv <- new.env()
@@
   npars %>% 
-    mutate(name = ifelse(name == "Hours per week", "Hours per week training", name)) %>%
+    mutate(name = ifelse(name == "Hours per week", "Training hours per week", name)) %>%
     slice(c(1,3,8, 4, 9:11)) %>% 
@@
   col.names = c(" ","Female (n = 120)", "Male (n = 231)"),
-  align = "c",
+  align = "lccc",
   escape = F
@@
 Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
-Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point.
+In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point (baseline NLE) was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].
 
@@
 Finally, arc direction was restricted between specific nodes within the explanatory variables.
-Arcs from competitive level $\rightarrow$ gender, baseline negative life events (NLE) $\rightarrow$ gender and baseline NLE $\rightarrow$ sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
+Arcs from competitive level $\rightarrow$ gender, baseline NLE $\rightarrow$ gender and baseline NLE $\rightarrow$ sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
 All subsequent models used the same blacklist.
@@
 Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
-TLE score was not included in the final structure.
+Given that NLE provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
+TLE was not included in the final structure.
 
@@
 Therefore, only RI was included in the final structure.
-Finally, both total score and asymmetry for balance were included in the initial network.
-However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
+Finally, both total score and asymmetry (percentage difference in score between limbs) for balance were included in the initial network.
+However, visual inspection of the network revealed no arcs between the balance asymmetry node and any other node in the network.
 Therefore, balance asymmetry was removed from the final network structure.
@@
 ```{r table2, results='asis'}
+
 varcutoffs %>% 
+  mutate(var = case_when(var == "clevel" ~ "Competitve level",
+                         var == "gender" ~"Gender",
+                         var == "pi" ~ "Previous injury",
+                         var == "hours" ~ "Training hours",
+                         var == "ind_team" ~ "Sport type",
+                         var == "nlebase" ~ "Baseline NLE",
+                         var == "stiffness" ~ "Stiffness",
+                         var == "balance" ~ "Balance",
+                         TRUE ~ var))  %>% 
   filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
@@
        col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
-  kable_styling(latex_options = "HOLD_position") %>% 
+  kable_styling(latex_options = "HOLD_position", font_size = 11) %>% 
   column_spec(2, width = c("5.2cm")) 
@@
 The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) $\rightarrow$ Fight-Flight-Freeze System (FFFS) and HRV $\rightarrow$ FFFS.
+Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) $\rightarrow$ Fight-Flight-Freeze System (FFFS) and heart arte variability (HRV) $\rightarrow$ FFFS.
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
@@
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
-The "individual or team sport" node had strong arcs to hours per week (`r str_finder("ind_team", "hours")`) and baseline negative life events (NLE) (`r str_finder("ind_team", "nlebase")`).
-Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
+The "Sport type" node had strong arcs to training hours (`r str_finder("ind_team", "hours")`) and baseline negative life events (NLE) (`r str_finder("ind_team", "nlebase")`).
+Individual athletes were more likely to have "High" training hours (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
 Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
@@
 Irrespective of sport or gender, strong connections were found between explanatory variables.
-The arc from baseline NLE (nlebase) $\rightarrow$ Reward Interest (RI\_1) had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI\_1 being in the "High" state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from "Low" to "High".
+The arc from baseline NLE $\rightarrow$ Reward Interest (RI\_1) had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI\_1 being in the "High" state increased from `r x[2,1]` to `r x[3,1]` when baseline NLE increased from "Low" to "High".
 
@@
 The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
-renderGraph(gR)
+#renderGraph(gR)
+include_graphics("figures_doc/Fig1.jpeg")
 ```
@@
 
-The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week, negative life events (NLE\_1), muscle stiffness (stiffness\_1), current competitive level and balance (balance\_1).
+The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Figure \ref{fig:fig2}), which contained hours spent training per week, negative life events (NLE\_1), muscle stiffness (stiffness\_1), competitive level and balance (balance\_1).
 The arc between NLE\_1 and injured\_1 was fixed in the network, so has the maximum strength of 1.
@@
   select(c(1, 2, 3)) %>%
+  mutate(variable = case_when(variable == "hours" ~ "Training hours",
+                              variable == "NLE_1" ~ "Negative life events_1",
+                              TRUE ~ str_to_sentence(variable))) %>% 
   mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
@@
   select(7, 2:6) %>% 
-  rename(Probability = "prob") %>%
+  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
+  rename(Probability = "prob",
+         "Competitive level" = clevel,
+         "Training hours" = hours,
+         "Negative life events_1" = "NLE_1") %>%
+  rename_all(~str_to_sentence(.x))%>%
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
-  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
   kable(
@@
   ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
+  kable_styling(latex_options = "HOLD_position",font_size = 11) %>%
   pack_rows("Highest", 1, 3) %>%
   pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
+  row_spec(0,bold=TRUE) %>% 
+  column_spec(3:4, width = "2cm")
 ```
@@
   mutate(Low = format(Low, nsmall = 2)) %>% 
+  mutate(variable = case_when(variable == "FFFS_1" ~ "Fight-Flight-Freeze System_1",
+                              variable == "NLE_2" ~ "Negative life events_2",
+                              variable == "HRV_2" ~ "Heart rate variability_2",
+                              TRUE ~ str_to_sentence(variable))) %>% 
   mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
@@
     escape = F,
-    align = "cccccc"
+    align = "lccccc"
   ) %>%
@@
   select(7,2:6) %>%
-  rename(Probability = "prob") %>%
+  rename(Probability = "prob",
+         "Fight-Flight-Freeze System_1" = FFFS_1,
+         "Negative life events_2 "= NLE_2,
+         "Heart rate variability_2" = "HRV_2",
+         Stiffness_2 = stiffness_2,
+         Balance_2 = balance_2) %>%
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
@@
   ) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 11) %>%
   pack_rows("Highest", 1, 3) %>%
   pack_rows("Lowest", 4, 6) %>%
-  row_spec(0,bold=TRUE)
+  row_spec(0,bold=TRUE) %>% 
+  column_spec(c(2,3),width = "2cm") %>% 
+  column_spec(5, width = "2.8cm")
 ```
@@
 bmtable %>% 
+  mutate(Term = case_when(Term == "BIS" ~ "Behavioural Inhibition System (BIS)",
+                          Term == "HRV" ~ "Heart rate variability (HRV)",
+                          TRUE ~ Term)) %>% 
   kable(
     .,
-    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and HRV.",
+    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing Fight-Flight-Freeze System, Behavioural Inhibition System and Heart rate variability",
     col.names = c("Term", "Estimate", "Error", "95\\% CI"),
@@
     .,
-    col.names = c("Probability", "nlec", "Stiffness"),
-    caption = "The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.",
+    col.names = c("Probability", "Negative life events", "Stiffness"),
+    caption = "The probability of injury with values of stiffness and negative life events held at 1SD below the mean change, at the mean change and 1 SD above the mean change.",
     align = "ccc"
@@
 Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
-The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
-In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).
+The combination of 1 SD above the mean change for negative life events (NLE)  and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
+In contrast, below average changes in NLE and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).
 
@@
   select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
+  rename(Probability = "prob",
+         "Training hours" = hours,
+         "Previous injury" = pi,
+         "Negative life events" = nle,
+         Stiffness = stiffness) %>%
   kable(
@@
   pack_rows("Lowest", 4, 6) %>%
-  kable_styling(latex_options = "HOLD_position") %>%
-  row_spec(0,bold=TRUE)
+  kable_styling(latex_options = "HOLD_position", font_size = 11) %>%
+  row_spec(0,bold=TRUE) %>% 
+  column_spec(c(2,4), width = "2.3cm")
 ```

### 2020-10-20 manuscript.tex
post editor comments

@@
 \centering
-\begin{tabular}[t]{c|c|c}
+\begin{tabular}[t]{l|c|c}
 \hline
@@
 \hline
-\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
+\hspace{1em}Training hours per week & 8.5 (4.5) & 11.2 (8.8)\\
 \hline
@@
 Given the limited support for a relationship between positive life events and injury (15), only negative and total life events were included in the network.
-Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
-In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).
+Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point.
+In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point (baseline NLE) was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).
 
@@
 Finally, arc direction was restricted between specific nodes within the explanatory variables.
-Arcs from competitive level \(\rightarrow\) gender, baseline negative life events (NLE) \(\rightarrow\) gender and baseline NLE \(\rightarrow\) sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
+Arcs from competitive level \(\rightarrow\) gender, baseline NLE \(\rightarrow\) gender and baseline NLE \(\rightarrow\) sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
 All subsequent models used the same blacklist.
@@
 Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
-Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
-TLE score was not included in the final structure.
+Given that NLE provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
+TLE was not included in the final structure.
 
@@
 \centering
+\fontsize{11}{13}\selectfont
 \begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
@@
 \hline
-clevel & Current competitive level & Club\_university\_county & National\_international\\
-\hline
-gender & Gender of the participant & Female & Male\\
-\hline
-hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
-\hline
-ind\_team & Participate in an individual or team based sport & Individual & Team\\
-\hline
-pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
-\hline
-nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
+Competitve level & Current competitive level & Club\_university\_county & National\_international\\
+\hline
+Gender & Gender of the participant & Female & Male\\
+\hline
+Training hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
+\hline
+Sport type & Participate in an individual or team based sport & Individual & Team\\
+\hline
+Previous injury & Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
+\hline
+Baseline NLE & Untransformed NLE at the first time point & 0-13 (Low) & >13-93 (High)\\
 \hline
@@
 \hline
-stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
+Stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
 \hline
@@
 \hline
-balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
-\hline
-NLE\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
-\hline
-NLE\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
-\hline
-NLE\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
-\hline
-TLE\_1 & Log TLE at TP 1 & 1.79-3.4 (Low) & >3.4-4.88 (High)\\
-\hline
-TLE\_2 & Log TLE at TP 2 & 1.79-3.74 (Low) & >3.74-5.42 (High)\\
-\hline
-TLE\_3 & Log TLE at TP 3 & 1.79-3.81 (Low) & >3.81-5.18 (High)\\
+Balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
+\hline
+NLE\_1 & Log Negative life events (NLE) at time 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
+\hline
+NLE\_2 & Log NLE at time 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
+\hline
+NLE\_3 & Log NLE at time 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
+\hline
+TLE\_1 & Log Total life events (TLE) at time 1 & 1.79-3.4 (Low) & >3.4-4.88 (High)\\
+\hline
+TLE\_2 & Log TLE at time 2 & 1.79-3.74 (Low) & >3.74-5.42 (High)\\
+\hline
+TLE\_3 & Log TLE at time 3 & 1.79-3.81 (Low) & >3.81-5.18 (High)\\
 \hline
@@
 The Markov blanket of the ``injured'' node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
-Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) and HRV \(\rightarrow\) FFFS.
+Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) and heart arte variability (HRV) \(\rightarrow\) FFFS.
 To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
@@
 Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
-The ``individual or team sport'' node had strong arcs to hours per week (0.94) and baseline negative life events (NLE) (0.78).
-Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
+The ``Sport type'' node had strong arcs to training hours (0.94) and baseline negative life events (NLE) (0.78).
+Individual athletes were more likely to have ``High'' training hours (0.84) compared to team-based athletes (0.60).
 Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
@@
 Irrespective of sport or gender, strong connections were found between explanatory variables.
-The arc from baseline NLE (nlebase) \(\rightarrow\) Reward Interest (RI\_1) had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.
+The arc from baseline NLE \(\rightarrow\) Reward Interest (RI\_1) had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when baseline NLE increased from ``Low'' to ``High''.
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/fig1-1.pdf}
+\includegraphics{figures_doc/Fig1.jpeg}
 \caption{\label{fig:fig1}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
@@
 \hline
-balance\_1 & 0.21 & 0.30\\
-\hline
-hours & 0.18 & 0.28\\
-\hline
-NLE\_1 & 0.24 & 0.26\\
-\hline
-stiffness\_1 & 0.17 & 0.31\\
+Balance\_1 & 0.21 & 0.30\\
+\hline
+Training hours & 0.18 & 0.28\\
+\hline
+Negative life events\_1 & 0.24 & 0.26\\
+\hline
+Stiffness\_1 & 0.17 & 0.31\\
 \hline
@@
 \centering
-\begin{tabular}[t]{c|c|c|c|c|c}
-\hline
-\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{NLE\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
+\fontsize{11}{13}\selectfont
+\begin{tabular}[t]{c|c|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|c|c}
+\hline
+\textbf{Probability} & \textbf{Competitive level} & \textbf{Training hours} & \textbf{Negative life events\_1} & \textbf{Stiffness\_1} & \textbf{Balance\_1}\\
 \hline
@@
 \hline
-balance\_2 & 0.17 & 0.27\\
-\hline
-FFFS\_1 & 0.30 & 0.11\\
-\hline
-HRV\_2 & 0.26 & 0.17\\
-\hline
-NLE\_2 & 0.23 & 0.19\\
-\hline
-stiffness\_2 & 0.13 & 0.27\\
+Balance\_2 & 0.17 & 0.27\\
+\hline
+Fight-Flight-Freeze System\_1 & 0.30 & 0.11\\
+\hline
+Heart rate variability\_2 & 0.26 & 0.17\\
+\hline
+Negative life events\_2 & 0.23 & 0.19\\
+\hline
+Stiffness\_2 & 0.13 & 0.27\\
 \hline
@@
 \centering
-\begin{tabular}[t]{c|c|c|c|c|c}
-\hline
-\textbf{Probability} & \textbf{FFFS\_1} & \textbf{NLE\_2} & \textbf{stiffness\_2} & \textbf{HRV\_2} & \textbf{balance\_2}\\
+\fontsize{11}{13}\selectfont
+\begin{tabular}[t]{c|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|c|>{\centering\arraybackslash}p{2.8cm}|c}
+\hline
+\textbf{Probability} & \textbf{Fight-Flight-Freeze System\_1} & \textbf{Negative life events\_2 } & \textbf{Stiffness\_2} & \textbf{Heart rate variability\_2} & \textbf{Balance\_2}\\
 \hline
@@
 
-\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and HRV.}
+\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing Fight-Flight-Freeze System, Behavioural Inhibition System and Heart rate variability}
 \centering
@@
 \hline
-BIS & 0.41 & 0.03 & [0.36, 0.47]\\
-\hline
-HRV & -0.19 & 0.03 & [-0.25, -0.13]\\
+Behavioural Inhibition System (BIS) & 0.41 & 0.03 & [0.36, 0.47]\\
+\hline
+Heart rate variability (HRV) & -0.19 & 0.03 & [-0.25, -0.13]\\
 \hline
@@
 
-\caption{\label{tab:table9}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
+\caption{\label{tab:table9}The probability of injury with values of stiffness and negative life events held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
 \centering
@@
 \hline
-\textbf{Probability} & \textbf{nlec} & \textbf{Stiffness}\\
+\textbf{Probability} & \textbf{Negative life events} & \textbf{Stiffness}\\
 \hline
@@
 Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
-The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
-In contrast, below average changes in nlec and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.11).
+The combination of 1 SD above the mean change for negative life events (NLE) and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
+In contrast, below average changes in NLE and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.11).
 
@@
 \centering
-\begin{tabular}[t]{c|c|c|c|c}
-\hline
-\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
+\fontsize{11}{13}\selectfont
+\begin{tabular}[t]{c|>{\centering\arraybackslash}p{2.3cm}|c|>{\centering\arraybackslash}p{2.3cm}|c}
+\hline
+\textbf{Probability} & \textbf{Training hours} & \textbf{Previous injury} & \textbf{Negative life events} & \textbf{Stiffness}\\
 \hline

### 2020-10-20 supplimentary_tables.Rmd
post editor comments

@@
 # source("r-files/results_code.R")
-load("all_data2.RData")
+load("all_data2x.RData")
 ```
@@
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
+    mutate(across(1:2, .fns = ~case_when(.x == "pi" ~ "Previous\ninjury",
+                                       .x == "gender" ~ "Gender",
+                                       .x == "clevel" ~ "Competitive\nlevel",
+                                       .x == "hours" ~"Training hours",
+                                       .x == "ind_team" ~ "Sport type",
+                                       .x == "nlebase" ~ "Baseline\nNLE",
+                                       .x == "stiffness_1" ~ "Stiffness_1",
+                                       .x == "stiffness_2" ~ "Stiffness_2",
+                                       .x == "NLE_1" ~ "Negative life\nevents_1",
+                                       .x == "NLE_2" ~ "Negative life\nevents_2",
+                                       .x == "RI_1" ~ "Reward\nInterest_1",
+                                       .x == "RI_2" ~ "Reward\nInterest_2",
+                                       .x == "BIS_1" ~ "Behavioural\nInhibition\nSystem_1",
+                                       .x == "BIS_2" ~ "Behavioural\nInhibition\nSystem_2",
+                                       .x == "HRV_1" ~ "Heart rate\nvariability_1",
+                                       .x == "HRV_2" ~ "Heart rate\nvariability_2",
+                                       .x == "balance_1" ~ "Balance_1",
+                                       .x == "balance_2" ~ "Balance_2",
+                                       .x == "FFFS_1" ~ "Fight-Flight-\nFreeze\nSystem_1",
+                                       .x == "FFFS_2" ~ "Fight-Flight-\nFreeze\nSystem_2",
+                                       .x == "injured_1" ~ "Injured_1",
+                                       .x == "injured_2" ~ "Injured_2"
+  ))) %>% 
    mutate_at(c("from", "to"), ~str_replace_all(., "_", "\\\\_")) %>% 
@@
   arrange(-strength) %>% 
+  mutate(across(1:2, .fns = ~case_when(.x == "pi" ~ "Previous\ninjury",
+                                       .x == "gender" ~ "Gender",
+                                       .x == "clevel" ~ "Competitive\nlevel",
+                                       .x == "hours" ~"Training\nhours",
+                                       .x == "stiffness" ~ "Stiffness",
+                                       .x == "nlec" ~ "Negative life\nevents",
+                                       .x == "RI" ~ "Reward\ninterest",
+                                       .x == "BIS" ~ "Behavioural\nInhibition\nSystem",
+                                       .x == "HRV" ~ "Heart rate\nvariability",
+                                       .x == "balance" ~ "Balance",
+                                       .x == "FFFS" ~ "Fight-Flight-\nFreeze\nSystem",
+                                       .x == "injured" ~ "Injured"))) %>%
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
@@
   select(7, 2:6) %>% 
+    rename(Probability = "prob",
+         "Competitive level" = clevel,
+         "Training hours" = hours,
+         "Negative life events_1" = "NLE_1") %>% 
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>% 
@@
   caption = "",
-    escape = F
+    escape = F,
+  align = "c"
 ) %>%
@@
   select(7, 2:6) %>%
+    rename(Probability = "prob",
+         "Fight-Flight-Freeze System_1" = FFFS_1,
+         "Negative life events_2 "= NLE_2,
+         "Heart rate variability_2" = "HRV_2",
+         Stiffness_2 = stiffness_2,
+         Balance_2 = balance_2) %>%
   `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_"))
@@
     caption = "",
-    escape = F
+    escape = F,
+    align = "c"
   ) %>%
-  kable_styling(latex_options = "HOLD_position")
+  kable_styling(latex_options = "HOLD_position") %>% 
+      column_spec(c(2,3),width = "2cm") %>% 
+  column_spec(5, width = "2.8cm")
 
@@
   select(prob, hours, pi, nle, stiffness) %>%
-  rename(Probability = "prob") %>%
+    rename(Probability = "prob",
+         "Training hours" = hours,
+         "Previous injury" = pi,
+         "Negative life events" = nle,
+         Stiffness = stiffness) %>%
   dendroTools::round_df(2) %>%
@@
   kable_styling(latex_options = "HOLD_position") %>%
-
   row_spec(0,bold=TRUE)

### 2020-10-20 r-files/functions.R

@@ -184,14 +184,14 @@ newprobtable <- function(tallc, vars, outcome, state, model, repeats = 500000) {
   
   str3 = rep(paste("(", outcome, " == '", state, "')", sep = ""), each = length(str1)/length(outcome))  # repeat each outcome for the length of combos
   
-  fitted <-  bn.fit(avg30, tallc, method = "bayes", iss = 1) # fit the model with bayes method
+  fitted <-  bn.fit(model, tallc, method = "bayes", iss = 1) # fit the model with bayes method
   
-
+  
   
   cmd = paste("cpquery(fitted, ", str3, ", ", str1, ", method = 'lw', n = ", repeats, ")", sep = "") # join all elements of string together
- 
   
-   prob <-  rep(0, length(str1)) # empty vector for probabilities 
+  
+  prob <-  rep(0, length(str1)) # empty vector for probabilities 
   for (i in seq(length(cmd))){
     prob[i] <- eval(parse(text = cmd[i]))
   } # for each combination of strings, what is the probability of outcome

### 2020-10-20 r-files/methods_code.R

@@ -142,9 +142,9 @@ icc <- irr::icc(d, model = "twoway", type = "agreement")
 
 
 
-data <- read_csv("data/full_data_raw.csv") %>%
+data <- read_csv("data/fulldatajoined.csv") %>%
   filter(time != 4) %>% 
-  rename(hrv= rmssd)# remove last time point as only questionnaire data is available
+  rename(HRV= rmssd)# remove last time point as only questionnaire data is available
 
 ####################### Cleaning up variables + impute #################################
 
@@ -183,7 +183,7 @@ voi <- data %>%
     stiffness,
     frequency,
     decrement,
-    hrv,
+    HRV,
     sdnn,
     meanHR,
     bis,
@@ -203,16 +203,16 @@ ids <- data %>%
 ## check how many missing
 
 missingcount <- tibble(
-  var = c("hrv", "myoton"),
-  missing = c(map(voi, ~ sum(is.na(.)))$hrv, map(voi, ~ sum(is.na(.)))$stiffness),
+  var = c("HRV", "myoton"),
+  missing = c(map(voi, ~ sum(is.na(.)))$HRV, map(voi, ~ sum(is.na(.)))$stiffness),
   total = c(668, 668)
 ) %>%
   mutate(percdiff = missing / total * 100)
 
 missingtot <- voi %>%
-  select(id, hrv, stiffness) %>%
+  select(id, HRV, stiffness) %>%
   mutate(
-    hrv = tidyr::replace_na(hrv, 0),
+    HRV = tidyr::replace_na(HRV, 0),
     stiffness = tidyr::replace_na(stiffness, 0)
   )
 
@@ -299,7 +299,7 @@ datanet <- datanew %>%
   ungroup() %>%
   mutate(negsev = ifelse(negsev == "NaN", 0, negsev)) %>%
   ungroup() %>%
-  mutate(hrv = round(log(hrv), digits = 2)) %>%
+  mutate(HRV = round(log(HRV), digits = 2)) %>%
   group_by(time) %>%
   mutate(
     totneg_d = splitr(totneg),
@@ -332,7 +332,7 @@ nlename <- list(
       GDP,
       RI,
       sdnn,
-      hrv,
+      HRV,
       balance,
       FFFS
     ) %>% names(),
@@ -355,7 +355,7 @@ nlename <- list(
       GDP,
       RI,
       sdnn,
-      hrv,
+      HRV,
       balance,
       FFFS
[…]

### 2020-10-20 r-files/results_code.R

@@ -33,7 +33,9 @@ inj_mechanism <- injtable %>%
 
 
 
-tallc <- networkprepare(networkdata)
+tallc <- networkprepare(networkdata) %>% 
+  rename(NLE_1 = nlelg_1,
+         NLE_2 = nlelg_2)
 ## extract names for blacklist + plotting
 
 g1 <- tallc %>%
@@ -69,7 +71,7 @@ bl <- expand.grid(from = g2, to = g1, stringsAsFactors = FALSE) %>%
   )
 ## create whitelist
 wl <- data.frame(
-  from = c("nlelg_1", "nlelg_2"),
+  from = c("NLE_1", "NLE_2"),
   to = c("injured_1", "injured_2")
 )
 
@@ -117,11 +119,114 @@ gR <- renderGraph(gR)
 jpeg("./figures_doc/Fig1.jpeg", res = 300, width = 20, height = 15, units = "cm")
 renderGraph(gR)
 dev.off()
+
+# with full labels
+
+zz <- c("Previous\ninjury",
+        "Baseline\nNLE",
+        "Gender",
+        "Sport type",
+        "Competitive\nlevel",
+        "Training hours",
+        "Injured_1",
+        "Negative life\nevents_1",
+        "Stiffness_1",
+        "Reward\nInterest_1",
+        "Behavioural\nInhibition\nSystem_1",
+        "Heart rate\nvariability_1",
+        "Balance_1",
+        "Fight-Flight-\nFreeze\nSystem_1",
+        "Injured_2",
+        "Negative life\nevents_2",
+        "Stiffness_2",
+        "Reward\nInterest_2",
+        "Behavioural\nInhibition\nSystem_2",
+        "Heart rate\nvariability_2",
+        "Balance_2",
+        "Fight-Flight-\nFreeze\nSystem_2")
+
+avg30_x <- avg30
+
+nodes(avg30_x) <- zz
+
+str_x <- str %>% 
+  mutate(across(1:2, .fns = ~case_when(.x == "pi" ~ "Previous\ninjury",
+                                       .x == "gender" ~ "Gender",
+                                       .x == "clevel" ~ "Competitive\nlevel",
+                                       .x == "hours" ~"Training hours",
+                                       .x == "ind_team" ~ "Sport type",
+                                       .x == "nlebase" ~ "Baseline\nNLE",
+                                       .x == "stiffness_1" ~ "Stiffness_1",
[…]

### 2020-11-15 manuscript.Rmd
post reviewer 4 comments

@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
-# source("r-files/functions.R")
-# source("r-files/methods_code.R")
-# source("r-files/results_code.R")
-
-load("all_data2x.RData")
-#myEnv <- new.env()
-#attach(myEnv, name="sourced_scripts")
+source("r-files/functions.R")
+source("r-files/methods_code.R")
+source("r-files/results_code.R")
 
@@
 For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence [@Lee2017; @Djaoui2017] and may account for the unexplained variance from the psychological predictors.
-In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's [@Williams1998] model.
+In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna [@Appaneal2014] proposed the biopsychosocial model of stress, athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's [@Williams1998] model.
 The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioural factors that impact sports participation [@Appaneal2014].
@@
 
-Using the frameworks provided by Williams and Andersen's [@Williams1998] stress injury model and Appaneal's BMSAIH model [@Appaneal2014], the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial and physiological stress-related factors with injury occurrence.
-A prospective, repeated measures design incorporating field-based physiological (heart rate variability, muscle stiffness and postural control) and psychosocial measures (Life Event Survey for Collegiate Athletes and Reinforcement Sensitivity Questionnaire Personality Questionnaire) combined with a BN modelling analysis was used to address the study aim.
+Using the frameworks provided by Williams and Andersen's [@Williams1998] stress injury model and Appaneal's BMSAIH model [@Appaneal2014], the aim of this longitudinal interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial (life event stress and reinforcement sensitivity) and physiological stress-related factors (heart rate variability, muscle stiffness and postural control) with injury occurrence.
+Due to the inevitable constraints of a longitudinal interdisciplinary study of this kind, we have not been able to incorporate a number of other variables that might also be of interest (e.g., coping resources, emotional states and attentional deficits).
+A prospective, repeated measures design incorporating the physiological and psychosocial measures combined with BN modelling was used to address the study aim.
 
@@
 Participants hands were placed on hips at the level of the iliac crests.
-A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.
+A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants' performance during the mBESS.
 
@@
 varcutoffs %>% 
-  mutate(var = case_when(var == "clevel" ~ "Competitve level",
+  mutate(var = case_when(var == "clevel" ~ "Competitive level",
                          var == "gender" ~"Gender",
@@
 
-```{r fig1, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn.
+```{r fig1, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the "strength.plot" function in bnlearn.
 The strength of each arc is shown graphically by the style of the arc.
 Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1).
-White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables.
-The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
+White nodes in the network indicate the explanatory variables, blue nodes indicate T1\\_1 and T2\\_1 variables, and red nodes indicated T2\\_2 and T3\\_2 variables.
+The injured\\_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
 #renderGraph(gR)
@@
     .,
-    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable",
+    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
     col.names = c("Variable", "Low", "High"),
@@
 Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+As advocated by Johnson et al.
+(2014), future research should continue to use repeated measure designs, including designs with more frequent monitoring, such as daily and weekly measures to gain further insight into the dynamic interplay between stress-related factors and injury.
+Sport injury research has also been criticised for adopting analytic approaches that are reductionist in nature [@Bittencourt2016] that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
 The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
@@
 
-While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
+While the BNs provided a contemporary approach that improved upon traditional methods such as logistic regression [@Olmedilla2018], a number of assumptions were made that potentially limited the approach employed in the study.
 Firstly, the choice was made to binarise variables in the first network so only "Low" and "High" states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner [@Qian2015] highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+Although binarising variables is a common procedure in BN analysis and has several advantages, Qian and Miltner [@Qian2015] highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
 For example, the meaning of a "Low" and "High" value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
+The use of a median split to determine the "Low" and "High" states could also be improved by using known cut-off scores that are associated with a higher risk of injury.
+While the measures used in the current research did not have any clear cut-off scores supported by the literature, future research should aim to identify potential cut-off scores that would consequently enable a more meaningful splitting procedure compared to using the group median.
 Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
@@
 The variables used in the present study were not definitive.
-Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], could help to further elucidate the relationship between stress and injury.
+Additional measures relating to coping, injury prevention behaviours, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress [@Appaneal2014; @Perna1995], could help to further elucidate the relationship between stress and injury.
+Future research should continue to explore the relationship between these measures, as well as their effect on the risk of injury.
+In particular, there is scope to build upon Appaneal and Perna's (2014) biopsychosocial model of stress, athletic injury and health which provides a framework for examining the interplay between psychological, behavioural and training-related factors.
 Future developments in the capture of life event stress using the LESCA are also warranted.

### 2020-11-15 manuscript.tex
post reviewer 4 comments

@@
 For example, a large body of research has suggested that training-related stress is also likely to contribute to the stress response and injury occurrence (26,27) and may account for the unexplained variance from the psychological predictors.
-In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna (28) proposed the biopsychosocial model of stress athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (14) model.
+In an attempt to combine the psychosocial factors proposed by Williams and Andersen (1998) and potential markers of training-related stress, Appaneal and Perna (28) proposed the biopsychosocial model of stress, athletic injury and health (BMSAIH) to serve as an extension to Williams and Andersen's (14) model.
 The BMSAIH enhances our understanding of the mediating pathways between the stress response and injury alongside other health outcomes and behavioural factors that impact sports participation (28).
@@
 
-Using the frameworks provided by Williams and Andersen's (14) stress injury model and Appaneal's BMSAIH model (28), the aim of this interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial and physiological stress-related factors with injury occurrence.
-A prospective, repeated measures design incorporating field-based physiological (heart rate variability, muscle stiffness and postural control) and psychosocial measures (Life Event Survey for Collegiate Athletes and Reinforcement Sensitivity Questionnaire Personality Questionnaire) combined with a BN modelling analysis was used to address the study aim.
+Using the frameworks provided by Williams and Andersen's (14) stress injury model and Appaneal's BMSAIH model (28), the aim of this longitudinal interdisciplinary study was to develop new understanding of the multifaceted interactions of psychosocial (life event stress and reinforcement sensitivity) and physiological stress-related factors (heart rate variability, muscle stiffness and postural control) with injury occurrence.
+Due to the inevitable constraints of a longitudinal interdisciplinary study of this kind, we have not been able to incorporate a number of other variables that might also be of interest (e.g., coping resources, emotional states and attentional deficits).
+A prospective, repeated measures design incorporating the physiological and psychosocial measures combined with BN modelling was used to address the study aim.
 
@@
 \hline
-Competitve level & Current competitive level & Club\_university\_county & National\_international\\
+Competitive level & Current competitive level & Club\_university\_county & National\_international\\
 \hline
@@
 \begin{figure}
-\centering
-\includegraphics{figures_doc/Fig1.jpeg}
-\caption{\label{fig:fig1}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn.
+\includegraphics[width=7.87in]{figures_doc/Fig1} \caption{The full Bayesian network structure was plotted using the "strength.plot" function in bnlearn.
 The strength of each arc is shown graphically by the style of the arc.
@@
 White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables.
-The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
+The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}\label{fig:fig1}
 \end{figure}
@@
 
-\caption{\label{tab:table4}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
+\caption{\label{tab:table4}Probability of injured\_1 being in the ``injured'' state, conditional on each variable.}
 \centering
@@
 Although there are unique and significant challenges with research employing such designs, a more fine-grained understanding of the dynamic relationships between stress-related factors and injury occurrence in athletes was achieved when compared to traditional cross-sectional, single time point research.
-Sport injury research has been criticised for adopting analytic approaches that are reductionist in nature (39) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
+As advocated by Johnson et al.~(2014), future research should continue to use repeated measure designs, including designs with more frequent monitoring, such as daily and weekly measures to gain further insight into the dynamic interplay between stress-related factors and injury.
+Sport injury research has also been criticised for adopting analytic approaches that are reductionist in nature (39) that fail to account for the complex, emergent behaviour that is characteristic of injury occurrence.
 The use of an interdisciplinary framework combined with a BN modelling approach in the study facilitated extended insight into the complex interplay that exists between psychosocial and physiological markers of stress and injury occurrence.
@@
 
-While the BN's provided a contemporary approach that improved upon traditional methods such as logistic regression (41), a number of assumptions were made that potentially limited the approach employed in the study.
+While the BNs provided a contemporary approach that improved upon traditional methods such as logistic regression (41), a number of assumptions were made that potentially limited the approach employed in the study.
 Firstly, the choice was made to binarise variables in the first network so only ``Low'' and ``High'' states were observed.
-Although binarising variables is a common procedure in BN analysis and has several advantages Qian and Miltner (100) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
+Although binarising variables is a common procedure in BN analysis and has several advantages, Qian and Miltner (100) highlighted that both a loss of statistical accuracy and potential difficulty in the subsequent interpretation of the model may arise.
 For example, the meaning of a ``Low'' and ``High'' value in this study was only meaningful for the population studied, and there could be additional levels within each category that were not investigated.
+The use of a median split to determine the ``Low'' and ``High'' states could also be improved by using known cut-off scores that are associated with a higher risk of injury.
+While the measures used in the current research did not have any clear cut-off scores supported by the literature, future research should aim to identify potential cut-off scores that would consequently enable a more meaningful splitting procedure compared to using the group median.
 Furthermore, in order to collect data on a large sample of participants, suitable measures were required to ensure the viability of the data collection.
@@
 The variables used in the present study were not definitive.
-Additional measures relating to coping, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress (28,29), could help to further elucidate the relationship between stress and injury.
+Additional measures relating to coping, injury prevention behaviours, injury-specific biomechanics and stress hormones such as cortisol, which been found to be a marker of both psychological and training-related stress (28,29), could help to further elucidate the relationship between stress and injury.
+Future research should continue to explore the relationship between these measures, as well as their effect on the risk of injury.
+In particular, there is scope to build upon Appaneal and Perna's (2014) biopsychosocial model of stress, athletic injury and health which provides a framework for examining the interplay between psychological, behavioural and training-related factors.
 Future developments in the capture of life event stress using the LESCA are also warranted.

### 2020-11-15 r-files/methods_code.R

@@ -596,3 +596,29 @@ varcutoffs[23, 2] <- "Log TLE at time 3"
 ####
 
 
+# datanet %>%
+#   select(time, gender, stiffness, balance, HRV, FFFS, BIS, RI) %>% 
+#   group_by(time, gender) %>% 
+#   summarise(across(.cols = 1:6, .fns = list(mean = mean))) %>% 
+#   pivot_longer(cols = c(-time, -gender), names_to = "median", values_to = "variable") %>% 
+#   mutate(median = str_remove(median, "_[a-z]*")) %>% 
+#   pivot_wider(names_from = gender, values_from = variable)
+#   view
+# 
+# datanet %>%
+#     select(gender, stiffness) 
+# 
+# 
+# 
+# 
+# 
+# 
+# %>% 
+#     group_by(time, gender) %>% 
+#     summarise(across(.cols = 1:6, .fns = list(mean = mean))) %>% 
+#     pivot_longer(cols = c(-time, -gender), names_to = "median", values_to = "variable") %>% 
+#     mutate(median = str_remove(median, "_[a-z]*")) %>% 
+#     pivot_wider(names_from = gender, values_from = variable)
+#   view
+#   
+

### 2020-11-15 r-files/results_code.R

@@ -459,6 +459,7 @@ labels <- as.character(strq1$strength)
 labels[4] <- str_pad(labels[4], side = "left", width = 2)
 
 edgenames <- edgeNames(q1p)
+edgenames <- edgenames[c(1,5,2,3,4)]
 names(labels) <- edgenames
 nodeRenderInfo(q1p)$fill[c("Injured_1", "Negative\nlife\nevents_1")] = "gold"
 edgeRenderInfo(q1p) <- list(lwd = c("Negative life\nevents_1~Injured_1" = 6,
@@ -469,7 +470,7 @@ edgeRenderInfo(q1p) <- list(lwd = c("Negative life\nevents_1~Injured_1" = 6,
 
 q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
 graph.par(list(edges=list(fontsize = 15), 
-               nodes=list(fontsize = 16)))
+               nodes=list(fontsize = 15)))
 
 nodeRenderInfo(q1p)$fill[c("Negative life\nevents_1", "Stiffness_1", "Balance_1")] = "deepskyblue"
 nodeRenderInfo(q1p)$fill["Injured_1"] = "gold"
@@ -479,7 +480,7 @@ renderGraph(q1p)
 dev.off()
 
 
-plotsg2 <- model2network('[Negative life\nevents_2][Gender][Fight-Flight-Freeze\nSystem_1][Injured_2|Fight-Flight-Freeze\nSystem_1:Negative life\nevents_2][Stiffness_2|Gender:Injured_2][Balance_2|Injured_2][Heart rate\nvariability_2|Injured_2]')
+plotsg2 <- model2network('[Negative life\nevents_2][Gender][Fight Flight\nFreeze\nSystem_1][Injured_2|Fight Flight\nFreeze\nSystem_1:Negative life\nevents_2][Stiffness_2|Gender:Injured_2][Balance_2|Injured_2][Heart rate\nvariability_2|Injured_2]')
 q2p <- graphviz.plot(plotsg2,
                      render = F,
                      shape = "rectangle",
@@ -495,36 +496,36 @@ strq2 <- str %>%
 labels <- as.character(strq2$strength)
 labels[6] <- str_pad(labels[6], side = "left", width = 2)
 edgenames <- c("Gender~Stiffness_2", 
-               "Fight-Flight-Freeze\nSystem_1~Injured_2",
+               "Fight Flight\nFreeze\nSystem_1~Injured_2",
[…]
